### 2014-01-24 www/Rennes.tex
adding some files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@267 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
%
%
% Retirez le caractere "%" au debut de la ligne ci--dessous si votre
% editeur de texte utilise des caracteres accentues
% \usepackage[latin1]{inputenc}

%
% Retirez le caractere "%" au debut des lignes ci--dessous si vous
% utiisez les symboles et macros de l'AMS
\usepackage{amsmath}
\usepackage{amsfonts}
%
%
\setlength{\textwidth}{16cm}
\setlength{\textheight}{21cm}
\setlength{\hoffset}{-1.4cm}
%
%
\begin{document}

%--------------------------------------------------------------------------

\begin{center}
{\Large
	{\sc  Regression for correlated variables :\\ Application in steel industry}
}
\bigskip

 Cl\'ement Th\'ery $^{1}$
\bigskip

{\it
$^{1}$ ArcelorMittal Dunkerque, Inria Lille, Université de Lille 1, clement.thery@arcelormittal.com

}
\end{center}
\bigskip

%--------------------------------------------------------------------------

{\bf R\'esum\'e.} La r\'egression lin\'eaire suppose en g\'en\'eral l'usage de variables explicatives ind\'ependantes. Les variables pr\'esentes dans les bases de donn\'ees d'origine industrielle sont souvent tr\`es fortement corr\'el\'ees (de par le process, diverses lois physiques, etc). Le mod\`ele g\'en\'eratif propos\'e consiste à expliciter les corr\'elations pr\'esentes sous la forme d'une de sous-r\'egressions lin\'eaires. La structure est ensuite utilis\'ee pour obtenir un mod\`ele lib\'er\'e des corr\'elations, facilement interpr\'etable et compatible avec les m\'ethodes de s\'election de variables. La structure de corr\'elations est d\'etermin\'ee à l'aide d'un algorithme de type MCMC. Un package R (CorReg) permet la mise en oeuvre de cette m\'ethode. 
\smallskip

{\bf Mots-cl\'es.} R\'egression, corr\'elations, industrie, s\'election de variables, mod\`eles g\'en\'eratifs, SEM (Structural Equation Model) \ldots
\bigskip\bigskip

{\bf Abstract.} Linear regression generally suppose independence between the covariates. Datasets found in industrial context often contains many highly correlated covariates (due to the process, physcial laws, etc). The proposed generative model consists in explicit modeling of the correlations with a structure of sub-regressions between the covariates. This structure is then used to obtain a model with independent covariates, easily interpreted, and compatible with any variable selection method. The structure of correlations is found with an MCMC algorithm. A R package (CorReg) implements this new method.  
\smallskip

{\bf Keywords.} Regression, correlations, industry, variable selection, generative models, Structural Equation Model \ldots

%--------------------------------------------------------------------------

\section{Le contexte}
	La régression linéaire classique suppose l'indépendance des covariables. Les corr\'elations posent des probl \`emes.
	\begin{eqnarray}
		Y&=&XA+\varepsilon \ \ \ \varepsilon\sim \mathcal{N}(0,\sigma^2) \\
		Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont lin\'eairement corr\'el\'ees}
	\end{eqnarray}

\section{Le mod \`ele g\'en\'eratif}
On dispose de $p$ variables $X$ fortement corrélées pour expliquer une variable réponse $Y$.
On rend explicite les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
	\begin{eqnarray}
		I_1&=&(I_1^1,\dots,I_1^{p_2}) \textrm{ avec}		\\
		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices des covariables qui expliquent $X^j$} \\
		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables dépendantes} \\
		p_2&=& \sharp I_2 \\
		\forall 1\leq j\leq p :  p_1^j&=&\sharp I_1^j
	\end{eqnarray}
	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendentes dans $X$ n'en expliquent pas d'autres.

	On note $I_2^c=\{1,\dots,p\}\setminus I_2$
Then our generative model can be written :
\begin{eqnarray}
	Y_{|X,S}=Y_{|X}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ with } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ with } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ free law}	
\end{eqnarray}
Where $B_{I_1^j}^j$ is the $p_1^j$-sized vector of the coefficients of the subregression.

We note that (\ref{MainR}) and (\ref{SR}) also give :
\begin{eqnarray}
	Y&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
					&=& X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}=X\tilde{A}+ \tilde{\varepsilon}\label{Trueexpl} \\
			\textrm{where }		\tilde{A}_{I_2}&=&0 \\
					\tilde{A}_{I_2^c}&=&A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j} \\
					\tilde{\varepsilon}&=&\sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
\end{eqnarray}
\section{Estimateur}
	Classical methods like Ordinary Least Squares (OLS) estimate $Y|X$ and obtain (Maximum Likelihood Estimation): 
		\begin{equation}
			\hat A = (X'X)^{-1}X'Y \textrm{ (ill-conditoned matrix to inverse)}
		\end{equation}
		With following properties :
		\begin{eqnarray}
			E[\hat{A}|X]&=&A \\
			Var[\hat{A}|X]&=& \sigma_Y^2(X'X)^{-1}
		\end{eqnarray}				
		And when correlations are strong, the matrix to invert is ill-conditioned and the variance explodes.

		Our idea is to reduce the variance so we explain $Y$ only with $X^{I_1}$ knowing (\ref{SR}) and (\ref{Trueexpl})
			\begin{equation}
				Y= X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}\label{explicatif}
			\end{equation}							
		So the new estimator simply is : 
		\begin{eqnarray}
			\hat{\tilde{A}}_{I_2^c} &=& (X'_{I_2^c} X^{I_2^c})^{-1}X'_{I_2^c}Y \\
			\hat{\tilde{A}}_{I_2} &=& 0
		\end{eqnarray}
		and we get the following properties :
		\begin{eqnarray}
			E[\hat{\tilde{A}}|X]&=&\tilde{A} \\
			Var[\hat{\tilde{A}}_{I_2^c}|X]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
			Var[\hat{\tilde{A}}_{I_2}|X]&=& 0 
		\end{eqnarray}
		We see that the variance is reduced (no correlations and smaller matrix give better conditioning) for small values of $\sigma_j$ $i.e.$ strong correlations.

		Both classical and our new estimators of $Y$ are unbiased (true model)\cite{saporta2006probabilites}.
	\\	
			This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used. 
		 \\				
			The explicit structure between the covariates helps to understand the model and the complex link between the covariate and the response variable so we call this model explicative.

			When we use a variable selection method on it we obtain two kinds of 0 :
			\begin{enumerate}
			\item Because of the structure we coerce $\hat{\tilde{A}}^{I_2} = 0 $. This kind of zero means redundant information but the covariate can be correlated with the response variable. So we don't have the grouping effect (so we are more parsimonious ) and we don't suffer from false interpretation (LASSO would).
			\item Variable selection methods can lead to get some exact zeros in $\hat{\tilde{A}}^{I_1}$. This kind of zero means that implied covariate has no significant effect on the response variable. And because variables in $X^{I_1}$ are orthogonal, we know that it is not misleading interpretation due to correlations.
			\end{enumerate}
\section{Recherche de structure}

		\subsection{Comparaison des structures}
	On utilise le Bayesian Information criterion (BIC)~\cite{BIChuard}. But BIC tends to give too complex structures because we test a great range of models. Thus we choose to penalise the complexity a bit more with specific a priori laws(uniform laws for the number of subregression and the complexity of each subregression instead of uniform law on S) :
	\begin{eqnarray}
		P(S)&=&P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2) \\
		P(I_1 | p_1,I_2,p_2)&=&\prod_{j =1}^{p_2}P(I_1^j|p_1^j,I_2,p_2) \\
		P(I_1^j|p_1^j,I_2,p_2)&=&\left(\begin{array}{c}
			p-p_2 \\ 
			p_1^j
			\end{array}  \right)^{-1} =\frac{p_1^j ! (p-p_2-p_1^j)!}{(p-p_2)!}\\
		P(p_1|I_2,p_2)&=&\prod_{j =1}^{p_2}P(p_1^j|I_2,p_2)		\\
		P(p_1^j|I_2,p_2)&=&\frac{1}{p-p_2}  \\
		P(I_2|p_2)&=&\left(\begin{array}{c}
			p \\ 
			p_2
			\end{array}  \right)^{-1}=\frac{p_2!(p-p_2)!}{p!}\\
		P(p_2) &=&\frac{1}{p_2} \\
		P(S)&=&\left(\prod_{j =1}^{p_2}\left(\begin{array}{c}
			p-p_2 \\ 
			p_1^j
			\end{array}  \right)^{-1}\right) \left(\frac{1}{p-p_2}\right)^{p_2}\frac{p_2!(p-p_2)!}{p!}\frac{1}{p_2} \\
			\ln P(S) &=& -\sum_{j=1}^{p_2}	\ln\left(\begin{array}{c}
			p-p_2 \\ 
			p_1^j
			\end{array}  \right)
			-p_2\ln (p-p_2)
			-\ln\left(\begin{array}{c}
			p \\ 
			p_2
			\end{array}  \right)
			-\ln( p_2	)
	\end{eqnarray}
	Then we have 
	\begin{eqnarray}
		P(S|X)&\propto &P(X|S)P(S) \\
		\ln(P(S|X))&=&\ln(P(X|S))+\ln(P(S))+cste \\
		%&=&BIC +\ln(P(S))+cste \\
		BIC^*&=&BIC +\ln(P(S))
	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ thus in the following we will use $BIC*$ under this hypothesis (that becomes a constraint in the MCMC).			
	\subsection{The Markov chain}
	First we remember that $S$ is completely described with $I_1$ :
	So we will only describe the variations in $I_1$ at each step and other parts of $S$ will follow according to the previous definition.
	for each step, starting from $S \in \mathcal{S}$ we define a neighbourhood $\mathcal{V}_{S,j}$ with $j \sim \mathcal{U}(\{1,\dots,p\}) $ like this  :	
	\begin{eqnarray}
		\mathcal{V}_{S,j}&=&\{ S^{(i,j)} | 1\leq i\leq p \} \cup\{S \}
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_i^j$ (add): 
			\begin{itemize}
				\item $I_1^j=I_1^j\cup \{i\}$
				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item if $i \in I_1^j$ (remove): $I_1^j=I_1^j\setminus \{i\}$
	\end{itemize}
	And then :

	We have $|\mathcal{V}_{S,j}|=p$ but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the subregressions or the whole structure, a maximum number of sub-regressions, etc). CorReg allows to modify this neighbourhood to better fit users constraints.

	We make a first approximation (\ref{Bicstar}) : 
	\begin{equation}
		P(S|X)\approx exp(BIC^*(S))
	\end{equation}
	We define ~\cite{BIChuard}, :
	\begin{equation}
		q(\tilde{S},\mathcal{V}_{S,j})=\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))}
	\end{equation}

	Where $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $ is the gap between a structure and the worst structure in the neighbourhood in terms of BIC.
	\newline

	 And then we can note $\forall (S,\tilde{S}) \in \mathcal{S}^2 $ :
		\begin{displaymath}
			\mathcal{P}(S,\tilde{S})= \frac{1}{p} \sum_{j=1}^p q(\tilde{S},\mathcal{V}_{S,j})
		\end{displaymath}
	The output will be the best structure seen in terms of BIC. If we have some knowledge about some sub-regressions (physical models for example) we can add them in the found structure. So the model is really expert-friendly.

 \subsubsection{initialisation}
L'intialisation peut par exemple utiliser la matrice des corrélations et/ou la méthode du Graphical Lasso\cite{friedman2008sparse}.		
La grande dimension de l'espace parcouru rend préférable l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (pour un temps de calcul égal). Accessoirement, les multiples chaînes permettent de paralléliser la recherche, ce qui peut être très appréciable.

	\subsubsection{properties}
	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité

%	if 	$\forall k, \tilde{Z} \not\in \mathcal{V}_{Z,k} $ then 
We want 
		\begin{equation}
			\mathcal{P}(S,\tilde{S})=\mathbf{1}_{[\exists j, \tilde{S} \in \mathcal{V}_{S,j} ]} P(\tilde{S}|X)
		\end{equation}
			Because the walk follows a regular and thus ergodic markov chain with a finite state space, it has exactly one stationary distribution \cite{grinstead1997introduction} : $\pi$ and every rows of $\operatorname{lim}_{k\rightarrow \infty}\mathcal{P}^k=W$ equals $\pi$.

\section{R\'esultats}	
\section{Conclusion et perspectives}
	CorReg est fonctionnel et disponible
	Besoin d'\'elargir à la gestion des valeurs manquantes tr \`es pr\'esentes dans l'industrie
\section{Exemple de r\'ef\'erences bibliographiques}
La n\'ecessit\'e de produire des r\'esum\'es clairs et bien
r\'ef\'erenc\'es a \'et\'e d\'emontr\'ee par Achin et Quidont~(2000). Le
r\'ecent article de Noteur~(2003) met en \'evidence \dots
%Quelques rappels :
%%
%\begin{center}
%%
%\begin{tabular}{lr} \hline
%%
%Accent aigu :              &  \'e; \\
%Accent grave :             &  \`a;\\
%Accent circonflexe :       &  \^o mais \^{\i};\\
%Tr\'emas :                 &  \"o mais \"{\i};\\
%C\'edille :                &  \c{c}. \\ \hline
%\end{tabular}
%%
%\end{center}
%--------------------------------------------------------------------------
\section*{Bibliographie}
\cite{marquardt1975ridge}
\bibliography{biblio}{}
\bibliographystyle{plain}
à recopier dans le bon ordre comme demand\'e ci-dessous.
\noindent [1] Auteurs (ann\'ee), Titre, revue, localisation.
\noindent [2] Achin, M. et Quidont, C. (2000), {\it Th\'eorie des
Catalogues}, Editions du Soleil, Montpellier.
\noindent [3] Noteur, U. N. (2003), Sur l'int\'er\^et des
r\'esum\'es, {\it Revue des Organisateurs de Congr\`es}, 34, 67--89.
\end{document}

### 2014-01-24 www/article9.tex
adding some files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@267 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage[table]{xcolor}
 \graphicspath{{figures/}}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Clément THERY}
\title{Regression for correlated variables : application in steel industry}

%%%% debut macro %%%%
\newenvironment{changemargin}[2]{\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{0pt}%
\setlength{\rightmargin}{0pt}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setlength{\parsep}{0pt plus 1pt}%
\addtolength{\leftmargin}{#1}%
\addtolength{\rightmargin}{#2}%
}\item }{\end{list}}
%%%% fin macro %%%%

\definecolor{darkgreen}{rgb}{0,0.4,0}
	 \definecolor{darkred}{rgb}{0.75,0,0}
	 \definecolor{darkblue}{rgb}{0,0,0.4}

\begin{document}
\maketitle
	\begin{itemize}
		\item resume
		\item keywords : linear regression, correlations, sem, selection, graphs, pretreatment, plugin
	\end{itemize}
\section{Introduction}
%la régression et ses problèmes

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability and is simple to learn for non-statistician. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation that faces an also classical problem : the variance of the estimators.
This variance increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model  : the more covariates you have the greater variance you get.
	\item The correlations within the covariates : strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

	With the rise of informatic, datasets contains more and more covariates and thus more and more useless covariates. So dimension reduction becomes a necessity. Moreover, when you use more covariates, you increase the chance to have correlated ones. For example, this work takes place in an industrial context with a big set of covariates (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc). In such a context, variance of the estimators can lead to arbitrary results or even no results at all. Prediction and interpretation are both strongly needed, with a preference for interpretation in industrial context (better to improve the process when possible than to only predict defects).

		~\\	~\\

%bibliographie

	When estimating the parameters of the regression we have to compute the inverse of a matrix\cite{saporta2006probabilites} which will be ill-conditioned or even singular if some covariates depend linearly from each other. For a model defined by \begin{equation}
	Y=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2)$ 
	we have the following Ordinary Least Squares (OLS) estimators :
	\begin{equation}
		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
	\end{equation}

	Because it is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
	Ridge regression\cite{marquardt1975ridge} proposes a biased estimator :
	\begin{equation}
		\hat{\beta}=(X'X+kI)^{-1}X'Y \textrm{ with } k\geq 0 
	\end{equation}
	But Ridge regression is not efficient to select covariates (it's an assumed choice) because coefficients tends to 0 but don't reach 0.
	So it gives difficult interpretations and is not adapted for our industrial context. We need to reduce the dimension of the model. Our goal is not just to predict but also to understand the response variable.

	Real datasets implies many irrelevant variables (datasets based on the whole process without any a priori) so we have to use variable selection methods.

	We note classical norms: $\parallel\beta\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\beta \parallel_1=\sum_{i=1}^p|\beta_i| $

	The Least Absolute Shrinkage and Selection Operator (LASSO)\cite{tibshirani1996regression} consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression\cite{efron2004least} (LAR) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates. And it  really selects covariates with coeffients set exactly to 0.
	 But LASSO also faces consistency problems\cite{Zhao2006MSC} when confronted  with correlated covariates. This point will be developped further (see \ref{consistency}) with numerical results. Another limitation of the LASSO is that it preserves at most $n$ predictors (troublesome when in high dimension).

	Elastic net\cite{zou2005regularization}	is a method developped to be a compromise between Ridge regression and the LASSO. 
%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
%	\begin{equation}
%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
%\end{equation}		
	Elastic net can be written:
	\begin{equation}
		\hat{\beta}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel Y-X\beta \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\beta\parallel_1+\alpha\parallel\beta\parallel_2^2\leq t \textrm{ for some } t
	\end{equation}
	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$
	It seems to give good predictions. But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.

	The CLusterwise Effect REgression\cite{yengo2012variable} (CLERE) tries to reduce the dimension by considering the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith $\beta$ following a Gaussian Mixture distribution. 
\begin{eqnarray}
	\beta_j|\mathbf{z}_j &\sim& \mathcal{N}(\sum_{k=1}^gb_kz_{jk},\gamma^2) \textrm{ with }\\
	\mathbf{z}_j&=&(z_{j1},\dots,z_{jg})\sim \mathcal{M}(\pi_1,\dots,\pi_g) \textrm{ and }\\
	\forall k=1,\dots ,g &&\sum_{j=1}^pz_{jk}\geq 1
\end{eqnarray}

The idea is to hope that this mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\hat{\beta}$. But we need to suppose having many covariates with the same level of effect on the response variable and seems to be less efficient in prediction than elastic net.

	~\\	~\\
% Principe de la méthode

 When some try to reduce the dimension and then just hope to have small correlations in the remaining dimensions, we propose to focus on the correlations, giving a model with orthogonal covariates. In fact we search the greatest set of orthogonal covariates to keep the maximum but with an orthogonality constraint. This can be viewed as a pretreatment on the dataset allowing to use then dimension reduction tools without suffering from correlations. We only consider strong correlations (i.e. : problematic ones) thus we keep most of the information contained in the dataset. We will in a second time be able to use the remaining part of the information (sequential approach).

	Our work is based on the assumption that if we know that correlations are a problem and if we precisely know the correlations, we could use this knowledge to avoid the problem.
	The idea is to suppose explicitly a linear structure between the covariates. It gives a recursive Structural Equation Model (SEM)\cite{davidson1993estimation}. That can be viewed as a system of linear regression. 	
		\begin{eqnarray}
		Y&=& X\beta_Y+\varepsilon_Y \\
		X&=&XB+ \varepsilon_{X} 
\end{eqnarray}

	Recursive sem don't really have a specific estimator because general system estimators (Seemingly Unrelated Regression (SUR)\cite{SURzellner} and Two-Stage Least Squares (2SLS)) are equivalent to independent Ordinary Least Squares when applied to recursive SEM \cite{TIMM}. Our work can be viewed as a new way of estimating recursive SEM based on their own structure.
	  In this work, we decide to distinguish the response variable from the other variables that are on the left of a regression. Thus we don't have a system of regressions but one regression on our response variable and a system of subregressions (without the response variable).
	  The structure is supposed to be the source of the correlations and allows us to define a reduced set of independent covariates. Thus we reduce dimension and correlations in the same time. The structure justifies the eviction of the redundant covariates without significant information loss. It can be seen as a pretreatment on the dataset based on the hypothesis of a strong structure between the covariates (i.e. : small $\varepsilon_X$).

	We can use any variable selection method on the reduced dataset with improved efficiency (reduced variance) due to dimension reduction and correlation suppression. So we obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.

 	We then observe that the reduced model only uses the partition given by the structure (who is explained and who explains) but not the structure itself (how they interacts). We also notice that even if variables are highly correlated, each can have a specific effect ($\varepsilon_X$) on the response variable.

 	So in a second step we propose further usage of the structure, taking back correlated variables to estimate the residuals of the reduced model. This estimation is also a linear regression that can take profit from any variable selection method. The idea is to estimate the complete model under the constraint of the structure. It can be done with Constrained Least Squares \cite{amemiya1985advanced} or sequentially with classical OLS (non optimal but easier to use selection methods).

But to work, we need an explicit structure between the covariates. SEM are often used in social sciences and economy where a structure is supposed "by hand" but here we want to be able to find a structure without any a priori (possibility to include some known structure remains). Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix). It consists in a selection in the precision matrix, setting some covariances to zero. But the resulting matrix is symmetric and we need an oriented graph for our SEM. So we developped an MCMC algorithm to find it (R package CorReg on Rforge). However, Graphical LASSO can be used in the initialization step of our MCMC. This structure is based on gaussian mixture models to fit better real datasets and to allow identifiability of the structure in terms of complexity (number of parameters).

 	~\\

 	%plan
 	This paper will first present the reduced model, its properties and the algorithm used to find the structure. Then we talk about further usage of the method, both estimating residuals of the reduced model and managing missing values in the dataset. 
 	We will finish with some numerical results on simulated and real industrial datasets before concluding and giving some perspectives.

\section{Model to decorrelate the covariates}

We have a $p$ correlated covariates $X$ to explain a response variabe $Y$.
Let $Z$ be the adjacency matrix that defines which covariates is depending on which others. That is $Z_{i,j}=\mathbf{1}_{(X^j depends\ on X^i)}$. We consider dependencies in a generative point of view ("depends on" means "is generated according to" ) so $Z$ is not symmetric and has no cycles.

	We can describe the structure $Z$ by $S=(p_2,I_2,p_1,I_1)$ defined by :
	\begin{eqnarray}
		p_2&=& \sum_{j=1}^p\mathbf{1}_{(\exists i, Z_{i,j}\neq 0)} \textrm{ the number of sub-regressions}\\
		I_2&=&(I_2^1,\dots,I_2^{p_2}) \textrm{ vector of the indices of the dependent covariates}\\
		I_1&=&(I_1^1,\dots,I_1^{p_2}) \textrm{ with}		\\
		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices of the covariates explaining $X^j$} \\
		p_1&=&(p_1^1,\dots,p_1^{p_2}) \textrm{ where }p_1^j=\sharp I_1^j 
	\end{eqnarray}
	We suppose $I_1\cap I_2=\emptyset$, $i.e.$ dependent variables don't explain other variables in $X$.

	We note $I_2^c=\{1,\dots,p\}\setminus I_2$
Then our generative model can be written :
\begin{eqnarray}
	Y_{|X,S}=Y_{|X}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ with } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ with } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ free law}	
\end{eqnarray}
Where $B_{I_1^j}^j$ is the $p_1^j$-sized vector of the coefficients of the subregression.

We note that (\ref{MainR}) and (\ref{SR}) also give :
\begin{eqnarray}
	Y&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
					&=& X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}=X\tilde{A}+ \tilde{\varepsilon}\label{Trueexpl} \\
			\textrm{where }		\tilde{A}_{I_2}&=&0 \\
					\tilde{A}_{I_2^c}&=&A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j} \\
					\tilde{\varepsilon}&=&\sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
\end{eqnarray}
%We can also use a matricial notation for the subregressions :
%\begin{eqnarray}
%	X^{I_2}&=&X^{I_2^c}B^{I_2}_{I_2^c}+ \varepsilon_{I_2}
%\end{eqnarray}
%				Where $B^{I_2}_{I_2^c}$ is a $(p-p_2) \times p_2$  matrix and $\varepsilon_{X^{I_2}}$ a $n \times p_2$ noise matrix. 
%				Column $i$ : $\varepsilon_X(i)\sim \mathcal{N}(0,d_i^2I_n)$. And $\varepsilon_X(i)\perp\varepsilon_X(j),i\neq j$.
\subsection{Structure identifiability}
	\begin{equation}
		\mathbf{P}(X)=\mathbf{P}(X^{I_2^c},X^{I_2})=\mathbf{P}(X^{I_2}|X^{I_2^c})\mathbf{P}(X^{I_2^c})\label{decompprobas}
	\end{equation}	
		When estimating the quality of the model we look at the likelihood (to compute the Bayesian Information Criterion (BIC)):
			\begin{equation}
				\mathcal{L}_B(X)=\prod_{\substack{1 \leq i \leq n \\ j \in I_2}} \mathbf{P}(X_{i,j} |X^{I_1^j})\prod_{\substack{1 \leq i \leq n \\j \in I_1^j}}\mathbf{P}(X_{i,j}) \label{likelihoodind}
			\end{equation} 
	We suppose $\forall j \in I_2^c : X^{j}$  follows gaussian \underline{mixture} models to better fit industrial variables (Figure \ref{figmixture}) estimated separately (we use the Rmixmod package for R). In the following, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
	\begin{figure}
	\begin{center}				
				\includegraphics[width=150px]{figures/mixmod.png} 
				\caption{Industrial variables seem to follow gaussian mixture models}\label{figmixture}
	\end{center}
	\end{figure}

		\begin{eqnarray}
			\forall j \in I_2 : \  X^j_{|X^{I_1^j}}&\sim & \mathcal{N}(X^{I_1^j}B_{I_1^j}^j,\sigma^2_j) \label{densitycondgauche}\\
			\forall j \in I_2^c : \  X^j &\sim & \mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j \label{densitydroite}\\
			\textrm{And we obtain, } \forall j \in I_2 : \  X^j&\sim & \mathcal{GM}(\bigotimes_{\substack{i \in I_1^j \\ Z_{i,j}=1 }}\pi_i \ ; 
			                                                         \bigoplus_{\substack{i \in I_1^j \\ Z_{i,j}=1 }} B_{i,j} \mu_i \ ; 
			                                                          \sigma_j^2+\bigoplus_{\substack{i \in I_1^j \\ Z_{i,j}=1 }}B_{i,j}^2 \sigma_i^2 )\label{densitygauche}
		\end{eqnarray}
		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of classes in $I_2$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective). We call this difference in component number "identifiability" in the sense that we try to find the model with the fewest component in $X^{I_1}$ (we use the BIC so when comparing two models with the same likelihood, the one with lesser parameters will win). So in this case we have a "better" model (the simplest one) in the group of equivalent models (permuting variables in each subregression).

		Remark : the uncrossing constraint ($I_2\cap I_1 = \emptyset$) significatively reduces the number of feasible models and thus increases identifiability (it also reduces the number of models equivalent in interpretation).

		And if we try to permute one subregression we obtain for $k \in I_2$ and $j \in I_1^k$:	
	\begin{eqnarray}
		X^k-\varepsilon_X^k-\sum_{\substack{i \in I_1^k \\ i\neq j }}X^iB_{i,k} &\sim&  \mathcal{GM}\left(\tilde{\pi}_j \ ; \tilde{\mu}_j \ ; \tilde{\sigma}^2_j \right) \textrm{ where} \\
		 \tilde{\pi}_j &=&  (\bigotimes_{i \in I_1}\pi_i)\otimes \bigotimes_{\substack{i \in I_1 \\ i\neq j }}\pi_i \\
		 \tilde{\mu}_j &=&   \frac{(\bigoplus_{i\in I_1} B_{i,k} \mu_i)\oplus \bigoplus_{\substack{i \in I_1 \\ i\neq j }} (-B_{i,k}\mu_i)}{B_{j,k}} \\
		 \tilde{\sigma}^2_j&=&\frac{ 2d_k^2+(\bigoplus_{i\in I_1}B_{i,k}^2 \sigma_i^2)\oplus \bigoplus_{\substack{i \in I_1 \\ i\neq j }} B_{i,k}^2 \sigma_i^2 }{B_{j,k}^2}
	\end{eqnarray}
	Thus if we call identifiability  the ability to find the model with the smallest variance on the subregression residuals, we have identifiability  on the structure if $ X^k-\varepsilon_X^k-\sum_{\substack{i \in I_1 \\ i\neq j }}X^iB_{i,k} \nsim  X^j$. That is the case if $d_k^2>0$ (non-exact subregression) because the variance of the noise of the subregression is doubled when the subregression is permuted. So even with gaussian variables, non-exact subregression are a sufficient condition for identifiability (in the meaning of smallest-variance subregressions). But you need to look the marginal laws because the joint one relies on the structure and does not use marginal laws of the left-sided covariates. 
	%en fait on est conditionnel à X donc permutable mais la permutation va à l'encontre de l'hypothèse mixmod...en gros. Par contre l'algo ne s'en rend pas compte pour l'instant car il travaille sur la loi jointe et pas sur les lois marginales induites.

	If there are exact subregressions, classical methods will fail (singular matrix) and the structure won't be identifiable. But it only means that several structure will have the same likelihood and they will have the same interpretation. So it's not a problem. Moreover, when an exact subregression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete. CorReg (Our R package) prints a warning to point out exact subregressions when found.

	\subsection{Estimator and properties}

		Classical methods like Ordinary Least Squares (OLS) estimate $Y|X$ and obtain (Maximum Likelihood Estimation): 
		\begin{equation}
			\hat A = (X'X)^{-1}X'Y \textrm{ (ill-conditoned matrix to inverse)}
		\end{equation}
		With following properties :
		\begin{eqnarray}
			E[\hat{A}|X]&=&A \\
			Var[\hat{A}|X]&=& \sigma_Y^2(X'X)^{-1}
		\end{eqnarray}				
		And when correlations are strong, the matrix to invert is ill-conditioned and the variance explodes.

		Our idea is to reduce the variance so we explain $Y$ only with $X^{I_1}$ knowing (\ref{SR}) and (\ref{Trueexpl})
			\begin{equation}
				Y= X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}\label{explicatif}
			\end{equation}							
		So the new estimator simply is : 
		\begin{eqnarray}
			\hat{\tilde{A}}_{I_2^c} &=& (X'_{I_2^c} X^{I_2^c})^{-1}X'_{I_2^c}Y \\
			\hat{\tilde{A}}_{I_2} &=& 0
		\end{eqnarray}
		and we get the following properties :
		\begin{eqnarray}
			E[\hat{\tilde{A}}|X]&=&\tilde{A} \\
			Var[\hat{\tilde{A}}_{I_2^c}|X]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
			Var[\hat{\tilde{A}}_{I_2}|X]&=& 0 
		\end{eqnarray}
		We see that the variance is reduced (no correlations and smaller matrix give better conditioning) for small values of $\sigma_j$ $i.e.$ strong correlations.

		Both classical and our new estimators of $Y$ are unbiased~\cite{saporta2006probabilites} (true model).
	\\

			There is no theoretical guarantee that our model is better. It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis. Therefore we made some simulations to compare both methods (see the end of this paper). This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used. Hence we hope to obtain a parsimonious model. 
		 \\

			The explicit structure between the covariates helps to understand the model and the complex link between the covariate and the response variable so we call this model explicative.

			When we use a variable selection method on it we obtain two kinds of 0 :
			\begin{enumerate}
			\item Because of the structure we coerce $\hat{\tilde{A}}^{I_2} = 0 $. This kind of zero means redundant information but the covariate can be correlated with the response variable. So we don't have the grouping effect (so we are more parsimonious ) and we don't suffer from false interpretation (LASSO would).
			\item Variable selection methods can lead to get some exact zeros in $\hat{\tilde{A}}^{I_1}$. This kind of zero means that implied covariate has no significant effect on the response variable. And because variables in $X^{I_1}$ are orthogonal, we know that it is not misleading interpretation due to correlations.
			\end{enumerate}

		\subsection{Why grouping effect is misleading}

			In industrial context, when a model explain why things go wrong, one will try to fix the problem.
			If $X_1=X_2+e$ and we have the grouping effect, we will obtain a model like $Y=aX_1+aX_2$. Then when one will try to modify Y he will modify one of the covariates and both will change so he won't get expected results. Nothing constrains us to give only one equation. It is clearly better to give the user another equation (or system for more complex models) describing the correlations. So you get the following model : $Y=aX_1+aX_2$ \underline{AND} $X_1=X_2+e$. So you have more information and are able to decide better actions. With such a model, grouping effect is no more useful because when saying $Y=2aX_2$ \underline{AND} $X_1=X_2+e$ you clearly show that $X_2$ is correlated with both $Y$ and $X_1$. So it is possible to combine the advantages of grouping effect and selection just giving several equations. Each equation here is very simple so you don't really increase complexity of the model. Uncrossed model (nilpotent Z) guarantee to keep a simple structure easily interpretable.

\section{Estimating subregressions}

  All our work is based on a linear structure between the covariates.
	Let's define $\mathcal{S}$ and $\mathcal{Z}$ the ensemble of feasible structures and the ensemble of corresponding adjacency matrices.

	$Z \in \mathcal{Z} \Leftrightarrow : $
	\begin{itemize}
		\item $Z$ is binary
		\item $Z Z=0 $ ($Z$ is not crossed). Equivalent to $I_1\cap I_2 =\emptyset$.
	\end{itemize}

	So  $\mathcal{Z}$ is just the set of the binary square nilpotent matrices of size $p$. $Z$ is an adjacency matrix and we know \cite{biggs1993algebraic} that $Z^p$ shows the number of paths of length $p$ (linking $p+1$ vertices). So we suppose that $Z$ is nilpotent, meaning it does not contain any non-trivial path. This strong hypothesis also strongly reduces the size of $\mathcal{Z}$.

	Now we have made hypothesis on the distribution, we can use them to compare the structures with the Bayesian Information criterion (BIC)~\cite{BIChuard}. But BIC tends to give too complex structures because we test a great range of models. Thus we choose to penalise the complexity a bit more with specific a priori laws(uniform laws for the number of subregression and the complexity of each subregression instead of uniform law on S) :
	\begin{eqnarray}
		P(S)&=&P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2) \\
		P(I_1 | p_1,I_2,p_2)&=&\prod_{j =1}^{p_2}P(I_1^j|p_1^j,I_2,p_2) \\
		P(I_1^j|p_1^j,I_2,p_2)&=&\left(\begin{array}{c}
			p-p_2 \\ 
			p_1^j
			\end{array}  \right)^{-1} =\frac{p_1^j ! (p-p_2-p_1^j)!}{(p-p_2)!}\\
		P(p_1|I_2,p_2)&=&\prod_{j =1}^{p_2}P(p_1^j|I_2,p_2)		\\
		P(p_1^j|I_2,p_2)&=&\frac{1}{p-p_2}  \\
		P(I_2|p_2)&=&\left(\begin{array}{c}
			p \\ 
			p_2
			\end{array}  \right)^{-1}=\frac{p_2!(p-p_2)!}{p!}\\
		P(p_2) &=&\frac{1}{p_2} \\
		P(S)&=&\left(\prod_{j =1}^{p_2}\left(\begin{array}{c}
			p-p_2 \\ 
			p_1^j
			\end{array}  \right)^{-1}\right) \left(\frac{1}{p-p_2}\right)^{p_2}\frac{p_2!(p-p_2)!}{p!}\frac{1}{p_2} \\
			\ln P(S) &=& -\sum_{j=1}^{p_2}	\ln\left(\begin{array}{c}
			p-p_2 \\ 
			p_1^j
			\end{array}  \right)
			-p_2\ln (p-p_2)
			-\ln\left(\begin{array}{c}
			p \\ 
			p_2
			\end{array}  \right)
			-\ln( p_2	)
	\end{eqnarray}
	Then we have 
	\begin{eqnarray}
		P(S|X)&\propto &P(X|S)P(S) \\
		\ln(P(S|X))&=&\ln(P(X|S))+\ln(P(S))+cste \\
		%&=&BIC +\ln(P(S))+cste \\
		BIC^*&=&BIC +\ln(P(S)) \label{Bicstar}
	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ thus in the following we will use $BIC*$ under this hypothesis (that becomes a constraint in the MCMC).

	\begin{equation}
		BIC(X|S) = \sum_{j=1}^p BIC(X^j|S)
	\end{equation}		
    Where 
    \begin{equation}
    	BIC(X^j|S)=-2\mathcal{L}_{|S}(X^j,\theta_j)+K_j\operatorname{log}(n)
    \end{equation}
	Where $K_j$ is the number of parameters to estimate.
	We will now use the following notation : $BIC(S)=BIC(X|S)$

	If we have some hypothesis on the distribution of some variables (exponentially distributed for example) we can compute corresponding $BIC$ separately and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).

	\subsection{The Markov chain}
		\subsubsection{Formalism with Z}
		For each step, starting from $Z \in \mathcal{Z}$ we define a neighbourhood $\mathcal{V}_{Z,j}$ with $j \sim \mathcal{U}(\{1,\dots,p\}) $ like this  :
		\begin{eqnarray*}
		\mathcal{V}_{Z,j}= \{ \tilde{Z} \in \mathcal{Z} &|&\exists ! i,  \tilde{Z}_{i,j}=1-Z_{i,j}, \textrm{ and } \forall (k,l) \neq (i,j) : \\
				&&  \tilde{Z}_{j,l}=(1-\tilde{Z}_{i,j})Z_{j,l} \textrm{ (row-wise relaxation)}, \\
				&& \tilde{Z}_{k,i}=(1-\tilde{Z}_{i,j}) Z_{k,i} \textrm{ (column-wise relaxation)} \\
				&&  \tilde{Z}_{k,l}=Z_{k,l} \\
				&& \} \cup\{Z \}
	\end{eqnarray*}	
	We have $|\mathcal{V}_{Z,j}|=p$ but some other constraints can be added on the definition of $\mathcal{Z}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the subregressions or the whole structure, a maximum number of sub-regressions, etc). CorReg allows to modify this neighbourhood to better fit users constraints. 
		\subsubsection{Formalism with S}
	First we remember that $S$ is completely described with $I_1$ :
	So we will only describe the variations in $I_1$ at each step and other parts of $S$ will follow according to the previous definition.
	for each step, starting from $S \in \mathcal{S}$ we define a neighbourhood $\mathcal{V}_{S,j}$ with $j \sim \mathcal{U}(\{1,\dots,p\}) $ like this  :	
	\begin{eqnarray}
		\mathcal{V}_{S,j}&=&\{ S^{(i,j)} | 1\leq i\leq p \} \cup\{S \}
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_i^j$ (add): 
			\begin{itemize}
				\item $I_1^j=I_1^j\cup \{i\}$
				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item if $i \in I_1^j$ (remove): $I_1^j=I_1^j\setminus \{i\}$
	\end{itemize}
	And then :
	\begin{eqnarray}
		I_2&=&\{j |\sharp I_1^j>0 \}\\
		p_2&=& \sharp I_2 \\
		\forall 1\leq j\leq p :  p_1^j&=&\sharp I_1^j
	\end{eqnarray}
	We have $|\mathcal{V}_{S,j}|=p$ but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the subregressions or the whole structure, a maximum number of sub-regressions, etc). CorReg allows to modify this neighbourhood to better fit users constraints.

	\subsubsection{properties}
	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité

%	if 	$\forall k, \tilde{Z} \not\in \mathcal{V}_{Z,k} $ then 
We want 
		\begin{equation}
			\mathcal{P}(S,\tilde{S})=\mathbf{1}_{[\exists j, \tilde{S} \in \mathcal{V}_{S,j} ]} P(\tilde{S}|X)
		\end{equation}
			Because the walk follows a regular and thus ergodic markov chain with a finite state space, it has exactly one stationary distribution \cite{grinstead1997introduction} : $\pi$ and every rows of $\operatorname{lim}_{k\rightarrow \infty}\mathcal{P}^k=W$ equals $\pi$.

	With $\forall S \in  \mathcal{S}$ :
	\begin{eqnarray}		
		0 \leq &\pi (S)& \leq 1 \nonumber \\
		\sum_{S \in \mathcal{S}}\pi(S) &=&1 \nonumber \\
		\pi (S) &=&\sum_{\tilde{S}\in \mathcal{S}} \pi(\tilde{S})\mathcal{P}(\tilde{S},S) \\%définition de la lois stationnaire
	\end{eqnarray}

	We make a first approximation (\ref{Bicstar}) : 
	\begin{equation}
		P(S|X)\approx exp(BIC^*(S))
	\end{equation}
	We define ~\cite{BIChuard}, :
	\begin{equation}
		q(\tilde{S},\mathcal{V}_{S,j})=\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))}
	\end{equation}

	Where $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $ is the gap between a structure and the worst structure in the neighbourhood in terms of BIC.
	\newline

	 And then we can note $\forall (S,\tilde{S}) \in \mathcal{S}^2 $ :
		\begin{displaymath}
			\mathcal{P}(S,\tilde{S})= \frac{1}{p} \sum_{j=1}^p q(\tilde{S},\mathcal{V}_{S,j})
		\end{displaymath}
	The output will be the best structure seen in terms of BIC. If we have some knowledge about some sub-regressions (physical models for example) we can add them in the found structure. So the model is really expert-friendly.

%ergodic=irreductible=melangeant -> par construction, tout est défaisable et reconstructible un point à la fois car toute sous-structure est réalisable
%regular : il existe k un nombre d'étapes qui permet d'aller de partout à partout. Par mélangeance, en prenant k au moins égal au plus long chemin possible (k>=p convient) et en faisant du sur place une fois arrivé.
%aperiodic : permet la stationnarité donc pour tout trajet, on peut faire le même un 1 pas de plus (un pas sur place).
%	Because the walk follows a regular and thus ergodic markov chain with a finite state space, it has exactly one stationary distribution \cite{grinstead1997introduction} : $\pi$ and every rows of $\operatorname{lim}_{k\rightarrow \infty}\mathcal{P}^k=W$ equals $\pi$.
%	
%	
%	With $\forall Z \in  \mathcal{Z}$ :
%	\begin{eqnarray}		
%		0 \leq &\pi (Z)& \leq 1 \\
%		\sum_{Z \in \mathcal{Z}}\pi(Z) &=&1 \\
%		\pi (Z) &=&\sum_{\tilde{Z}\in \mathcal{Z}} \pi(\tilde{Z})\mathcal{P}(\tilde{Z},Z) \\%définition de la lois stationnaire
%			&=&\sum_{\tilde{Z}\in \mathcal{Z}} \pi(\tilde{Z})\frac{1}{p} \sum_{j=1}^p q(Z,\mathcal{V}_{\tilde{Z},j}) \\
%			&=&\sum_{\tilde{Z}\in \mathcal{Z}} \pi(\tilde{Z})\frac{1}{p} \sum_{j=1}^p \mathbf{1}_{ \{Z\in \mathcal{V}_{\tilde{Z},j}\} }\frac{exp(\frac{-1}{2}\Delta BIC(Z,\mathcal{V}_{\tilde{Z},j}))}{\sum_{Z_l\in \mathcal{V}_{\tilde{Z},j}}exp(\frac{-1}{2}\Delta BIC(Z_l,\mathcal{V}_{\tilde{Z},j}))} \\
%			&=& \sum_{\tilde{Z}\in \mathcal{Z}} \pi(\tilde{Z})\frac{1}{p} \sum_{j=1}^p \mathbf{1}_{ \{Z\in \mathcal{V}_{\tilde{Z},j}\} }\frac{exp(\frac{-1}{2}BIC(Z))}{\sum_{Z_l\in \mathcal{V}_{\tilde{Z},j}}exp(\frac{-1}{2} BIC(Z_l))}
%	\end{eqnarray}

%par définition de la loi stationnaire : le mode de la loi stationnaire est le point le plus attractif dans l'absolu.
%la question est de savoir s'il correspond au meilleur BIC
%Or par stationnarité, Pi est invariant après une étape.
% donc les modifications par le rapport local des BIC laissent Pi invariant
%donc le plus attractif global reste aussi attractif quand on lui applique le local

 \subsubsection{initialisation}
The initial structure can be based on a first warming algorithm taking the correlations into account. Ones are randomly placed into $Z$, weighted by the absolute value of the correlations. Then this structure can be reduced by the hadamard product with the binary matrix obtained by Graphical Lasso\cite{friedman2008sparse}.		
	\subsubsection{reduced neighbourhood}
	If the algorithm did not have time to converge, it can be finished with a few step fr which the neighboorhood would only contain smaller candidates. It is equivalent to ask for each element in the structure if the criterion would be better without it. Thus it can be seen as a final cleaning step. But in fact, it's just contui
	\subsubsection{multiple try}
	One would rather test multiple short chains than lose time in initialisation or long chains
\section{Numerical results on simulated datasets}
	Here are some results on simulated datasets. 
	\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure ?}
			The first criterion is $BIC^*$ which is minimised in the MCMC. We can compare this criterion for both the true Structure and the found one. But if the structures differ, the $BIC^*$ will also differ and comparison won't show how far the found structure is from the true one. So we need more precise criterions to compare the true structure $S$ and the found one $\tilde{S}$.
			Global indicators :
			\begin{itemize}
				\item True left : the number of found dependent variables that really are dependent $TL=\sharp(I_2\cap \tilde{I}_2)$ 
				\item Wrong left : the number of found dependent variables that are not dependent $WL=\sharp(\tilde{I}_2)-TL$
				\item Missing left : the number of really dependent variables not found $ML=\sharp{I_2}-TL$
				\item $\Delta p_2$ : the gap between the number of sub-regression in both model : $\Delta p_2=\sharp{I_2}-\sharp(\tilde{I}_2)$. The sign defines if $\tilde{S}$ is too complex or too simple
				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_2}p_1^j-\sum_{j \in \tilde{p}_2}\tilde{p}_1^j$
			\end{itemize}
			And we have aussi some criterion to compare between the structures :				
				\begin{itemize}
					\item Mean  true found $R^2$ (or $sigma^2$) : the mean of $R^2$ (in the true model) for sub-regressions found by the algorithm and existing in the true model (comparing only the left-side variable). $MTF_{R^2}=\frac{1}{TL}\sum_{j \in I_2\cap \tilde{I}_2}R^2(j)$.
				\item Mean Missing $R^2$ (or $sigma^2$) : the mean of $R^2$ for sub-regressions  only in the true structure (comparing only the left-side variable).$MM_{R^2}=\frac{1}{ML}\sum_{j \in I_2 \setminus (I_2\cap \tilde{I}_2)}R^2(j)$.
				\item Mean found $R^2$ (or $sigma^2$) : the mean of $R^2$ (in the found model) for sub-regressions found by the algorithm and existing in the true model (comparing only the left-side variable). $MF_{R^2}=\frac{1}{TL}\sum_{j \in I_2\cap \tilde{I}_2}\tilde{R}^2(j)$.
				\item Mean Wrong $R^2$ (or $sigma^2$) : the mean of $R^2$ for sub-regressions in only in the found structure (comparing only the left-side variable).$MW_{R^2}=\frac{1}{WL}\sum_{j \in \tilde{I}_2 \setminus (I_2\cap \tilde{I}_2)}\tilde{R}^2(j)$.
				\item Mean $R^2$ : the mean of $R^2$ of all sub-regression for a given structure. We can then compare this value for both structures
				\item Mean $\sigma^2$ : the mean of $\sigma^2$ of all sub-regression for a given structure. We can then compare this value for both structures
				\item Complexity : global complexity of a model : $compl=\sum_{j \in p_2}p_1^j$. We can then compare both global complexities.
			\end{itemize}

		\subsubsection{Results}	
	Tableau de la forme :	
	\begin{table}[h!]
\centering
\begin{tabular}{|r|rrrr|rrrrrr|}
  \hline
n & time &trueBIC & BICempty & BIC\_opt & True1 & False1 & missing1 & $\Delta$p2 &True\_left & False\_left \\ 
  \hline
40 &??& ?? & ?? & ?? &?? & ?? & ??& ?? &??& ?? \\ 
  60 &??& ?? & ?? & ?? & ?? & ?? & ?? & ?? & ?? & ?? \\ 
  100 &??& ?? & ?? &?? &?? & ?? & ?? &?? & ?? & ?? \\ 
   \hline
\end{tabular}
\caption{p variables. Markov chain was XX seconds long for $n=100$(mean observed). }
\end{table}

Ordre des critères de comparaison : MSE sur X, Vraigauche, fauxgauche, bics (les 3), vrais1, faux1, missing1,deltap2

	L'idée serait de n'avoir qu'une seule configuration (vu qu'ici on ne dépend pas de Y) et garder la même pour tous les tableaux suivants (pour pouvoir s'appuyer sur celui-ci dans l'interprétation). Tous les tableaux seraient générés en même temps. Pour chaque base générée, on génèrerait plusieurs Y de plusieurs manières pour avoir tous les cas sur les mêmes données.
	La parallélisation des expériences se ferait alors sur le nombre de réplications. Les résultats seraient toujours basés sur Zchapeau (et donc Bchapeau).

	On devrait constater que quand l'algo a le temps de converger, on trouve pour n petit des BICs meilleurs que le vrai modèle (d'où un bruit sur la structure). quand n augmente, ce surapprentissage devrait disparaître et on devrait donc converger vers le vrai Z (et le vrai BIC).

	\subsection{$Y$ depends only on some covariates in $X^{I_1}$}

		\subsubsection{without selection}
			\begin{table}[h!]
			\centering
			\begin{tabular}{|r|rr|rr|rr|}
			  \hline
			n & OLS & (sd) & explicative & (sd) & predictive & (sd) \\ 
			  \hline
			 40 & NA & NA & ?? & ??& ?? & ?? \\ 
			  60 & ??& ?? & ?? & ?? & ?? & ?? \\ 
			  100 & ?? & ?? & ?? & ?? & ?? & ?? \\ 
			   \hline
			\end{tabular}
			\caption{$Y$ only depends on $X^{I_1}$. $p=50$ and $\operatorname{Var}(Y)\simeq 3.10^8$}
			\end{table}

			On doit constater qu'on est meilleurs que OLS, que l'explicatif gagne (vrai modèle possible) mais que le prédictif reste bon. On doit aussi voir que quand n grandit OLS commence à redevenir correct.
		\subsubsection{with selection}
			Avec le même Y que pour le cas sans sélection, (et les mêmes données) on teste simplement d'autres modèles :
			\begin{itemize}
			\item package lm
			\item modèle complet avec lasso (et LAR)
			\item modèle complet avec elastic net
			\item modèle complet ridge
			\item modèle explicatif elastic net
			\item modèle prédictif elastic net
			\item modèle explicatif LASSO
			\item modèle prédictif LASSO
\end{itemize}

	Il y aurait un tableau par valeur de n pour pouvoir donner en plus des MSE des valeurs de sparsité et de validité du modèle (comparaison des positions des 0).
	Je n'ai pas pour l'instant de quoi utiliser CLERE mais l'article CLERE montre qu'elastic net est meilleur en prédiction donc pour les MSE ce n'est pas trop un problème.
	\subsection{$Y$ depends only on some covariates in $X^{I_2}$}
		même chose qu'avant mais on part avec un handicap. les notions d'explicatif et prédictif finaux devraient alors prendre tout leur sens.

	\subsection{global case}		
		Y dépend un peu de tout le monde... me semble trop compliqué car beaucoup trop de cas possibles. la conclusion étant de toute manières qu'on sera quelque part entre les deux cas précédents.			Je mettrais bien des exemples simples et poussés (3 variables explicatives comme dans l'article sur la consistance du lasso) pour que les gens puissent facilement refaire le test chez eux, même sans notre package (hypothèse du vrai Z).
		la simplicité de l'exemple permettrait aussi de voir ce qui se passe si Zchapeau est une version permutée du vraiZ. 
		On testerait là aussi tous les modèles concurrents abordés plus haut.

		Attention : on a une variabilité due à la validation croisée. Sur les mêmes données, quand on lance plusieurs fois la sélection on ne trouve pas toujours exactement les mêmes 0 (tout de même relativement stable, peut s'arranger en choisissant un meilleur K pour la validation croisée).

	\subsection{Consistency Issues}\label{consistency}
		Consistency issues of the LASSO are well known and Zhao \cite{Zhao2006MSC} gives a very simple example to illustrate it.
		We have taken the same example to show how our method is more consistent.
		Here $p=3$ and $n=1000$.We define $X_1,X_2,\varepsilon_Y,\varepsilon_{X} i.i.d. \sim \mathcal{N}(0,1)$ and then $X_3=\frac{2}{3}X_1+\frac{2}{3}X_2+\frac{1}{3}\varepsilon_X$ and $Y=2X_1+3X_2+\varepsilon_Y$.
		We compare consistencies of complete,explicative and predictive model with LASSO (and LAR) for selection.
		It happens that the algorithm don't find the true strucure but a permuted one so we look at the results obtained with the true $Z$ (but $\hat{B}$ is used) and with the structure found by the Markov chain after a few seconds.

		True $Z$ is found $340$ times on $1000$ tries.

		\begin{table}[h!]
		\centering
		\begin{tabular}{|c|c|c|c|}
		\hline 
		 & Classical LASSO & Explicative & Predictive \\ 
		\hline 
		True $Z$ &  1.006479 & \textbf{1.005468} & \textbf{1.006093} \\ 
		\hline 
		$\hat{Z}$ & \textbf{1.006479} & 1.884175 & 1.006517 \\ 
		\hline 
		\end{tabular} 
		\caption{MSE observer on a validation sample (1000 individuals)}
		\end{table}

		We observe as we hoped that explicative model is better when using true $Z$ (coercing real zeros) and that explicative with $\hat{Z}$ is penalized (coercing wrong coefficients to be zeros).
		But the main point is that the predictive model stay better than the classical one whith the true $Z$ and corrects enough the explicative model to follow the classical LASSO closely when using $\hat{Z}$. 
		And when we look at the consistency :
		\begin{table}[h!]	
		\centering
		\begin{tabular}{|c|c|c|c|}
		\hline 
		 & Classical LASSO & Explicative & Predictive \\ 
		\hline 
		True $Z$ &  0 & 1000 & 830 \\ 
		\hline 
		$\hat{Z}$ & 0 & 340 & \textbf{621} \\ 
		\hline 
		\end{tabular} 
		\caption{number of consistent model found ($Y$ depending on $X_1,X_2$ and only them) on $1000$ tries}
		\end{table}

		$299$ times on $1000$ tries, the predictive model using $\hat{Z}$ is better than classical LASSO in terms of MSE \underline{and} consistent (classical LASSO is never consistent).

		We also made the same experiment but with $X_1,X_2$ (and consequently $X_3$) following gaussian mixtures (to improve identifiability) randomly generated by our R package. 
		True $Z$ is now found $714$ times on $1000$ tries \label{testidentifiable}. So it confirms that non-gaussian models are easier to identify.

		\begin{table}[h!]
		\centering
		\begin{tabular}{|c|c|c|c|}
		\hline 
		 & Classical LASSO & Explicative & Predictive \\ 
		\hline 
		True $Z$ &  1.571029 & \textbf{1.569559} & \textbf{1.570801} \\ 
		\hline 
		$\hat{Z}$ & 1.005402 & 1.465768 & \textbf{1.005066} \\ 
		\hline 
		\end{tabular} 
		\caption{MSE observed on a validation sample (1000 individuals)}
		\end{table}

		And when we look at the consistency :
		\begin{table}[h!]
		\centering	
		\begin{tabular}{|c|c|c|c|}
		\hline 
		 & Classical LASSO & Explicative & Predictive \\ 
		\hline 
		True $Z$ &  0 & 1000 & 789 \\ 
		\hline 
		$\hat{Z}$ & 0 & 714 & \textbf{608} \\ 
		\hline 
		\end{tabular} 
		\caption{number of consistent model found ($Y$ depending on $X_1,X_2$ and only them) on $1000$ tries}
		\end{table}

		$299$ times on $1000$ tries, the predictive model using $\hat{Z}$ is better than classical LASSO in terms of MSE \underline{and} consistent (classical LASSO is never consistent).

	\clearpage	
\section{Numerical results on real datasets}
This work takes place in steel industry context, with quality oriented objective : to understand and prevent quality problems on finished product, knowing the whole process. In particular, we focus on regression problems.
        \begin{center}
          \begin{tabular}{ccc}
         \includegraphics[width=70px,height=70px]{liquid.jpg} & \includegraphics[width=70px,height=70px]{Brame1.jpg} & \includegraphics[width=70px,height=70px]{Brame.jpg} \\
          	\includegraphics[width=70px,height=70px]{ecras_moy.jpg} &\includegraphics[width=70px,height=70px]{tcc2.jpg} & \includegraphics[width=70px,height=70px]{bobines.jpg}
          \end{tabular}
        \end{center}
        Industrial context often means specific issues :
  \begin{itemize}
			\item Highly correlated parameters (parameters depends on the targeted product, physical models,...).
			\item Sometimes more variables than individuals.
			\item Missing values
		\end{itemize}
				\begin{figure}[h]
		\caption{Missing values on a real industrial dataset}
		\centering
		\includegraphics[width=150px]{figures/missing.png} \label{missingplot}
		\end{figure}

\subsection{The dataset}
		\begin{itemize}
			\item variables from the whole process
			\item The stakes : hundreds euros per ton (for information: Dunkerque site produces up to 7.5 millions tons a year)
		\end{itemize}

		\begin{figure}[hbtp]
			\caption{Correlation matrix of the dataset ($p=293$, $n=3000$)}
			\centering
			\includegraphics[width=150px]{figures/exfos_blanc2.png}
			\end{figure}
			Some observed correlations with physical meaning :
			\begin{itemize}
				\item Width and Weight : 0.905
				\item Temperature before and after a tool : 0.983 
				\item Roughness of both faces : 0.919
				\item Mean \& Max of a curve : 0.911
			\end{itemize}
		The method was tested on 205 variables without missing values.

	\subsection{Results}
		The algorithm gives a structure with 82 subregressions with a mean of 5 regressors.
			Some found subregressions with physical meaning :
			\begin{itemize}
				\item Mean.weight = f (Min.weight , Max.weight , Sigma.weight ) and other same-shaped subregressions.
				\item Width = f (Mean.flow , Mean.speed.CC)

True Physical model (not linear) :\begin{itemize}
\item Width = flow / (speed * thickness) (thickness is constant)
\end{itemize} 
			\end{itemize}

			Some of the other subression represent physical models used to regulate the process and that were forgotten by the metallurgist we worked with.
		Found model has selected relevant variables (verified with metallurgist).

		\begin{figure}[hbtp]
		\begin{minipage}[c]{.40\linewidth}
			\includegraphics[width=200px]{figures/boxplotssregexfobrut.png}
   \end{minipage} \hfill
   \begin{minipage}[c]{.52\linewidth}
			\includegraphics[width=250px]{figures/histogramR2exfobrut.png}
   \end{minipage}
			\caption{Adjusted R2 of found subregressions (industrial dataset)}
			\centering
			\end{figure}
%		
%\begin{figure}[hbtp]
%			\caption{}
%			\centering
%			\includegraphics[width=200px]{figures/histogramR2exfobrut.png}
%			\end{figure}
%

				We used Elastic Net\cite{zou2005regularization} on this dataset for selection (get better results than LASSO). Here are the observed MSE on a $n=847$ validation sample.
Predictive model (sequential elastic net base on estimated structure and using all the variables) is $5,82\%$ better (Figure \ref{compareMSEexfos}) than elastic net computed on the whole dataset.
		\begin{figure}[h]
						\centering
						\label{compareMSEexfos}
						\includegraphics[width=350px]{figures/elasticnet.png}
						\caption{MSE comparison on industrial dataset}
				\end{figure}						
		\clearpage
\section{Conclusion}
	We have seen that correlations can lead to serious estimation and variable selection problems in linear regression and that in such a context, it can be useful to explicitly model the structure between the covariates and to use this structure (even sequentially) to avoid correlations issues. We also show that real industrial context faces this kind of situations so our model can help to interprete and predict physical phenomenon efficiently and to help to manage missing values. But for now we still need a full dataset to learn the structure between the covariates and the method only works with numerical values. Further work is needed to face these two challenges.

\bibliography{biblio}{}
\bibliographystyle{plain}
\end{document}

### 2014-01-24 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@268 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 {\it
-$^{1}$ ArcelorMittal Dunkerque, Inria Lille, Université de Lille 1, clement.thery@arcelormittal.com
+$^{1}$ ArcelorMittal Dunkerque, Inria Lille, Universit\'e de Lille 1, clement.thery@arcelormittal.com
 
@@
 \section{Le contexte}
-	La régression linéaire classique suppose l'indépendance des covariables. Les corr\'elations posent des probl \`emes.
+	La r\'egression lin\'eaire classique suppose l'ind\'ependance des covariables. Les corr\'elations posent des probl\`emes.
 	\begin{eqnarray}
@@
 
-\section{Le mod \`ele g\'en\'eratif}
-On dispose de $p$ variables $X$ fortement corrélées pour expliquer une variable réponse $Y$.
-On rend explicite les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
+\section{Le mod\`ele g\'en\'eratif}
+On dispose de $p$ variables $X$ fortement corr\'el\'ees pour expliquer une variable r\'eponse $Y$.
+On rend explicite les corr\'elations au sein de $X$ sous la forme d'une structure de sous-r\'egressions lin\'eaires $S=(p_2,I_2,p_1,I_1)$ d\'efinie ainsi :
 	\begin{eqnarray}
@@
 		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices des covariables qui expliquent $X^j$} \\
-		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables dépendantes} \\
+		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables d\'ependantes} \\
 		p_2&=& \sharp I_2 \\
@@
 	\end{eqnarray}
-	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendentes dans $X$ n'en expliquent pas d'autres.
+	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables d\'ependentes dans $X$ n'en expliquent pas d'autres.
 
@@
 			\textrm{where }		\tilde{A}_{I_2}&=&0 \\
-					\tilde{A}_{I_2^c}&=&A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j} \\
-					\tilde{\varepsilon}&=&\sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
+					\tilde{A}_{I_2^c}&=&A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j} 
 \end{eqnarray}
@@
  \subsubsection{initialisation}
-L'intialisation peut par exemple utiliser la matrice des corrélations et/ou la méthode du Graphical Lasso\cite{friedman2008sparse}.		
-La grande dimension de l'espace parcouru rend préférable l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (pour un temps de calcul égal).
-Accessoirement, les multiples chaînes permettent de paralléliser la recherche, ce qui peut être très appréciable.
+L'intialisation peut par exemple utiliser la matrice des corr\'elations et/ou la m\'ethode du Graphical Lasso\cite{friedman2008sparse}.		
+La grande dimension de l'espace parcouru rend pr\'ef\'erable l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (pour un temps de calcul \'egal).
+Accessoirement, les multiples chaînes permettent de parall\'eliser la recherche, ce qui peut être très appr\'eciable.
 
@@
 	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
-	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité
+	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non sym\'etrique mais ne remets  pas en cause l'homog\'en\'eit\'e
 
@@
 	CorReg est fonctionnel et disponible
-	Besoin d'\'elargir à la gestion des valeurs manquantes tr \`es pr\'esentes dans l'industrie
+	Besoin d'\'elargir à la gestion des valeurs manquantes tr\`es pr\'esentes dans l'industrie
 \section{Exemple de r\'ef\'erences bibliographiques}
 La n\'ecessit\'e de produire des r\'esum\'es clairs et bien
-r\'ef\'erenc\'es a \'et\'e d\'emontr\'ee par Achin et Quidont~(2000). Le
-r\'ecent article de Noteur~(2003) met en \'evidence \dots
+r\'ef\'erenc\'es a\'et\'e d\'emontr\'ee par Achin et Quidont~(2000). Le
+r\'ecent article de Noteur~(2003) met en\'evidence \dots
 %Quelques rappels :

### 2014-01-24 www/Rennes.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@269 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		p_2&=& \sharp I_2 \\
-		\forall 1\leq j\leq p :  p_1^j&=&\sharp I_1^j
+		p_1&=&(p_1^1,\dots,p_1^{p_2}) \textrm{ avec }p_1^j=\sharp I_1^j 
 	\end{eqnarray}

### 2014-01-24 www/Rennes.tex
reduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@270 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-		\subsection{Comparaison des structures}
-	On utilise le Bayesian Information criterion (BIC)~\cite{BIChuard}.
-But BIC tends to give too complex structures because we test a great range of models.
-Thus we choose to penalise the complexity a bit more with specific a priori laws(uniform laws for the number of subregression and the complexity of each subregression instead of uniform law on S) :
-	\begin{eqnarray}
-		P(S)&=&P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2) \\
-		P(I_1 | p_1,I_2,p_2)&=&\prod_{j =1}^{p_2}P(I_1^j|p_1^j,I_2,p_2) \\
-		P(I_1^j|p_1^j,I_2,p_2)&=&\left(\begin{array}{c}
-			p-p_2 \\ 
-			p_1^j
-			\end{array}  \right)^{-1} =\frac{p_1^j ! (p-p_2-p_1^j)!}{(p-p_2)!}\\
-		P(p_1|I_2,p_2)&=&\prod_{j =1}^{p_2}P(p_1^j|I_2,p_2)		\\
-		P(p_1^j|I_2,p_2)&=&\frac{1}{p-p_2}  \\
-		P(I_2|p_2)&=&\left(\begin{array}{c}
-			p \\ 
-			p_2
-			\end{array}  \right)^{-1}=\frac{p_2!(p-p_2)!}{p!}\\
-		P(p_2) &=&\frac{1}{p_2} \\
-		P(S)&=&\left(\prod_{j =1}^{p_2}\left(\begin{array}{c}
-			p-p_2 \\ 
-			p_1^j
-			\end{array}  \right)^{-1}\right) \left(\frac{1}{p-p_2}\right)^{p_2}\frac{p_2!(p-p_2)!}{p!}\frac{1}{p_2} \\
-			\ln P(S) &=& -\sum_{j=1}^{p_2}	\ln\left(\begin{array}{c}
-			p-p_2 \\ 
-			p_1^j
-			\end{array}  \right)
-			-p_2\ln (p-p_2)
-			-\ln\left(\begin{array}{c}
-			p \\ 
-			p_2
-			\end{array}  \right)
-			-\ln( p_2	)
-	\end{eqnarray}
-	Then we have 
+		\subsection{P\'enalisation de la vraisemblance}
+		On va s'appuie sur la vraisemblance p\'enalis\'ee de la structure à la manière du critère BIC~\cite{BIChuard}. 
 		\begin{eqnarray}
@@
 		%&=&BIC +\ln(P(S))+cste \\
-		BIC^*&=&BIC +\ln(P(S))
+		BIC^*&=&BIC +\ln(P(S)) \label{Bicstar}
 	\end{eqnarray}	
-	It increases penalty on complexity for $p_2<\frac{p}{2}$ thus in the following we will use $BIC*$ under this hypothesis (that becomes a constraint in the MCMC).			
-	\subsection{The Markov chain}
-	First we remember that $S$ is completely described with $I_1$ :
-	So we will only describe the variations in $I_1$ at each step and other parts of $S$ will follow according to the previous definition.
-	for each step, starting from $S \in \mathcal{S}$ we define a neighbourhood $\mathcal{V}_{S,j}$ with $j \sim \mathcal{U}(\{1,\dots,p\}) $ like this  :	
+	Pour \'eviter une surcomplexit\'e de la structure trouv\'ee, on peut alors faire des hypothèses a priori sur $P(S)$.
+Par exemple, au lieu de supposer l'\'equiprobabilit\'e pour tous les $S$, on peut supposer l'\'equiprobabilit\'e des $p_2$ et $p_1^j$, ce qui vient p\'enaliser davantage la compl\'exit\'e sous l'hypothèse $p_2<\frac{p}{2}$ (qui devient alors une contrainte suppl\'ementaire dans l'algorithme de recherche).
+	On a
+	\begin{eqnarray}
+		P(S)&=&P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2) % \\
+%		P(I_1 | p_1,I_2,p_2)&=&\prod_{j =1}^{p_2}P(I_1^j|p_1^j,I_2,p_2) \\
+%		P(I_1^j|p_1^j,I_2,p_2)&=&\left(\begin{array}{c}
+%			p-p_2 \\ 
+%			p_1^j
+%			\end{array}  \right)^{-1} =\frac{p_1^j ! (p-p_2-p_1^j)!}{(p-p_2)!}\\
+%		P(p_1|I_2,p_2)&=&\prod_{j =1}^{p_2}P(p_1^j|I_2,p_2)		\\
+%		P(p_1^j|I_2,p_2)&=&\frac{1}{p-p_2}  \\
+%		P(I_2|p_2)&=&\left(\begin{array}{c}
+%			p \\ 
+%			p_2
+%			\end{array}  \right)^{-1}=\frac{p_2!(p-p_2)!}{p!}\\
+%		P(p_2) &=&\frac{1}{p_2} \\
+%		P(S)&=&\left(\prod_{j =1}^{p_2}\left(\begin{array}{c}
+%			p-p_2 \\ 
+%			p_1^j
+%			\end{array}  \right)^{-1}\right) \left(\frac{1}{p-p_2}\right)^{p_2}\frac{p_2!(p-p_2)!}{p!}\frac{1}{p_2} \\
+%			\ln P(S) &=& -\sum_{j=1}^{p_2}	\ln\left(\begin{array}{c}
+%			p-p_2 \\ 
+%			p_1^j
+%			\end{array}  \right)
+%			-p_2\ln (p-p_2)
+%			-\ln\left(\begin{array}{c}
+%			p \\ 
+%			p_2
+%			\end{array}  \right)
+%			-\ln( p_2	)
+	\end{eqnarray}		
+%	It increases penalty on complexity for $p_2<\frac{p}{2}$ thus in the following we will use $BIC*$ under this hypothesis (that becomes a constraint in the MCMC).			
+	\subsection{Parcours Markovien de saut}
+	$S$ est entièrement d\'efini à partir de $I_1$ donc on se contente ici de modifier $I_1$ (les autres \'el\'ements de $S$ seront ensuite recalcul\'es selon leur d\'efinition).
+	A chaque \'etape, pour $S \in \mathcal{S}$ on d\'efinit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ de la manière suivante  :	
 	\begin{eqnarray}
@@
 	\end{eqnarray}	
-	With $S^{(i,j)}$ defined by the following algorithm :
+	avec $S^{(i,j)}$ obtenu selon l'algorithme :
 	\begin{itemize}
-		\item if $i \notin I_i^j$ (add): 
+		\item Si $i \notin I_i^j$ (ajout): 
 			\begin{itemize}
 				\item $I_1^j=I_1^j\cup \{i\}$
-				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
-				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
+				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : relaxation colonne)
+				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : relaxation ligne) 
 			\end{itemize}			 
-		\item if $i \in I_1^j$ (remove): $I_1^j=I_1^j\setminus \{i\}$
+		\item Sinon ( $i \in I_1^j$ (suppression)): $I_1^j=I_1^j\setminus \{i\}$
 	\end{itemize}
-	And then :
 

### 2014-01-24 www/Rennes.tex
reduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@271 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % editeur de texte utilise des caracteres accentues
-% \usepackage[latin1]{inputenc}
-
+\usepackage[utf8]{inputenc}
 %
@@
 
- Cl\'ement Th\'ery $^{1}$
+ Clément Théry $^{1}$
 \bigskip
@@
 {\it
-$^{1}$ ArcelorMittal Dunkerque, Inria Lille, Universit\'e de Lille 1, clement.thery@arcelormittal.com
+$^{1}$ ArcelorMittal Dunkerque, Inria Lille, Université de Lille 1, clement.thery@arcelormittal.com
 
@@
 
-{\bf R\'esum\'e.} La r\'egression lin\'eaire suppose en g\'en\'eral l'usage de variables explicatives ind\'ependantes.
-Les variables pr\'esentes dans les bases de donn\'ees d'origine industrielle sont souvent tr\`es fortement corr\'el\'ees (de par le process, diverses lois physiques, etc).
-Le mod\`ele g\'en\'eratif propos\'e consiste à expliciter les corr\'elations pr\'esentes sous la forme d'une de sous-r\'egressions lin\'eaires.
-La structure est ensuite utilis\'ee pour obtenir un mod\`ele lib\'er\'e des corr\'elations, facilement interpr\'etable et compatible avec les m\'ethodes de s\'election de variables.
-La structure de corr\'elations est d\'etermin\'ee à l'aide d'un algorithme de type MCMC.
-Un package R (CorReg) permet la mise en oeuvre de cette m\'ethode.
+{\bf Résumé.} La régression linéaire suppose en général l'usage de variables explicatives indépendantes.
+Les variables présentes dans les bases de données d'origine industrielle sont souvent très fortement corrélées (de par le process, diverses lois physiques, etc).
+Le modèle génératif proposé consiste à expliciter les corrélations présentes sous la forme d'une de sous-régressions linéaires.
+La structure est ensuite utilisée pour obtenir un modèle libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
+La structure de corrélations est déterminée à l'aide d'un algorithme de type MCMC.
+Un package R (CorReg) permet la mise en oeuvre de cette méthode.
 \smallskip
 
-{\bf Mots-cl\'es.} R\'egression, corr\'elations, industrie, s\'election de variables, mod\`eles g\'en\'eratifs, SEM (Structural Equation Model) \ldots
+{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs, SEM (Structural Equation Model) \ldots
 \bigskip\bigskip
@@
 \section{Le contexte}
-	La r\'egression lin\'eaire classique suppose l'ind\'ependance des covariables. Les corr\'elations posent des probl\`emes.
+	La régression linéaire classique suppose l'indépendance des covariables. Les corrélations posent des problèmes.
 	\begin{eqnarray}
 		Y&=&XA+\varepsilon \ \ \ \varepsilon\sim \mathcal{N}(0,\sigma^2) \\
-		Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont lin\'eairement corr\'el\'ees}
+		Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont linéairement corrélées}
 	\end{eqnarray}
 
-\section{Le mod\`ele g\'en\'eratif}
-On dispose de $p$ variables $X$ fortement corr\'el\'ees pour expliquer une variable r\'eponse $Y$.
-On rend explicite les corr\'elations au sein de $X$ sous la forme d'une structure de sous-r\'egressions lin\'eaires $S=(p_2,I_2,p_1,I_1)$ d\'efinie ainsi :
+\section{Le modèle génératif}
+On dispose de $p$ variables $X$ fortement corrélées pour expliquer une variable réponse $Y$.
+On rend explicite les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
 	\begin{eqnarray}
@@
 		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices des covariables qui expliquent $X^j$} \\
-		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables d\'ependantes} \\
+		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables dépendantes} \\
 		p_2&=& \sharp I_2 \\
@@
 	\end{eqnarray}
-	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables d\'ependentes dans $X$ n'en expliquent pas d'autres.
+	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendentes dans $X$ n'en expliquent pas d'autres.
 
@@
 
-		\subsection{P\'enalisation de la vraisemblance}
-		On va s'appuie sur la vraisemblance p\'enalis\'ee de la structure à la manière du critère BIC~\cite{BIChuard}. 
+	%	\subsection{Pénalisation de la vraisemblance}
+		On va s'appuie sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
 		\begin{eqnarray}
@@
 	\end{eqnarray}	
-	Pour \'eviter une surcomplexit\'e de la structure trouv\'ee, on peut alors faire des hypothèses a priori sur $P(S)$.
-Par exemple, au lieu de supposer l'\'equiprobabilit\'e pour tous les $S$, on peut supposer l'\'equiprobabilit\'e des $p_2$ et $p_1^j$, ce qui vient p\'enaliser davantage la compl\'exit\'e sous l'hypothèse $p_2<\frac{p}{2}$ (qui devient alors une contrainte suppl\'ementaire dans l'algorithme de recherche).
+	Pour éviter une surcomplexité de la structure trouvée, on peut alors faire des hypothèses a priori sur $P(S)$.
+Par exemple, au lieu de supposer l'équiprobabilité pour tous les $S$, on peut supposer l'équiprobabilité des $p_2$ et $p_1^j$, ce qui vient pénaliser davantage la compléxité sous l'hypothèse $p_2<\frac{p}{2}$ (qui devient alors une contrainte supplémentaire dans l'algorithme de recherche).
 	On a
@@
 %	It increases penalty on complexity for $p_2<\frac{p}{2}$ thus in the following we will use $BIC*$ under this hypothesis (that becomes a constraint in the MCMC).			
-	\subsection{Parcours Markovien de saut}
-	$S$ est entièrement d\'efini à partir de $I_1$ donc on se contente ici de modifier $I_1$ (les autres \'el\'ements de $S$ seront ensuite recalcul\'es selon leur d\'efinition).
-	A chaque \'etape, pour $S \in \mathcal{S}$ on d\'efinit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ de la manière suivante  :	
+%	\subsection{Parcours Markovien de saut}
+	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de modifier $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
+	A chaque étape, pour $S \in \mathcal{S}$ on définit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ :% de la manière suivante  :	
 	\begin{eqnarray}
@@
 			\begin{itemize}
-				\item $I_1^j=I_1^j\cup \{i\}$
-				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : relaxation colonne)
-				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : relaxation ligne) 
+				\item $I_1^j :=I_1^j\cup \{i\}$, et pour garder $I_1\cap I_2=\emptyset$ :
+				\item $I_1^i :=\emptyset$ et $I_1:=I_1 \setminus \{j\}$
 			\end{itemize}			 
@@
 
-	We have $|\mathcal{V}_{S,j}|=p$ but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the subregressions or the whole structure, a maximum number of sub-regressions, etc).
-CorReg allows to modify this neighbourhood to better fit users constraints.
+	On a donc $p$ candidats à chaque étape. Mais CorReg (notre package) permet à l'utilisateur de modifier ce voisinage.
 
@@
 	\end{equation}
-	We define ~\cite{BIChuard}, :
 	\begin{equation}
@@
 
-	Where $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $ is the gap between a structure and the worst structure in the neighbourhood in terms of BIC.
+	Where $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $.
 	\newline
@@
 		\end{displaymath}
-	The output will be the best structure seen in terms of BIC.
-If we have some knowledge about some sub-regressions (physical models for example) we can add them in the found structure.
-So the model is really expert-friendly.
+		La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire.
+		Le résultat obtenu est la meilleur structure rencontrée en termes de $BIC^*$ (vraisemblance pénalisée).
+Certaines parties de $S$ peuvent être contraintes pour insérer par exemple des modèles physiques.
+De ce fait, la méthode permet de tenir compte d'éventuels modèles experts.
 
- \subsubsection{initialisation}
-L'intialisation peut par exemple utiliser la matrice des corr\'elations et/ou la m\'ethode du Graphical Lasso\cite{friedman2008sparse}.		
-La grande dimension de l'espace parcouru rend pr\'ef\'erable l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (pour un temps de calcul \'egal).
-Accessoirement, les multiples chaînes permettent de parall\'eliser la recherche, ce qui peut être très appr\'eciable.
+L'intialisation peut se faire en utilisant la matrice des corrélations et/ou la méthode du Graphical Lasso\cite{friedman2008sparse}.		
+La grande dimension de l'espace parcouru rend préférable l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (pour un temps de calcul égal).
+Accessoirement, les multiples chaînes permettent de paralléliser la recherche, ce qui peut être très appréciable.
 
@@
 	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
-	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non sym\'etrique mais ne remets  pas en cause l'homog\'en\'eit\'e
+	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité
 
@@
 
-\section{R\'esultats}	
+\section{Résultats}	
 \section{Conclusion et perspectives}
 	CorReg est fonctionnel et disponible
-	Besoin d'\'elargir à la gestion des valeurs manquantes tr\`es pr\'esentes dans l'industrie
-\section{Exemple de r\'ef\'erences bibliographiques}
-La n\'ecessit\'e de produire des r\'esum\'es clairs et bien
-r\'ef\'erenc\'es a\'et\'e d\'emontr\'ee par Achin et Quidont~(2000). Le
-r\'ecent article de Noteur~(2003) met en\'evidence \dots
-%Quelques rappels :
-%%
-%\begin{center}
-%%
-%\begin{tabular}{lr} \hline
-%%
-%Accent aigu :              & \'e; \\
-%Accent grave :             &  \`a;\\
-%Accent circonflexe :       &  \^o mais \^{\i};\\
-%Tr\'emas :                 &  \"o mais \"{\i};\\
-%C\'edille :                &  \c{c}. \\ \hline
-%\end{tabular}
-%%
-%\end{center}
-%--------------------------------------------------------------------------
+	Besoin d'élargir à la gestion des valeurs manquantes très présentes dans l'industrie
+\section{Exemple de références bibliographiques}
+La nécessité de produire des résumés clairs et bien
+référencés aété démontrée par Achin et Quidont~(2000). Le
+récent article de Noteur~(2003) met enévidence \dots
+
 \section*{Bibliographie}
-\cite{marquardt1975ridge}
 \bibliography{biblio}{}
 \bibliographystyle{plain}
-à recopier dans le bon ordre comme demand\'e ci-dessous.
-\noindent [1] Auteurs (ann\'ee), Titre, revue, localisation.
-\noindent [2] Achin, M. et Quidont, C. (2000), {\it Th\'eorie des
+à recopier dans le bon ordre comme demandé ci-dessous.
+
+\noindent [1] Auteurs (année), Titre, revue, localisation.
+
+\noindent [2] Achin, M. et Quidont, C. (2000), {\it Théorie des
 Catalogues}, Editions du Soleil, Montpellier.
-\noindent [3] Noteur, U. N. (2003), Sur l'int\'er\^et des
-r\'esum\'es, {\it Revue des Organisateurs de Congr\`es}, 34, 67--89.
+
+\noindent [3] Noteur, U. N. (2003), Sur l'intér\^et des
+résumés, {\it Revue des Organisateurs de Congrès}, 34, 67--89.
 \end{document}

### 2014-01-24 www/Rennes.tex
resultats
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@272 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %
+\usepackage{graphicx}
+
+ \graphicspath{{figures/}}
+
 \begin{document}
@@
 	Y&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
-					&=& X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}=X\tilde{A}+ \tilde{\varepsilon}\label{Trueexpl} \\
-			\textrm{where }		\tilde{A}_{I_2}&=&0 \\
-					\tilde{A}_{I_2^c}&=&A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j} 
+					&=& X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}=X\tilde{A}+ \tilde{\varepsilon}\label{Trueexpl} 			
 \end{eqnarray}
@@
 This kind of zero means redundant information but the covariate can be correlated with the response variable.
-So we don't have the grouping effect (so we are more parsimonious ) and we don't suffer from false interpretation (LASSO would).
+So we don't have the grouping effect (so we are more parsimonious ) and we don't suffer from false interpretation (LASSO would~\cite{Zhao2006MSC}).
 			\item Variable selection methods can lead to get some exact zeros in $\hat{\tilde{A}}^{I_1}$.
@@
 \section{Résultats}
+
+Some observed correlations with physical meaning :
+			\begin{itemize}
+				\item Width and Weight : 0.905
+				\item Temperature before and after a tool : 0.983 
+				\item Roughness of both faces : 0.919
+				\item Mean \& Max of a curve : 0.911
+			\end{itemize}
+
+Some found subregressions with physical meaning :
+			\begin{itemize}
+				\item Mean.weight = f (Min.weight , Max.weight , Sigma.weight ) and other same-shaped subregressions.
+				\item Width = f (Mean.flow , Mean.speed.CC)
+
+True Physical model (not linear) :\begin{itemize}
+\item Width = flow / (speed * thickness) (thickness is constant)
+\end{itemize} 
+			\end{itemize}
+
+			Some of the other subregression represent physical models used to regulate the process and that were forgotten by the metallurgist we worked with.
+
+Exemple détaillé :
+
+		\begin{figure}	
+		\centering		
+			\includegraphics[width=100px]{figures/histcor_auchan.JPG} 
+			\caption{Histogramme des corrélations observées (117 individus pour 168 variables)}
+		\end{figure}
+
 \section{Conclusion et perspectives}

### 2014-01-24 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@274 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 {\Large
-	{\sc  Regression for correlated variables :\\ Application in steel industry}
+	{\sc  CorReg : Régression sur variables corrélées\\ et Application à l'industrie Sidérurgique}
 }
@@
 
-	\subsubsection{properties}
-	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
-	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité
+\section{Résultats}	
+Les données industrielles sont fortement corrélées de manière naturelle : Largeur et poids d'une brame d'acier ($\rho=0.905$), Température avant et après un outil (($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
 
-%	if 	$\forall k, \tilde{Z} \not\in \mathcal{V}_{Z,k} $ then 
-We want 
-		\begin{equation}
-			\mathcal{P}(S,\tilde{S})=\mathbf{1}_{[\exists j, \tilde{S} \in \mathcal{V}_{S,j} ]} P(\tilde{S}|X)
-		\end{equation}
-			Because the walk follows a regular and thus ergodic markov chain with a finite state space, it has exactly one stationary distribution \cite{grinstead1997introduction} : $\pi$ and every rows of $\operatorname{lim}_{k\rightarrow \infty}\mathcal{P}^k=W$ equals $\pi$.
+Certaines des sous-régressions obtenues par CorReg ont une interprétation physique:
+\begin{itemize}
+	\item Moyenne = f (Min , Max , Sigma ) pour des données courbes
+	\item Largeur du produit= f (débit de fonte , vitesse de la coulée continue)	
+Vrai modèle physique (non linéaire) :
 
-\section{Résultats}
-
-Some observed correlations with physical meaning :
-			\begin{itemize}
-				\item Width and Weight : 0.905
-				\item Temperature before and after a tool : 0.983 
-				\item Roughness of both faces : 0.919
-				\item Mean \& Max of a curve : 0.911
+	 Largeur = $\frac{\textrm{débit}}{\textrm{vitesse } \times \textrm{ épaisseur}}$ (Mais dans ce cas précis l'épaisseur est constante)
 			\end{itemize}
 
-Some found subregressions with physical meaning :
-			\begin{itemize}
-				\item Mean.weight = f (Min.weight , Max.weight , Sigma.weight ) and other same-shaped subregressions.
-				\item Width = f (Mean.flow , Mean.speed.CC)
+			D'autres sous-régressions traduisent des modèles physiques (modèles ballistiques, etc) utilisés pour réguler le process.
+\\
 
-True Physical model (not linear) :\begin{itemize}
-\item Width = flow / (speed * thickness) (thickness is constant)
-\end{itemize} 
-			\end{itemize}
-
-			Some of the other subregression represent physical models used to regulate the process and that were forgotten by the metallurgist we worked with.
-
-Exemple détaillé :
-
-		\begin{figure}	
-		\centering		
-			\includegraphics[width=100px]{figures/histcor_auchan.JPG} 
-			\caption{Histogramme des corrélations observées (117 individus pour 168 variables)}
+Exemple de régression finale :
+\begin{figure}[!h]
+	\begin{minipage}[c]{.40\linewidth}
+			\includegraphics[width=200px]{figures/histcor_auchan.JPG} 
+	\end{minipage} \hfill
+   \begin{minipage}[c]{.52\linewidth}
+		\begin{tabular}{|c|c|c|}
+		\hline 
+		  & MSE  & nombre de régresseurs  \\ 
+		\hline
+		LASSO (lars) & 0.80 & 54 \\ 
+		\hline 
+		CorReg (et lars) & 0.53 & 24  \\ 
+		\hline 
+		\end{tabular} 
+   \end{minipage}
+   \caption{résultats obtenus sur données réelles : $n=117$ et $p=168$.
+A gauche les corrélations entre variables explicatives, à droite les MSE et la complexité (nombre de variables explicatives) des modèles obtenus pour expliquer $Y$ }
 \end{figure}
@@
 \section{Conclusion et perspectives}
-	CorReg est fonctionnel et disponible
-	Besoin d'élargir à la gestion des valeurs manquantes très présentes dans l'industrie
+	CorReg est fonctionnel et disponible. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
+	La force de CorReg est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles simples (parsimonieux) et facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
+	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, très présentes dans l'industrie.
+Cet aspect est envisagé sérieusement pour la prochaine version de CorReg.
+
 \section{Exemple de références bibliographiques}
 La nécessité de produire des résumés clairs et bien
-référencés aété démontrée par Achin et Quidont~(2000). Le
-récent article de Noteur~(2003) met enévidence \dots
+référencés a été démontrée par Achin et Quidont~(2000). Le
+récent article de Noteur~(2003) met en évidence \dots
 

### 2014-01-24 www/Rennes.tex
traduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@275 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 {\Large
-	{\sc  CorReg : Régression sur variables corrélées\\ et Application à l'industrie Sidérurgique}
+	{\sc  CorReg : Régression sur variables Corrélées\\ et Application à l'industrie Sidérurgique}
 }
@@
 Les variables présentes dans les bases de données d'origine industrielle sont souvent très fortement corrélées (de par le process, diverses lois physiques, etc).
-Le modèle génératif proposé consiste à expliciter les corrélations présentes sous la forme d'une de sous-régressions linéaires.
+Le modèle génératif proposé consiste à expliciter les corrélations présentes sous la forme d'une structure de sous-régressions linéaires.
 La structure est ensuite utilisée pour obtenir un modèle libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
@@
 
-{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs, SEM (Structural Equation Model) \ldots
+{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs, SEM (Structural Equation Model), \ldots
 \bigskip\bigskip
@@
 
-{\bf Keywords.} Regression, correlations, industry, variable selection, generative models, Structural Equation Model \ldots
+{\bf Keywords.} Regression, correlations, industry, variable selection, generative models, Structural Equation Model, \ldots
 
@@
 \section{Le contexte}
-	La régression linéaire classique suppose l'indépendance des covariables. Les corrélations posent des problèmes.
+	La régression linéaire classique suppose l'indépendance des covariables. Les corrélations posent en effet des problèmes, tant au niveau de l'interprétation qu'en termes de variance des estimateurs.
 	\begin{eqnarray}
-		Y&=&XA+\varepsilon \ \ \ \varepsilon\sim \mathcal{N}(0,\sigma^2) \\
+		Y&=&XA+\varepsilon \ \ \textrm{ avec } \ \varepsilon\sim \mathcal{N}(0,\sigma^2) \\
 		Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont linéairement corrélées}
@@
 	\end{eqnarray}
-	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendentes dans $X$ n'en expliquent pas d'autres.
+	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendantes dans $X$ n'en expliquent pas d'autres.
 
 	On note $I_2^c=\{1,\dots,p\}\setminus I_2$
-Then our generative model can be written :
+Le modèle génératif s'écrit alors :
 \begin{eqnarray}
@@
 
-	On a donc $p$ candidats à chaque étape. Mais CorReg (notre package) permet à l'utilisateur de modifier ce voisinage.
+	On a donc $p$ candidats à chaque étape. Mais le package CorReg permet à l'utilisateur de modifier ce voisinage.
 

### 2014-01-24 www/Rennes.tex
translate
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@276 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{eqnarray}
-	Y_{|X,S}=Y_{|X}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ with } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
-	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ with } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
-    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ free law}	
+	Y_{|X,S}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
+	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
+    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ loi quelconque}	
 \end{eqnarray}
-Where $B_{I_1^j}^j$ is the $p_1^j$-sized vector of the coefficients of the subregression.
+Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression de $X^j$.
 
-We note that (\ref{MainR}) and (\ref{SR}) also give :
+On remarque que (\ref{MainR}) et (\ref{SR}) impliquent :
 \begin{eqnarray}

### 2014-01-24 www/Rennes.tex
translate
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@277 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{document}
+
+     \def\Var{{\rm Var}\,}
 
@@
 		Y&=&XA+\varepsilon \ \ \textrm{ avec } \ \varepsilon\sim \mathcal{N}(0,\sigma^2) \\
-		Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont linéairement corrélées}
+		\Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont linéairement corrélées}
 	\end{eqnarray}
@@
 \section{Estimateur}
-	Classical methods like Ordinary Least Squares (OLS) estimate $Y|X$ and obtain (Maximum Likelihood Estimation): 
+	Les Moindres Carrés Ordinaires (MCO) donnent (maximum de vraisemblance): 
 		\begin{equation}
-			\hat A = (X'X)^{-1}X'Y \textrm{ (ill-conditoned matrix to inverse)}
+			\hat A = (X'X)^{-1}X'Y \textrm{ (matrice à inverser mal conditionnée)}
 		\end{equation}
-		With following properties :
+		avec les propriétés suivantes :
 		\begin{eqnarray}
-			E[\hat{A}|X]&=&A \\
-			Var[\hat{A}|X]&=& \sigma_Y^2(X'X)^{-1}
+			\rm{E}[\hat{A}|X]&=&A \\
+			\Var[\hat{A}|X]&=& \sigma_Y^2(X'X)^{-1}
 		\end{eqnarray}				
-		And when correlations are strong, the matrix to invert is ill-conditioned and the variance explodes.
+		La variance de l'estimateur explose quand les corrélations sont fortes (matrice quasi-singulière).
 
-		Our idea is to reduce the variance so we explain $Y$ only with $X^{I_1}$ knowing (\ref{SR}) and (\ref{Trueexpl})
+		CorReg réduit la variance de en estimant $Y$ seulement avec $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
 			\begin{equation}
@@
 			\end{equation}							
-		So the new estimator simply is : 
+		Ainsi, l'estimateur devient : 
 		\begin{eqnarray}
@@
 		\end{eqnarray}
-		and we get the following properties :
+		avec les propriétés suivantes :
 		\begin{eqnarray}
-			E[\hat{\tilde{A}}|X]&=&\tilde{A} \\
-			Var[\hat{\tilde{A}}_{I_2^c}|X]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
-			Var[\hat{\tilde{A}}_{I_2}|X]&=& 0 
+			\rm{E}[\hat{\tilde{A}}|X]&=&\tilde{A} \\
+			\Var[\hat{\tilde{A}}_{I_2^c}|X]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
+			\Var[\hat{\tilde{A}}_{I_2}|X]&=& 0 
 		\end{eqnarray}
-		We see that the variance is reduced (no correlations and smaller matrix give better conditioning) for small values of $\sigma_j$ $i.e.$ strong correlations.
+		La variance est réduite(retrait des corrélations et réduction de la dimension améliorent drastiquement le conditionnement) pour les faibles valeurs de $\sigma_j$ $i.e.$ les fortes corrélations.
 
-		Both classical and our new estimators of $Y$ are unbiased (true model)\cite{saporta2006probabilites}.
+		Le modèle complet et le nôtre prédisent tous les deux $Y$ sans biais (vrai modèle)\cite{saporta2006probabilites}.
 	\\	
-			This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used. 
+	Ce nouveau modèle est de dimension réduite et consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
 		 \\		
-			The explicit structure between the covariates helps to understand the model and the complex link between the covariate and the response variable so we call this model explicative.
+		 La structure explicite entre les variables permet de mieux comprendre les phénomènes en jeu et la parsimonie du modèle facilite son interprétation.
 
-			When we use a variable selection method on it we obtain two kinds of 0 :
+		En ajoutant une étape de sélection de variable on obtient deux types de $0$ :
 			\begin{enumerate}

### 2014-01-24 www/Rennes.tex
translate
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@278 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
translate
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@279 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 			\begin{enumerate}
-			\item Because of the structure we coerce $\hat{\tilde{A}}^{I_2} = 0 $.
-This kind of zero means redundant information but the covariate can be correlated with the response variable.
-So we don't have the grouping effect (so we are more parsimonious ) and we don't suffer from false interpretation (LASSO would~\cite{Zhao2006MSC}).
-			\item Variable selection methods can lead to get some exact zeros in $\hat{\tilde{A}}^{I_1}$.
-This kind of zero means that implied covariate has no significant effect on the response variable.
-And because variables in $X^{I_1}$ are orthogonal, we know that it is not misleading interpretation due to correlations.
+			\item La structure implique $\hat{\tilde{A}}^{I_2} = 0 $.
+Ce type de 0 est à interpréter comme une redondance d'information, mais la variable associée n'est pas pour autant jugée indépendante de $Y$.
+On obtient un modèle parcimonieux sans effet groupe, mais sans erreur d'interprétation, contrairement au LASSO ~\cite{Zhao2006MSC} car on connaît la signification de ce type de $0$.
+			\item Une phase de sélection de variables peut également produire des coefficients nuls dans $\hat{\tilde{A}}^{I_1}$.
+Ces $0$ sont alors des $0$ d'indépendance vis-à-vis de la réponse qui s'interprètent donc classiquement.Et comme les variables de $X^{I_1}$ sont orthogonales, on peut avoir confiance dans cette interprétation.
 			\end{enumerate}
+		Le modèle obtenu est donc sans biais de prédiction, parcimonieux, et consistant en interprétation.	
 \section{Recherche de structure}
@@
 	%	\subsection{Pénalisation de la vraisemblance}
-		On va s'appuie sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
+		On va s'appuier sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
 		\begin{eqnarray}
@@
 
-	We make a first approximation (\ref{Bicstar}) : 
+	On fait l'approximation suivante  (\ref{Bicstar}) : 
 	\begin{equation}
@@
 
-	Where $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $.
+	où $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $.
 	\newline
 
-	 And then we can note $\forall (S,\tilde{S}) \in \mathcal{S}^2 $ :
+	 Ce qui donne enfin $\forall (S,\tilde{S}) \in \mathcal{S}^2 $ :
 		\begin{displaymath}

### 2014-01-27 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@280 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The structure of correlations is found with an MCMC algorithm.
-A R package (CorReg) implements this new method.
+An R package (CorReg) implements this new method.
 \smallskip
@@
 \section{Le modèle génératif}
-On dispose de $p$ variables $X$ fortement corrélées pour expliquer une variable réponse $Y$.
+On dispose de $X=(X^1,\dots,X^p)$ variables fortement corrélées pour expliquer une variable réponse $Y$.
 On rend explicite les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
@@
 
-		CorReg réduit la variance de en estimant $Y$ seulement avec $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
+		CorReg réduit la variance de l'estimateur en estimant $Y$ seulement avec $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
 			\begin{equation}
@@
 		\end{eqnarray}
-		La variance est réduite(retrait des corrélations et réduction de la dimension améliorent drastiquement le conditionnement) pour les faibles valeurs de $\sigma_j$ $i.e.$ les fortes corrélations.
+		La variance est réduite (retrait des corrélations et réduction de la dimension améliorent drastiquement le conditionnement) pour les faibles valeurs de $\sigma_j$ $i.e.$ les fortes corrélations.
 
@@
 	\end{equation}
-	\begin{equation}
-		q(\tilde{S},\mathcal{V}_{S,j})=\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))}
-	\end{equation}
+	\begin{eqnarray}
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \frac{1}{p} \sum_{j=1}^p q(\tilde{S},\mathcal{V}_{S,j}) \\
+	\textrm{où }	q(\tilde{S},\mathcal{V}_{S,j})&=&\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))} \\
+	\textrm{ avec} \Delta BIC(S,\mathcal{V}_{S,j})&=&BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \}
+	\end{eqnarray}
 
-	où $\Delta BIC(S,\mathcal{V}_{S,j})=BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \} $.
-	\newline
-
-	 Ce qui donne enfin $\forall (S,\tilde{S}) \in \mathcal{S}^2 $ :
-		\begin{displaymath}
-			\mathcal{P}(S,\tilde{S})= \frac{1}{p} \sum_{j=1}^p q(\tilde{S},\mathcal{V}_{S,j})
-		\end{displaymath}
 		La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire.
 		Le résultat obtenu est la meilleur structure rencontrée en termes de $BIC^*$ (vraisemblance pénalisée).
-Certaines parties de $S$ peuvent être contraintes pour insérer par exemple des modèles physiques.
-De ce fait, la méthode permet de tenir compte d'éventuels modèles experts.
 
 L'intialisation peut se faire en utilisant la matrice des corrélations et/ou la méthode du Graphical Lasso\cite{friedman2008sparse}.		
-La grande dimension de l'espace parcouru rend préférable l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (pour un temps de calcul égal).
-Accessoirement, les multiples chaînes permettent de paralléliser la recherche, ce qui peut être très appréciable.
+La grande dimension de l'espace parcouru rend préférable  (pour un temps de calcul égal) l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (permet aussi la parallélisation).
 
 \section{Résultats}	
-Les données industrielles sont fortement corrélées de manière naturelle : Largeur et poids d'une brame d'acier ($\rho=0.905$), Température avant et après un outil (($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
+Les données industrielles sont fortement corrélées de manière naturelle : Largeur et poids d'une brame ($\rho=0.905$), Température avant et après un outil (($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
 
@@
 
-			D'autres sous-régressions traduisent des modèles physiques (modèles ballistiques, etc) utilisés pour réguler le process.
-\\
+			D'autres sous-régressions traduisent des modèles physiques qui régulent le process.
 
-Exemple de régression finale :
+%Exemple de régression finale :
 \begin{figure}[!h]
@@
    \end{minipage}
-   \caption{résultats obtenus sur données réelles : $n=117$ et $p=168$.
-A gauche les corrélations entre variables explicatives, à droite les MSE et la complexité (nombre de variables explicatives) des modèles obtenus pour expliquer $Y$ }
+   \caption{résultats obtenus sur données réelles : $n=117$ et $p=168$.    }
 \end{figure}
@@
 Cet aspect est envisagé sérieusement pour la prochaine version de CorReg.
-
-\section{Exemple de références bibliographiques}
-La nécessité de produire des résumés clairs et bien
-référencés a été démontrée par Achin et Quidont~(2000). Le
-récent article de Noteur~(2003) met en évidence \dots
 

### 2014-01-27 www/Rennes.tex
upd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@281 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		 \\		
-		 La structure explicite entre les variables permet de mieux comprendre les phénomènes en jeu et la parsimonie du modèle facilite son interprétation.
+		 La structure explicite entre les variables permet de mieux comprendre les phénomènes en jeu et la parcimonie du modèle facilite son interprétation.
 
@@
 			\item Une phase de sélection de variables peut également produire des coefficients nuls dans $\hat{\tilde{A}}^{I_1}$.
-Ces $0$ sont alors des $0$ d'indépendance vis-à-vis de la réponse qui s'interprètent donc classiquement.Et comme les variables de $X^{I_1}$ sont orthogonales, on peut avoir confiance dans cette interprétation.
+Ces $0$ sont alors des $0$ d'indépendance vis-à-vis de la réponse qui s'interprètent donc classiquement.
+Et comme les variables de $X^{I_1}$ sont orthogonales, on peut avoir confiance dans cette interprétation.
 			\end{enumerate}
-		Le modèle obtenu est donc sans biais de prédiction, parcimonieux, et consistant en interprétation.	
+		Le modèle obtenu est donc sans biais de prédiction, parcimonieux et consistant en interprétation.	
 \section{Recherche de structure}
@@
 	%	\subsection{Pénalisation de la vraisemblance}
-		On va s'appuier sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
+		On va s'appuyer sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
 		\begin{eqnarray}
@@
 	CorReg est fonctionnel et disponible. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
-	La force de CorReg est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles simples (parsimonieux) et facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
+	La force de CorReg est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles simples (parcimonieux) et facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
 	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, très présentes dans l'industrie.

### 2014-01-27 www/Rennes.tex
upd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@282 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-	On a donc $p$ candidats à chaque étape. Mais le package CorReg permet à l'utilisateur de modifier ce voisinage.
+	On a donc $p$ candidats à chaque étape. Mais le package \texttt{CorReg} permet à l'utilisateur de modifier ce voisinage.
 
@@
 	\begin{equation}
-		P(S|X)\approx exp(BIC^*(S))
+		P(S|X)\approx \exp(BIC^*(S))
 	\end{equation}
@@
 			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \frac{1}{p} \sum_{j=1}^p q(\tilde{S},\mathcal{V}_{S,j}) \\
-	\textrm{où }	q(\tilde{S},\mathcal{V}_{S,j})&=&\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))} \\
+	\textrm{où }	q(\tilde{S},\mathcal{V}_{S,j})&=&\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))} \\
 	\textrm{ avec} \Delta BIC(S,\mathcal{V}_{S,j})&=&BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \}
@@
 		La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire.
-		Le résultat obtenu est la meilleur structure rencontrée en termes de $BIC^*$ (vraisemblance pénalisée).
+		Le résultat obtenu est la meilleure structure rencontrée en termes de $BIC^*$ (vraisemblance pénalisée).
 
@@
 \section{Résultats}	
-Les données industrielles sont fortement corrélées de manière naturelle : Largeur et poids d'une brame ($\rho=0.905$), Température avant et après un outil (($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
+Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
 
-Certaines des sous-régressions obtenues par CorReg ont une interprétation physique:
+Certaines des sous-régressions obtenues par CorReg ont une interprétation physique :
 \begin{itemize}
 	\item Moyenne = f (Min , Max , Sigma ) pour des données courbes
-	\item Largeur du produit= f (débit de fonte , vitesse de la coulée continue)	
+	\item Largeur du produit = f (débit de fonte , vitesse de la coulée continue)	
 Vrai modèle physique (non linéaire) :

### 2014-01-27 www/Rennes.tex
u
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@283 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 La structure de corrélations est déterminée à l'aide d'un algorithme de type MCMC.
-Un package R (CorReg) permet la mise en oeuvre de cette méthode.
+Un package R (\texttt{CorReg}) permet la mise en oeuvre de cette méthode.
 \smallskip
@@
 The structure of correlations is found with an MCMC algorithm.
-An R package (CorReg) implements this new method.
+An R package (\texttt{CorReg}) implements this new method.
 \smallskip
@@
 
-		CorReg réduit la variance de l'estimateur en estimant $Y$ seulement avec $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
+		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement avec $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
 			\begin{equation}
@@
 			D'autres sous-régressions traduisent des modèles physiques qui régulent le process.
+\\
 
-%Exemple de régression finale :
+Exemple de régression sur une variable réponse dans le cadre des données réelles :
 \begin{figure}[!h]

### 2014-01-27 www/Rennes.tex
wrong indice
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@284 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Les variables présentes dans les bases de données d'origine industrielle sont souvent très fortement corrélées (de par le process, diverses lois physiques, etc).
-Le modèle génératif proposé consiste à expliciter les corrélations présentes sous la forme d'une structure de sous-régressions linéaires.
-La structure est ensuite utilisée pour obtenir un modèle libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
+Le modèle génératif proposé ici consiste à expliciter les corrélations présentes sous la forme d'une structure de sous-régressions linéaires.
+La structure est ensuite utilisée pour obtenir un modèle parcimonieux libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
 La structure de corrélations est déterminée à l'aide d'un algorithme de type MCMC.
-Un package R (\texttt{CorReg}) permet la mise en oeuvre de cette méthode.
+Un package R (\textsc{CorReg}) permet la mise en oeuvre de cette méthode.
 \smallskip
@@
 The proposed generative model consists in explicit modeling of the correlations with a structure of sub-regressions between the covariates.
-This structure is then used to obtain a model with independent covariates, easily interpreted, and compatible with any variable selection method.
+This structure is then used to obtain a reduced model with independent covariates, easily interpreted, and compatible with any variable selection method.
 The structure of correlations is found with an MCMC algorithm.
-An R package (\texttt{CorReg}) implements this new method.
+An R package (\textsc{CorReg}) implements this new method.
 \smallskip
@@
 On dispose de $X=(X^1,\dots,X^p)$ variables fortement corrélées pour expliquer une variable réponse $Y$.
-On rend explicite les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
+On rend explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
 	\begin{eqnarray}
-		I_1&=&(I_1^1,\dots,I_1^{p_2}) \textrm{ avec}		\\
+		I_1&=&(I_1^1,\dots,I_1^{p}) \textrm{ avec}		\\
 		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices des covariables qui expliquent $X^j$} \\
@@
 
-	On a donc $p$ candidats à chaque étape. Mais le package \texttt{CorReg} permet à l'utilisateur de modifier ce voisinage.
+	On a donc $p$ candidats à chaque étape. Mais le package \textsc{CorReg} permet à l'utilisateur de modifier ce voisinage.
 
@@
 	\textrm{où }	q(\tilde{S},\mathcal{V}_{S,j})&=&\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))} \\
-	\textrm{ avec} \Delta BIC(S,\mathcal{V}_{S,j})&=&BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \}
+	\textrm{ avec } \Delta BIC(S,\mathcal{V}_{S,j})&=&BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \}
 	\end{eqnarray}
@@
 
-Certaines des sous-régressions obtenues par CorReg ont une interprétation physique :
+Sous-régressions obtenues par \textsc{CorReg} ayant interprétation physique :
 \begin{itemize}
@@
 \section{Conclusion et perspectives}
-	CorReg est fonctionnel et disponible. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
-	La force de CorReg est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles simples (parcimonieux) et facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
+	\textsc{CorReg} est fonctionnel et disponible. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
+	La force de \textsc{CorReg} est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles simples (parcimonieux) et facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
 	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, très présentes dans l'industrie.
-Cet aspect est envisagé sérieusement pour la prochaine version de CorReg.
+Cet aspect est envisagé sérieusement pour la prochaine version de \textsc{CorReg}.
 

### 2014-01-27 pkg/CorReg/R/newtheta.R
cleaning
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@285 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -7,7 +7,7 @@
 #' @param A vector of the coefficient for the main regression
 #' @param lambda lagrange multiplicators
 #' @param nbit number of iteration for Newton
-#' @export
+# ' @export
 newtheta<-function(X=X,Z=Z,B=B,Sigma=Sigma,A=A,lambda=NULL,Atilde=Atilde,nbit=1){
    X=cbind(1,X)
    I2=which(colSums(Z)!=0)

### 2014-01-27 pkg/CorReg/R/MakeJ.R
cleaning
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@285 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -4,7 +4,7 @@
 #' @param A Fixed p+1 coefficient vector for the main regression
 #' @param Z p squared binary matrix describing the structure (adjacency matrix)
 #' @param Sigma p2 vector of standard deviations for each subregression
-#'@export
+# '@export
 MakeJ<-function(X=X,Z=Z,B=B,Sigma=Sigma,A=A){
    X=cbind(1,X)
    I2=which(colSums(Z)!=0)

### 2014-01-27 pkg/CorReg/R/correg.R

@@ -18,7 +18,6 @@
 #' @param compl boolean to decide if the complete modele is computed
 #' @param expl boolean to decide if the explicative model is in the output
 #' @param pred boolean to decide if the predictive model is computed
-#' @param pred2 boolean to define if the new predictive model is computed
 #' @param select selection method in ("lar","lasso","forward.stagewise","stepwise", "elasticnet", "NULL","ridge")
 #' @param criterion the criterion used to compare the models
 #' @param K the number of clusters for cross-validation
@@ -212,44 +211,6 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = TRUE,
                                  beta = A_final)
         
       }
-      if(pred2){
-         if (is.null(B)) {
-            B = hatB(Z = Z, X = X)
-         }else{B=as.matrix(B)}
-         I1star=which(rowSums(Z)!=0)
-         I2=which(colSums(Z)!=0)
-         R=matrix(0,ncol=ncol(Z),nrow=length(I1star))
-         R[cbind(1:length(I1star),I1star)]=1
-         R[,I2]=B[I1star+1,I2]      
-        
-         beta_OLS=as.matrix(OLS(X=X,Y=Y,intercept=intercept)$beta)
-         R=as.matrix(R)
-         Xloc=X
-         if(intercept){
-            Xloc=cbind(1,X)
-            R=rbind(0,R)
-            R[1,I2]=B[1,I2]
-            r=res$expl$A[c(1,I1star+1)]
-            R=cbind(0,R)
-            R[1,1]=1
-         }else{
-            r=res$expl$A[I1star]
-         }
-         if(ncol(X)>nrow(X)){
-            id=diag(ncol(Xloc))
-            beta_cc=beta_OLS+ginv(t(Xloc)%*%Xloc)%*%t(R)%*%solve(R%*%ginv(t(Xloc)%*%Xloc)%*%t(R))%*%(r-R%*%beta_OLS)
-            nb0=ncol(X)-nrow(X)
-            if(nb0<=length(I1star)){
-               qui0=I1star[1:nb0]
-               Xloc=Xloc[,-(qui0+intercept)]
-            }
-         }else{
-            beta_cc=beta_OLS+solve(t(Xloc)%*%Xloc)%*%t(R)%*%solve(R%*%solve(t(Xloc)%*%Xloc)%*%t(R))%*%(r-R%*%beta_OLS)
-         }
[…]

### 2014-01-27 www/Rennes.tex
u
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@286 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\begin{eqnarray}
-		I_1&=&(I_1^1,\dots,I_1^{p}) \textrm{ avec}		\\
+		I_1&=&\{I_1^1,\dots,I_1^{p}\} \textrm{ avec}		\\
 		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices des covariables qui expliquent $X^j$} \\

### 2014-01-27 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@287 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-27 www/Rennes.tex
@@
 	\begin{eqnarray}
-		I_1&=&(I_1^1,\dots,I_1^{p}) \textrm{ avec}		\\
-		I_1^j &=& \{i |Z_{i,j}=1 \} \textrm{ indices des covariables qui expliquent $X^j$} \\
+		I_1&=&\{I_1^1,\dots,I_1^{p}\} \textrm{ avec}		\\
+		I_1^j &=&\{ \textrm{ indices des covariables qui expliquent $X^j$} \} \\ % \{i |Z_{i,j}=1 \} 
 		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables dépendantes} \\
@@
 	\end{eqnarray}
+%	Où $Z$ est la matrice d'adjacence du graphe qui définit les dépendances entres variables (SEM récursif).
 	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendantes dans $X$ n'en expliquent pas d'autres.

### 2014-01-27 www/article9.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@287 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
- 	We then observe that the reduced model only uses the partition given by the structure (who is explained and who explains) but not the structure itself (how they interacts).
+% 	We then observe that the reduced model only uses the partition given by the structure (who is explained and who explains) but not the structure itself (how they interacts).
 We also notice that even if variables are highly correlated, each can have a specific effect ($\varepsilon_X$) on the response variable.
-
- 	So in a second step we propose further usage of the structure, taking back correlated variables to estimate the residuals of the reduced model.
+% 	
+% 	So in a second step we propose further usage of the structure, taking back correlated variables to estimate the residuals of the reduced model.
 This estimation is also a linear regression that can take profit from any variable selection method.
@@
 It can be done with Constrained Least Squares \cite{amemiya1985advanced} or sequentially with classical OLS (non optimal but easier to use selection methods).
+% 	
+%
 
@@
 This structure is based on gaussian mixture models to fit better real datasets and to allow identifiability of the structure in terms of complexity (number of parameters).
-
- 	~\\
+ 	\\
 

### 2014-01-27 www/Rennes.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@288 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\begin{eqnarray}
-		I_1&=&\{I_1^1,\dots,I_1^{p}\} \textrm{ avec}		\\
-		I_1^j &=&\{ \textrm{ indices des covariables qui expliquent $X^j$} \} \\ % \{i |Z_{i,j}=1 \} 
+		I_1&=&\{I_1^1,\dots,I_1^{p}\} \textrm{ avec $I_1^j$ l'ensemble des indices des covariables qui expliquent $X^j$ \ }		\\
+%		I_1^j &=&\{ \textrm{ indices des covariables qui expliquent $X^j$} \} \\ % \{i |Z_{i,j}=1 \} 
 		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables dépendantes} \\
@@
 \end{eqnarray}
-Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression de $X^j$.
+Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$.
 
@@
 
-		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement avec $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
+		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement à partir de $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
 			\begin{equation}
@@
 	%	\subsection{Pénalisation de la vraisemblance}
-		On va s'appuyer sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
+		Le choix de structure s'appuie sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
 		\begin{eqnarray}
@@
 %	\subsection{Parcours Markovien de saut}
-	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de modifier $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
-	A chaque étape, pour $S \in \mathcal{S}$ on définit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ :% de la manière suivante  :	
+
+	A chaque étape, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ :% de la manière suivante  :	
 	\begin{eqnarray}
@@
 	\end{itemize}
-
 	On a donc $p$ candidats à chaque étape. Mais le package \textsc{CorReg} permet à l'utilisateur de modifier ce voisinage.
+	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de modifier $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
 

### 2014-01-27 www/Rennes.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@289 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	%	\subsection{Pénalisation de la vraisemblance}
-		Le choix de structure s'appuie sur la vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
+		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
 		\begin{eqnarray}
@@
 
-	A chaque étape, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ :% de la manière suivante  :	
+	A chaque étape de l'algorithme MCMC, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ :% de la manière suivante  :
 	\begin{eqnarray}

### 2014-01-27 www/Rennes.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@290 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@291 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	On a donc $p$ candidats à chaque étape. Mais le package \textsc{CorReg} permet à l'utilisateur de modifier ce voisinage.
-	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de modifier $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
+	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de décrire les modifications dans $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
 
-	On fait l'approximation suivante  (\ref{Bicstar}) : 
+	On fait l'approximation suivante : 
 	\begin{equation}
@@
 	\end{equation}
+	On définit alors
 	\begin{eqnarray}
@@
 Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
-
-Sous-régressions obtenues par \textsc{CorReg} ayant interprétation physique :
+Exemples de Sous-régressions obtenues par \textsc{CorReg} ayant interprétation physique :
 \begin{itemize}
 	\item Moyenne = f (Min , Max , Sigma ) pour des données courbes
-	\item Largeur du produit = f (débit de fonte , vitesse de la coulée continue)	
+	\item Largeur du produit = f (débit de fonte , vitesse de la coulée continue)	\\
 Vrai modèle physique (non linéaire) :
@@
 
-			D'autres sous-régressions traduisent des modèles physiques qui régulent le process.
+			D'autres sous-régressions traduisent des modèles physiques qui régulent le process...
 \\
@@
 		\hline 
-		  & MSE  & nombre de régresseurs  \\ 
+		  & MSE  & Variables retenues  \\ 
 		\hline

### 2014-01-29 pkg/CorReg/R/BicZcurve.R
new function to choose p2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@295 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,39 @@
+#' Curve of the BIC for each possible p2 with a fixed Z
+#' @export
+BicZcurve<-function(X=X,Z=Z,Bic_null_vect=Bic_null_vect,plot=T,star=F,trunc=NULL){
+   p2=sum(colSums(Z)!=0)
+   curve=sum(BicZ(X=X,Z=Z,Bic_null_vect=Bic_null_vect,star=star))
+   
+   if(p2>0){
+      I2=which(colSums(Z)!=0)
+      sigmavect=R2Z(Z=Z,X=X,crit="R2",adj=T)
+      ordre=order(sigmavect[I2],decreasing=T)
+      for (i in 1:p2){
+         Zloc=Z;Zloc[,-I2[ordre[1:i]]]=0
+         curve=c(curve,sum(BicZ(X=X,Z=Zloc,Bic_null_vect=Bic_null_vect,star=star)))
+      }
+      if(plot){
+         plot(curve[-1])
+         abline(h=curve[1])
+      }
+   }   
+   quimin=which.min(curve)
+   if(quimin>1){
+      quimin=quimin-1
+      Zopt=Z;Zopt[,-I2[ordre[1:quimin]]]=0
+      if(plot){
+         abline(v=quimin)
+      }
+   }else{
+      Zopt=0*Z
+   }
+   if(!is.null(trunc)){
+      trunc=min(trunc,p2)
+      trunc=max(0,trunc)
+      Zopt=Z;Zopt[,-I2[ordre[1:trunc]]]=0
+      if(plot){
+         abline(v=quimin,col="red")
+      }
+   }
+   return(list(curve=curve,Zopt=Zopt))
+}

### 2014-01-29 pkg/CorReg/R/BicZcurve.R
more features
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@296 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -2,7 +2,7 @@
 #' @export
 BicZcurve<-function(X=X,Z=Z,Bic_null_vect=Bic_null_vect,plot=T,star=F,trunc=NULL){
    p2=sum(colSums(Z)!=0)
-   curve=sum(BicZ(X=X,Z=Z,Bic_null_vect=Bic_null_vect,star=star))
+   curve=sum(BicZ(X=X,Z=0*Z,Bic_null_vect=Bic_null_vect,star=star))
    
    if(p2>0){
       I2=which(colSums(Z)!=0)
@@ -32,7 +32,7 @@ BicZcurve<-function(X=X,Z=Z,Bic_null_vect=Bic_null_vect,plot=T,star=F,trunc=NULL
       trunc=max(0,trunc)
       Zopt=Z;Zopt[,-I2[ordre[1:trunc]]]=0
       if(plot){
-         abline(v=quimin,col="red")
+         abline(v=trunc,col="red")
       }
    }
    return(list(curve=curve,Zopt=Zopt))

### 2014-01-29 pkg/CorReg/R/generateur_melange.R
changing default value
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@297 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -24,7 +24,7 @@
 generateur_melange<-function(n=130,
                                 p=100,
                                 ratio=0.4,
-                                max_compl=3,
+                                max_compl=1,
                                 valid=1000,
                                 positif=0.6,
                                 sigma_Y=10,

### 2014-01-29 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@298 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
- Clément Théry $^{1}$
+  Clément Théry$^{1}$ \& Christophe Biernacki$^{2}$ \& Gaétan Loridant$^{3}$
 \bigskip
@@
 {\it
-$^{1}$ ArcelorMittal Dunkerque, Inria Lille, Université de Lille 1, clement.thery@arcelormittal.com
+$^{1}$ ArcelorMittal, Université Lille 1, Inria, CNRS, clement.thery@inria.fr
 
+$^{2}$ Université Lille 1, Inria, CNRS, christophe.biernacki@math.univ-lille1.fr
+
+$^{3}$ Etudes Industrielles ArcelorMittal Dunkerque, gaetan.loridant@arcelormittal.com\textbf{}
 }
@@
 
-\section{Le contexte}
-	La régression linéaire classique suppose l'indépendance des covariables. Les corrélations posent en effet des problèmes, tant au niveau de l'interprétation qu'en termes de variance des estimateurs.
+\section{Décorrélation par modèle génératif}
+	La régression linéaire classique suppose l'indépendance des covariables.
+Les corrélations posent en effet des problèmes, tant au niveau de l'interprétation qu'en termes de variance des estimateurs.
+La régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ qui explose si les colonnes de $X$ sont linéairement corrélées
+
+		On suppose le modèle génératif suivant :
 	\begin{eqnarray}
-		Y&=&XA+\varepsilon \ \ \textrm{ avec } \ \varepsilon\sim \mathcal{N}(0,\sigma^2) \\
-		\Var(\hat{A}|X)&=& \sigma^2(X'X)^{-1} \textrm{ explose si les colonnes de x sont linéairement corrélées}
-	\end{eqnarray}
-
-\section{Le modèle génératif}
-On dispose de $X=(X^1,\dots,X^p)$ variables fortement corrélées pour expliquer une variable réponse $Y$.
-On rend explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$ définie ainsi :
-	\begin{eqnarray}
-		I_1&=&\{I_1^1,\dots,I_1^{p}\} \textrm{ avec $I_1^j$ l'ensemble des indices des covariables qui expliquent $X^j$ \ }		\\
-%		I_1^j &=&\{ \textrm{ indices des covariables qui expliquent $X^j$} \} \\ % \{i |Z_{i,j}=1 \} 
-		I_2&=&\{j |\sharp I_1^j>0 \}  \textrm{ indices des variables dépendantes} \\
-		p_2&=& \sharp I_2 \\
-		p_1&=&(p_1^1,\dots,p_1^{p_2}) \textrm{ avec }p_1^j=\sharp I_1^j 
-	\end{eqnarray}
-%	Où $Z$ est la matrice d'adjacence du graphe qui définit les dépendances entres variables (SEM récursif).
-	On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendantes dans $X$ n'en expliquent pas d'autres.
-
-	On note $I_2^c=\{1,\dots,p\}\setminus I_2$
-Le modèle génératif s'écrit alors :
-\begin{eqnarray}
-	Y_{|X,S}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
+	Y_{|X,S}&=&XA+\varepsilon_Y= X^{I_2^c}A_{1}+X^{I_2}A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
 	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
@@
 \end{eqnarray}
-Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$.
+Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$, $I_1=\{I_1^1,\dots,I_1^{p}\}$, $I_2=\{j |\sharp I_1^j>0 \}$.
+On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendantes dans $X$ n'en expliquent pas d'autres.
+On note $p_2= \sharp I_2$, $p_1=(p_1^1,\dots,p_1^{p_2})$ et $I_2^c=\{1,\dots,p\}\setminus I_2$.
+
+On a donc rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$.
 
@@
 \begin{eqnarray}
-	Y&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
+	Y_{|X,S}&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
 					&=& X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}=X\tilde{A}+ \tilde{\varepsilon}\label{Trueexpl} 			
@@
 
-	A chaque étape de l'algorithme MCMC, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S,j}$ avec $j \sim \mathcal{U}(\{1,\dots,p\}) $ :% de la manière suivante  :
-	\begin{eqnarray}
-		\mathcal{V}_{S,j}&=&\{ S^{(i,j)} | 1\leq i\leq p \} \cup\{S \}
-	\end{eqnarray}	
-	avec $S^{(i,j)}$ obtenu selon l'algorithme :
-	\begin{itemize}
-		\item Si $i \notin I_i^j$ (ajout): 
-			\begin{itemize}
-				\item $I_1^j :=I_1^j\cup \{i\}$, et pour garder $I_1\cap I_2=\emptyset$ :
-				\item $I_1^i :=\emptyset$ et $I_1:=I_1 \setminus \{j\}$
-			\end{itemize}			 
-		\item Sinon ( $i \in I_1^j$ (suppression)): $I_1^j=I_1^j\setminus \{i\}$
-	\end{itemize}
-	On a donc $p$ candidats à chaque étape. Mais le package \textsc{CorReg} permet à l'utilisateur de modifier ce voisinage.
+	A chaque étape de l'algorithme MCMC, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S}$ de $p$ candidats , mais le package \textsc{CorReg} permet à l'utilisateur de modifier ce voisinage.
 	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de décrire les modifications dans $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
@@
 \section{Résultats}
+
 Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
@@
    \end{minipage}
-   \caption{résultats obtenus sur données réelles : $n=117$ et $p=168$.    }
+   \caption{résultats obtenus sur données réelles : $n=117$ et $p=168$. l'erreur est réduite d'un tiers alors que la complexité du modèle est divisée par $2,5$.   }
 \end{figure}

### 2014-01-29 pkg/CorReg/R/correg.R
remove useless parameter
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@300 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -34,7 +34,7 @@
 #' @param deltamin criterion to stop alternance
 #' 
 correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = TRUE, 
-                pred = TRUE, pred2=FALSE,prednew=FALSE,
+                pred = TRUE,prednew=FALSE,
                 select = "lar",
                 criterion = c("MSE", "BIC"),
                 X_test = NULL, Y_test = NULL, intercept = TRUE, 

### 2014-01-29 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@301 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Les corrélations posent en effet des problèmes, tant au niveau de l'interprétation qu'en termes de variance des estimateurs.
-La régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ qui explose si les colonnes de $X$ sont linéairement corrélées
+La régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ qui explose si les colonnes de $X$ sont linéairement corrélées.
 
@@
 	\begin{eqnarray}
-	Y_{|X,S}&=&XA+\varepsilon_Y= X^{I_2^c}A_{1}+X^{I_2}A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
+	Y_{|X,S}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
 	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
-    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ loi quelconque}	
+    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ mélanges gaussiens orthogonaux à $k_j$ composantes} 	
 \end{eqnarray}
@@
 	Y_{|X,S}&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
-					&=& X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}=X\tilde{A}+ \tilde{\varepsilon}\label{Trueexpl} 			
+					&=& X^{I_2^c}\alpha_{I_2^c}+ \varepsilon_{\alpha}=X\alpha+ \varepsilon_{\alpha} \label{Trueexpl} 			
 \end{eqnarray}
 \section{Estimateur}
-	Les Moindres Carrés Ordinaires (MCO) donnent (maximum de vraisemblance): 
+		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement à partir de $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
+			On a ainsi : 
 		\begin{equation}
-			\hat A = (X'X)^{-1}X'Y \textrm{ (matrice à inverser mal conditionnée)}
+			\hat{\alpha}_{I_2^c} = (X'_{I_2^c} X^{I_2^c})^{-1}X'_{I_2^c}Y \textrm{ et } \hat{\alpha}_{I_2} = 0
 		\end{equation}
-		avec les propriétés suivantes :
+		estimateur sans biais \cite{saporta2006probabilites} avec :
 		\begin{eqnarray}
-			\rm{E}[\hat{A}|X]&=&A \\
-			\Var[\hat{A}|X]&=& \sigma_Y^2(X'X)^{-1}
-		\end{eqnarray}				
-		La variance de l'estimateur explose quand les corrélations sont fortes (matrice quasi-singulière).
-
-		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement à partir de $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
-			\begin{equation}
-				Y= X^{I_2^c}\tilde{A}_{I_2^c}+ \tilde{\varepsilon}\label{explicatif}
-			\end{equation}							
-		Ainsi, l'estimateur devient : 
-		\begin{eqnarray}
-			\hat{\tilde{A}}_{I_2^c} &=& (X'_{I_2^c} X^{I_2^c})^{-1}X'_{I_2^c}Y \\
-			\hat{\tilde{A}}_{I_2} &=& 0
-		\end{eqnarray}
-		avec les propriétés suivantes :
-		\begin{eqnarray}
-			\rm{E}[\hat{\tilde{A}}|X]&=&\tilde{A} \\
 			\Var[\hat{\tilde{A}}_{I_2^c}|X]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
-			\Var[\hat{\tilde{A}}_{I_2}|X]&=& 0 
 		\end{eqnarray}
@@
 
-		Le modèle complet et le nôtre prédisent tous les deux $Y$ sans biais (vrai modèle)\cite{saporta2006probabilites}.
+		Le modèle complet et le nôtre prédisent tous les deux $Y$ sans biais (vrai modèle).
+La décorrélation se fait au prix d'un bruit blanc supplémentaire $\sum_{j \in I_2}\varepsilon_{j}A_{j} $ qui est d'autant plus faible que les corrélations sont fortes.
 	\\	
-	Ce nouveau modèle est de dimension réduite et consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
+	Ce nouveau modèle consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
 		 \\		
@@
 
-		En ajoutant une étape de sélection de variable on obtient deux types de $0$ :
-			\begin{enumerate}
-			\item La structure implique $\hat{\tilde{A}}^{I_2} = 0 $.
-Ce type de 0 est à interpréter comme une redondance d'information, mais la variable associée n'est pas pour autant jugée indépendante de $Y$.
-On obtient un modèle parcimonieux sans effet groupe, mais sans erreur d'interprétation, contrairement au LASSO ~\cite{Zhao2006MSC} car on connaît la signification de ce type de $0$.
-			\item Une phase de sélection de variables peut également produire des coefficients nuls dans $\hat{\tilde{A}}^{I_1}$.
-Ces $0$ sont alors des $0$ d'indépendance vis-à-vis de la réponse qui s'interprètent donc classiquement.
-Et comme les variables de $X^{I_1}$ sont orthogonales, on peut avoir confiance dans cette interprétation.
-			\end{enumerate}
+		En ajoutant une étape de sélection de variable on obtient deux types de $0$ : les $0$ de corrélation issus de la structure et qui sont à interpréter comme des $0$ de redondance d'information (qui ne signifient donc en rien l'indépendance avec $Y$) et les $0$ de sélection issus de l'éventuelle méthode de sélection de variables (type LASSO) et qui sont à interpréter comme l'indépendance entre la variable explicative concernée et la variable réponse.
+			\\
 		Le modèle obtenu est donc sans biais de prédiction, parcimonieux et consistant en interprétation.	
@@
 	%	\subsection{Pénalisation de la vraisemblance}
-		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}. 
+		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}, mais en prenant comme loi a priori sur $S$ une loi uniforme hiérarchique $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$ plutôt qu'une loi uniforme simple.
 		\begin{eqnarray}
 		P(S|X)&\propto &P(X|S)P(S) \\
-		\ln(P(S|X))&=&\ln(P(X|S))+\ln(P(S))+cste \\
+%		\ln(P(S|X))&=&\ln(P(X|S))+\ln(P(S))+cste \\
 		%&=&BIC +\ln(P(S))+cste \\
@@
 	\end{eqnarray}	
-	Pour éviter une surcomplexité de la structure trouvée, on peut alors faire des hypothèses a priori sur $P(S)$.
-Par exemple, au lieu de supposer l'équiprobabilité pour tous les $S$, on peut supposer l'équiprobabilité des $p_2$ et $p_1^j$, ce qui vient pénaliser davantage la compléxité sous l'hypothèse $p_2<\frac{p}{2}$ (qui devient alors une contrainte supplémentaire dans l'algorithme de recherche).
+	L'équiprobabilité ainsi supposée des $p_2$ et $p_1^j$ vient pénaliser davantage la compléxité sous l'hypothèse $p_2<\frac{p}{2}$ (qui devient alors une contrainte supplémentaire dans l'algorithme de recherche).
 	On a
-	\begin{eqnarray}
-		P(S)&=&P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2) % \\
-%		P(I_1 | p_1,I_2,p_2)&=&\prod_{j =1}^{p_2}P(I_1^j|p_1^j,I_2,p_2) \\
-%		P(I_1^j|p_1^j,I_2,p_2)&=&\left(\begin{array}{c}
-%			p-p_2 \\ 
-%			p_1^j
-%			\end{array}  \right)^{-1} =\frac{p_1^j ! (p-p_2-p_1^j)!}{(p-p_2)!}\\
-%		P(p_1|I_2,p_2)&=&\prod_{j =1}^{p_2}P(p_1^j|I_2,p_2)		\\
-%		P(p_1^j|I_2,p_2)&=&\frac{1}{p-p_2}  \\
-%		P(I_2|p_2)&=&\left(\begin{array}{c}
-%			p \\ 
-%			p_2
-%			\end{array}  \right)^{-1}=\frac{p_2!(p-p_2)!}{p!}\\
-%		P(p_2) &=&\frac{1}{p_2} \\
-%		P(S)&=&\left(\prod_{j =1}^{p_2}\left(\begin{array}{c}
-%			p-p_2 \\ 
-%			p_1^j
-%			\end{array}  \right)^{-1}\right) \left(\frac{1}{p-p_2}\right)^{p_2}\frac{p_2!(p-p_2)!}{p!}\frac{1}{p_2} \\
-%			\ln P(S) &=& -\sum_{j=1}^{p_2}	\ln\left(\begin{array}{c}
-%			p-p_2 \\ 
-%			p_1^j
-%			\end{array}  \right)
-%			-p_2\ln (p-p_2)
-%			-\ln\left(\begin{array}{c}
-%			p \\ 
-%			p_2
-%			\end{array}  \right)
-%			-\ln( p_2	)
-	\end{eqnarray}		
-%	It increases penalty on complexity for $p_2<\frac{p}{2}$ thus in the following we will use $BIC*$ under this hypothesis (that becomes a constraint in the MCMC).			
-%	\subsection{Parcours Markovien de saut}
 
-	A chaque étape de l'algorithme MCMC, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S}$ de $p$ candidats , mais le package \textsc{CorReg} permet à l'utilisateur de modifier ce voisinage.
-	$S$ est entièrement défini à partir de $I_1$ donc on se contente ici de décrire les modifications dans $I_1$. %(les autres éléments de $S$ seront ensuite recalculés selon leur définition).
+	A chaque étape de l'algorithme MCMC, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S}$ de $p$ candidats (le package \textsc{CorReg} permet à l'utilisateur de choisir parmi plusieurs types de voisinage).
 
@@
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \frac{1}{p} \sum_{j=1}^p q(\tilde{S},\mathcal{V}_{S,j}) \\
-	\textrm{où }	q(\tilde{S},\mathcal{V}_{S,j})&=&\mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(\frac{-1}{2}\Delta BIC(\tilde{S},\mathcal{V}_{S,j}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(\frac{-1}{2}\Delta BIC(S_l,\mathcal{V}_{S,j}))} \\
-	\textrm{ avec } \Delta BIC(S,\mathcal{V}_{S,j})&=&BIC(S)-\min\{BIC(\tilde{S})| \tilde{S} \in \mathcal{V}_{S,j} \}
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \frac{1}{p} \sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(\frac{-1}{2} BIC(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(\frac{-1}{2} BIC(S_l))} \\
 	\end{eqnarray}
@@
 \section{Résultats}
+
+\begin{tabular}{|c|c|c|c|c|}
+\hline 
+$n$ & $p_2$ & Complet & CORREG & meilleur \\ 
+\hline 
+30 & 0 & • & • & • \\ 
+\hline 
+30 & 8 & • & • & • \\ 
+\hline 
+30 & 24 & • & • & • \\ 
+\hline 
+50 & 0 & • & • & • \\ 
+\hline 
+50 & 8 & • & • & • \\ 
+\hline 
+50 & 24 & • & • & • \\ 
+\hline 
+\end{tabular}
 

### 2014-01-29 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@302 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@303 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 On a donc rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$.	
+Cette structure est identifiable au sens de la complexité de la structure et des mélanges gaussiens via un critère de type BIC sous certaines conditions simples.
+Du point de vue de l'interprétation, les cas non identifiables (triviaux) sont équivalents et ne posent pas de problème.
 
@@
 		\begin{eqnarray}
-			\Var[\hat{\tilde{A}}_{I_2^c}|X]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
+			\Var[\hat{\alpha}_{I_2^c}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
 		\end{eqnarray}
@@
 
-		En ajoutant une étape de sélection de variable on obtient deux types de $0$ : les $0$ de corrélation issus de la structure et qui sont à interpréter comme des $0$ de redondance d'information (qui ne signifient donc en rien l'indépendance avec $Y$) et les $0$ de sélection issus de l'éventuelle méthode de sélection de variables (type LASSO) et qui sont à interpréter comme l'indépendance entre la variable explicative concernée et la variable réponse.
+		En ajoutant une étape de sélection de variable on obtient deux types de $0$ : ceux de corrélation, issus de la structure et qui sont à interpréter comme des $0$ de redondance d'information (qui ne signifient donc en rien l'indépendance avec $Y$) et les $0$ de sélection, issus de l'éventuelle méthode de sélection de variables (type LASSO) et qui sont à interpréter comme l'indépendance entre la variable explicative concernée et la variable réponse.
 			\\
-		Le modèle obtenu est donc sans biais de prédiction, parcimonieux et consistant en interprétation.	
+		Le modèle obtenu est donc sans biais de prédiction pour $Y$, parcimonieux et consistant en interprétation.
+
 \section{Recherche de structure}
@@
 
+	En pratique, on commence par estimer pour chaque variable de $X$ sa densité sous l'hypothèse d'un mélange gaussien.
+On peut alors calculer la loi jointe de $X$ pour chaque structure réalisable rencontrée durant l'algorithme MCMC.
 \section{Résultats}
 
-\begin{tabular}{|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c|c|}
 \hline 
-$n$ & $p_2$ & Complet & CORREG & meilleur \\ 
+$n$ & $p_2$ & $\sharp$ complet & $\sharp$ \textsc{CorReg} &Complet & \textsc{CorReg} & meilleur \\ 
 \hline  
-30 & 0 & • & • & • \\ 
+30 & 0 & • & • & • &&\\
 \hline 
-30 & 8 & • & • & • \\ 
+30 & 8 & • & • & • &&\\ 
 \hline 
-30 & 24 & • & • & • \\ 
+30 & 16 & 22.16 & 18.84 & 65 033.19 & 1 202.52 & 0.69\\ 
 \hline 
-50 & 0 & • & • & • \\ 
+30 & 24 & 23.05 (3.7) & 17.29 (2.1) & 1 723.67 (9 300) & 987.75 (6 994)& 0.86 (0.34)\\ 
 \hline 
-50 & 8 & • & • & • \\ 
+50 & 0 & • & • & •&& \\ 
 \hline 
-50 & 24 & • & • & • \\ 
+50 & 8 & • & • & • &&\\ 
+\hline 
+50 & 24 & • & • & • && \\ 
 \hline 

### 2014-01-30 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@304 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \section{Résultats}	
+Pour chaque simulation présentée ci-dessous, chacune des configurations à été simulée $100$ fois.
+Les tableaux affichent les moyennes observées et écarts-types.
+Pour l'ensemble des simulations $p=40$, les $X$ indépendants suivent des lois normales dont la moyenne suit une loi de poisson de paramètre $\lambda=5$ et d'écart-type $\lambda$.
+Les $B_{i,j}$ suivent la même loi mais sont ensuite divisés par le produit $\lambda \times p_1^j$ pour que les moyennes des variables dépendantes ne trahissent pas leur dépendance.
+On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ valent tous 1 et \textsc{CorReg} a été contraint à ne chercher que des sous-régressions 2 à 2.
+Cependant, la normalité des variables à été cachée à
+\textsc{CorReg} qui a donc travaillé avec $K$ libre.
 
+\begin{table}
 \begin{tabular}{|c|c|c|c|c|c|c|}
 \hline 
-$n$ & $p_2$ & $\sharp$ complet & $\sharp$ \textsc{CorReg} &Complet & \textsc{CorReg} & meilleur \\ 
+$n$ & $p_2$ & $\sharp$ LARS & $\sharp$ \textsc{CorReg} &LARS & \textsc{CorReg} & meilleur \\ 
 \hline  
@@
 \hline 
-30 & 16 & 22.16 & 18.84 & 65 033.19 & 1 202.52 & 0.69\\ 
+30 & 23.98 (4.1) & 20.12 (3.8) & 18.84 & 204 896.03 & 497.97 (484.3) & 0.77 \\ 
 \hline 
-30 & 24 & 23.05 (3.7) & 17.29 (2.1) & 1 723.67 (9 300) & 987.75 (6 994)& 0.86 (0.34)\\ 
+30 & 24 & 23.05 (3.7) & 17.29 (2.1) & 1 723.67 (9 300) & 987.75 (6 994)& 0.86 \\ 
 \hline 
@@
 \hline 
+50 & 32 & 15.4 (6.2) & 11.06 (1.6) & 179.81 (105.7) & 133.49 (18.8) & 0.78 (0.4) \\ 
+\hline 
 \end{tabular} 
+\caption{Efficacité en prédiction de \textsc{CorReg} par rapport au LASSO suivant le LARS. $\sigma_j=0.01$}
+\end{table}
 

### 2014-01-30 pkg/CorReg/R/generateur_melange.R
change B generation and default values
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@305 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -31,7 +31,7 @@ generateur_melange<-function(n=130,
                                 sigma_sousreg=0.25,
                                 meanvar=NULL,
                                 sigmavar=NULL,
-                                lambda=5,#pour l enombre de composantes des m?langes gaussiens
+                                lambda=0,#pour l enombre de composantes des m?langes gaussiens
                                 Amax=15,
                                 lambdapois=5,#pour les valeurs des coefs
                                 gamma=F,
@@ -51,8 +51,8 @@ generateur_melange<-function(n=130,
   if(R>0){
     list_X2=sample(qui,R) # melange R individus pri au hasard 
     B[1,list_X2]=rpois(R,lambdapois)*(rep(-1,R)+2*rbinom(R,1,positif)) # 1ere ligne de B = ?????
-    for(j in list_X2){
-      B[sample(qui[-c(list_X2-1)],size=max_compl),j]=(1/max_compl)*rpois(max_compl,5)*(rep(-1,max_compl)+2*rbinom(max_compl,1,positif))
+    for(j in list_X2){#remplissage aléatoire de max_compl éléments de B sur chaque colonne à gauche
+      B[sample(qui[-c(list_X2-1)],size=max_compl),j]=(1/max_compl*lambdapois)*rpois(max_compl,lambdapois)*(rep(-1,max_compl)+2*rbinom(max_compl,1,positif))
     }
     #ajout de G
     G=diag(p+1)

### 2014-01-30 pkg/CorReg/R/BicZcurve.R
some doc
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@306 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,4 +1,10 @@
-#' Curve of the BIC for each possible p2 with a fixed Z
+#' Curve of the BIC for each possible p2 with a fixed Z and truncature of Z
+#' @param X matrix containing the dataset
+#' @param Z adjacency matrix (binary) describing the structure between the variables
+#' @param Bic_null_vect vector of the BIC for each variable. used when the variable is independant
+#' @param plot boolean to plot or not the curve
+#' @param star boolean to use BIC* (hierarchical uniform law on the structure)
+#' @param trunc number of sub-regression to keep (best R2). if NULL the min of BIC is kept
 #' @export
 BicZcurve<-function(X=X,Z=Z,Bic_null_vect=Bic_null_vect,plot=T,star=F,trunc=NULL){
    p2=sum(colSums(Z)!=0)

### 2014-01-30 pkg/CorReg/R/generateur_melange.R
bad weighting
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@307 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -52,7 +52,7 @@ generateur_melange<-function(n=130,
     list_X2=sample(qui,R) # melange R individus pri au hasard 
     B[1,list_X2]=rpois(R,lambdapois)*(rep(-1,R)+2*rbinom(R,1,positif)) # 1ere ligne de B = ?????
     for(j in list_X2){#remplissage aléatoire de max_compl éléments de B sur chaque colonne à gauche
-      B[sample(qui[-c(list_X2-1)],size=max_compl),j]=(1/max_compl*lambdapois)*rpois(max_compl,lambdapois)*(rep(-1,max_compl)+2*rbinom(max_compl,1,positif))
+      B[sample(qui[-c(list_X2-1)],size=max_compl),j]=(1/max_compl)*rpois(max_compl,lambdapois)*(rep(-1,max_compl)+2*rbinom(max_compl,1,positif))
     }
     #ajout de G
     G=diag(p+1)

### 2014-01-30 www/Rennes.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@308 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Les tableaux affichent les moyennes observées et écarts-types.
-Pour l'ensemble des simulations $p=40$, les $X$ indépendants suivent des lois normales dont la moyenne suit une loi de poisson de paramètre $\lambda=5$ et d'écart-type $\lambda$.
-Les $B_{i,j}$ suivent la même loi mais sont ensuite divisés par le produit $\lambda \times p_1^j$ pour que les moyennes des variables dépendantes ne trahissent pas leur dépendance.
+Pour l'ensemble des simulations $p=40$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
+Les $B_{i,j}$ suivent la même loi de poisson mais avec un signe aléatoire.
 On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ valent tous 1 et \textsc{CorReg} a été contraint à ne chercher que des sous-régressions 2 à 2.
-Cependant, la normalité des variables à été cachée à
-\textsc{CorReg} qui a donc travaillé avec $K$ libre.
+Cependant, la normalité des variables à été cachée à \textsc{CorReg} qui a donc travaillé avec $K$ libre.
+$Y$ dépend de $15$ variables choisies uniformément dans $X$ sans tenir compte de la structure.
 
 \begin{table}
+\centering
 \begin{tabular}{|c|c|c|c|c|c|c|}
@@
 \hline  
-30 & 0 & • & • & • &&\\
+30 & 0 & 25.49 (3) & 25.49 (3) & 472 385 (1 327 228) & 472 385 (1 327 228) & 1 \\
 \hline 
-30 & 8 & • & • & • &&\\ 
+30 & 8 & 21.03 (6.5) & 16.65 (5.8) & 640.61 (837.6) & 177271.56 (882162.4) & 0.61 \\ 
 \hline 
-30 & 23.98 (4.1) & 20.12 (3.8) & 18.84 & 204 896.03 & 497.97 (484.3) & 0.77 \\ 
+30 & 16 & 23.4 (4.4) & 18.57 (3.8) & 11083.81 (94368.9) & 381154.09 (1932655.1) & 0.66 \\ 
 \hline 

### 2014-01-30 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@309 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \centering
+
+\end{table}
+
+\begin{table}[h!]
+\centering 
 \begin{tabular}{|c|c|c|c|c|c|c|}
 \hline 
-$n$ & $p_2$ & $\sharp$ LARS & $\sharp$ \textsc{CorReg} &LARS & \textsc{CorReg} & meilleur \\ 
+$n$ & $p_2$ & $\sharp$ LARS & $\sharp$ \textsc{CorReg} &LARS & \textsc{CorReg} & $\tau$ \\ 
 \hline  
-30 & 0 & 25.49 (3) & 25.49 (3) & 472 385 (1 327 228) & 472 385 (1 327 228) & 1 \\
+30 & 0 & 25.49 (3) & 25.49 (3) & \textbf{472 385 (1 327 228)} & \textbf{472 385 (1 327 228)} & \textbf{1} \\
 \hline 
-30 & 8 & 21.03 (6.5) & 16.65 (5.8) & 640.61 (837.6) & 177271.56 (882162.4) & 0.61 \\ 
+30 & 8 & 20.88 (6.3) & 19.4 (6.8) & \textbf{519.18 (401.4)} & 864.18 (2093.9) &\textbf{ 0.57}  \\ 
 \hline 
-30 & 16 & 23.4 (4.4) & 18.57 (3.8) & 11083.81 (94368.9) & 381154.09 (1932655.1) & 0.66 \\ 
+30 & 16 & 23.83 (3.8) & 18.81 (3.6) & 25 671.62 (191 415.1) & \textbf{18 843.54 (184 681.9)} & \textbf{0.82} \\ 
 \hline 
-30 & 24 & 23.05 (3.7) & 17.29 (2.1) & 1 723.67 (9 300) & 987.75 (6 994)& 0.86 \\ 
+30 & 24 & 22.34 (4.3) & 16.91 (2.3) & 1 163.62 (3 944.4) & \textbf{920.79 (4 206.2)} & \textbf{0.8} \\ 
 \hline 
-50 & 0 & • & • & •&& \\ 
+30 & 32 & 17.83 (6.4) & 12.1 (2.5) & 779.32 (1494.9) & \textbf{201.41 (95.5)} & \textbf{0.81}\\ 
 \hline 
-50 & 8 & • & • & • &&\\ 
+50 & 0 & 24.53 (4.7) & 24.53 (4.7) & \textbf{247.05 (121.9)} & \textbf{247.05 (121.9)} & \textbf{1}  \\ 
 \hline 
-50 & 24 & • & • & • && \\ 
+50 & 8 & 16.31 (6.1) & 13.71 (3.9) & \textbf{183.38 (73.1)} & 1808.79 (14 178.1) & \textbf{0.66}\\ 
 \hline 
-50 & 32 & 15.4 (6.2) & 11.06 (1.6) & 179.81 (105.7) & 133.49 (18.8) & 0.78 (0.4) \\ 
+50 & 16 & 20.11 (4) & 16.71 (2.8) & \textbf{201.11 (81.3)} & 268.63 (1021.1) & \textbf{0.77}\\ 
+\hline 
+50 & 24 &20.61 (4) & 16.57 (2.5) & \textbf{199.36 (99.6)} & 242.14 (842.3) & \textbf{0.83} \\ 
+\hline 
+50 & 32 & 15.58 (6) & 11.46 (1.9) & 185.48 (127.3) & \textbf{137.85 (25.6)} & \textbf{0.76 } \\ 
+\hline 
+400 & 32 & 14.01 (3.6) & 11.22 (1.8) & 104.96 (4.8) & \textbf{103.51 (4.7)} & \textbf{0.76}\\ 
 \hline 
 \end{tabular} 
-\caption{Efficacité en prédiction de \textsc{CorReg} par rapport au LASSO suivant le LARS. $\sigma_j=0.01$}
+\caption{Efficacité en prédiction (echantillons de validation de 1 000 individus) de \textsc{CorReg} par rapport au LASSO suivant le LARS.
+$\tau$ est le taux de victoires (ou égalité) de \textsc{CorReg}, $\sigma_j=0.01$ et $\sigma_Y=10$.} \label{tableMSEsim}
 \end{table}
+Le tableau \ref{tableMSEsim} montre que \textsc{CorReg} est égal au LASSO en l'absence de structure et le bat quand les corrélations sont fortes. On note également que les modèles proposés par 
+\textsc{CorReg} sont parfois plus simples que le vrai modèle (15 variables) car celui-ci possède une forme réduite (estimée par \textsc{CorReg}).
+ON retrouve également le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
+On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle, comme pour le LASSO.
 

### 2014-01-30 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@315 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\begin{eqnarray}
-	Y_{|X,S}&=&XA+\varepsilon_Y= X^{I_2^c}A_{I_2^c}+X^{I_2}A_{I_2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
+	Y_{|X,S}&=&XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
 	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
@@
 \end{eqnarray}
-Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$, $I_1=\{I_1^1,\dots,I_1^{p}\}$, $I_2=\{j |\sharp I_1^j>0 \}$.
-On suppose $I_1\cap I_2=\emptyset$, $i.e.$ Les variables dépendantes dans $X$ n'en expliquent pas d'autres.
-On note $p_2= \sharp I_2$, $p_1=(p_1^1,\dots,p_1^{p_2})$ et $I_2^c=\{1,\dots,p\}\setminus I_2$.
+Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$,\\ $I_1=\{I_1^1,\dots,I_1^{p}\}$, $I_2=\{j |\sharp I_1^j>0 \}$, $X=(X^1,\dots,X^p)=(X_1,X_2)$ où $X_2=X^{I_2}$, $A=(A^1,\dots,A^p)=(A_1,A_2)$ où $A_2=A^{I_2}$.\\
+On suppose $I_1\cap I_2=\emptyset$, $i.e.$ les variables dépendantes dans $X$ n'en expliquent pas d'autres.
+On note $p_2= \sharp I_2$ et $p_1=(p_1^1,\dots,p_1^{p_2})$.
 
@@
 \begin{eqnarray}
-	Y_{|X,S}&=&X^{I_2^c} (A_{I_2^c}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y \\
-					&=& X^{I_2^c}\alpha_{I_2^c}+ \varepsilon_{\alpha}=X\alpha+ \varepsilon_{\alpha} \label{Trueexpl} 			
+	Y_{|X,S}&=&X_1 (A_{1}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
+					= X_1\alpha_{1}+ \varepsilon_{\alpha}=X\alpha+ \varepsilon_{\alpha} \label{Trueexpl} 			
 \end{eqnarray}
@@
 		\begin{equation}
-			\hat{\alpha}_{I_2^c} = (X'_{I_2^c} X^{I_2^c})^{-1}X'_{I_2^c}Y \textrm{ et } \hat{\alpha}_{I_2} = 0
+			\hat{\alpha}_{1} = (X'_{1} X_1)^{-1}X'_{1}Y \textrm{ et } \hat{\alpha}_{2} = 0
 		\end{equation}
@@
 		\begin{eqnarray}
-			\Var[\hat{\alpha}_{I_2^c}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{I_2^c} X^{I_2^c})^{-1} \\
+			\Var[\hat{\alpha}_{1}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1} \\
 		\end{eqnarray}
@@
 	%	\subsection{Pénalisation de la vraisemblance}
+
 		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}, mais en prenant comme loi a priori sur $S$ une loi uniforme hiérarchique $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$ plutôt qu'une loi uniforme simple.
@@
 Les $B_{i,j}$ suivent la même loi de poisson mais avec un signe aléatoire.
-On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ valent tous 1 et \textsc{CorReg} a été contraint à ne chercher que des sous-régressions 2 à 2.
-Cependant, la normalité des variables à été cachée à \textsc{CorReg} qui a donc travaillé avec $K$ libre.
+On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1.
+\textsc{CorReg}  a  travaillé avec $K$ et $p_1$ libres.
 $Y$ dépend de $15$ variables choisies uniformément dans $X$ sans tenir compte de la structure.
@@
 \centering
-
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline 
+$n$ & $p_2$ & $\sum_{j=1}^{p}\sharp(I_1^j \cap \hat{I_1^j})$ & $\sum_{j=1}^{p}\sharp(\hat{I_1^j} \setminus I_1^j)$ & $\sum_{j=1}^{p}\sharp(I_1^j \setminus \hat{I_1^j})$ & $p_2-\hat{p_2}$ & $\sharp(I_2 \cap \hat{I_2})$ & $\sharp(\hat{I_2} \setminus I_2)$ \\
+\hline  
+30 & 0 & 0 (0) & 0 (0) & 0 (0) & 0 (0) & 0 (0) & 0 (0) \\
+\hline 
+30 & 8 & 7.1 (3) & 15.13 (7.8) & 16.74 (3.1) & 3.44 (1) & 2.64 (1.2) & 1.92 (1.1)\\ 
+\hline 
+30 & 16 &  11.41 (4.2) & 28.39 (11.1) & 36.25 (4.2) & 8.77 (1.2) & 4.72 (1.5) & 2.51 (1.4)\\ 
+\hline 
+30 & 24 & 17.49 (5) & 42.69 (16.1) & 54.07 (5.1) & 14.03 (1.3) & 7.78 (2) & 2.19 (1.5)\\ 
+\hline 
+30 & 32 &19.89 (5.6) & 69.28 (14.8) & 75.5 (5.6) & 18.48 (1.3) & 12.35 (1.4) & 1.17 (1) \\ 
+\hline 
+\hline 
+50 & 0 &   &&&&&\\ 
+\hline 
+50 & 8 & &&&&&\\ 
+\hline 
+50 & 16 & &&&&&\\ 
+\hline 
+50 & 24 & &&&&&\\ 
+\hline 
+50 & 32 & &&&&&\\ 
+\hline 
+400 & 32 & &&&&&\\ 
+\hline 
+\end{tabular} 
+\caption{Comparaison de la vraie structure et de celle proposée par \textsc{CorReg}.
+Les variables dépendentes sont bien choisies mais les sous-régressions sont bruitées.
+Ici, les $p_1^j$ non nuls valent tous 3}
 \end{table}
@@
 \hline 
+\hline 
 50 & 0 & 24.53 (4.7) & 24.53 (4.7) & \textbf{247.05 (121.9)} & \textbf{247.05 (121.9)} & \textbf{1}  \\ 
@@
 50 & 32 & 15.58 (6) & 11.46 (1.9) & 185.48 (127.3) & \textbf{137.85 (25.6)} & \textbf{0.76 } \\ 
+\hline 
 \hline 

### 2014-01-31 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@318 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-On a donc rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(p_2,I_2,p_1,I_1)$.	
+On a donc rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(I_1,I_2,p_1,p_2)$.	
 Cette structure est identifiable au sens de la complexité de la structure et des mélanges gaussiens via un critère de type BIC sous certaines conditions simples.
@@
 		\begin{eqnarray}
-			\Var[\hat{\alpha}_{1}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1} \\
+			\Var[\hat{\alpha}_{1}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1}
 		\end{eqnarray}
@@
 \hline 
-$n$ & $p_2$ & $\sum_{j=1}^{p}\sharp(I_1^j \cap \hat{I_1^j})$ & $\sum_{j=1}^{p}\sharp(\hat{I_1^j} \setminus I_1^j)$ & $\sum_{j=1}^{p}\sharp(I_1^j \setminus \hat{I_1^j})$ & $p_2-\hat{p_2}$ & $\sharp(I_2 \cap \hat{I_2})$ & $\sharp(\hat{I_2} \setminus I_2)$ \\
+$n$ & $p_2$ & bon gauche & faux gauche    & MSEcompl  &    MSEexpl & MSEexplvraiZ\\ 
 \hline  
-30 & 0 & 0 (0) & 0 (0) & 0 (0) & 0 (0) & 0 (0) & 0 (0) \\
+30 & 0 & 0 (0) & 0.06 (0.2) & 2.6084e+07 (1.5e+07) & 2.616e+07  (1.5e+07) &  2.6084e+07  (1.5e+07) \\
 \hline 
-30 & 8 & 7.1 (3) & 15.13 (7.8) & 16.74 (3.1) & 3.44 (1) & 2.64 (1.2) & 1.92 (1.1)\\ 
+30 & 16 &  9.8 (1.7) & 6.17 (1.7) & 3019282.1 (4.5e+06) & 10688.46 (5.8e+04) & 663.35 (5.2e+02)\\ 
 \hline 
-30 & 16 &  11.41 (4.2) & 28.39 (11.1) & 36.25 (4.2) & 8.77 (1.2) & 4.72 (1.5) & 2.51 (1.4)\\ 
-\hline 
-30 & 24 & 17.49 (5) & 42.69 (16.1) & 54.07 (5.1) & 14.03 (1.3) & 7.78 (2) & 2.19 (1.5)\\ 
-\hline 
-30 & 32 &19.89 (5.6) & 69.28 (14.8) & 75.5 (5.6) & 18.48 (1.3) & 12.35 (1.4) & 1.17 (1) \\ 
+30 & 32 &22.21 (1.4e+00) & 5.21 (1.4e+00) & 466.81 (6.1e+02) & 160.46 (4.0e+01) & 143.55 (2.6e+01) \\ 
 \hline 
@@
 \hline 
-50 & 8 & &&&&&\\ 
-\hline 
 50 & 16 & &&&&&\\ 
 \hline 
-50 & 24 & &&&&&\\ 
-\hline 
 50 & 32 & &&&&&\\ 
-\hline 
+\hline \hline
 400 & 32 & &&&&&\\ 
-\hline 
-\end{tabular} 
-\caption{Comparaison de la vraie structure et de celle proposée par \textsc{CorReg}.
-Les variables dépendentes sont bien choisies mais les sous-régressions sont bruitées.
-Ici, les $p_1^j$ non nuls valent tous 3}
-\end{table}
-
-\begin{table}[h!]
-\centering 
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_2$ & $\sharp$ LARS & $\sharp$ \textsc{CorReg} &LARS & \textsc{CorReg} & $\tau$ \\ 
-\hline  
-30 & 0 & 25.49 (3) & 25.49 (3) & \textbf{472 385 (1 327 228)} & \textbf{472 385 (1 327 228)} & \textbf{1} \\
-\hline 
-30 & 8 & 20.88 (6.3) & 19.4 (6.8) & \textbf{519.18 (401.4)} & 864.18 (2093.9) &\textbf{ 0.57}  \\ 
-\hline 
-30 & 16 & 23.83 (3.8) & 18.81 (3.6) & 25 671.62 (191 415.1) & \textbf{18 843.54 (184 681.9)} & \textbf{0.82} \\ 
-\hline 
-30 & 24 & 22.34 (4.3) & 16.91 (2.3) & 1 163.62 (3 944.4) & \textbf{920.79 (4 206.2)} & \textbf{0.8} \\ 
-\hline 
-30 & 32 & 17.83 (6.4) & 12.1 (2.5) & 779.32 (1494.9) & \textbf{201.41 (95.5)} & \textbf{0.81}\\ 
-\hline 
-\hline 
-50 & 0 & 24.53 (4.7) & 24.53 (4.7) & \textbf{247.05 (121.9)} & \textbf{247.05 (121.9)} & \textbf{1}  \\ 
-\hline 
-50 & 8 & 16.31 (6.1) & 13.71 (3.9) & \textbf{183.38 (73.1)} & 1808.79 (14 178.1) & \textbf{0.66}\\ 
-\hline 
-50 & 16 & 20.11 (4) & 16.71 (2.8) & \textbf{201.11 (81.3)} & 268.63 (1021.1) & \textbf{0.77}\\ 
-\hline 
-50 & 24 &20.61 (4) & 16.57 (2.5) & \textbf{199.36 (99.6)} & 242.14 (842.3) & \textbf{0.83} \\ 
-\hline 
-50 & 32 & 15.58 (6) & 11.46 (1.9) & 185.48 (127.3) & \textbf{137.85 (25.6)} & \textbf{0.76 } \\ 
-\hline 
-\hline 
-400 & 32 & 14.01 (3.6) & 11.22 (1.8) & 104.96 (4.8) & \textbf{103.51 (4.7)} & \textbf{0.76}\\ 
 \hline 
@@
 \end{table}
+
 Le tableau \ref{tableMSEsim} montre que \textsc{CorReg} est égal au LASSO en l'absence de structure et le bat quand les corrélations sont fortes. On note également que les modèles proposés par 

### 2014-01-31 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@319 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@320 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \hline 
-30 & 0 & 0 (0) & 0.06 (0.2) & 2.6084e+07 (1.5e+07) & 2.616e+07  (1.5e+07) &  2.6084e+07  (1.5e+07) \\
+30 & 16 &  9.8 (1.7) & 6.17 (1.7) & 3 019 282.1 & 10 688.46  & 663.35\\ 
 \hline 
-30 & 16 &  9.8 (1.7) & 6.17 (1.7) & 3019282.1 (4.5e+06) & 10688.46 (5.8e+04) & 663.35 (5.2e+02)\\ 
-\hline 
-30 & 32 &22.21 (1.4e+00) & 5.21 (1.4e+00) & 466.81 (6.1e+02) & 160.46 (4.0e+01) & 143.55 (2.6e+01) \\ 
+30 & 32 &22.21 (1.4) & 5.21 (1.4) & 466.81  & 160.46  & 143.55) \\ 
 \hline 
 \hline 
-50 & 0 &   &&&&&\\ 
+50 & 0 & 0 (0) & 0.02 (0.1) & 617.51  & 914.93  & 617.51 \\ 
 \hline 
-50 & 16 & &&&&&\\ 
+50 & 16 & 10.43 (1.6) & 5.44 (1.6) & 302.88  & 543.84  & 194.28 \\ 
 \hline 
-50 & 32 & &&&&&\\ 
+50 & 32 & 22.2 (1.4) & 5.07 (1.6) & 178.78 & 129.41  & 120.93 \\ 
 \hline \hline
-400 & 32 & &&&&&\\ 
+400 & 32 &21.85 (1.5) & 5.35 (1.4) & 103.76  & 102.23  & 101.49 \\ 
 \hline 
 \end{tabular} 
-\caption{Efficacité en prédiction (echantillons de validation de 1 000 individus) de \textsc{CorReg} par rapport au LASSO suivant le LARS.
-$\tau$ est le taux de victoires (ou égalité) de \textsc{CorReg}, $\sigma_j=0.01$ et $\sigma_Y=10$.} \label{tableMSEsim}
+\caption{$Y$ ne dépend pas de $X_2$ et $\sigma_Y=10$.} \label{tableMSEsimp20}
 \end{table}

### 2014-01-31 pkg/CorReg/R/GM_Loglikelihood.R
CRAN check
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@321 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -13,7 +13,7 @@ GM_Loglikelihood<-function(Y=Y,X=X,B=B,sigma=sigma,M=NULL,mixmod,log=T,intercept
     M[is.na(M)]=1
     X[is.na(X)]=0
   }
-  return(.Call( "GM_likelihood",Y,X,B,sigma,M,mixmod$nbclust,mixmod$details,log,intercept, package = "CorReg"))
+  return(.Call( "GM_likelihood",Y,X,B,sigma,M,mixmod$nbclust,mixmod$details,log,intercept, PACKAGE = "CorReg"))
 #   res=0
 #   #on calcule d'abord la partie fixe qui sera commune (ajoutée) à toutes les classes
 #   quibon=which(!is.na(X))

### 2014-01-31 pkg/CorReg/R/correg.R

@@ -1,7 +1,5 @@
 #' Estimates the response variable using a structure
-#' @useDynLib CorReg
-#' @import Rcpp
-#' @import RcppEigen 
+#'  @import Rcpp
 #' @import Rmixmod
 #' @import lars
 #' @import elasticnet
@@ -9,6 +7,9 @@
 #' @import mclust 
 #' @import ridge
 #' @import MASS
+#' @import parcor
+# ' @import RcppEigen
+#' @useDynLib CorReg
 #' @export
 #' @param B the (p+1)xp matrix assiated to Z and that contains the parameters of the sub-regressions
 #' @param lambda parameter for elasticnet (quadratic penalty)
@@ -32,6 +33,7 @@
 #' @param prednew alternate optimisation for predictive
 #' @param nbalter number of alternance for prednew
 #' @param deltamin criterion to stop alternance
+#' 
 #' 
 correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = TRUE, 
                 pred = TRUE,prednew=FALSE,

### 2014-01-31 pkg/CorReg/R/BicLoc.R
CRAN check
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@321 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -2,6 +2,6 @@
 # ' 
 
 BicLoc<-function(X=X,Y=Y,intercept=T,sigma=T,SumSquare=F,methode=1){
-  ret= .Call( "BicLoc", X,Y,intercept,sigma,SumSquare,methode, package = "CorReg")
+  ret= .Call( "BicLoc", X,Y,intercept,sigma,SumSquare,methode, PACKAGE = "CorReg")
   return(ret)
 }

### 2014-01-31 pkg/CorReg/R/calcul_BIC_mixmod.R

@@ -6,7 +6,8 @@
 #' @param details boolean to give the details of the mixtures found
 #' @param max boolean. Use an heuristic to shrink nbclustmax according to the number of individuals in the dataset
 #' @param mclust boolean. Use mclust instead of Rmixmod
-calcul_BIC_mixmod<-function(X=X,nbclustmax=10,bla=FALSE,details=FALSE,max=TRUE,mclust=TRUE){
+#' @param nbini number of initial points for Rmixmod
+calcul_BIC_mixmod<-function(X=X,nbclustmax=10,bla=FALSE,details=FALSE,max=TRUE,mclust=TRUE,nbini=50){
   #X est la matrice sans constante
   n=nrow(X)
   if(max){
@@ -21,7 +22,7 @@ calcul_BIC_mixmod<-function(X=X,nbclustmax=10,bla=FALSE,details=FALSE,max=TRUE,m
   if(mclust==F){#si on veut utiliser mixmod
     for (i in 1:p){
       vect=X[!is.na(X[,i]),i]#donnees observees seulement
-      res=mixmodCluster(data=vect,criterion="BIC",nbCluster=c(1:nbclustmax))["bestResult"]
+      res=mixmodCluster(data=vect,criterion="BIC",nbCluster=c(1:nbclustmax),strategy=mixmodStrategy(nbTryInInit=nbini))["bestResult"]
       if(bla){print(res)}
       nbclust[i]=res[1]
       BIC_vect[i]=res[3]

### 2014-01-31 pkg/CorReg/R/correg.R
CRAN
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@322 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -8,7 +8,7 @@
 #' @import ridge
 #' @import MASS
 #' @import parcor
-# ' @import RcppEigen
+#' @import RcppEigen
 #' @useDynLib CorReg
 #' @export
 #' @param B the (p+1)xp matrix assiated to Z and that contains the parameters of the sub-regressions

### 2014-01-31 pkg/CorReg/src/BicLoc_cpp.cpp
CRAN
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@323 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -68,7 +68,7 @@ double BicLoc_cpp(Eigen::MatrixXd X,Eigen::MatrixXd Y,bool intercept,int methode
         double nbsum=Sum(0,0);
         
         double log_likelihood = (-Y.rows()/2)*log(2*PI*sumc*sumc)-nbsum/(2*sumc*sumc);
-        BicTheta=(-2*log_likelihood+(k*log(Y.rows())));
+        BicTheta=(-2*log_likelihood+(k*log(double(Y.rows()))));
       }   
 return BicTheta;
 }

### 2014-01-31 pkg/CorReg/src/BicTheta.cpp

@@ -60,7 +60,7 @@ BEGIN_RCPP
         
     float log_likelihood = (-vecY.rows()/2)*log(2*PI*sumc*sumc)-nbsum/(2*sumc*sumc);
         
-    float BicTheta=(-2*log_likelihood+(k*log(vecY.rows())));
+    float BicTheta=(-2*log_likelihood+(k*log(float(vecY.rows()))));
   
   
     return List::create(

### 2014-01-31 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@324 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Les tableaux affichent les moyennes observées et écarts-types.
-Pour l'ensemble des simulations $p=40$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
+Pour l'ensemble des simulations $p=40$, $\sigma_Y=10$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
 Les $B_{i,j}$ suivent la même loi de poisson mais avec un signe aléatoire.
@@
 \hline 
-$n$ & $p_2$ & bon gauche & faux gauche    & MSEcompl  &    MSEexpl & MSEexplvraiZ\\ 
+$n$ & $p_2$ & bon gauche & faux gauche    & LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} vrai $S$\\ 
 \hline 
-30 & 16 &  9.8 (1.7) & 6.17 (1.7) & 3 019 282.1 & 10 688.46  & 663.35\\ 
+30 & 16 &  8 & 5.39 & 2 466 225.35 & 13 796.03 & 588.9\\ 
 \hline 
-30 & 32 &22.21 (1.4) & 5.21 (1.4) & 466.81  & 160.46  & 143.55) \\ 
+30 & 32 & 17.05 & 2.7 & 979.16 & 196.33 & 141.2\\ 
 \hline 
 \hline 
-50 & 0 & 0 (0) & 0.02 (0.1) & 617.51  & 914.93  & 617.51 \\ 
+50 & 0 & 0 & 0 & 499.18 & 499.18 & 499.18 \\ 
 \hline 
-50 & 16 & 10.43 (1.6) & 5.44 (1.6) & 302.88  & 543.84  & 194.28 \\ 
+50 & 16 & 9.18 & 4.94 & 315.34 & 202.64 & 193.38 \\ 
 \hline 
-50 & 32 & 22.2 (1.4) & 5.07 (1.6) & 178.78 & 129.41  & 120.93 \\ 
+50 & 32 & 19.13 & 2.24 & 179.89 & 138.96 & 120.21 \\ 
 \hline \hline
-400 & 32 &21.85 (1.5) & 5.35 (1.4) & 103.76  & 102.23  & 101.49 \\ 
+400 & 32 &23.62 & 1.17 & 105 & 103.68 & 102.52 \\ 
 \hline 
 \end{tabular} 
-\caption{$Y$ ne dépend pas de $X_2$ et $\sigma_Y=10$.} \label{tableMSEsimp20}
+\caption{$Y$ ne dépend pas de $X_2$.\textsc{CorReg} gagne même en se trompant un peu sur $S$.} \label{tableMSEsimp20}
 \end{table}
 
-Le tableau \ref{tableMSEsim} montre que \textsc{CorReg} est égal au LASSO en l'absence de structure et le bat quand les corrélations sont fortes. On note également que les modèles proposés par 
-\textsc{CorReg} sont parfois plus simples que le vrai modèle (15 variables) car celui-ci possède une forme réduite (estimée par \textsc{CorReg}).
-ON retrouve également le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
+\begin{table}
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline 
+$n$ & $p_2$ & bon gauche & faux gauche    & LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} vrai $S$\\ 
+\hline 
+30 & 16 &  &  & &   &  \\ 
+\hline 
+30 & 32 &  &  & &   & \\ 
+\hline 
+\hline 
+50 & 0 &  &  & &   & \\ 
+\hline 
+50 & 16 &  &  & &   &\\ 
+\hline 
+50 & 32 &  &  & &   &\\ 
+\hline \hline
+400 & 32 &  &  & &   & \\ 
+\hline 
+\end{tabular} 
+\caption{$Y$  dépend  de $X$ entier.} \label{tableMSEsimtout}
+\end{table}
+
+Les tableaux \ref{tableMSEsimp20} et \ref{tableMSEsimtout} montrent que \textsc{CorReg} est égal au LASSO en l'absence de structure et le bat quand les corrélations sont fortes.
+On retrouve le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
 On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle, comme pour le LASSO.

### 2014-01-31 pkg/CorReg/src/OLS.cpp
SOLARIS update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@326 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -78,7 +78,7 @@ SEXP OLS(SEXP X,SEXP Y,SEXP intercept,SEXP sigma,SEXP Bic,SEXP methode){
         
         float log_likelihood = (-vecY.rows()/2)*log(2*PI*sumc*sumc)-nbsum/(2*sumc*sumc);
         
-        BicTheta=(-2*log_likelihood+(k*log(vecY.rows())));
+        BicTheta=(-2*log_likelihood+(k*float(log(vecY.rows()))));
         
       }
 

### 2014-01-31 pkg/CorReg/src/ProbaZ_cpp.cpp
SOLARIS update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@326 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -26,10 +26,10 @@ double ProbaZ_cpp(Eigen::MatrixXd Z){
    if(p2>0){
       for (int j = 0; j < p; j++){//pour chaque colonne de Z
         if(SumCol(j)>0){// si variable à gauche
-            logproba=logproba-log(R::choose((p-p2),SumCol(j)));
+            logproba=logproba-log(double(R::choose((p-p2),SumCol(j))));
         }
       }
-      logproba=logproba-log(p2)-p2*log(p-p2)-log(R::choose(p,p2));
+      logproba=logproba-log(double(p2))-double(p2)*log(double(p-p2))-log(double(R::choose(p,p2)));
    }  
    return logproba;
 }

### 2014-01-31 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@325 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@327 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 La régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ qui explose si les colonnes de $X$ sont linéairement corrélées.
+Les méthodes de sélection comme le LASSO~\cite{tibshiranilasso} muni du LAR~\cite{efron2004least} sont également touchées par ce problème de corrélation~\cite{Zhao2006MSC}.
 
@@
 \hline \hline
-400 & 32 &23.62 & 1.17 & 105 & 103.68 & 102.52 \\ 
+400 & 32 & 23.66 & 1.13 & 105.38 & 103.88 & 102.81\\ 
 \hline 
 \end{tabular} 
-\caption{$Y$ ne dépend pas de $X_2$.\textsc{CorReg} gagne même en se trompant un peu sur $S$.} \label{tableMSEsimp20}
+\caption{$Y$ ne dépend pas de $X_2$.\textsc{CorReg} gagne même en se trompant un peu sur $S$.} \label{tableMSEsimdroite}
 \end{table}
@@
 \hline 
-30 & 16 &  &  & &   &  \\ 
+30 & 16 & 8.48 & 4.88 & 3 511 185.23 & 10 686.62 & 738.89 \\ 
 \hline 
-30 & 32 &  &  & &   & \\ 
+30 & 32 & 16.89 & 2.78 & 565.51 & 189.54 & 139.24\\ 
 \hline 
 \hline 
-50 & 0 &  &  & &   & \\ 
+50 & 0 & 0 & 0 & 529.94 & 529.94 & 529.94 \\ 
 \hline 
-50 & 16 &  &  & &   &\\ 
+50 & 16 & 8.89 & 5.4 & 347.59 & 233.99 & 197.95\\ 
 \hline 
-50 & 32 &  &  & &   &\\ 
+50 & 32 & 18.95 & 2.44 & 163.7 & 139.39 & 121.56 \\ 
 \hline \hline
-400 & 32 &  &  & &   & \\ 
+400 & 32 & 23.49 & 1.06 & 104.52 & 103.6 & 102.67\\ 
 \hline 
@@
 \caption{$Y$  dépend  de $X$ entier.} \label{tableMSEsimtout}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline 
+$n$ & $p_2$ & bon gauche & faux gauche    & LAR  &    \textsc{CorReg} $\hat S$ & \textsc{CorReg} vrai $S$\\ 
+\hline 
+30 & 16 & 8.29 & 5 & 5 851.45 & 559.58 & 340.29\\ 
+\hline 
+30 & 32 & 17 & 2.59 & 893 & 196.01 & 135.78 \\ 
+\hline 
+\hline 
+50 & 16 &  8.98 & 5.19 & 201.56 & 164.58 & 162.49\\ 
+\hline 
+50 & 32 & 19.05 & 2.32 & 172.93 & 136.77 & 121.19\\ 
+\hline \hline
+400 & 32 & 23.51 & 1.09 & 104.49 & 103.02 & 102.26 \\ 
+\hline 
+\end{tabular} 
+\caption{$Y$  dépend  de $X_2$ uniquement. cas normalement défavorable à \textsc{CorReg}.} \label{tableMSEsimgauche}
 \end{table}

### 2014-01-31 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@328 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 La régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ qui explose si les colonnes de $X$ sont linéairement corrélées.
-Les méthodes de sélection comme le LASSO~\cite{tibshiranilasso} muni du LAR~\cite{efron2004least} sont également touchées par ce problème de corrélation~\cite{Zhao2006MSC}.
+Les méthodes de sélection comme le LASSO [5] muni du LAR [1] sont également touchées par ce problème de corrélation [6].
 
@@
 		\end{equation}
-		estimateur sans biais \cite{saporta2006probabilites} avec :
+		estimateur sans biais [4] avec :
 		\begin{eqnarray}
@@
 
-		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC~\cite{BIChuard}, mais en prenant comme loi a priori sur $S$ une loi uniforme hiérarchique $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$ plutôt qu'une loi uniforme simple.
+		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC [3], mais en prenant comme loi a priori sur $S$ une loi uniforme hiérarchique $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$ plutôt qu'une loi uniforme simple.
 		\begin{eqnarray}
@@
 
-L'intialisation peut se faire en utilisant la matrice des corrélations et/ou la méthode du Graphical Lasso\cite{friedman2008sparse}.		
+L'intialisation peut se faire en utilisant la matrice des corrélations et/ou la méthode du Graphical Lasso [2].		
 La grande dimension de l'espace parcouru rend préférable  (pour un temps de calcul égal) l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (permet aussi la parallélisation).
@@
 \section{Résultats}	
-Pour chaque simulation présentée ci-dessous, chacune des configurations à été simulée $100$ fois.
-Les tableaux affichent les moyennes observées et écarts-types.
-Pour l'ensemble des simulations $p=40$, $\sigma_Y=10$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
+Pour les simulations présentées dans les tableaux \ref{tableMSEsimdroite}, \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
+Les tableaux affichent le nombre de variable dépendantes trouvées, le nombre de variables jugées dépendantes à tort et les erreures moyennes en prédiction (MSE) sur $Y$ à partir d'échantillons de validation de 1 000 individus.
+Pour l'ensemble des simulations $p=40$, $\sigma_Y=10$, $\sigma=0.001$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
 Les $B_{i,j}$ suivent la même loi de poisson mais avec un signe aléatoire.
-On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1.
-\textsc{CorReg}  a  travaillé avec $K$ et $p_1$ libres.
-$Y$ dépend de $15$ variables choisies uniformément dans $X$ sans tenir compte de la structure.
+On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1 dans le vrai modèle.
+\textsc{CorReg}  a  travaillé avec $K$,$p_2$ et $p_1$ libres.
 
@@
 
-Les tableaux \ref{tableMSEsimp20} et \ref{tableMSEsimtout} montrent que \textsc{CorReg} est égal au LASSO en l'absence de structure et le bat quand les corrélations sont fortes.
+Les résultats (tableaux 1 à 3) montrent que \textsc{CorReg} est équivalent au LASSO en l'absence de corrélations et le bat quand les corrélations sont fortes.
 On retrouve le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
@@
 
-			D'autres sous-régressions traduisent des modèles physiques qui régulent le process...
+			D'autres sous-régressions traduisent des modèles physiques qui régulent le process industriel sur la base d'expertises métallurgiques.
 \\
@@
 \section*{Bibliographie}
-\bibliography{biblio}{}
-\bibliographystyle{plain}
-à recopier dans le bon ordre comme demandé ci-dessous.
+%\bibliography{biblio}{}
+%\bibliographystyle{plain}
+%à recopier dans le bon ordre comme demandé ci-dessous.
 
-\noindent [1] Auteurs (année), Titre, revue, localisation.
+\noindent [1] B. Efron, T. Hastie, I. Johnstone, and R. Tibshirani (2004). Least angle regression. {\it The
+Annals of statistics}, 32(2):407-499.
 
-\noindent [2] Achin, M. et Quidont, C. (2000), {\it Théorie des
-Catalogues}, Editions du Soleil, Montpellier.
+\noindent [2] J. Friedman, T. Hastie, and R. Tibshirani (2008). Sparse inverse covariance estimation with
+the graphical lasso.  {\it Biostatistics}, 9(3):432-441 .
 
-\noindent [3] Noteur, U. N. (2003), Sur l'intér\^et des
-résumés, {\it Revue des Organisateurs de Congrès}, 34, 67--89.
+\noindent [3] E. Lebarbier and T. Mary-Huard (2006). Une introduction au critère bic: fondements
+théoriques et interprétation.  {\it Journal de la SFdS }, 147(1):39-57.
+
+\noindent [4] G. Saporta (2006). Probabilités, analyse des données et statistique.  {\it Editions Technip}.
+
+\noindent[5] R. Tibshirani (1996). Regression shrinkage and selection via the lasso.  {\it Journal of the Royal
+Statistical Society}. Series B (Methodological), pages 267-288.
+
+\noindent [6] Peng Zhao and Bin Yu (December 2006). On model selection consistency of lasso. {\it J. Mach. Learn.
+Res.} 7:2541-2563.
+
+%
+%\noindent [1] Auteurs (année), Titre, revue, localisation.
+%
+%\noindent [2] Achin, M. et Quidont, C. (2000), {\it Théorie des
+%Catalogues}, Editions du Soleil, Montpellier.
+%
+%\noindent [3] Noteur, U. N. (2003), Sur l'intér\^et des
+%résumés, {\it Revue des Organisateurs de Congrès}, 34, 67--89.
 \end{document}

### 2014-02-03 www/Rennes.tex
update intro + concl
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@329 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-{\bf Résumé.} La régression linéaire suppose en général l'usage de variables explicatives indépendantes.
-Les variables présentes dans les bases de données d'origine industrielle sont souvent très fortement corrélées (de par le process, diverses lois physiques, etc).
-Le modèle génératif proposé ici consiste à expliciter les corrélations présentes sous la forme d'une structure de sous-régressions linéaires.
-La structure est ensuite utilisée pour obtenir un modèle parcimonieux libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
+{\bf Résumé.} La régression linéaire suppose en général l'usage de variables explicatives décorrélées, hypothèse souvent irréaliste pour les bases de données d'origine industrielle où de nombreuses corrélations sont dues au process, à des lois physiques, {\it etc}.
+Le modèle  proposé explicite les corrélations présentes sous la forme d'une famille de régressions linéaires entre covariables, permettant d'obtenir par marginalisation un modèle de régression parcimonieux libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
 La structure de corrélations est déterminée à l'aide d'un algorithme de type MCMC.
-Un package R (\textsc{CorReg}) permet la mise en oeuvre de cette méthode.
+Un package R (\textsc{CorReg}) permet la mise en oeuvre de cette méthode qui sera illustrée sur données simulées et sur données réelles issues de l'industrie sidérurgique.
 \smallskip
 
-{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs, SEM (Structural Equation Model), \ldots
+{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs, \ldots
 \bigskip\bigskip
@@
 
-{\bf Keywords.} Regression, correlations, industry, variable selection, generative models, Structural Equation Model, \ldots
+{\bf Keywords.} Regression, correlations, industry, variable selection, generative models,  \ldots
 
@@
 \end{tabular} 
-\caption{$Y$  dépend  de $X_2$ uniquement. cas normalement défavorable à \textsc{CorReg}.} \label{tableMSEsimgauche}
+\caption{$Y$  dépend  de $X_2$ uniquement (cas normalement défavorable à \textsc{CorReg} ).} \label{tableMSEsimgauche}
 \end{table}
@@
 
-Exemple de régression sur une variable réponse dans le cadre des données réelles :
+Exemple de régression sur une variable réponse dans le cadre de données réelles :
 \begin{figure}[!h]
@@
 \section{Conclusion et perspectives}
-	\textsc{CorReg} est fonctionnel et disponible. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
-	La force de \textsc{CorReg} est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles simples (parcimonieux) et facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
+	\textsc{CorReg} est fonctionnel et disponible sur R-forge. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
+	La force de \textsc{CorReg} est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles de régression simples et donc facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
 	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, très présentes dans l'industrie.
 Cet aspect est envisagé sérieusement pour la prochaine version de \textsc{CorReg}.
+En effet, le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire.
 
@@
 théoriques et interprétation.  {\it Journal de la SFdS }, 147(1):39-57.
-
-\noindent [4] G. Saporta (2006). Probabilités, analyse des données et statistique.  {\it Editions Technip}.
 

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@330 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Décorrélation par modèle génératif}
-	La régression linéaire classique suppose l'indépendance des covariables.
-Les corrélations posent en effet des problèmes, tant au niveau de l'interprétation qu'en termes de variance des estimateurs.
-La régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ qui explose si les colonnes de $X$ sont linéairement corrélées.
-Les méthodes de sélection comme le LASSO [5] muni du LAR [1] sont également touchées par ce problème de corrélation [6].
+\section{Introduction}
+	La régression linéaire classique suppose la décorrélation des covariables, source de problèmes en termes de variance des estimateurs.En effet, pour une variable réponse$Y$ et un ensemble de covariables $X \in \mathcal{R}^p$, la régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ dégénéré si les colonnes de $X$ sont linéairement corrélées.
+Les méthodes de sélection comme le LASSO [4] muni du LAR [1] sont elles-mêmes touchées par ce problème de corrélation [5].
+		\\
 
+		Notre idée est de modéliser explicitement les corrélations présentes entre covariables sous la forme d'une famille de régression entre celles-ci.
+Nous présenterons donc le modèle génératif associé et l'algorithme MCMC permettant de choisir la famille de régression à utiliser avant d'illustrer l'efficacité de la méthode sur des données simulées puis sur des données réelles.
+
+\section{Modèle supprimant les co-variables corrélées}		
 	On suppose le modèle génératif suivant :
-	\begin{eqnarray}
-	Y_{|X,S}&=&XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}\\
-	\forall j \in I_2 : \  X^j_{|X^{I_1^j},S}&=&X^{I_1^j}B_{I_1^j}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}\\
-    \forall j \notin I_2 : \ X^j &=& f(\theta_j) \textrm{ mélanges gaussiens orthogonaux à $k_j$ composantes} 	
-\end{eqnarray}
-Où $B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$,\\ $I_1=\{I_1^1,\dots,I_1^{p}\}$, $I_2=\{j |\sharp I_1^j>0 \}$, $X=(X^1,\dots,X^p)=(X_1,X_2)$ où $X_2=X^{I_2}$, $A=(A^1,\dots,A^p)=(A_1,A_2)$ où $A_2=A^{I_2}$.\\
+
+	\begin{itemize}
+		\item Régression principale :
+			 \begin{equation}
+			 	Y_{|X,S}=XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}
+			 \end{equation}
+		\item Famille de $p_2$ régressions entre covariables corrélées :
+			\begin{equation}
+			\forall j \in I_2 : \  X^j_{|X_1,S}=X_1B_{1}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}
+			\end{equation}
+		\item Mélanges gaussiens indépendants pour les covariables non corrélées :
+			\begin{equation}
+				\forall j \notin I_2 : \ X^j \sim \sum_{k=1}^{k_j}\pi_k\mathcal{N}(\mu_{k_j},\sigma_{k_j}^2)
+			\end{equation}
+	\end{itemize}
+
+Où %$B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$,\\ 
+$I_1=\{I_1^1,\dots,I_1^{p}\}$ est le vecteur des indices des variables à droite dans (\ref{SR}) , $I_2=\{j |\sharp I_1^j>0 \}$ est l'ensemble des indices des variables corrélées à gauche dans (\ref{SR}).
+On a donc une partition des données $X=(X_1,X_2)$ où $X_2=X^{I_2}$ et $X_1=X\setminus X_2$(pour alléger les notations).\\
 On suppose $I_1\cap I_2=\emptyset$, $i.e.$ les variables dépendantes dans $X$ n'en expliquent pas d'autres.
-On note $p_2= \sharp I_2$ et $p_1=(p_1^1,\dots,p_1^{p_2})$.
+On note $p_2= \sharp I_2$ le nombre de régressions entre covariables et $p_1=(p_1^1,\dots,p_1^{p_2})$.
 
-On a donc rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(I_1,I_2,p_1,p_2)$.	
-Cette structure est identifiable au sens de la complexité de la structure et des mélanges gaussiens via un critère de type BIC sous certaines conditions simples.
-Du point de vue de l'interprétation, les cas non identifiables (triviaux) sont équivalents et ne posent pas de problème.
+On a ainsi rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(I_1,I_2,p_1,p_2)$.	
+Ce modèle génératif est identifiable sous certaines conditions simples (sur les $k_i$) non détaillées ici.
 
@@
 \end{eqnarray}
-\section{Estimateur}
 		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement à partir de $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@331 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Le modèle  proposé explicite les corrélations présentes sous la forme d'une famille de régressions linéaires entre covariables, permettant d'obtenir par marginalisation un modèle de régression parcimonieux libéré des corrélations, facilement interprétable et compatible avec les méthodes de sélection de variables.
-La structure de corrélations est déterminée à l'aide d'un algorithme de type MCMC.
+La structure de corrélations est estimée à l'aide d'un algorithme de type MCMC.
 Un package R (\textsc{CorReg}) permet la mise en oeuvre de cette méthode qui sera illustrée sur données simulées et sur données réelles issues de l'industrie sidérurgique.
@@
 
-{\bf Abstract.} Linear regression generally suppose independence between the covariates.
-Datasets found in industrial context often contains many highly correlated covariates (due to the process, physcial laws, etc).
-The proposed generative model consists in explicit modeling of the correlations with a structure of sub-regressions between the covariates.
-This structure is then used to obtain a reduced model with independent covariates, easily interpreted, and compatible with any variable selection method.
+{\bf Abstract.} Linear regression generally suppose to have decorrelated covariates.
+This hypothesis is often irrealist with industrial datasets that contains many highly correlated covariates (due to the process, physcial laws,  {\it etc}).
+The proposed generative model consists in explicit modeling of the correlations with a family of linear regressions between the covariates permitting to obtain by marginalization a parsimonious correlation-free regression model, easily understandable and compatible with variable selection methods.
 The structure of correlations is found with an MCMC algorithm.
-An R package (\textsc{CorReg}) implements this new method.
+An R package (\textsc{CorReg}) implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry.
 \smallskip
@@
 
-On remarque que (\ref{MainR}) et (\ref{SR}) impliquent :
+On remarque que (\ref{MainR}) et (\ref{SR}) impliquent par simple intégration sur $X_2$, un modèle de régression en $Y$ s'exprimant uniquement en fonction des variables non corrélées $X_1$ :
 \begin{eqnarray}
 	Y_{|X,S}&=&X_1 (A_{1}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
-					= X_1\alpha_{1}+ \varepsilon_{\alpha}=X\alpha+ \varepsilon_{\alpha} \label{Trueexpl} 			
+					= X_1\alpha_{1}+ \varepsilon_{\alpha} \label{Trueexpl} 			
 \end{eqnarray}
-		\textsc{CorReg} réduit la variance de l'estimateur en estimant $Y$ seulement à partir de $X^{I_1}$, sachant (\ref{SR}) et (\ref{Trueexpl}).
-			On a ainsi : 
-		\begin{equation}
-			\hat{\alpha}_{1} = (X'_{1} X_1)^{-1}X'_{1}Y \textrm{ et } \hat{\alpha}_{2} = 0
-		\end{equation}
-		estimateur sans biais [4] avec :
+
+L'estimateur classique du Maximum de Vraisemblance de $\alpha$ est sans biais et s'exprime par	
 \begin{eqnarray}
+	\hat{\alpha}_{1}&=& (X'_{1} X_1)^{-1}X'_{1}Y \textrm{ et sa matrice de variance} \\
 	\Var[\hat{\alpha}_{1}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1}
 \end{eqnarray}
-		La variance est réduite (retrait des corrélations et réduction de la dimension améliorent drastiquement le conditionnement) pour les faibles valeurs de $\sigma_j$ $i.e.$ les fortes corrélations.
+		peut être notablement mieux conditionnée que celle de $\hat A$ initial.
 
-		Le modèle complet et le nôtre prédisent tous les deux $Y$ sans biais (vrai modèle).
-La décorrélation se fait au prix d'un bruit blanc supplémentaire $\sum_{j \in I_2}\varepsilon_{j}A_{j} $ qui est d'autant plus faible que les corrélations sont fortes.
+	En outre, ce nouveau modèle réduit consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
 		 \\		
-	Ce nouveau modèle consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
+		 Notons enfin que la structure explicite entre les variables permet de mieux comprendre les phénomènes en jeu et la parcimonie du modèle facilite son interprétation. 
 \\
-		 La structure explicite entre les variables permet de mieux comprendre les phénomènes en jeu et la parcimonie du modèle facilite son interprétation.
 
-		En ajoutant une étape de sélection de variable on obtient deux types de $0$ : ceux de corrélation, issus de la structure et qui sont à interpréter comme des $0$ de redondance d'information (qui ne signifient donc en rien l'indépendance avec $Y$) et les $0$ de sélection, issus de l'éventuelle méthode de sélection de variables (type LASSO) et qui sont à interpréter comme l'indépendance entre la variable explicative concernée et la variable réponse.
-			\\
-		Le modèle obtenu est donc sans biais de prédiction pour $Y$, parcimonieux et consistant en interprétation.
+		\textbf{\underline{Remarque}} : En ajoutant une étape de sélection de variables on obtient ainsi deux types de $0$ : ceux issus de l'étape de décorrélation et ceux issus de cette éventuelle étape de sélection de type LASSO.
 
-\section{Recherche de structure}
+\section{Estimation de la structure de corrélation}
 
@@
 
-		Le choix de structure s'appuie sur $BIC^*$, vraisemblance pénalisée de la structure à la manière du critère BIC [3], mais en prenant comme loi a priori sur $S$ une loi uniforme hiérarchique $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$ plutôt qu'une loi uniforme simple.
+		Le choix de structure s'appuie sur un critère noté $BIC^*$ et qui correspond à la vraisemblance pénalisée de la structure à la manière du critère BIC [3], mais en prenant comme loi {\it a priori} sur $S$ une loi uniforme hiérarchique $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$ plutôt qu'une loi uniforme simple.
+On a donc :
 		\begin{eqnarray}
-		P(S|X)&\propto &P(X|S)P(S) \\
+		%P(S|X)&\propto &P(X|S)P(S) \\
 %		\ln(P(S|X))&=&\ln(P(X|S))+\ln(P(S))+cste \\
@@
 	\end{eqnarray}	
-	L'équiprobabilité ainsi supposée des $p_2$ et $p_1^j$ vient pénaliser davantage la compléxité sous l'hypothèse $p_2<\frac{p}{2}$ (qui devient alors une contrainte supplémentaire dans l'algorithme de recherche).
+	L'équiprobabilité ainsi supposée des $p_2$ et $p_1^j$ vient pénaliser davantage la compléxité sous l'hypothèse $p_2<\frac{p}{2}$ , hypothèse réaliste sur le nombre de régressions entre covariables.
+La recherche du meilleur $S$ selon $BIC^*$ n'est pas un problème simple et on va s'appuyer sur un algorithme MCMC.
 	On a
 
-	A chaque étape de l'algorithme MCMC, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S}$ de $p$ candidats (le package \textsc{CorReg} permet à l'utilisateur de choisir parmi plusieurs types de voisinage).
-
-	On fait l'approximation suivante : 
-	\begin{equation}
-		P(S|X)\approx \exp(BIC^*(S))
-	\end{equation}
-	On définit alors
+	A chaque étape de l'algorithme, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S}$ %de $p$ candidats (le package \textsc{CorReg} permet à l'utilisateur de choisir parmi plusieurs types de voisinage).
+		et donc la fonction de transition est guidée par $BIC^*$ de la façon suivante :	
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \frac{1}{p} \sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(\frac{-1}{2} BIC(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(\frac{-1}{2} BIC(S_l))} \\
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(\frac{-1}{2} BIC^*(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(\frac{-1}{2} BIC^*(S_l))}
 	\end{eqnarray}
 
-		La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire.
-		Le résultat obtenu est la meilleure structure rencontrée en termes de $BIC^*$ (vraisemblance pénalisée).
+		La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire dont le mode correspond à la structure de plus grande valeur de $BIC^*$
 
 L'intialisation peut se faire en utilisant la matrice des corrélations et/ou la méthode du Graphical Lasso [2].		
-La grande dimension de l'espace parcouru rend préférable  (pour un temps de calcul égal) l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (permet aussi la parallélisation).
+La grande dimension de l'espace parcouru rend préférable  (pour un temps de calcul égal) l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (permettant aussi la parallélisation).
 
-	En pratique, on commence par estimer pour chaque variable de $X$ sa densité sous l'hypothèse d'un mélange gaussien.
+	En pratique, on commence par estimer pour chaque variable de $X$ sa densité sous l'hypothèse d'un mélange gaussien (avec Mixmod [6]).
 On peut alors calculer la loi jointe de $X$ pour chaque structure réalisable rencontrée durant l'algorithme MCMC.
-\section{Résultats}	
-Pour les simulations présentées dans les tableaux \ref{tableMSEsimdroite}, \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
-Les tableaux affichent le nombre de variable dépendantes trouvées, le nombre de variables jugées dépendantes à tort et les erreures moyennes en prédiction (MSE) sur $Y$ à partir d'échantillons de validation de 1 000 individus.
+Sans cette hypothèse générative supplémentaire sur $X_1$, l'utilisation de $BIC^*$ serait compromise.
+Notons cependant la souplesse de cette hypothèse due à la grande flexibilité des mélanges gaussiens.
+\section{Résultats sur données simulées}	
+L'ensemble de la méthode à été programmé dans un package R dénommé \textsc{CorReg}.Pour les simulations présentées dans les tableaux \ref{tableMSEsimdroite}, \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
+Les tableaux affichent le nombre de variable dépendantes trouvées, le nombre de variables jugées dépendantes à tort et les erreurs moyennes en prédiction (MSE) sur $Y$ à partir d'échantillons de validation de 1 000 individus.
 Pour l'ensemble des simulations $p=40$, $\sigma_Y=10$, $\sigma=0.001$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
-Les $B_{i,j}$ suivent la même loi de poisson mais avec un signe aléatoire.
-On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1 dans le vrai modèle.
-\textsc{CorReg}  a  travaillé avec $K$,$p_2$ et $p_1$ libres.
+Les $B_{1}^j$ suivent la même loi de poisson mais avec un signe aléatoire.
+On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (fortes corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1 dans le vrai modèle.
+\textsc{CorReg}  a  travaillé avec, $p_2$ et $p_1$ libres et a utilisé Mixmod pour estimer les densités dans $X_1$.
 
@@
 On retrouve le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
-On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle, comme pour le LASSO.
+On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle de régression en $Y$, comme pour le LASSO.
 
 Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
-Exemples de Sous-régressions obtenues par \textsc{CorReg} ayant interprétation physique :
+Exemples de Sous-régressions obtenues par \textsc{CorReg} ayant une interprétation physique :
 \begin{itemize}
@@
    \end{minipage}
-   \caption{résultats obtenus sur données réelles : $n=117$ et $p=168$. l'erreur est réduite d'un tiers alors que la complexité du modèle est divisée par $2,5$.   }
+   \caption{Résultats obtenus sur données réelles : $n=117$ et $p=168$. L'erreur est réduite d'un tiers alors que la complexité du modèle est divisée par $2.5$.   }
 \end{figure}
@@
 
-\noindent [1] B. Efron, T. Hastie, I. Johnstone, and R. Tibshirani (2004). Least angle regression. {\it The
+\noindent [1] Efron, B., Hastie, T., Johnstone,I. et Tibshirani, R. (2004), Least angle regression. {\it The
 Annals of statistics}, 32(2):407-499.
 
-\noindent [2] J. Friedman, T. Hastie, and R. Tibshirani (2008). Sparse inverse covariance estimation with
+\noindent [2] Friedman, J., Hastie, T. et Tibshirani, R. (2008), Sparse inverse covariance estimation with
 the graphical lasso.  {\it Biostatistics}, 9(3):432-441 .
 
-\noindent [3] E. Lebarbier and T. Mary-Huard (2006). Une introduction au critère bic: fondements
+\noindent [3] Lebarbier, E. et Mary-Huard,T. (2006), Une introduction au critère bic: fondements
 théoriques et interprétation.  {\it Journal de la SFdS }, 147(1):39-57.
 
-\noindent[5] R. Tibshirani (1996). Regression shrinkage and selection via the lasso.  {\it Journal of the Royal
+\noindent[4] Tibshirani, R. (1996). Regression shrinkage and selection via the lasso,  {\it Journal of the Royal
 Statistical Society}. Series B (Methodological), pages 267-288.
 
-\noindent [6] Peng Zhao and Bin Yu (December 2006). On model selection consistency of lasso. {\it J. Mach. Learn.
+\noindent [5] Zhao,P. et Yu,B. (2006), On model selection consistency of lasso, {\it J. Mach. Learn.
 Res.} 7:2541-2563.
+
+\noindent [6] Biernacki, C., Celeux, G., Govaert, G., \& Langrognet, F.
+(2006), Model-based cluster and discriminant analysis with the MIXMOD software, Computational Statistics \& Data Analysis, 51(2), 587-600.
 

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@332 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@333 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs, \ldots
+{\bf Mots-clés.} Régression, corrélations, industrie, sélection de variables, modèles génératifs
 \bigskip\bigskip
@@
 {\bf Abstract.} Linear regression generally suppose to have decorrelated covariates.
-This hypothesis is often irrealist with industrial datasets that contains many highly correlated covariates (due to the process, physcial laws,  {\it etc}).
+This hypothesis is often irrealist with industrial datasets that contains many highly correlated covariates due to the process, physcial laws,  {\it etc}.
 The proposed generative model consists in explicit modeling of the correlations with a family of linear regressions between the covariates permitting to obtain by marginalization a parsimonious correlation-free regression model, easily understandable and compatible with variable selection methods.
@@
 
-{\bf Keywords.} Regression, correlations, industry, variable selection, generative models,  \ldots
+{\bf Keywords.} Regression, correlations, industry, variable selection, generative models
 
@@
 \section{Introduction}
-	La régression linéaire classique suppose la décorrélation des covariables, source de problèmes en termes de variance des estimateurs.En effet, pour une variable réponse$Y$ et un ensemble de covariables $X \in \mathcal{R}^p$, la régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2)$ donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ dégénéré si les colonnes de $X$ sont linéairement corrélées.
+	La régression linéaire classique suppose la décorrélation des covariables, source de problèmes en termes de variance des estimateurs.
+En effet, pour une variable réponse $Y$ et un ensemble de covariables $X \in \mathcal{R}^{n \times p}$, la régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2I_n)$ et $A$ vecteur des $p$ coefficients donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ dégénéré si les colonnes de $X$ sont linéairement corrélées.
 Les méthodes de sélection comme le LASSO [4] muni du LAR [1] sont elles-mêmes touchées par ce problème de corrélation [5].
@@
 		Notre idée est de modéliser explicitement les corrélations présentes entre covariables sous la forme d'une famille de régression entre celles-ci.
-Nous présenterons donc le modèle génératif associé et l'algorithme MCMC permettant de choisir la famille de régression à utiliser avant d'illustrer l'efficacité de la méthode sur des données simulées puis sur des données réelles.
+Nous présenterons donc en première partie le modèle génératif associé puis l'algorithme MCMC permettant d'estimer la famille de régression à utiliser avant d'illustrer dans les parties 3 et 4 l'efficacité de la méthode sur des données simulées puis sur des données réelles avant de conclure.
 
-\section{Modèle supprimant les co-variables corrélées}		
+\section{Modèle supprimant les covariables corrélées}		
 	On suppose le modèle génératif suivant :
@@
 	\begin{itemize}
-		\item Régression principale :
+		\item Régression principale entre $Y$ et $X$:
 			 \begin{equation}
-			 	Y_{|X,S}=XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2) \label{MainR}
+			 	Y_{|X,S}=XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2I_n) \label{MainR}
 			 \end{equation}
-		\item Famille de $p_2$ régressions entre covariables corrélées :
+		\item Famille de $p_2$ régressions entre covariables de $X$ corrélées :
 			\begin{equation}
-			\forall j \in I_2 : \  X^j_{|X_1,S}=X_1B_{1}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2) \label{SR}
+			\forall j \in I_2 : \  X^j_{|X_1,S}=X_1B_{1}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2I_n) \label{SR}
 			\end{equation}
@@
 			\begin{equation}
-				\forall j \notin I_2 : \ X^j \sim \sum_{k=1}^{k_j}\pi_k\mathcal{N}(\mu_{k_j},\sigma_{k_j}^2)
+				\forall j \notin I_2 : \ X^j \sim \sum_{k=1}^{k_j}\pi_k\mathcal{N}(\mu_{k_j},\sigma_{k_j}^2I_n)
 			\end{equation}
@@
 
-Où %$B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$,\\ 
+où %$B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$,\\ 
 $I_1=\{I_1^1,\dots,I_1^{p}\}$ est le vecteur des indices des variables à droite dans (\ref{SR}) , $I_2=\{j |\sharp I_1^j>0 \}$ est l'ensemble des indices des variables corrélées à gauche dans (\ref{SR}).
@@
 		%&=&BIC +\ln(P(S))+cste \\
-		BIC^*&=&BIC +\ln(P(S)) \label{Bicstar}
+		BIC^*&=&BIC +\ln(P(S)) \label{Bicstar} .
 	\end{eqnarray}	
@@
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(\frac{-1}{2} BIC^*(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(\frac{-1}{2} BIC^*(S_l))}
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(\frac{-1}{2} BIC^*(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(\frac{-1}{2} BIC^*(S_l))} .
 	\end{eqnarray}
@@
 \end{tabular} 
-\caption{$Y$ ne dépend pas de $X_2$.\textsc{CorReg} gagne même en se trompant un peu sur $S$.} \label{tableMSEsimdroite}
+\caption{$Y$ ne dépend pas de $X_2$.} \label{tableMSEsimdroite}
 \end{table}
@@
 On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle de régression en $Y$, comme pour le LASSO.
-
+\section{Résultats sur données réelles}
 Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@334 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 $I_1=\{I_1^1,\dots,I_1^{p}\}$ est le vecteur des indices des variables à droite dans (\ref{SR}) , $I_2=\{j |\sharp I_1^j>0 \}$ est l'ensemble des indices des variables corrélées à gauche dans (\ref{SR}).
-On a donc une partition des données $X=(X_1,X_2)$ où $X_2=X^{I_2}$ et $X_1=X\setminus X_2$(pour alléger les notations).\\
+Les $B_1^j$ sont les coefficients des régressions entre covariables.
+On a donc une partition des données $X=(X_1,X_2)$ où $X_2=X^{I_2}$ et $X_1=X\setminus X_2$.\\
 On suppose $I_1\cap I_2=\emptyset$, $i.e.$ les variables dépendantes dans $X$ n'en expliquent pas d'autres.
-On note $p_2= \sharp I_2$ le nombre de régressions entre covariables et $p_1=(p_1^1,\dots,p_1^{p_2})$.
+On note $p_2= \sharp I_2$ le nombre de régressions entre covariables et $p_1=(p_1^1,\dots,p_1^{p_2})$ qui est le vecteur des longueurs des régressions au sein de $X$.
 
 On a ainsi rendu explicites les corrélations au sein de $X$ sous la forme d'une structure de sous-régressions linéaires $S=(I_1,I_2,p_1,p_2)$.	
-Ce modèle génératif est identifiable sous certaines conditions simples (sur les $k_i$) non détaillées ici.
+Ce modèle génératif est identifiable sous certaines conditions simples (sur les $k_j$) non détaillées ici.
 
-On remarque que (\ref{MainR}) et (\ref{SR}) impliquent par simple intégration sur $X_2$, un modèle de régression en $Y$ s'exprimant uniquement en fonction des variables non corrélées $X_1$ :
+On remarque que alors (\ref{MainR}) et (\ref{SR}) impliquent par simple intégration sur $X_2$, un modèle de régression en $Y$ s'exprimant {\it uniquement en fonction des variables non corrélées $X_1$} :
 \begin{eqnarray}
-	Y_{|X,S}&=&X_1 (A_{1}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
-					= X_1\alpha_{1}+ \varepsilon_{\alpha} \label{Trueexpl} 			
+	Y_{|X_1,S}&=&X_1 (A_{1}+ \sum_{j \in I_2}B^{j}_{I_1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
+					= X_1\alpha_{1}+ \varepsilon_{\alpha}. \label{Trueexpl} 			
 \end{eqnarray}		
-
 L'estimateur classique du Maximum de Vraisemblance de $\alpha$ est sans biais et s'exprime par	
 \begin{eqnarray}
-	\hat{\alpha}_{1}&=& (X'_{1} X_1)^{-1}X'_{1}Y \textrm{ et sa matrice de variance} \\
+	\hat{\alpha}_{1}&=& (X'_{1} X_1)^{-1}X'_{1}Y  \\
+\textrm{et sa matrice de variance } && \nonumber \\
 	\Var[\hat{\alpha}_{1}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1}
@@
 		peut être notablement mieux conditionnée que celle de $\hat A$ initial.						
-
 	En outre, ce nouveau modèle réduit consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
@@
 
-		\textbf{\underline{Remarque}} : En ajoutant une étape de sélection de variables on obtient ainsi deux types de $0$ : ceux issus de l'étape de décorrélation et ceux issus de cette éventuelle étape de sélection de type LASSO.
+		\textbf{\underline{Remarque}} : En ajoutant une étape de sélection de variables on obtient ainsi deux  ``types de $0$"  : ceux issus de l'étape de décorrélation et ceux issus de cette éventuelle étape de sélection de type LASSO.
 
@@
 La recherche du meilleur $S$ selon $BIC^*$ n'est pas un problème simple et on va s'appuyer sur un algorithme MCMC.
-	On a
 
 	A chaque étape de l'algorithme, pour $S \in \mathcal{S}$ (ensemble des structures réalisables) on définit un voisinage $\mathcal{V}_{S}$ %de $p$ candidats (le package \textsc{CorReg} permet à l'utilisateur de choisir parmi plusieurs types de voisinage).
-		et donc la fonction de transition est guidée par $BIC^*$ de la façon suivante :	
+		et ensuite la fonction de transition est guidée par $BIC^*$ de la façon suivante :	
 	\begin{eqnarray}
@@
 	\end{eqnarray}
-
-		La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire dont le mode correspond à la structure de plus grande valeur de $BIC^*$
+La chaîne de Markov ainsi constituée est ergodique dans un espace d'états finis et possède une unique loi stationnaire dont le mode correspond à la structure de plus grande valeur de $BIC^*$.
 
@@
 
-	En pratique, on commence par estimer pour chaque variable de $X$ sa densité sous l'hypothèse d'un mélange gaussien (avec Mixmod [6]).
+	En pratique, on commence par estimer pour chaque variable de $X$ sa densité sous l'hypothèse d'un mélange gaussien (avec le package Rmixmod de Mixmod [6]).
 On peut alors calculer la loi jointe de $X$ pour chaque structure réalisable rencontrée durant l'algorithme MCMC.
 Sans cette hypothèse générative supplémentaire sur $X_1$, l'utilisation de $BIC^*$ serait compromise.
-Notons cependant la souplesse de cette hypothèse due à la grande flexibilité des mélanges gaussiens.
+Notons cependant la souplesse de cette hypothèse due à la grande flexibilité des mélanges gaussiens [7].
 \section{Résultats sur données simulées}	
-L'ensemble de la méthode à été programmé dans un package R dénommé \textsc{CorReg}.Pour les simulations présentées dans les tableaux \ref{tableMSEsimdroite}, \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
-Les tableaux affichent le nombre de variable dépendantes trouvées, le nombre de variables jugées dépendantes à tort et les erreurs moyennes en prédiction (MSE) sur $Y$ à partir d'échantillons de validation de 1 000 individus.
-Pour l'ensemble des simulations $p=40$, $\sigma_Y=10$, $\sigma=0.001$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
-Les $B_{1}^j$ suivent la même loi de poisson mais avec un signe aléatoire.
-On cherche ici à se comparer à la méthode $LASSO$ dans les cas où celle-ci est en difficulté (fortes corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1 dans le vrai modèle.
-\textsc{CorReg}  a  travaillé avec, $p_2$ et $p_1$ libres et a utilisé Mixmod pour estimer les densités dans $X_1$.
+L'ensemble de la méthode à été programmé dans un package R dénommé \textsc{CorReg}.
+Pour les simulations présentées dans les tableaux \ref{tableMSEsimdroite}, \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
+Les tableaux affichent le nombre de variables dépendantes trouvées (``bon gauche"), le nombre de variables jugées dépendantes à tort (``faux gauche") et les erreurs moyennes en prédiction (MSE) sur $Y$ à partir d'échantillons de validation de 1 000 individus.
+Pour l'ensemble des simulations $p=40$, $\sigma_Y=10$, $\sigma=0.001$, les $X$ indépendants suivent des mélanges gaussiens à $\lambda=5$ classes de moyenne selon une loi de Poisson de paramètre $\lambda$ et d'écart-type $\lambda$.
+Les $B_{1}^j$ suivent la même loi de Poisson mais avec un signe aléatoire.
+On cherche ici à se comparer à la méthode LASSO dans les cas où celle-ci est en difficulté (fortes corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1 dans le vrai modèle.
+\textsc{CorReg}  a  travaillé avec $p_2$ et $p_1$ libres et a utilisé Mixmod pour estimer les densités dans $X_1$.
 
@@
 On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle de régression en $Y$, comme pour le LASSO.
+
+On remarque que quand $p_2$ augmente le LASSO commence à se ressaisir car il y a de plus en plus de faux modèles proches du vrai en termes de prédiction.
+Le LASSO trouve donc par moment des modèles inconsistants en interprétation mais relativement corrects en prédiction.
+ATTENTION : dans le cadre industriel, l'interprétation du modèle mène à des actions sur le process et donc un modèle inconsistant en interprétation peut mener à des actions contre-productives (d'où l'importance de l'interprétabilité de \textsc{CorReg}).
 \section{Résultats sur données réelles}
@@
 
+\noindent [7] McLachlan, G. J., et Basford, K. E. (1988), Mixture models : Inference and applications to clustering. {\it Statistics: Textbooks and Monographs, New York: Dekker }, 1988, 1.
 %

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@335 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Le LASSO trouve donc par moment des modèles inconsistants en interprétation mais relativement corrects en prédiction.
-ATTENTION : dans le cadre industriel, l'interprétation du modèle mène à des actions sur le process et donc un modèle inconsistant en interprétation peut mener à des actions contre-productives (d'où l'importance de l'interprétabilité de \textsc{CorReg}).
+%ATTENTION : dans le cadre industriel, l'interprétation du modèle mène à des actions sur le process et donc un modèle inconsistant en interprétation peut mener à des actions contre-productives (d'où l'importance de l'interprétabilité de \textsc{CorReg}).
 \section{Résultats sur données réelles}
@@
 
-Exemple de régression sur une variable réponse dans le cadre de données réelles :
+%Exemple de régression sur une variable réponse dans le cadre de données réelles :
 \begin{figure}[!h]
@@
 \noindent [7] McLachlan, G. J., et Basford, K. E. (1988), Mixture models : Inference and applications to clustering. {\it Statistics: Textbooks and Monographs, New York: Dekker }, 1988, 1.
+
 %

### 2014-02-03 pkg/CorReg/src/OLS.cpp
SOLARIS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@336 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -78,8 +78,7 @@ SEXP OLS(SEXP X,SEXP Y,SEXP intercept,SEXP sigma,SEXP Bic,SEXP methode){
         
         float log_likelihood = (-vecY.rows()/2)*log(2*PI*sumc*sumc)-nbsum/(2*sumc*sumc);
         
-        BicTheta=(-2*log_likelihood+(k*float(log(vecY.rows()))));
-        
+        BicTheta=(-2*log_likelihood+(k*log(float(vecY.rows()))));
       }
 
   }

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@338 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \section{Résultats sur données simulées}	
-L'ensemble de la méthode à été programmé dans un package R dénommé \textsc{CorReg}.
-Pour les simulations présentées dans les tableaux \ref{tableMSEsimdroite}, \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
+L'ensemble de la méthode à été programmé dans un package R dénommé \textsc{CorReg}. Pour les simulations présentées dans les tableaux %\ref{tableMSEsimdroite},
+ \ref{tableMSEsimtout} et \ref{tableMSEsimgauche}, chacune des configurations à été simulée $100$ fois.
 Les tableaux affichent le nombre de variables dépendantes trouvées (``bon gauche"), le nombre de variables jugées dépendantes à tort (``faux gauche") et les erreurs moyennes en prédiction (MSE) sur $Y$ à partir d'échantillons de validation de 1 000 individus.
@@
 \textsc{CorReg}  a  travaillé avec $p_2$ et $p_1$ libres et a utilisé Mixmod pour estimer les densités dans $X_1$.
+%
+%\begin{table}
+%\centering
+%\begin{tabular}{|c|c|c|c|c|c|c|c|}
+%\hline
+% \multicolumn{2}{c}{}  & \multicolumn{2}{|c |}{Qualité de $S$} & \multicolumn{3}{c|}{Qualité de prédiction (MSE)} \\
+%\hline 
+%$n$ & $p_2$ & bon gauche & faux gauche    & LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} vrai $S$\\ 
+%\hline 
+%30 & 16 &  8 & 5.39 & 2 466 225.35 & 13 796.03 & 588.9\\ 
+%\hline 
+%30 & 32 & 17.05 & 2.7 & 979.16 & 196.33 & 141.2\\ 
+%\hline 
+%\hline 
+%50 & 0 & 0 & 0 & 499.18 & 499.18 & 499.18 \\ 
+%\hline 
+%50 & 16 & 9.18 & 4.94 & 315.34 & 202.64 & 193.38 \\ 
+%\hline 
+%50 & 32 & 19.13 & 2.24 & 179.89 & 138.96 & 120.21 \\ 
+%\hline \hline
+%400 & 32 & 23.66 & 1.13 & 105.38 & 103.88 & 102.81\\ 
+%\hline 
+%\end{tabular} 
+%\caption{$Y$ ne dépend pas de $X_2$.} \label{tableMSEsimdroite}
+%\end{table}
 
@@
 \begin{tabular}{|c|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_2$ & bon gauche & faux gauche    & LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} vrai $S$\\ 
-\hline 
-30 & 16 &  8 & 5.39 & 2 466 225.35 & 13 796.03 & 588.9\\ 
-\hline 
-30 & 32 & 17.05 & 2.7 & 979.16 & 196.33 & 141.2\\ 
-\hline 
-\hline 
-50 & 0 & 0 & 0 & 499.18 & 499.18 & 499.18 \\ 
-\hline 
-50 & 16 & 9.18 & 4.94 & 315.34 & 202.64 & 193.38 \\ 
-\hline 
-50 & 32 & 19.13 & 2.24 & 179.89 & 138.96 & 120.21 \\ 
-\hline \hline
-400 & 32 & 23.66 & 1.13 & 105.38 & 103.88 & 102.81\\ 
-\hline 
-\end{tabular} 
-\caption{$Y$ ne dépend pas de $X_2$.} \label{tableMSEsimdroite}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\cline{3-7}
+ \multicolumn{2}{c|}{}  & \multicolumn{2}{c|}{Qualité de $S$} & \multicolumn{3}{c|}{Qualité de prédiction (MSE)} \\
 \hline 
@@
 \end{tabular} 
-\caption{$Y$  dépend  de $X$ entier.} \label{tableMSEsimtout}
+\caption{$Y$  dépend  de $X$ entier. \textsc{CorReg} gagne logiquement.} \label{tableMSEsimtout}
 \end{table}
@@
 \begin{table}
-\centering
 \begin{tabular}{|c|c|c|c|c|c|c|c|}
+\cline{3-7}
+ \multicolumn{2}{c|}{}  & \multicolumn{2}{c|}{Qualité de $S$} & \multicolumn{3}{c|}{Qualité de prédiction (MSE)} \\
 \hline 
@@
 \end{tabular} 
-\caption{$Y$  dépend  de $X_2$ uniquement (cas normalement défavorable à \textsc{CorReg} ).} \label{tableMSEsimgauche}
+\caption{$Y$  dépend  de $X_2$ uniquement (cas normalement défavorable à \textsc{CorReg}). Les corrélations suffisent à faire gagner \textsc{CorReg}} \label{tableMSEsimgauche}
 \end{table}
 
-Les résultats (tableaux 1 à 3) montrent que \textsc{CorReg} est équivalent au LASSO en l'absence de corrélations et le bat quand les corrélations sont fortes.
+Les résultats (tableaux 1 et 2) montrent que \textsc{CorReg} est équivalent au LASSO en l'absence de corrélations et le bat quand les corrélations sont fortes.
 On retrouve le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
@@
 
-\noindent [7] McLachlan, G. J., et Basford, K. E. (1988), Mixture models : Inference and applications to clustering. {\it Statistics: Textbooks and Monographs, New York: Dekker }, 1988, 1.
+\noindent [7] McLachlan, G. J., et Basford, K. E. (1988), Mixture models : Inference and applications to clustering. {\it Statistics: Textbooks and Monographs, New York: Dekker }, 1.
 

### 2014-02-03 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@339 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %\begin{tabular}{|c|c|c|c|c|c|c|c|}
+%\cline{3-7}
+% \multicolumn{2}{c|}{}  & \multicolumn{2}{c|}{Qualité de $\hat S$} & \multicolumn{3}{c|}{Qualité de prédiction (MSE)} \\
 %\hline 
@@
 \cline{3-7}
- \multicolumn{2}{c|}{}  & \multicolumn{2}{c|}{Qualité de $S$} & \multicolumn{3}{c|}{Qualité de prédiction (MSE)} \\
+ \multicolumn{2}{c|}{}  & \multicolumn{2}{c|}{Qualité de $\hat S$} & \multicolumn{3}{c|}{Qualité de prédiction (MSE)} \\
 \hline 
@@
 \end{tabular} 
-\caption{$Y$  dépend  de $X_2$ uniquement (cas normalement défavorable à \textsc{CorReg}). Les corrélations suffisent à faire gagner \textsc{CorReg}} \label{tableMSEsimgauche}
+\caption{$Y$  dépend  de $X_2$ uniquement (cas normalement défavorable à \textsc{CorReg}). } \label{tableMSEsimgauche}
 \end{table}
@@
 On retrouve le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
-On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle de régression en $Y$, comme pour le LASSO.
+On constate enfin la convergence asymptotique de \textsc{CorReg} vers le vrai modèle de régression.% en $Y$, comme pour le LASSO.
 
@@
 Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
-Exemples de Sous-régressions obtenues par \textsc{CorReg} ayant une interprétation physique :
+Certaines régressions obtenues décrivent des règles process, d'autres sont plus évidentes :
 \begin{itemize}
@@
 			\end{itemize}
-
-			D'autres sous-régressions traduisent des modèles physiques qui régulent le process industriel sur la base d'expertises métallurgiques.
-\\
+		La figure 1 illustre (Corrélations et $R^2$ des régressions en $X$) les résultats obtenus lors d'une étude qualité chez ArcelorMittal.
+Le MSE obtenu par \textsc{CorReg} est $\mathbf{6.47\%}$ meilleur que celui du LASSO.
+Pour un échantillon d'apprentissage de 3 000 individus et un échantillon de validation (postérieur) de 847 individus.
+L'étude a été menée dans le cadre de la qualité produit.
+L'enjeu de ces quelques pourcents de gain se chiffre en dizaine de milliers d'euros annuels sans compter l'impact sur les parts de marché (non chiffrable mais bien plus considérable).
 
 %Exemple de régression sur une variable réponse dans le cadre de données réelles :
-\begin{figure}[!h]
-	\begin{minipage}[c]{.40\linewidth}
-			\includegraphics[width=200px]{figures/histcor_auchan.JPG} 
+\begin{figure}[h!]
+	\begin{minipage}[c]{.5\linewidth}
+			\includegraphics[width=200px]{figures/correlexfoshist.png} 
 	\end{minipage} \hfill
-   \begin{minipage}[c]{.52\linewidth}
-		\begin{tabular}{|c|c|c|}
-		\hline 
-		  & MSE  & Variables retenues  \\ 
-		\hline
-		LASSO (lars) & 0.80 & 54 \\ 
-		\hline 
-		CorReg (et lars) & 0.53 & 24  \\ 
-		\hline 
-		\end{tabular} 
+   \begin{minipage}[c]{.5\linewidth}
+			\includegraphics[width=200px]{figures/histR2exfos.png} 
    \end{minipage}
-   \caption{Résultats obtenus sur données réelles : $n=117$ et $p=168$. L'erreur est réduite d'un tiers alors que la complexité du modèle est divisée par $2.5$.   }
+  \caption{Etude qualité ArcelorMittal avec $205$ variables process. $\hat{p_2}=82$.    }
 \end{figure}

### 2014-02-04 www/Rennes.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@341 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	La régression linéaire classique suppose la décorrélation des covariables, source de problèmes en termes de variance des estimateurs.
-En effet, pour une variable réponse $Y$ et un ensemble de covariables $X \in \mathcal{R}^{n \times p}$, la régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2I_n)$ et $A$ vecteur des $p$ coefficients donne un estimateur de variance $ \Var(\hat{A}|X)= \sigma^2(X'X)^{-1}$ dégénéré si les colonnes de $X$ sont linéairement corrélées.
+En effet, pour une variable réponse $Y \in \mathcal{R}^{n}$ et un ensemble de covariables $X \in \mathcal{R}^{n \times p}$, la régression $Y=XA+\varepsilon$ avec $ \varepsilon\sim \mathcal{N}(0,\sigma^2I_n)$ (où $I_n$ est la matrice identité de taille $n$) et $A\in \mathcal{R}^{p}$ vecteur des $p$ coefficients donne un estimateur $\hat A$ de variance $ \Var(\hat{A}|X)= \sigma^2_Y(X'X)^{-1}$ dégénéré si les colonnes de $X$ sont linéairement corrélées.
 Les méthodes de sélection comme le LASSO [4] muni du LAR [1] sont elles-mêmes touchées par ce problème de corrélation [5].
@@
 
-		Notre idée est de modéliser explicitement les corrélations présentes entre covariables sous la forme d'une famille de régression entre celles-ci.
-Nous présenterons donc en première partie le modèle génératif associé puis l'algorithme MCMC permettant d'estimer la famille de régression à utiliser avant d'illustrer dans les parties 3 et 4 l'efficacité de la méthode sur des données simulées puis sur des données réelles avant de conclure.
+		Notre idée est de modéliser explicitement les corrélations présentes entre covariables sous la forme d'une famille de régressions entre celles-ci.
+Nous présenterons donc le modèle génératif associé puis en partie \ref{secMCMC} l'algorithme MCMC permettant d'estimer la famille de régressions à utiliser avant d'illustrer dans les parties \ref{secressim} et \ref{secresrel} l'efficacité de la méthode sur des données simulées puis sur des données réelles avant de conclure en partie \ref{secconcl}.
 
@@
 	On suppose le modèle génératif suivant :
-
 	\begin{itemize}
@@
 			 \begin{equation}
-			 	Y_{|X,S}=XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2I_n) \label{MainR}
+			 	Y_{|X,S}=XA+\varepsilon_Y= X_1A_{1}+X_2A_{2}+\varepsilon_Y \textrm{ avec } \varepsilon_Y \sim \mathcal{N}(0,\sigma_Y^2I_n); \label{MainR}
 			 \end{equation}
@@
 			\begin{equation}
-			\forall j \in I_2 : \  X^j_{|X_1,S}=X_1B_{1}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2I_n) \label{SR}
+			\forall j \in I_2 : \  X^j_{|X_1,S}=X_1B_{1}^j + \varepsilon_{j} \textrm{ avec } \varepsilon_j \sim \mathcal{N}(0,\sigma_j^2I_n); \label{SR}
 			\end{equation}
@@
 			\begin{equation}
-				\forall j \notin I_2 : \ X^j \sim \sum_{k=1}^{k_j}\pi_k\mathcal{N}(\mu_{k_j},\sigma_{k_j}^2I_n)
+				\forall j \notin I_2 : \ X^j \sim \sum_{k=1}^{k_j}\pi_k\mathcal{N}(\mu_{k_j},\sigma_{k_j}^2I_n);
 			\end{equation}
 	\end{itemize}
-
 	où %$B_{I_1^j}^j$ est le vecteur de taille $p_1^j$ des coefficients de la sous-régression en $X^j$,\\ 
 $I_1=\{I_1^1,\dots,I_1^{p}\}$ est le vecteur des indices des variables à droite dans (\ref{SR}) ,  $I_2=\{j |\sharp I_1^j>0 \}$ est l'ensemble des indices des variables corrélées à gauche dans (\ref{SR}).
-Les $B_1^j$ sont les coefficients des régressions entre covariables.
-On a donc une partition des données $X=(X_1,X_2)$ où $X_2=X^{I_2}$ et $X_1=X\setminus X_2$.\\
-On suppose $I_1\cap I_2=\emptyset$, $i.e.$ les variables dépendantes dans $X$ n'en expliquent pas d'autres.
+Les $B_1^j \in \mathcal{R}^{(p-p_2)}$ sont les coefficients des régressions entre covariables.
+On a donc une partition des données $X=(X_1,X_2)$ où $X_2=X^{I_2}$ et $X_1=X\setminus X_2$.
+On suppose en outre $I_1\cap I_2=\emptyset$, $i.e.$ les variables dépendantes dans $X$ n'en expliquent pas d'autres.
 On note $p_2= \sharp I_2$ le nombre de régressions entre covariables et $p_1=(p_1^1,\dots,p_1^{p_2})$ qui est le vecteur des longueurs des régressions au sein de $X$.
@@
 
-On remarque que alors (\ref{MainR}) et (\ref{SR}) impliquent par simple intégration sur $X_2$, un modèle de régression en $Y$ s'exprimant {\it uniquement en fonction des variables non corrélées $X_1$} :
+On remarque alors que (\ref{MainR}) et (\ref{SR}) impliquent par simple intégration sur $X_2$, un modèle de régression en $Y$ s'exprimant {\it uniquement en fonction des variables non corrélées $X_1$} :
 \begin{eqnarray}
@@
 L'estimateur classique du Maximum de Vraisemblance de $\alpha$ est sans biais et s'exprime par	
-\begin{eqnarray}
-	\hat{\alpha}_{1}&=& (X'_{1} X_1)^{-1}X'_{1}Y  \\
-\textrm{et sa matrice de variance } && \nonumber \\
-	\Var[\hat{\alpha}_{1}|X,S]&=& (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1}
-\end{eqnarray}
-		peut être notablement mieux conditionnée que celle de $\hat A$ initial.						
+\begin{equation}
+	\hat{\alpha}_{1}= (X'_{1} X_1)^{-1}X'_{1}Y  
+\end{equation}
+En particulier sa matrice de variance 
+\begin{equation}
+	\Var[\hat{\alpha}_{1}|X,S]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X_{1})^{-1}
+\end{equation}
+		peut être notablement mieux conditionnée que celle de $\hat A$ initial (dimension réduite et surtout variables orthogonales).						
 	En outre, ce nouveau modèle réduit consiste en une régression linéaire classique qui peut donc bénéficier des outils de sélection de variables au même titre que le modèle complet.			
-		 \\		
 		 Notons enfin que la structure explicite entre les variables permet de mieux comprendre les phénomènes en jeu et la parcimonie du modèle facilite son interprétation. 
 \\ 
+		\textbf{\underline{Remarque}} : En ajoutant une étape de sélection de variables (de type LASSO) on obtient ainsi deux  ``types de $0$"  : ceux issus de l'étape de décorrélation et ceux issus de la sélection.
 
-		\textbf{\underline{Remarque}} : En ajoutant une étape de sélection de variables on obtient ainsi deux  ``types de $0$"  : ceux issus de l'étape de décorrélation et ceux issus de cette éventuelle étape de sélection de type LASSO.
-
-\section{Estimation de la structure de corrélation}
+\section{Estimation de la structure de corrélation}\label{secMCMC}
 
@@
 	L'équiprobabilité ainsi supposée des $p_2$ et $p_1^j$ vient pénaliser davantage la compléxité sous l'hypothèse $p_2<\frac{p}{2}$ , hypothèse réaliste sur le nombre de régressions entre covariables.
-La recherche du meilleur $S$ selon $BIC^*$ n'est pas un problème simple et on va s'appuyer sur un algorithme MCMC.
+La recherche du meilleur $S$ selon $BIC^*$ n'est pas un problème simple et on va s'appuyer sur un algorithme MCMC pour le résoudre.
 
@@
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=& \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(\frac{-1}{2} BIC^*(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(\frac{-1}{2} BIC^*(S_l))} .
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : P(S,\tilde{S})&=& \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S}\} }\frac{\exp(-\frac{1}{2} BIC^*(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S}}\exp(-\frac{1}{2} BIC^*(S_l))}.
 	\end{eqnarray}
@@
 Notons cependant la souplesse de cette hypothèse due à la grande flexibilité des mélanges gaussiens [7].
-\section{Résultats sur données simulées}	
+\section{Résultats sur données simulées}\label{secressim}	
 L'ensemble de la méthode à été programmé dans un package R dénommé \textsc{CorReg}. Pour les simulations présentées dans les tableaux %\ref{tableMSEsimdroite},
@@
 %ATTENTION : dans le cadre industriel, l'interprétation du modèle mène à des actions sur le process et donc un modèle inconsistant en interprétation peut mener à des actions contre-productives (d'où l'importance de l'interprétabilité de \textsc{CorReg}).
-\section{Résultats sur données réelles}
-Les données industrielles sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), Moyenne et maximum d'une courbe ($\rho=0.911$).
-Certaines régressions obtenues décrivent des règles process, d'autres sont plus évidentes :
-\begin{itemize}
-	\item Moyenne = f (Min , Max , Sigma ) pour des données courbes
-	\item Largeur du produit = f (débit de fonte , vitesse de la coulée continue)	\\
-Vrai modèle physique (non linéaire) :
-
-	 Largeur = $\frac{\textrm{débit}}{\textrm{vitesse } \times \textrm{ épaisseur}}$ (Mais dans ce cas précis l'épaisseur est constante)
-			\end{itemize}
-		La figure 1 illustre (Corrélations et $R^2$ des régressions en $X$) les résultats obtenus lors d'une étude qualité chez ArcelorMittal.
-Le MSE obtenu par \textsc{CorReg} est $\mathbf{6.47\%}$ meilleur que celui du LASSO.
-Pour un échantillon d'apprentissage de 3 000 individus et un échantillon de validation (postérieur) de 847 individus.
+\section{Résultats sur données réelles}\label{secresrel}
+La figure 1 illustre les résultats obtenus lors d'une étude qualité chez ArcelorMittal.
+Les données sidérurgiques sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), moyenne et maximum d'une courbe ($\rho=0.911$).
+La base analysée ici comportait $205$ variables et 3 847 individus (apprentissage du modèle sur les 3 000 premiers et validation sur les 847 autres).
+%Certaines régressions obtenues décrivent des règles process, d'autres sont plus évidentes :
+%\begin{itemize}
+%	\item Moyenne = f (Min , Max , Sigma ) pour des données courbes
+%	\item Largeur du produit = f (débit de fonte , vitesse de la coulée continue)	\\
+%Vrai modèle physique (non linéaire) :
+%
+%	 Largeur = $\frac{\textrm{débit}}{\textrm{vitesse } \times \textrm{ épaisseur}}$ (Mais dans ce cas précis l'épaisseur est constante)
+%			\end{itemize}
+\textsc{CorReg} retrouve ($p_2=82$) en plus des corrélations ci-dessus des modèles de régulation du process et certains modèles physiques naturels.
+\\
+		Le MSE (sur l'échantillon de validation) obtenu par \textsc{CorReg} est $\mathbf{6.47\%}$ meilleur que celui du LASSO.
+\textsc{CorReg} donne un modèle à 34 variables contre 21 pour le LASSO avec 14 variables communes.
 L'étude a été menée dans le cadre de la qualité produit.
@@
    \end{minipage}
-  \caption{Etude qualité ArcelorMittal avec $205$ variables process. $\hat{p_2}=82$.    }
+  \caption{Histogrammes des corrélations et des $R^2_{adj}$ des régressions obtenues par \textsc{CorReg} }
 \end{figure}
 
-\section{Conclusion et perspectives}
+\section{Conclusion et perspectives}\label{secconcl}
 	\textsc{CorReg} est fonctionnel et disponible sur R-forge. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.

### 2014-02-04 pkg/CorReg/src/BicLoc_cpp.cpp
SOLARIS update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@343 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -51,7 +51,7 @@ double BicLoc_cpp(Eigen::MatrixXd X,Eigen::MatrixXd Y,bool intercept,int methode
       Eigen::MatrixXd residus = temp*inverse;
       Eigen::MatrixXd sig = (Y-residus);
       
-      //calcul de l'ecart type (sigma)
+      //calcul de ecart type (sigma)
       for (int i = 0; i < sig.rows(); i++){
         sumc=sumc+(sig(i,0)-sig.mean())*(sig(i,0)-sig.mean());
       }
@@ -60,7 +60,7 @@ double BicLoc_cpp(Eigen::MatrixXd X,Eigen::MatrixXd Y,bool intercept,int methode
       sumc=sqrt(sumc);
       if(sumc==0){//modele exact
         BicTheta=0;
-       Rcout<<"exact model found, please delete one variable \n";
+       Rcpp::Rcout<<"exact model found, please delete one variable \n";
       }else{
       //calcul du BIC
         int k=inverse.rows()+1;

### 2014-02-04 pkg/CorReg/src/CVMSE.cpp

@@ -17,8 +17,8 @@ using Eigen::VectorXd;
 SEXP CVMSE(SEXP RX,SEXP RY,SEXP RK,SEXP Rintercept,SEXP Rmethode,SEXP Rgroupe)
 {
   BEGIN_RCPP
-  //déclaration des paramètres
-  const Map<MatrixXd> X(as<Map<MatrixXd> >(RX));//on utilise le format EIgen pour la matrice pour bénéficier des outils matriciels de base
+  //déclaration des parametres
+  const Map<MatrixXd> X(as<Map<MatrixXd> >(RX));//on utilise le format EIgen pour la matrice pour beneficier des outils matriciels de base
   const Map<MatrixXd> Y(as<Map<MatrixXd> >(RY));
   const Map<VectorXi> groupe(as<Map<VectorXi> >(Rgroupe));
   
@@ -26,12 +26,12 @@ SEXP CVMSE(SEXP RX,SEXP RY,SEXP RK,SEXP Rintercept,SEXP Rmethode,SEXP Rgroupe)
   int methode = Rcpp::as<int>(Rmethode);     // length vector 
   bool intercept = Rcpp::as<bool>(Rintercept);     // length vector 
 
-  //initialisations hors paramètres
+  //initialisations hors parametres
   int n = X.rows();
   int p = X.cols(); 
   VectorXi compteur(K);//vecteur des groupes
   compteur <<VectorXi::Zero(K);
- // srand (time(NULL)); //initialisation de la graine aléatoire
+ // srand (time(NULL)); //initialisation de la graine aleatoire
   
   for(int i=0;i<n;i++){//recensement des groupes
     compteur[groupe[i]]++;
@@ -40,14 +40,14 @@ SEXP CVMSE(SEXP RX,SEXP RY,SEXP RK,SEXP Rintercept,SEXP Rmethode,SEXP Rgroupe)
   if(intercept){
     inter++;
   }
-  MatrixXd beta(p+inter,1);// déclaration du beta qui ne change jamais de taille
+  MatrixXd beta(p+inter,1);// declaration du beta qui ne change jamais de taille
   beta << MatrixXd::Zero(p+inter,1);
   double MSE=0;
   double MSEloc=0;
   int compt_test=0;
   int compt_appr=0;
   for(int j=0;j<K;j++){//pour chaque groupe
-    //déclarations locales
+    //declarations locales
[…]

### 2014-02-04 pkg/CorReg/src/OLS.cpp

@@ -82,7 +82,7 @@ SEXP OLS(SEXP X,SEXP Y,SEXP intercept,SEXP sigma,SEXP Bic,SEXP methode){
       }
 
   }
-  //retour des objets demandés
+  //retour des objets demandes
   if (isigma[0]==1){
     if (iBic[0]==1){
           return List::create(

### 2014-02-04 pkg/CorReg/src/cleanZ.cpp

@@ -28,15 +28,15 @@ SEXP cleanZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_BIC,SEXP plot,SEXP bla
   vec_iterator imet_BIC=met_BIC.begin(),iplot=Plot.begin(),ibla=Bla.begin();
   int p=matZ.cols();//nombre de colonne de la matrice Z
   Eigen::MatrixXd Zopt;//meilleur Z obtenu
-  double Bicbest;//BIC associé au meilleur modèle
+  double Bicbest;//BIC associe au meilleur modèle
   Eigen::VectorXd bicvect;//vecteur BIC des matrices Z retenues
-  Eigen::MatrixXd newZ;//matrice Z modifié à chaque étapes
+  Eigen::MatrixXd newZ;//matrice Z modifie a chaque etapes
   Eigen::MatrixXd list_cand;//matrice qui contient les coordonnées des candidats (liste candidats)
   int compte;//permet de cree la matrice liste en désignant le numéro du candidat (liste candidats)
   int nbcand;//nombre de candidats
   int numcand;//numero du candidat
-  int i_loc;//premiere coordonnée du candidat (modification Z)
-  int j_loc;//duexième coordonnée du candidat (modification Z)
+  int i_loc;//premiere coordonnee du candidat (modification Z)
+  int j_loc;//duexieme coordonnee du candidat (modification Z)
   Eigen::MatrixXd Zcand;//matrice Z du candidat (modification Z)
   Eigen::VectorXd BIC_cand;//vecteur qui contient les BIC de chaque colonne des la matrice Zcand (calcul du BIC)
   double Sum_BIC_cand;//somme des BIC de BIC_cand (calcul BIC)
@@ -49,7 +49,7 @@ SEXP cleanZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_BIC,SEXP plot,SEXP bla
   int iret;
   nbcand=matZ.sum();
   Eigen::VectorXd bic_etape(nbcand);//vecteur qui stock le BIC de chaque etapes
-  Eigen::VectorXd complexite_etape(nbcand);//vecteur qui stock la compléxité de Z a chaque etapes
+  Eigen::VectorXd complexite_etape(nbcand);//vecteur qui stock la complexite de Z a chaque etapes
[…]

### 2014-02-04 pkg/CorReg/src/cleancolZ.cpp

@@ -16,7 +16,7 @@ using Rcpp::as;
 SEXP cleancolZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_BIC,SEXP plot,SEXP bla,SEXP Rstar)
 {
   BEGIN_RCPP
-  //déclaration des variables
+  //declaration des variables
   const Map<MatrixXd> matZ(as<Map<MatrixXd> >(Z));//Z
   const Map<MatrixXd> matX(as<Map<MatrixXd> >(X));//X
   const Map<VectorXd> Bic_vide_vect(as<Map<VectorXd> >(bic_vide_vect));//bic_vide_vect
@@ -27,14 +27,14 @@ SEXP cleancolZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_BIC,SEXP plot,SEXP 
   vec_iterator imet_BIC=met_BIC.begin(),iplot=Plot.begin(),ibla=Bla.begin();
   int p=matZ.cols();//nombre de colonne de la matrice Z
   Eigen::MatrixXd Zopt;//meilleur Z obtenu
-  double Bicbest;//BIC associé au meilleur modèle
+  double Bicbest;//BIC associe au meilleur modele
   Eigen::VectorXd bicvect;//vecteur BIC des matrices Z retenues
-  Eigen::MatrixXd newZ;//matrice Z modifié à chaque étapes
+  Eigen::MatrixXd newZ;//matrice Z modifie a chaque etapes
   Eigen::VectorXd list_cand;//vecteur qui contient les indices des colonnes candidates (liste candidats)
-  int compte=0;//permet de cree la matrice liste en désignant le numéro du candidat (liste candidats)
+  int compte=0;//permet de cree la matrice liste en designant le numero du candidat (liste candidats)
   int nbcand;//nombre de candidats
   int numcand;//numero du candidat
-  int i_loc;//premiere coordonnée du candidat (modification Z)
+  int i_loc;//premiere coordonnee du candidat (modification Z)
   Eigen::MatrixXd Zcand;//matrice Z du candidat (modification Z)
   Eigen::ArrayXXd SumCol(1,p);//SumCol est un vecteur qui contient la somme de chaque colonne de Zcand (modification Z)
   Eigen::VectorXd BIC_cand;//vecteur qui contient les BIC de chaque colonne des la matrice Zcand (calcul du BIC)
[…]

### 2014-02-04 pkg/CorReg/src/kronecker_prod_cpp.cpp

@@ -6,7 +6,7 @@ using namespace std;
 using Eigen::Map;
 using Eigen::MatrixXd;
 using Rcpp::as;
-//V1 et V2 sont des vecteurs colonnes issus de matrices (sous-matrices donc)
+//V1 et V2 sont des vecteurs colonnes issus de matrices (sous matrices donc)
 RcppExport Rcpp::NumericVector kronecker_prod_cpp(const Rcpp::NumericVector &V1,const Rcpp::NumericVector &V2){
     int size1=V1.size();
     int size2=V2.size();

### 2014-02-04 pkg/CorReg/src/kronecker_sum_cpp.cpp

@@ -4,7 +4,7 @@ using namespace Rcpp ;
 using namespace Eigen;
 using namespace std;
 
-//V1 and V2 are column vectors from matrices (sub-matrices)
+//V1 and V2 are column vectors from matrices (sub matrices)
 RcppExport Rcpp::NumericVector kronecker_sum_cpp(const Rcpp::NumericVector &V1,const Rcpp::NumericVector &V2){
     int size1=V1.size();
     int size2=V2.size();

### 2014-02-04 pkg/CorReg/src/rechercheZ_sparse_relax.cpp

@@ -96,13 +96,16 @@ SEXP rechercheZ_sparse_relax(SEXP X,SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP bic_vid
     Zopt_si = Z_si;
     Zopt_sj = Z_sj;
   }
-  Rcout<<"Bicbest :"<<Bicbest<<"\n";
+  if (ibla[0]>0)
+  {
+   Rcout<<"Bicbest :"<<Bicbest<<"\n";
+  }
   newZ_zi=Z_zi;
   newZ_zj=Z_zj;
   newZ_si=Z_si;
   newZ_sj=Z_sj;
   int step =0;
-  Rcout<<"ici";
+  //Rcout<<"ici";
   if (imet_tirage[0]==0)//methode de changement de la ligne et de la colonne
   {
     nbcand=2*p-2;
@@ -136,23 +139,23 @@ SEXP rechercheZ_sparse_relax(SEXP X,SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP bic_vid
   nb_opt=0;
   stock_bool.resize(nbcand);
   stock_BIC.resize(nbcand);
-  for(v=0;v<newZ_sj.size();v++)
-  {
-    Rcout<<newZ_sj(v)<<" ";
-  }
-  Rcout<<"\n\n";
-  for(v=0;v<newZ_zj.size();v++)
-  {
-    Rcout<<newZ_zj(v)<<" ";
-  }
-
-  Rcout<<"\n";
-  for(v=0;v<newZ_zi.size();v++)
-  {
-    Rcout<<newZ_zi(v)<<" ";
-  }
-          
-  Rcout<<"\n\n";
+//  for(v=0;v<newZ_sj.size();v++)
+//  {
+//    Rcout<<newZ_sj(v)<<" ";
+//  }
+//  Rcout<<"\n\n";
+//  for(v=0;v<newZ_zj.size();v++)
+//  {
+//    Rcout<<newZ_zj(v)<<" ";
+//  }
+//
+//  Rcout<<"\n";
+//  for(v=0;v<newZ_zi.size();v++)
+//  {
+//    Rcout<<newZ_zi(v)<<" ";
+//  }
+//          
+//  Rcout<<"\n\n";
   while(step<imaxiter[0])
   // initialisation des variables
   
[…]

### 2014-02-04 pkg/CorReg/R/rechercheZ_sparse_relax.R
SOLARIS update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@343 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,26 +1,43 @@
-# ' recherche de structure au format creux et en mode relax
-# '@param X
-# '@param Zi numeros des lignes de chaque 1
-# '@param Zj numeros des colonnes de chaque 1
-# '@param Sj nombre de 1 dans chaque lignes
-# '@param Sj nombre de 1 dans chaque colonnes
-# '@param bic_vide_vect vecteur BIC de la matrice nulle
-# '@param methode_tirage 0:ligne et colonne,-1:seulement la colonne, entier>0:nombre aleatoire de candidats
-# '@param methode_BIC 1:utilisation de la fonction householderQr, 2:utilisation de la fonction colPivHouseholderQr
-# '@param Rmax complexite maximum d'une sous-regression
-# '@param Maxiter nombre d'etapes
-# '@param plot T:retourne le type de changement, la complexite et le BIC de chaque etapes
-# '@param best T:permet d'aller systematiquement au meilleur BIC si il est meilleur que tout les autres deja rencontres 
-# '@param better T:permet d'aller systematiquement au meilleur BIC si il est meilleur que l'etape precedente
-# '@param random F:permet de s'ameliorer ou de rester sur place
-# '@param bla 0:pas de messages, 1:affiche le BIC,le numero d'etape et la complexite de Z quand il y'a un meilleur BIC, 2:affiche le BIC,le numero d'etape,la complexite de Z,le nombre de candidats et le BIC minimum observe parmi les candidats quand il y'a un meilleur BIC, 3: affiche en plus de bla=1 la complexite locale et le BIC local
-# '@param nb_opt_max
-# '@param Mixmod
-# '@return etape 0:suppression,1 ajout,2 stationarite
+#' Sparse structure research
+#' @param X the dataset
+#' @param Zi indices of the rows of the 1
+#' @param Zj indices of the columns of the 1
+#' @param Sj nombre de 1 dans chaque lignes
+#' @param Sj nombre de 1 dans chaque colonnes
+#' @param bic_vide_vect vecteur BIC de la matrice nulle
+#' @param methode_tirage 0:ligne et colonne,-1:seulement la colonne, entier>0:nombre aleatoire de candidats
[…]

### 2014-02-04 pkg/CorReg/src/BICZsparse.cpp

@@ -39,7 +39,7 @@ SEXP BICZsparse(SEXP X,SEXP Z_Zi,SEXP Z_Sj,SEXP Bic_vide_vect,SEXP BicOld,SEXP m
     {
       if(Z_sj(k)!=0)
       {
-        //on créé X_loc et y_loc
+        //on fait X_loc et y_loc
         X_loc.resize(inrow[0],(Z_sj(k)));
         
         debut_colj=0;

### 2014-02-04 pkg/CorReg/src/BicTheta.cpp

@@ -22,7 +22,7 @@ BEGIN_RCPP
   Eigen::MatrixXd temp(size,sizecol+1);
   
   
-  //ajout de la constance si demandé
+  //ajout de la constance si demande
   if (iinter[0]==1){
     Eigen::ArrayXd vec1(size); 
      

### 2014-02-04 pkg/CorReg/src/BicZ.cpp

@@ -54,7 +54,7 @@ SEXP BicZ(SEXP X,SEXP Z,SEXP Bic_vide_vect,SEXP BicOld,SEXP methode,SEXP Zold)
         int newsize=colZ.sum();
         mat_var_a_droite.resize(sizerow,newsize);
         int k=0;
-        //on recherche l'emplacement des 1 dans le vecteur colonne (variables a droite)
+        //on recherche l emplacement des 1 dans le vecteur colonne (variables a droite)
         for(int j=0;j<=size-1;j++)
         {
           if (colZ(j,0)==1)
@@ -65,11 +65,11 @@ SEXP BicZ(SEXP X,SEXP Z,SEXP Bic_vide_vect,SEXP BicOld,SEXP methode,SEXP Zold)
           }//fin if
         }//fin for j
   
-        //calcul de son BIC en appellant OLS (les parametres sont : Y=X[,colonne],X=X[,matrice des variables à droite])
+        //calcul de son BIC en appellant OLS (les parametres sont : Y=X[,colonne],X=X[,matrice des variables a droite])
         BicCol=BicLoc_cpp(mat_var_a_droite,matX.block(0,i,sizerow,1),"T",method);
         if(BicCol==0){//a priori on a un modele exact (peu probable que le BIC soit vraiment nul)
         Rcout <<i+1<< " depends on";
-        //on recherche l'emplacement des 1 dans le vecteur colonne (variables a droite)
+        //on recherche l emplacement des 1 dans le vecteur colonne (variables a droite)
         for(int j=0;j<=size-1;j++)
         {
           if (colZ(j,0)==1)

### 2014-02-04 pkg/CorReg/src/BicZ_cpp.cpp

@@ -50,13 +50,13 @@ Eigen::MatrixXd val;
     //calcul de son BIC en appellant OLS (les parametres sont : Y=X[,colonne],X=X[,matrice des variables à droite])
    BicCol=BicLoc_cpp(mat_var_a_droite,X.block(0,i,sizerow,1),"T",methode);
    if(BicCol==0){//a priori on a un modele exact (peu probable que le BIC soit vraiment nul)
-        Rcout<<i+1<< " depends on";
+        Rcpp::Rcout<<i+1<< " depends on";
         //on recherche l'emplacement des 1 dans le vecteur colonne (variables a droite)
         for(int j=0;j<=size-1;j++)
         {
           if (colZ(j,0)==1)
           {
-           Rcout<< " "<< j +1<< " ";
+           Rcpp::Rcout<< " "<< j +1<< " ";
           }//fin if
         }//fin for j
   }

### 2014-02-04 pkg/CorReg/src/BicZmissRcpp.cpp

@@ -66,10 +66,10 @@ SEXP BicZmissRcpp(SEXP X,SEXP Z,SEXP M,SEXP Bic_vide_vect,SEXP BicOld,SEXP metho
           }//fin if
         }//fin for j
   
-        //calcul de son BIC en appellant OLS (les parametres sont : Y=X[,colonne],X=X[,matrice des variables Ã  droite])
+        //calcul de son BIC en appellant OLS (les parametres sont : Y=X[,colonne],X=X[,matrice des variables a  droite])
         BicCol=BicLoc_cpp(mat_var_a_droite,matX.block(0,i,sizerow,1),"T",method);
         if(BicCol==0){//a priori on a un modele exact (peu probable que le BIC soit vraiment nul)
-        Rcout <<i+1<< " depends on";
+        Rcpp::Rcout <<i+1<< " depends on";
         //on recherche l'emplacement des 1 dans le vecteur colonne (variables a droite)
         for(int j=0;j<=size-1;j++)
         {

### 2014-02-04 pkg/CorReg/src/GM_likelihood.cpp

@@ -59,23 +59,23 @@ extern "C" SEXP GM_likelihood(SEXP RY,SEXP RX,SEXP RBeta,SEXP RSigma,SEXP RM,SEX
         j++;
       }
     }
-    //on sait maintenant précisément qui manque et où trouver les paramètres associés
+    //on sait maintenant précisément qui manque et ou trouver les paramètres associés
     if(nbclass==1){// la loi est normale quand-même
       sigma=sigma*sigma;//on passe au carré car on va ajouter les variances
       for(int i=0;i<nbmank;i++){//pour chaque manquant
-        meanfix=meanfix+detailsmixmod(nbclustcumsum(quimank(i))-1,1)*beta(quimank(i)+inter);//colonne 1 est la deuxième colonne donc la moyenne
+        meanfix=meanfix+detailsmixmod(nbclustcumsum(quimank(i))-1,1)*beta(quimank(i)+inter);//colonne 1 est la deuxieme colonne donc la moyenne
         sigma=sigma+detailsmixmod(nbclustcumsum(quimank(i))-1,2)*beta(quimank(i)+inter)*beta(quimank(i)+inter);
       }
-      res=Rcpp::wrap(dnorm(Y,meanfix,sqrt(sigma)));//on a des variances donc on passe à la racine
+      res=Rcpp::wrap(dnorm(Y,meanfix,sqrt(sigma)));//on a des variances donc on passe a la racine
     }else{//plusieurs classes
       NumericVector proptot(1);
       NumericVector vartot(1);
-      NumericVector meantot(1);//en deux temps pour qu'il crée bien des vecteurs
+      NumericVector meantot(1);//en deux temps pour qu'il cree bien des vecteurs
       meantot(0)=meanfix;
       proptot(0)=1;
       vartot(0)=sigma*sigma;      
       for(int i=0;i<nbmank;i++){//pour chaque manquant
-        //on crée les vecteurs des paramètres
+        //on cree les vecteurs des paramètres
         NumericVector proploc(nbclustmixmod(quimank(i)));
         NumericVector meanloc(nbclustmixmod(quimank(i)));
         NumericVector varloc(nbclustmixmod(quimank(i)));
@@ -91,7 +91,7 @@ extern "C" SEXP GM_likelihood(SEXP RY,SEXP RX,SEXP RBeta,SEXP RSigma,SEXP RM,SEX
       double resloc;
       for(int i=0;i<nbclass;i++){
[…]

### 2014-02-04 pkg/CorReg/src/modsparse_cpp.cpp

@@ -69,7 +69,7 @@ SEXP modsparse_cpp(SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP complexite,SEXP p,SEXP i
         {
           if(ajout==false && (zj(k)>iJ[0] ||(zj(k)==iJ[0] && zi(k)>iI[0]) ))//on ajoute
           {            
-            zj_loc<<zj_loc,iJ[0];//on passe par zj_loc
+            zj_loc<<zj_loc,iJ[0];//on passe par zjloc
             zi_loc<<zi_loc,iI[0];
             ajout=true;
           }
@@ -96,7 +96,7 @@ SEXP modsparse_cpp(SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP complexite,SEXP p,SEXP i
           {
             if(ajout==false && (zj(k)>iJ[0] ||(zj(k)==iJ[0] && zi(k)>iI[0]) ))//on ajoute
             {            
-              zj_loc<<zj_loc,iJ[0];//on passe par zj_loc
+              zj_loc<<zj_loc,iJ[0];//on passe par zjloc
               zi_loc<<zi_loc,iI[0];
               ajout=true;
             }

### 2014-02-04 pkg/CorReg/src/rcppeigen_hello_world.h
cleaning
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@344 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,8 +1,0 @@
-#ifndef _CorReg_RCPP_HELLO_WORLD_H
-#define _CorReg_RCPP_HELLO_WORLD_H
-
-#include <RcppEigen.h>
-
-RcppExport SEXP rcppeigen_hello_world() ;
-
-#endif

### 2014-02-04 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@342 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update biblio
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@345 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 L'intialisation peut se faire en utilisant la matrice des corrélations et/ou la méthode du Graphical Lasso [2].		
-La grande dimension de l'espace parcouru rend préférable  (pour un temps de calcul égal) l'utilisation de multiples chaînes courtes plutôt qu'une seule très longue (permettant aussi la parallélisation).
+La grande dimension de l'espace parcouru rend préférable [8]  (pour un temps de calcul égal) l'utilisation de multiples chaînes courtes  plutôt qu'une seule très longue (permettant aussi la parallélisation).
 
@@
 En effet, le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire.
+Enfin, la famille de régressions pourrait être utilisée dans domaines autres que celui de la régression linéaire.
+La connaissance de la structure interne des données possède en effet une valeur intrinsèque que d'autres méthodes pourraient tenter de mettre à profit.
 
@@
 
+\noindent [8] Gilks, W. R., Richardson, S., \& Spiegelhalter, D. J. (Eds.). (1996). Markov chain Monte Carlo in practice (Vol. 2). CRC press.
 %

### 2014-02-05 www/Rennes.tex
update correlations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@346 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \textsc{CorReg} retrouve ($p_2=82$) en plus des corrélations ci-dessus des modèles de régulation du process et certains modèles physiques naturels.
+Les valeurs absolues des corrélations dans $X_1$ sont $5,45\%$ plus faibles que dans $X$, ce qui est significatif vue la dimension de la matrice des corrélations et sa moyenne très faible malgré de fortes corrélations.
+Le nombre de corrélations supérieures à $0.8$ en valeur absolue est réduit de $\mathbf{77,9\%}$ (respectivement $86$ et $19$ paires).
 \\
 		Le MSE (sur l'échantillon de validation) obtenu par \textsc{CorReg} est $\mathbf{6.47\%}$ meilleur que celui du LASSO.
-\textsc{CorReg} donne un modèle à 34 variables contre 21 pour le LASSO avec 14 variables communes.
+\textsc{CorReg} donne un modèle à 38 variables contre 20 pour le LASSO avec 13 variables communes.
+Le LASSO propose 7 variables différentes du modèle  \textsc{CorReg} mais elles sont toutes dans $X_2$ et expliquées par des variables du modèle  \textsc{CorReg} qui s'interdit l'usage de $X_2$.
 		 L'étude a été menée dans le cadre de la qualité produit.
@@
 
-%Exemple de régression sur une variable réponse dans le cadre de données réelles :
+%%Exemple de régression sur une variable réponse dans le cadre de données réelles :
 \begin{figure}[h!]
 	\begin{minipage}[c]{.5\linewidth}
-			\includegraphics[width=200px]{figures/correlexfoshist.png} 
+			\includegraphics[width=300px]{figures/boxplotXX1.png} 
 	\end{minipage} \hfill
-   \begin{minipage}[c]{.5\linewidth}
-			\includegraphics[width=200px]{figures/histR2exfos.png} 
+   \begin{minipage}[c]{.35\linewidth}
+			\includegraphics[width=165px]{figures/histR2exfos.png} 
    \end{minipage}
-  \caption{Histogrammes des corrélations et des $R^2_{adj}$ des régressions obtenues par \textsc{CorReg} }
+  \caption{Corrélations dans $X$ et $X_1$ et $R^2_{adj}$ des régressions obtenues par \textsc{CorReg} }
 \end{figure}
@@
 
-\noindent [6] Biernacki, C., Celeux, G., Govaert, G., \& Langrognet, F.
+\noindent [6] Biernacki, C., Celeux, G., Govaert, G., et Langrognet, F.
 (2006), Model-based cluster and discriminant analysis with the MIXMOD software, Computational Statistics \& Data Analysis, 51(2), 587-600.
 
-\noindent [7] McLachlan, G. J., et Basford, K. E. (1988), Mixture models : Inference and applications to clustering. {\it Statistics: Textbooks and Monographs, New York: Dekker }, 1.
+\noindent [7] McLachlan, G., et Peel, D. (2004). Finite mixture models. Wiley. com.
 
-\noindent [8] Gilks, W. R., Richardson, S., \& Spiegelhalter, D. J. (Eds.). (1996). Markov chain Monte Carlo in practice (Vol. 2). CRC press.
+\noindent [8] Gilks, W. R., Richardson, S., et Spiegelhalter, D. J. (Eds.). (1996). Markov chain Monte Carlo in practice (Vol. 2). CRC press.
 %

### 2014-02-05 www/Rennes.tex
update correlations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@347 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@348 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Les $B_{1}^j$ suivent la même loi de Poisson mais avec un signe aléatoire.
-On cherche ici à se comparer à la méthode LASSO dans les cas où celle-ci est en difficulté (fortes corrélations 2 à 2) donc les $p_1^j$ non nuls valent tous 1 dans le vrai modèle.
-\textsc{CorReg}  a  travaillé avec $p_2$ et $p_1$ libres et a utilisé Mixmod pour estimer les densités dans $X_1$.
+On cherche ici à se comparer à la méthode LASSO dans les cas où celle-ci est en difficulté le vrai modèle est constitué de corrélations 2 à 2.
+\textsc{CorReg}  a  travaillé avec $p_2$ et $p_1$ libres et a utilisé Rmixmod pour estimer les densités dans $X_1$.
 %
@@
 
-Les résultats (tableaux 1 et 2) montrent que \textsc{CorReg} est équivalent au LASSO en l'absence de corrélations et le bat quand les corrélations sont fortes.
+Les tableaux 1 et 2 montrent que \textsc{CorReg} est équivalent au LASSO en l'absence de corrélations et le bat quand les corrélations sont fortes.
 On retrouve le phénomène attendu du LASSO moins impacté par les corrélations quand $n$ grandit.
@@
 
-On remarque que quand $p_2$ augmente le LASSO commence à se ressaisir car il y a de plus en plus de faux modèles proches du vrai en termes de prédiction.
-Le LASSO trouve donc par moment des modèles inconsistants en interprétation mais relativement corrects en prédiction.
+On remarque que quand $p_2$ augmente le LASSO commence à se ressaisir car il y a de plus en plus de faux modèles proches du vrai en termes de prédiction donc le LASSO trouve des modèles inconsistants en interprétation mais relativement corrects en prédiction.
 %ATTENTION : dans le cadre industriel, l'interprétation du modèle mène à des actions sur le process et donc un modèle inconsistant en interprétation peut mener à des actions contre-productives (d'où l'importance de l'interprétabilité de \textsc{CorReg}).
-\section{Résultats sur données réelles}\label{secresrel}
-La figure 1 illustre les résultats obtenus lors d'une étude qualité chez ArcelorMittal.
-Les données sidérurgiques sont fortement corrélées de manière naturelle : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), moyenne et maximum d'une courbe ($\rho=0.911$).
-La base analysée ici comportait $205$ variables et 3 847 individus (apprentissage du modèle sur les 3 000 premiers et validation sur les 847 autres).
+\section{Résultats d'une Etude qualité chez ArcelorMittal}\label{secresrel}
+%%Exemple de régression sur une variable réponse dans le cadre de données réelles :
+\begin{figure}[h!]
+	\begin{minipage}[l]{.30\linewidth}
+			\includegraphics[height=150px,width=150px]{figures/correlXX1.png} 
+			\caption{Valeurs de $\rho$ pour $X$ (haut) et $X_1$ (bas).}
+	\end{minipage} \hfill
+	\begin{minipage}[c]{.30\linewidth}
+			\includegraphics[height=150px, width=150px]{figures/histR2exfos.png} 
+			\caption{$R^2_{adj}$ des 82 régressions obtenues.}
+	\end{minipage} \hfill
+   \begin{minipage}[r]{.30\linewidth}
+			\includegraphics[height=150px,width=150px]{figures/p1exfos.png} 
+			\caption{Longueur des régressions obtenues ($p_1$).}
+   \end{minipage}
+\end{figure}   	
+On note $\rho$ la valeur absolue des corrélations.
+Les données sidérurgiques étudiées ($p=205$ et $n=3000$) sont fortement corrélées de manière naturelle (Figure 1 : en haut) : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), moyenne et maximum d'une courbe ($\rho=0.911$).
 %Certaines régressions obtenues décrivent des règles process, d'autres sont plus évidentes :
@@
 %			\end{itemize}
-\textsc{CorReg} retrouve ($p_2=82$) en plus des corrélations ci-dessus des modèles de régulation du process et certains modèles physiques naturels.
-Les valeurs absolues des corrélations dans $X_1$ sont $5,45\%$ plus faibles que dans $X$, ce qui est significatif vue la dimension de la matrice des corrélations et sa moyenne très faible malgré de fortes corrélations.
-Le nombre de corrélations supérieures à $0.8$ en valeur absolue est réduit de $\mathbf{77,9\%}$ (respectivement $86$ et $19$ paires).
+\textsc{CorReg} trouve  en plus des corrélations ci-dessus des modèles de régulation du process et des modèles physiques naturels pour un total de $p_2=82$ régressions (Figure 2) de longueur moyenne $\bar p_1=5$ (Figure 3).
+%Les valeurs absolues des corrélations dans $X_1$ sont $5,45\%$ plus faibles que dans $X$, ce qui est significatif vue la dimension de la matrice des corrélations et sa moyenne très faible malgré de fortes corrélations.
+Entre $X$ et $X_1$ le nombre de $\rho > 0.7$ est réduit de $\mathbf{79,33\%}$  avec respectivement $150$ et $31$ paires de variables (Figure 1 en bas).
 \\
-		Le MSE (sur l'échantillon de validation) obtenu par \textsc{CorReg} est $\mathbf{6.47\%}$ meilleur que celui du LASSO.
-\textsc{CorReg} donne un modèle à 38 variables contre 20 pour le LASSO avec 13 variables communes.
-Le LASSO propose 7 variables différentes du modèle  \textsc{CorReg} mais elles sont toutes dans $X_2$ et expliquées par des variables du modèle  \textsc{CorReg} qui s'interdit l'usage de $X_2$.
+		Ici $Y$ est un indicateur qualité (confidentiel).
+Le MSE (sur échantillon de validation de 847 nouveaux individus) obtenu par \textsc{CorReg} est $\mathbf{1.55\%}$ meilleur que celui du LASSO, avec respectivement 31 et 20 variables dont 13 communes.
+LASSO propose 7 variables différentes de  \textsc{CorReg} mais elles sont toutes dans $X_2$ et expliquées par des variables du modèle  \textsc{CorReg} avec un $R^2_{adj}$ moyen de $0.82$.De plus $\rho$ est $13.9\%$ plus faible pour les variables de \textsc{CorReg} que pour LASSO malgré davantage de variables.
+		\\
+		En termes d'interprétation, accompagner la régression en $Y$ avec la famille de régressions permet de mieux comprendre les conséquences d'éventuelles mesures correctives sur l'ensemble du process.
+Cela perment aussi de déterminer les {\it actionneurs } qui influent sur $Y$ quand le LASSO fait ressortir des variables {\it subies}.
+On peut donc plus facilement corriger le process pour atteindre l'objectif fixé.
 		 L'étude a été menée dans le cadre de la qualité produit.
@@
 
-%%Exemple de régression sur une variable réponse dans le cadre de données réelles :
-\begin{figure}[h!]
-	\begin{minipage}[c]{.5\linewidth}
-			\includegraphics[width=300px]{figures/boxplotXX1.png} 
-	\end{minipage} \hfill
-   \begin{minipage}[c]{.35\linewidth}
-			\includegraphics[width=165px]{figures/histR2exfos.png} 
-   \end{minipage}
-  \caption{Corrélations dans $X$ et $X_1$ et $R^2_{adj}$ des régressions obtenues par \textsc{CorReg} }
-\end{figure}
-
 \section{Conclusion et perspectives}\label{secconcl}
-	\textsc{CorReg} est fonctionnel et disponible sur R-forge. L'outil a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
-	La force de \textsc{CorReg} est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles de régression simples et donc facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
+	\textsc{CorReg} est disponible sur R-forge et a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
+	Sa force est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles de régression simples et donc facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
 	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, très présentes dans l'industrie.
-Cet aspect est envisagé sérieusement pour la prochaine version de \textsc{CorReg}.
-En effet, le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire.
-Enfin, la famille de régressions pourrait être utilisée dans domaines autres que celui de la régression linéaire.
-La connaissance de la structure interne des données possède en effet une valeur intrinsèque que d'autres méthodes pourraient tenter de mettre à profit.
+D'ailleurs le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire.
+Enfin, le principe de \textsc{CorReg} qui est l'explicitation des régressions latentes entre covariables pourrai être appliqué à d'autres méthodes prédictives (régression logistique, {\it etc}).
+%La connaissance de la structure interne des données possède en effet une valeur intrinsèque que d'autres méthodes pourraient tenter de mettre à profit.
 

### 2014-02-05 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@349 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 On note $\rho$ la valeur absolue des corrélations.
-Les données sidérurgiques étudiées ($p=205$ et $n=3000$) sont fortement corrélées de manière naturelle (Figure 1 : en haut) : largeur et poids d'une brame ($\rho=0.905$), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho=0.919$), moyenne et maximum d'une courbe ($\rho=0.911$).
+Les données sidérurgiques étudiées ($p$=205 et $n$=3000) sont fortement corrélées de manière naturelle (Figure 1 : en haut) : largeur et poids d'une brame ($\rho$=0.905), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho$=0.919), moyenne et maximum d'une courbe ($\rho$=0.911).
 %Certaines régressions obtenues décrivent des règles process, d'autres sont plus évidentes :
@@
 Le MSE (sur échantillon de validation de 847 nouveaux individus) obtenu par \textsc{CorReg} est $\mathbf{1.55\%}$ meilleur que celui du LASSO, avec respectivement 31 et 20 variables dont 13 communes.
-LASSO propose 7 variables différentes de  \textsc{CorReg} mais elles sont toutes dans $X_2$ et expliquées par des variables du modèle  \textsc{CorReg} avec un $R^2_{adj}$ moyen de $0.82$.De plus $\rho$ est $13.9\%$ plus faible pour les variables de \textsc{CorReg} que pour LASSO malgré davantage de variables.
+LASSO propose 7 variables différentes de  \textsc{CorReg} mais elles sont toutes dans $X_2$ et \textsc{CorReg} reprend les variables explicatives des régressions correspondantes ($R^2_{adj}$ moyen de $0.82$).
+De plus $\rho$ est $13.9\%$ plus faible pour les variables de \textsc{CorReg} que pour LASSO malgré davantage de variables.
 		\\

### 2014-02-05 www/Rennes.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@350 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 D'ailleurs le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire.
-Enfin, le principe de \textsc{CorReg} qui est l'explicitation des régressions latentes entre covariables pourrai être appliqué à d'autres méthodes prédictives (régression logistique, {\it etc}).
+Enfin, le principe de \textsc{CorReg} qui est l'explicitation des régressions latentes entre covariables pourrait être appliqué à d'autres méthodes prédictives (régression logistique, {\it etc}).
 %La connaissance de la structure interne des données possède en effet une valeur intrinsèque que d'autres méthodes pourraient tenter de mettre à profit.

### 2014-02-06 pkg/CorReg/R/rechercheZ.R
doc update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@351 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,8 +1,8 @@
 #' recherche de structure
 #'@param X the dataset
-#'@param Z Z est une matrice nulle si on ne lui a pas mis de valeur
-#'@param Bic_null_vect vecteur BIC de la matrice nulle
-#'@param methode_tirage 0:ligne et colonne,-1:seulement la colonne, entier>0:nombre aleatoire de candidats, -2 : tout le monde (hors diagonale), -3 : uniquement les non-nuls
+#'@param Z binary adjacency matrix of size p. if NULL zero matrix is used
+#'@param Bic_null_vect vector of the BIC for each covariate
+#'@param methode_tirage 0:row and column,-1:column only, int>0:random int candidates, -2 : all (but the diag), -3 : non-zeros
 #'@param reject 0:mode relax, 1:mode reject
 #'@param methode_BIC 1:utilisation de la fonction householderQr, 2:utilisation de la fonction colPivHouseholderQr
 #'@param Rmax complexite maximum d'une sous-regression

### 2014-02-06 www/Rennes.tex
final update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@352 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-02-05 www/Rennes.tex
@@
 %ATTENTION : dans le cadre industriel, l'interprétation du modèle mène à des actions sur le process et donc un modèle inconsistant en interprétation peut mener à des actions contre-productives (d'où l'importance de l'interprétabilité de \textsc{CorReg}).
-\section{Résultats d'une Etude qualité chez ArcelorMittal}\label{secresrel}
+\section{Résultats d'une étude qualité chez ArcelorMittal}\label{secresrel}
 %%Exemple de régression sur une variable réponse dans le cadre de données réelles :
@@
 On note $\rho$ la valeur absolue des corrélations.
-Les données sidérurgiques étudiées ($p$=205 et $n$=3000) sont fortement corrélées de manière naturelle (Figure 1 : en haut) : largeur et poids d'une brame ($\rho$=0.905), température avant et après un outil ($\rho=0.983$), rugosité des deux faces du produit ($\rho$=0.919), moyenne et maximum d'une courbe ($\rho$=0.911).
+Les données sidérurgiques étudiées ($p$=205 et $n$=3000) sont fortement corrélées de manière naturelle (Figure 1 en haut), comme la largeur et poids d'une brame ($\rho$=0.905), la température avant et après un outil ($\rho=0.983$), la rugosité des deux faces du produit ($\rho$=0.919), une moyenne et un maximum ($\rho$=0.911).
 %Certaines régressions obtenues décrivent des règles process, d'autres sont plus évidentes :
@@
 \\
-		Ici $Y$ est un indicateur qualité (confidentiel).
+		Ici $Y$ est un indicateur qualité produit (confidentiel).
 Le MSE (sur échantillon de validation de 847 nouveaux individus) obtenu par \textsc{CorReg} est $\mathbf{1.55\%}$ meilleur que celui du LASSO, avec respectivement 31 et 20 variables dont 13 communes.
@@
 		En termes d'interprétation, accompagner la régression en $Y$ avec la famille de régressions permet de mieux comprendre les conséquences d'éventuelles mesures correctives sur l'ensemble du process.
-Cela perment aussi de déterminer les {\it actionneurs } qui influent sur $Y$ quand le LASSO fait ressortir des variables {\it subies}.
+Cela permet typiquement de déterminer les {\it actionneurs } qui influent sur $Y$ quand le LASSO fait ressortir des variables {\it subies}.
 On peut donc plus facilement corriger le process pour atteindre l'objectif fixé.
-		 L'étude a été menée dans le cadre de la qualité produit.
 		L'enjeu de ces quelques pourcents de gain se chiffre en dizaine de milliers d'euros annuels sans compter l'impact sur les parts de marché (non chiffrable mais bien plus considérable).
@@
 	\textsc{CorReg} est disponible sur R-forge et a d'ores et déjà montré son efficacité sur de vraies problématiques de régression en entreprise.
-	Sa force est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs modèles de régression simples et donc facilement accessibles aux non statisticiens (régressions linéaires) tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
-	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, très présentes dans l'industrie.
-D'ailleurs le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire.
-Enfin, le principe de \textsc{CorReg} qui est l'explicitation des régressions latentes entre covariables pourrai être appliqué à d'autres méthodes prédictives (régression logistique, {\it etc}).
+Sa force est la grande interprétabilité du modèle proposé, qui est constitué de plusieurs régression linéaires courtes et donc facilement accessibles aux non statisticiens tout en luttant efficacement contre les problématiques de corrélations, omniprésentes dans l'industrie.
+	On note néanmoins le besoin d'élargir le champ d'application à la gestion des valeurs manquantes, aussi très présentes dans l'industrie.
+D'ailleurs le modèle génératif actuel permettrait cette nouvelle fonctionnalité sans hypothèse supplémentaire, ce qui renforce encore son intérêt.
+Enfin, le principe de \textsc{CorReg} qui est l'explicitation des régressions latentes entre covariables pourrait être appliqué à d'autres méthodes prédictives (régression logistique, {\it etc}.).
 %La connaissance de la structure interne des données possède en effet une valeur intrinsèque que d'autres méthodes pourraient tenter de mettre à profit.

### 2014-03-14 pkg/CorReg/R/confint_coef.R
minor update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@446 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -22,16 +22,24 @@ confint_coef<-function(modele=NULL,n=NULL,prop=NULL,mean=NULL,alpha=0.05,labels=
       abline(h=0,col="red")
       print(paste("AIC : ",AIC(modele)))
    }else if(!is.null(prop)){
-      coef=c()
+      coef=list()
+      if(length(prop)>1 & !is.null(labels)){
+         un=round(prop*n)
+         zero=n-un
+         names(zero)=labels;names(un)=labels
+         X=rbind(cbind(1,rep(labels,times=un)),cbind(0,rep(labels,times=zero)))
+         coef=chisq.test(X[,1],X[,2])
+      }
+      coef$int=c()
       for (i in 1:length(prop)){
-         coef=rbind(coef,c(prop[i],icproportion(prop=prop[i],n=n[i],alpha=alpha)))
+         coef$int=rbind(coef$int,c(prop[i],icproportion(prop=prop[i],n=n[i],alpha=alpha)))
       }
       colnames(coef)=c("values","borne inf IC", "borne sup IC")
-      barplot(coef[,1],ylim=c(min(coef[,2]),max(coef[,3])),col="cyan",main="Intervalle de confiance des proportions",sub="",names.arg=labels)
-      x0=seq(from=1,length.out=nrow(coef),by=1.2)-0.3
-      y0=coef[,2]
+      barplot(coef$int[,1],ylim=c(min(coef$int[,2]),max(coef$int[,3])),col="cyan",main="Intervalle de confiance des proportions",sub="",names.arg=labels)
+      x0=seq(from=1,length.out=nrow(coef$int),by=1.2)-0.3
+      y0=coef$int[,2]
       x1=x0
-      y1=coef[,3]
+      y1=coef$int[,3]
       arrows(x0=x0,y0=y0,x1=x1,y1=y1,angle=90,code=3,lwd=2)
       if(!is.null(mean)){
          abline(h=mean,col="red")

### 2014-03-14 pkg/CorReg/R/recursive_tree.R

@@ -28,5 +28,5 @@ recursive_tree<-function(data=data,Y="Y",modele=NULL,kill=NULL,index=NULL,print=
    if(print){
       print(arbre)
    }
-   return(list(tree=arbre,modele=modele))
+   return(list(modele=modele,tree=arbre))
 }

### 2014-03-14 www/article13.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@447 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage[table]{xcolor}
 \graphicspath{{figures/}}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Clément THERY}
\title{CorReg pretreatment : Regression for correlated variables and application in steel industry}

%%%% debut macro %%%%
\newenvironment{changemargin}[2]{\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{0pt}%
\setlength{\rightmargin}{0pt}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setlength{\parsep}{0pt plus 1pt}%
\addtolength{\leftmargin}{#1}%
\addtolength{\rightmargin}{#2}%
}\item }{\end{list}}
%%%% fin macro %%%%

\definecolor{darkgreen}{rgb}{0,0.4,0}
	 \definecolor{darkred}{rgb}{0.75,0,0}
	 \definecolor{darkblue}{rgb}{0,0,0.4}

\begin{document}
\maketitle
\bigskip
{\bf Abstract.} Linear regression generally suppose to have decorrelated covariates. This hypothesis is often irrealist with industrial datasets that contains many highly correlated covariates due to the process, physcial laws,  {\it etc}. The proposed generative model consists in explicit modeling of the correlations with a family of linear regressions between the covariates permitting to obtain by marginalization a parsimonious correlation-free regression model, easily understandable and compatible with variable selection methods. The structure of correlations is found with an MCMC algorithm. An R package (\textsc{CorReg}) available on the CRAN implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry.
\smallskip

{\bf Keywords.} Regression, correlations, industry, variable selection, generative models

\section{Introduction}
%la régression et ses problèmes

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability and is simple to learn for non-statistician. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation that faces an also classical problem : the variance of the estimators.
When estimating the parameters of the regression we have to compute the inverse of a matrix\cite{saporta2006probabilites} which will be ill-conditioned or even singular if some covariates depend linearly from each other. For a model defined by 
	\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2)$ 
	we have the following Ordinary Least Squares (OLS) estimators :
	\begin{equation}
		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
	\end{equation}
	With variance 
	\begin{equation}
		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1}
	\end{equation}
	And when correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance explodes.
This variance increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model  : the more covariates you have the greater variance you get.
	\item The correlations within the covariates : strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

	With the rise of informatic, datasets contains more and more covariates and thus more and more useless covariates. So dimension reduction becomes a necessity. Moreover, when you use more covariates, you increase the chance to have correlated ones. For example, this work takes place in an industrial (steel industry) context with a big set of covariates (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc). In such a context, variance of the estimators can lead to arbitrary results or even no results at all. Prediction and interpretation are both strongly needed, with a preference for interpretation in industrial context (better to improve the process when possible than to only predict defects).
		~\\	~\\

%bibliographie	
	Because OLS is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
Moreover, real datasets implies many irrelevant variables (datasets based on the whole process without any a priori) so we have to use variable selection methods.

In the following we note classical norms: $\parallel\beta\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\beta \parallel_1=\sum_{i=1}^p|\beta_i| $.

	Ridge regression\cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:
	\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_2^2\leq k
	\end{equation}
	But Ridge regression is not efficient to select covariates (it's an assumed choice) because coefficients tends to 0 but don't reach 0.
	So it gives difficult interpretations for large values of $p$ and is not adapted for our industrial context. We need to reduce the dimension of the model. Our goal is not just to predict but also to understand the response variable.

	The Least Absolute Shrinkage and Selection Operator (LASSO)\cite{tibshirani1996regression} consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression\cite{efron2004least} (LAR) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates. And it  really selects covariates with coeffients set exactly to 0.
	 But LASSO also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. Another limitation of the LASSO is that it preserves at most $n$ predictors (troublesome when in high dimension). Some recent variants of the LASSO does exist for the choice of the penalization coefficient like the adaptative LASSO \cite{zou2006adaptive} or the random LASSO \cite{wang2011random}.
	 \\

	Elastic net\cite{zou2005regularization}	is a method developped to be a compromise between Ridge regression and the LASSO. 
%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
%	\begin{equation}
%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
%\end{equation}		
	Elastic net can be written:
	\begin{equation}
		\hat{\beta}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel Y-X\beta \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\beta\parallel_1+\alpha\parallel\beta\parallel_2^2\leq t \textrm{ for some } t
	\end{equation}
	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
	It seems to give good predictions. But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model. 
	\\

	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (OSCAR) \cite{bondell2008simultaneous}.
	The CLusterwise Effect REgression\cite{yengo2012variable} (CLERE) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith grouped $\beta_j$ following a Gaussian Mixture distribution.

The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\hat{\beta}$. But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net. Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute OLS but has no specific protection against correlations issues.
	~\\	~\\
% Principe de la méthode

 When some try to reduce the dimension and then just hope to have small correlations in the remaining dimensions,  we propose to focus on the correlations, giving a model with orthogonal covariates and an explicit structure between covariates. In fact we search the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint. This can be viewed as a pretreatment on the dataset allowing to use then other dimension reduction tools without suffering from correlations. We only consider strong correlations (i.e. : problematic ones) thus we keep most of the information contained in the dataset. 
 %We will in a second time be able to use the remaining part of the information (sequential approach).

	Our work is based on the assumption that if we know that correlations are a problem and if we precisely know the correlations, we could use this knowledge to avoid the problem.
	The idea is to suppose explicitly a linear structure between the covariates. 
	It gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (SEM)\cite{davidson1993estimation}. Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive SEM).
  	Other methods already rely on linear structure and start to take into account correlations but they only consider covariances between the residuals (SUR) \cite{SURzellner} or covariances between the endogenous variables like SPRING (Structured selection of Primordial Relationships IN the General linear model) \cite{chiquetconf}.

	  In this work, we decide to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we don't have a system of regressions but one regression on our response variable and a system of subregressions (without the response variable). And we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a reduced set of independent covariates. Thus we reduce dimension and correlations in the same time. The structure justifies the eviction of the redundant covariates without significant information loss. It can be seen as a pretreatment on the dataset based on the hypothesis of a strong structure between the covariates (i.e. : small $\varepsilon_X$).

	We can use any variable selection method on the reduced dataset with improved efficiency (reduced variance) due to dimension reduction and correlation suppression. So we obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.

But to work, we need an explicit structure between the covariates. SEM are often used in social sciences and economy where a structure is supposed "by hand" but here we want to be able to find a structure without any a priori (possibility to include some known structure remains). Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix). It consists in a selection in the precision matrix, setting some covariances to zero. But the resulting matrix is symmetric and we need an oriented graph for our SEM. So we developped an MCMC algorithm to find it (R package CorReg on CRAN). However, Graphical LASSO can be used in the initialization step of our MCMC. This structure is based on gaussian mixture models to fit better real datasets and to allow identifiability of the structure in terms of complexity (number of parameters).
 	\\

 	%plan
 	This paper will first present the reduced model and its properties before describing in Section 3 the algorithm used to find the structure.
 	We will then look at some numerical results on simulated (Section 4) and real industrial datasets (Section 5) before concluding and giving some perspectives in the sixth part.

\section{Model to decorrelate the covariates}

\subsection{Modelisation of the correlations}
Let $Y \in \mathcal{R}^n$ be a response variable we want to explain with a set $X \in \mathcal{R}^{n\times p}$ of $p$ correlated covariates.
We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$. Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $p_1=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression. Thus we have $p_2=| I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.

In the following, we note $X^j$ the $j^{th}$ column of a matrix $X$.
For lighter notation we define $X_2=X^{I_2}$ the matrix of the endogenous covariates and $X_1=X\setminus X_2$ the matrix of the remaining exogenous covariates.

We can now write the generative model:
\begin{itemize}
	\item Main regression between $Y$ and $X$:
		\begin{equation}
			Y_{|S}=XA+\varepsilon_Y =X_1A_1+X_2A_2+\varepsilon_Y \textrm{ whith } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
		\end{equation}
		where $A=(A_1,A_2) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $I_n$ the identity matrix, 
	\item Family of $p_2$ regressions within correlated covariates in $X$:
		\begin{equation}
			\forall j \in I_2: X^j_{|X_1,S}=X_1B^j_1+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n); \label{SR}
		\end{equation}
		where $B_1^j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
	\item  $p-p_2$ remaining independent exogenous variables:
		\begin{equation}
			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)%=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1v0}
		\end{equation}
\end{itemize}

We make the hypothesis of the uncrossing rule $I_1\cap I_2=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus we have a partition $X=[X_1,X_2]$.

This generative model is conditionnal to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian. These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $B_1^j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
\\
If there are exact regressions ($\sigma^2_j=0$) in (\ref{SR}), the structure won't be identifiable. But it only means that several structure will have the same likelihood and they will have the same interpretation. So it's not really a problem. Moreover, when an exact subregression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete. CorReg (Our R package) prints a warning to point out exact regressions when found.

\subsection{Estimator and properties}
	We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $X_2$ a regression model on $Y$ {\it depending only on uncorrelated covariates $X_1$}:
	\begin{equation}
		Y=X_1 (A_{1}+ \sum_{j \in I_2}B^{j}_{1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
		 = X_1\alpha_1+ \varepsilon_{\alpha}\label{Trueexpl}
	\end{equation}			
		So we have the unbiased estimator: 
		\begin{equation}
			\hat{\alpha}_{1} = (X'_{1} X^{1})^{-1}X'_{1}Y % \textrm{ and }\hat{\alpha}_{2} = 0
		\end{equation}
		with variance:
		\begin{equation}
			\operatorname{Var}[\hat{\alpha}_{1}|X]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X^{1})^{-1} % \textrm{ and }\operatorname{Var}[\hat{\alpha}_{2}|X]= 0 
		\end{equation}
		We see that the variance is reduced (no correlations and smaller matrix give better conditioning) for small values of $\sigma_j$ $i.e.$ strong correlations.					
			%There is no theoretical guarantee that our model is better. It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis. Therefore we made some simulations to compare both methods (see the end of this paper).
			Moreover, the explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset. Expert knowledge can even be added to the structure.
			 This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used. Hence we hope to obtain a parsimonious model with { \it two kinds of zeros}: those from decorrelating step and those from selection step, with different meanings.
			 \\
			 The \textsc{CorReg} package proposes to make selection with both LASSO (with LAR), elasticnet, adaptative lasso, and others more classical methods. A last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}).

	\subsection{Advantages for interpretation}

		Grouping effect is the fact that correlated covariates get similar coefficient and are selected together. It's the case with elasticnet \cite{zou2005regularization} for example.	
			If $X_2=X_1+e$ and we have the grouping effect, we will obtain a model like $\hat{Y}=\hat{a}X_1+\hat{a}X_2$. 
			Then, if you try to modify the response value, you will modify one of the covariates and both will change so you won't get expected results.

		At the opposite, methods like the LASSO with LAR will only keep one covariate for each group of correlated covariates  so we get $\hat{Y}=2aX_1$ and think that $X_2 \perp Y$ but $Y$ depends on $X_2$.

			Nothing constrains us to give only one equation. It is clearly better to give the user another equation (or system for more complex models) describing the correlations. 
			So you get the following model : $Y=aX_1+aX_2$ and $X_1=X_2+e$. Then you have more information and are able to decide better actions. 
			With such a model, grouping effect is no more useful because when saying $Y=2aX_1$ and $X_2=X_1+e$ with the information that $X_2$ has been removed locally because of its correlation with $X_1$, you don't get misleading interpretations anymore. In fact, we have two kinds of zeros (those from selection and those from correlations) and we know for each zero which kind is it.
			So it is possible to combine the advantages of grouping effect and selection just giving several equations. 
			Each equation here is very simple (only linear regressions) so you don't really increase complexity of the model. 
			Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable.

	\section{Estimating structure of sub-regressions with a Markov chain}

	\subsection{Structure comparison} \label{compstruct}
  All our work is based on $S$, the linear structure between the covariates.
	Our generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood according to the complexity of the structure~\cite{BIChuard}. We will prefer this kind of comparison criterion instead of cross-validation that is very time-consuming and thus not friendly with combinatory problematics.
	We note $\Theta$ the set of the parameters of the generative model
	\begin{eqnarray}
		P(S|X)&\propto& P(X|S)P(S) \\
		BIC&=&-2\mathcal{L}(X,S,\Theta)+|\Theta|\log(n) \approx -2\log P(X|S)
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus \mbox{\textsc{CorReg}} allows to choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	\textsc{CorReg} adds this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ in the MCMC when hierarchical hypothesis is made (it will be the case in the followings).			
	But we can imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context. 
\\

		We will now use the following notation: $\psi(S)=\psi(X|S)$	where $\psi(X|S)$ is the chosen criterion, that will be $BIC$ with hierarchical uniform hypothesis in our numerical results.
\subsection{Detailed generative model}

	To better fit industrial variables (Figure \ref{graphMixmod}), we suppose  that variables in $X_1$ follow Gaussian mixtures. The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust. But one can use other laws if needed. Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.
\begin{equation}
			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1}
		\end{equation}
		Variables in $X_1$ are supposed to be independent.
	Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to compute corresponding $\psi$ separately, give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).

	\subsection{The neighbourhood}
		\subsubsection{Definition}
	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
	\\
	For each step, starting from $S \in \mathcal{S}$ we define a neighbourhood:
		\begin{eqnarray}
		\mathcal{V}_{S,j}&=& \{S \}\cup \{ S^{(i,j)} |1\leq i \leq p, i\neq j  \} \\
		\textrm{where }\ \ j &\sim & \mathcal{U}(\{1,\dots,p\}) 
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_i^j$ (add): 
			\begin{itemize}
				\item $I_1^j=I_1^j\cup \{i\}$
				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item else (remove): $I_1^j=I_1^j\setminus \{i\}$
	\end{itemize}

	\smallskip
	At every moment, coherence between $I_1$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_1^j=|I_1^j|$, $I_2=\{j |p_1^j>0 \}$, $p_2= |I_2|$, .
		\subsubsection{Alternatives}
	We have here at each step $|\mathcal{V}_{S,j}|=p$ candidates but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the internal regressions or the whole structure, a maximum number of internal regressions, {\it etc.}). \textsc{CorReg} allows to modify this neighbourhood to better fit users constraints. Relaxation (column-wise and row-wise) is optional but gives more stability to the number of feasible candidates at each step and allows to modify several parts of $I_1$ in only one step when needed. Hence it improves efficiency by a significant reinforcement of the irreductibility of the Markov chain. Rejecting candidates instead of doing the relaxation steps will  however reduce the number of evaluated candidates and thus accelerate the walk. So it can be used for a warming phase when $n$ is great and time is missing.

	The hierchical uniform hypothesis made above for $P(S)$ implies $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ so candidates may be rejected to satisfy this hypothesis. Strongest constraints on $p_2$ and/or $p_1$ can be given in \textsc{CorReg} if relevant.

If the algorithm did not have time to converge, it can be continued with a few step for which the neighboorhood would only contain smaller candidates (in terms of complexity). It is equivalent to ask for each element in $I_1$ if the criterion $\psi$ would be better without it. Thus it can be seen as a final cleaning step. But in fact, it's just continuing the MCMC with a reduced neighbourhood.	
	\subsection{The walk}
	\subsubsection{Transition probabilities}
	We first make the approximation
	\begin{equation}
		P(S|X)\approx exp(\psi(S)).
	\end{equation}
	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
	At each step the markov chain moves with probabiliy:
	\begin{eqnarray}
			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2} \psi(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}\psi(S_l))} \\
	\end{eqnarray}
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité

Because the walk follows a regular and thus ergodic markov chain with a finite state space, it has exactly one stationary distribution \cite{grinstead1997introduction} %: $\pi$ and every rows of $\operatorname{lim}_{k\rightarrow \infty}\mathcal{P}^k=W$ equals $\pi$.
%	
%	
%	With $\forall S \in  \mathcal{S}$ :
%	\begin{eqnarray}		
%		0 \leq &\pi (S)& \leq 1 \nonumber \\
%		\sum_{S \in \mathcal{S}}\pi(S) &=&1 \nonumber \\
%		\pi (S) &=&\sum_{\tilde{S}\in \mathcal{S}} \pi(\tilde{S})\mathcal{P}(\tilde{S},S) \\%définition de la lois stationnaire
%	\end{eqnarray}
%		
and the output will be the best structure in terms of $\psi$ which weights each candidate. Practically speaking, \textsc{CorReg} returns the best structure seen during the walk.
Numerical results (Section 4) illustrates the efficiency of the walk when the true model really contains a linear structure or no structure at all (Table (\ref{compZvrai})) and when the structure is not linear (Table \ref{compZnonlin})).

 \subsubsection{Initialisation(s)}

 If we have some knowledge about some sub-regressions (physical models for example) we can add them in the found and/or initial structure. So the model is really expert-friendly.
The initial structure can be based on a first warming algorithm taking the correlations into account. coefficients are randomly placed into $I_1$, weighted by the absolute value of the correlations. We do so in the followings. Then this structure can be reduced by the hadamard product with the binary matrix obtained by Graphical Lasso\cite{friedman2008sparse}. Graphical LASSO is time consuming so we only used it on real datasets (each simulation was computed 100 times).

	One would rather test multiple short chains than lose time in initialisation or long chains \cite{gilks1996markov}. It also helps to face local extrema. In the followings, the chain was launched with twenty initialisations.

\section{Numerical results on simulated datasets}
	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} show the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} et \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation ,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$. The $B_1^j$ are generated according to the same Poisson law but with a random sign. $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.  
	We used \textsc{Rmixmod} to estimate the densities of each covariate.

		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure ?}
			The first criterion is $\psi$ which is minimised in the MCMC. But in our case, $\psi$ is based on the likelihood whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
			Global indicators :
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_2\cap \hat{I}_2|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_2|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_2|-TL$
				\item $\Delta p_2$ : the gap between the number of sub-regression in both model : $\Delta p_2=|I_2|-|\hat{I}_2|$. The sign defines if $\hat{S}$ is too complex or too simple
				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_2}p_1^j-\sum_{j \in \hat{p}_2}\hat{p}_1^j$
			\end{itemize}
		\subsubsection{Results on $S$}	\label{compZ}
In table \ref{compZvrai} we compare found structures in different contexts with both Uniform (U) and Hierarchical Uniform (HU) a priori law on $P(S)$. 
We see that (HU) hypothesis gives sparser models even if $\max (\hat{p}_1^j)=5$. $p=40$ so the maximum value for $\hat{p_2}$ is $20$ in (HU) case and we see that this max is not reached. The stronger penalty implied by (HU) really is efficient. The datasets used for (U) and (HU) are the same to keep the comparison  meaningful.

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
$n$ & $p_2$ & $\psi$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
30 & 0 & U&0.4104 & 3.2428 & 0 & 5.43 & 0 & -5.43 & 22.55  \\
& & & (0.0275) & (0.3711) & (0) & (1.9346) & (0) & (1.9346) & (8.0884) \\
 &  &HU &0.4104 & 8.2338 & 0 & 0.53 & 0 & -0.53 & 2.27  \\
& & & (0.0275) & (0.9045) & (0) & (0.7844) & (0) & (0.7844) & (3.3024) \\
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&0.4182 & 2.7735 & 10.96 & 5.93 & 4.98 & -0.95 & 38.16  \\
& & & (0.0329) & (0.1498) & (1.9844) & (2.0313) & (1.9948) & (0.9987) & (6.499) \\
 &  &HU &0.4182 & 4.1876 & 11.61 & 4.57 & 4.33 & -0.24 & 16.48  \\
& & & (0.0329) & (0.2178) & (1.8743) & (1.9502) & (1.8752) & (0.4948) & (5.4892) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 0.4456 & 2.9154 & 25.23 & 1.92 & 6.5 & 4.58 & 28  \\
& & & (0.0429) & (0.1331) & (1.4761) & (1.0888) & (1.4668) & (0.9866) & (5.0831) \\
 &  & HU & 0.4456 & 4.0233 & 16.96 & 3.04 & 14.77 & 11.73 & 4.35  \\
& & & (0.0429) & (0.1091) & (1.3993) & (1.3993) & (1.4692) & (0.5478) & (5.8833) \\
\hline
\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
50 & 0 & U&0.5229 & 4.7068 & 0 & 4.2 & 0 & -4.2 & 13.35  \\
& & & (0.0519) & (0.5865) & (0) & (1.7233) & (0) & (1.7233) & (5.6468) \\
 &  & HU &0.5229 & 10.1198 & 0 & 0.13 & 0 & -0.13 & 0.32  \\
& & & (0.0519) & (0.5541) & (0) & (0.3667) & (0) & (0.3667) & (0.9732) \\
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&0.5205 & 3.3681 & 11.15 & 5.42 & 4.72 & -0.7 & 22.85  \\	
& & & (0.0451) & (0.3123) & (1.93) & (1.9132) & (1.886) & (0.7317) & (5.7742) \\
 &  &HU &0.5205 & 4.909 & 11.42 & 4.59 & 4.45 & -0.14 & 7.55  \\
& & & (0.0451) & (0.4556) & (1.9079) & (1.8968) & (1.8333) & (0.3487) & (4.0611) \\
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U& 0.5833 & 3.2683 & 28.17 & 1.38 & 3.7 & 2.32 & 12.74  \\
& & & (0.0628) & (0.316) & (1.3711) & (0.9077) & (1.3143) & (0.8394) & (4.3359) \\
 &  &HU &0.5833 & 4.3599 & 17.27 & 2.73 & 14.6 & 11.87 & -2.61  \\
& & & (0.0628) & (0.3729) & (1.1708) & (1.1708) & (1.2792) & (0.338) & (4.4854) \\
\hline
\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
100 & 0 & U& 0.9623 & 12.9373 & 0 & 2.83 & 0 & -2.83 & 6.23  \\
& & & (0.077) & (1.7778) & (0) & (1.2953) & (0) & (1.2953) & (3.1999) \\
 &  &HU & 0.9623 & 20.9817 & 0 & 0.01 & 0 & -0.01 & 0.02  \\
& & & (0.077) & (1.9421) & (0) & (0.1) & (0) & (0.1) & (0.2) \\
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U& 1.1223 & 6.9647 & 11.67 & 4.8 & 4.25 & -0.55 & 12.58  \\
& & & (0.1122) & (0.5473) & (2.0003) & (2.0646) & (1.956) & (0.7833) & (3.9471) \\
 &  &HU & 1.1223 & 8.8486 & 12.04 & 3.95 & 3.88 & -0.07 & 3.75  \\
& & & (0.1122) & (0.7174) & (1.9223) & (1.9404) & (1.9137) & (0.2564) & (2.2625) \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U& 1.4343 & 5.9626 & 30.14 & 0.84 & 1.61 & 0.77 & 6.96  \\
& & & (0.2528) & (0.3136) & (1.3928) & (0.8495) & (1.2941) & (0.7086) & (3.0975) \\
 &  &HU & 1.4343 & 7.3741 & 17.49 & 2.51 & 14.26 & 11.75 & -3.76  \\
& & & (0.2528) & (0.2748) & (1.1849) & (1.1849) & (1.2441) & (0.4794) & (4.4859) \\
\hline
\end{tabular} 
\caption{Results of the Markov chain with  constraint $\hat{p}_1\leq 5$. Mean observed and standard deviation (sd). } \label{compZvrai}
\end{table}
With no constraints on $\hat{p}_1$ when $n<p$ we can have a perfect overlearning, so it is recommended to constrain each $\hat{p}_1^j <n$. Table \ref{compZvraip1max5} shows the impact on the research of such a constraint.

For  \ref{compZnonlin} we have generated non linear structures. Variables in $X_2$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $X_1$ . Dependencies are still real but non linear.
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
$n$ & $p_2$ &  $BIC(S)$ & $BIC(\emptyset)$ &$BIC(\hat{S})$  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
\hline
30 & 16 & ? & ? & ?  & ?  &?  & ?  & ? & ? \\
& &(?) & (?) & (?) & (?) & (?) & (?) & (?) & (?) \\
\hline
\end{tabular} 
\caption{Results of the Markov chain for non linear structure ($\log$ and square). Mean observed and standard deviation (sd). } \label{compZnonlin}
\end{table}

\clearpage

	\subsection{$Y$ depends on all variables in $X$}	 \label{tableMSEsimtout}	
We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure). The datasets used here were those from \ref{compZvrai}.
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$& \psi & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
30 & 0 & U& MSE (sd) & 262627.57 (732019) & 5928332.4 (49005690.2) & 262627.57 (732019) \\
& & & cpl (sd) & 30 (0) & 29.99 (0.1) & 30 (0) \\
 &  &HU &MSE (sd) & 262627.57 (732019) & 10381962.64 (90962496.9) & 262627.57 (732019) \\
& & & cpl (sd) & 30 (0) & 30 (0) & 30 (0) \\
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 510747.53 (2287539.8) & 635.42 (335.2) & 610.32 (424.9) \\
& & & cpl (sd) & 30 (0) & 24.11 (1) & 25.06 (0.2) \\
 &  &HU &MSE (sd) & 510747.53 (2287539.8) & 603.24 (415.6) & 610.32 (424.9) \\
& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25.06 (0.2) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
50 & 0 & U&
& & & 
 &  & HU &
& & & 
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
100 & 0 & U&   \\
& & & \\
 &  &HU &   \\
& & &  \\
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins. When $n<p$ the dataset was reduced to $p=n$ automatically by a $QR$ decomposition as the lm function of R does. Without selection, all models have min$(n,p)$ non-zero coefficients.} \label{YXlinOLS}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
30 & 0 & U&MSE (sd) & 1246.35 (350.5) & 1433.98 (526.7) & 1246.35 (350.5) \\
& & & cpl (sd) & 17.84 (5.5) & 15.79 (5.6) & 17.84 (5.5) \\
 &  &HU &MSE (sd) & 1246.35 (350.5) & 1248.84 (341.4) & 1246.35 (350.5) \\
& & & cpl (sd) & 17.84 (5.5) & 17.63 (5.6) & 17.84 (5.5) \\ \\
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&
& & & 
 &  &HU &
& & & 
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & MSE (sd) & 712.79 (405.2) & 566.32 (226.4) & 554.01 (257.1) \\
& & & cpl (sd) & 16.68 (4.4) & 15.35 (4.3) & 16.14 (4.5) \\
 &  & HU & MSE (sd) & 712.79 (405.2) & 551.92 (242.4) & 554.01 (257.1) \\
& & & cpl (sd) & 16.68 (4.4) & 15.96 (4.4) & 16.14 (4.5) \\
\hline
\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
50 & 0 & U&
& & & 
 &  & HU &
& & & 
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
100 & 0 & U&   \\
& & & \\
 &  &HU &   \\
& & &  \\
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinLASSO}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
30 & 0 & U&MSE (sd) & 1326.54 (388.4) & 1356.98 (331.4) & 1326.54 (388.4) \\
& & & cpl (sd) & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
 &  &HU & MSE (sd) & 1326.54 (388.4) & 1307.96 (356.6) & 1326.54 (388.4) \\
& & & cpl (sd) & 12.14 (5.1) & 12.27 (5.1) & 12.14 (5.1) \\
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 1400.56 (1598.2) & 668.57 (274.6) & 653.59 (283.7) \\
& & & cpl (sd) & 13.86 (6.8) & 14.31 (5.7) & 14.83 (5.3) \\
 &  &HU &MSE (sd) & 1400.56 (1598.2) & 643.25 (277.9) & 653.59 (283.7) \\
& & & cpl (sd) & 13.86 (6.8) & 14.83 (5.6) & 14.83 (5.3) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
50 & 0 & U&
& & & 
 &  & HU &
& & & 
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
100 & 0 & U&   \\
& & & \\
 &  &HU &   \\
& & &  \\
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinenet}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
30 & 0 & U& MSE (sd) & 1919.43 (861.8) & 2014.13 (683.6) & 1919.43 (861.8) \\
& & &cpl (sd) & 22.8 (3.3) & 19.34 (4.9) & 22.8 (3.3) \\
 &  &HU &MSE (sd) & 1919.43 (861.8) & 1885.39 (814) & 1919.43 (861.8) \\
& & & cpl (sd) & 22.8 (3.3) & 22.73 (3.1) & 22.8 (3.3) \\
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 696.46 (492.9) & 662.07 (285.9) & 660.93 (354) \\
& & & cpl (sd) & 15.9 (3.8) & 15.1 (3.7) & 15.76 (3.7) \\
 &  &HU &MSE (sd) & 696.46 (492.9) & 655.94 (325.7) & 660.93 (354) \\
& & & cpl (sd) & 15.9 (3.8) & 15.72 (3.6) & 15.76 (3.7) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
50 & 0 & U&
& & & 
 &  & HU &
& & & 
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
100 & 0 & U&   \\
& & & \\
 &  &HU &   \\
& & &  \\
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinstep}
\end{table}

\clearpage
	\subsection{$Y$ depends only on covariates in $X_2$ (worst case for us)}	 \label{tableMSEsimgauche}
We now try the method with a response depending only on variables in $X_2$. The datasets used here were still those from \ref{compZvrai}.
Depending only on $X_2$ imply sparsity and impossibility to obtain the true model when using the true structure.

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 81781.54 (247281.7) & 507.1 (269.2) & 593.03 (428.3) \\
& & & cpl (sd) & 30 (0) & 24.11 (1) & 25.06 (0.2) \\
 &  &HU &MSE (sd) & 81781.54 (247281.7) & 587.07 (425.6) & 593.03 (428.3) \\
& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25.06 (0.2) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X2$. } \label{YX2linOLS}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 367.45 (198.4) & 298.32 (122.4) & 292.53 (102.2) \\
& & & cpl (sd) & 14.37 (5.1) & 12.9 (3.8) & 12.84 (3.8) \\
 &  &HU &MSE (sd) & 367.45 (198.4) & 298.1 (119.1) & 292.53 (102.2) \\
& & & cpl (sd) & 14.37 (5.1) & 12.81 (3.8) & 12.84 (3.8) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X2$. }\label{YX2linLASSO}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 499.32 (218.8) & 311.41 (137.4) & 305.09 (123.5) \\
& & & cpl (sd) & 14.08 (6.1) & 11.75 (4.8) & 11.49 (4.7) \\
 &  &HU &MSE (sd) & 499.32 (218.8) & 307.6 (128.7) & 305.09 (123.5) \\
& & & cpl (sd) & 14.08 (6.1) & 11.63 (4.7) & 11.49 (4.7) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X2$.}\label{YX2linenet}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
$n$ & $p_2$&  \psi &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & U&MSE (sd) & 394.59 (454.6) & 341.47 (153.2) & 351.14 (148) \\
& & & cpl (sd) & 13.09 (3.2) & 12.9 (2.9) & 12.92 (3) \\
 &  &HU &MSE (sd) & 394.59 (454.6) & 353.99 (157) & 351.14 (148) \\
& & & cpl (sd) & 13.09 (3.2) & 12.83 (2.9) & 12.92 (3) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & U & 
& & & 
 &  & HU & 
& & & 
\hline
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & U&	
& & & 
 &  &HU &
& & & 
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & U&
& & & 
 &  &HU &
& & &
\hline
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & U&  \\
& & &  \\
 &  &HU &  \\
& & & \\
\hline
\end{tabular} 
\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X2$.}\label{YX2linstep}
\end{table}

	\clearpage	
\section{Numerical results on real datasets}
\subsection{Quality case study}
This work takes place in steel industry context, with quality oriented objective : to understand and prevent quality problems on finished product, knowing the whole process.

We have :
		\begin{itemize}
			\item a quality parameter (confidential) as response variable,
			\item 205 variables from the whole process to explain it.
			\item The stakes : a hundred euros per ton (for information: Dunkerque's  site aims to produce up to 7.5 millions tons a year)
		\end{itemize}

\begin{figure}[h!]
	\begin{minipage}[l]{.30\linewidth}
			\includegraphics[height=150px,width=150px]{figures/mixmod.png} 
			\caption{Example of non-Gaussian real variable easily modeled by a Gaussian mixture}\label{graphMixmod}
	\end{minipage} \hfill
	\begin{minipage}[c]{.30\linewidth}
			\includegraphics[height=150px, width=150px]{figures/histR2exfos.png} 
			\caption{$R^2_{adj}$ of the 76 sub-regressions.}
	\end{minipage} \hfill
   \begin{minipage}[r]{.30\linewidth}
			\includegraphics[height=150px,width=150px]{figures/correlexfoshist.png} 
			\caption{Histogram of correlations in $X$.} \label{compareMSEexfos}
   \end{minipage}
\end{figure}   			
	We get a training set of $n=3 000$ products described by $p=205$ variables from the industrial process and a validation sample of $847$ products.
	Let's note $\rho$ the absolute value of correlations between two covariates. Industrial variables are naturally highly correlated as the width and the weight of a steel slab ($\rho=0.905$), the temperature before and after some tool ($\rho=0.983$), the  roughness of both faces of the product ($\rho= 0.919$), a mean and a max ($\rho=0.911$). \textsc{CorReg} also found more complex structures describing physical models, like   Width = f (Mean.flow , Mean.speed.CC) even if the true Physcial model is not linear : Width = flow / (speed * thickness) (here thickness is constant). Regulation models used to optimize the process were also found. These first results are easily understandable and meet metallurgists expertise.  
			The algorithm gives a structure of $p_2=76$ subregressions with a mean of $\bar{p_1}=5.17$ regressors. In $X_1$ the number of $\rho>0.7$ is $\textbf{79.33\%}$ smaller than in $X$.

			It is now time to look at the predictive results (Figure \ref{compareMSEexfos}).
				The best model found when not using \textsc{CorReg} is given by the LASSO. But when using \textsc{CorReg} elasticnet produces a better model in terms of prediction. LASSO gives a model with 21 non-zero coefficients and elasticnet with \textsc{CorReg} gives a model with 40 non-zero parameters but $6.40\%$ better in prediction on the validation sample (847 products). $14$ non-zero coefficients are common between the two models.
				Elasticnet alone get a model with 78 parameters that is improved by $9.75\%$ in prediction when used with \textsc{CorReg}. When using LASSO with \textsc{CorReg} we obtain a model with 24 non-zero coefficients that is $4.11\%$ better than LASSO alone. We also computed the OLS model (without selection) and the naive one (estimating the response by the mean of the learning set). All the MSE were modified here to obtain a value of 100 for the best (to preserve confidentiality). Elasticnet with \textsc{CorReg} is $13.51\%$ better than OLS.
		\begin{figure}[h]
			\centering
				\label{compareMSEexfos}
				\includegraphics[width=400px]{figures/MSEfinal.png}
			\caption{MSE comparison on industrial dataset. Learning set : 3 000 products, validation set : 847 products}
		\end{figure}		
		\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|}
	\hline 
	Model & MSE & Complexity (with intercept) \\ 
	\hline 
	OLS & 115.63 & 206 \\ 
	\hline 
	\textsc{CorReg} + OLS & 109.59&130 \\ 
	\hline  
	LASSO & 106.84 & 21 \\ 
	\hline 
	\textsc{CorReg} + LASSO & 102.45 &24 \\ 
	\hline 
	elasticnet & 110.81 & 78\\ 
	\hline 
	\textsc{CorReg} + elasticnet & 100 &40 \\ 
	\hline 
\end{tabular} 
\caption{Results obtained on a validation sample.}	
\end{table}

		In terms of interpretation, the main regression comes with the family of regression so it gives a better understanding of the consequences of corrective actions on the whole process. It typically permits to determine the \textit{tuning parameters} whereas LASSO would point variables we can't directly act on.	So it becomes easier to take corrective actions on the process to reach the goal. The stakes are so important that even a little improvement leads to consequent benefits, and we don't even talk of the impact on the market shares that is even more important.
		\clearpage

		\subsection{Production case study}
This second example is about a phenomenon that impacts the productivity of a steel plan.
We have :
		\begin{itemize}
			\item a (confidential)  response variable,
			\item $p=145$ variables from the whole process to explain it but only $n=100$ individuals.
			\item The stakes : $20\%$ of productivity to gain
		\end{itemize}

\begin{figure}[h!]
	\begin{minipage}[l]{.30\linewidth}
			\includegraphics[width=150px]{figures/correl0BVBI.png} 
			\caption{Correlations between the covariates}
	\end{minipage} \hfill
	\begin{minipage}[c]{.30\linewidth}
			\includegraphics[height=150px, width=150px]{figures/R2corregBVBI.png} 
			\caption{$R^2_{adj}$ of the 67 sub-regressions.}
	\end{minipage} \hfill
   \begin{minipage}[r]{.30\linewidth}
			\includegraphics[height=150px,width=150px]{figures/histcorrelBVBI.png} 
			\caption{Histogram of correlations in $X$.} 
   \end{minipage}
\end{figure}   			
	Here $n<p$ so we only have the leave one out cross-validation MSE.
	\textsc{CorReg} improves LASSO by $5.24\%$ and elasticnet by $8.60\%$. \textsc{CorReg} combined with LASSO gives the best result but it is only a leave on out MSE.
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|}
	\hline 
	Model & MSE & Complexity (with intercept) \\ 
	\hline 
	LASSO & 105.54 & 34 \\ 
	\hline 
	\textsc{CorReg} + LASSO & 100 & 18 \\ 
	\hline 
	elasticnet & 129.94 & 13 \\ 
	\hline 
	\textsc{CorReg} + elasticnet & 118.76 & 21 \\ 
	\hline 
\end{tabular} 
\caption{Results obtained with leave-one out cross-validation. $n=100, p=145$.}	
\end{table}
In this precise case, \textsc{CorReg} found a structure that helped to decorrelate covariates in interpretation and to find the relevant part of the process to optimize.

\section{Conclusion and perspectives}
	We have seen that correlations can lead to serious estimation and variable selection problems in linear regression and that in such a context, it can be useful to explicitly model the structure between the covariates and to use this structure (even sequentially) to avoid correlations issues. We also show that real industrial context faces this kind of situations so our model can help to interprete and predict physical phenomenon efficiently and to help to manage missing values. But for now we still need a full dataset to learn the structure between the covariates and even if correlations are strong, some information is lost. Further work is needed to face these two challenges.

	\textsc{CorReg} is accessible on CRAN and has already proved its efficiency on real regression problematics in industry. \textsc{CorReg}'s strength is its great interpretability of the model, composed of several short linear regression easily managed by non-statisticians while strongly reducing correlations issues that are everywhere in industry.
	Nevertheless, we need to enlarge its application field to missing values, also very commons in industry. The actual generative model allows such a functionality without supplementary hypothesis and this also is a strength of \textsc{CorReg}.

	Another perspective would be to take back lost information (the residual of each sub-regression) to improve predictive efficiency when needed. It would only consists in a second step of linear regression between the residuals and would thus still be able to use any selection method.

	This paper only treats linear regression but such a pretreatment could be used for logistic regression, {\it etc.}
	So the subject is still wide opened.

\bibliography{biblio}{}
\bibliographystyle{plain}

\end{document}

### 2014-03-14 www/article13.tex
update numerical results
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@448 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \hline 
-$n$ & $p_2$& \psi & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$& $\psi$ & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 178323.95 (1426610.2) & 180.02 (44.9) & 141.03 (27.8) \\
+& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9.27 (0.5) \\
+ &  & HU & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
+& & & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
 \hline
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&
-& & & 
- &  & HU &
-& & & 
+50 & 0 & U&MSE (sd) & 528.08 (228.6) & 886.54 (364.3) & 528.08 (228.6) \\
+& & & cpl (sd) & 41 (0) & 36.8 (1.7) & 41 (0) \\
+ &  & HU &MSE (sd) & 528.08 (228.6) & 542.29 (235) & 528.08 (228.6) \\
+& & & cpl (sd) & 41 (0) & 40.87 (0.4) & 41 (0) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&MSE (sd) & 612.72 (291.5) & 239.17 (89.4) & 200.32 (42.6) \\	
+& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25.13 (0.4) \\
+ &  &HU &MSE (sd) & 612.72 (291.5) & 207.6 (51.3) & 200.32 (42.6) \\
+& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25.13 (0.4) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 555.44 (262.5) & 128.98 (18.1) & 121.08 (11.9) \\
+& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9.13 (0.3) \\
+ &  &HU &MSE (sd) & 555.44 (262.5) & 171.9 (31.1) & 121.08 (11.9) \\
+& & &cpl (sd) & 41 (0) & 21 (0) & 9.13 (0.3) \\
 \hline
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U&   \\
-& & & \\
- &  &HU &   \\
-& & &  \\
+100 & 0 & U& MSE (sd) & 167.71 (20.6) & 323.44 (124.1) & 167.71 (20.6) \\
+& & & cpl (sd) & 41 (0) & 38.17 (1.3) & 41 (0) \\
+ &  &HU &MSE (sd) & 167.71 (20.6) & 167.98 (20.8) & 167.71 (20.6) \\  
+& & & cpl (sd) & 41 (0) & 40.99 (0.1) & 41 (0) \\
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U&  MSE (sd) & 168.68 (22.4) & 158.51 (51.6) & 133.49 (12.2) \\
+& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25.08 (0.3) \\
+ &  &HU &  MSE (sd) & 168.68 (22.4) & 137.03 (20.3) & 133.49 (12.2) \\
+& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25.08 (0.3) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U&  MSE (sd) & 173.25 (22.9) & 112.8 (10.7) & 110.63 (6.9) \\\\
+& & & cpl (sd) & 41 (0) & 10.02 (0.8) & 9.25 (0.5) \\
+ &  &HU & MSE (sd) & 173.25 (22.9) & 127.4 (11.9) & 110.63 (6.9) \\
+& & & cpl (sd) & 41 (0) & 21 (0) & 9.25 (0.5) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&
-& & & 
- &  &HU &
-& & & 
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 712.79 (405.2) & 566.32 (226.4) & 554.01 (257.1) \\
+30 &16& U & MSE (sd) & 712.79 (405.2) & 566.32 (226.4) & 554.01 (257.1) \\
 & & & cpl (sd) & 16.68 (4.4) & 15.35 (4.3) & 16.14 (4.5) \\
@@
 & & & cpl (sd) & 16.68 (4.4) & 15.96 (4.4) & 16.14 (4.5) \\
+\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
+30 & 32 & U & MSE (sd) & 216.12 (128.3) & 157.78 (39.4) & 147.04 (28.4) \\
+& & & cpl (sd) & 11.23 (4.4) & 8.56 (1.8) & 7.91 (1.3) \\
+ &  & HU & MSE (sd) & 216.12 (128.3) & 178.88 (68.9) & 147.04 (28.4) \\
+& & & cpl (sd) & 11.23 (4.4) & 9.86 (3) & 7.91 (1.3) \\
 \hline
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&
-& & & 
- &  & HU &
-& & & 
+50 & 0 & U&MSE (sd) & 658.38 (221.4) & 872.32 (239.4) & 658.38 (221.4) \\
+& & & cpl (sd) & 28.2 (6) & 22.95 (5.8) & 28.2 (6) \\
+ &  & HU &MSE (sd) & 658.38 (221.4) & 652.72 (211.9) & 658.38 (221.4) \\
+& & & cpl (sd) & 28.2 (6) & 28.36 (5.9) & 28.2 (6) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&MSE (sd) & 274.34 (101.7) & 268.76 (106.3) & 225.64 (67.5) \\	
+& & & cpl (sd) & 22.14 (4.1) & 19.52 (2.9) & 20.47 (2.8) \\
+ &  &HU &MSE (sd) & 274.34 (101.7) & 233.8 (78.9) & 225.64 (67.5) \\
+& & & cpl (sd) & 22.14 (4.1) & 20.19 (2.8) & 20.47 (2.8) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 165.6 (73.7) & 128.34 (18.8) & 123.73 (12.8) \\
+& & & cpl (sd) & 12.93 (6.4) & 8.51 (1.5) & 8.09 (1.1) \\
+ &  &HU &MSE (sd) & 165.6 (73.7) & 135.84 (24.1) & 123.73 (12.8) \\
+& & &cpl (sd) & 12.93 (6.4) & 10.17 (3.1) & 8.09 (1.1) \\
 \hline
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U&   \\
-& & & \\
- &  &HU &   \\
-& & &  \\
+100 & 0 & U&  MSE (sd) & 183.33 (31.1) & 357.15 (133.4) & 183.33 (31.1) \\
+& & & cpl (sd) & 37.78 (2.4) & 32.93 (3.7) & 37.78 (2.4) \\
+ &  &HU &  MSE (sd) & 183.33 (31.1) & 183.78 (31.6) & 183.33 (31.1) \\
+& & &  cpl (sd) & 37.78 (2.4) & 37.75 (2.4) & 37.78 (2.4) \\
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U& MSE (sd) & 148.83 (18.6) & 164.44 (54.3) & 139.57 (16.1) \\
+& & & cpl (sd) & 25.22 (4) & 21.72 (2) & 22.3 (1.7) \\
+ &  &HU &  MSE (sd) & 148.83 (18.6) & 142.97 (22.1) & 139.57 (16.1) \\
+& & & cpl (sd) & 25.22 (4) & 22.24 (1.8) & 22.3 (1.7) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U& MSE (sd) & 124.29 (19.9) & 113.36 (11) & 111.54 (7.2) \\
+& & & cpl (sd) & 13.18 (5.9) & 8.68 (1.2) & 8.53 (1) \\
+ &  &HU & MSE (sd) & 124.29 (19.9) & 118.33 (12) & 111.54 (7.2) \\
+& & & cpl (sd) & 13.18 (5.9) & 11.02 (2.9) & 8.53 (1) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
 30 & 0 & U&MSE (sd) & 1326.54 (388.4) & 1356.98 (331.4) & 1326.54 (388.4) \\
-& & & cpl (sd) & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
+& & & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
  &  &HU & MSE (sd) & 1326.54 (388.4) & 1307.96 (356.6) & 1326.54 (388.4) \\
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 855.74 (582.6) & 181.08 (51.9) & 146.79 (32.7) \\
+& & & cpl (sd) & 15.57 (6.3) & 11.19 (2.4) & 8.37 (1.3) \\
+ &  & HU & MSE (sd) & 855.74 (582.6) & 311.57 (163.9) & 146.79 (32.7) \\
+& & & cpl (sd) & 15.57 (6.3) & 14.87 (3.9) & 8.37 (1.3) \\
 \hline
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&
-& & & 
- &  & HU &
-& & & 
+50 & 0 & U&MSE (sd) & 738.94 (254.5) & 914.56 (283.7) & 738.94 (254.5) \\
+& & & cpl (sd) & 25.85 (8.7) & 20.97 (7.8) & 25.85 (8.7) \\
+ &  & HU &MSE (sd) & 738.94 (254.5) & 751.06 (262.2) & 738.94 (254.5) \\
+& & & cpl (sd) & 25.85 (8.7) & 25.43 (8.8) & 25.85 (8.7) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&	MSE (sd) & 516.36 (226.6) & 276.6 (133.6) & 228.85 (89.3) \\
+& & & cpl (sd) & 26.72 (7.4) & 20.49 (3.8) & 21.8 (3.3) \\
+ &  &HU &MSE (sd) & 516.36 (226.6) & 239.77 (103.7) & 228.85 (89.3) \\
+& & & cpl (sd) & 26.72 (7.4) & 21.57 (3.4) & 21.8 (3.3) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 328.7 (146.3) & 130.61 (20.6) & 124.01 (15.4) \\
+& & & cpl (sd) & 24.7 (7.1) & 9.95 (1.5) & 8.22 (1.2) \\
+ &  &HU &MSE (sd) & 328.7 (146.3) & 169.13 (40.2) & 124.01 (15.4) \\
+& & &cpl (sd) & 24.7 (7.1) & 16.04 (3.8) & 8.22 (1.2) \\
 \hline
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U&   \\
-& & & \\
- &  &HU &   \\
-& & &  \\
+100 & 0 & U& MSE (sd) & 175.93 (25.9) & 355.79 (143.8) & 175.93 (25.9) \\
+& & & cpl (sd) & 39.71 (1.7) & 34.11 (4.6) & 39.71 (1.7) \\
+ &  &HU &  MSE (sd) & 175.93 (25.9) & 176.2 (26.3) & 175.93 (25.9) \\
+& & &  cpl (sd) & 39.71 (1.7) & 39.69 (1.7) & 39.71 (1.7) \\
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U&  MSE (sd) & 173.65 (25.1) & 164.96 (59) & 139.51 (17.6) \\
+& & &  cpl (sd) & 35.54 (4.2) & 22.58 (2.2) & 23.26 (1.9) \\
+ &  &HU &  MSE (sd) & 173.65 (25.1) & 143.04 (23.2) & 139.51 (17.6) \\
+& & & cpl (sd) & 35.54 (4.2) & 23.16 (1.9) & 23.26 (1.9) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U&  MSE (sd) & 161.92 (25.2) & 113.61 (10.9) & 111.46 (7.2) \\
+& & &  cpl (sd) & 30.78 (6.4) & 9.25 (1.2) & 8.67 (1) \\
+ &  &HU &  MSE (sd) & 161.92 (25.2) & 127.44 (13.3) & 111.46 (7.2) \\
+& & &cpl (sd) & 30.78 (6.4) & 17.85 (2.7) & 8.67 (1) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 394.93 (356) & 155.18 (36.4) & 147.81 (30.3) \\
+& & &cpl (sd) & 16.1 (6.3) & 8.26 (1.7) & 7.77 (1.3) \\
+ &  & HU &  MSE (sd) & 394.93 (356) & 189.49 (79) & 147.81 (30.3) \\
+& & & cpl (sd) & 16.1 (6.3) & 10.23 (3) & 7.77 (1.3) \\
 \hline
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&
-& & & 
- &  & HU &
-& & & 
+50 & 0 & U&MSE (sd) & 745.88 (241.7) & 962.12 (319.3) & 745.88 (241.7) \\
+& & & cpl (sd) & 27.19 (4.6) & 22.9 (4.5) & 27.19 (4.6) \\
+ &  & HU &MSE (sd) & 745.88 (241.7) & 749.08 (244.2) & 745.88 (241.7) \\
+& & & cpl (sd) & 27.19 (4.6) & 27.21 (4.5) & 27.19 (4.6) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&	MSE (sd) & 279.82 (151.1) & 290.13 (117.3) & 239.61 (69.5) \\
+& & & cpl (sd) & 21.24 (5.2) & 18.13 (2.8) & 19.23 (2.7) \\
+ &  &HU &MSE (sd) & 279.82 (151.1) & 252.35 (108.5) & 239.61 (69.5) \\
+& & & cpl (sd) & 21.24 (5.2) & 18.97 (2.7) & 19.23 (2.7) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 199.3 (158.7) & 126.81 (17.4) & 124.19 (13) \\
+& & & cpl (sd) & 14.31 (6.2) & 8.16 (1.2) & 8.04 (1) \\
+ &  &HU &MSE (sd) & 199.3 (158.7) & 136.32 (25.2) & 124.19 (13) \\
+& & &cpl (sd) & 14.31 (6.2) & 9.59 (2.6) & 8.04 (1) \\
 \hline
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U&   \\
-& & & \\
- &  &HU &   \\
-& & &  \\
+100 & 0 & U&   MSE (sd) & 193.26 (34.9) & 387.18 (151.6) & 193.26 (34.9) \\
+& & & cpl (sd) & 36.3 (2.5) & 30.94 (4.2) & 36.3 (2.5) \\
+ &  &HU &   MSE (sd) & 193.26 (34.9) & 194.43 (37.3) & 193.26 (34.9) \\
+& & &  cpl (sd) & 36.3 (2.5) & 36.22 (2.7) & 36.3 (2.5) \\
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U&  MSE (sd) & 145.46 (20.1) & 165.74 (56.1) & 140.27 (17.5) \\
+& & & cpl (sd) & 23.44 (3.8) & 21.23 (2.1) & 21.93 (1.7) \\
+ &  &HU &  MSE (sd) & 145.46 (20.1) & 143.74 (23.2) & 140.27 (17.5) \\
+& & & cpl (sd) & 23.44 (3.8) & 21.86 (1.7) & 21.93 (1.7) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U& MSE (sd) & 128.44 (18.4) & 113.27 (10.7) & 111.77 (7.2) \\
+& & &  cpl (sd) & 13.58 (5.2) & 8.59 (1.1) & 8.48 (1) \\
+ &  &HU &  MSE (sd) & 128.44 (18.4) & 118.61 (11.5) & 111.77 (7.2) \\
+& & & cpl (sd) & 13.58 (5.2) & 10.75 (2.8) & 8.48 (1) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 74136.81 (297716.1) & 189.45 (48.1) & 144.51 (31.4) \\
+& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9.27 (0.5) \\
+ &  & HU & MSE (sd) & 74136.81 (297716.1) & 340.53 (169.4) & 144.51 (31.4) \\
+& & & cpl (sd) & 30 (0) & 21 (0) & 9.27 (0.5) \\
 \hline
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&	MSE (sd) & 684.04 (438.1) & 190.97 (37.7) & 197.32 (40) \\
+& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25.13 (0.4) \\
+ &  &HU &MSE (sd) & 684.04 (438.1) & 196.36 (40.5) & 197.32 (40) \\
+& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25.13 (0.4) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 596.32 (323) & 126.94 (15.5) & 119.39 (12.4) \\
+& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9.13 (0.3) \\
+ &  &HU &MSE (sd) & 596.32 (323) & 168.03 (27.7) & 119.39 (12.4) \\
+& & &cpl (sd) & 41 (0) & 21 (0) & 9.13 (0.3) \\
 \hline
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U&  MSE (sd) & 168.35 (20.1) & 133.75 (12.4) & 135.08 (13.1) \\
+& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25.08 (0.3) \\
+ &  &HU &  MSE (sd) & 168.35 (20.1) & 135.04 (13.1) & 135.08 (13.1) \\
+& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25.08 (0.3) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U&  MSE (sd) & 168.07 (21.4) & 109.61 (7.4) & 108.64 (7.2) \\
+& & &  cpl (sd) & 41 (0) & 10.02 (0.8) & 9.25 (0.5) \\
+ &  &HU & MSE (sd) & 168.07 (21.4) & 124.23 (11.1) & 108.64 (7.2) \\
+& & & cpl (sd) & 41 (0) & 21 (0) & 9.25 (0.5) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 273.12 (267.2) & 166.93 (49) & 152.45 (39.5) \\
+& & & cpl (sd) & 11.7 (5.8) & 8.7 (2.5) & 7.45 (1.4) \\
+ &  & HU & MSE (sd) & 273.12 (267.2) & 175.68 (57.5) & 152.45 (39.5) \\
+& & & cpl (sd) & 11.7 (5.8) & 9.19 (2.9) & 7.45 (1.4) \\
 \hline
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&	MSE (sd) & 189.64 (52.5) & 171.97 (45.9) & 176.58 (46.8) \\
+& & & cpl (sd) & 14.15 (3.9) & 13.4 (2.8) & 13.5 (3.1) \\
+ &  &HU &MSE (sd) & 189.64 (52.5) & 174.36 (45.3) & 176.58 (46.8) \\
+& & & cpl (sd) & 14.15 (3.9) & 13.57 (3) & 13.5 (3.1) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 163.88 (67.4) & 124.53 (17) & 121.99 (16.6) \\
+& & & cpl (sd) & 13.06 (7.2) & 8.29 (1.6) & 7.73 (1.2) \\
+ &  &HU &MSE (sd) & 163.88 (67.4) & 133.7 (26.4) & 121.99 (16.6) \\
+& & &cpl (sd) & 13.06 (7.2) & 9.65 (3.3) & 7.73 (1.2) \\
 \hline
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U& MSE (sd) & 129.65 (14.8) & 127.11 (12.1) & 127.46 (12.2) \\
+& & &  cpl (sd) & 14.86 (3.1) & 13.97 (2.5) & 14.04 (2.5) \\
+ &  &HU &  MSE (sd) & 129.65 (14.8) & 127.46 (12.2) & 127.46 (12.2) \\
+& & & cpl (sd) & 14.86 (3.1) & 14.02 (2.5) & 14.04 (2.5) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U&  MSE (sd) & 121.21 (20.1) & 109.66 (7.9) & 109.04 (7.8) \\
+& & & cpl (sd) & 12.84 (5.2) & 8.42 (1.1) & 8.11 (0.8) \\
+ &  &HU &  MSE (sd) & 121.21 (20.1) & 112.56 (11.2) & 109.04 (7.8) \\
+& & & cpl (sd) & 12.84 (5.2) & 9.77 (2.3) & 8.11 (0.8) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 691.55 (503.2) & 193.45 (54.2) & 150.17 (35) \\
+& & & cpl (sd) & 12.54 (7.2) & 10.26 (2.7) & 7.81 (1.4) \\
+ &  & HU & MSE (sd) & 691.55 (503.2) & 279.21 (117.3) & 150.17 (35) \\
+& & & cpl (sd) & 12.54 (7.2) & 13 (4.3) & 7.81 (1.4) \\
 \hline
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&	MSE (sd) & 282.05 (134.8) & 180.64 (47.6) & 181.49 (48.6) \\
+& & & cpl (sd) & 20.37 (6.3) & 13.68 (4) & 14.05 (4) \\
+ &  &HU &MSE (sd) & 282.05 (134.8) & 180.84 (47.5) & 181.49 (48.6) \\
+& & & cpl (sd) & 20.37 (6.3) & 14.03 (4) & 14.05 (4) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 308.13 (142.9) & 129.39 (19.6) & 122.71 (17.8) \\
+& & & cpl (sd) & 22.94 (7.7) & 9.37 (2.1) & 7.91 (1.4) \\
+ &  &HU &MSE (sd) & 308.13 (142.9) & 159.14 (30.4) & 122.71 (17.8) \\
+& & &cpl (sd) & 22.94 (7.7) & 15.42 (4) & 7.91 (1.4) \\
 \hline
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U& MSE (sd) & 150.14 (19.7) & 127.48 (14) & 128.21 (15) \\
+& & &  cpl (sd) & 25.6 (5.4) & 14.73 (3.8) & 14.73 (3.9) \\
+ &  &HU &  MSE (sd) & 150.14 (19.7) & 128.21 (14.9) & 128.21 (15) \\
+& & & cpl (sd) & 25.6 (5.4) & 14.7 (3.9) & 14.73 (3.9) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U& MSE (sd) & 157.18 (23.5) & 110.24 (8) & 109.24 (7.8) \\
+& & &  cpl (sd) & 30.41 (6.7) & 8.96 (1.3) & 8.3 (1) \\
+ &  &HU &  MSE (sd) & 157.18 (23.5) & 123.8 (12.5) & 109.24 (7.8) \\
+& & & cpl (sd) & 30.41 (6.7) & 17.18 (3.3) & 8.3 (1) \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  \psi &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $\psi$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 
-& & & 
- &  & HU & 
-& & & 
+30 & 32 & U & MSE (sd) & 410.11 (343.5) & 163.05 (45.4) & 152.08 (38.1) \\
+& & & cpl (sd) & 15.8 (6.5) & 8.29 (2.2) & 7.34 (1.3) \\
+ &  & HU & MSE (sd) & 410.11 (343.5) & 178.94 (69.7) & 152.08 (38.1) \\
+& & & cpl (sd) & 15.8 (6.5) & 8.89 (3) & 7.34 (1.3) \\
 \hline
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	
-& & & 
- &  &HU &
-& & & 
+50 & 16 & U&	MSE (sd) & 180.67 (52.7) & 173.02 (38.3) & 174.84 (36.8) \\
+& & & cpl (sd) & 13.56 (3.7) & 12.78 (2.4) & 13.07 (2.4) \\
+ &  &HU &MSE (sd) & 180.67 (52.7) & 175.3 (38.1) & 174.84 (36.8) \\
+& & & cpl (sd) & 13.56 (3.7) & 13.01 (2.4) & 13.07 (2.4) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&
-& & & 
- &  &HU &
-& & &
+50 & 32 & U&MSE (sd) & 188.91 (88.5) & 124.38 (17.4) & 122.38 (16.7) \\
+& & & cpl (sd) & 14.74 (6.6) & 8.07 (1.6) & 7.65 (1.2) \\
+ &  &HU &MSE (sd) & 188.91 (88.5) & 134.22 (23.6) & 122.38 (16.7) \\
+& & &cpl (sd) & 14.74 (6.6) & 9.5 (2.8) & 7.65 (1.2) \\
 \hline
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 16 & U&  MSE (sd) & 128.18 (13.7) & 127.43 (13.7) & 127.71 (13.3) \\
+& & &  cpl (sd) & 13.78 (2.8) & 13.51 (2) & 13.59 (2.1) \\
+ &  &HU & MSE (sd) & 128.18 (13.7) & 127.71 (13.3) & 127.71 (13.3) \\
+& & & cpl (sd) & 13.78 (2.8) & 13.59 (2.1) & 13.59 (2.1) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  \\
-& & &  \\
- &  &HU &  \\
-& & & \\
+100 & 32 & U& MSE (sd) & 126.15 (20.6) & 109.47 (8) & 109.04 (7.8) \\
+& & & cpl (sd) & 13.25 (4.4) & 8.3 (1) & 8.12 (0.8) \\
+ &  &HU &  MSE (sd) & 126.15 (20.6) & 113.96 (11.4) & 109.04 (7.8) \\
+& & & cpl (sd) & 13.25 (4.4) & 10.08 (2.6) & 8.12 (0.8) \\
 \hline

### 2014-03-14 www/article13.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@449 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{table}
-With no constraints on $\hat{p}_1$ when $n<p$ we can have a perfect overlearning, so it is recommended to constrain each $\hat{p}_1^j <n$.
-Table \ref{compZvraip1max5} shows the impact on the research of such a constraint.
-
-For  \ref{compZnonlin} we have generated non linear structures.
-Variables in $X_2$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $X_1$ .
-Dependencies are still real but non linear.
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-\hline
-$n$ & $p_2$ &  $BIC(S)$ & $BIC(\emptyset)$ &$BIC(\hat{S})$  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
-\hline
-30 & 16 & ? & ? & ?  & ?  &?  & ?  & ? & ? \\
-& &(?) & (?) & (?) & (?) & (?) & (?) & (?) & (?) \\
-\hline
-\end{tabular} 
-\caption{Results of the Markov chain for non linear structure ($\log$ and square). Mean observed and standard deviation (sd). } \label{compZnonlin}
-\end{table}
 
@@
 The datasets used here were those from \ref{compZvrai}.
+
+We observe that \textsc{CorReg} is better than classical methods especially whith the Hierarchical Uniform law.
+When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods.
+We also observe that simple methods like stepwise (here from the package \textsc{lars}) give good results in prediction.
+
+It is also notable that (HU) has a greater computational cost than (U). But next versions of the package may optimize a bit more the penalization step and reduce this gap.
 \begin{table}[h!]
@@
  &  &HU &MSE (sd) & 1246.35 (350.5) & 1248.84 (341.4) & 1246.35 (350.5) \\
-& & & cpl (sd) & 17.84 (5.5) & 17.63 (5.6) & 17.84 (5.5) \\ \\
+& & & cpl (sd) & 17.84 (5.5) & 17.63 (5.6) & 17.84 (5.5) \\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U&MSE (sd) & 1326.54 (388.4) & 1356.98 (331.4) & 1326.54 (388.4) \\
+30 & 0 & U&MSE (sd) & 1 326.54 (388.4) & 1 356.98 (331.4) & 1 326.54 (388.4) \\
 & & & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
@@
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 1400.56 (1598.2) & 668.57 (274.6) & 653.59 (283.7) \\
+30 & 16 & U&MSE (sd) & 1 400.56 (1598.2) & 668.57 (274.6) & 653.59 (283.7) \\
 & & & cpl (sd) & 13.86 (6.8) & 14.31 (5.7) & 14.83 (5.3) \\
- &  &HU &MSE (sd) & 1400.56 (1598.2) & 643.25 (277.9) & 653.59 (283.7) \\
+ &  &HU &MSE (sd) & 1 400.56 (1598.2) & 643.25 (277.9) & 653.59 (283.7) \\
 & & & cpl (sd) & 13.86 (6.8) & 14.83 (5.6) & 14.83 (5.3) \\
@@
 \end{tabular} 
-\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinenet}
+\caption{Elasticnet combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinenet}
 \end{table}
@@
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U& MSE (sd) & 1919.43 (861.8) & 2014.13 (683.6) & 1919.43 (861.8) \\
+30 & 0 & U& MSE (sd) & 1 919.43 (861.8) & 2 014.13 (683.6) & 1 919.43 (861.8) \\
 & & &cpl (sd) & 22.8 (3.3) & 19.34 (4.9) & 22.8 (3.3) \\
- &  &HU &MSE (sd) & 1919.43 (861.8) & 1885.39 (814) & 1919.43 (861.8) \\
+ &  &HU &MSE (sd) & 1 919.43 (861.8) & 1 885.39 (814) & 1 919.43 (861.8) \\
 & & & cpl (sd) & 22.8 (3.3) & 22.73 (3.1) & 22.8 (3.3) \\
@@
 We now try the method with a response depending only on variables in $X_2$. The datasets used here were still those from \ref{compZvrai}.
-Depending only on $X_2$ imply sparsity and impossibility to obtain the true model when using the true structure.
+Depending only on $X_2$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
 
@@
 \clearpage
+\subsection{Non linear case}
+ we have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
+Variables in $X_2$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $X_1$ .
+Dependencies are still real but non linear.
+\textsc{CorReg} still found dependencies.
+One example of non-linear structure found with \textsc{CorReg} on real datasets is given in section \ref{sectionexfos}.
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+\hline
+$n$ & $p_2$ & $\psi$ &Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
+\hline
+30 & 16 & U& 0.4313 & 2.7769 & 9.96 & 2.45 & 5.95 & 3.5 & 29.42  \\
+& & &(0.0391) & (0.0883) & (1.3175) & (1.3808) & (1.3056) & (1.453) & (6.2639) \\
+ &  & HU & 0.4313 & 4.9079 & 7.28 & 1.25 & 8.63 & 7.38 & 5.77  \\ 
+& & &(0.0391) & (0.4353) & (1.5769) & (0.9987) & (1.5548) & (1.6316) & (6.1297) \\ 
+\hline
+\end{tabular} 
+\caption{Results of the Markov chain for non linear structure ($\log$ and square). Mean observed and standard deviation (sd). } \label{compZnonlin}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline 
+$n$ & $p_2$&  $\psi$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+\hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
+30 & 0 & U&MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
+& & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\
+ &  &HU &MSE (sd) & 996.06 (323.7) & 1094.5 (621.8) & 1501.25 (905.9) \\
+& & & cpl (sd) & 17.66 (5.4) & 16.98 (4.8) & 11.97 (4.9) \\ 
+\hline
+\end{tabular} 
+\caption{ $Y$ depends on all variables in $X$. \textsc{CorReg} is not too far from LASSO.} \label{resYnonlin}
+\end{table}
+
+	\clearpage	
 \section{Numerical results on real datasets}
-\subsection{Quality case study}
+\subsection{Quality case study} \label{sectionexfos}
 This work takes place in steel industry context, with quality oriented objective : to understand and prevent quality problems on finished product, knowing the whole process.

### 2014-03-14 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@450 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@451 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X2$. } \label{YX2linOLS}
+\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$. } \label{YX2linOLS}
 \end{table}
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X2$. }\label{YX2linLASSO}
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$. }\label{YX2linLASSO}
 \end{table}
@@
 \end{tabular} 
-\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X2$.}\label{YX2linenet}
+\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$.}\label{YX2linenet}
 \end{table}
@@
 \end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X2$.}\label{YX2linstep}
+\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X_2$.}\label{YX2linstep}
 \end{table}
@@
 \textsc{CorReg} also found more complex structures describing physical models, like   Width = f (Mean.flow , Mean.speed.CC) even if the true Physcial model is not linear : Width = flow / (speed * thickness) (here thickness is constant).
-Regulation models used to optimize the process were also found.
+Non linear regulation models used to optimize the process were also found (but are confidential).
 These first results are easily understandable and meet metallurgists expertise.

### 2014-03-14 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@452 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		where $B_1^j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
-	\item  $p-p_2$ remaining independent exogenous variables:
-		\begin{equation}
-			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)%=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1v0}
-		\end{equation}
 \end{itemize}
+It remains $p-p_2$ independent exogenous variables : $\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)$
 
@@
 
-This generative model is conditionnal to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian.
+This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian.
 These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $B_1^j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
@@
 So it's not really a problem.
-Moreover, when an exact subregression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
-CorReg (Our R package) prints a warning to point out exact regressions when found.
+Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
+\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
 
@@
 
-		At the opposite, methods like the LASSO with LAR will only keep one covariate for each group of correlated covariates  so we get $\hat{Y}=2aX_1$ and think that $X_2 \perp Y$ but $Y$ depends on $X_2$.
+		At the opposite, methods like the LASSO with LAR will only keep one covariate for each group of correlated covariates  so we get $\hat{Y}=2aX_1$ and think that $X_2 \perp Y$ but $Y$ depends on $X_2$ so it is worse than grouping effect for interpretation.
 
@@
 			Each equation here is very simple (only linear regressions) so you don't really increase complexity of the model. 
-			Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable.
+			Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect).
 
@@
 $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.
-	We used \textsc{Rmixmod} to estimate the densities of each covariate.
+	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.
 

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@471 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-\author{Clément THERY}
-\title{CorReg pretreatment : Regression for correlated variables and application in steel industry}
+\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
+\title{CorReg: model based pretreatment for regression with correlated variables. Application in steel industry}
 
@@
 When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression.
-It provides a predictive model with a good interpretability and is simple to learn for non-statistician.
+It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn.
 Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
-Linear regression is a very classic situation that faces an also classical problem : the variance of the estimators.
-When estimating the parameters of the regression we have to compute the inverse of a matrix\cite{saporta2006probabilites} which will be ill-conditioned or even singular if some covariates depend linearly from each other.
-For a model defined by
+Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
+We note the linear regression model:
 \begin{equation}
@@
 	\end{equation}
-	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2)$ 
-	we have the following Ordinary Least Squares (OLS) estimators :
-	\begin{equation}
-		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
-	\end{equation}
-	With variance 
-	\begin{equation}
-		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1}
-	\end{equation}
-	And when correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance explodes.
-This variance increases based on two aspects :
+	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
+The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (OLS) by \cite{saporta2006probabilites}
+	\begin{equation}
+		\hat{\beta}=(X'X)^{-1}X'Y.
+\end{equation}
+
+When estimating the parameters of the regression we have to compute the inverse of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
+Variance of $\hat{\beta}$ increases based on two aspects :
 \begin{itemize}
@@
 Moreover, when you use more covariates, you increase the chance to have correlated ones.
-For example, this work takes place in an industrial (steel industry) context with a big set of covariates (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
+It will be illustrated with examples from steel industry in the last section of this paper with a big set of covariates (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
 In such a context, variance of the estimators can lead to arbitrary results or even no results at all.
-Prediction and interpretation are both strongly needed, with a preference for interpretation in industrial context (better to improve the process when possible than to only predict defects).
+Prediction and interpretation are both strongly needed, depending on the context.
+For instance, in an industrial context interpretation could be favored to improve the process instead of only predict defects.
 		~\\	~\\
@@
 	Because OLS is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
-Moreover, real datasets implies many irrelevant variables (datasets based on the whole process without any a priori) so we have to use variable selection methods.
+Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.
 
@@
 	\end{equation}
-	But Ridge regression is not efficient to select covariates (it's an assumed choice) because coefficients tends to 0 but don't reach 0.
-	So it gives difficult interpretations for large values of $p$ and is not adapted for our industrial context.
-We need to reduce the dimension of the model.
-Our goal is not just to predict but also to understand the response variable.
+	But Ridge regression do not aim at selecting covariates because coefficients tends to 0 but don't reach 0.
+	So it gives difficult interpretations for large values of $p$.
 
@@
 It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates.
-And it  really selects covariates with coeffients set exactly to 0.
+And it  really selects covariates with coefficients set exactly to 0.
 	 But LASSO also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates.
@@
 
-	Elastic net\cite{zou2005regularization}	is a method developped to be a compromise between Ridge regression and the LASSO. 
+	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the LASSO. 
 %	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
@@
 	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
-	It seems to give good predictions.
+	%It seems to give good predictions. 
 	But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
@@
 
- When some try to reduce the dimension and then just hope to have small correlations in the remaining dimensions,  we propose to focus on the correlations, giving a model with orthogonal covariates and an explicit structure between covariates.
+ Where some try to reduce the dimension and then just hope to have small correlations in the remaining dimensions,  we propose to focus on the correlations, giving a model with orthogonal covariates and an explicit structure between covariates.
 In fact we search the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
 This can be viewed as a pretreatment on the dataset allowing to use then other dimension reduction tools without suffering from correlations.
-We only consider strong correlations (i.e.
+%We only consider strong correlations (i.e.
 : problematic ones) thus we keep most of the information contained in the dataset.
@@
 However, Graphical LASSO can be used in the initialization step of our MCMC.
-This structure is based on gaussian mixture models to fit better real datasets and to allow identifiability of the structure in terms of complexity (number of parameters).
+This structure is based on a generative model to be able to compare likelihood-based criterion	 in the MCMC.
  	\\
@@
 
-\section{Model to decorrelate the covariates}
-
+\section{Model to select decorrelated covariates}
+\subsection{The decorrelation problem in regression}
+For a model defined by 
+	\begin{equation}
+		Y_{|X}=X\beta + \varepsilon
+	\end{equation}
+	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
+	we have the following Ordinary Least Squares (OLS) estimators:
+	\begin{equation}
+		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
+	\end{equation}
+	With variance 
+	\begin{equation}
+		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1} \label{eqOLS}
+	\end{equation}
+	 The Mean Squared Error (MSE) on $\hat{\beta}$ is:
+	\begin{eqnarray}
+		MSE(\hat{\beta}|\beta,X)&=&Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
+			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
+	\end{eqnarray}	 
+	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
 \subsection{Modelisation of the correlations}
-Let $Y \in \mathcal{R}^n$ be a response variable we want to explain with a set $X \in \mathcal{R}^{n\times p}$ of $p$ correlated covariates.
+Let now $X \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
 We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$.
@@
 
-We can now write the generative model:
-\begin{itemize}
-	\item Main regression between $Y$ and $X$:
-		\begin{equation}
-			Y_{|S}=XA+\varepsilon_Y =X_1A_1+X_2A_2+\varepsilon_Y \textrm{ whith } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
-		\end{equation}
-		where $A=(A_1,A_2) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $I_n$ the identity matrix, 
-	\item Family of $p_2$ regressions within correlated covariates in $X$:
-		\begin{equation}
-			\forall j \in I_2: X^j_{|X_1,S}=X_1B^j_1+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n); \label{SR}
+We define a family of $p_2$ regressions within correlated covariates in $X$:
+	\begin{equation}
+		\forall j \in I_2: X^j_{|X_1,S}=X_1B^j_1+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n) \label{SR}
 	\end{equation}
 		where $B_1^j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
-\end{itemize}
 It remains $p-p_2$ independent exogenous variables : $\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)$
@@
 
-This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian.
+This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian (see the appendices).
 These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $B_1^j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
@@
 Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
-\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
-
+%\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
+
+\subsection{Marginal regression model with decorrelated covariates}
+Let $Y \in \mathcal{R}^n$ be a response variable we want to explain with $X$.
+We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $X_2$ a regression model on $Y$ {\it depending only on uncorrelated covariates $X_1$}:
+		\begin{equation}
+			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y \textrm{ whith } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
+		\end{equation}
+		where $\beta=(\beta_1,\beta_2) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $I_n$ the identity matrix. 
+Knowing (\ref{SR}) we then obtain a marginal regression model:
+\begin{eqnarray}
+	Y_{|X_1}&=&X_1 (\beta_{1}+ \sum_{j \in I_2}B^{j}_{1}\beta_{j})+  \sum_{j \in I_2}\varepsilon_{j}\beta_{j}+\varepsilon_Y \label{Trueexpl} \\
+		&=&X_1 \alpha_1 + X_2\alpha_2 + \varepsilon_{\alpha} \textrm{ where} \alpha_2  = 0.
+\end{eqnarray}
+We note that it simply is a linear regression on some of the original covariates so we only made a pretreatment on the dataset by setting some coefficients to $0$ because of correlations.
+These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates.
+It isa variable pre-selection independent of the response $Y$.
+As a pretreatment, it allows usage of any method in a seconde time to estimate $\alpha$.
 \subsection{Estimator and properties}
-	We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $X_2$ a regression model on $Y$ {\it depending only on uncorrelated covariates $X_1$}:
-	\begin{equation}
-		Y=X_1 (A_{1}+ \sum_{j \in I_2}B^{j}_{1}A_{j})+  \sum_{j \in I_2}\varepsilon_{j}A_{j}+\varepsilon_Y 
-		 = X_1\alpha_1+ \varepsilon_{\alpha}\label{Trueexpl}
-	\end{equation}			
-		So we have the unbiased estimator: 
-		\begin{equation}
-			\hat{\alpha}_{1} = (X'_{1} X^{1})^{-1}X'_{1}Y % \textrm{ and }\hat{\alpha}_{2} = 0
+	The marginal regression model simply is
+	\begin{equation}
+		Y_{|X_1}= X_1\alpha_1+ \varepsilon_{\alpha}
+	\end{equation}			
+		So we have the unbiased estimator of $\alpha$: 
+		\begin{equation}
+			\hat{\alpha}_{1} = (X'_{1} X_{1})^{-1}X'_{1}Y  \textrm{ and }\hat{\alpha}_{2} = 0
+		\end{equation}
+		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $Y$ and $\alpha$
+		but in terms of $\beta$ this estimator is biased:
+		\begin{equation}
+			\operatorname{E}[\hat{\alpha_1}|X_1]=\beta_1+\sum_{j \in I_2}B^{j}_{1}\beta_{j}\textrm{ and }\operatorname{E}[\hat{\alpha_2}|X_1]=0
 		\end{equation}
@@
 		\begin{equation}
-			\operatorname{Var}[\hat{\alpha}_{1}|X]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}A_{j}^2 )(X'_{1} X^{1})^{-1} % \textrm{ and }\operatorname{Var}[\hat{\alpha}_{2}|X]= 0 
-		\end{equation}
-		We see that the variance is reduced (no correlations and smaller matrix give better conditioning) for small values of $\sigma_j$ $i.e.$ strong correlations.					
+			\operatorname{Var}[\hat{\alpha}_{1}|X_1]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 )(X'_{1} X_{1})^{-1}  \textrm{ and }\operatorname{Var}[\hat{\alpha}_{2}|X_1]= 0 
+		\end{equation}
+		We see that the variance is reduced compared to OLS described in equation (\ref{eqOLS})(no correlations and smaller matrix give better conditioning ) for small values of $\sigma_j$ $i.e.$ strong correlations.
 			%There is no theoretical guarantee that our model is better.
@@
 Expert knowledge can even be added to the structure.
-			 This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used.
+			 This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used then,like LASSO or even stepwise \cite{seber2012linear}.
 Hence we hope to obtain a parsimonious model with { \it two kinds of zeros}: those from decorrelating step and those from selection step, with different meanings.
 			 \\
-			 The \textsc{CorReg} package proposes to make selection with both LASSO (with LAR), elasticnet, adaptative lasso, and others more classical methods.
 			  A last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
@@
 
+	To better illustrate the bias-variance compromise, we look at a simple example with $p=3$ variables.
+$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=\frac{1}{2}x_1+\frac{1}{2}x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
+We also have $\beta=(\frac{1}{3},\frac{1}{3},\frac{1}{3})$ and $\sigma_Y=10$.
+Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
+Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
+	\begin{equation}
+		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+\frac{1}{2}}
+	\end{equation}
+
+\begin{figure}[h!]
+	\begin{minipage}[l]{.32\linewidth}
+			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
+			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
+	\end{minipage} \hfill
+	\begin{minipage}[c]{.32\linewidth}
+			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
+			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
+	\end{minipage} \hfill
+   \begin{minipage}[r]{.32\linewidth}
+			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
+			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
+   \end{minipage} 
+\end{figure} 
+It is clear in Figures \ref{MQE1} to \ref{MQE3} that the reduced model is more robust than OLS.
+And even when sub-regression are week (small value of $1-\mathcal{R}^2$) it does not explode.
+We also see that the error implied by strong correlations shrinks whit the rise of $n$.
+$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$.
+both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
+Thus, as $\operatorname{Var}(\hat{\alpha}_{1})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus OLS.
+It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
+
 	\subsection{Advantages for interpretation}
@@
 		Grouping effect is the fact that correlated covariates get similar coefficient and are selected together. It's the case with elasticnet \cite{zou2005regularization} for example.	
-			If $X_2=X_1+e$ and we have the grouping effect, we will obtain a model like $\hat{Y}=\hat{a}X_1+\hat{a}X_2$. 
+			If $X_2=X_1+e$ and we have the grouping effect, we will obtain a model like $\hat{Y}=\hat{\beta}X_1+\hat{\beta}X_2$. 
 			Then, if you try to modify the response value, you will modify one of the covariates and both will change so you won't get expected results.
 
-		At the opposite, methods like the LASSO with LAR will only keep one covariate for each group of correlated covariates  so we get $\hat{Y}=2aX_1$ and think that $X_2 \perp Y$ but $Y$ depends on $X_2$ so it is worse than grouping effect for interpretation.
-
-			Nothing constrains us to give only one equation. It is clearly better to give the user another equation (or system for more complex models) describing the correlations. 
-			So you get the following model : $Y=aX_1+aX_2$ and $X_1=X_2+e$. Then you have more information and are able to decide better actions. 
-			With such a model, grouping effect is no more useful because when saying $Y=2aX_1$ and $X_2=X_1+e$ with the information that $X_2$ has been removed locally because of its correlation with $X_1$, you don't get misleading interpretations anymore.
+		%At the opposite, methods like the LASSO with LAR will only keep one covariate for each group of correlated covariates  so we get $\hat{Y}=2aX_1$ and think that $X_2 \perp Y$ but $Y$ depends on $X_2$ so it is worse than grouping effect for interpretation.
+
+			Nothing constrains us to give only one equation. %It is clearly better to give the user another equation (or system for more complex models) describing the correlations. 
+			If you have the following model : $Y=\beta X_1+\beta X_2$ and $X_1=X_2+e$. Then you have more information and are able to decide better actions. 
+			With such a model, grouping effect is no more useful because when saying $Y=2\beta X_1$ and $X_2=X_1+e$ with the information that $X_2$ has been removed locally because of its correlation with $X_1$, you don't get misleading interpretations anymore.
 In fact, we have two kinds of zeros (those from selection and those from correlations) and we know for each zero which kind is it.
@@
 			Each equation here is very simple (only linear regressions) so you don't really increase complexity of the model. 
-			Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect).
+			Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
 
 	\section{Estimating structure of sub-regressions with a Markov chain}	
-
+\subsection{Detailed generative model}
+	To be able to compare structures with probabilistic criterions, we need a full generative model. We suppose that variables in $X_1$ follow Gaussian mixtures. 
+	\begin{equation}
+			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1}
+		\end{equation}
+		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed.
+Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.
+
+		Variables in $X_1$ are supposed to be independent.
+	Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways.
+%compute corresponding $\psi$ according to it.
+%and then improve the walk (will keep a structure only if it is really relevant).%and give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).
+	We now have a full generative model.
 	\subsection{Structure comparison} \label{compstruct}
-  All our work is based on $S$, the linear structure between the covariates.
-	Our generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood according to the complexity of the structure~\cite{BIChuard}.
+  To obtain the marginal regression model we need $S$, the linear structure between the covariates.
+	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood according to the complexity of the structure~\cite{BIChuard}.
 We will prefer this kind of comparison criterion instead of cross-validation that is very time-consuming and thus not friendly with combinatory problematics.
@@
 		P(S|X)&\propto& P(X|S)P(S) \\
-		BIC&=&-2\mathcal{L}(X,S,\Theta)+|\Theta|\log(n) \approx -2\log P(X|S)
+		-2\log P(X|S)&\approx & BIC=-2\mathcal{L}(X,S,\Theta)+|\Theta|\log(n)\\  
 	\end{eqnarray}
 	But $BIC$ tends to give too complex structures because we test a great range of models. 
-	Thus \mbox{\textsc{CorReg}} allows to choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
+	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
 %	 Thus we have :
@@
 	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
-	\textsc{CorReg} adds this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ in the MCMC when hierarchical hypothesis is made (it will be the case in the followings).			
+	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
 	But we can imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context.
 \\	
-
 		We will now use the following notation: $\psi(S)=\psi(X|S)$	where $\psi(X|S)$ is the chosen criterion, that will be $BIC$ with hierarchical uniform hypothesis in our numerical results.
-\subsection{Detailed generative model}
-
-	To better fit industrial variables (Figure \ref{graphMixmod}), we suppose  that variables in $X_1$ follow Gaussian mixtures.
-The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust.
-But one can use other laws if needed.
-Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.
-\begin{equation}
-			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1}
-		\end{equation}
-		Variables in $X_1$ are supposed to be independent.
-	Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to compute corresponding $\psi$ separately, give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).
+We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.
 
 	\subsection{The neighbourhood}
-		\subsubsection{Definition}
 	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
@@
 	At every moment, coherence between $I_1$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_1^j=|I_1^j|$, $I_2=\{j |p_1^j>0 \}$, $p_2= |I_2|$, .
-		\subsubsection{Alternatives}
-	We have here at each step $|\mathcal{V}_{S,j}|=p$ candidates but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the internal regressions or the whole structure, a maximum number of internal regressions, {\it etc.}).
-\textsc{CorReg} allows to modify this neighbourhood to better fit users constraints.
-Relaxation (column-wise and row-wise) is optional but gives more stability to the number of feasible candidates at each step and allows to modify several parts of $I_1$ in only one step when needed.
-Hence it improves efficiency by a significant reinforcement of the irreductibility of the Markov chain.
-Rejecting candidates instead of doing the relaxation steps will  however reduce the number of evaluated candidates and thus accelerate the walk.
-So it can be used for a warming phase when $n$ is great and time is missing.
-
-	The hierchical uniform hypothesis made above for $P(S)$ implies $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ so candidates may be rejected to satisfy this hypothesis.
-Strongest constraints on $p_2$ and/or $p_1$ can be given in \textsc{CorReg} if relevant.
-
-If the algorithm did not have time to converge, it can be continued with a few step for which the neighboorhood would only contain smaller candidates (in terms of complexity).
-It is equivalent to ask for each element in $I_1$ if the criterion $\psi$ would be better without it.
-Thus it can be seen as a final cleaning step.
-But in fact, it's just continuing the MCMC with a reduced neighbourhood.
+
 	\subsection{The walk}
@@
 We do so in the followings.
-Then this structure can be reduced by the hadamard product with the binary matrix obtained by Graphical Lasso\cite{friedman2008sparse}.
-Graphical LASSO is time consuming so we only used it on real datasets (each simulation was computed 100 times).
+Then this structure could be for example reduced by the hadamard product with the binary matrix obtained by Graphical Lasso\cite{friedman2008sparse} that makes selection in the precision matrix but it is time consuming.
 
@@
 	The \textsc{CorReg} package has been tested on simulated datasets. 
-Section \ref{compZ} show the results obtained in terms of $\hat{S}$.
-Sections \ref{tableMSEsimtout} et \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods.
+Section \ref{compZ} shows the results obtained in terms of $\hat{S}$.
+Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods.
 Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals.
-For each simulation ,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
-The $B_1^j$ are generated according to the same Poisson law but with a random sign.
+For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
+The $B_1^j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign.
 $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.
@@
 		\subsection{Finding the structure}
-		\subsubsection{How to evaluate found structure ?}
+		\subsubsection{How to evaluate found structure?}
 			The first criterion is $\psi$ which is minimised in the MCMC.
@@
 The datasets used for (U) and (HU) are the same to keep the comparison  meaningful.
-
+These datasets and $\hat{S}$ are those used for tables
+ \ref{YXlinOLS} to \ref{YX2linstep}.
+
+ It is also notable that (HU) has a greater computational cost than (U). But next versions of the package may optimize a bit more the penalization step and reduce this gap.
+	We also notice that the MCMC is faster when there are numerous correlations (rejecting more candidates). 
 \begin{table}[h!]
@@
 \clearpage
-
-	\subsection{$Y$ depends on all variables in $X$}	 \label{tableMSEsimtout}	
+\subsection{Results on prediction}
+	\subsubsection{$Y$ depends on all variables in $X$}	 \label{tableMSEsimtout}	
 We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure).
@@
 When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods.
-We also observe that simple methods like stepwise (here from the package \textsc{lars}) give good results in prediction.
-
-It is also notable that (HU) has a greater computational cost than (U). But next versions of the package may optimize a bit more the penalization step and reduce this gap.
+We also observe in table \ref{YXlinstep} that simple methods like stepwise (here from the package \textsc{lars}) can give good results in prediction.
+
 \begin{table}[h!]
@@
 30 & 16 & U&MSE (sd) & 510747.53 (2287539.8) & 635.42 (335.2) & 610.32 (424.9) \\
-& & & cpl (sd) & 30 (0) & 24.11 (1) & 25.06 (0.2) \\
+& & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
  &  &HU &MSE (sd) & 510747.53 (2287539.8) & 603.24 (415.6) & 610.32 (424.9) \\
-& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25.06 (0.2) \\
+& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25 (0) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
 30 & 32 & U & MSE (sd) & 178323.95 (1426610.2) & 180.02 (44.9) & 141.03 (27.8) \\
-& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9.27 (0.5) \\
+& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
  &  & HU & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
-& & & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
+& & & cpl (sd) & 30 (0) & 21 (0) & 9 (0) \\
 \hline
@@
 50 & 16 & U&MSE (sd) & 612.72 (291.5) & 239.17 (89.4) & 200.32 (42.6) \\	
-& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25.13 (0.4) \\
+& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
  &  &HU &MSE (sd) & 612.72 (291.5) & 207.6 (51.3) & 200.32 (42.6) \\
-& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25.13 (0.4) \\
+& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25 (0) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
 50 & 32 & U&MSE (sd) & 555.44 (262.5) & 128.98 (18.1) & 121.08 (11.9) \\
-& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9.13 (0.3) \\
+& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
  &  &HU &MSE (sd) & 555.44 (262.5) & 171.9 (31.1) & 121.08 (11.9) \\
-& & &cpl (sd) & 41 (0) & 21 (0) & 9.13 (0.3) \\
+& & &cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
 \hline
@@
 100 & 16 & U&  MSE (sd) & 168.68 (22.4) & 158.51 (51.6) & 133.49 (12.2) \\
-& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25.08 (0.3) \\
+& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
  &  &HU &  MSE (sd) & 168.68 (22.4) & 137.03 (20.3) & 133.49 (12.2) \\
-& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25.08 (0.3) \\
+& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25 (0) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  MSE (sd) & 173.25 (22.9) & 112.8 (10.7) & 110.63 (6.9) \\\\
-& & & cpl (sd) & 41 (0) & 10.02 (0.8) & 9.25 (0.5) \\
+100 & 32 & U&  MSE (sd) & 173.25 (22.9) & 112.8 (10.7) & 110.63 (6.9) \\
+& & & cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
  &  &HU & MSE (sd) & 173.25 (22.9) & 127.4 (11.9) & 110.63 (6.9) \\
-& & & cpl (sd) & 41 (0) & 21 (0) & 9.25 (0.5) \\
+& & & cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
 \hline
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinLASSO}
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$.
+\textsc{CorReg} logically wins.
+LASSO is better than OLS but is improved by \textsc{CorReg} even for large values of $n$.}\label{YXlinLASSO}
 \end{table}
@@
 \end{tabular} 
-\caption{Elasticnet combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinenet}
+\caption{Elasticnet combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins. LASSO was better.}\label{YXlinenet}
 \end{table}
@@
 \end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X$. \textsc{CorReg} logically wins}\label{YXlinstep}
+\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X$. \textsc{CorReg} logically wins but stepwise is quite good compared to elasticnet.}\label{YXlinstep}
 \end{table}
@@
 \clearpage
-	\subsection{$Y$ depends only on covariates in $X_2$ (worst case for us)}	 \label{tableMSEsimgauche}
+	\subsubsection{$Y$ depends only on covariates in $X_2$ (worst case for us)}	 \label{tableMSEsimgauche}
 We now try the method with a response depending only on variables in $X_2$. The datasets used here were still those from \ref{compZvrai}.
@@
 30 & 16 & U&MSE (sd) & 81781.54 (247281.7) & 507.1 (269.2) & 593.03 (428.3) \\
-& & & cpl (sd) & 30 (0) & 24.11 (1) & 25.06 (0.2) \\
+& & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
  &  &HU &MSE (sd) & 81781.54 (247281.7) & 587.07 (425.6) & 593.03 (428.3) \\
-& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25.06 (0.2) \\
+& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25 (0) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
 30 & 32 & U & MSE (sd) & 74136.81 (297716.1) & 189.45 (48.1) & 144.51 (31.4) \\
-& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9.27 (0.5) \\
+& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
  &  & HU & MSE (sd) & 74136.81 (297716.1) & 340.53 (169.4) & 144.51 (31.4) \\
-& & & cpl (sd) & 30 (0) & 21 (0) & 9.27 (0.5) \\
+& & & cpl (sd) & 30 (0) & 21 (0) & 9 (0) \\
 \hline
@@
 50 & 16 & U&	MSE (sd) & 684.04 (438.1) & 190.97 (37.7) & 197.32 (40) \\
-& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25.13 (0.4) \\
+& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
  &  &HU &MSE (sd) & 684.04 (438.1) & 196.36 (40.5) & 197.32 (40) \\
-& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25.13 (0.4) \\
+& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25 (0) \\
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
 50 & 32 & U&MSE (sd) & 596.32 (323) & 126.94 (15.5) & 119.39 (12.4) \\
-& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9.13 (0.3) \\
+& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
  &  &HU &MSE (sd) & 596.32 (323) & 168.03 (27.7) & 119.39 (12.4) \\
-& & &cpl (sd) & 41 (0) & 21 (0) & 9.13 (0.3) \\
+& & &cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
 \hline
@@
 100 & 16 & U&  MSE (sd) & 168.35 (20.1) & 133.75 (12.4) & 135.08 (13.1) \\
-& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25.08 (0.3) \\
+& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
  &  &HU &  MSE (sd) & 168.35 (20.1) & 135.04 (13.1) & 135.08 (13.1) \\
-& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25.08 (0.3) \\
+& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25 (0) \\
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
 100 & 32 & U&  MSE (sd) & 168.07 (21.4) & 109.61 (7.4) & 108.64 (7.2) \\
-& & &  cpl (sd) & 41 (0) & 10.02 (0.8) & 9.25 (0.5) \\
+& & &  cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
  &  &HU & MSE (sd) & 168.07 (21.4) & 124.23 (11.1) & 108.64 (7.2) \\
-& & & cpl (sd) & 41 (0) & 21 (0) & 9.25 (0.5) \\
-\hline
-\end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$. } \label{YX2linOLS}
+& & & cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
+\hline
+\end{tabular} 
+\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$.
+Sometimes $\hat{S}$ gives better results than $S$ because $S$ is penalized by the fact that it relies only on covariates not in the true model.
+} \label{YX2linOLS}
 \end{table}
@@
 \clearpage
-\subsection{Non linear case}
- we have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
+\subsection{Robustess of the model}
+ We have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
 Variables in $X_2$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $X_1$ .
@@
 \end{tabular} 
-\caption{ $Y$ depends on all variables in $X$. \textsc{CorReg} is not too far from LASSO.} \label{resYnonlin}
+\caption{ $Y$ depends on all variables in $X$. \textsc{CorReg} is not too far from LASSO even if it stays behind. } \label{resYnonlin}
 \end{table}
@@
 	\begin{minipage}[l]{.30\linewidth}
-			\includegraphics[width=150px]{figures/correl0BVBI.png} 
-			\caption{Correlations between the covariates}
+			\includegraphics[width=150px]{figures/correlbeforeafter.png} 
+			\caption{Correlations between the covariates in $X$ (upper) and $\hat{X}_1$ (lower).}
 	\end{minipage} \hfill
@@
 \end{figure}   			
-	Here $n<p$ so we only have the leave one out cross-validation MSE.
-	\textsc{CorReg} improves LASSO by $5.24\%$ and elasticnet by $8.60\%$. \textsc{CorReg} combined with LASSO gives the best result but it is only a leave on out MSE.
+	Here $n<p$ so we only compare the leave-one-out cross-validation MSE.
+	\textsc{CorReg} improves LASSO by $5.24\%$ and elasticnet by $8.60\%$. \textsc{CorReg} combined with LASSO gives the best result but it is only a leave-on-out MSE.
 \begin{table}[h!]
@@
 \bibliographystyle{plain}
+\section{Appendices}
+	\subsection{Identifiability of the structure}
+	The model presented above relies on a discrete structure $S$ between the covariates.
+But to find it we need identifiability property to insure the MCMC will asymptotically find the true model.
+Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions ?
+The answer is no.
+Thus $S$ is identifiable.
+
+\subsubsection{Alternative neighbourhoods for the MCMC}
+	We have here at each step $|\mathcal{V}_{S,j}|=p$ candidates but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the internal regressions or the whole structure, a maximum number of internal regressions, {\it etc.}).
+\textsc{CorReg} allows to modify this neighbourhood to better fit users constraints.
+Relaxation (column-wise and row-wise) is optional but gives more stability to the number of feasible candidates at each step and allows to modify several parts of $I_1$ in only one step when needed.
+Hence it improves efficiency by a significant reinforcement of the irreductibility of the Markov chain.
+Rejecting candidates instead of doing the relaxation steps will  however reduce the number of evaluated candidates and thus accelerate the walk.
+So it can be used for a warming phase when $n$ is great and time is missing.
+
+	The hierchical uniform hypothesis made above for $P(S)$ implies $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ so candidates may be rejected to satisfy this hypothesis.
+Stronger constraints on $p_2$ and/or $p_1$ can be given in \textsc{CorReg} if relevant.
+
+If the algorithm did not have time to converge (stationnarity), it can be continued with a few step for which the neighboorhood would only contain smaller candidates (in terms of complexity).
+It is equivalent to ask for each element in $I_1$ if the criterion $\psi$ would be better without it.
+Thus it can be seen as a final cleaning step.
+But in fact, it's just continuing the MCMC with a reduced neighbourhood.
 

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@472 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
-\title{CorReg: model based pretreatment for regression with correlated variables. Application in steel industry}
+\title{CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry}
 
@@
 \bigskip
-{\bf Abstract.} Linear regression generally suppose to have decorrelated covariates.
-This hypothesis is often irrealist with industrial datasets that contains many highly correlated covariates due to the process, physcial laws,  {\it etc}.
-The proposed generative model consists in explicit modeling of the correlations with a family of linear regressions between the covariates permitting to obtain by marginalization a parsimonious correlation-free regression model, easily understandable and compatible with variable selection methods.
+{\bf Abstract.} Linear regression outcomes are known to be damaged by correlated covariates.
+However many modern datasets are expected to convey more and more highly correlated covariates.
+We propose to explicitly model the correlations by a family of linear regressions between the covariates.
+Marginalization allows then to obtain a parsimonious correlation-free regression model, easily understandable and from which it is then possible to perform standard linear estimation methods including variables selection procedures for instance.
 The structure of correlations is found with an MCMC algorithm.
-An R package (\textsc{CorReg}) available on the CRAN implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry.
+An R package (\textsc{CorReg}) available on the CRAN implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry where correlated variables are frequent.
 \smallskip

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@473 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@474 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  	This paper will first present the reduced model and its properties before describing in Section 3 the algorithm used to find the structure.
- 	We will then look at some numerical results on simulated (Section 4) and real industrial datasets (Section 5) before concluding and giving some perspectives in the sixth part.
+ 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in the last part.
 
@@
 	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
+	Another problem is that matricial inversion requires $n\geq p$.
+When it is not the case, a frequently used alternative is the Moore-Penrose generalized inverse \cite{katsikis2008fast}.
+In the followings, matrices inverses when $n<p$ will refer to this generalized inverse.
+Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
 \subsection{Modelisation of the correlations}
@@
 
-\section{Numerical results on simulated datasets}
+\section{Numerical results on simulated datasets} \label{sectionsimul}
 	\subsection{The datasets}	
@@
 	\clearpage	
-\section{Numerical results on real datasets}
+\section{Numerical results on real datasets} \label{sectionrealcase}
 \subsection{Quality case study} \label{sectionexfos}

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@475 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
- Where some try to reduce the dimension and then just hope to have small correlations in the remaining dimensions,  we propose to focus on the correlations, giving a model with orthogonal covariates and an explicit structure between covariates.
+ Where some try to reduce the dimension and then have to suppose having no correlations between the remaining covariates,  we propose to focus on the correlations, giving a model with orthogonal covariates and an explicit structure between covariates.
 In fact we search the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@476 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 And we consider correlations between the explicative covariates of the main regression, not between the residuals.
-	  The structure is supposed to be the source of the correlations and allows us to define a reduced set of independent covariates.
-Thus we reduce dimension and correlations in the same time.
-The structure justifies the eviction of the redundant covariates without significant information loss.
-It can be seen as a pretreatment on the dataset based on the hypothesis of a strong structure between the covariates (i.e.
-: small $\varepsilon_X$).
-
-	We can use any variable selection method on the reduced dataset with improved efficiency (reduced variance) due to dimension reduction and correlation suppression.
-So we obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
+	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
+It can be seen as a pretreatment on the dataset because other regression methods can the be used on this reduced and correlation-free set covariates.
+So we may obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
 This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.
@@
 
-But to work, we need an explicit structure between the covariates.
-SEM are often used in social sciences and economy where a structure is supposed "by hand" but here we want to be able to find a structure without any a priori (possibility to include some known structure remains).
+SEM are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
 Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix).
@@
 However, Graphical LASSO can be used in the initialization step of our MCMC.
-This structure is based on a generative model to be able to compare likelihood-based criterion in the MCMC.
+The structure is based on a full generative model to be able to compare likelihood-based criterion in the MCMC.
  	\\
@@
  	%plan
- 	This paper will first present the reduced model and its properties before describing in Section 3 the algorithm used to find the structure.
+ 	This paper will first present the marginal regression model and its properties before describing in Section 3 the random walk used to find the structure.
  	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in the last part.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@477 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
-The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (OLS) by \cite{saporta2006probabilites}
+The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS} \cite{saporta2006probabilites}) by
 	\begin{equation}
@@
 %bibliographie	
-	Because OLS is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
+	Because \textsc{OLS} is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
 Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.
@@
 
-	The Least Absolute Shrinkage and Selection Operator (LASSO)\cite{tibshirani1996regression} consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
+	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
 		\begin{equation}
@@
 		\end{equation}	
-	 The Least Angle Regression\cite{efron2004least} (LAR) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
+	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
 It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates.
@@
 	 But LASSO also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates.
-Another limitation of the LASSO is that it preserves at most $n$ predictors (troublesome when in high dimension).
-Some recent variants of the LASSO does exist for the choice of the penalization coefficient like the adaptative LASSO \cite{zou2006adaptive} or the random LASSO \cite{wang2011random}.
+Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension).
+Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random LASSO \cite{wang2011random}.
 	 \\
 
-	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the LASSO. 
+	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}. 
 %	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
@@
 
-	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (OSCAR) \cite{bondell2008simultaneous}.
-	The CLusterwise Effect REgression\cite{yengo2012variable} (CLERE) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith grouped $\beta_j$ following a Gaussian Mixture distribution.
+	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}).
+	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith grouped $\beta_j$ following a Gaussian Mixture distribution.
 
@@
 But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net.
-Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute OLS but has no specific protection against correlations issues.
+Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
 	~\\	~\\
@@
 
-	Our work is based on the assumption that if we know that correlations are a problem and if we precisely know the correlations, we could use this knowledge to avoid the problem.
-	The idea is to suppose explicitly a linear structure between the covariates. 
-	It gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (SEM)\cite{davidson1993estimation}.
-Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive SEM).
-  	Other methods already rely on linear structure and start to take into account correlations but they only consider covariances between the residuals (SUR) \cite{SURzellner} or covariances between the endogenous variables like SPRING (Structured selection of Primordial Relationships IN the General linear model) \cite{chiquetconf}.
-
-	  In this work, we decide to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
-Thus we don't have a system of regressions but one regression on our response variable and a system of subregressions (without the response variable).
-And we consider correlations between the explicative covariates of the main regression, not between the residuals.
+	Our work is based on the assumption that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
+	The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
+Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
+  	Other methods take into account correlations but they only consider covariances between the residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or the endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).
+
+	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
+Thus we don't have a system of regressions but one regression on our response variable and a system of sub-regressions (without the response variable).
+Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
 	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@478 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
-The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS} \cite{saporta2006probabilites}) by
-	\begin{equation}
-		\hat{\beta}=(X'X)^{-1}X'Y.
-\end{equation}
-
-When estimating the parameters of the regression we have to compute the inverse of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
+The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}).
+As shown in \ref{sectionOLS}, estimation of the parameters of the regression we have to compute the inverse of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other.
 Variance of $\hat{\beta}$ increases based on two aspects :
@@
 Moreover, when you use more covariates, you increase the chance to have correlated ones.
-It will be illustrated with examples from steel industry in the last section of this paper with a big set of covariates (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
+It will be illustrated with examples from steel industry in the last section of this paper (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
 	 In such a context, variance of the estimators can lead to arbitrary results or even no results at all.
@@
 	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
-It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates.
+It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates.
 And it  really selects covariates with coefficients set exactly to 0.
-	 But LASSO also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates.
+	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates.
 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension).
-Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random LASSO \cite{wang2011random}.
+Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
 	 \\
@@
 Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
-  	Other methods take into account correlations but they only consider covariances between the residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or the endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).
+  	Other methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).
 
@@
 	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
-It can be seen as a pretreatment on the dataset because other regression methods can the be used on this reduced and correlation-free set covariates.
+It can be seen as a pretreatment on the dataset because other regression methods can then be used.
 So we may obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
@@
 
-SEM are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
+\textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
 Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix).
 It consists in a selection in the precision matrix, setting some covariances to zero.
-But the resulting matrix is symmetric and we need an oriented graph for our SEM.
-So we developped an MCMC algorithm to find it (R package CorReg on CRAN).
-However, Graphical LASSO can be used in the initialization step of our MCMC.
-The structure is based on a full generative model to be able to compare likelihood-based criterion in the MCMC.
+But the resulting matrix is symmetric and we need an oriented structure for our \textsc{SEM}.
+So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN).
+The structure is based on a full generative model on $X$ to be able to compare likelihood-based criterion in the \textsc{MCMC}.
  	\\
@@
 \section{Model to select decorrelated covariates}
-\subsection{The decorrelation problem in regression}
+\subsection{The decorrelation problem in regression}\label{sectionOLS}
 For a model defined by 
@@
 	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
-	we have the following Ordinary Least Squares (OLS) estimators:
+	we have the following Ordinary Least Squares (\textsc{OLS}) estimators \cite{saporta2006probabilites}:
 	\begin{equation}
@@
 	\end{equation}
-	 The Mean Squared Error (MSE) on $\hat{\beta}$ is:
+	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
 	\begin{eqnarray}
-		MSE(\hat{\beta}|\beta,X)&=&Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
+		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
 			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@479 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}).
-As shown in \ref{sectionOLS}, estimation of the parameters of the regression we have to compute the inverse of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other.
+As shown in section \ref{sectionOLS}, estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other.
 Variance of $\hat{\beta}$ increases based on two aspects :
@@
 	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
-It can be seen as a pretreatment on the dataset because other regression methods can then be used.
-So we may obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
+So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
 This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@480 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@481 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
-Thus we don't have a system of regressions but one regression on our response variable and a system of sub-regressions (without the response variable).
+Thus we have one regression on our response variable and a system of sub-regressions (without the response variable).
 Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
@@
 	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
-	we have the following Ordinary Least Squares (\textsc{OLS}) estimators \cite{saporta2006probabilites}:
+	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
 	\begin{equation}
@@
 	\end{equation}
-	With variance 
+	With variance matrix
 	\begin{equation}
@@
 
-We define a family of $p_2$ regressions within correlated covariates in $X$:
-	\begin{equation}
-		\forall j \in I_2: X^j_{|X_1,S}=X_1B^j_1+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n) \label{SR}
-	\end{equation}
-		where $B_1^j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
-It remains $p-p_2$ independent exogenous variables : $\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)$
+The family of $p_2$ regressions within correlated covariates in $X$ is noted:
+	\begin{equation}
+		\forall j \in I_2: X^j_{|X_1,S}=X_1\delta_j+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n) \label{SR}
+	\end{equation}
+		where $\delta_j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
+It remains $p-p_2$ independent exogenous variables $X_1$: $\forall j \notin I_2 : X^j \sim f(\theta_j)$
 
@@
 This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian (see the appendices).
-These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $B_1^j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
+These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $\delta_j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
 \\
-If there are exact regressions ($\sigma^2_j=0$) in (\ref{SR}), the structure won't be identifiable.
-But it only means that several structure will have the same likelihood and they will have the same interpretation.
-So it's not really a problem.
-Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
-%\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
 
@@
 But to find it we need identifiability property to insure the MCMC will asymptotically find the true model.
-Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions ?
+Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions?
 The answer is no.
 Thus $S$ is identifiable.
+
+	If there are exact regressions ($\sigma^2_j=0$) in (\ref{SR}), the structure won't be identifiable.
+But it only means that several structure will have the same likelihood and they will have the same interpretation.
+So it's not really a problem.
+Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
+%\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
+	In the followings we suppose $\sigma^2_j\neq 0$.
 

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@482 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian (see the appendices).
-These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $\delta_j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
+These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate the $\delta_j$ and $\sigma_j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
 \\
@@
 \subsection{Marginal regression model with decorrelated covariates}
-Let $Y \in \mathcal{R}^n$ be a response variable we want to explain with $X$.
+Let $Y \in \mathcal{R}^n$ be a response variable we want to explain with $X$:
+	\begin{equation}
+			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y \textrm{ whith } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
+		\end{equation}
+		where $\beta=(\beta_1,\beta_2) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $I_n$ the identity matrix. 
 We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $X_2$ a regression model on $Y$ {\it depending only on uncorrelated covariates $X_1$}:
 		\begin{equation}
-			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y \textrm{ whith } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
-		\end{equation}
-		where $\beta=(\beta_1,\beta_2) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $I_n$ the identity matrix. 
+			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y
+		\end{equation}
+
 Knowing (\ref{SR}) we then obtain a marginal regression model:
@@
 	Y_{|X_1}&=&X_1 (\beta_{1}+ \sum_{j \in I_2}B^{j}_{1}\beta_{j})+  \sum_{j \in I_2}\varepsilon_{j}\beta_{j}+\varepsilon_Y \label{Trueexpl} \\
-		&=&X_1 \alpha_1 + X_2\alpha_2 + \varepsilon_{\alpha} \textrm{ where} \alpha_2  = 0.
+		&=&X_1 \alpha_1 + X_2\alpha_2 + \varepsilon_{\alpha} \textrm{ where } \alpha_2  = 0.
 \end{eqnarray}
@@
 For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
-The $B_1^j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign.
+The $\delta_j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign.
 $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@483 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $X_2$ a regression model on $Y$ {\it depending only on uncorrelated covariates $X_1$}:
-		\begin{equation}
-			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y
-		\end{equation}
-
-Knowing (\ref{SR}) we then obtain a marginal regression model:
 \begin{eqnarray}
-	Y_{|X_1}&=&X_1 (\beta_{1}+ \sum_{j \in I_2}B^{j}_{1}\beta_{j})+  \sum_{j \in I_2}\varepsilon_{j}\beta_{j}+\varepsilon_Y \label{Trueexpl} \\
+	Y_{|X_1}&=&X_1 (\beta_{1}+ \sum_{j \in I_2}\beta_{j}\delta_j)+  \sum_{j \in I_2}\beta_{j}\varepsilon_{j}+\varepsilon_Y \label{Trueexpl} \\
 		&=&X_1 \alpha_1 + X_2\alpha_2 + \varepsilon_{\alpha} \textrm{ where } \alpha_2  = 0.
@@
 These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates.
-It isa variable pre-selection independent of the response $Y$.
+It is a variable pre-selection independent of the response $Y$.
 As a pretreatment, it allows usage of any method in a seconde time to estimate $\alpha$.
@@
 	\end{equation}			
-		So we have the unbiased estimator of $\alpha$: 
+		So we have the \textsc{OLS} unbiased estimator of $\alpha$: 
 		\begin{equation}
@@
 		\begin{equation}
-			\operatorname{E}[\hat{\alpha_1}|X_1]=\beta_1+\sum_{j \in I_2}B^{j}_{1}\beta_{j}\textrm{ and }\operatorname{E}[\hat{\alpha_2}|X_1]=0
+			\operatorname{E}[\hat{\alpha_1}|X_1]=\beta_1+\sum_{j \in I_2}\beta_{j}\delta_j \textrm{ and }\operatorname{E}[\hat{\alpha_2}|X_1]=0
 		\end{equation}

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@484 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
+Grouping effect is the fact that correlated covariates get similar coefficient and are selected together.
 	\\
@@
 It is a variable pre-selection independent of the response $Y$.
-As a pretreatment, it allows usage of any method in a seconde time to estimate $\alpha$.
+As a pretreatment, it allows usage of any method in a second time to estimate $\alpha$like LASSO or even stepwise \cite{seber2012linear}.
+
+\subsection{Advantages for interpretation}\label{interpretation}
+The explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset.
+Expert knowledge can even be added to the structure.
+
+After selection and estimation we will obtain a model with { \it two kinds of zeros}: those from decorrelating step and those from selection step, with different meanings. 
+Thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.
+
+Each equation here is very simple (only linear regressions). 
+Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
+
 \subsection{Estimator and properties}
@@
 Therefore we made some simulations to compare both methods (see the end of this paper).
-			Moreover, the explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset.
-Expert knowledge can even be added to the structure.
-			 This new model is reduced even without variable selection and is just a linear regression so every method for variable selection in linear regression can be used then,like LASSO or even stepwise \cite{seber2012linear}.
-Hence we hope to obtain a parsimonious model with { \it two kinds of zeros}: those from decorrelating step and those from selection step, with different meanings.
+			Moreover, 
 			 \\
@@
 It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
-
-	\subsection{Advantages for interpretation}
-
-		Grouping effect is the fact that correlated covariates get similar coefficient and are selected together. It's the case with elasticnet \cite{zou2005regularization} for example.	
-			If $X_2=X_1+e$ and we have the grouping effect, we will obtain a model like $\hat{Y}=\hat{\beta}X_1+\hat{\beta}X_2$. 
-			Then, if you try to modify the response value, you will modify one of the covariates and both will change so you won't get expected results.
-
-		%At the opposite, methods like the LASSO with LAR will only keep one covariate for each group of correlated covariates  so we get $\hat{Y}=2aX_1$ and think that $X_2 \perp Y$ but $Y$ depends on $X_2$ so it is worse than grouping effect for interpretation.
-
-			Nothing constrains us to give only one equation. %It is clearly better to give the user another equation (or system for more complex models) describing the correlations. 
-			If you have the following model : $Y=\beta X_1+\beta X_2$ and $X_1=X_2+e$. Then you have more information and are able to decide better actions. 
-			With such a model, grouping effect is no more useful because when saying $Y=2\beta X_1$ and $X_2=X_1+e$ with the information that $X_2$ has been removed locally because of its correlation with $X_1$, you don't get misleading interpretations anymore.
-In fact, we have two kinds of zeros (those from selection and those from correlations) and we know for each zero which kind is it.
-			So it is possible to combine the advantages of grouping effect and selection just giving several equations. 
-			Each equation here is very simple (only linear regressions) so you don't really increase complexity of the model. 
-			Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
 

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@485 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Therefore we made some simulations to compare both methods (see the end of this paper).
-			Moreover, 
 			 \\
-			  A last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
-This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}).
 
@@
 
+When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
+This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}).
 \begin{table}[h!]

### 2014-04-01 www/article13.tex
uppdate MQE formulas
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@486 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	To better illustrate the bias-variance compromise, we look at a simple example with $p=3$ variables.
-$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=\frac{1}{2}x_1+\frac{1}{2}x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
-We also have $\beta=(\frac{1}{3},\frac{1}{3},\frac{1}{3})$ and $\sigma_Y=10$.
+$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
+We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
 Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
@@
 	\begin{equation}
-		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+\frac{1}{2}}
+		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
 	\end{equation}
@@
 
-\subsubsection{Alternative neighbourhoods for the MCMC}
+\subsection{Alternative neighbourhoods for the MCMC}
 	We have here at each step $|\mathcal{V}_{S,j}|=p$ candidates but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the internal regressions or the whole structure, a maximum number of internal regressions, {\it etc.}).

### 2014-04-01 www/article13.tex
up MQE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@487 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		We see that the variance is reduced compared to OLS described in equation (\ref{eqOLS})(no correlations and smaller matrix give better conditioning ) for small values of $\sigma_j$ $i.e.$ strong correlations.
+So we play on the bias-variance tradeoff, reducing the variance by adding a bias.
 			%There is no theoretical guarantee that our model is better.
@@
 			 \\
-
-	To better illustrate the bias-variance compromise, we look at a simple example with $p=3$ variables.
-$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
-We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
-Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
-Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
-	\begin{equation}
-		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
-	\end{equation}
-
-\begin{figure}[h!]
-	\begin{minipage}[l]{.32\linewidth}
-			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
-			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
-	\end{minipage} \hfill
-	\begin{minipage}[c]{.32\linewidth}
-			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
-			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
-	\end{minipage} \hfill
-   \begin{minipage}[r]{.32\linewidth}
-			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
-			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
-   \end{minipage} 
-\end{figure} 
-It is clear in Figures \ref{MQE1} to \ref{MQE3} that the reduced model is more robust than OLS.
-And even when sub-regression are week (small value of $1-\mathcal{R}^2$) it does not explode.
-We also see that the error implied by strong correlations shrinks whit the rise of $n$.
-$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$.
-both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
-Thus, as $\operatorname{Var}(\hat{\alpha}_{1})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus OLS.
-It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
 
@@
 \section{Numerical results on simulated datasets} \label{sectionsimul}
+
+	\subsection{Theoretical Mean squared errors on $\hat \beta$} 
+	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables.
+$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
+We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
+Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
+Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
+	\begin{equation}
+		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
+	\end{equation}
+
+\begin{figure}[h!]
+	\begin{minipage}[l]{.32\linewidth}
+			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
+			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
+	\end{minipage} \hfill
+	\begin{minipage}[c]{.32\linewidth}
+			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
+			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
+	\end{minipage} \hfill
+   \begin{minipage}[r]{.32\linewidth}
+			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
+			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
+   \end{minipage} 
+\end{figure} 
+It is clear in Figures \ref{MQE1} to \ref{MQE3} that the reduced model is more robust than OLS.
+And even when sub-regression are week (small value of $1-\mathcal{R}^2$) it does not explode.
+We also see that the error implied by strong correlations shrinks whit the rise of $n$.
+$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$.
+both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
+Thus, as $\operatorname{Var}(\hat{\alpha}_{2})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
+It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
+
 	\subsection{The datasets}	

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@488 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{figure} 
-It is clear in Figures \ref{MQE1} to \ref{MQE3} that the reduced model is more robust than OLS.
-And even when sub-regression are week (small value of $1-\mathcal{R}^2$) it does not explode.
-We also see that the error implied by strong correlations shrinks whit the rise of $n$.
-$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$.
-both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
+It is clear in Figures \ref{MQE1} to \ref{MQE3} that the marginal model is more robust than \textsc{OLS} on $X$.
+And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it does not explode until extreme values (sub-regression nearly fully explained by the noise).
+We also see that the error implied by strong correlations shrinks whith the rise of $n$.
+$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$ where both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
 Thus, as $\operatorname{Var}(\hat{\alpha}_{2})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@489 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
-\title{CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry}
+\title{Model-based pretreatment for regression with correlated variables.}
 
@@
 Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
-We note the linear regression model:
-\begin{equation}
-		Y_{|X}=X\beta + \varepsilon
-	\end{equation}
-	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
-The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}).
-As shown in section \ref{sectionOLS}, estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other.
-Variance of $\hat{\beta}$ increases based on two aspects :
+In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
+\begin{equation}
+		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
+	\end{equation}
+	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
+The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}).
+%As shown in section \ref{sectionOLS},
+	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
+Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
 \begin{itemize}
-	\item The dimension $p$ (number of covariates) of the model  : the more covariates you have the greater variance you get.
-	\item The correlations within the covariates : strongly correlated covariates give bad-conditioning and increase variance of the estimators .
+	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
+	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
 \end{itemize}
@@
 
-In the following we note classical norms: $\parallel\beta\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\beta \parallel_1=\sum_{i=1}^p|\beta_i| $.
+In the following we note classical norms: $\parallel\boldsymbol{\beta}\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\boldsymbol{\beta} \parallel_1=\sum_{i=1}^p|\beta_i| $.
 
@@
 	\begin{equation}
-		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_2^2\leq k
+		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel \boldsymbol{\beta} \parallel_2^2\leq k
 	\end{equation}
@@
 		\begin{equation}
-		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_1\leq \lambda
+		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda
 		\end{equation}	
@@
 	\begin{equation}
-		\hat{\beta}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel Y-X\beta \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\beta\parallel_1+\alpha\parallel\beta\parallel_2^2\leq t \textrm{ for some } t
+		\boldsymbol{\hat{\beta}}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta} \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\boldsymbol{\beta}\parallel_1+\alpha\parallel\boldsymbol{\beta}\parallel_2^2\leq t \textrm{ for some } t
 	\end{equation}
@@
 	%It seems to give good predictions. 
-	But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
-Grouping effect is the fact that correlated covariates get similar coefficient and are selected together.
+	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
 	\\
@@
 \section{Model to select decorrelated covariates}
-\subsection{The decorrelation problem in regression}\label{sectionOLS}
+\subsection{A classical problem: correlations in regression}\label{sectionOLS}
 For a model defined by 

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@490 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@491 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
-Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
+Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even without bias from modelisation. 
 In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
@@
 	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}).
-	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith grouped $\beta_j$ following a Gaussian Mixture distribution.
+	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables with grouped $\beta_j$ following a Gaussian Mixture distribution.
 
 The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$.
-In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\hat{\beta}$.
+In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\boldsymbol{\hat{\beta}}$.
 But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net.
@@
 So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN).
-The structure is based on a full generative model on $X$ to be able to compare likelihood-based criterion in the \textsc{MCMC}.
+The structure is based on a full generative model on $\boldsymbol{X}$ to be able to compare different size structures with a likelihood-based criterion in the \textsc{MCMC}.
  	\\
@@
 	\begin{equation}
-		Y_{|X}=X\beta + \varepsilon
-	\end{equation}
-	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
+		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X\beta} + \boldsymbol{\varepsilon}
+	\end{equation}
+	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ with $\boldsymbol{I_n}$ the $n$-sized identity matrix,
 	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
 	\begin{equation}
-		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
+		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
 	\end{equation}
@@
 	\begin{equation}
-		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1} \label{eqOLS}
-	\end{equation}
-	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
-	\begin{eqnarray}
-		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
-			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
-	\end{eqnarray}	 
-	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
+		\operatorname{Var}(\boldsymbol{\hat{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
+	\end{equation}
+	and without any bias.
+	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance explodes.
 	Another problem is that matricial inversion requires $n\geq p$.
@@
 Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
-\subsection{Modelisation of the correlations}
-Let now $X \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
-We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$.
-Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $p_1=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression.
-Thus we have $p_2=| I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.
-
-In the following, we note $X^j$ the $j^{th}$ column of a matrix $X$.
-For lighter notation we define $X_2=X^{I_2}$ the matrix of the endogenous covariates and $X_1=X\setminus X_2$ the matrix of the remaining exogenous covariates.
-
-The family of $p_2$ regressions within correlated covariates in $X$ is noted:
-	\begin{equation}
-		\forall j \in I_2: X^j_{|X_1,S}=X_1\delta_j+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n) \label{SR}
-	\end{equation}
-		where $\delta_j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
-It remains $p-p_2$ independent exogenous variables $X_1$: $\forall j \notin I_2 : X^j \sim f(\theta_j)$
-
-We make the hypothesis of the uncrossing rule $I_1\cap I_2=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus we have a partition $X=[X_1,X_2]$.
-
-This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian (see the appendices).
-These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate the $\delta_j$ and $\sigma_j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
+\subsection{Our proposal: modelisation of the correlations}
+Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
+We propose to explicitly define a family of $p_2$ internal regressions between covariates with $I_2$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$.
+Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $\boldsymbol{p_1}=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression.
+Thus we have $p_2=|I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.
+
+In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of a matrix $\boldsymbol{X}$.
+For lighter notation we define $\boldsymbol{X_2}=\boldsymbol{X^{I_2}}$ the matrix of the endogenous covariates and $\boldsymbol{X_1}=\boldsymbol{X}\setminus \boldsymbol{X_2}$ the matrix of the remaining exogenous covariates.
+
+The family of $p_2$ regressions within correlated covariates in $\boldsymbol{X}$ is noted:
+	\begin{equation}
+		\boldsymbol{X_2}_{|\boldsymbol{X_1},S} \textrm{ defined by }\forall j \in I_2: \boldsymbol{X^j}_{|\boldsymbol{X_1},S}=\boldsymbol{X_1\alpha_j}+\boldsymbol{\varepsilon_j} \textrm{ with } \boldsymbol{\varepsilon_j} \sim(\boldsymbol{0},\sigma^2_j\boldsymbol{I_n}) \label{SR}
+	\end{equation}
+		where $\boldsymbol{\delta_j} \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
+
+We make the hypothesis of the uncrossing rule $I_1\cap I_2=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus we have a partition $\boldsymbol{X}=\left[\boldsymbol{X_1},\boldsymbol{X_2}\right]$.
+
+The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon_j}$) would not stay Gaussian (see the appendices).
+These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate the $\boldsymbol{\delta_j}$ and $\sigma_j$ like \textsc{SUR} with Feasible Generalized Least Squares (\textsc{FGLS}) by Zellner \cite{SURzellner} or \textsc{SPRING} \cite{chiquetconf} as mentioned in the introduction.
 \\
 
-\subsection{Marginal regression model with decorrelated covariates}
-Let $Y \in \mathcal{R}^n$ be a response variable we want to explain with $X$:
-	\begin{equation}
-			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y \textrm{ whith } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
+\subsection{A by-product model: marginal regression with decorrelated covariates}
+Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $X$:
+	\begin{equation}
+			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y \textrm{ with } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
 		\end{equation}
@@
 
-\subsection{Advantages for interpretation}\label{interpretation}
+\subsection{Strategy of use: pretreatment before classical selection methods}\label{interpretation}
 The explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset.
@@
 
-\subsection{Estimator and properties}
 	The marginal regression model simply is
@@
 			 \\
+	\subsection{Illustration of the tradeoff conveyed by the pretreatment}		  
+	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
+	\begin{eqnarray}
+		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
+			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
+	\end{eqnarray}	 
+	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables.
+$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
+We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
+Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
+Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
+	\begin{equation}
+		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
+	\end{equation}
+
+\begin{figure}[h!]
+	\begin{minipage}[l]{.32\linewidth}
+			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
+			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
+	\end{minipage} \hfill
+	\begin{minipage}[c]{.32\linewidth}
+			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
+			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
+	\end{minipage} \hfill
+   \begin{minipage}[r]{.32\linewidth}
+			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
+			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
+   \end{minipage} 
+\end{figure} 
+It is clear in Figures \ref{MQE1} to \ref{MQE3} that the marginal model is more robust than \textsc{OLS} on $X$.
+And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it does not explode until extreme values (sub-regression nearly fully explained by the noise).
+We also see that the error implied by strong correlations shrinks with the rise of $n$.
+$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$ where both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
+Thus, as $\operatorname{Var}(\hat{\alpha}_{2})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
+It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
 
@@
 
-	\subsection{Theoretical Mean squared errors on $\hat \beta$} 
-	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables.
-$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
-We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
-Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
-Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
-	\begin{equation}
-		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
-	\end{equation}
-
-\begin{figure}[h!]
-	\begin{minipage}[l]{.32\linewidth}
-			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
-			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
-	\end{minipage} \hfill
-	\begin{minipage}[c]{.32\linewidth}
-			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
-			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
-	\end{minipage} \hfill
-   \begin{minipage}[r]{.32\linewidth}
-			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
-			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
-   \end{minipage} 
-\end{figure} 
-It is clear in Figures \ref{MQE1} to \ref{MQE3} that the marginal model is more robust than \textsc{OLS} on $X$.
-And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it does not explode until extreme values (sub-regression nearly fully explained by the noise).
-We also see that the error implied by strong correlations shrinks whith the rise of $n$.
-$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$ where both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
-Thus, as $\operatorname{Var}(\hat{\alpha}_{2})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
-It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
-
 	\subsection{The datasets}	
@@
 
-We observe that \textsc{CorReg} is better than classical methods especially whith the Hierarchical Uniform law.
+We observe that \textsc{CorReg} is better than classical methods especially with the Hierarchical Uniform law.
 When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods.
@@
 			\centering
-				\label{compareMSEexfos}
+				\label{barplotMSEexfos}
 				\includegraphics[width=400px]{figures/MSEfinal.png}

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@492 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{A by-product model: marginal regression with decorrelated covariates}
-Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $X$:
-	\begin{equation}
-			Y_{|S,X}=X\beta+\varepsilon_Y =X_1\beta_1+X_2\beta_2+\varepsilon_Y \textrm{ with } \varepsilon_Y \sim \mathcal{N}(0,\sigma^2_YI_n);	\label{MainR}
-		\end{equation}
-		where $\beta=(\beta_1,\beta_2) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $I_n$ the identity matrix. 
-We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $X_2$ a regression model on $Y$ {\it depending only on uncorrelated covariates $X_1$}:
+Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}$:
+	\begin{equation}
+			\boldsymbol{Y}_{|\boldsymbol{X_1},\boldsymbol{X_2},S}=\boldsymbol{X_1\beta_1}+\boldsymbol{X_2\beta_2}+\boldsymbol{\varepsilon_Y} \textrm{ with } \boldsymbol{\varepsilon_Y} \sim \mathcal{N}(\boldsymbol{0},\sigma^2_Y\boldsymbol{I_n});	\label{MainR}
+		\end{equation}
+		where $\boldsymbol{\beta}=(\boldsymbol{\beta_1},\boldsymbol{\beta_2}) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $\boldsymbol{I_n}$ the identity matrix. 
+We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $\boldsymbol{X_2}$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X_1}$}:
 \begin{eqnarray}
-	Y_{|X_1}&=&X_1 (\beta_{1}+ \sum_{j \in I_2}\beta_{j}\delta_j)+  \sum_{j \in I_2}\beta_{j}\varepsilon_{j}+\varepsilon_Y \label{Trueexpl} \\
-		&=&X_1 \alpha_1 + X_2\alpha_2 + \varepsilon_{\alpha} \textrm{ where } \alpha_2  = 0.
+	\boldsymbol{Y}_{|\boldsymbol{X_1},S}&=&\boldsymbol{X_1} (\boldsymbol{\beta_{1}}+ \sum_{j \in I_2}\beta_{j}\boldsymbol{\alpha_j})+  \sum_{j \in I_2}\beta_{j}\boldsymbol{\varepsilon_{j}}+\boldsymbol{\varepsilon_Y} \label{Trueexpl} \\
+	&=&\boldsymbol{X_1}\boldsymbol{\beta_1^*}+\boldsymbol{\varepsilon_Y^*}=\boldsymbol{X}\boldsymbol{\beta^*}+\boldsymbol{\varepsilon_Y^*} \textrm{ where }\boldsymbol{\beta^*}=(\boldsymbol{\beta_1^*},\boldsymbol{\beta_2^*}) \textrm{ and } \boldsymbol{\beta_2^*}=\boldsymbol{0}
 \end{eqnarray}
@@
 These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates.
-It is a variable pre-selection independent of the response $Y$.
+It is a variable pre-selection independent of the response $\boldsymbol{Y}$.
+
+\subsection{Strategy of use: pretreatment before classical selection methods}\label{interpretation}
 As a pretreatment, it allows usage of any method in a second time to estimate $\alpha$like LASSO or even stepwise \cite{seber2012linear}.
 
-\subsection{Strategy of use: pretreatment before classical selection methods}\label{interpretation}
 The explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset.
@@
 
+\section{Acknowledgements}
+	We want to thanks ArcelorMittal Atlantique \& Lorraine that has granted this work, given the chance to use \textsc{CorReg} on real dataset and authorized the package to be open-sourced licensed (\textsc{CeCILL}), especially Dunkerque's site where most of the work has been done.
 \bibliography{biblio}{}

### 2014-04-01 www/article13.tex
up structure
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@493 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-	\section{Estimating structure of sub-regressions with a Markov chain}	
-\subsection{Detailed generative model}
+\section{Sub-regressions model selection}	
+	\subsection{Modeling the uncorrelated covariates: a full generative approach}
 	To be able to compare structures with probabilistic criterions, we need a full generative model. We suppose that variables in $X_1$ follow Gaussian mixtures. 
@@
 	We now have a full generative model.
-	\subsection{Structure comparison} \label{compstruct}
+	\subsection{Penalization of the integrated likelihood} \label{compstruct}
   To obtain the marginal regression model we need $S$, the linear structure between the covariates.
@@
 
-	\subsection{The neighbourhood}
+\subsection{MCMC algorithm}
+	\subsubsection{The neighbourhood}
 	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
@@
 
-	\subsection{The walk}
 	\subsubsection{Transition probabilities}

### 2014-04-07 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@496 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{document}
-\maketitle
+
+\begin{center}
+{\Large
+	{\sc  CorReg : Régression sur variables Corrélées\\ et Application à l'industrie Sidérurgique}
+}
+\bigskip
+
+  Clément Théry$^{1}$ \& Christophe Biernacki$^{2}$ \& Gaétan Loridant$^{3}$
+\bigskip
+
+{\it
+$^{1}$ ArcelorMittal, Université Lille 1, Inria, CNRS, clement.thery@arcelormittal.com
+
+$^{2}$ Université Lille 1, Inria, CNRS, christophe.biernacki@math.univ-lille1.fr
+
+$^{3}$ Etudes Industrielles ArcelorMittal Dunkerque, gaetan.loridant@arcelormittal.com\textbf{}
+}
+\end{center}
+\bigskip
+april 2014
 \bigskip

### 2014-04-07 www/article13.tex
title
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@499 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
+%\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
 \title{Model-based pretreatment for regression with correlated variables.}
@@
 \begin{document}
-
 \begin{center}
 {\Large
-	{\sc  CorReg : Régression sur variables Corrélées\\ et Application à l'industrie Sidérurgique}
+	{\sc Model-based pretreatment for regression \\ with correlated variables.}
 }
@@
 \bigskip
-april 2014
-\bigskip
+
 {\bf Abstract.} Linear regression outcomes are known to be damaged by correlated covariates.

### 2014-04-07 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@500 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
-The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}).
+The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated with Ordinary Least Squares (\textsc{OLS}).
 %As shown in section \ref{sectionOLS},
@@
 
-	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
+	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty to obtain zeros in $\hat{\boldsymbol{\beta}}$:
 		\begin{equation}
@@
 	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. 
-Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension).
+%	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
 	 Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.

### 2014-04-07 www/article13.tex
up intro
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@501 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables with grouped $\beta_j$ following a Gaussian Mixture distribution.
-
 The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$.
 In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\boldsymbol{\hat{\beta}}$.
+
 	But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net.
@@
 
- Where some try to reduce the dimension and then have to suppose having no correlations between the remaining covariates,  we propose to focus on the correlations, giving a model with orthogonal covariates and an explicit structure between covariates.
+ Where some reduce the dimension without looking at correlations and then have to suppose having no correlations between the remaining covariates,  we propose to focus on the correlations, giving a marginal model with orthogonal covariates and an explicit structure between covariates.
+Our work is based on the assumption that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
  In fact we search the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
-This can be viewed as a pretreatment on the dataset allowing to use then other dimension reduction tools without suffering from correlations.
+This can be viewed as a pretreatment on the dataset allowing to use then other tools for dimension reduction and estimation without suffering from correlations.
+Practically speaking we use a system of linear regressions between the covariates to model the structure of the correlations.
+This structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint law on the dataset $\boldsymbol{X}$.
 %We only consider strong correlations (i.e.
@@
  %We will in a second time be able to use the remaining part of the information (sequential approach).
-
-	Our work is based on the assumption that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
-	The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
-Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
-  	Other methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).
-
-	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
-Thus we have one regression on our response variable and a system of sub-regressions (without the response variable).
-Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
-	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
-So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
-This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.
-So we dodge the drawbacks of both grouping effect and variable selection.
-
-\textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
-Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix).
-It consists in a selection in the precision matrix, setting some covariances to zero.
-But the resulting matrix is symmetric and we need an oriented structure for our \textsc{SEM}.
-So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN).
-The structure is based on a full generative model on $\boldsymbol{X}$ to be able to compare different size structures with a likelihood-based criterion in the \textsc{MCMC}.
- 	\\
 
@@
 The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon_j}$) would not stay Gaussian (see the appendices).
-These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate the $\boldsymbol{\delta_j}$ and $\sigma_j$ like \textsc{SUR} with Feasible Generalized Least Squares (\textsc{FGLS}) by Zellner \cite{SURzellner} or \textsc{SPRING} \cite{chiquetconf} as mentioned in the introduction.
+
+%ancienne intro
+
+The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
+Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
+  	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).
+  	\\ Here we suppose independency between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\delta_j}$ and $\sigma_j$.
+
+	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
+Thus we have one regression on our response variable and a system of sub-regressions (without the response variable).
+Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
+	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
+So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
+This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.
+So we dodge the drawbacks of both grouping effect and variable selection.
  	\\
+%ancienne intro
 
@@
 	\subsection{Modeling the uncorrelated covariates: a full generative approach}
+%ancienne intro	
+\textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
+Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix).
+It consists in a selection in the precision matrix, setting some covariances to zero.
+But the resulting matrix is symmetric and we need an oriented structure for our \textsc{SEM}.
+So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN).
+The structure is based on a full generative model on $\boldsymbol{X}$ to be able to compare different size structures with a likelihood-based criterion in the \textsc{MCMC}.
+	%ancienne intro
+
 	To be able to compare structures with probabilistic criterions, we need a full generative model. We suppose that variables in $X_1$ follow Gaussian mixtures. 

### 2014-04-07 www/article13.tex
update notation and MSE and tradeoff
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@502 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
+
+%ancienne intro
+
+\subsection{A by-product model: marginal regression with decorrelated covariates}
  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
@@
  	\\
-%ancienne intro
-
-\subsection{A by-product model: marginal regression with decorrelated covariates}
 Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}$:
@@
 	\begin{equation}
-		Y_{|X_1}= X_1\alpha_1+ \varepsilon_{\alpha}
+		\boldsymbol{Y}_{|\boldsymbol{X_1}}= \boldsymbol{X_1}\boldsymbol{\beta_1^*}+ \boldsymbol{\varepsilon_Y^*}
 	\end{equation}			
@@
 		\begin{equation}
-			\hat{\alpha}_{1} = (X'_{1} X_{1})^{-1}X'_{1}Y  \textrm{ and }\hat{\alpha}_{2} = 0
-		\end{equation}
-		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $Y$ and $\alpha$
-		but in terms of $\beta$ this estimator is biased:
-		\begin{equation}
-			\operatorname{E}[\hat{\alpha_1}|X_1]=\beta_1+\sum_{j \in I_2}\beta_{j}\delta_j \textrm{ and }\operatorname{E}[\hat{\alpha_2}|X_1]=0
+			\boldsymbol{\hat\beta_1^*} = (\boldsymbol{X_1}' \boldsymbol{X_1})^{-1}\boldsymbol{X_1}'\boldsymbol{Y}  \textrm{ and }\boldsymbol{\hat\beta_2^*} = \boldsymbol{0}
+		\end{equation}
+		We see in (\ref{Trueexpl}) that it gives an unbiased estimation of $\boldsymbol{Y}$ and $\boldsymbol{\beta^*}$
+		but in terms of $\boldsymbol{\beta}$ this estimator is biased:
+		\begin{equation}
+			\operatorname{E}[\boldsymbol{\hat\beta_1^*}|\boldsymbol{X_1}]=\boldsymbol{\beta_1}+\sum_{j \in I_2}\beta_{j}\boldsymbol{\alpha_j} \textrm{ and }\operatorname{E}[\boldsymbol{\hat\beta_2^*}|\boldsymbol{X_1}]=\boldsymbol{0}
 		\end{equation}
@@
 		\begin{equation}
-			\operatorname{Var}[\hat{\alpha}_{1}|X_1]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 )(X'_{1} X_{1})^{-1}  \textrm{ and }\operatorname{Var}[\hat{\alpha}_{2}|X_1]= 0 
+			\operatorname{Var}[\boldsymbol{\hat\beta_1^*}|\boldsymbol{X_1}]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 )(\boldsymbol{X_1}' \boldsymbol{X_1})^{-1}  \textrm{ and }\operatorname{Var}[\boldsymbol{\hat\beta_2^*}|\boldsymbol{X_1}]= \boldsymbol{0}
 		\end{equation}
@@
 	\subsection{Illustration of the tradeoff conveyed by the pretreatment}		  
-	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
+	 The Mean Squared Error (\textsc{MSE}) on $\boldsymbol{\hat{\beta}}$ is:
 	\begin{eqnarray}
-		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
-			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
+		\textsc{MSE}(\boldsymbol{\hat{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}})) \\
+			&=&0+\sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS and } 
 	\end{eqnarray}	 
@@
 We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
-Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model: $MSE(\hat{\beta}|\beta,X)=Bias^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))$ for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
+Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
 Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:

### 2014-04-07 www/article13.tex
tradeoff updated
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@503 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\begin{eqnarray}
-		\textsc{MSE}(\boldsymbol{\hat{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}})) \\
-			&=&0+\sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS and } 
+		\textsc{MSE}(\boldsymbol{\hat{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}})) \\
+			&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
+			&=& \parallel\sum_{j \in I_2}\beta_{j}\boldsymbol{\alpha_j} \parallel_2^2 +\parallel \boldsymbol{\beta_2}\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X_1}' \boldsymbol{X_1})^{-1})
 	\end{eqnarray}	 
 	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables.
-$X_1$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $X_3=x_1+x_2+\varepsilon_3$ where $\varepsilon_3\sim{\mathcal{N}(0,\sigma_3^2)}$.
-We also have $\beta=(1,1,1)$ and $\sigma_Y=10$.
-Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg} model for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
+$\boldsymbol{X_1}$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $\boldsymbol{x_3}=\boldsymbol{x_1}+\boldsymbol{x_2}+\boldsymbol{\varepsilon_3}$ where $\boldsymbol{\varepsilon_3}\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I_n})}$.
+We also have $\boldsymbol{\beta}=(1,1,1)$ and $\sigma_Y=10$.
+Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg}'s marginal  model for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
 Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
 	\begin{equation}
-		1-\mathcal{R}^2=\frac{\operatorname{Var}(\varepsilon_3)}{\operatorname{Var}(x_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
+		1-\mathcal{R}^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon_3)}}{\operatorname{Var}(\boldsymbol{x_3})}=\frac{\sigma_3^2}{\sigma_3^2+2}
 	\end{equation}
@@
 We also see that the error implied by strong correlations shrinks with the rise of $n$.
-$MSE_{OLS}=0+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}))=\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{1}))+\operatorname{Tr}(\operatorname{Var}(\hat{\beta}_{2}))$ where both terms are multiplied by $\sigma_Y^2$ and so is $\operatorname{Tr}(\operatorname{Var}(\hat{\alpha}_{1}))$.
-Thus, as $\operatorname{Var}(\hat{\alpha}_{2})=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
+We see that $\sigma_y$ multplies $\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}}))=\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{1}}))+\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{2}}))$ for both models but for the marginal model $\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{2}}))=0$.
+ Thus, as $\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{2}}))=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
 It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).

### 2014-04-07 www/article13.tex
full generative
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@504 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 This structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint law on the dataset $\boldsymbol{X}$.
+This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{CRAN}.
 %We only consider strong correlations (i.e.
@@
 %ancienne intro	
-\textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
-Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix).
-It consists in a selection in the precision matrix, setting some covariances to zero.
-But the resulting matrix is symmetric and we need an oriented structure for our \textsc{SEM}.
+Structural equations models like \textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
+Graphical LASSO \cite{friedman2008sparse} offers a method to obtain a structure based on the precision matrix (inverse of the variance-covariance matrix), setting some coefficients of the precision matrix to zero.
+But the resulting matrix is symmetric and we need an oriented structure for $S$ to avoid cycles.
 So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN).
-The structure is based on a full generative model on $\boldsymbol{X}$ to be able to compare different size structures with a likelihood-based criterion in the \textsc{MCMC}.
 	%ancienne intro
 
-	To be able to compare structures with probabilistic criterions, we need a full generative model. We suppose that variables in $X_1$ follow Gaussian mixtures. 
-	\begin{equation}
-			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1}
-		\end{equation}
-		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed.
+	To be able to compare structures with probabilistic likelihood-based criterions, we need a full generative model on $\boldsymbol{X}$ .
+We suppose that variables in $\boldsymbol{X_1}$ follow Gaussian mixtures.
+	\begin{equation}
+			\forall j \notin I_2 : \boldsymbol{X^j}_{|S} \sim f(\boldsymbol{\theta_j})=\mathcal{GM}(\boldsymbol{\pi_j};\boldsymbol{\mu_j};\boldsymbol{\sigma^2_j}) \textrm{ with } \boldsymbol{\pi_j},\boldsymbol{\mu_j},\boldsymbol{\sigma^2_j} \textrm{ vectors of size } k_j.
+\label{mixtureX1}
+		\end{equation}
+		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed (exponentially distributed for example).
 Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.
 
-		Variables in $X_1$ are supposed to be independent.
-	Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways.
+		Variables in $\boldsymbol{X_1}$ are supposed to be independent.
+	%Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways.
 %compute corresponding $\psi$ according to it.

### 2014-04-07 www/article13.tex
bic*
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@505 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   To obtain the marginal regression model we need $S$, the linear structure between the covariates.
-	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood according to the complexity of the structure~\cite{BIChuard}.
-We will prefer this kind of comparison criterion instead of cross-validation that is very time-consuming and thus not friendly with combinatory problematics.
-	We note $\Theta$ the set of the parameters of the generative model
+  Cross-validation is very time-consuming and thus not friendly with combinatory problematics.
+Moreover, we need a criterion compatible with structures of different sizes (varying $p_2$) and not related with $Y$ because the structure is inherent to $\boldsymbol{X}$ only.
+Thus it must be a global criterion.
+	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood of the joint law on $\boldsymbol{X}$ according to the complexity of the structure~\cite{BIChuard}.
+
+	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
 	\begin{eqnarray}

### 2014-04-07 www/article13.tex
MCMC
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@506 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\begin{eqnarray}
-		P(S|X)&\propto& P(X|S)P(S) \\
-		-2\log P(X|S)&\approx & BIC=-2\mathcal{L}(X,S,\Theta)+|\Theta|\log(n)\\  
+		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \\
+		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)\\  
 	\end{eqnarray}
@@
 	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
-	But we can imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context.
-\\	
-		We will now use the following notation: $\psi(S)=\psi(X|S)$	where $\psi(X|S)$ is the chosen criterion, that will be $BIC$ with hierarchical uniform hypothesis in our numerical results.
 		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.
+
+	We can also imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context.
 
@@
 	\subsubsection{Transition probabilities}
-	We first make the approximation
-	\begin{equation}
-		P(S|X)\approx exp(\psi(S)).
-	\end{equation}
-	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
+
+	The walk follows a time-homogeneous Markov Chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we just compute the probabilities when we need them).
 	At each step the markov chain moves with probabiliy:
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2} \psi(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}\psi(S_l))} \\
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}P(\tilde{S}|\boldsymbol{X}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}P(S_l|\boldsymbol{X}))} \\
 	\end{eqnarray}
@@
 %		
-and the output will be the best structure in terms of $\psi$ which weights each candidate. Practically speaking, \textsc{CorReg} returns the best structure seen during the walk.
+and the output will be the best structure in terms of $P(S|\boldsymbol{X})$ which weights each candidate. Practically speaking, \textsc{CorReg} returns the best structure seen during the walk.
 Numerical results (Section 4) illustrates the efficiency of the walk when the true model really contains a linear structure or no structure at all (Table (\ref{compZvrai})) and when the structure is not linear (Table \ref{compZnonlin})).

### 2014-04-08 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@507 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\begin{eqnarray}
-		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \\
+		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \label{approxBIC}\\
 		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)\\  
@@
 For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
-The $\delta_j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign.
+The $\boldsymbol{\alpha_j}$ and $\beta_j$ are generated according to the same Poisson law but with a random sign.
 $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.
@@
 		\subsubsection{How to evaluate found structure?}
-			The first criterion is $\psi$ which is minimised in the MCMC.
-But in our case, $\psi$ is based on the likelihood whose value don't have any intrinsic meaning.
+			The first criterion is $P(S|\boldsymbol{X})$ which is maximized in the MCMC.
+But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning.
 To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
@@
 We see that (HU) hypothesis gives sparser models even if $\max (\hat{p}_1^j)=5$.
-$p=40$ so the maximum value for $\hat{p_2}$ is $20$ in (HU) case and we see that this max is not reached.
-The stronger penalty implied by (HU) really is efficient.
+$p=40$ so the maximum value for $\hat{p_2}$ is $20$ in (HU) case and we see that this max is not reached, meaning that the stronger penalty implied by (HU) really is efficient.
 The datasets used for (U) and (HU) are the same to keep the comparison  meaningful.
@@
 
- It is also notable that (HU) has a greater computational cost than (U). But next versions of the package may optimize a bit more the penalization step and reduce this gap.
+ It is also notable that (HU) has a greater computational cost than (U). 
 	We also notice that the MCMC is faster when there are numerous correlations (rejecting more candidates). 
@@
 \hline
-$n$ & $p_2$ & $\psi$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
+\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
+\hline
+$n$ & $p_2$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure).
-The datasets used here were those from \ref{compZvrai}.
+The datasets used here were those from table \ref{compZvrai}.
 
 We observe that \textsc{CorReg} is better than classical methods especially with the Hierarchical Uniform law.
-When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods.
+When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods, so it is robust.
 We also observe in table \ref{YXlinstep} that simple methods like stepwise (here from the package \textsc{lars}) can give good results in prediction.
@@
 When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
-This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}).
+This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
 \begin{table}[h!]

### 2014-04-08 www/article13.tex
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@508 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Thus it must be a global criterion.
+	Because it is about model selection and we are able to provide a full generative model, we decide to follow a Bayesian approach (\cite{raftery1995bayesian}, \cite{andrieu1999joint},\cite{chipman2001practical}).
 	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood of the joint law on $\boldsymbol{X}$ according to the complexity of the structure~\cite{BIChuard}.
@@
 		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \label{approxBIC}\\
-		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)\\  
+		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)  
 	\end{eqnarray}
@@
  \subsubsection{Initialisation(s)}
-
  If we have some knowledge about some sub-regressions (physical models for example) we can add them in the found and/or initial structure. So the model is really expert-friendly.

### 2014-04-08 www/article13.tex
updated formalism
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@509 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	But $BIC$ tends to give too complex structures because we test a great range of models. 
-	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
+	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
 %	 Thus we have :
@@
 \subsection{Results on prediction}
-	\subsubsection{$Y$ depends on all variables in $X$}	 \label{tableMSEsimtout}	
+	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 \label{tableMSEsimtout}	
 We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure).
@@
 \hline 
-$n$ & $p_2$& $\psi$ & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$& $P(S)$ & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$.
+\caption{OLS and OLS combined with constrained \textsc{CorReg}.
+$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
 \textsc{CorReg} logically wins.
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$.
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
 \textsc{CorReg} logically wins.
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $P(S)$&indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \end{tabular} 
-\caption{Elasticnet combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X$. \textsc{CorReg} logically wins. LASSO was better.}\label{YXlinenet}
+\caption{Elasticnet combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$. \textsc{CorReg} logically wins. LASSO was better.}\label{YXlinenet}
 \end{table}
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X$. \textsc{CorReg} logically wins but stepwise is quite good compared to elasticnet.}\label{YXlinstep}
+\caption{Stepwise  combined with constrained \textsc{CorReg}.
+$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$.
+\textsc{CorReg} logically wins but stepwise is quite good compared to elasticnet.}\label{YXlinstep}
 \end{table}
@@
 \clearpage
-	\subsubsection{$Y$ depends only on covariates in $X_2$ (worst case for us)}	 \label{tableMSEsimgauche}
-We now try the method with a response depending only on variables in $X_2$. The datasets used here were still those from \ref{compZvrai}.
-Depending only on $X_2$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
+	\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_2}$ (worst case for us)}	 \label{tableMSEsimgauche}
+We now try the method with a response depending only on variables in $\boldsymbol{X_2}$. The datasets used here were still those from \ref{compZvrai}.
+Depending only on $\boldsymbol{X_2}$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
 
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&$P(S)$ &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$.
+\caption{OLS and OLS combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_2}$.
 Sometimes $\hat{S}$ gives better results than $S$ because $S$ is penalized by the fact that it relies only on covariates not in the true model.
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $P(S)$ &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$. }\label{YX2linLASSO}
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_2}$. }\label{YX2linLASSO}
 \end{table}
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $P(S)$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$Y$  depends on all variables in $X_2$.}\label{YX2linenet}
+\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_2}$.}\label{YX2linenet}
 \end{table}
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $Y$ depends on all variables in $X_2$.}\label{YX2linstep}
+\caption{Stepwise  combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X_2}$.}\label{YX2linstep}
 \end{table}
@@
  We have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
-Variables in $X_2$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $X_1$ .
+Variables in $\boldsymbol{X_2}$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $\boldsymbol{X_1}$ .
 Dependencies are still real but non linear.
@@
 \hline
-$n$ & $p_2$ & $\psi$ &Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
+\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
+\hline
+$n$ & $p_2$ & $P(S)$&Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
 \hline
@@
 \hline 
-$n$ & $p_2$&  $\psi$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_2$& $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
@@
 \end{tabular} 
-\caption{ $Y$ depends on all variables in $X$. \textsc{CorReg} is not too far from LASSO even if it stays behind. } \label{resYnonlin}
+\caption{ $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$. \textsc{CorReg} is not too far from LASSO even if it stays behind. } \label{resYnonlin}
 \end{table}
@@
 			\includegraphics[height=150px,width=150px]{figures/correlexfoshist.png} 
-			\caption{Histogram of correlations in $X$.} \label{compareMSEexfos}
+			\caption{Histogram of correlations in $\boldsymbol{X}$.} \label{compareMSEexfos}
    \end{minipage}
@@
 These first results are easily understandable and meet metallurgists expertise.
-			The algorithm gives a structure of $p_2=76$ subregressions with a mean of $\bar{p_1}=5.17$ regressors. In $X_1$ the number of $\rho>0.7$ is $\textbf{79.33\%}$ smaller than in $X$.
+			The algorithm gives a structure of $p_2=76$ subregressions with a mean of $\bar{p_1}=5.17$ regressors.
+In $\boldsymbol{X_1}$ the number of $\rho>0.7$ is $\textbf{79.33\%}$ smaller than in $\boldsymbol{X}$.
 
@@
 			\includegraphics[width=150px]{figures/correlbeforeafter.png} 
-			\caption{Correlations between the covariates in $X$ (upper) and $\hat{X}_1$ (lower).}
+			\caption{Correlations between the covariates in $\boldsymbol{X}$ (upper) and $\boldsymbol{\hat{X}_1}$ (lower).}
 	\end{minipage} \hfill
@@
 			\includegraphics[height=150px,width=150px]{figures/histcorrelBVBI.png} 
-			\caption{Histogram of correlations in $X$.} 
+			\caption{Histogram of correlations in $\boldsymbol{X}$.} 
    \end{minipage}
@@
 If the algorithm did not have time to converge (stationnarity), it can be continued with a few step for which the neighboorhood would only contain smaller candidates (in terms of complexity).
-It is equivalent to ask for each element in $I_1$ if the criterion $\psi$ would be better without it.
+It is equivalent to ask for each element in $I_1$ if the criterion $P(S|\boldsymbol{X})$ would be better without it.
 Thus it can be seen as a final cleaning step.

### 2014-04-09 www/article13.tex
up resume
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@517 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-{\bf Abstract.} Linear regression outcomes are known to be damaged by correlated covariates.
-However many modern datasets are expected to convey more and more highly correlated covariates.
-We propose to explicitly model the correlations by a family of linear regressions between the covariates.
-Marginalization allows then to obtain a parsimonious correlation-free regression model, easily understandable and from which it is then possible to perform standard linear estimation methods including variables selection procedures for instance.
+{\bf Abstract.} Linear regression outcomes are known to be damaged by highly correlated covariates.
+However many modern datasets are expected to convey more and more highly correlated covariates due to the global increase of the amount of variables in datasets.
+We propose to explicitly model the correlations by a family of linear regressions between the covariates, some covariates explaining others.
+It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable and from which it is possible to perform standard linear estimation methods including variables selection procedures for instance.
 The structure of correlations is found with an MCMC algorithm.

### 2014-04-09 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@518 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 We propose to explicitly model the correlations by a family of linear regressions between the covariates, some covariates explaining others.
-It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable and from which it is possible to perform standard linear estimation methods including variables selection procedures for instance.
+It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable.
+This pretreatment by variable selection gives then the possibility to perform standard linear estimation methods including variables selection procedures for instance.
 The structure of correlations is found with an MCMC algorithm.
@@
 It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn.
-Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
-Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even without bias from modelisation. 
-In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
+Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
+Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators. We note the linear regression model:
 \begin{equation}

### 2014-04-09 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@519 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn.
-Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
+Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from industry and astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
 Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators. We note the linear regression model:

### 2014-04-09 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@522 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\end{equation}
-	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$.
-The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated with Ordinary Least Squares (\textsc{OLS}).
+	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y >0$.
+The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{OLS}).
 %As shown in section \ref{sectionOLS},
-	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
+	Estimation of $\boldsymbol{\beta}$ requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
 Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :

### 2014-04-10 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@524 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\end{equation}
-	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y >0$.
+	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I}_n)$ the noise of the regression, with $\boldsymbol{I}_n$ the $n$-sized identity matrix and $\sigma_Y >0$.
 The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{OLS}).
@@
 	Estimation of $\boldsymbol{\beta}$ requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
-Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
-\begin{itemize}
-	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
-	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
-\end{itemize}
-
-	With the rise of informatic, datasets contains more and more covariates and thus more and more useless covariates.
-So dimension reduction becomes a necessity.
-Moreover, when you use more covariates, you increase the chance to have correlated ones.
+Conditionning of $\boldsymbol{X}'\boldsymbol{X}$ get worse based on two aspects : the dimension $p$ (number of covariates) of the model (the more covariates you have the greater variance you get)
+	 and the correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
+
+	With the rise of informatics, datasets contain more and more covariates and thus dimension reduction becomes a necessity.
+Moreover, increasing the number of covariates also increases the chance to have correlated ones.
 It will be illustrated with examples from steel industry in the last section of this paper (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
-	 In such a context, variance of the estimators can lead to arbitrary results or even no results at all.
-Prediction and interpretation are both strongly needed, depending on the context.
-For instance, in an industrial context interpretation could be favored to improve the process instead of only predict defects.
+	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
+For instance, in an industrial context interpretation could be favored to improve the process instead of only predict defects, so robust estimation is needed.
 		~\\	~\\
@@
 %bibliographie	
-	Because \textsc{OLS} is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
-Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.
+	Because \textsc{OLS} is the minimum-variance linear unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
 
@@
 	\begin{equation}
-		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel \boldsymbol{\beta} \parallel_2^2\leq k
+		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel \boldsymbol{\beta} \parallel_2^2\leq \lambda \textrm{ with } \lambda>0
 	\end{equation}
@@
 	So it gives difficult interpretations for large values of $p$.	
+	Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.
 
@@
 		\begin{equation}
-		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda
+		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda \textrm{ with } \lambda>0
 		\end{equation}	
@@
 And it  really selects covariates with coefficients set exactly to 0.
-	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. 
+
 %	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
-	 Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
+	 Some recent variants of the \textsc{LASSO} do exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
 	 \\
@@
 	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
+Hence interpretations are not the same and nothing explains why.
+So it can be very confusing.
 	\\
@@
 Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
+	 \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates and we see that none of the above methods take the correlations into account.
+
 	~\\	~\\
@@
 
- Where some reduce the dimension without looking at correlations and then have to suppose having no correlations between the remaining covariates,  we propose to focus on the correlations, giving a marginal model with orthogonal covariates and an explicit structure between covariates.
+ Where some reduce the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates that helps to define a marginal model with uncorrelated covariates.
+Thus we have variable selection with explanation and improve interpretation and not only prediction.
  Our work is based on the assumption that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
@@
  	This paper will first present the marginal regression model and its properties before describing in Section 3 the random walk used to find the structure.
- 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in the last part.
+ 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in Section \ref{conclusion}.
 
@@
 	and without any bias.
-	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance explodes.
+	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
 	Another problem is that matricial inversion requires $n\geq p$. 
-When it is not the case, a frequently used alternative is the Moore-Penrose generalized inverse \cite{katsikis2008fast}.
-In the followings, matrices inverses when $n<p$ will refer to this generalized inverse.
-Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
 \subsection{Our proposal: modelisation of the correlations}
 Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
-We propose to explicitly define a family of $p_2$ internal regressions between covariates with $I_2$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$.
-Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $\boldsymbol{p_1}=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression.
-Thus we have $p_2=|I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.
-
 In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of a matrix $\boldsymbol{X}$.
-For lighter notation we define $\boldsymbol{X_2}=\boldsymbol{X^{I_2}}$ the matrix of the endogenous covariates and $\boldsymbol{X_1}=\boldsymbol{X}\setminus \boldsymbol{X_2}$ the matrix of the remaining exogenous covariates.
-
-The family of $p_2$ regressions within correlated covariates in $\boldsymbol{X}$ is noted:
-	\begin{equation}
-		\boldsymbol{X_2}_{|\boldsymbol{X_1},S} \textrm{ defined by }\forall j \in I_2: \boldsymbol{X^j}_{|\boldsymbol{X_1},S}=\boldsymbol{X_1\alpha_j}+\boldsymbol{\varepsilon_j} \textrm{ with } \boldsymbol{\varepsilon_j} \sim(\boldsymbol{0},\sigma^2_j\boldsymbol{I_n}) \label{SR}
-	\end{equation}
-		where $\boldsymbol{\delta_j} \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
-
-We make the hypothesis of the uncrossing rule $I_1\cap I_2=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus we have a partition $\boldsymbol{X}=\left[\boldsymbol{X_1},\boldsymbol{X_2}\right]$.
-
-The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon_j}$) would not stay Gaussian (see the appendices).
+
+We propose to explicitly define a family of $p_r$ internal regressions between covariates with $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
+We note $\boldsymbol{X}_r=\boldsymbol{X}^{I_r}$ the sub-matrix of the redundant covariates (explained by others) and  $\boldsymbol{X}_f$ the remaining submatrix of free covariates (independent explaining covariates).
+We focus here on the expression of $P(\boldsymbol{X}_r|\boldsymbol{X}_f)$.
+
+We have $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$.
+Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $\boldsymbol{p_f}=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each internal regression.
+Thus we have $p_r=|I_r|$ and $p_f^j=|I_f^j|$ where $|.|$ represents the cardinal of an ensemble.
+
+The family of $p_r$ regressions within correlated covariates in $\boldsymbol{X}$ is noted:
+	\begin{equation}
+		\boldsymbol{X}_{r|\boldsymbol{X}_f,S} \textrm{ defined by }\forall j \in I_r: \boldsymbol{X}^j_{|\boldsymbol{X}_f,S}=\boldsymbol{X}_f\boldsymbol{\alpha}_j+\boldsymbol{\varepsilon}_j \textrm{ with } \boldsymbol{\varepsilon}_j \sim(\boldsymbol{0},\sigma^2_j\boldsymbol{I}_n) \label{SR}
+	\end{equation}
+		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_f^j$).
+
+We make the hypothesis of the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus $\boldsymbol{X}_f$ and $\boldsymbol{X}_f$ give a partition of $\boldsymbol{X}$, so $\boldsymbol{X}_f=\boldsymbol{X}\setminus \boldsymbol{X}_r$.
+The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
 
@@
 Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
-  	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).
-  	\\ Here we suppose independency between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\delta_j}$ and $\sigma_j$.
+  	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) without explicit expression of the correlations and thus a smaller interpretation potential.
+  	\\ Here we suppose independency between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\alpha}_j$ and $\sigma_j$.
 
 	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
-Thus we have one regression on our response variable and a system of sub-regressions (without the response variable).
+Thus we have one regression on our response variable ($P(\boldsymbol{Y}|\boldsymbol{X})$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$).
 Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
@@
 \subsection{A by-product model: marginal regression with decorrelated covariates}
- The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss.
-So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables.
-This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.
-So we dodge the drawbacks of both grouping effect and variable selection.
+Now we know $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$ by the structure of sub-regressions, we are able to define a marginal regression model $P(\boldsymbol{Y}|\boldsymbol{X}_f,S)$ based on the reduced set of independent covariates $\boldsymbol{X}_f$ without significant information loss.
  	\\
-Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}$:
-	\begin{equation}
-			\boldsymbol{Y}_{|\boldsymbol{X_1},\boldsymbol{X_2},S}=\boldsymbol{X_1\beta_1}+\boldsymbol{X_2\beta_2}+\boldsymbol{\varepsilon_Y} \textrm{ with } \boldsymbol{\varepsilon_Y} \sim \mathcal{N}(\boldsymbol{0},\sigma^2_Y\boldsymbol{I_n});	\label{MainR}
-		\end{equation}
-		where $\boldsymbol{\beta}=(\boldsymbol{\beta_1},\boldsymbol{\beta_2}) \in  \mathcal{R}^p$ is the vector of the regression coefficients and $\boldsymbol{I_n}$ the identity matrix. 
-We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $\boldsymbol{X_2}$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X_1}$}:
+Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$:
+	\begin{equation}
+			\boldsymbol{Y}_{|\boldsymbol{X}_f,\boldsymbol{X}_r,S}=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon_Y} \textrm{ with } \boldsymbol{\varepsilon_Y} \sim \mathcal{N}(\boldsymbol{0},\sigma^2_Y\boldsymbol{I}_n);	\label{MainR}
+		\end{equation}
+		where $\boldsymbol{\beta}=(\boldsymbol{\beta}_f,\boldsymbol{\beta}_r) \in  \mathcal{R}^p$ is the vector of the regression coefficients associated respectively to $\boldsymbol{X}_f$ and $\boldsymbol{I}_n$ the identity matrix.
+We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $\boldsymbol{X}_r$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X}_f$}:
 \begin{eqnarray}
-	\boldsymbol{Y}_{|\boldsymbol{X_1},S}&=&\boldsymbol{X_1} (\boldsymbol{\beta_{1}}+ \sum_{j \in I_2}\beta_{j}\boldsymbol{\alpha_j})+  \sum_{j \in I_2}\beta_{j}\boldsymbol{\varepsilon_{j}}+\boldsymbol{\varepsilon_Y} \label{Trueexpl} \\
-	&=&\boldsymbol{X_1}\boldsymbol{\beta_1^*}+\boldsymbol{\varepsilon_Y^*}=\boldsymbol{X}\boldsymbol{\beta^*}+\boldsymbol{\varepsilon_Y^*} \textrm{ where }\boldsymbol{\beta^*}=(\boldsymbol{\beta_1^*},\boldsymbol{\beta_2^*}) \textrm{ and } \boldsymbol{\beta_2^*}=\boldsymbol{0}
+	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_2^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}
 \end{eqnarray}
@@
 \subsection{Strategy of use: pretreatment before classical selection methods}\label{interpretation}
-As a pretreatment, it allows usage of any method in a second time to estimate $\alpha$like LASSO or even stepwise \cite{seber2012linear}.
+As a pretreatment, it allows usage of any method in a second time to estimate $\boldsymbol{\beta}^*$, even with variable selection methods like LASSO or even stepwise \cite{seber2012linear}.
+
+After selection and estimation we will obtain a model with { \it two steps of variable selection}: the decorrelation step (coerced zeros associated to redundant information) and the classical selection step, with different meanings for obtained zeros in $\hat{\boldsymbol{\beta}}^*$.
+ This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.
+So we dodge the drawbacks of both grouping effect and variable selection.
 
@@
 
-After selection and estimation we will obtain a model with { \it two kinds of zeros}: those from decorrelating step and those from selection step, with different meanings. 
-Thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.
-
-Each equation here is very simple (only linear regressions). 
 Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
@@
 	\begin{equation}
-		\boldsymbol{Y}_{|\boldsymbol{X_1}}= \boldsymbol{X_1}\boldsymbol{\beta_1^*}+ \boldsymbol{\varepsilon_Y^*}
-	\end{equation}			
-		So we have the \textsc{OLS} unbiased estimator of $\alpha$: 
-		\begin{equation}
-			\boldsymbol{\hat\beta_1^*} = (\boldsymbol{X_1}' \boldsymbol{X_1})^{-1}\boldsymbol{X_1}'\boldsymbol{Y}  \textrm{ and }\boldsymbol{\hat\beta_2^*} = \boldsymbol{0}
+		\boldsymbol{Y}_{|\boldsymbol{X}_f}= \boldsymbol{X}_f\boldsymbol{\beta}_f^*+ \boldsymbol{\varepsilon}_Y^*
+	\end{equation}			
+		So we have the \textsc{OLS} unbiased estimator of $\boldsymbol{\beta}^*$: 
+		\begin{equation}
+			\hat{\boldsymbol{\beta}}_f^* = (\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{Y}  \textrm{ and }\boldsymbol{\hat\beta}_r^* = \boldsymbol{0}
 		\end{equation}
@@
 		\begin{equation}
-			\operatorname{E}[\boldsymbol{\hat\beta_1^*}|\boldsymbol{X_1}]=\boldsymbol{\beta_1}+\sum_{j \in I_2}\beta_{j}\boldsymbol{\alpha_j} \textrm{ and }\operatorname{E}[\boldsymbol{\hat\beta_2^*}|\boldsymbol{X_1}]=\boldsymbol{0}
+			\operatorname{E}[\hat{\boldsymbol{\beta}}_f^*|\boldsymbol{X}_f]=\boldsymbol{\beta}_f+\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \textrm{ and }\operatorname{E}[\hat{\boldsymbol{\beta}}_r^*|\boldsymbol{X}_f]=\boldsymbol{0}
 		\end{equation}
@@
 		\begin{equation}
-			\operatorname{Var}[\boldsymbol{\hat\beta_1^*}|\boldsymbol{X_1}]= (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 )(\boldsymbol{X_1}' \boldsymbol{X_1})^{-1}  \textrm{ and }\operatorname{Var}[\boldsymbol{\hat\beta_2^*}|\boldsymbol{X_1}]= \boldsymbol{0}
+			\operatorname{Var}[\hat{\boldsymbol{\beta}}_f^*|\boldsymbol{X}_f]= (\sigma^2_Y+\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2 )(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}  \textrm{ and }\operatorname{Var}[\hat{\boldsymbol{\beta}}_r^*|\boldsymbol{X}_f]= \boldsymbol{0}
 		\end{equation}
@@
 	\subsection{Illustration of the tradeoff conveyed by the pretreatment}		  
-	 The Mean Squared Error (\textsc{MSE}) on $\boldsymbol{\hat{\beta}}$ is:
+	 The Mean Squared Error (\textsc{MSE}) on $\hat{\boldsymbol{\beta}}$ is:
 	\begin{eqnarray}
-		\textsc{MSE}(\boldsymbol{\hat{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}})) \\
+		\textsc{MSE}(\hat{\boldsymbol{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}})) \\
 			&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
-			&=& \parallel\sum_{j \in I_2}\beta_{j}\boldsymbol{\alpha_j} \parallel_2^2 +\parallel \boldsymbol{\beta_2}\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X_1}' \boldsymbol{X_1})^{-1})
+			&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1})
 	\end{eqnarray}	 
 	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables.
-$\boldsymbol{X_1}$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $\boldsymbol{x_3}=\boldsymbol{x_1}+\boldsymbol{x_2}+\boldsymbol{\varepsilon_3}$ where $\boldsymbol{\varepsilon_3}\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I_n})}$.
-We also have $\boldsymbol{\beta}=(1,1,1)$ and $\sigma_Y=10$.
+$\boldsymbol{X}_f$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$.
+We also have $\boldsymbol{\beta}=(1,1,1)$ and $\sigma_Y \in \{1,10\}$  .
 Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg}'s marginal  model for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
-Figures \ref{MQE1} to \ref{MQE3} show the theoretical MSE evolution with the strength of the sub-regression:
-	\begin{equation}
-		1-\mathcal{R}^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon_3)}}{\operatorname{Var}(\boldsymbol{x_3})}=\frac{\sigma_3^2}{\sigma_3^2+2}
+Figure \ref{MQE1} shows the theoretical MSE evolution with the strength of the sub-regression:
+	\begin{equation}
+		1-\mathcal{R}^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon)_3}}{\operatorname{Var}(\boldsymbol{x}_3)}=\frac{\sigma_3^2}{\sigma_3^2+2}
 	\end{equation}
@@
 \begin{figure}[h!]
-	\begin{minipage}[l]{.32\linewidth}
-			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
-			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
-	\end{minipage} \hfill
-	\begin{minipage}[c]{.32\linewidth}
-			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
-			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
-	\end{minipage} \hfill
-   \begin{minipage}[r]{.32\linewidth}
-			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
-			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
-   \end{minipage} 
+%	\begin{minipage}[l]{.32\linewidth}
+%			\includegraphics[width=170px]{figures/MQEn15sigmaY10.png} 
+%			\caption{For $n=15$. Dotted: \textsc{Correg}, plain: OLS}\label{MQE1}
+%	\end{minipage} \hfill
+%	\begin{minipage}[c]{.32\linewidth}
+%			\includegraphics[ width=170px]{figures/MQEn100sigmaY10.png} 
+%			\caption{For $n=100$. Dotted: \textsc{Correg}, plain: OLS}
+%	\end{minipage} \hfill
+%   \begin{minipage}[r]{.32\linewidth}
+%			\includegraphics[width=170px]{figures/MQEn1000sigmaY10.png} 
+%			\caption{For $n=1000$. Dotted: \textsc{Correg}, plain: OLS.} \label{MQE3}
+%   \end{minipage} 
+	\includegraphics[width=500px]{figures/MQEexplOLS.png}\label{MQE1}
+	\caption{MSE of OLS and CorReg (dotted) estimators for varying $(1-R^2)$ of the sub-regression, $n$ and $\sigma_Y$.}
 \end{figure} 
-It is clear in Figures \ref{MQE1} to \ref{MQE3} that the marginal model is more robust than \textsc{OLS} on $X$.
+It is clear in Figure \ref{MQE1} that the marginal model is more robust than \textsc{OLS} on $X$.
 And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it does not explode until extreme values (sub-regression nearly fully explained by the noise).
 We also see that the error implied by strong correlations shrinks with the rise of $n$.
-We see that $\sigma_y$ multplies $\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}}))=\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{1}}))+\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{2}}))$ for both models but for the marginal model $\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{2}}))=0$.
- Thus, as $\operatorname{Tr}(\operatorname{Var}(\boldsymbol{\hat{\beta}_{2}}))=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
+We see that $\sigma_Y$ multiplies $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}))=\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{f}))+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))$ for both models but for the marginal model $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$.
+ Thus, as $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
 It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
@@
 	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.
-
+	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$ (see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
 		\subsection{Finding the structure}
@@
 
-\section{Conclusion and perspectives}
+\section{Conclusion and perspectives} \label{conclusion}
 	We have seen that correlations can lead to serious estimation and variable selection problems in linear regression and that in such a context, it can be useful to explicitly model the structure between the covariates and to use this structure (even sequentially) to avoid correlations issues.
@@
 
-\subsection{Alternative neighbourhoods for the MCMC}
+	\subsection{The \textsc{CorReg} package}
+\subsubsection{Alternative neighbourhoods for the MCMC}
 	We have here at each step $|\mathcal{V}_{S,j}|=p$ candidates but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the internal regressions or the whole structure, a maximum number of internal regressions, {\it etc.}).

### 2014-04-14 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@525 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 We have $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$.
-Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $\boldsymbol{p_f}=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each internal regression.
-Thus we have $p_r=|I_r|$ and $p_f^j=|I_f^j|$ where $|.|$ represents the cardinal of an ensemble.
+Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each internal regression  and $p_f^j=|I_f^j|$.
+Here $|.|$ represents the cardinal of an ensemble.
 
@@
 	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
-	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_2^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_2^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
 \end{eqnarray}
@@
 
-Moreover, the uncrossing constraint ($I_1\cap I_2=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
-
-	The marginal regression model simply is
-	\begin{equation}
-		\boldsymbol{Y}_{|\boldsymbol{X}_f}= \boldsymbol{X}_f\boldsymbol{\beta}_f^*+ \boldsymbol{\varepsilon}_Y^*
-	\end{equation}			
-		So we have the \textsc{OLS} unbiased estimator of $\boldsymbol{\beta}^*$: 
+Moreover, the uncrossing constraint ($I_f\cap I_r=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
+
+	For the marginal regression model defined in (\ref{modexpl})
+%	\begin{equation}
+%		\boldsymbol{Y}_{|\boldsymbol{X}_f}= \boldsymbol{X}_f\boldsymbol{\beta}_f^*+ \boldsymbol{\varepsilon}_Y^*
+%	\end{equation}			
+%		So 
+we have the \textsc{OLS} unbiased estimator of $\boldsymbol{\beta}^*$: 
 		\begin{equation}
@@
 It is clear in Figure \ref{MQE1} that the marginal model is more robust than \textsc{OLS} on $X$.
-And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it does not explode until extreme values (sub-regression nearly fully explained by the noise).
+And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it remains stable until extreme values (sub-regression nearly fully explained by the noise).
 We also see that the error implied by strong correlations shrinks with the rise of $n$.
@@
 \section{Sub-regressions model selection}	
-	\subsection{Modeling the uncorrelated covariates: a full generative approach}
-%ancienne intro	
 Structural equations models like \textsc{SEM} are often used in social sciences and economy where a structure is supposed "by hand" but here we want to find it automatically.
@@
 But the resulting matrix is symmetric and we need an oriented structure for $S$ to avoid cycles.
-So we developped an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN).
-	%ancienne intro
-
-	To be able to compare structures with probabilistic likelihood-based criterions, we need a full generative model on $\boldsymbol{X}$ .
-We suppose that variables in $\boldsymbol{X_1}$ follow Gaussian mixtures.
-	\begin{equation}
-			\forall j \notin I_2 : \boldsymbol{X^j}_{|S} \sim f(\boldsymbol{\theta_j})=\mathcal{GM}(\boldsymbol{\pi_j};\boldsymbol{\mu_j};\boldsymbol{\sigma^2_j}) \textrm{ with } \boldsymbol{\pi_j},\boldsymbol{\mu_j},\boldsymbol{\sigma^2_j} \textrm{ vectors of size } k_j.
+
+Cross-validation is very time-consuming and thus not friendly with combinatory problematics.
+Moreover, we need a criterion compatible with structures of different sizes (varying $p_2$) and not related with $Y$ because the structure is inherent to $\boldsymbol{X}$ only.
+Thus it must be a global criterion.
+
+We want to find the structure $S$ that maximizes $P(S|\boldsymbol{X})$ and we have:	
+	\begin{eqnarray}
+	 \label{approxBIC} P(S|\boldsymbol{X})&\propto & P(\boldsymbol{X}|S)P(S)
+	=P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)P(\boldsymbol{X}_f|S)P(S)
+	\end{eqnarray}
+
+	\subsection{Modeling the uncorrelated covariates: a full generative approach on $P(\boldsymbol{X})$}
+	To be able to compare structures with probabilistic likelihood-based criterions, we need a full generative model on $\boldsymbol{X}$.
+Sub-regressions give $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S) $ but $P(\boldsymbol{X}_f|S)$ is still undefined.
+We suppose that variables in $\boldsymbol{X}_f$ follow Gaussian mixtures of $k_j \in \mathbf{N}^*$ components:
+	\begin{equation}
+			\forall j \notin I_r : \boldsymbol{X}^j_{|S} \sim f(\boldsymbol{\theta}_j)=\mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j.
 \label{mixtureX1}
@@
 
-		Variables in $\boldsymbol{X_1}$ are supposed to be independent.
+		Variables in $\boldsymbol{X}_f$ are in the followings supposed to be independent.
 	%Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways.
@@
 	We now have a full generative model.
-	\subsection{Penalization of the integrated likelihood} \label{compstruct}
-  To obtain the marginal regression model we need $S$, the linear structure between the covariates.
-  Cross-validation is very time-consuming and thus not friendly with combinatory problematics.
-Moreover, we need a criterion compatible with structures of different sizes (varying $p_2$) and not related with $Y$ because the structure is inherent to $\boldsymbol{X}$ only.
-Thus it must be a global criterion.
+
+	\subsection{Penalization of the integrated likelihood by $P(S)$} \label{compstruct}
+
 Because it is about model selection and we are able to provide a full generative model, we decide to follow a Bayesian approach (\cite{raftery1995bayesian}, \cite{andrieu1999joint},\cite{chipman2001practical}).
+
   Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood of the joint law on $\boldsymbol{X}$ according to the complexity of the structure~\cite{BIChuard}.
@@
 	\begin{eqnarray}
-		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \label{approxBIC}\\
 		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)  
@@
 	But $BIC$ tends to give too complex structures because we test a great range of models. 
-	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
+	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
 %	 Thus we have :
@@
 %	\end{eqnarray}		
-	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
-	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
-		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.
-
-	We can also imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context.
+	It increases penalty on complexity for $p_r<\frac{p}{2}$ and $p_f^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
+	this constraint on $\hat{p}_r$ and $\hat{p}_f^j$ is given in the research algorithm when the Hierarchical Uniform (HU) hypothesis is made instead of Uniform one (U) in numerical experiments (section \ref{sectionsimul} and \ref{sectionrealcase}).
+		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we obtain better results when (HU) this constraint is verified.
+
+%	We can also imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context.
+%
+
+	%ancienne intro	
+	%ancienne intro
 
 \subsection{MCMC algorithm}
+	Now we have a comparison criterion, we define an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN). 
 	\subsubsection{The neighbourhood}
-	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
+	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_f\cap I_r=\emptyset$).
 	\\
@@
 	\begin{itemize}
-		\item if $i \notin I_i^j$ (add): 
+		\item if $i \notin I_f^j$ (add): 
 			\begin{itemize}
-				\item $I_1^j=I_1^j\cup \{i\}$
-				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
-				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
+				\item $I_f^j=I_f^j\cup \{i\}$
+				\item $I_f^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
+				\item $I_f=I_f \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
 			\end{itemize}			 
-		\item else (remove): $I_1^j=I_1^j\setminus \{i\}$
+		\item else (remove): $I_f^j=I_f^j\setminus \{i\}$
 	\end{itemize}
@@
 	\smallskip
-	At every moment, coherence between $I_1$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_1^j=|I_1^j|$, $I_2=\{j |p_1^j>0 \}$, $p_2= |I_2|$, .
+	At every moment, coherence between $I_f$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_f^j=|I_f^j|$, $I_r=\{j |p_f^j>0 \}$, $p_r= |I_r|$ .
 
@@
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}P(\tilde{S}|\boldsymbol{X}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}P(S_l|\boldsymbol{X}))} \\
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}P(\tilde{S}|\boldsymbol{X}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}P(S_l|\boldsymbol{X}))}
 	\end{eqnarray}
@@
 Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals.
-For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
-The $\boldsymbol{\alpha_j}$ and $\beta_j$ are generated according to the same Poisson law but with a random sign.
-$S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.
+For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
+The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign.
+$S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_f^j)=5$.
 	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.
@@
 			\begin{itemize}
-				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_2\cap \hat{I}_2|$ 
-				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_2|-TL$
-				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_2|-TL$
-				\item $\Delta p_2$ : the gap between the number of sub-regression in both model : $\Delta p_2=|I_2|-|\hat{I}_2|$. The sign defines if $\hat{S}$ is too complex or too simple
-				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_2}p_1^j-\sum_{j \in \hat{p}_2}\hat{p}_1^j$
+				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_r\cap \hat{I}_r|$ 
+				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_r|-TL$
+				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_r|-TL$
+				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple
+				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
 			\end{itemize}
@@
 In table \ref{compZvrai} we compare found structures in different contexts with both Uniform (U) and Hierarchical Uniform (HU) a priori law on $P(S)$. 
-We see that (HU) hypothesis gives sparser models even if $\max (\hat{p}_1^j)=5$.
-$p=40$ so the maximum value for $\hat{p_2}$ is $20$ in (HU) case and we see that this max is not reached, meaning that the stronger penalty implied by (HU) really is efficient.
+We see that (HU) hypothesis gives sparser models even with $\max (\hat{p}_f^j)=5$.
+Moreover, this constraint is not active because observed comlexity are smaller (this constraint only serves to accelerate the walk by reducing the dimension of $\mathcal{S}$ because each configuration was computed a hundred times), meaning that the stronger penalty implied by (HU) really is efficient.
 The datasets used for (U) and (HU) are the same to keep the comparison  meaningful.
@@
 \hline
-$n$ & $p_2$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
+$n$ & $p_r$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_r$ & $\Delta compl$ \\
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \end{tabular} 
-\caption{Results of the Markov chain with  constraint $\hat{p}_1\leq 5$. Mean observed and standard deviation (sd). } \label{compZvrai}
+\caption{Results of the Markov chain with  constraint $\hat{p}_f\leq 5$. Mean observed and standard deviation (sd). } \label{compZvrai}
 \end{table}
@@
 \hline 
-$n$ & $p_2$& $P(S)$ & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$& $P(S)$ & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \hline 
-$n$ & $p_2$&  $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.
+$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
 \textsc{CorReg} logically wins.
@@
 \hline 
-$n$ & $p_2$&  $P(S)$&indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$&indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \hline 
-$n$ & $p_2$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
@@
 \clearpage
-	\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_2}$ (worst case for us)}	 \label{tableMSEsimgauche}
-We now try the method with a response depending only on variables in $\boldsymbol{X_2}$. The datasets used here were still those from \ref{compZvrai}.
-Depending only on $\boldsymbol{X_2}$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
+	\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_r}$ (worst case for us)}	 \label{tableMSEsimgauche}
+We now try the method with a response depending only on variables in $\boldsymbol{X_r}$. The datasets used here were still those from \ref{compZvrai}.
+Depending only on $\boldsymbol{X_r}$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
 
@@
 \hline 
-$n$ & $p_2$&$P(S)$ &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&$P(S)$ &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_2}$.
+\caption{OLS and OLS combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_r}$.
 Sometimes $\hat{S}$ gives better results than $S$ because $S$ is penalized by the fact that it relies only on covariates not in the true model.
@@
 \hline 
-$n$ & $p_2$&  $P(S)$ &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_2}$. }\label{YX2linLASSO}
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_r}$. }\label{YX2linLASSO}
 \end{table}
@@
 \hline 
-$n$ & $p_2$&  $P(S)$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \hline 
-$n$ & $p_2$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
@@
 \end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X_2}$.}\label{YX2linstep}
+\caption{Stepwise  combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X_r}$.}\label{YX2linstep}
 \end{table}
@@
  We have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
-Variables in $\boldsymbol{X_2}$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $\boldsymbol{X_1}$ .
+Variables in $\boldsymbol{X_r}$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $\boldsymbol{X_f}$ .
 Dependencies are still real but non linear.
@@
 \hline
-$n$ & $p_2$ & $P(S)$&Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
+$n$ & $p_r$ & $P(S)$&Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
 \hline
@@
 \hline 
-$n$ & $p_2$& $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$& $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
 \hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
@@
 These first results are easily understandable and meet metallurgists expertise.
-			The algorithm gives a structure of $p_2=76$ subregressions with a mean of $\bar{p_1}=5.17$ regressors.
-In $\boldsymbol{X_1}$ the number of $\rho>0.7$ is $\textbf{79.33\%}$ smaller than in $\boldsymbol{X}$.
+			The algorithm gives a structure of $p_r=76$ subregressions with a mean of $\bar{\boldsymbol{p}}_f=5.17$ regressors.
+In $\boldsymbol{X}_f$ the number of $\rho>0.7$ is $\textbf{79.33\%}$ smaller than in $\boldsymbol{X}$.
 
@@
 			\includegraphics[width=150px]{figures/correlbeforeafter.png} 
-			\caption{Correlations between the covariates in $\boldsymbol{X}$ (upper) and $\boldsymbol{\hat{X}_1}$ (lower).}
+			\caption{Correlations between the covariates in $\boldsymbol{X}$ (upper) and $\hat{\boldsymbol{X}}_f$ (lower).}
 	\end{minipage} \hfill
@@
 \textsc{CorReg} allows to modify this neighbourhood to better fit users constraints.
-Relaxation (column-wise and row-wise) is optional but gives more stability to the number of feasible candidates at each step and allows to modify several parts of $I_1$ in only one step when needed.
+Relaxation (column-wise and row-wise) is optional but gives more stability to the number of feasible candidates at each step and allows to modify several parts of $I_f$ in only one step when needed.
 Hence it improves efficiency by a significant reinforcement of the irreductibility of the Markov chain.
@@
 
-	The hierchical uniform hypothesis made above for $P(S)$ implies $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ so candidates may be rejected to satisfy this hypothesis.
-Stronger constraints on $p_2$ and/or $p_1$ can be given in \textsc{CorReg} if relevant.
+	The hierchical uniform hypothesis made above for $P(S)$ implies $p_r<\frac{p}{2}$ and $p_f^j<\frac{p}{2}$ so candidates may be rejected to satisfy this hypothesis.
+Stronger constraints on $p_r$ and/or $p_f^j$ can be given in \textsc{CorReg} if relevant.
 
 If the algorithm did not have time to converge (stationnarity), it can be continued with a few step for which the neighboorhood would only contain smaller candidates (in terms of complexity).
-It is equivalent to ask for each element in $I_1$ if the criterion $P(S|\boldsymbol{X})$ would be better without it.
+It is equivalent to ask for each element in $I_f$ if the criterion $P(S|\boldsymbol{X})$ would be better without it.
 Thus it can be seen as a final cleaning step.

### 2014-04-14 pkg/CorReg/R/compare_beta.R
changinf default value
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@528 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -6,7 +6,8 @@
 
 compare_beta<-function(truebeta=truebeta,hatbeta=hatbeta,taux=FALSE){
    resloc=compare_zero(trueA=truebeta,Aalgo=hatbeta,taux=FALSE)
-   resloc=list(resloc,compare_sign(trueA=truebeta,Aalgo=hatbeta))
+   resloc=c(resloc,compare_sign(trueA=truebeta,Aalgo=hatbeta))
    resloc$biais=sum((truebeta-hatbeta)^2)
    return(resloc)
+   
 }

### 2014-04-14 pkg/CorReg/R/correg.R
changinf default value
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@528 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -44,7 +44,7 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = TRUE,
                 select = "lar",
                 criterion = c("MSE", "BIC"),
                 X_test = NULL, Y_test = NULL, intercept = TRUE, 
-                K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=TRUE,#final=FALSE,
+                K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=FALSE,#final=FALSE,
                 nbalter=10,deltamin=0.01,alpha=NULL,g=5) 
 {
   res = list()

### 2014-04-14 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@529 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable.
-This pretreatment by variable selection gives then the possibility to perform standard linear estimation methods including variables selection procedures for instance.
-The structure of correlations is found with an MCMC algorithm.
+It corresponds to a kind of variable selection preliminary step which has then to be followed by standard linear estimation methods including variables selection procedures for instance.
+The structure of correlations is found with an MCMC algorithm aiming at optimizing a specific BIC criterion.
  An R package (\textsc{CorReg}) available on the CRAN implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry where correlated variables are frequent.
@@
 %la régression et ses problèmes
-
-When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression.
-It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn.
+Linear regression is a very standard and efficient method providing a predictive model with a good interpretability even  for non-statistician.
 Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from industry and astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
@@
 \begin{equation}
-		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
+		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
 	\end{equation}
@@
 	\end{equation}
-	With variance matrix
+	with variance matrix
 	\begin{equation}
@@
 \subsection{Our proposal: modelisation of the correlations}
-Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
-In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of a matrix $\boldsymbol{X}$.
+Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ covariates.
+In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of $\boldsymbol{X}$.
 
@@
  	\\
-Let $\boldsymbol{Y} \in \mathcal{R}^n$ be a response variable we want to explain with $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$:
+Using the partition $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$ we can rewrite \ref{regressionsimple}:
 	\begin{equation}
@@
 		where $\boldsymbol{\beta}=(\boldsymbol{\beta}_f,\boldsymbol{\beta}_r) \in  \mathcal{R}^p$ is the vector of the regression coefficients associated respectively to $\boldsymbol{X}_f$ and $\boldsymbol{I}_n$ the identity matrix.
-We note that (\ref{MainR}) and (\ref{SR}) also give by simple integration on $\boldsymbol{X}_r$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X}_f$}:
+We note that (\ref{SR}) and (\ref{MainR}) give also by simple integration on $\boldsymbol{X}_r$ a marginal regression model on $\boldsymbol{Y}$ {\it depending only on uncorrelated covariates $\boldsymbol{X}_f$}:
 \begin{eqnarray}
 	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
-	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_2^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_r^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
 \end{eqnarray}
-We note that it simply is a linear regression on some of the original covariates so we only made a pretreatment on the dataset by setting some coefficients to $0$ because of correlations.
+We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by setting some coefficients to $0$ because of correlations.
 These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates.
@@
 
-\subsection{Strategy of use: pretreatment before classical selection methods}\label{interpretation}
+\subsection{Strategy of use: pretreatment before classical estimation/selection methods}\label{interpretation}
 As a pretreatment, it allows usage of any method in a second time to estimate $\boldsymbol{\beta}^*$, even with variable selection methods like LASSO or even stepwise \cite{seber2012linear}.
@@
 \hline
-\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
+\multicolumn{3}{|c}{Configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
 \hline

### 2014-04-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@530 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@531 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from industry and astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
-Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators. We note the linear regression model:
+However, linear regression has to face the problem of the variance of the estimators.
+%Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators. 
+We note the linear regression model:
 \begin{equation}
@@
 
-	With the rise of informatics, datasets contain more and more covariates and thus dimension reduction becomes a necessity.
-Moreover, increasing the number of covariates also increases the chance to have correlated ones.
+	With the rise of informatics, datasets contain more and more covariates,% and thus dimension reduction becomes a necessity. Moreover, 
+	increasing the chance to have correlated ones.
 It will be illustrated with examples from steel industry in the last section of this paper (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
 	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
-For instance, in an industrial context interpretation could be favored to improve the process instead of only predict defects, so robust estimation is needed.
+For instance, in an industrial context interpretation could be favoured to improve the process instead of only predict defects, so robust estimation is needed.
 		~\\	~\\
@@
 	\end{equation}
-	But Ridge regression do not aim at selecting covariates because coefficients tends to 0 but don't reach 0.
-	So it gives difficult interpretations for large values of $p$.	
-	Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.
+	But this penalty is not guided by the correlations.
+It is the same for each covariates and will be too large for independent covariates and/or too small for correlated ones.
+So the efficiency is of such a method is limited.
+	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large values of $p$.
+Real datasets often imply many irrelevant variables so variable selection would be favoured when possible.
 

### 2014-04-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@532 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 It is the same for each covariates and will be too large for independent covariates and/or too small for correlated ones.
-So the efficiency is of such a method is limited.
+So the efficiency of such a method is limited.
 	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large values of $p$.
@@
 		\end{equation}	
-	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
+	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
 It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates.
-And it  really selects covariates with coefficients set exactly to 0.
-
+Like the ridge regression, the penalty does not distinguish correlated and independent covariates, but the selection reduces the dimension and may select less correlated covariates.
 %	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
 	 Some recent variants of the \textsc{LASSO} do exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
+But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates.
 	 \\
 
-	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}. 
+	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}: 
 %	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
@@
 %\end{equation}		
-	Elastic net can be written:
+	%Elastic net can be written:
 	\begin{equation}
@@
 	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
-Hence interpretations are not the same and nothing explains why.
+%Hence, when comparing the two models, interpretations are not the same and nothing explicitly explains why.
 So it can be very confusing.
@@
 Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
-	 \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates and we see that none of the above methods take the correlations into account.
+	We see that none of the above methods takes explicitly the correlations into account.
 
@@
 
- Where some reduce the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates that helps to define a marginal model with uncorrelated covariates.
-Thus we have variable selection with explanation and improve interpretation and not only prediction.
- Our work is based on the assumption that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
- In fact we search the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
-This can be viewed as a pretreatment on the dataset allowing to use then other tools for dimension reduction and estimation without suffering from correlations.
+ Where some reduce the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
 Practically speaking we use a system of linear regressions between the covariates to model the structure of the correlations.
-This structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint law on the dataset $\boldsymbol{X}$.
+It helps to define the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
+Then we can define a marginal model with uncorrelated covariates. Thus we have variable selection with explanation and improve interpretation and not only prediction.
+ This can be viewed as a pretreatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
+
+  The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint law on the dataset $\boldsymbol{X}$.
 This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{CRAN}.

### 2014-04-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@533 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  	%plan
- 	This paper will first present the marginal regression model and its properties before describing in Section 3 the random walk used to find the structure.
+ 	This paper will first present the linear modelisation of the correlations and the by-product marginal regression model before describing in Section 3 the random walk used to find the structure.
  	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in Section \ref{conclusion}.
@@
 \subsection{A classical problem: correlations in regression}\label{sectionOLS}
-For a model defined by 
-	\begin{equation}
-		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X\beta} + \boldsymbol{\varepsilon}
-	\end{equation}
-	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ with $\boldsymbol{I_n}$ the $n$-sized identity matrix,
-	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
+For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
 	\begin{equation}
@@
 	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
-	Another problem is that matricial inversion requires $n\geq p$. 
+	Another problem is that matrix inversion requires $n\geq p$. 
 \subsection{Our proposal: modelisation of the correlations}
@@
 
-We propose to explicitly define a family of $p_r$ internal regressions between covariates with $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
+We make the hypothesis that correlations in $\boldsymbol{X}$ are generated by a family of $p_r$ internal regressions between covariates with $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
 We note $\boldsymbol{X}_r=\boldsymbol{X}^{I_r}$ the sub-matrix of the redundant covariates (explained by others) and  $\boldsymbol{X}_f$ the remaining submatrix of free covariates (independent explaining covariates).
@@
 
-We make the hypothesis of the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus $\boldsymbol{X}_f$ and $\boldsymbol{X}_f$ give a partition of $\boldsymbol{X}$, so $\boldsymbol{X}_f=\boldsymbol{X}\setminus \boldsymbol{X}_r$.
+Now we have an explicit structure of correlations and we know which covariates will cause ill-conditioning.
+The explicit form also gives the ability to understand phenomenon inside the dataset and to see why some covariates are not selected by methods like the LASSO when all $\boldsymbol{X}$ is provided (redundent variables).
+
+We make the hypothesis of the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus $\boldsymbol{X}_r$ and $\boldsymbol{X}_f$ give a partition of $\boldsymbol{X}$, so $\boldsymbol{X}_f=\boldsymbol{X}\setminus \boldsymbol{X}_r$.
+This hypothesis ensures that $S$ contains no cycle and is straightforward readable (no need to order $I_r$).
+This hypothesis is not so restrictive because cyclic structures have no sense and any non-cyclic structure can be associated with a structure that verifies the uncrossing constraint by just replacing covariates in $\boldsymbol{X}_r$ by their expression in $S$ when on the right of a sub-regression.
+
 The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
@@
 Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
+We see that the $S$ does not depend on $Y$ so it can be learnt independently, even with a larger dataset (if missing values in $Y$).
 

### 2014-04-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@534 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@536 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %bibliographie	
-	Because \textsc{OLS} is the minimum-variance linear unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
+	Because \textsc{OLS} is the minimum-variance linear unbiased estimator, other methods can be seen as a pretreatment on the dataset to improve $\boldsymbol{X}'\boldsymbol{X}$ conditioning (by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}).
 
@@
 
-	Ridge regression\cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:
+	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
+	Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:
 	\begin{equation}
@@
 	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large values of $p$.
-Real datasets often imply many irrelevant variables so variable selection would be favoured when possible.
-
+
+	Real datasets often imply many irrelevant variables so variable selection would be favoured when possible. Moreover, selection may keep only uncorrelated covariates and thus resolve the problem.	
 	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty to obtain zeros in $\hat{\boldsymbol{\beta}}$:
@@
 It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates.
-Like the ridge regression, the penalty does not distinguish correlated and independent covariates, but the selection reduces the dimension and may select less correlated covariates.
+But like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates.
 %	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
@@
 
-	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}: 
+	Elastic net \cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}: 
 %	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
@@
 	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
+Once again, nothing specifically aims to reduce the correlations.
 %Hence, when comparing the two models, interpretations are not the same and nothing explicitly explains why.
@@
 
-	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}).
+	Another way of improving the conditioning is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}) to reduce the dimension and also the correlations if correlated covariates are in the same clusters.
 	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables with grouped $\beta_j$ following a Gaussian Mixture distribution.
@@
 
- Where some reduce the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
-Practically speaking we use a system of linear regressions between the covariates to model the structure of the correlations.
+ Instead of reducing the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the correlations, we could use this knowledge to avoid the problem.
+Practically speaking we use a system of linear regressions for the joint distribution of $\boldsymbol{X}$ to model the structure of the correlations.
 It helps to define the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
-Then we can define a marginal model with uncorrelated covariates. Thus we have variable selection with explanation and improve interpretation and not only prediction.
+Then we can define a marginal model with independent covariates. Thus we have variable selection with explanation and improve interpretation and not only prediction.
  This can be viewed as a pretreatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
-
-  The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint law on the dataset $\boldsymbol{X}$.
+This pretreatment is specifically done to decorrelate the covariates, unlike methods described above.
+
+  The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint distribution on the dataset $\boldsymbol{X}$.
 This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{CRAN}.
@@
  	\\
-Using the partition $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$ we can rewrite \ref{regressionsimple}:
-	\begin{equation}
-			\boldsymbol{Y}_{|\boldsymbol{X}_f,\boldsymbol{X}_r,S}=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon_Y} \textrm{ with } \boldsymbol{\varepsilon_Y} \sim \mathcal{N}(\boldsymbol{0},\sigma^2_Y\boldsymbol{I}_n);	\label{MainR}
+Using the partition $\boldsymbol{X}=[\boldsymbol{X}_f,\boldsymbol{X}_r]$ we can rewrite (\ref{regressionsimple}):
+	\begin{equation}
+			\boldsymbol{Y}_{|\boldsymbol{X}_f,\boldsymbol{X}_r,S}=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon_Y} \label{MainR}
 		\end{equation}
@@
 
-We want to find the structure $S$ that maximizes $P(S|\boldsymbol{X})$ and we have:	
+We want to find the most probable structure $S$ knowing the dataset, so we search for the structure that maximizes $P(S|\boldsymbol{X})$ and we have:	
 	\begin{eqnarray}

### 2014-04-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@537 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \section{Model to select decorrelated covariates}
-\subsection{A classical problem: correlations in regression}\label{sectionOLS}
-For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
-	\begin{equation}
-		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
-	\end{equation}
-	with variance matrix
-	\begin{equation}
-		\operatorname{Var}(\boldsymbol{\hat{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
-	\end{equation}
-	and without any bias.
-	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
-	Another problem is that matrix inversion requires $n\geq p$. 
+%\subsection{A classical problem: correlations in regression}\label{sectionOLS}
+
 \subsection{Our proposal: modelisation of the correlations}
-Let now $\boldsymbol{X} \in \mathcal{R}^{n\times p}$ be a set of $p$ covariates.
-In the following, we note $\boldsymbol{X^j}$ the $j^{th}$ column of $\boldsymbol{X}$.
-
-We make the hypothesis that correlations in $\boldsymbol{X}$ are generated by a family of $p_r$ internal regressions between covariates with $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
-We note $\boldsymbol{X}_r=\boldsymbol{X}^{I_r}$ the sub-matrix of the redundant covariates (explained by others) and  $\boldsymbol{X}_f$ the remaining submatrix of free covariates (independent explaining covariates).
-We focus here on the expression of $P(\boldsymbol{X}_r|\boldsymbol{X}_f)$.
-
-We have $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$.
-Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each internal regression  and $p_f^j=|I_f^j|$.
-Here $|.|$ represents the cardinal of an ensemble.
-
+We make the hypothesis that $\boldsymbol{X}$ can be described by a partition $\boldsymbol{X}=(\boldsymbol{X}_f,\boldsymbol{X}_r) $ where variables in $\boldsymbol{X}_r$ are endogenous covariates resulting from linear sub-regressions based on $\boldsymbol{X}_f$, the submatrix of mutually independent exogenous covariates.
+Then $\boldsymbol{X}_r$ is the $n\times p_r$ submatrix of $0\leq p_r <p$ redundent covariates and $\boldsymbol{X}_f$ the $n\times (p-p_r)$ submatrix of the free (independent) covariates.
+In the following, we note $\boldsymbol{X}^j$ the $j^{th}$ column of $\boldsymbol{X}$ and $|.|$ represents the cardinal of an ensemble.
 The family of $p_r$ regressions within correlated covariates in $\boldsymbol{X}$ is noted:
 	\begin{equation}
-		\boldsymbol{X}_{r|\boldsymbol{X}_f,S} \textrm{ defined by }\forall j \in I_r: \boldsymbol{X}^j_{|\boldsymbol{X}_f,S}=\boldsymbol{X}_f\boldsymbol{\alpha}_j+\boldsymbol{\varepsilon}_j \textrm{ with } \boldsymbol{\varepsilon}_j \sim(\boldsymbol{0},\sigma^2_j\boldsymbol{I}_n) \label{SR}
-	\end{equation}
-		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_f^j$).
-
-Now we have an explicit structure of correlations and we know which covariates will cause ill-conditioning.
-The explicit form also gives the ability to understand phenomenon inside the dataset and to see why some covariates are not selected by methods like the LASSO when all $\boldsymbol{X}$ is provided (redundent variables).
-
-We make the hypothesis of the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates, thus $\boldsymbol{X}_r$ and $\boldsymbol{X}_f$ give a partition of $\boldsymbol{X}$, so $\boldsymbol{X}_f=\boldsymbol{X}\setminus \boldsymbol{X}_r$.
+		\boldsymbol{X}_{r|\boldsymbol{X}_f,S} \textrm{ defined by }\forall \boldsymbol{X}^j \subset \boldsymbol{X}_r: \boldsymbol{X}^j_{|\boldsymbol{X}_f,S}=\boldsymbol{X}_f\boldsymbol{\alpha}_j+\boldsymbol{\varepsilon}_j \textrm{ with } \boldsymbol{\varepsilon}_j \sim\mathcal{N}(\boldsymbol{0},\sigma^2_j\boldsymbol{I}_n) \label{SR}
+	\end{equation}
+		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_r)}$ are the vectors of the regression coefficients between the covariates.
+So we now have the expression of $P(\boldsymbol{X}_r|\boldsymbol{X}_f) $
+\\
+\\
+We note $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
+We also define $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$. 
+We see that $I_f$ defines the non-null coefficients in $\boldsymbol{\alpha}_j$ (each sub-regression can be very parsimonious).
+Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each sub-regression  and $p_f^j=|I_f^j|$.
+\\
+\\
+
+The partition of $\boldsymbol{X}$ implies the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates.
 This hypothesis ensures that $S$ contains no cycle and is straightforward readable (no need to order $I_r$).
-This hypothesis is not so restrictive because cyclic structures have no sense and any non-cyclic structure can be associated with a structure that verifies the uncrossing constraint by just replacing covariates in $\boldsymbol{X}_r$ by their expression in $S$ when on the right of a sub-regression.
-
-The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
+It is not so restrictive because cyclic structures have no sense and any non-cyclic structure can be associated with a structure that verifies the uncrossing constraint by just successively replacing endogenous covariates by their sub-regression when they are also exogenous in some other sub-regressions.
 
@@
 
-	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression).
-Thus we have one regression on our response variable ($P(\boldsymbol{Y}|\boldsymbol{X})$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$).
+	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a sub-regression).
+Thus we have one regression on our response variable ($P(\boldsymbol{Y}|\boldsymbol{X}))$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$).
 Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
-We see that the $S$ does not depend on $Y$ so it can be learnt independently, even with a larger dataset (if missing values in $Y$).
+We see that the $S$ does not depend on $\boldsymbol{Y}$ so it can be learnt independently, even with a larger dataset (if missing values in $\boldsymbol{Y}$).
 
@@
 \subsection{Strategy of use: pretreatment before classical estimation/selection methods}\label{interpretation}
+Now we have an explicit structure of correlations and we know which covariates will cause ill-conditioning.
+The explicit form also gives the ability to understand phenomenon inside the dataset and to see why some covariates are not selected by methods like the LASSO when all $\boldsymbol{X}$ is provided (redundent variables).
+
 As a pretreatment, it allows usage of any method in a second time to estimate $\boldsymbol{\beta}^*$, even with variable selection methods like LASSO or even stepwise \cite{seber2012linear}.
@@
 	\subsection{Illustration of the tradeoff conveyed by the pretreatment}	
+	Here we look at the OLS case.
+For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
+	\begin{equation}
+		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
+	\end{equation}
+	with variance matrix
+	\begin{equation}
+		\operatorname{Var}(\hat{\boldsymbol{\beta}}_{OLS})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
+	\end{equation}
+	and without any bias.
+	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
+	Another problem is that matrix inversion requires $n\geq p$.
+
 	 The Mean Squared Error (\textsc{MSE}) on $\hat{\boldsymbol{\beta}}$ is:
 	\begin{eqnarray}
-		\textsc{MSE}(\hat{\boldsymbol{\beta}}|\boldsymbol{\beta},\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}})) \\
-			&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
-			&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1})
+		\textsc{MSE}(\hat{\boldsymbol{\beta}}|\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}})) \\
+			\textsc{MSE}(\hat{\boldsymbol{\beta}}_{OLS}|\boldsymbol{X})&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
+			\textsc{MSE}(\hat{\boldsymbol{\beta}}^*_{OLS}|\boldsymbol{X})&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1})
 	\end{eqnarray}	 
@@
 Thus it must be a global criterion.
+
+The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
 

### 2014-04-16 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@538 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  Instead of reducing the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the correlations, we could use this knowledge to avoid the problem.
-Practically speaking we use a system of linear regressions for the joint distribution of $\boldsymbol{X}$ to model the structure of the correlations.
+More precisely, we model the correlations with a system of linear regressions for the joint distribution of $\boldsymbol{X}$ to model the structure of the correlations.
 It helps to define the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
@@
 In the following, we note $\boldsymbol{X}^j$ the $j^{th}$ column of $\boldsymbol{X}$ and $|.|$ represents the cardinal of an ensemble.
-The family of $p_r$ regressions within correlated covariates in $\boldsymbol{X}$ is noted:
+The structure $S$ of $p_r$ regressions within correlated covariates in $\boldsymbol{X}$ is described by:
 	\begin{equation}
@@
 	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
-	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_r^*) \textrm{ and } \hat{\boldsymbol{\beta}}_r^*=\boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_r^*) \textrm{ and } \boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
 \end{eqnarray}
@@
 		\textsc{MSE}(\hat{\boldsymbol{\beta}}|\boldsymbol{X})&=&\parallel \operatorname{Bias}\parallel_2^2+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}})) \\
-			\textsc{MSE}(\hat{\boldsymbol{\beta}}_{OLS}|\boldsymbol{X})&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) \textrm{ for OLS, and then for the marginal model:} \\
+			\textsc{MSE}(\hat{\boldsymbol{\beta}}_{OLS}|\boldsymbol{X})&=& 0 + \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}) %\textrm{ for OLS, and then for the marginal model:}
+			 \\
 			\textsc{MSE}(\hat{\boldsymbol{\beta}}^*_{OLS}|\boldsymbol{X})&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1})

### 2014-04-16 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@539 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 We make the hypothesis that $\boldsymbol{X}$ can be described by a partition $\boldsymbol{X}=(\boldsymbol{X}_f,\boldsymbol{X}_r) $ where variables in $\boldsymbol{X}_r$ are endogenous covariates resulting from linear sub-regressions based on $\boldsymbol{X}_f$, the submatrix of mutually independent exogenous covariates.
+So we model the correlations by $P(\boldsymbol{X}_r|\boldsymbol{X}_f) $ with $\boldsymbol{X}_f$ orthogonals.
  Then $\boldsymbol{X}_r$ is the $n\times p_r$ submatrix of $0\leq p_r <p$ redundent covariates and $\boldsymbol{X}_f$ the $n\times (p-p_r)$ submatrix of the free (independent) covariates.
@@
 		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_r)}$ are the vectors of the regression coefficients between the covariates. 
-So we now have the expression of $P(\boldsymbol{X}_r|\boldsymbol{X}_f) $
 \\
@@
   	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) without explicit expression of the correlations and thus a smaller interpretation potential.
-  	\\ Here we suppose independency between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\alpha}_j$ and $\sigma_j$.
+  	\\ Here we suppose independence between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\alpha}_j$ and $\sigma_j$.
 
@@
 \end{eqnarray}
-We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by setting some coefficients to $0$ because of correlations.
+We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by selecting
+
+setting some coefficients to $0$ because of correlations.
 These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates.

### 2014-04-16 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@540 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{eqnarray}
-We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by selecting
-
-setting some coefficients to $0$ because of correlations.
-These $0$ won't be interpreted as independence with the response variable but as redundant information due to correlations between the covariates.
+We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by selecting $\boldsymbol{X}_f$ because of the correlations given by $S$.
 It is a variable pre-selection independent of the response $\boldsymbol{Y}$.
-
 \subsection{Strategy of use: pretreatment before classical estimation/selection methods}\label{interpretation}
-Now we have an explicit structure of correlations and we know which covariates will cause ill-conditioning.
-The explicit form also gives the ability to understand phenomenon inside the dataset and to see why some covariates are not selected by methods like the LASSO when all $\boldsymbol{X}$ is provided (redundent variables).
 
@@
 
-After selection and estimation we will obtain a model with { \it two steps of variable selection}: the decorrelation step (coerced zeros associated to redundant information) and the classical selection step, with different meanings for obtained zeros in $\hat{\boldsymbol{\beta}}^*$.
- This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more.
-So we dodge the drawbacks of both grouping effect and variable selection.
-
-The explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset.
-Expert knowledge can even be added to the structure.
+After selection and estimation we will obtain a model with { \it two steps of variable selection}: the decorrelation step (coerced selection associated to redundant information defined in $S$) and the classical selection step, with different meanings for obtained zeros in $\hat{\boldsymbol{\beta}}^*_f$ (irrelevant covariates) and for $\hat{\boldsymbol{\beta}}^*_r=0$ (redundant information).
+ Thus we are able to distinguish the reasons of selection and consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.
+
+The explicit structure is parsimonious and simply consists in linear regressions and thus is easily understood by non statistician, allowing them to have a better knowledge of the phenomenon inside the dataset and to take better actions.
+Expert knowledge can even be added to the structure, physical models for example.
 
 Moreover, the uncrossing constraint ($I_f\cap I_r=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
+
+			There is no theoretical guarantee that our model is better.
+It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis.
+We play on the bias-variance tradeoff.
+
+	\subsection{Illustration of the tradeoff conveyed by the pretreatment}	
+	As an example, we look at the OLS estimator on $\boldsymbol{X}$ with and without the pretreatment that is $\boldsymbol{X}_f$ selection.
+For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
+	\begin{equation}
+		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
+	\end{equation}
+	with variance matrix
+	\begin{equation}
+		\operatorname{Var}(\hat{\boldsymbol{\beta}}_{OLS})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
+	\end{equation}
+	and without any bias.
+	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
+	Another problem is that matrix inversion requires $n\geq p$.
 
@@
 So we play on the bias-variance tradeoff, reducing the variance by adding a bias.
-			%There is no theoretical guarantee that our model is better.
-It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis.
-Therefore we made some simulations to compare both methods (see the end of this paper).
-			 \\
-	\subsection{Illustration of the tradeoff conveyed by the pretreatment}	
-	Here we look at the OLS case.
-For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
-	\begin{equation}
-		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
-	\end{equation}
-	with variance matrix
-	\begin{equation}
-		\operatorname{Var}(\hat{\boldsymbol{\beta}}_{OLS})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
-	\end{equation}
-	and without any bias.
-	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
-	Another problem is that matrix inversion requires $n\geq p$.
 
@@
 \end{figure} 
-It is clear in Figure \ref{MQE1} that the marginal model is more robust than \textsc{OLS} on $X$.
+It is clear in Figure \ref{MQE1} that the marginal model is more robust than \textsc{OLS} on $\boldsymbol{X}$.
 And when sub-regression get weaker ($1-\mathcal{R}^2$ tends to 1) it remains stable until extreme values (sub-regression nearly fully explained by the noise).
@@
 It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
+Further results are provided in sections \ref{sectionsimul} and \ref{sectionrealcase}.
 

### 2014-04-16 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@541 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Cross-validation is very time-consuming and thus not friendly with combinatory problematics.
-Moreover, we need a criterion compatible with structures of different sizes (varying $p_2$) and not related with $Y$ because the structure is inherent to $\boldsymbol{X}$ only.
+Moreover, we need a criterion compatible with structures of different sizes (varying $p_r$) and not related with $\boldsymbol{Y}$ because the structure is inherent to $\boldsymbol{X}$ only.
 Thus it must be a global criterion.
-
-The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
 
@@
 	\end{eqnarray}
+
+	The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
 

### 2014-04-16 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@543 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\end{eqnarray}
-
+So we will try to maximize $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)P(S)$.
 	The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
@@
 	\subsection{Modeling the uncorrelated covariates: a full generative approach on $P(\boldsymbol{X})$}
-	To be able to compare structures with probabilistic likelihood-based criterions, we need a full generative model on $\boldsymbol{X}$.
+	To be able to compare structures with $P(S|\boldsymbol{X})$, we need a full generative model on $\boldsymbol{X}$.
 Sub-regressions give $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S) $ but $P(\boldsymbol{X}_f|S)$ is still undefined.
@@
 		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we obtain better results when (HU) this constraint is verified.
+Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)$ so it is equivalent to a maximization of the $BIC$.
+With the Hierarchical Uniform hypothesis we maximize $\psi(\boldsymbol{X},S)\approx BIC + P(S)$.
 
@@
 	\begin{eqnarray}
-			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}P(\tilde{S}|\boldsymbol{X}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}P(S_l|\boldsymbol{X}))}
+			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2}\psi(\boldsymbol{X},\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}\psi(\boldsymbol{X},S_l))}
 	\end{eqnarray}
@@
 The initial structure can be based on a first warming algorithm taking the correlations into account.
-coefficients are randomly placed into $I_1$, weighted by the absolute value of the correlations.
+Coefficients are randomly placed into $I_f$, weighted by the absolute value of the correlations.
 We do so in the followings.
-Then this structure could be for example reduced by the hadamard product with the binary matrix obtained by Graphical Lasso\cite{friedman2008sparse} that makes selection in the precision matrix but it is time consuming.
+Then this structure could be for example reduced by the Hadamard product with the binary matrix obtained by Graphical Lasso\cite{friedman2008sparse} that makes selection in the precision matrix but it is time consuming.
 
@@
 It also helps to face local extrema.
-In the followings, the chain was launched with twenty initialisations.
+In the followings, the chain wzs launched with twenty initialisations each time.
 

### 2014-04-17 www/article13.tex
up title
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@544 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
new title
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@545 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
-\title{Model-based pretreatment for regression with correlated variables.}
+\title{Model-based variable selection for regression with highly correlated variables.}
 
@@
 {\Large
-	{\sc Model-based pretreatment for regression \\ with correlated variables.}
+	{\sc Model-based pretreatment for regression \\ with highly correlated variables.}
 }
@@
 
-	With the rise of informatics, datasets contain more and more covariates,% and thus dimension reduction becomes a necessity. Moreover, 
+	With the rise of informatics, datasets contain more and more covariates, % and thus dimension reduction becomes a necessity. Moreover, 
 	increasing the chance to have correlated ones.
-It will be illustrated with examples from steel industry in the last section of this paper (many parameters of the whole process without any a priori) highly correlated (physical laws, process rules, etc).
+It will be illustrated with examples from steel industry in the last section of this paper (many parameters of the whole process without any a priori and highly correlated: physical laws, process rules, etc).
 	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
@@
 %bibliographie	
-	Because \textsc{OLS} is the minimum-variance linear unbiased estimator, other methods can be seen as a pretreatment on the dataset to improve $\boldsymbol{X}'\boldsymbol{X}$ conditioning (by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}).
+	Because \textsc{OLS} is the minimum-variance linear unbiased estimator, some other methods can be seen as a modification of the dataset to improve the robustess of the estimator (by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}).
 
@@
 
-	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
+	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance tradeoff and get better prediction.
 	Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:

### 2014-04-18 www/article13.tex
up title
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@546 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-%\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
-\title{Model-based variable selection for regression with highly correlated variables.}
+%%\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
+%\title{Model-based variable selection for regression with highly correlated variables.}
 
@@
 {\Large
-	{\sc Model-based pretreatment for regression \\ with highly correlated variables.}
+	{\sc Model-based variable selection for regression \\ with highly correlated variables.}
 }

### 2014-04-24 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@549 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 So we will try to maximize $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)P(S)$.
-	The discrete structure $S$ is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$, the residuals ($\boldsymbol{\varepsilon}_j$) would not stay Gaussian (see the appendices).
+	The discrete structure $S$ is identifiable under some hypothesis detailed in appendix, that is we can't permute some sub-regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$.
 
@@
 Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions?
-The answer is no.
-Thus $S$ is identifiable.
-
-	If there are exact regressions ($\sigma^2_j=0$) in (\ref{SR}), the structure won't be identifiable.
+
+		If there are exact sub-regressions ($\sigma^2_j=0$), the structure won't be identifiable.
 But it only means that several structure will have the same likelihood and they will have the same interpretation.
@@
 	In the followings we suppose $\sigma^2_j\neq 0$.
+
+	We introduce the matricial notation
+		$\boldsymbol{X}_r=\boldsymbol{X}_f\boldsymbol{\alpha} + \boldsymbol{\varepsilon}$
+		we note $\Theta$ the parameter of the model. $\boldsymbol{\alpha}$ is the $(p-p_r)\times p_r$ matrix whose columns are the $\boldsymbol{\alpha}_j$.
+		\\ We want to know if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|\tilde{S},\tilde{\Theta})$ does imply $(S,\Theta)=(\tilde{S},\tilde{\Theta})$.
+
+	\subsubsection{Identifiability of $\Theta$}		
+		We know by identifiability of the Gaussian mixtures and linear regression ($\boldsymbol{X}_f'\boldsymbol{X}_f$ is of full rank) that if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|S,\tilde{\Theta})$ then $\Theta=\tilde{\Theta}$.
+But the identifiability of the discrete parameter $S$ is still pending.
+
+	\subsubsection{Sufficient conditions for Identifiability of $S$}
+
+	In the following, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
+	\begin{eqnarray}
+			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
+			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j  \label{densitydroite}\\
+			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
+						\mathcal{GM}(\bigotimes_{\substack{i \in I_f^j  }}\boldsymbol{\pi}_i \ ; 
+			                         \bigoplus_{\substack{i \in I_f^j  }} \alpha_{i,j} \boldsymbol{\mu}_i \ ; 
+			                         \boldsymbol{\sigma}_j^2+\bigoplus_{\substack{i \in I_f^j  }}\alpha_{i,j}^2 \boldsymbol{\sigma}_i^2 )\label{densitygauche}
+		\end{eqnarray}
+		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
+We call this difference in component number "identifiability" in the sense that we try to find the model with the fewest component in $X_f$ (we use the BIC so when comparing two models with the same likelihood, the one with fewer parameters will win).
+So in this case we have a "better" model (the simplest one) in the group of equivalent-meaning models (permuting variables in some sub-regression).
+	So a sufficient condition is $\forall j \in I_r: k_j\geq 2$ with at least two distinct couples $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$
+
+	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$,then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained by others (\cite{leamer1978least},\cite{cameron2005microeconometrics})
+
+	Remark : the uncrossing constraint ($I_2\cap I_1 = \emptyset$) significantly reduces the number of feasible models and thus increases identifiability (because it also reduces the number of models equivalent in interpretation).
+
+	If there are exact sub-regressions, classical methods will fail (singular matrix) and the structure won't be identifiable.
+But it only means that several structure will have the same likelihood and they will have the same interpretation.
+So it's not a problem.
+Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
+CorReg (Our R package) prints a warning to point out exact sub-regressions when found.
+
+%	 As a set of Gaussian mixtures and linear combination of Gaussian mixtures, $\boldsymbol{X}$ follows a multivariate Gaussian mixture model with $K$ components:
+%		\begin{eqnarray}
+%			f(\boldsymbol{X}|S,\theta)=\sum_{k=1}^Kp_k \Phi(\boldsymbol{X}_r|\boldsymbol{X}_f\boldsymbol{\alpha},S,\theta)\Phi(\boldsymbol{X}_f|S,\theta)
+%		\end{eqnarray}	
+%
 

### 2014-04-25 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@551 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
-%\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
+\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
 	In the followings we suppose $\sigma^2_j\neq 0$.
@@
 
-	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$,then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained by others (\cite{leamer1978least},\cite{cameron2005microeconometrics})
+	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$, then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained in books and papers (\cite{leamer1978least},\cite{cameron2005microeconometrics}).
 
 	Remark : the uncrossing constraint ($I_2\cap I_1 = \emptyset$) significantly reduces the number of feasible models and thus increases identifiability (because it also reduces the number of models equivalent in interpretation).
-
-	If there are exact sub-regressions, classical methods will fail (singular matrix) and the structure won't be identifiable.
-But it only means that several structure will have the same likelihood and they will have the same interpretation.
-So it's not a problem.
-Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete.
-CorReg (Our R package) prints a warning to point out exact sub-regressions when found.
 

### 2014-04-27 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@557 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable.
-It corresponds to a kind of variable selection preliminary step which has then to be followed by standard linear estimation methods including variables selection procedures for instance.
+It corresponds to a kind of variable selection preliminary step which has then to be followed by standard linear estimation methods including classical variables selection procedures for instance.
 The structure of correlations is found with an MCMC algorithm aiming at optimizing a specific BIC criterion.
@@
 %la régression et ses problèmes
-Linear regression is a very standard and efficient method providing a predictive model with a good interpretability even  for non-statistician.
-Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from industry and astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
-However, linear regression has to face the problem of the variance of the estimators.
+Linear regression is a very standard and efficient method providing a predictive model with a good interpretability even  for non-statisticians.
+Therefore, it is used in nearly all the fields where statistics are made \cite{montgomery2012introduction},  industry (illustrated in the present paper), astronomy \cite{isobe1990linear}, sociology \cite{longford2012revision} \dots
+%However, linear regression has to face the problem of the variance of the estimators.
 %Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators. 
+%We note the linear regression model:
+%\begin{equation}
+%		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
+%	\end{equation}
+%	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I}_n)$ the noise of the regression, with $\boldsymbol{I}_n$ the $n$-sized identity matrix and $\sigma_Y >0$.
+The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{OLS}).
+%As shown in section \ref{sectionOLS},
+%	Estimation of $\boldsymbol{\beta}$ requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
+%Conditionning of $\boldsymbol{X}'\boldsymbol{X}$ get worse based on two aspects : the dimension $p$ (number of covariates) of the model (the more covariates you have the greater variance you get)
+%	 and the correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
+
+	With the rise of informatics, datasets contain more and more covariates, % and thus dimension reduction becomes a necessity. Moreover, 
+	increasing the chance to have correlated ones.
+It will be illustrated with examples on real datasets in the last section of this paper (steel industry: many parameters of the whole process without any a priori and highly correlated because of physical laws, process rules, etc).
+	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
+For instance, interpretation could be favoured to improve a situation instead of only predict problems, so robust estimation is a real stake.
+		~\\	~\\
+
+%bibliographie	
+	The Ordinary Least Squares (\textsc{OLS}) estimator is the minimum-variance linear unbiased estimator but it suffers from ill-conditioned matrices inversion when the covariates are highly correlated (section \ref{sectionOLS}).
+Therefore, some other methods can be seen as a modification of the dataset to improve the robustess of the estimator (by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}).
+
+%In the following we note classical norms: $\parallel\boldsymbol{\beta}\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\boldsymbol{\beta} \parallel_1=\sum_{i=1}^p|\beta_i| $.
+
+	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance tradeoff and get better prediction.
+	Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty.
+	But this penalty is not guided by the correlations.
+It is the same for each covariate and will be too large for independent covariates and/or too small for correlated ones.
+So the efficiency of such a method is limited.
+	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large number of covariates.
+
+	Real datasets often imply many irrelevant variables so variable selection would be favoured when possible. Moreover, selection may keep only uncorrelated covariates and thus resolve the problem.	
+	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a parametric $L_1$ penalty to shrink some coefficients exactly to zero.
+%	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
+It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates.
+But like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates.
+%	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
+	% Some recent variants of the \textsc{LASSO} do exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
+But
+	 \textsc{LASSO} is also known to face consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates.
+	 \\
+
+	Elastic net \cite{zou2005regularization} is a method developed to be a compromise between Ridge regression and the \textsc{LASSO} by a linear combination of $L_1$ and $L_2$ penalties.
+%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
+%	\begin{equation}
+%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
+%\end{equation}		
+	%Elastic net can be written:
+%	\begin{equation}
+%		\boldsymbol{\hat{\beta}}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta} \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\boldsymbol{\beta}\parallel_1+\alpha\parallel\boldsymbol{\beta}\parallel_2^2\leq t \textrm{ for some } t
+%	\end{equation}
+%	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
+	%It seems to give good predictions. 
+	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain similar predictions with a more parsimonious model.
+Once again, nothing specifically aims to reduce the correlations.
+%Hence, when comparing the two models, interpretations are not the same and nothing explicitly explains why.
+So it can be very confusing.
+	\\
+
+	Another way of improving the conditioning is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}) to reduce the dimension and also the correlations if correlated covariates are in the same clusters.
+	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the regression coefficients no longer as fixed effect parameters but as unobserved independent random variables with grouped coefficients following a Gaussian Mixture distribution.
+The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than the number of covariates.
+In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on the coefficients estimator.
+
+	But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net.
+Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
+	We see that none of the above methods takes explicitly the correlations into account.
+
+	~\\	~\\
+% Principe de la méthode
+
+ Instead of reducing the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the correlations, we could use this knowledge to avoid the problem.
+More precisely, we model the correlations with a system of linear regressions for the joint distribution of the covariates to model the structure of the correlations.
+It helps to define the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
+Then we can define a marginal model with independent covariates. 
+Thus we have an explained variable selection guided by the correlations and it improves interpretation, not only prediction.
+ This can be viewed as a pretreatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
+This pretreatment is specifically done to decorrelate the covariates, unlike methods described above.
+The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint distribution on the covariates.
+This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{CRAN}.
+%We only consider strong correlations (i.e.
+: problematic ones) thus we keep most of the information contained in the dataset.
+ %We will in a second time be able to use the remaining part of the information (sequential approach).
+
+ 	%plan
+ 	This paper will first present the linear modelisation of the correlations and the by-product marginal regression model before describing in Section 3 the random walk used to find the structure.
+ 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in Section \ref{conclusion}.
+
+\section{Model to select decorrelated covariates}
+\subsection{A classical problem: correlations in regression}\label{sectionOLS}
 We note the linear regression model:
@@
 
-	With the rise of informatics, datasets contain more and more covariates, % and thus dimension reduction becomes a necessity. Moreover, 
-	increasing the chance to have correlated ones.
-It will be illustrated with examples from steel industry in the last section of this paper (many parameters of the whole process without any a priori and highly correlated: physical laws, process rules, etc).
-	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
-For instance, in an industrial context interpretation could be favoured to improve the process instead of only predict defects, so robust estimation is needed.
-		~\\	~\\
-
-%bibliographie	
-	Because \textsc{OLS} is the minimum-variance linear unbiased estimator, some other methods can be seen as a modification of the dataset to improve the robustess of the estimator (by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}).
-
-In the following we note classical norms: $\parallel\boldsymbol{\beta}\parallel_2^2=\sum_{i=1}^p(\beta_i)^2$ and $\parallel\boldsymbol{\beta} \parallel_1=\sum_{i=1}^p|\beta_i| $.
-
-	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance tradeoff and get better prediction.
-	Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty:
-	\begin{equation}
-		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel \boldsymbol{\beta} \parallel_2^2\leq \lambda \textrm{ with } \lambda>0
-	\end{equation}
-	But this penalty is not guided by the correlations.
-It is the same for each covariates and will be too large for independent covariates and/or too small for correlated ones.
-So the efficiency of such a method is limited.
-	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large values of $p$.
-
-	Real datasets often imply many irrelevant variables so variable selection would be favoured when possible. Moreover, selection may keep only uncorrelated covariates and thus resolve the problem.	
-	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty to obtain zeros in $\hat{\boldsymbol{\beta}}$:
-		\begin{equation}
-		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda \textrm{ with } \lambda>0
-		\end{equation}	
-	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive.
-It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates.
-But like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates.
-%	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
-	 Some recent variants of the \textsc{LASSO} do exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
-But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates.
-	 \\
-
-	Elastic net \cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}: 
-%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
-%	\begin{equation}
-%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
-%\end{equation}		
-	%Elastic net can be written:
-	\begin{equation}
-		\boldsymbol{\hat{\beta}}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta} \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\boldsymbol{\beta}\parallel_1+\alpha\parallel\boldsymbol{\beta}\parallel_2^2\leq t \textrm{ for some } t
-	\end{equation}
-	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
-	%It seems to give good predictions. 
-	But it is based on the grouping effect so correlated covariates get similar coefficients and are selected together whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model.
-Once again, nothing specifically aims to reduce the correlations.
-%Hence, when comparing the two models, interpretations are not the same and nothing explicitly explains why.
-So it can be very confusing.
-	\\
-
-	Another way of improving the conditioning is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}) to reduce the dimension and also the correlations if correlated covariates are in the same clusters.
-	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables with grouped $\beta_j$ following a Gaussian Mixture distribution.
-The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$.
-In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\boldsymbol{\hat{\beta}}$.
-
-	But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net.
-Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
-	We see that none of the above methods takes explicitly the correlations into account.
-
-	~\\	~\\
-% Principe de la méthode
-
- Instead of reducing the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the correlations, we could use this knowledge to avoid the problem.
-More precisely, we model the correlations with a system of linear regressions for the joint distribution of $\boldsymbol{X}$ to model the structure of the correlations.
-It helps to define the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
-Then we can define a marginal model with independent covariates. Thus we have variable selection with explanation and improve interpretation and not only prediction.
- This can be viewed as a pretreatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
-This pretreatment is specifically done to decorrelate the covariates, unlike methods described above.
-
-  The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint distribution on the dataset $\boldsymbol{X}$.
-This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{CRAN}.
-%We only consider strong correlations (i.e.
-: problematic ones) thus we keep most of the information contained in the dataset.
- %We will in a second time be able to use the remaining part of the information (sequential approach).
-
- 	%plan
- 	This paper will first present the linear modelisation of the correlations and the by-product marginal regression model before describing in Section 3 the random walk used to find the structure.
- 	We will then look at some numerical results on simulated (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) before concluding and giving some perspectives in Section \ref{conclusion}.
-
-\section{Model to select decorrelated covariates}
-%\subsection{A classical problem: correlations in regression}\label{sectionOLS}
-
 \subsection{Our proposal: modelisation of the correlations}
-We make the hypothesis that $\boldsymbol{X}$ can be described by a partition $\boldsymbol{X}=(\boldsymbol{X}_f,\boldsymbol{X}_r) $ where variables in $\boldsymbol{X}_r$ are endogenous covariates resulting from linear sub-regressions based on $\boldsymbol{X}_f$, the submatrix of mutually independent exogenous covariates.
+We make the hypothesis that $\boldsymbol{X}$ can be described by a partition $\boldsymbol{X}=(\boldsymbol{X}_f,\boldsymbol{X}_r) $ given by an explicit structure $S$ where variables in $\boldsymbol{X}_r$ are endogenous covariates resulting from linear sub-regressions based on $\boldsymbol{X}_f$, the submatrix of mutually independent exogenous covariates.
 So we model the correlations by $P(\boldsymbol{X}_r|\boldsymbol{X}_f) $ with $\boldsymbol{X}_f$ orthogonals.
  Then $\boldsymbol{X}_r$ is the $n\times p_r$ submatrix of $0\leq p_r <p$ redundent covariates and $\boldsymbol{X}_f$ the $n\times (p-p_r)$ submatrix of the free (independent) covariates.
-In the following, we note $\boldsymbol{X}^j$ the $j^{th}$ column of $\boldsymbol{X}$ and $|.|$ represents the cardinal of an ensemble.
+In the following, we note $\boldsymbol{X}^j$ the $j^{th}$ column of $\boldsymbol{X}$.
 The structure $S$ of $p_r$ regressions within correlated covariates in $\boldsymbol{X}$ is described by:
@@
 	\end{equation}
-		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_r)}$ are the vectors of the regression coefficients between the covariates. 
+		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_r)}$ are the sparse vectors of the regression coefficients between the covariates (each sub-regression freely implies different covariates). 
 \\
 \\
-We note $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
-We also define $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$. 
-We see that $I_f$ defines the non-null coefficients in $\boldsymbol{\alpha}_j$ (each sub-regression can be very parsimonious).
-Then we have an explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each sub-regression  and $p_f^j=|I_f^j|$.
-\\
-\\
-
-The partition of $\boldsymbol{X}$ implies the uncrossing rule $I_f\cap I_r=\emptyset$ {\it i.e.} endogenous variables don't explain other covariates.
-This hypothesis ensures that $S$ contains no cycle and is straightforward readable (no need to order $I_r$).
+
+The partition of $\boldsymbol{X}$ implies the uncrossing rule % $I_f\cap I_r=\emptyset$ 
+{\it i.e.} endogenous variables don't explain other covariates.
+This hypothesis ensures that $S$ contains no cycle and is straightforward readable (no need to order the sub-regressions).
 It is not so restrictive because cyclic structures have no sense and any non-cyclic structure can be associated with a structure that verifies the uncrossing constraint by just successively replacing endogenous covariates by their sub-regression when they are also exogenous in some other sub-regressions.
@@
 The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
-Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
+Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when it is basic (recursive \textsc{SEM}).
   	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) without explicit expression of the correlations and thus a smaller interpretation potential.
@@
 	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a sub-regression).
-Thus we have one regression on our response variable ($P(\boldsymbol{Y}|\boldsymbol{X}))$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$).
+Thus we have one regression on the response variable ($P(\boldsymbol{Y}|\boldsymbol{X}))$ and a system of sub-regressions (without the response variable: $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$).
 Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
@@
 
-Moreover, the uncrossing constraint ($I_f\cap I_r=\emptyset$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
+Moreover, the uncrossing constraint (partition of $\boldsymbol{X}$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
 
@@
 Thus it must be a global criterion.
+Because it is about model selection and we are able to provide a full generative model (section \ref{sectionfullgen}), we decide to follow a Bayesian approach (\cite{raftery1995bayesian}, \cite{andrieu1999joint},\cite{chipman2001practical}).
 
@@
 
-	\subsection{Modeling the uncorrelated covariates: a full generative approach on $P(\boldsymbol{X})$}
+	\subsection{Modeling the uncorrelated covariates: a full generative approach on $P(\boldsymbol{X})$} \label{sectionfullgen}
 	To be able to compare structures with $P(S|\boldsymbol{X})$, we need a full generative model on $\boldsymbol{X}$.
@@
 	\begin{equation}
-			\forall j \notin I_r : \boldsymbol{X}^j_{|S} \sim f(\boldsymbol{\theta}_j)=\mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j.
+			\forall \boldsymbol{X}^j \notin \boldsymbol{X}_r : \boldsymbol{X}^j_{|S} \sim f(\boldsymbol{\theta}_j)=\mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j.
 \label{mixtureX1}
@@
 
-Because it is about model selection and we are able to provide a full generative model, we decide to follow a Bayesian approach (\cite{raftery1995bayesian}, \cite{andrieu1999joint},\cite{chipman2001practical}).
-
   Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood of the joint law on $\boldsymbol{X}$ according to the complexity of the structure~\cite{BIChuard}.
@@
 	\end{eqnarray}
-	But $BIC$ tends to give too complex structures because we test a great range of models. 
-	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
+	But $BIC$ tends to give too complex structures because we test a great range of models. Thus we choose to penalise the complexity a bit more.
+
+	We note $I_r$ the set of indices of endogenous variables in $\boldsymbol{X}$ (explained ones).
+We also define $I_f=\{I_f^1,\dots,I_f^p \}$ the set of the sets of indices of exogenous covariates (explaining ones $=\boldsymbol{X}_f$) with $\forall j \notin I_r, I_f^j=\emptyset$. 
+We see that $I_f$ defines the non-null coefficients in $\boldsymbol{\alpha}_j$ (each sub-regression can be very parsimonious).
+Then we have the explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each sub-regression  and $p_f^j=|I_f^j|$, with $|.|$ the cardinal of an ensemble.
+
+	 We suppose a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
 %	 Thus we have :

### 2014-04-28 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@558 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
-30 & 0 & U&MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
+30 & 16 & U&MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
 & & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\

### 2014-04-28 www/article13.tex
up ident
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@559 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$, then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained in books and papers (\cite{leamer1978least},\cite{cameron2005microeconometrics}).
-
-	Remark : the uncrossing constraint ($I_2\cap I_1 = \emptyset$) significantly reduces the number of feasible models and thus increases identifiability (because it also reduces the number of models equivalent in interpretation).
+	In the followings, $N_{11}$ and $N_{12}$ Gaussian (multivariate or not). $\boldsymbol{X}_r$ is here supposed to be univariate (identifiability in such a case will be generalisable).
+		\begin{eqnarray}
+			f(\boldsymbol{X_f})&=& p_{11} N_{11}+p_{12}N_{12}\\
+			\boldsymbol{X}_r&=&\boldsymbol{X}_f\boldsymbol{\alpha}_r +\boldsymbol{\varepsilon}_r \\
+			f(\boldsymbol{X}_r)&=&p_{11}(N_{11}+\boldsymbol{\varepsilon})+p_{12}(N_{12}+\boldsymbol{\varepsilon}) \\
+			f(\boldsymbol{X}_r|\boldsymbol{X}_f)&=&\boldsymbol{\varepsilon}\\
+			f(\boldsymbol{X}_r,\boldsymbol{X}_f)&=&f(\boldsymbol{X}_r|\boldsymbol{X}_f)f(\boldsymbol{X}_f)=f(\boldsymbol{X}_f|\boldsymbol{X}_r)f(\boldsymbol{X}_r) \\
+			f(\boldsymbol{X}_f|\boldsymbol{X}_r)&=&\frac{f(\boldsymbol{X}_r|\boldsymbol{X}_f)f(\boldsymbol{X}_f)}{f(\boldsymbol{X}_r)} \\
+			&=&\frac{(p_{11} N_{11}+p_{12}N_{12})\boldsymbol{\varepsilon}}{p_{11}(N_{11}+\boldsymbol{\varepsilon})+p_{12}(N_{12}+\boldsymbol{\varepsilon})} \textrm{ with }p_{11}+p_{12}=1\\
+			&=&\frac{(p_{11} N_{11}+p_{12}N_{12})\boldsymbol{\varepsilon}}{(p_{11} N_{11}+p_{12}N_{12})+\boldsymbol{\varepsilon}}
+		\end{eqnarray}
+	So we have a quotient of two multivariate Gaussian mixtures and the question is to know if it is Gaussian (in such a case we wouldn't have identifiability).
+If $\boldsymbol{X}_f$ is a Gaussian mixture with $K>1$ distinct components, then so will be $\boldsymbol{X}_r$.
+	We know that for a Gaussian multivariate distribution $(\boldsymbol{X}_r,\boldsymbol{X}_f)$ the conditional distributions $(\boldsymbol{X}_r|\boldsymbol{X}_f)$ and $(\boldsymbol{X}_f|\boldsymbol{X}_r)$ are both Gaussian.
+
+		We decompose the joint distribution based on the components $k$.
+
+	\begin{eqnarray}
+			f(\boldsymbol{X}_r)&=&f(\boldsymbol{X}_f\boldsymbol{\alpha}_r +\boldsymbol{\varepsilon}_r) (MG to define explicitely) \\
+			\forall 1\leq j \leq K,  f(\boldsymbol{X}_r|k=j)&=& Gaussian to define
+			\end{eqnarray}
+		\begin{equation}
+			f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{j=1}^K p_jf(\boldsymbol{X}_r,\boldsymbol{X}_f|k=j)
+	\end{equation}
+			where  $\forall 1\leq j \leq K$, $ f(\boldsymbol{X}_r,\boldsymbol{X}_f|k=j)$  is Gaussian multivariate  and thus
+			\begin{equation}
+			f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{j=1}^K p_jf(\boldsymbol{X}_f|\boldsymbol{X}_r,k=j)f(\boldsymbol{X}_r|k=j)
+	\end{equation}
+				where  $\forall 1\leq j \leq K$, $ f(\boldsymbol{X}_r,\boldsymbol{X}_f|k=j)$  is Gaussian and $f(\boldsymbol{X}_f|\boldsymbol{X}_r,k=j)$ is Gaussian
+
+				Now we know that $\forall 1\leq j \leq K$,  $f(\boldsymbol{X}_f|\boldsymbol{X}_r,k=j)$ is Gaussian:
+				\begin{equation}
+				f(\boldsymbol{X}_f|\boldsymbol{X}_r)=\sum_{j=1}^Kp_jf(\boldsymbol{X}_f|\boldsymbol{X}_r,k=j)
+				\end{equation}
+				is a Gaussian mixture, so we have identifiability of the model even for trivial sub-regression if we keep the constraint to have distinct components.
+			%f(\boldsymbol{X}_r,\boldsymbol{X}_f)&=& sum_{j=1}^K
+
+%	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$, then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained in books and papers (\cite{leamer1978least},\cite{cameron2005microeconometrics}).
+%	
+	Remark : the uncrossing constraint ($I_r\cap I_f = \emptyset$) significantly reduces the number of feasible models and thus increases identifiability (because it also reduces the number of models equivalent in interpretation).
 

### 2014-05-01 www/article13.tex
up ident
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@560 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up identi
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@561 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-04-28 www/article13.tex
@@
 
-%ancienne intro
-
-The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
-Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when it is basic (recursive \textsc{SEM}).
-  	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) without explicit expression of the correlations and thus a smaller interpretation potential.
-  	\\ Here we suppose independence between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\alpha}_j$ and $\sigma_j$.
-
 	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a sub-regression).
@@
 Expert knowledge can even be added to the structure, physical models for example.
+
+The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
+Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when it is basic (recursive \textsc{SEM}).
+  	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) without explicit expression of the correlations and thus a smaller interpretation potential.
+  	\\ Here we suppose independence between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\alpha}_j$ and $\sigma_j$.
 
@@
 	\begin{equation}
-			\forall \boldsymbol{X}^j \notin \boldsymbol{X}_r : \boldsymbol{X}^j_{|S} \sim f(\boldsymbol{\theta}_j)=\mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j.
+			\forall \boldsymbol{X}^j \notin \boldsymbol{X}_r : \boldsymbol{X}^j_{|S} \sim f(\boldsymbol{\theta}_j)=\mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j.
 \label{mixtureX1}
@@
 		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed (exponentially distributed for example).
-Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.
+Gaussian case is just a special case ($K_j=1$) of Gaussian mixture so it is included in our hypothesis but identifiability of $S$ requires to have at least one Gaussian mixture with at least two distinct components in each sub-regression (theorem and proof in appendix \ref{preuveident}).
+Identifiability if $S$ is not necessary to use a given structure but helps to find it.
 
@@
 We see that $I_f$ defines the non-null coefficients in $\boldsymbol{\alpha}_j$ (each sub-regression can be very parsimonious).
-Then we have the explicit structure $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each sub-regression  and $p_f^j=|I_f^j|$, with $|.|$ the cardinal of an ensemble.
-
+Then we have the explicit structure characterized by $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each sub-regression  and $p_f^j=|I_f^j|$, with $|.|$ the cardinal of an ensemble.
+\\
+Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)$ so it is equivalent to a maximization of the $BIC$.
 	 We suppose a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
-%	 Thus we have :
-%		\begin{eqnarray}
-%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
-%	\end{eqnarray}		
+	 Thus we have :
+		\begin{eqnarray}
+		BIC_+(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
+	\end{eqnarray}		
 	It increases penalty on complexity for $p_r<\frac{p}{2}$ and $p_f^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
-	this constraint on $\hat{p}_r$ and $\hat{p}_f^j$ is given in the research algorithm when the Hierarchical Uniform (HU) hypothesis is made instead of Uniform one (U) in numerical experiments (section \ref{sectionsimul} and \ref{sectionrealcase}).
-		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we obtain better results when (HU) this constraint is verified.
-Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)$ so it is equivalent to a maximization of the $BIC$.
+	this constraint on $\hat{p}_r$ and $\hat{p}_f^j$ is given in the research algorithm when the Hierarchical Uniform hypothesis is made instead of Uniform one in numerical experiments (section \ref{sectionsimul} and \ref{sectionrealcase}).
+		We do not change $BIC$ but only $P(S)$ so the properties of $BIC_+$ are the same as classical $BIC$ but we obtain better results when the constraints on the complexity are verified.
 With the Hierarchical Uniform hypothesis we maximize $\psi(\boldsymbol{X},S)\approx BIC + P(S)$.
@@
 %
-
-	%ancienne intro	
-	%ancienne intro
 
@@
 \textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
-	In the followings we suppose $\sigma^2_j\neq 0$.
+	In the followings we suppose $\sigma^2_j\neq 0$, then $\boldsymbol{X}_f'\boldsymbol{X}_f$ and $\boldsymbol{X}'\boldsymbol{X}$ are of full rank (but the later is ill-conditioned for small values of $\sigma^2_j$)  .
 
@@
 	\subsubsection{Identifiability of $\Theta$}		
-		We know by identifiability of the Gaussian mixtures and linear regression ($\boldsymbol{X}_f'\boldsymbol{X}_f$ is of full rank) that if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|S,\tilde{\Theta})$ then $\Theta=\tilde{\Theta}$.
+		We know by identifiability of the Gaussian mixtures and linear regression ($\boldsymbol{X}_f'\boldsymbol{X}_f$ is of full rank) that if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|S,\tilde{\Theta})$ then $\Theta=\tilde{\Theta}$ up to a permutation of the components.
 But the identifiability of the discrete parameter $S$ is still pending.
 
-	\subsubsection{Sufficient conditions for Identifiability of $S$}
+	\subsubsection{Identifiability of $S$, quick look}
 
@@
 			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
-			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } k_j  \label{densitydroite}\\
+			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j  \label{densitydroite}\\
 			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
@@
 		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
-We call this difference in component number "identifiability" in the sense that we try to find the model with the fewest component in $X_f$ (we use the BIC so when comparing two models with the same likelihood, the one with fewer parameters will win).
+We call this difference in component number "identifiability" in the sense that we try to find the model with the fewest component in $\boldsymbol{X}_f$ (we use the BIC so when comparing two models with the same likelihood, the one with fewer parameters will win).
 So in this case we have a "better" model (the simplest one) in the group of equivalent-meaning models (permuting variables in some sub-regression).
-	So a sufficient condition is $\forall j \in I_r: k_j\geq 2$ with at least two distinct couples $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$
-
-	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$, then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained in books and papers (\cite{leamer1978least},\cite{cameron2005microeconometrics}).
-
-	Remark : the uncrossing constraint ($I_2\cap I_1 = \emptyset$) significantly reduces the number of feasible models and thus increases identifiability (because it also reduces the number of models equivalent in interpretation).
+	\\
+
+	\subsubsection{Identifiability: Theorem and proof} \label{preuveident}
+		A necessary and sufficient condition for identifiability of $S$ is $\forall j \in I_r: K_j\geq 2$ with at least two distinct components in $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$
+\\
+	{\it Proof}:
+	We know that for a Gaussian multivariate distribution $(\boldsymbol{X}_r,\boldsymbol{X}_f)$ the conditional distributions $(\boldsymbol{X}_r|\boldsymbol{X}_f)$ and $(\boldsymbol{X}_f|\boldsymbol{X}_r)$ are both Gaussian (multivariate or not according to $p_r$).
+
+		If $\exists j \in I_r: k_j=1$ (Gaussian) then all the covariates on the right of the associated sub-regression (with non-zero coefficients) are Gaussian and then the sub-regression can be permuted and is not identifiable.
+So if $\boldsymbol{X}^j$ are Gaussian mixtures, it is a necessary condition for identifiability to have $\forall j \in I_r: K_j\geq 2$ with at least two distinct components in $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$.
+
+	Gaussian mixtures in $\boldsymbol{X_r}$ with at least two distinct components does imply the presence of at least one Gaussian mixture in each sub-regression (with at least two components).
+%	\begin{equation}
+%	 \forall j \in I_f :    \boldsymbol{X}^j_{|S}=\sum_{k=1}^{K_j}\pi_j(k)\mathcal{N}\left(\mu_j(k),\sigma_j^2(k)\boldsymbol{I}_n\right) \\
+%	\end{equation}
+%	$\forall j \in I_r$ we can define $K_{j}$ the number of distinct components of $\boldsymbol{X}_j$.
+%	\begin{eqnarray}
+%				\boldsymbol{X}^{j}_{|S}&=&\sum_{k=1}^{K_j}\pi_j(k)\mathcal{N}\left(\mu_j(k),\sigma_j^2(k)\boldsymbol{I}_n\right)
+%	\end{eqnarray}
+	The joint distribution of the $p$ Gaussian mixtures is a multivariate Gaussian mixture $\boldsymbol{X}$ with $K\geq 2$ distinct components $(\boldsymbol{m}_k,\boldsymbol{\Sigma}_k)$ where $\boldsymbol{m}_k \in \mathbf{R}^p$ and $\boldsymbol{\Sigma}_k$ is a $p \times p$ matrix.
+		We decompose the joint distribution based on the components:
+
+		\begin{equation}
+			f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{k=1}^K p_kf(\boldsymbol{X}_r,\boldsymbol{X}_f|k)
+	\end{equation}
+			where  $\forall 1\leq k \leq K$, $ f(\boldsymbol{X}_r,\boldsymbol{X}_f|k)$  is a multivariate Gaussian.
+			\\
+			Then,  $\forall j \in I_r, \forall i \in I_f^j$ we have:
+			\begin{equation}
+			 f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{k=1}^K p_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)f(\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i|k)
+	\end{equation}
+				where the $f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)$ are Gaussian.
+
+Hence,
+\begin{equation}
+				f(\boldsymbol{X}^i|\boldsymbol{X}\setminus\boldsymbol{X}^i)=f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i)=\sum_{k=1}^Kp_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)
+				\end{equation}
+
+				is a Gaussian mixture.
+
+%				Identifiability of $S$ means we don't have a linear regression of $\boldsymbol{X}^i$ on $(\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i)$.
+So we need a non-linear relationship or a not-Gaussian residual $f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}\setminus\boldsymbol{X}^i)$, {\it i.e.
+} the Gaussian mixture $f(\boldsymbol{X}^i|\boldsymbol{X}\setminus\boldsymbol{X}^i)$ should have at least two distinct components.
+%	Linear relationship means $\exists \boldsymbol{\gamma} \in \mathbf{R}^{p-1}$ and $\sigma_i \in \mathbf{R}$ for which
+%	\begin{equation}
+%		\forall 1\leq k \leq K, \boldsymbol{X}^i_{|\boldsymbol{X}_r,\boldsymbol{X}\setminus\boldsymbol{X}^i,k}=\mathcal{N}(\boldsymbol{m}_k^{\bar{i}}\boldsymbol{\gamma},\sigma_i^2)
+%	\end{equation}
+%					
+%				
+%				\begin{equation}
+%				f(\boldsymbol{X}^i|\boldsymbol{X}\setminus\boldsymbol{X}^i)=f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i)=\sum_{k=1}^Kp_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)
+%				\end{equation}
+%				
+				If it is a Gaussian mixture with at least two distinct components when $\boldsymbol{X}^j \in \boldsymbol{X}_r$ has at least two distinct components.
+				Then $\boldsymbol{X}^i$ can't be explained by a linear regression on the other covariates (non-Gaussian residual).		
+				 So we have identifiability of the structure $S$ even for trivial sub-regressions if we keep the constraint to have distinct components in $\boldsymbol{X}_r$ (equivalent to have at least one Gaussian mixture with at least two components in each sub-regression).
+%			%f(\boldsymbol{X}_r,\boldsymbol{X}_f)&=& sum_{j=1}^K 
+%
+%The distribution of $\boldsymbol{X}_r$ is a Gaussian mixture defined in (\ref{densitygauche}) thus
+%			$\forall 1\leq k \leq K$,  $(\boldsymbol{X}_r|k)$ is Gaussian: 
+%		
+%
+%
+%In the followings, $N_{11}$ and $N_{12}$ Gaussian (multivariate or not). $\boldsymbol{X}_r$ is here supposed to be univariate (identifiability in such a case will be generalisable).
+%		\begin{eqnarray}
+%			f(\boldsymbol{X_f})&=& p_{11} N_{11}+p_{12}N_{12}\\
+%			\boldsymbol{X}_r&=&\boldsymbol{X}_f\boldsymbol{\alpha}_r +\boldsymbol{\varepsilon}_r \\
+%			f(\boldsymbol{X}_r)&=&p_{11}(N_{11}+\boldsymbol{\varepsilon})+p_{12}(N_{12}+\boldsymbol{\varepsilon}) \\
+%			f(\boldsymbol{X}_r|\boldsymbol{X}_f)&=&\boldsymbol{\varepsilon}\\
+%			f(\boldsymbol{X}_r,\boldsymbol{X}_f)&=&f(\boldsymbol{X}_r|\boldsymbol{X}_f)f(\boldsymbol{X}_f)=f(\boldsymbol{X}_f|\boldsymbol{X}_r)f(\boldsymbol{X}_r) \\
+%			f(\boldsymbol{X}_f|\boldsymbol{X}_r)&=&\frac{f(\boldsymbol{X}_r|\boldsymbol{X}_f)f(\boldsymbol{X}_f)}{f(\boldsymbol{X}_r)} \\
+%			&=&\frac{(p_{11} N_{11}+p_{12}N_{12})\boldsymbol{\varepsilon}}{p_{11}(N_{11}+\boldsymbol{\varepsilon})+p_{12}(N_{12}+\boldsymbol{\varepsilon})} \textrm{ with }p_{11}+p_{12}=1\\
+%			&=&\frac{(p_{11} N_{11}+p_{12}N_{12})\boldsymbol{\varepsilon}}{(p_{11} N_{11}+p_{12}N_{12})+\boldsymbol{\varepsilon}}
+%		\end{eqnarray}
+%	So we have a quotient of two multivariate Gaussian mixtures and the question is to know if it is Gaussian (in such a case we wouldn't have identifiability).
+%If $\boldsymbol{X}_f$ is a Gaussian mixture with $K>1$ distinct components, then so will be $\boldsymbol{X}_r$.
+%	
+%	If the sufficient condition is not satisfied but we have different variances for $\boldsymbol{X}_f$ and $\boldsymbol{X}_r$, then we don't have identifiability but reverse regressions don't give the same results as the true model, as illustrated in the well-known salary discrimination study \cite{goldberger1984reverse} and explained in books and papers (\cite{leamer1978least},\cite{cameron2005microeconometrics}).
+%
 

### 2014-05-01 www/article13.tex
up intro + tableaux
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@562 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
- Instead of reducing the dimension without looking at correlations,  we propose to focus on the correlations, giving an explicit structure between covariates with the idea that if we know explicitly the correlations, we could use this knowledge to avoid the problem.
-More precisely, we model the correlations with a system of linear regressions for the joint distribution of the covariates to model the structure of the correlations.
+ Instead of reducing the dimension without looking at correlations, Selvarclust \cite{maugis2009variable} proposes to focus on the correlations, giving an explicit structure between covariates by linear sub-regressions between them.
+But is in a clustering context, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even it is known to be often unstable \cite{miller2002subset}.
+We propose to develop this method for linear regression with a specifically adapted algorithm to find the structure of sub-regression.
+
+  The idea is that if we know explicitly the correlations, we could use this knowledge to avoid the problem.
+More precisely, we model the correlations with a system of linear sub-regressions for the joint distribution of the covariates to model the structure of the correlations.
 It helps to define the greatest set of orthogonal covariates to keep the maximum information but with an orthogonality constraint.
@@
 
+Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)\propto P(\boldsymbol{X}|S)$ so it is equivalent to a maximization of the $BIC$.
 	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
@@
 \\
-Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)$ so it is equivalent to a maximization of the $BIC$.
-	 We suppose a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of a simple uniform law on $S$ that is generally used and provides no penalty.
+	 We suppose a hierarchical uniform {\it a priori} distribution $P(S)=P(I_f | \boldsymbol{p}_f,I_r,p_r)P(\boldsymbol{p}_f|I_r,p_r)P(I_r|p_r)P(p_r)$  instead of the simple uniform law on $S$ that is generally used and provides no penalty.
 	 Thus we have :
@@
 	this constraint on $\hat{p}_r$ and $\hat{p}_f^j$ is given in the research algorithm when the Hierarchical Uniform hypothesis is made instead of Uniform one in numerical experiments (section \ref{sectionsimul} and \ref{sectionrealcase}).
-		We do not change $BIC$ but only $P(S)$ so the properties of $BIC_+$ are the same as classical $BIC$ but we obtain better results when the constraints on the complexity are verified.
-With the Hierarchical Uniform hypothesis we maximize $\psi(\boldsymbol{X},S)\approx BIC + P(S)$.
+		$BIC_+$ does not change $BIC$ but only $P(S)$ so the properties of $BIC_+$ are the same as classical $BIC$ but we obtain better results when the constraints on the complexity are verified.
+%With the Hierarchical Uniform hypothesis we maximize $\psi(\boldsymbol{X},S)\approx BIC + P(S)$.
 
@@
 \subsection{MCMC algorithm}
-	Now we have a comparison criterion, we define an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN). 
+	Now we have a comparison criterion $\psi(\boldsymbol{X},S)$, we define an MCMC algorithm to find the structure (R package \textsc{CorReg} on CRAN). 
 	\subsubsection{The neighbourhood}
@@
 		\subsubsection{How to evaluate found structure?}
-			The first criterion is $P(S|\boldsymbol{X})$ which is maximized in the MCMC.
+			The first criterion is $\psi(\boldsymbol{X},S)$ which is maximized in the MCMC.
 But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning.
@@
 		\subsubsection{Results on $S$}	\label{compZ}
-In table \ref{compZvrai} we compare found structures in different contexts with both Uniform (U) and Hierarchical Uniform (HU) a priori law on $P(S)$. 
-We see that (HU) hypothesis gives sparser models even with $\max (\hat{p}_f^j)=5$.
-Moreover, this constraint is not active because observed comlexity are smaller (this constraint only serves to accelerate the walk by reducing the dimension of $\mathcal{S}$ because each configuration was computed a hundred times), meaning that the stronger penalty implied by (HU) really is efficient.
-The datasets used for (U) and (HU) are the same to keep the comparison  meaningful.
+In table \ref{compZvrai} we compare found structures in different contexts with both Uniform and Hierarchical Uniform {\it a priori} law on $P(S)$, noted respectively $BIC$ and $BIC_+$. 
+We see that usage of $BIC_+$ gives sparser models even with $\max (\hat{p}_f^j)=5$.
+Moreover, this constraint is not active because observed complexities are smaller (this constraint only serves to accelerate the walk by reducing the dimension of $\mathcal{S}$ because each configuration was computed a hundred times), meaning that the stronger penalty implied by $BIC_+$ really is efficient.
+The datasets used for $BIC$ and $BIC_+$ are the same to keep the comparison  meaningful.
 These datasets and $\hat{S}$ are those used for tables
@@
 
- It is also notable that (HU) has a greater computational cost than (U). 
-	We also notice that the MCMC is faster when there are numerous correlations (rejecting more candidates). 
+ It is also notable that $BIC_+$ has a greater computational cost than $BIC$. 
+	We notice that the MCMC is faster when there are numerous correlations (rejecting more candidates). 
 \begin{table}[h!]
@@
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U&0.4104 & 3.2428 & 0 & 5.43 & 0 & -5.43 & 22.55  \\
+30 & 0 & $BIC$ &0.4104 & 3.2428 & 0 & 5.43 & 0 & -5.43 & 22.55  \\
 & & & (0.0275) & (0.3711) & (0) & (1.9346) & (0) & (1.9346) & (8.0884) \\
@@
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&0.4182 & 2.7735 & 10.96 & 5.93 & 4.98 & -0.95 & 38.16  \\
+30 & 16 & $BIC$ &0.4182 & 2.7735 & 10.96 & 5.93 & 4.98 & -0.95 & 38.16  \\
 & & & (0.0329) & (0.1498) & (1.9844) & (2.0313) & (1.9948) & (0.9987) & (6.499) \\
@@
 & & & (0.0429) & (0.1331) & (1.4761) & (1.0888) & (1.4668) & (0.9866) & (5.0831) \\
- &  & HU & 0.4456 & 4.0233 & 16.96 & 3.04 & 14.77 & 11.73 & 4.35  \\
+ &  & $BIC_+$ & 0.4456 & 4.0233 & 16.96 & 3.04 & 14.77 & 11.73 & 4.35  \\
 & & & (0.0429) & (0.1091) & (1.3993) & (1.3993) & (1.4692) & (0.5478) & (5.8833) \\
@@
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&0.5229 & 4.7068 & 0 & 4.2 & 0 & -4.2 & 13.35  \\
+50 & 0 & $BIC$ &0.5229 & 4.7068 & 0 & 4.2 & 0 & -4.2 & 13.35  \\
 & & & (0.0519) & (0.5865) & (0) & (1.7233) & (0) & (1.7233) & (5.6468) \\
- &  & HU &0.5229 & 10.1198 & 0 & 0.13 & 0 & -0.13 & 0.32  \\
+ &  & $BIC_+$ &0.5229 & 10.1198 & 0 & 0.13 & 0 & -0.13 & 0.32  \\
 & & & (0.0519) & (0.5541) & (0) & (0.3667) & (0) & (0.3667) & (0.9732) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&0.5205 & 3.3681 & 11.15 & 5.42 & 4.72 & -0.7 & 22.85  \\	
+50 & 16 & $BIC$ &0.5205 & 3.3681 & 11.15 & 5.42 & 4.72 & -0.7 & 22.85  \\	
 & & & (0.0451) & (0.3123) & (1.93) & (1.9132) & (1.886) & (0.7317) & (5.7742) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U& 0.5833 & 3.2683 & 28.17 & 1.38 & 3.7 & 2.32 & 12.74  \\
+50 & 32 & $BIC$ & 0.5833 & 3.2683 & 28.17 & 1.38 & 3.7 & 2.32 & 12.74  \\
 & & & (0.0628) & (0.316) & (1.3711) & (0.9077) & (1.3143) & (0.8394) & (4.3359) \\
@@
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U& 0.9623 & 12.9373 & 0 & 2.83 & 0 & -2.83 & 6.23  \\
+100 & 0 & $BIC$ & 0.9623 & 12.9373 & 0 & 2.83 & 0 & -2.83 & 6.23  \\
 & & & (0.077) & (1.7778) & (0) & (1.2953) & (0) & (1.2953) & (3.1999) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U& 1.1223 & 6.9647 & 11.67 & 4.8 & 4.25 & -0.55 & 12.58  \\
+100 & 16 & $BIC$ & 1.1223 & 6.9647 & 11.67 & 4.8 & 4.25 & -0.55 & 12.58  \\
 & & & (0.1122) & (0.5473) & (2.0003) & (2.0646) & (1.956) & (0.7833) & (3.9471) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U& 1.4343 & 5.9626 & 30.14 & 0.84 & 1.61 & 0.77 & 6.96  \\
+100 & 32 & $BIC$ & 1.4343 & 5.9626 & 30.14 & 0.84 & 1.61 & 0.77 & 6.96  \\
 & & & (0.2528) & (0.3136) & (1.3928) & (0.8495) & (1.2941) & (0.7086) & (3.0975) \\
@@
 \hline 
-$n$ & $p_r$& $P(S)$ & indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$& $P(S)$ & indicator &OLS  &    \textsc{CorReg} $\hat S$ + OLS & \textsc{CorReg} $S$ + OLS\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U& MSE (sd) & 262627.57 (732019) & 5928332.4 (49005690.2) & 262627.57 (732019) \\
+30 & 0 & $BIC$ & MSE (sd) & 262627.57 (732019) & 5928332.4 (49005690.2) & 262627.57 (732019) \\
 & & & cpl (sd) & 30 (0) & 29.99 (0.1) & 30 (0) \\
@@
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 510747.53 (2287539.8) & 635.42 (335.2) & 610.32 (424.9) \\
+30 & 16 & $BIC$ &MSE (sd) & 510747.53 (2287539.8) & 635.42 (335.2) & 610.32 (424.9) \\
 & & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
@@
 & & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
- &  & HU & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
+ &  & $BIC_+$ & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
 & & & cpl (sd) & 30 (0) & 21 (0) & 9 (0) \\
@@
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&MSE (sd) & 528.08 (228.6) & 886.54 (364.3) & 528.08 (228.6) \\
+50 & 0 & $BIC$ &MSE (sd) & 528.08 (228.6) & 886.54 (364.3) & 528.08 (228.6) \\
 & & & cpl (sd) & 41 (0) & 36.8 (1.7) & 41 (0) \\
- &  & HU &MSE (sd) & 528.08 (228.6) & 542.29 (235) & 528.08 (228.6) \\
+ &  & $BIC_+$ &MSE (sd) & 528.08 (228.6) & 542.29 (235) & 528.08 (228.6) \\
 & & & cpl (sd) & 41 (0) & 40.87 (0.4) & 41 (0) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&MSE (sd) & 612.72 (291.5) & 239.17 (89.4) & 200.32 (42.6) \\	
+50 & 16 & $BIC$ &MSE (sd) & 612.72 (291.5) & 239.17 (89.4) & 200.32 (42.6) \\	
 & & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 555.44 (262.5) & 128.98 (18.1) & 121.08 (11.9) \\
+50 & 32 & $BIC$ &MSE (sd) & 555.44 (262.5) & 128.98 (18.1) & 121.08 (11.9) \\
 & & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
@@
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U& MSE (sd) & 167.71 (20.6) & 323.44 (124.1) & 167.71 (20.6) \\
+100 & 0 & $BIC$ & MSE (sd) & 167.71 (20.6) & 323.44 (124.1) & 167.71 (20.6) \\
 & & & cpl (sd) & 41 (0) & 38.17 (1.3) & 41 (0) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  MSE (sd) & 168.68 (22.4) & 158.51 (51.6) & 133.49 (12.2) \\
+100 & 16 & $BIC$ &  MSE (sd) & 168.68 (22.4) & 158.51 (51.6) & 133.49 (12.2) \\
 & & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  MSE (sd) & 173.25 (22.9) & 112.8 (10.7) & 110.63 (6.9) \\
+100 & 32 & $BIC$ &  MSE (sd) & 173.25 (22.9) & 112.8 (10.7) & 110.63 (6.9) \\
 & & & cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
@@
 \hline 
-$n$ & $p_r$&  $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$ + LASSO& \textsc{CorReg} $S$ + LASSO\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U&MSE (sd) & 1246.35 (350.5) & 1433.98 (526.7) & 1246.35 (350.5) \\
+30 & 0 & $BIC$ &MSE (sd) & 1246.35 (350.5) & 1433.98 (526.7) & 1246.35 (350.5) \\
 & & & cpl (sd) & 17.84 (5.5) & 15.79 (5.6) & 17.84 (5.5) \\
@@
 & & & cpl (sd) & 16.68 (4.4) & 15.35 (4.3) & 16.14 (4.5) \\
- &  & HU & MSE (sd) & 712.79 (405.2) & 551.92 (242.4) & 554.01 (257.1) \\
+ &  & $BIC_+$ & MSE (sd) & 712.79 (405.2) & 551.92 (242.4) & 554.01 (257.1) \\
 & & & cpl (sd) & 16.68 (4.4) & 15.96 (4.4) & 16.14 (4.5) \\
@@
 & & & cpl (sd) & 11.23 (4.4) & 8.56 (1.8) & 7.91 (1.3) \\
- &  & HU & MSE (sd) & 216.12 (128.3) & 178.88 (68.9) & 147.04 (28.4) \\
+ &  & $BIC_+$ & MSE (sd) & 216.12 (128.3) & 178.88 (68.9) & 147.04 (28.4) \\
 & & & cpl (sd) & 11.23 (4.4) & 9.86 (3) & 7.91 (1.3) \\
@@
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&MSE (sd) & 658.38 (221.4) & 872.32 (239.4) & 658.38 (221.4) \\
+50 & 0 & $BIC$ &MSE (sd) & 658.38 (221.4) & 872.32 (239.4) & 658.38 (221.4) \\
 & & & cpl (sd) & 28.2 (6) & 22.95 (5.8) & 28.2 (6) \\
- &  & HU &MSE (sd) & 658.38 (221.4) & 652.72 (211.9) & 658.38 (221.4) \\
+ &  & $BIC_+$ &MSE (sd) & 658.38 (221.4) & 652.72 (211.9) & 658.38 (221.4) \\
 & & & cpl (sd) & 28.2 (6) & 28.36 (5.9) & 28.2 (6) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&MSE (sd) & 274.34 (101.7) & 268.76 (106.3) & 225.64 (67.5) \\	
+50 & 16 & $BIC$ &MSE (sd) & 274.34 (101.7) & 268.76 (106.3) & 225.64 (67.5) \\	
 & & & cpl (sd) & 22.14 (4.1) & 19.52 (2.9) & 20.47 (2.8) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 165.6 (73.7) & 128.34 (18.8) & 123.73 (12.8) \\
+50 & 32 & $BIC$ &MSE (sd) & 165.6 (73.7) & 128.34 (18.8) & 123.73 (12.8) \\
 & & & cpl (sd) & 12.93 (6.4) & 8.51 (1.5) & 8.09 (1.1) \\
@@
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U&  MSE (sd) & 183.33 (31.1) & 357.15 (133.4) & 183.33 (31.1) \\
+100 & 0 & $BIC$ &  MSE (sd) & 183.33 (31.1) & 357.15 (133.4) & 183.33 (31.1) \\
 & & & cpl (sd) & 37.78 (2.4) & 32.93 (3.7) & 37.78 (2.4) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U& MSE (sd) & 148.83 (18.6) & 164.44 (54.3) & 139.57 (16.1) \\
+100 & 16 & $BIC$ & MSE (sd) & 148.83 (18.6) & 164.44 (54.3) & 139.57 (16.1) \\
 & & & cpl (sd) & 25.22 (4) & 21.72 (2) & 22.3 (1.7) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U& MSE (sd) & 124.29 (19.9) & 113.36 (11) & 111.54 (7.2) \\
+100 & 32 & $BIC$ & MSE (sd) & 124.29 (19.9) & 113.36 (11) & 111.54 (7.2) \\
 & & & cpl (sd) & 13.18 (5.9) & 8.68 (1.2) & 8.53 (1) \\
@@
 \hline 
-$n$ & $p_r$&  $P(S)$&indicator &Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$&indicator &Elasticnet  &    \textsc{CorReg} $\hat S$ + Elasticnet& \textsc{CorReg} $S$ + Elasticnet\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U&MSE (sd) & 1 326.54 (388.4) & 1 356.98 (331.4) & 1 326.54 (388.4) \\
+30 & 0 & $BIC$ &MSE (sd) & 1 326.54 (388.4) & 1 356.98 (331.4) & 1 326.54 (388.4) \\
 & & & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
@@
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 1 400.56 (1598.2) & 668.57 (274.6) & 653.59 (283.7) \\
+30 & 16 & $BIC$ &MSE (sd) & 1 400.56 (1598.2) & 668.57 (274.6) & 653.59 (283.7) \\
 & & & cpl (sd) & 13.86 (6.8) & 14.31 (5.7) & 14.83 (5.3) \\
@@
 & & & cpl (sd) & 15.57 (6.3) & 11.19 (2.4) & 8.37 (1.3) \\
- &  & HU & MSE (sd) & 855.74 (582.6) & 311.57 (163.9) & 146.79 (32.7) \\
+ &  & $BIC_+$ & MSE (sd) & 855.74 (582.6) & 311.57 (163.9) & 146.79 (32.7) \\
 & & & cpl (sd) & 15.57 (6.3) & 14.87 (3.9) & 8.37 (1.3) \\
@@
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&MSE (sd) & 738.94 (254.5) & 914.56 (283.7) & 738.94 (254.5) \\
+50 & 0 & $BIC$ &MSE (sd) & 738.94 (254.5) & 914.56 (283.7) & 738.94 (254.5) \\
 & & & cpl (sd) & 25.85 (8.7) & 20.97 (7.8) & 25.85 (8.7) \\
- &  & HU &MSE (sd) & 738.94 (254.5) & 751.06 (262.2) & 738.94 (254.5) \\
+ &  & $BIC_+$ &MSE (sd) & 738.94 (254.5) & 751.06 (262.2) & 738.94 (254.5) \\
 & & & cpl (sd) & 25.85 (8.7) & 25.43 (8.8) & 25.85 (8.7) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	MSE (sd) & 516.36 (226.6) & 276.6 (133.6) & 228.85 (89.3) \\
+50 & 16 & $BIC$ &	MSE (sd) & 516.36 (226.6) & 276.6 (133.6) & 228.85 (89.3) \\
 & & & cpl (sd) & 26.72 (7.4) & 20.49 (3.8) & 21.8 (3.3) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 328.7 (146.3) & 130.61 (20.6) & 124.01 (15.4) \\
+50 & 32 & $BIC$ &MSE (sd) & 328.7 (146.3) & 130.61 (20.6) & 124.01 (15.4) \\
 & & & cpl (sd) & 24.7 (7.1) & 9.95 (1.5) & 8.22 (1.2) \\
@@
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U& MSE (sd) & 175.93 (25.9) & 355.79 (143.8) & 175.93 (25.9) \\
+100 & 0 & $BIC$ & MSE (sd) & 175.93 (25.9) & 355.79 (143.8) & 175.93 (25.9) \\
 & & & cpl (sd) & 39.71 (1.7) & 34.11 (4.6) & 39.71 (1.7) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  MSE (sd) & 173.65 (25.1) & 164.96 (59) & 139.51 (17.6) \\
+100 & 16 & $BIC$ &  MSE (sd) & 173.65 (25.1) & 164.96 (59) & 139.51 (17.6) \\
 & & &  cpl (sd) & 35.54 (4.2) & 22.58 (2.2) & 23.26 (1.9) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  MSE (sd) & 161.92 (25.2) & 113.61 (10.9) & 111.46 (7.2) \\
+100 & 32 & $BIC$ &  MSE (sd) & 161.92 (25.2) & 113.61 (10.9) & 111.46 (7.2) \\
 & & &  cpl (sd) & 30.78 (6.4) & 9.25 (1.2) & 8.67 (1) \\
@@
 \hline 
-$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$ + Stepwise& \textsc{CorReg} $S$ + Stepwise\\ 
 \hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & U& MSE (sd) & 1 919.43 (861.8) & 2 014.13 (683.6) & 1 919.43 (861.8) \\
+30 & 0 & $BIC$ & MSE (sd) & 1 919.43 (861.8) & 2 014.13 (683.6) & 1 919.43 (861.8) \\
 & & &cpl (sd) & 22.8 (3.3) & 19.34 (4.9) & 22.8 (3.3) \\
@@
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 696.46 (492.9) & 662.07 (285.9) & 660.93 (354) \\
+30 & 16 & $BIC$ &MSE (sd) & 696.46 (492.9) & 662.07 (285.9) & 660.93 (354) \\
 & & & cpl (sd) & 15.9 (3.8) & 15.1 (3.7) & 15.76 (3.7) \\
@@
 & & &cpl (sd) & 16.1 (6.3) & 8.26 (1.7) & 7.77 (1.3) \\
- &  & HU &  MSE (sd) & 394.93 (356) & 189.49 (79) & 147.81 (30.3) \\
+ &  & $BIC_+$ &  MSE (sd) & 394.93 (356) & 189.49 (79) & 147.81 (30.3) \\
 & & & cpl (sd) & 16.1 (6.3) & 10.23 (3) & 7.77 (1.3) \\
@@
 \hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & U&MSE (sd) & 745.88 (241.7) & 962.12 (319.3) & 745.88 (241.7) \\
+50 & 0 & $BIC$ &MSE (sd) & 745.88 (241.7) & 962.12 (319.3) & 745.88 (241.7) \\
 & & & cpl (sd) & 27.19 (4.6) & 22.9 (4.5) & 27.19 (4.6) \\
- &  & HU &MSE (sd) & 745.88 (241.7) & 749.08 (244.2) & 745.88 (241.7) \\
+ &  & $BIC_+$ &MSE (sd) & 745.88 (241.7) & 749.08 (244.2) & 745.88 (241.7) \\
 & & & cpl (sd) & 27.19 (4.6) & 27.21 (4.5) & 27.19 (4.6) \\
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	MSE (sd) & 279.82 (151.1) & 290.13 (117.3) & 239.61 (69.5) \\
+50 & 16 & $BIC$ &	MSE (sd) & 279.82 (151.1) & 290.13 (117.3) & 239.61 (69.5) \\
 & & & cpl (sd) & 21.24 (5.2) & 18.13 (2.8) & 19.23 (2.7) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 199.3 (158.7) & 126.81 (17.4) & 124.19 (13) \\
+50 & 32 & $BIC$ &MSE (sd) & 199.3 (158.7) & 126.81 (17.4) & 124.19 (13) \\
 & & & cpl (sd) & 14.31 (6.2) & 8.16 (1.2) & 8.04 (1) \\
@@
 \hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & U&   MSE (sd) & 193.26 (34.9) & 387.18 (151.6) & 193.26 (34.9) \\
+100 & 0 & $BIC$ &   MSE (sd) & 193.26 (34.9) & 387.18 (151.6) & 193.26 (34.9) \\
 & & & cpl (sd) & 36.3 (2.5) & 30.94 (4.2) & 36.3 (2.5) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  MSE (sd) & 145.46 (20.1) & 165.74 (56.1) & 140.27 (17.5) \\
+100 & 16 & $BIC$ &  MSE (sd) & 145.46 (20.1) & 165.74 (56.1) & 140.27 (17.5) \\
 & & & cpl (sd) & 23.44 (3.8) & 21.23 (2.1) & 21.93 (1.7) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U& MSE (sd) & 128.44 (18.4) & 113.27 (10.7) & 111.77 (7.2) \\
+100 & 32 & $BIC$ & MSE (sd) & 128.44 (18.4) & 113.27 (10.7) & 111.77 (7.2) \\
 & & &  cpl (sd) & 13.58 (5.2) & 8.59 (1.1) & 8.48 (1) \\
@@
 	\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_r}$ (worst case for us)}	 \label{tableMSEsimgauche}
-We now try the method with a response depending only on variables in $\boldsymbol{X_r}$. The datasets used here were still those from \ref{compZvrai}.
-Depending only on $\boldsymbol{X_r}$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
+We now try the method with a response depending only on variables in $\boldsymbol{X}_r$. The datasets used here were still those from \ref{compZvrai}.
+Depending only on $\boldsymbol{X}_r$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
 
@@
 \hline 
-$n$ & $p_r$&$P(S)$ &indicator &OLS  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&$P(S)$ &indicator &OLS  &    \textsc{CorReg} $\hat S$ + OLS& \textsc{CorReg} $S$ + OLS\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 81781.54 (247281.7) & 507.1 (269.2) & 593.03 (428.3) \\
+30 & 16 & $BIC$ &MSE (sd) & 81781.54 (247281.7) & 507.1 (269.2) & 593.03 (428.3) \\
 & & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
@@
 & & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
- &  & HU & MSE (sd) & 74136.81 (297716.1) & 340.53 (169.4) & 144.51 (31.4) \\
+ &  & $BIC_+$ & MSE (sd) & 74136.81 (297716.1) & 340.53 (169.4) & 144.51 (31.4) \\
 & & & cpl (sd) & 30 (0) & 21 (0) & 9 (0) \\
@@
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	MSE (sd) & 684.04 (438.1) & 190.97 (37.7) & 197.32 (40) \\
+50 & 16 & $BIC$ &	MSE (sd) & 684.04 (438.1) & 190.97 (37.7) & 197.32 (40) \\
 & & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 596.32 (323) & 126.94 (15.5) & 119.39 (12.4) \\
+50 & 32 & $BIC$ &MSE (sd) & 596.32 (323) & 126.94 (15.5) & 119.39 (12.4) \\
 & & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  MSE (sd) & 168.35 (20.1) & 133.75 (12.4) & 135.08 (13.1) \\
+100 & 16 & $BIC$ &  MSE (sd) & 168.35 (20.1) & 133.75 (12.4) & 135.08 (13.1) \\
 & & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  MSE (sd) & 168.07 (21.4) & 109.61 (7.4) & 108.64 (7.2) \\
+100 & 32 & $BIC$ &  MSE (sd) & 168.07 (21.4) & 109.61 (7.4) & 108.64 (7.2) \\
 & & &  cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
@@
 \hline 
-$n$ & $p_r$&  $P(S)$ &indicator &LAR  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &LAR  &    \textsc{CorReg} $\hat S$ + LAR& \textsc{CorReg} $S$ + LAR\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 367.45 (198.4) & 298.32 (122.4) & 292.53 (102.2) \\
+30 & 16 & $BIC$ &MSE (sd) & 367.45 (198.4) & 298.32 (122.4) & 292.53 (102.2) \\
 & & & cpl (sd) & 14.37 (5.1) & 12.9 (3.8) & 12.84 (3.8) \\
@@
 & & & cpl (sd) & 11.7 (5.8) & 8.7 (2.5) & 7.45 (1.4) \\
- &  & HU & MSE (sd) & 273.12 (267.2) & 175.68 (57.5) & 152.45 (39.5) \\
+ &  & $BIC_+$ & MSE (sd) & 273.12 (267.2) & 175.68 (57.5) & 152.45 (39.5) \\
 & & & cpl (sd) & 11.7 (5.8) & 9.19 (2.9) & 7.45 (1.4) \\
@@
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	MSE (sd) & 189.64 (52.5) & 171.97 (45.9) & 176.58 (46.8) \\
+50 & 16 & $BIC$ &	MSE (sd) & 189.64 (52.5) & 171.97 (45.9) & 176.58 (46.8) \\
 & & & cpl (sd) & 14.15 (3.9) & 13.4 (2.8) & 13.5 (3.1) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 163.88 (67.4) & 124.53 (17) & 121.99 (16.6) \\
+50 & 32 & $BIC$ &MSE (sd) & 163.88 (67.4) & 124.53 (17) & 121.99 (16.6) \\
 & & & cpl (sd) & 13.06 (7.2) & 8.29 (1.6) & 7.73 (1.2) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U& MSE (sd) & 129.65 (14.8) & 127.11 (12.1) & 127.46 (12.2) \\
+100 & 16 & $BIC$ & MSE (sd) & 129.65 (14.8) & 127.11 (12.1) & 127.46 (12.2) \\
 & & &  cpl (sd) & 14.86 (3.1) & 13.97 (2.5) & 14.04 (2.5) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U&  MSE (sd) & 121.21 (20.1) & 109.66 (7.9) & 109.04 (7.8) \\
+100 & 32 & $BIC$ &  MSE (sd) & 121.21 (20.1) & 109.66 (7.9) & 109.04 (7.8) \\
 & & & cpl (sd) & 12.84 (5.2) & 8.42 (1.1) & 8.11 (0.8) \\
@@
 \hline 
-$n$ & $p_r$&  $P(S)$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$ + Elasticnet& \textsc{CorReg} $S$ + Elasticnet\\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 499.32 (218.8) & 311.41 (137.4) & 305.09 (123.5) \\
+30 & 16 & $BIC$ &MSE (sd) & 499.32 (218.8) & 311.41 (137.4) & 305.09 (123.5) \\
 & & & cpl (sd) & 14.08 (6.1) & 11.75 (4.8) & 11.49 (4.7) \\
@@
 & & & cpl (sd) & 12.54 (7.2) & 10.26 (2.7) & 7.81 (1.4) \\
- &  & HU & MSE (sd) & 691.55 (503.2) & 279.21 (117.3) & 150.17 (35) \\
+ &  & $BIC_+$ & MSE (sd) & 691.55 (503.2) & 279.21 (117.3) & 150.17 (35) \\
 & & & cpl (sd) & 12.54 (7.2) & 13 (4.3) & 7.81 (1.4) \\
@@
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	MSE (sd) & 282.05 (134.8) & 180.64 (47.6) & 181.49 (48.6) \\
+50 & 16 & $BIC$ &	MSE (sd) & 282.05 (134.8) & 180.64 (47.6) & 181.49 (48.6) \\
 & & & cpl (sd) & 20.37 (6.3) & 13.68 (4) & 14.05 (4) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 308.13 (142.9) & 129.39 (19.6) & 122.71 (17.8) \\
+50 & 32 & $BIC$ &MSE (sd) & 308.13 (142.9) & 129.39 (19.6) & 122.71 (17.8) \\
 & & & cpl (sd) & 22.94 (7.7) & 9.37 (2.1) & 7.91 (1.4) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U& MSE (sd) & 150.14 (19.7) & 127.48 (14) & 128.21 (15) \\
+100 & 16 & $BIC$ & MSE (sd) & 150.14 (19.7) & 127.48 (14) & 128.21 (15) \\
 & & &  cpl (sd) & 25.6 (5.4) & 14.73 (3.8) & 14.73 (3.9) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U& MSE (sd) & 157.18 (23.5) & 110.24 (8) & 109.24 (7.8) \\
+100 & 32 & $BIC$ & MSE (sd) & 157.18 (23.5) & 110.24 (8) & 109.24 (7.8) \\
 & & &  cpl (sd) & 30.41 (6.7) & 8.96 (1.3) & 8.3 (1) \\
@@
 \hline 
-$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$ + stepwise & \textsc{CorReg} $S$ + stepwise \\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & U&MSE (sd) & 394.59 (454.6) & 341.47 (153.2) & 351.14 (148) \\
+30 & 16 & $BIC$ &MSE (sd) & 394.59 (454.6) & 341.47 (153.2) & 351.14 (148) \\
 & & & cpl (sd) & 13.09 (3.2) & 12.9 (2.9) & 12.92 (3) \\
@@
 & & & cpl (sd) & 15.8 (6.5) & 8.29 (2.2) & 7.34 (1.3) \\
- &  & HU & MSE (sd) & 410.11 (343.5) & 178.94 (69.7) & 152.08 (38.1) \\
+ &  & $BIC_+$ & MSE (sd) & 410.11 (343.5) & 178.94 (69.7) & 152.08 (38.1) \\
 & & & cpl (sd) & 15.8 (6.5) & 8.89 (3) & 7.34 (1.3) \\
@@
 \hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & U&	MSE (sd) & 180.67 (52.7) & 173.02 (38.3) & 174.84 (36.8) \\
+50 & 16 & $BIC$ &	MSE (sd) & 180.67 (52.7) & 173.02 (38.3) & 174.84 (36.8) \\
 & & & cpl (sd) & 13.56 (3.7) & 12.78 (2.4) & 13.07 (2.4) \\
@@
 \hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & U&MSE (sd) & 188.91 (88.5) & 124.38 (17.4) & 122.38 (16.7) \\
+50 & 32 & $BIC$ &MSE (sd) & 188.91 (88.5) & 124.38 (17.4) & 122.38 (16.7) \\
 & & & cpl (sd) & 14.74 (6.6) & 8.07 (1.6) & 7.65 (1.2) \\
@@
 \hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & U&  MSE (sd) & 128.18 (13.7) & 127.43 (13.7) & 127.71 (13.3) \\
+100 & 16 & $BIC$ &  MSE (sd) & 128.18 (13.7) & 127.43 (13.7) & 127.71 (13.3) \\
 & & &  cpl (sd) & 13.78 (2.8) & 13.51 (2) & 13.59 (2.1) \\
@@
 \hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & U& MSE (sd) & 126.15 (20.6) & 109.47 (8) & 109.04 (7.8) \\
+100 & 32 & $BIC$ & MSE (sd) & 126.15 (20.6) & 109.47 (8) & 109.04 (7.8) \\
 & & & cpl (sd) & 13.25 (4.4) & 8.3 (1) & 8.12 (0.8) \\
@@
 \hline
-30 & 16 & U& 0.4313 & 2.7769 & 9.96 & 2.45 & 5.95 & 3.5 & 29.42  \\
+30 & 16 & $BIC$ & 0.4313 & 2.7769 & 9.96 & 2.45 & 5.95 & 3.5 & 29.42  \\
 & & &(0.0391) & (0.0883) & (1.3175) & (1.3808) & (1.3056) & (1.453) & (6.2639) \\
- &  & HU & 0.4313 & 4.9079 & 7.28 & 1.25 & 8.63 & 7.38 & 5.77  \\ 
+ &  & $BIC_+$ & 0.4313 & 4.9079 & 7.28 & 1.25 & 8.63 & 7.38 & 5.77  \\ 
 & & &(0.0391) & (0.4353) & (1.5769) & (0.9987) & (1.5548) & (1.6316) & (6.1297) \\ 
@@
 \hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
-30 & 16 & U&MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
+30 & 16 & $BIC$ &MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
 & & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\

### 2014-05-01 www/article13.tex
up ident
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@563 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 So we will try to maximize $\psi(\boldsymbol{X},S)=P(\boldsymbol{X}|S)P(S)$.
-	The discrete structure $S$ is identifiable under some hypothesis detailed in appendix, that is we can't permute some sub-regressions in (\ref{SR}) and obtain the same joint distribution $P(\boldsymbol{X},\boldsymbol{Y})$.
 
@@
 		\end{equation}
-		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed (exponentially distributed for example).
-Gaussian case is just a special case ($K_j=1$) of Gaussian mixture so it is included in our hypothesis but identifiability of $S$ requires to have at least one Gaussian mixture with at least two distinct components in each sub-regression (theorem and proof in appendix \ref{preuveident}).
-Identifiability if $S$ is not necessary to use a given structure but helps to find it.
-
-		Variables in $\boldsymbol{X}_f$ are in the followings supposed to be independent.
+		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust.
+Gaussian case is just a special case ($K_j=1$) of Gaussian mixture so it is included in our hypothesis but identifiability of $S$ requires to have at least one Gaussian mixture with at least two distinct components in each sub-regression (derived from the identifiability of the SR model in \cite{maugis2009variable}, more details in the Appendices \ref{preuveident}).
+
+		Remark:  Identifiability if $S$ is not necessary to use a given structure but helps to find it.
+
+		Variables in $\boldsymbol{X}_f$ are in the followings supposed to be independent Gaussian mixtures with at least two distinct components each.
 	%Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways.
@@
 \section{Appendices}
-	\subsection{Identifiability of the structure}
+	\subsection{Identifiability of the structure} \label{preuveident}
 	The model presented above relies on a discrete structure $S$ between the covariates.
 But to find it we need identifiability property to insure the MCMC will asymptotically find the true model.
-Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions?
+Identifiability of the structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions?
 
@@
 
-	\subsubsection{Identifiability of $\Theta$}		
-		We know by identifiability of the Gaussian mixtures and linear regression ($\boldsymbol{X}_f'\boldsymbol{X}_f$ is of full rank) that if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|S,\tilde{\Theta})$ then $\Theta=\tilde{\Theta}$ up to a permutation of the components.
+%	\subsubsection{Identifiability of $\Theta$}		
+%		We know by identifiability of the Gaussian mixtures and linear regression ($\boldsymbol{X}_f'\boldsymbol{X}_f$ is of full rank) that if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|S,\tilde{\Theta})$ then $\Theta=\tilde{\Theta}$ up to a permutation of the components.
 But the identifiability of the discrete parameter $S$ is still pending.
-
-	\subsubsection{Identifiability of $S$, quick look}
-
-	In the following, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
-	\begin{eqnarray}
-			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
-			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j  \label{densitydroite}\\
-			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
-						\mathcal{GM}(\bigotimes_{\substack{i \in I_f^j  }}\boldsymbol{\pi}_i \ ; 
-			                         \bigoplus_{\substack{i \in I_f^j  }} \alpha_{i,j} \boldsymbol{\mu}_i \ ; 
-			                         \boldsymbol{\sigma}_j^2+\bigoplus_{\substack{i \in I_f^j  }}\alpha_{i,j}^2 \boldsymbol{\sigma}_i^2 )\label{densitygauche}          
-		\end{eqnarray}
-		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
+%		
+%	\subsubsection{Identifiability of $S$, quick look}
+%	
+%	In the following, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
+%	\begin{eqnarray}
+%			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
+%			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j  \label{densitydroite}\\
+%			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
+%						\mathcal{GM}(\bigotimes_{\substack{i \in I_f^j  }}\boldsymbol{\pi}_i \ ; 
+%			                         \bigoplus_{\substack{i \in I_f^j  }} \alpha_{i,j} \boldsymbol{\mu}_i \ ; 
+%			                         \boldsymbol{\sigma}_j^2+\bigoplus_{\substack{i \in I_f^j  }}\alpha_{i,j}^2 \boldsymbol{\sigma}_i^2 )\label{densitygauche}          
+%		\end{eqnarray}
+%		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
 We call this difference in component number "identifiability" in the sense that we try to find the model with the fewest component in $\boldsymbol{X}_f$ (we use the BIC so when comparing two models with the same likelihood, the one with fewer parameters will win).
 So in this case we have a "better" model (the simplest one) in the group of equivalent-meaning models (permuting variables in some sub-regression).
-	\\
-
-	\subsubsection{Identifiability: Theorem and proof} \label{preuveident}
-		A necessary and sufficient condition for identifiability of $S$ is $\forall j \in I_r: K_j\geq 2$ with at least two distinct components in $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$
-\\
-	{\it Proof}:
-	We know that for a Gaussian multivariate distribution $(\boldsymbol{X}_r,\boldsymbol{X}_f)$ the conditional distributions $(\boldsymbol{X}_r|\boldsymbol{X}_f)$ and $(\boldsymbol{X}_f|\boldsymbol{X}_r)$ are both Gaussian (multivariate or not according to $p_r$).
-
-		If $\exists j \in I_r: k_j=1$ (Gaussian) then all the covariates on the right of the associated sub-regression (with non-zero coefficients) are Gaussian and then the sub-regression can be permuted and is not identifiable.
+%	\\
+%	
+If all the variables in $\boldsymbol{X}$ are Gaussian mixtures with at least two distinct components, then the model of sub-regression is identifiable (from the identifiability of the SR model by Maugis \cite{maugis2009variable}).
+
+But making this hypothesis on $\boldsymbol{X}_r$ is sufficient.
+In the followings, we call Gaussian mixture the Gaussian mixtures with at least two distinct components, and Gaussian mixtures with equal components are called Gaussian (in fact they are).
+
+First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there is at least one Gaussian mixture on the right of each sub-regression.
+So there is a non-empty subset of $\boldsymbol{X} \subset \boldsymbol{X}_f$ containing Gaussian mixtures.
+We note $\boldsymbol{X}_{S^c}$ its complement (empty or not) in $\boldsymbol{X}_f$
+
+The theorem from Maugis guarantee that a sub-regression between Gaussian mixtures is identifiable.
+		For each sub-regression, 
+%	\subsubsection{Identifiability: Theorem and proof} 
+%		A necessary and sufficient condition for identifiability of $S$ is $\forall j \in I_r: K_j\geq 2$ with at least two distinct components in $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$
+%\\
+%	{\it Proof}:
+%	We know that for a Gaussian multivariate distribution $(\boldsymbol{X}_r,\boldsymbol{X}_f)$ the conditional distributions $(\boldsymbol{X}_r|\boldsymbol{X}_f)$ and $(\boldsymbol{X}_f|\boldsymbol{X}_r)$ are both Gaussian (multivariate or not according to $p_r$).
+%	
+%		If $\exists j \in I_r: k_j=1$ (Gaussian) then all the covariates on the right of the associated sub-regression (with non-zero coefficients) are Gaussian and then the sub-regression can be permuted and is not identifiable.
 So if $\boldsymbol{X}^j$ are Gaussian mixtures, it is a necessary condition for identifiability to have $\forall j \in I_r: K_j\geq 2$ with at least two distinct components in $(\boldsymbol{\mu}_j,\boldsymbol{\sigma}_j)$.
-
-	Gaussian mixtures in $\boldsymbol{X_r}$ with at least two distinct components does imply the presence of at least one Gaussian mixture in each sub-regression (with at least two components).
+%
+%	Gaussian mixtures in $\boldsymbol{X_r}$ with at least two distinct components does imply the presence of at least one Gaussian mixture in each sub-regression (with at least two components).
+%%	\begin{equation}
+%%	 \forall j \in I_f :    \boldsymbol{X}^j_{|S}=\sum_{k=1}^{K_j}\pi_j(k)\mathcal{N}\left(\mu_j(k),\sigma_j^2(k)\boldsymbol{I}_n\right) \\
+%%	\end{equation}
+%%	$\forall j \in I_r$ we can define $K_{j}$ the number of distinct components of $\boldsymbol{X}_j$.
+%%	\begin{eqnarray}
+%%				\boldsymbol{X}^{j}_{|S}&=&\sum_{k=1}^{K_j}\pi_j(k)\mathcal{N}\left(\mu_j(k),\sigma_j^2(k)\boldsymbol{I}_n\right)
+%%	\end{eqnarray}
+%	The joint distribution of the $p$ Gaussian mixtures is a multivariate Gaussian mixture $\boldsymbol{X}$ with $K\geq 2$ distinct components $(\boldsymbol{m}_k,\boldsymbol{\Sigma}_k)$ where $\boldsymbol{m}_k \in \mathbf{R}^p$ and $\boldsymbol{\Sigma}_k$ is a $p \times p$ matrix.
+%		We decompose the joint distribution based on the components:
+%		
 %		\begin{equation}
-%	 \forall j \in I_f :    \boldsymbol{X}^j_{|S}=\sum_{k=1}^{K_j}\pi_j(k)\mathcal{N}\left(\mu_j(k),\sigma_j^2(k)\boldsymbol{I}_n\right) \\
+%			f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{k=1}^K p_kf(\boldsymbol{X}_r,\boldsymbol{X}_f|k)
 %	\end{equation}
-%	$\forall j \in I_r$ we can define $K_{j}$ the number of distinct components of $\boldsymbol{X}_j$.
-%	\begin{eqnarray}
-%				\boldsymbol{X}^{j}_{|S}&=&\sum_{k=1}^{K_j}\pi_j(k)\mathcal{N}\left(\mu_j(k),\sigma_j^2(k)\boldsymbol{I}_n\right)
-%	\end{eqnarray}
-	The joint distribution of the $p$ Gaussian mixtures is a multivariate Gaussian mixture $\boldsymbol{X}$ with $K\geq 2$ distinct components $(\boldsymbol{m}_k,\boldsymbol{\Sigma}_k)$ where $\boldsymbol{m}_k \in \mathbf{R}^p$ and $\boldsymbol{\Sigma}_k$ is a $p \times p$ matrix.
-		We decompose the joint distribution based on the components:
-
-		\begin{equation}
-			f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{k=1}^K p_kf(\boldsymbol{X}_r,\boldsymbol{X}_f|k)
-	\end{equation}
-			where  $\forall 1\leq k \leq K$, $ f(\boldsymbol{X}_r,\boldsymbol{X}_f|k)$  is a multivariate Gaussian.
-			\\
-			Then,  $\forall j \in I_r, \forall i \in I_f^j$ we have:
-			\begin{equation}
-			 f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{k=1}^K p_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)f(\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i|k)
-	\end{equation}
-				where the $f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)$ are Gaussian.
-
-Hence,
-\begin{equation}
-				f(\boldsymbol{X}^i|\boldsymbol{X}\setminus\boldsymbol{X}^i)=f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i)=\sum_{k=1}^Kp_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)
-				\end{equation}
-
-				is a Gaussian mixture.
-
+%			where  $\forall 1\leq k \leq K$, $ f(\boldsymbol{X}_r,\boldsymbol{X}_f|k)$  is a multivariate Gaussian.
+%			\\
+%			Then,  $\forall j \in I_r, \forall i \in I_f^j$ we have:
+%			\begin{equation}
+%			 f(\boldsymbol{X}_r,\boldsymbol{X}_f)= \sum_{k=1}^K p_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)f(\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i|k)
+%	\end{equation}
+%				where the $f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)$ are Gaussian.
+%				
+%Hence,
+%\begin{equation}
+%				f(\boldsymbol{X}^i|\boldsymbol{X}\setminus\boldsymbol{X}^i)=f(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i)=\sum_{k=1}^Kp_kf(\boldsymbol{X}^i|\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i,k)
+%				\end{equation}
+%				
+%				is a Gaussian mixture.				
+%				
 %				Identifiability of $S$ means we don't have a linear regression of $\boldsymbol{X}^i$ on $(\boldsymbol{X}_r,\boldsymbol{X}_f\setminus\boldsymbol{X}^i)$.
@@
 %				
-				If it is a Gaussian mixture with at least two distinct components when $\boldsymbol{X}^j \in \boldsymbol{X}_r$ has at least two distinct components.
-				Then $\boldsymbol{X}^i$ can't be explained by a linear regression on the other covariates (non-Gaussian residual).		
-				 So we have identifiability of the structure $S$ even for trivial sub-regressions if we keep the constraint to have distinct components in $\boldsymbol{X}_r$ (equivalent to have at least one Gaussian mixture with at least two components in each sub-regression).
+%				If it is a Gaussian mixture with at least two distinct components when $\boldsymbol{X}^j \in \boldsymbol{X}_r$ has at least two distinct components.
+%				Then $\boldsymbol{X}^i$ can't be explained by a linear regression on the other covariates (non-Gaussian residual).		
+%				 So we have identifiability of the structure $S$ even for trivial sub-regressions if we keep the constraint to have distinct components in $\boldsymbol{X}_r$ (equivalent to have at least one Gaussian mixture with at least two components in each sub-regression).
 %			%f(\boldsymbol{X}_r,\boldsymbol{X}_f)&=& sum_{j=1}^K 

### 2014-05-01 www/article13.tex
up ident
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@564 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there is at least one Gaussian mixture on the right of each sub-regression.
-So there is a non-empty subset of $\boldsymbol{X} \subset \boldsymbol{X}_f$ containing Gaussian mixtures.
-We note $\boldsymbol{X}_{S^c}$ its complement (empty or not) in $\boldsymbol{X}_f$
+So we have $\boldsymbol{X}^G \subsetneq \boldsymbol{X}_f$ containing Gaussian variables and we note the Gaussian mixtures $\boldsymbol{X}^{G^c}\neq \emptyset$ its complement in $\boldsymbol{X}_f$.
 
 The theorem from Maugis guarantee that a sub-regression between Gaussian mixtures is identifiable.
-		For each sub-regression, 
+		\begin{eqnarray}
+			\forall j \in I_r, \boldsymbol{X}^j_{|\boldsymbol{X}^G,\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^G\boldsymbol{\alpha}_j^G+\boldsymbol{X}^{G^c}\boldsymbol{\alpha}_j^{G^c}+ \boldsymbol{\varepsilon}_j \\
+			\boldsymbol{X}^j_{|\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^{G^c}\boldsymbol{\alpha}_j^{G^c} + \tilde{\boldsymbol{\varepsilon}}_j \textrm{ is identifiable where} \\
+			\tilde{\boldsymbol{\varepsilon}}_j&=&\boldsymbol{X}^{G}\boldsymbol{\alpha}_j^{G}+ \boldsymbol{\varepsilon}_j \textrm{ is Gaussian. And we have the unique decomposition} \\
+			f(\boldsymbol{X}^j|\boldsymbol{X}^G,\boldsymbol{X}^{G^c})&=&f(\boldsymbol{X}^j|\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G})
+		\end{eqnarray}
+		We have the unique (Gaussian are allowed only on the right of the sub-regressions) decomposition $f(\boldsymbol{X}_r|\boldsymbol{X}_f)=f(\boldsymbol{X}_r|\boldsymbol{X}^G,\boldsymbol{X}^{G^c})=f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G}) $
+		where  $f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})$ is identifiable. Moreover, $\boldsymbol{X}^{G}$ is a set (empty or not) of independent Gaussian variables, whose parameters are identifiable.
+		So $\forall j \in I_r, f(\boldsymbol{X}^j|\boldsymbol{X}_f)$. And the structure is identifiable.
+
 %	\subsubsection{Identifiability: Theorem and proof} 

### 2014-05-02 www/article13.tex
up intro
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@565 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	increasing the chance to have correlated ones.  
-It will be illustrated with examples on real datasets in the last section of this paper (steel industry: many parameters of the whole process without any a priori and highly correlated because of physical laws, process rules, etc).
 	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
@@
 	The Ordinary Least Squares (\textsc{OLS}) estimator is the minimum-variance linear unbiased estimator but it suffers from ill-conditioned matrices inversion when the covariates are highly correlated (section \ref{sectionOLS}).
-Therefore, some other methods can be seen as a modification of the dataset to improve the robustess of the estimator (by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}).
+Therefore, some other methods try to improve the robustess of the estimator by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}.
 
@@
 	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large number of covariates.
-
-	Real datasets often imply many irrelevant variables so variable selection would be favoured when possible. Moreover, selection may keep only uncorrelated covariates and thus resolve the problem.	
+The absence of variable selection is not compatible with the need to find small set of relevant covariates to explain the response variable.
+
+	Real datasets often imply many irrelevant variables so variable selection would be favoured when possible to reduce the dimension and obtain models small enough to be understood.
+Moreover, selection may keep only uncorrelated covariates and thus resolve the correlations problem.
 	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a parametric $L_1$ penalty to shrink some coefficients exactly to zero.
@@
 But like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates.
+It only produces a parsimonious model, that is a gain for interpretation but only half the way.
 %	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
@@
 But
-	 \textsc{LASSO} is also known to face consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates.
+	 Indeed, \textsc{LASSO} is also known to face consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates. So the quality of interpretation is compromised.
 	 \\
@@
 Once again, nothing specifically aims to reduce the correlations.
-%Hence, when comparing the two models, interpretations are not the same and nothing explicitly explains why.
-So it can be very confusing.
+Interpretation will be misleading in both LASSO and Elastic net cases because nothing differentiates correlated and uncorrelated covariates.
+	%Hence, when comparing the two models, interpretations are not the same and nothing explicitly explains why. So it can be very confusing. 
 	\\
 
-	Another way of improving the conditioning is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}) to reduce the dimension and also the correlations if correlated covariates are in the same clusters.
+	Another way of improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}) to reduce the dimension and also the correlations if correlated covariates are in the same clusters.
 	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the regression coefficients no longer as fixed effect parameters but as unobserved independent random variables with grouped coefficients following a Gaussian Mixture distribution.
@@
 	 Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
-	We see that none of the above methods takes explicitly the correlations into account.
+	We see that none of the above methods takes explicitly the correlations into account, even if the clustering methods introduce a kind of relationship between the covariates.
 
@@
  Instead of reducing the dimension without looking at correlations, Selvarclust \cite{maugis2009variable} proposes to focus on the correlations, giving an explicit structure between covariates by linear sub-regressions between them.
-But is in a clustering context, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even it is known to be often unstable \cite{miller2002subset}.
+But it is made in a clustering context, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even it is known to be often unstable \cite{miller2002subset}.
 We propose to develop this method for linear regression with a specifically adapted algorithm to find the structure of sub-regression.
@@
 This pretreatment is specifically done to decorrelate the covariates, unlike methods described above.
-The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint distribution on the covariates.
+The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint distribution on the covariates, independently from the response variable.
 This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{CRAN}.
@@
 This work takes place in steel industry context, with quality oriented objective : to understand and prevent quality problems on finished product, knowing the whole process.
+The correlations are strong here (many parameters of the whole process without any a priori and highly correlated because of physical laws, process rules, {\it etc.}).
 
@@
 	We have seen that correlations can lead to serious estimation and variable selection problems in linear regression and that in such a context, it can be useful to explicitly model the structure between the covariates and to use this structure (even sequentially) to avoid correlations issues.
-We also show that real industrial context faces this kind of situations so our model can help to interprete and predict physical phenomenon efficiently and to help to manage missing values.
+We also show that real industrial context faces this kind of situations so our model can help to understand and predict physical phenomenon efficiently.
 But for now we still need a full dataset to learn the structure between the covariates and even if correlations are strong, some information is lost.
@@
 	Nevertheless, we need to enlarge its application field to missing values, also very commons in industry.
-The actual generative model allows such a functionality without supplementary hypothesis and this also is a strength of \textsc{CorReg}.
+The structure can be used to estimate missing values in $\boldsymbol{X}_r$ but the actual generative model allows to go further (to manage missing values even in the MCMC algorithm) without supplementary hypothesis and this also is a strength of \textsc{CorReg}.
 
@@
 
-First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there is at least one Gaussian mixture on the right of each sub-regression.
+First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there must be at least one Gaussian mixture on the right of each sub-regression.
 So we have $\boldsymbol{X}^G \subsetneq \boldsymbol{X}_f$ containing Gaussian variables and we note the Gaussian mixtures $\boldsymbol{X}^{G^c}\neq \emptyset$ its complement in $\boldsymbol{X}_f$.
@@
 		where  $f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})$ is identifiable. Moreover, $\boldsymbol{X}^{G}$ is a set (empty or not) of independent Gaussian variables, whose parameters are identifiable.
-		So $\forall j \in I_r, f(\boldsymbol{X}^j|\boldsymbol{X}_f)$. And the structure is identifiable.
+		So the structure is identifiable.
 

### 2014-05-06 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@566 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I}_n)$ the noise of the regression, with $\boldsymbol{I}_n$ the $n$-sized identity matrix and $\sigma_Y >0$.
-The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{OLS}).
-%As shown in section \ref{sectionOLS},
+The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{OLS}): %As shown in section \ref{sectionOLS},
+	\begin{equation}
+		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
+	\end{equation}
+	with variance matrix
+	\begin{equation}
+		\operatorname{Var}(\hat{\boldsymbol{\beta}}_{OLS})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
+	\end{equation}
+	and without any bias.
 	Estimation of $\boldsymbol{\beta}$ requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
-Conditionning of $\boldsymbol{X}'\boldsymbol{X}$ get worse based on two aspects : the dimension $p$ (number of covariates) of the model (the more covariates you have the greater variance you get)
+Conditionning of $\boldsymbol{X}'\boldsymbol{X}$ get worse based on two aspects: the dimension $p$ (number of covariates) of the model (the more covariates you have the greater variance you get)
 	 and the correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
+	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
+	Another problem is that matrix inversion requires $n\geq p$. 	
+As a running example, we look at a simple case with $p=5$ variables defined by four independent scaled Gaussian $\mathcal{N}(0,1)$ named $\boldsymbol{x}_1,\boldsymbol{x}_2$ and $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$.
+We also define another couple $\boldsymbol{x}_4,\boldsymbol{x}_5$ of covariates that are {\it i.i.d.
+} with $(\boldsymbol{x}_1,\boldsymbol{x}_2)$ and two {\it scenarii} for $\boldsymbol{Y}$ with $\boldsymbol{\beta}=(1,1,1,1,1)$ and $\sigma_Y \in \{10,20\}$ .
+It is clear that $\boldsymbol{X}'\boldsymbol{X}$ will become more ill-conditioned as $\sigma_3$ gets smaller.
 
@@
  Then $\boldsymbol{X}_r$ is the $n\times p_r$ submatrix of $0\leq p_r <p$ redundent covariates and $\boldsymbol{X}_f$ the $n\times (p-p_r)$ submatrix of the free (independent) covariates.
+
+In our running example, $\boldsymbol{X}_r=\boldsymbol{x}_3$ and $\boldsymbol{X}_f=\{\boldsymbol{x}_1,\boldsymbol{x}_2,\boldsymbol{x}_4,\boldsymbol{x}_5 \}$
+
 In the following, we note $\boldsymbol{X}^j$ the $j^{th}$ column of $\boldsymbol{X}$.
@@
 		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_r)}$ are the sparse vectors of the regression coefficients between the covariates (each sub-regression freely implies different covariates).
+In our running example, $p_r=1$ and $\boldsymbol{\alpha}_3=(1,1,0,0)'$
 \\
@@
 
-The partition of $\boldsymbol{X}$ implies the uncrossing rule % $I_f\cap I_r=\emptyset$ 
+The partition of $\boldsymbol{X}$ implies the uncrossing rule  $\boldsymbol{X}_r \cap \boldsymbol{X}_f$ 
 {\it i.e.} endogenous variables don't explain other covariates.
@@
 It is a variable pre-selection independent of the response $\boldsymbol{Y}$.
+\\
+Our running example gives here $\boldsymbol{Y}= 2\boldsymbol{x}_1+2\boldsymbol{x}_2+\boldsymbol{x}_4+\boldsymbol{x}_5+\boldsymbol{\varepsilon}_3 +\boldsymbol{\varepsilon}_Y$
 \subsection{Strategy of use: pretreatment before classical estimation/selection methods}\label{interpretation}
 
-As a pretreatment, it allows usage of any method in a second time to estimate $\boldsymbol{\beta}^*$, even with variable selection methods like LASSO or even stepwise \cite{seber2012linear}.
+As a pretreatment, it allows usage of any method in a second time to estimate $\boldsymbol{\beta}^*$, even with variable selection methods like LASSO or stepwise \cite{seber2012linear}.
 
@@
 	\subsection{Illustration of the tradeoff conveyed by the pretreatment}	
-	As an example, we look at the OLS estimator on $\boldsymbol{X}$ with and without the pretreatment that is $\boldsymbol{X}_f$ selection.
-For a model defined by (\ref{regressionsimple}), we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
-	\begin{equation}
-		\boldsymbol{\hat{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}
-	\end{equation}
-	with variance matrix
-	\begin{equation}
-		\operatorname{Var}(\hat{\boldsymbol{\beta}}_{OLS})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1} \label{eqOLS}
-	\end{equation}
-	and without any bias.
-	When correlations between covariates are strong, the matrix to invert is ill-conditioned and the variance increases, giving unstable and unusable estimator \cite{hoerl1970ridge}.
-	Another problem is that matrix inversion requires $n\geq p$.
+	We compare the OLS estimator on $\boldsymbol{X}$ defined in section \ref{sectionOLS} with the estimator obtained by the pretreatment that is $\boldsymbol{X}_f$ selection.
 
@@
 	\end{eqnarray}	 
-	To better illustrate the bias-variance tradeoff, we look at a simple example with $p=3$ variables.
-$\boldsymbol{X}_f$ is composed by two independent scaled Gaussian $\mathcal{N}(0,1)$, $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$.
-We also have $\boldsymbol{\beta}=(1,1,1)$ and $\sigma_Y \in \{1,10\}$  .
-Then we observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg}'s marginal  model for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
+	To better illustrate the bias-variance tradeoff, we look at the running example.
+We observe the theoretical Mean Squared Error (MSE) of the estimator of both OLS and \textsc{CorReg}'s marginal  model for several values of $\sigma_3$ (strength of the sub-regression) and $n$.
 Figure \ref{MQE1} shows the theoretical MSE evolution with the strength of the sub-regression:
@@
 %   \end{minipage} 
-	\includegraphics[width=500px]{figures/MQEexplOLS.png}\label{MQE1}
+	\includegraphics[width=500px]{figures/MQEexplOLSp5.png}\label{MQE1}
 	\caption{MSE of OLS and CorReg (dotted) estimators for varying $(1-R^2)$ of the sub-regression, $n$ and $\sigma_Y$.}
@@
 We see that $\sigma_Y$ multiplies $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}))=\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{f}))+\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))$ for both models but for the marginal model $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$.
- Thus, as $\operatorname{Tr}(\operatorname{Var}(\hat{\boldsymbol{\beta}}_{r}))=0$ when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
+ Thus, when $\sigma_Y^2$ rises it increases the advantage of \textsc{CorReg} versus \textsc{OLS}.
 It illustrates the importance of dimension reduction when the model has a strong noise (very usual case on real datasets where true model is not even exactly linear).
@@
 
-		Remark:  Identifiability if $S$ is not necessary to use a given structure but helps to find it.
+		Remark:  Identifiability of $S$ is not necessary to use a given structure but helps to find it.
 
@@
 
-Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)\propto P(\boldsymbol{X}|S)$ so it is equivalent to a maximization of the $BIC$.
+Uniform law on $P(S)$ gives $\psi(\boldsymbol{X},S)\propto P(\boldsymbol{X}|S)$ so it is equivalent to a minimization of the $BIC$.
 	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
@@
 Then we have the explicit structure characterized by $S=(I_f,I_r,p_f,p_r)$ where $p_r=|I_r|$, $\boldsymbol{p}_f=(p_f^1,\dots,p_f^{p_r})$ is the vector of the number of covariates in each sub-regression  and $p_f^j=|I_f^j|$, with $|.|$ the cardinal of an ensemble.
+Our running example is then described by $S= \left( \{ \{1,2\}\},\{3\},(2),(1)\right)$
 \\
@@
 		\begin{eqnarray}
-		BIC_+(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
+		BIC_+(X|S)&=&BIC(X|S) -\ln(P(S)) \label{Bicstar}
 	\end{eqnarray}		

### 2014-05-06 pkg/CorReg/R/ProbaZ.R
up comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@567 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -31,8 +31,10 @@ ProbaZ<-function(Z=NULL,p=NULL,proba=FALSE,star=TRUE){
       }
       nb=1#modele vide
       if(p>1){
-         #calcul du nombre de modèles
-         for (i in 1:(p-1)){
+         #calcul du nombre de modeles
+         for (i in 1:(p-1)){#pour chaque nombre de sous-regression possible
+            #choix de qui est à gauche, puis ayant une partition, tout devient possible à droite 
+            #(sauf le cas vide qui enfreindrait le nombre de sous-regression donc on fait -1)
             nb=nb+choose(p,i)*(2^(p-i)-1)^i
          }
       }

### 2014-05-06 pkg/CorReg/R/searchZ.R
up comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@567 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -29,8 +29,8 @@ searchZ<-function(X=X,Z=NULL,Bic_null_vect=NULL,candidates=-1,reject=0,methode=1
     p2max=ncol(X)+1 
   }
   if(star){
-     p2max=min(p2max,ncol(X)/2)
-     p1max=min(p1max,ncol(X)/2)
+     p2max=floor(min(p2max,ncol(X)/2))
+     p1max=floor(min(p1max,ncol(X)/2))
   }
   if(is.null(nb_opt_max)){
     nb_opt_max=Maxiter

### 2014-05-06 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@568 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@569 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-		Variables in $\boldsymbol{X}_f$ are in the followings supposed to be independent Gaussian mixtures with at least two distinct components each.
+		Variables in $\boldsymbol{X}$ are in the followings supposed to be independent Gaussian mixtures with at least two distinct components each. 
 	%Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways.
@@
 	\end{eqnarray}		
-	It increases penalty on complexity for $p_r<\frac{p}{2}$ and $p_f^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
+	It increases penalty on complexity for $p_r\leq\frac{p}{2}$ and $p_f^j\leq\frac{p}{2}$ . Hence %when using $BIC*$ 
 	this constraint on $\hat{p}_r$ and $\hat{p}_f^j$ is given in the research algorithm when the Hierarchical Uniform hypothesis is made instead of Uniform one in numerical experiments (section \ref{sectionsimul} and \ref{sectionrealcase}).

### 2014-05-06 www/article13.tex
up comment
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@570 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 & & & (0.0275) & (0.3711) & (0) & (1.9346) & (0) & (1.9346) & (8.0884) \\
- &  &HU &0.4104 & 8.2338 & 0 & 0.53 & 0 & -0.53 & 2.27  \\
+ &  & $BIC_+$ &0.4104 & 8.2338 & 0 & 0.53 & 0 & -0.53 & 2.27  \\
 & & & (0.0275) & (0.9045) & (0) & (0.7844) & (0) & (0.7844) & (3.3024) \\
@@
 & & & (0.0329) & (0.1498) & (1.9844) & (2.0313) & (1.9948) & (0.9987) & (6.499) \\
- &  &HU &0.4182 & 4.1876 & 11.61 & 4.57 & 4.33 & -0.24 & 16.48  \\
+ &  & $BIC_+$ &0.4182 & 4.1876 & 11.61 & 4.57 & 4.33 & -0.24 & 16.48  \\
 & & & (0.0329) & (0.2178) & (1.8743) & (1.9502) & (1.8752) & (0.4948) & (5.4892) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & 0.4456 & 2.9154 & 25.23 & 1.92 & 6.5 & 4.58 & 28  \\
+30 & 32 & $BIC$ & 0.4456 & 2.9154 & 25.23 & 1.92 & 6.5 & 4.58 & 28  \\
 & & & (0.0429) & (0.1331) & (1.4761) & (1.0888) & (1.4668) & (0.9866) & (5.0831) \\
@@
 & & & (0.0451) & (0.3123) & (1.93) & (1.9132) & (1.886) & (0.7317) & (5.7742) \\
- &  &HU &0.5205 & 4.909 & 11.42 & 4.59 & 4.45 & -0.14 & 7.55  \\
+ &  & $BIC_+$ &0.5205 & 4.909 & 11.42 & 4.59 & 4.45 & -0.14 & 7.55  \\
 & & & (0.0451) & (0.4556) & (1.9079) & (1.8968) & (1.8333) & (0.3487) & (4.0611) \\
@@
 & & & (0.0628) & (0.316) & (1.3711) & (0.9077) & (1.3143) & (0.8394) & (4.3359) \\
- &  &HU &0.5833 & 4.3599 & 17.27 & 2.73 & 14.6 & 11.87 & -2.61  \\
+ &  & $BIC_+$ &0.5833 & 4.3599 & 17.27 & 2.73 & 14.6 & 11.87 & -2.61  \\
 & & & (0.0628) & (0.3729) & (1.1708) & (1.1708) & (1.2792) & (0.338) & (4.4854) \\
@@
 & & & (0.077) & (1.7778) & (0) & (1.2953) & (0) & (1.2953) & (3.1999) \\
- &  &HU & 0.9623 & 20.9817 & 0 & 0.01 & 0 & -0.01 & 0.02  \\
+ &  & $BIC_+$ & 0.9623 & 20.9817 & 0 & 0.01 & 0 & -0.01 & 0.02  \\
 & & & (0.077) & (1.9421) & (0) & (0.1) & (0) & (0.1) & (0.2) \\
@@
 & & & (0.1122) & (0.5473) & (2.0003) & (2.0646) & (1.956) & (0.7833) & (3.9471) \\
- &  &HU & 1.1223 & 8.8486 & 12.04 & 3.95 & 3.88 & -0.07 & 3.75  \\
+ &  & $BIC_+$ & 1.1223 & 8.8486 & 12.04 & 3.95 & 3.88 & -0.07 & 3.75  \\
 & & & (0.1122) & (0.7174) & (1.9223) & (1.9404) & (1.9137) & (0.2564) & (2.2625) \\
@@
 & & & (0.2528) & (0.3136) & (1.3928) & (0.8495) & (1.2941) & (0.7086) & (3.0975) \\
- &  &HU & 1.4343 & 7.3741 & 17.49 & 2.51 & 14.26 & 11.75 & -3.76  \\
+ &  & $BIC_+$ & 1.4343 & 7.3741 & 17.49 & 2.51 & 14.26 & 11.75 & -3.76  \\
 & & & (0.2528) & (0.2748) & (1.1849) & (1.1849) & (1.2441) & (0.4794) & (4.4859) \\
@@
 & & & cpl (sd) & 30 (0) & 29.99 (0.1) & 30 (0) \\
- &  &HU &MSE (sd) & 262627.57 (732019) & 10381962.64 (90962496.9) & 262627.57 (732019) \\
+ &  & $BIC_+$ &MSE (sd) & 262627.57 (732019) & 10381962.64 (90962496.9) & 262627.57 (732019) \\
 & & & cpl (sd) & 30 (0) & 30 (0) & 30 (0) \\
@@
 & & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
- &  &HU &MSE (sd) & 510747.53 (2287539.8) & 603.24 (415.6) & 610.32 (424.9) \\
+ &  & $BIC_+$ &MSE (sd) & 510747.53 (2287539.8) & 603.24 (415.6) & 610.32 (424.9) \\
 & & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25 (0) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 178323.95 (1426610.2) & 180.02 (44.9) & 141.03 (27.8) \\
+30 & 32 & $BIC$ & MSE (sd) & 178323.95 (1426610.2) & 180.02 (44.9) & 141.03 (27.8) \\
 & & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
- &  &HU &MSE (sd) & 612.72 (291.5) & 207.6 (51.3) & 200.32 (42.6) \\
+ &  & $BIC_+$ &MSE (sd) & 612.72 (291.5) & 207.6 (51.3) & 200.32 (42.6) \\
 & & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
- &  &HU &MSE (sd) & 555.44 (262.5) & 171.9 (31.1) & 121.08 (11.9) \\
+ &  & $BIC_+$ &MSE (sd) & 555.44 (262.5) & 171.9 (31.1) & 121.08 (11.9) \\
 & & &cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 38.17 (1.3) & 41 (0) \\
- &  &HU &MSE (sd) & 167.71 (20.6) & 167.98 (20.8) & 167.71 (20.6) \\  
+ &  & $BIC_+$ &MSE (sd) & 167.71 (20.6) & 167.98 (20.8) & 167.71 (20.6) \\  
 & & & cpl (sd) & 41 (0) & 40.99 (0.1) & 41 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
- &  &HU &  MSE (sd) & 168.68 (22.4) & 137.03 (20.3) & 133.49 (12.2) \\
+ &  & $BIC_+$ &  MSE (sd) & 168.68 (22.4) & 137.03 (20.3) & 133.49 (12.2) \\
 & & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
- &  &HU & MSE (sd) & 173.25 (22.9) & 127.4 (11.9) & 110.63 (6.9) \\
+ &  & $BIC_+$ & MSE (sd) & 173.25 (22.9) & 127.4 (11.9) & 110.63 (6.9) \\
 & & & cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
@@
 & & & cpl (sd) & 17.84 (5.5) & 15.79 (5.6) & 17.84 (5.5) \\
- &  &HU &MSE (sd) & 1246.35 (350.5) & 1248.84 (341.4) & 1246.35 (350.5) \\
+ &  & $BIC_+$ &MSE (sd) & 1246.35 (350.5) & 1248.84 (341.4) & 1246.35 (350.5) \\
 & & & cpl (sd) & 17.84 (5.5) & 17.63 (5.6) & 17.84 (5.5) \\ 
 \hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 &16& U & MSE (sd) & 712.79 (405.2) & 566.32 (226.4) & 554.01 (257.1) \\
+30 &16& $BIC$ & MSE (sd) & 712.79 (405.2) & 566.32 (226.4) & 554.01 (257.1) \\
 & & & cpl (sd) & 16.68 (4.4) & 15.35 (4.3) & 16.14 (4.5) \\
@@
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 216.12 (128.3) & 157.78 (39.4) & 147.04 (28.4) \\
+30 & 32 & $BIC$ & MSE (sd) & 216.12 (128.3) & 157.78 (39.4) & 147.04 (28.4) \\
 & & & cpl (sd) & 11.23 (4.4) & 8.56 (1.8) & 7.91 (1.3) \\
@@
 & & & cpl (sd) & 22.14 (4.1) & 19.52 (2.9) & 20.47 (2.8) \\
- &  &HU &MSE (sd) & 274.34 (101.7) & 233.8 (78.9) & 225.64 (67.5) \\
+ &  & $BIC_+$ &MSE (sd) & 274.34 (101.7) & 233.8 (78.9) & 225.64 (67.5) \\
 & & & cpl (sd) & 22.14 (4.1) & 20.19 (2.8) & 20.47 (2.8) \\
@@
 & & & cpl (sd) & 12.93 (6.4) & 8.51 (1.5) & 8.09 (1.1) \\
- &  &HU &MSE (sd) & 165.6 (73.7) & 135.84 (24.1) & 123.73 (12.8) \\
+ &  & $BIC_+$ &MSE (sd) & 165.6 (73.7) & 135.84 (24.1) & 123.73 (12.8) \\
 & & &cpl (sd) & 12.93 (6.4) & 10.17 (3.1) & 8.09 (1.1) \\
@@
 & & & cpl (sd) & 37.78 (2.4) & 32.93 (3.7) & 37.78 (2.4) \\
- &  &HU &  MSE (sd) & 183.33 (31.1) & 183.78 (31.6) & 183.33 (31.1) \\
+ &  & $BIC_+$ &  MSE (sd) & 183.33 (31.1) & 183.78 (31.6) & 183.33 (31.1) \\
 & & &  cpl (sd) & 37.78 (2.4) & 37.75 (2.4) & 37.78 (2.4) \\
@@
 & & & cpl (sd) & 25.22 (4) & 21.72 (2) & 22.3 (1.7) \\
- &  &HU &  MSE (sd) & 148.83 (18.6) & 142.97 (22.1) & 139.57 (16.1) \\
+ &  & $BIC_+$ &  MSE (sd) & 148.83 (18.6) & 142.97 (22.1) & 139.57 (16.1) \\
 & & & cpl (sd) & 25.22 (4) & 22.24 (1.8) & 22.3 (1.7) \\
@@
 & & & cpl (sd) & 13.18 (5.9) & 8.68 (1.2) & 8.53 (1) \\
- &  &HU & MSE (sd) & 124.29 (19.9) & 118.33 (12) & 111.54 (7.2) \\
+ &  & $BIC_+$ & MSE (sd) & 124.29 (19.9) & 118.33 (12) & 111.54 (7.2) \\
 & & & cpl (sd) & 13.18 (5.9) & 11.02 (2.9) & 8.53 (1) \\
@@
 & & & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
- &  &HU & MSE (sd) & 1326.54 (388.4) & 1307.96 (356.6) & 1326.54 (388.4) \\
+ &  & $BIC_+$ & MSE (sd) & 1326.54 (388.4) & 1307.96 (356.6) & 1326.54 (388.4) \\
 & & & cpl (sd) & 12.14 (5.1) & 12.27 (5.1) & 12.14 (5.1) \\
@@
 & & & cpl (sd) & 13.86 (6.8) & 14.31 (5.7) & 14.83 (5.3) \\
- &  &HU &MSE (sd) & 1 400.56 (1598.2) & 643.25 (277.9) & 653.59 (283.7) \\
+ &  & $BIC_+$ &MSE (sd) & 1 400.56 (1598.2) & 643.25 (277.9) & 653.59 (283.7) \\
 & & & cpl (sd) & 13.86 (6.8) & 14.83 (5.6) & 14.83 (5.3) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 855.74 (582.6) & 181.08 (51.9) & 146.79 (32.7) \\
+30 & 32 & $BIC$ & MSE (sd) & 855.74 (582.6) & 181.08 (51.9) & 146.79 (32.7) \\
 & & & cpl (sd) & 15.57 (6.3) & 11.19 (2.4) & 8.37 (1.3) \\
@@
 & & & cpl (sd) & 26.72 (7.4) & 20.49 (3.8) & 21.8 (3.3) \\
- &  &HU &MSE (sd) & 516.36 (226.6) & 239.77 (103.7) & 228.85 (89.3) \\
+ &  & $BIC_+$ &MSE (sd) & 516.36 (226.6) & 239.77 (103.7) & 228.85 (89.3) \\
 & & & cpl (sd) & 26.72 (7.4) & 21.57 (3.4) & 21.8 (3.3) \\
@@
 & & & cpl (sd) & 24.7 (7.1) & 9.95 (1.5) & 8.22 (1.2) \\
- &  &HU &MSE (sd) & 328.7 (146.3) & 169.13 (40.2) & 124.01 (15.4) \\
+ &  & $BIC_+$ &MSE (sd) & 328.7 (146.3) & 169.13 (40.2) & 124.01 (15.4) \\
 & & &cpl (sd) & 24.7 (7.1) & 16.04 (3.8) & 8.22 (1.2) \\
@@
 & & & cpl (sd) & 39.71 (1.7) & 34.11 (4.6) & 39.71 (1.7) \\
- &  &HU &  MSE (sd) & 175.93 (25.9) & 176.2 (26.3) & 175.93 (25.9) \\
+ &  & $BIC_+$ &  MSE (sd) & 175.93 (25.9) & 176.2 (26.3) & 175.93 (25.9) \\
 & & &  cpl (sd) & 39.71 (1.7) & 39.69 (1.7) & 39.71 (1.7) \\
@@
 & & &  cpl (sd) & 35.54 (4.2) & 22.58 (2.2) & 23.26 (1.9) \\
- &  &HU &  MSE (sd) & 173.65 (25.1) & 143.04 (23.2) & 139.51 (17.6) \\
+ &  & $BIC_+$ &  MSE (sd) & 173.65 (25.1) & 143.04 (23.2) & 139.51 (17.6) \\
 & & & cpl (sd) & 35.54 (4.2) & 23.16 (1.9) & 23.26 (1.9) \\
@@
 & & &  cpl (sd) & 30.78 (6.4) & 9.25 (1.2) & 8.67 (1) \\
- &  &HU &  MSE (sd) & 161.92 (25.2) & 127.44 (13.3) & 111.46 (7.2) \\
+ &  & $BIC_+$ &  MSE (sd) & 161.92 (25.2) & 127.44 (13.3) & 111.46 (7.2) \\
 & & &cpl (sd) & 30.78 (6.4) & 17.85 (2.7) & 8.67 (1) \\
@@
 & & &cpl (sd) & 22.8 (3.3) & 19.34 (4.9) & 22.8 (3.3) \\
- &  &HU &MSE (sd) & 1 919.43 (861.8) & 1 885.39 (814) & 1 919.43 (861.8) \\
+ &  & $BIC_+$ &MSE (sd) & 1 919.43 (861.8) & 1 885.39 (814) & 1 919.43 (861.8) \\
 & & & cpl (sd) & 22.8 (3.3) & 22.73 (3.1) & 22.8 (3.3) \\
@@
 & & & cpl (sd) & 15.9 (3.8) & 15.1 (3.7) & 15.76 (3.7) \\
- &  &HU &MSE (sd) & 696.46 (492.9) & 655.94 (325.7) & 660.93 (354) \\
+ &  & $BIC_+$ &MSE (sd) & 696.46 (492.9) & 655.94 (325.7) & 660.93 (354) \\
 & & & cpl (sd) & 15.9 (3.8) & 15.72 (3.6) & 15.76 (3.7) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 394.93 (356) & 155.18 (36.4) & 147.81 (30.3) \\
+30 & 32 & $BIC$ & MSE (sd) & 394.93 (356) & 155.18 (36.4) & 147.81 (30.3) \\
 & & &cpl (sd) & 16.1 (6.3) & 8.26 (1.7) & 7.77 (1.3) \\
@@
 & & & cpl (sd) & 21.24 (5.2) & 18.13 (2.8) & 19.23 (2.7) \\
- &  &HU &MSE (sd) & 279.82 (151.1) & 252.35 (108.5) & 239.61 (69.5) \\
+ &  & $BIC_+$ &MSE (sd) & 279.82 (151.1) & 252.35 (108.5) & 239.61 (69.5) \\
 & & & cpl (sd) & 21.24 (5.2) & 18.97 (2.7) & 19.23 (2.7) \\
@@
 & & & cpl (sd) & 14.31 (6.2) & 8.16 (1.2) & 8.04 (1) \\
- &  &HU &MSE (sd) & 199.3 (158.7) & 136.32 (25.2) & 124.19 (13) \\
+ &  & $BIC_+$ &MSE (sd) & 199.3 (158.7) & 136.32 (25.2) & 124.19 (13) \\
 & & &cpl (sd) & 14.31 (6.2) & 9.59 (2.6) & 8.04 (1) \\
@@
 & & & cpl (sd) & 36.3 (2.5) & 30.94 (4.2) & 36.3 (2.5) \\
- &  &HU &   MSE (sd) & 193.26 (34.9) & 194.43 (37.3) & 193.26 (34.9) \\
+ &  & $BIC_+$ &   MSE (sd) & 193.26 (34.9) & 194.43 (37.3) & 193.26 (34.9) \\
 & & &  cpl (sd) & 36.3 (2.5) & 36.22 (2.7) & 36.3 (2.5) \\
@@
 & & & cpl (sd) & 23.44 (3.8) & 21.23 (2.1) & 21.93 (1.7) \\
- &  &HU &  MSE (sd) & 145.46 (20.1) & 143.74 (23.2) & 140.27 (17.5) \\
+ &  & $BIC_+$ &  MSE (sd) & 145.46 (20.1) & 143.74 (23.2) & 140.27 (17.5) \\
 & & & cpl (sd) & 23.44 (3.8) & 21.86 (1.7) & 21.93 (1.7) \\
@@
 & & &  cpl (sd) & 13.58 (5.2) & 8.59 (1.1) & 8.48 (1) \\
- &  &HU &  MSE (sd) & 128.44 (18.4) & 118.61 (11.5) & 111.77 (7.2) \\
+ &  & $BIC_+$ &  MSE (sd) & 128.44 (18.4) & 118.61 (11.5) & 111.77 (7.2) \\
 & & & cpl (sd) & 13.58 (5.2) & 10.75 (2.8) & 8.48 (1) \\
@@
 & & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
- &  &HU &MSE (sd) & 81781.54 (247281.7) & 587.07 (425.6) & 593.03 (428.3) \\
+ &  & $BIC_+$ &MSE (sd) & 81781.54 (247281.7) & 587.07 (425.6) & 593.03 (428.3) \\
 & & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25 (0) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 74136.81 (297716.1) & 189.45 (48.1) & 144.51 (31.4) \\
+30 & 32 & $BIC$ & MSE (sd) & 74136.81 (297716.1) & 189.45 (48.1) & 144.51 (31.4) \\
 & & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
- &  &HU &MSE (sd) & 684.04 (438.1) & 196.36 (40.5) & 197.32 (40) \\
+ &  & $BIC_+$ &MSE (sd) & 684.04 (438.1) & 196.36 (40.5) & 197.32 (40) \\
 & & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
- &  &HU &MSE (sd) & 596.32 (323) & 168.03 (27.7) & 119.39 (12.4) \\
+ &  & $BIC_+$ &MSE (sd) & 596.32 (323) & 168.03 (27.7) & 119.39 (12.4) \\
 & & &cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
@@
 & & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
- &  &HU &  MSE (sd) & 168.35 (20.1) & 135.04 (13.1) & 135.08 (13.1) \\
+ &  & $BIC_+$ &  MSE (sd) & 168.35 (20.1) & 135.04 (13.1) & 135.08 (13.1) \\
 & & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25 (0) \\
@@
 & & &  cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
- &  &HU & MSE (sd) & 168.07 (21.4) & 124.23 (11.1) & 108.64 (7.2) \\
+ &  & $BIC_+$ & MSE (sd) & 168.07 (21.4) & 124.23 (11.1) & 108.64 (7.2) \\
 & & & cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
@@
 & & & cpl (sd) & 14.37 (5.1) & 12.9 (3.8) & 12.84 (3.8) \\
- &  &HU &MSE (sd) & 367.45 (198.4) & 298.1 (119.1) & 292.53 (102.2) \\
+ &  & $BIC_+$ &MSE (sd) & 367.45 (198.4) & 298.1 (119.1) & 292.53 (102.2) \\
 & & & cpl (sd) & 14.37 (5.1) & 12.81 (3.8) & 12.84 (3.8) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 273.12 (267.2) & 166.93 (49) & 152.45 (39.5) \\
+30 & 32 & $BIC$ & MSE (sd) & 273.12 (267.2) & 166.93 (49) & 152.45 (39.5) \\
 & & & cpl (sd) & 11.7 (5.8) & 8.7 (2.5) & 7.45 (1.4) \\
@@
 & & & cpl (sd) & 14.15 (3.9) & 13.4 (2.8) & 13.5 (3.1) \\
- &  &HU &MSE (sd) & 189.64 (52.5) & 174.36 (45.3) & 176.58 (46.8) \\
+ &  & $BIC_+$ &MSE (sd) & 189.64 (52.5) & 174.36 (45.3) & 176.58 (46.8) \\
 & & & cpl (sd) & 14.15 (3.9) & 13.57 (3) & 13.5 (3.1) \\
@@
 & & & cpl (sd) & 13.06 (7.2) & 8.29 (1.6) & 7.73 (1.2) \\
- &  &HU &MSE (sd) & 163.88 (67.4) & 133.7 (26.4) & 121.99 (16.6) \\
+ &  & $BIC_+$ &MSE (sd) & 163.88 (67.4) & 133.7 (26.4) & 121.99 (16.6) \\
 & & &cpl (sd) & 13.06 (7.2) & 9.65 (3.3) & 7.73 (1.2) \\
@@
 & & &  cpl (sd) & 14.86 (3.1) & 13.97 (2.5) & 14.04 (2.5) \\
- &  &HU &  MSE (sd) & 129.65 (14.8) & 127.46 (12.2) & 127.46 (12.2) \\
+ &  & $BIC_+$ &  MSE (sd) & 129.65 (14.8) & 127.46 (12.2) & 127.46 (12.2) \\
 & & & cpl (sd) & 14.86 (3.1) & 14.02 (2.5) & 14.04 (2.5) \\
@@
 & & & cpl (sd) & 12.84 (5.2) & 8.42 (1.1) & 8.11 (0.8) \\
- &  &HU &  MSE (sd) & 121.21 (20.1) & 112.56 (11.2) & 109.04 (7.8) \\
+ &  & $BIC_+$ &  MSE (sd) & 121.21 (20.1) & 112.56 (11.2) & 109.04 (7.8) \\
 & & & cpl (sd) & 12.84 (5.2) & 9.77 (2.3) & 8.11 (0.8) \\
@@
 & & & cpl (sd) & 14.08 (6.1) & 11.75 (4.8) & 11.49 (4.7) \\
- &  &HU &MSE (sd) & 499.32 (218.8) & 307.6 (128.7) & 305.09 (123.5) \\
+ &  & $BIC_+$ &MSE (sd) & 499.32 (218.8) & 307.6 (128.7) & 305.09 (123.5) \\
 & & & cpl (sd) & 14.08 (6.1) & 11.63 (4.7) & 11.49 (4.7) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 691.55 (503.2) & 193.45 (54.2) & 150.17 (35) \\
+30 & 32 & $BIC$ & MSE (sd) & 691.55 (503.2) & 193.45 (54.2) & 150.17 (35) \\
 & & & cpl (sd) & 12.54 (7.2) & 10.26 (2.7) & 7.81 (1.4) \\
@@
 & & & cpl (sd) & 20.37 (6.3) & 13.68 (4) & 14.05 (4) \\
- &  &HU &MSE (sd) & 282.05 (134.8) & 180.84 (47.5) & 181.49 (48.6) \\
+ &  & $BIC_+$ &MSE (sd) & 282.05 (134.8) & 180.84 (47.5) & 181.49 (48.6) \\
 & & & cpl (sd) & 20.37 (6.3) & 14.03 (4) & 14.05 (4) \\
@@
 & & & cpl (sd) & 22.94 (7.7) & 9.37 (2.1) & 7.91 (1.4) \\
- &  &HU &MSE (sd) & 308.13 (142.9) & 159.14 (30.4) & 122.71 (17.8) \\
+ &  & $BIC_+$ &MSE (sd) & 308.13 (142.9) & 159.14 (30.4) & 122.71 (17.8) \\
 & & &cpl (sd) & 22.94 (7.7) & 15.42 (4) & 7.91 (1.4) \\
@@
 & & &  cpl (sd) & 25.6 (5.4) & 14.73 (3.8) & 14.73 (3.9) \\
- &  &HU &  MSE (sd) & 150.14 (19.7) & 128.21 (14.9) & 128.21 (15) \\
+ &  & $BIC_+$ &  MSE (sd) & 150.14 (19.7) & 128.21 (14.9) & 128.21 (15) \\
 & & & cpl (sd) & 25.6 (5.4) & 14.7 (3.9) & 14.73 (3.9) \\
@@
 & & &  cpl (sd) & 30.41 (6.7) & 8.96 (1.3) & 8.3 (1) \\
- &  &HU &  MSE (sd) & 157.18 (23.5) & 123.8 (12.5) & 109.24 (7.8) \\
+ &  & $BIC_+$ &  MSE (sd) & 157.18 (23.5) & 123.8 (12.5) & 109.24 (7.8) \\
 & & & cpl (sd) & 30.41 (6.7) & 17.18 (3.3) & 8.3 (1) \\
@@
 & & & cpl (sd) & 13.09 (3.2) & 12.9 (2.9) & 12.92 (3) \\
- &  &HU &MSE (sd) & 394.59 (454.6) & 353.99 (157) & 351.14 (148) \\
+ &  & $BIC_+$ &MSE (sd) & 394.59 (454.6) & 353.99 (157) & 351.14 (148) \\
 & & & cpl (sd) & 13.09 (3.2) & 12.83 (2.9) & 12.92 (3) \\
 \hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & U & MSE (sd) & 410.11 (343.5) & 163.05 (45.4) & 152.08 (38.1) \\
+30 & 32 & $BIC$ & MSE (sd) & 410.11 (343.5) & 163.05 (45.4) & 152.08 (38.1) \\
 & & & cpl (sd) & 15.8 (6.5) & 8.29 (2.2) & 7.34 (1.3) \\
@@
 & & & cpl (sd) & 13.56 (3.7) & 12.78 (2.4) & 13.07 (2.4) \\
- &  &HU &MSE (sd) & 180.67 (52.7) & 175.3 (38.1) & 174.84 (36.8) \\
+ &  & $BIC_+$ &MSE (sd) & 180.67 (52.7) & 175.3 (38.1) & 174.84 (36.8) \\
 & & & cpl (sd) & 13.56 (3.7) & 13.01 (2.4) & 13.07 (2.4) \\
@@
 & & & cpl (sd) & 14.74 (6.6) & 8.07 (1.6) & 7.65 (1.2) \\
- &  &HU &MSE (sd) & 188.91 (88.5) & 134.22 (23.6) & 122.38 (16.7) \\
+ &  & $BIC_+$ &MSE (sd) & 188.91 (88.5) & 134.22 (23.6) & 122.38 (16.7) \\
 & & &cpl (sd) & 14.74 (6.6) & 9.5 (2.8) & 7.65 (1.2) \\
@@
 & & &  cpl (sd) & 13.78 (2.8) & 13.51 (2) & 13.59 (2.1) \\
- &  &HU & MSE (sd) & 128.18 (13.7) & 127.71 (13.3) & 127.71 (13.3) \\
+ &  & $BIC_+$ & MSE (sd) & 128.18 (13.7) & 127.71 (13.3) & 127.71 (13.3) \\
 & & & cpl (sd) & 13.78 (2.8) & 13.59 (2.1) & 13.59 (2.1) \\
@@
 & & & cpl (sd) & 13.25 (4.4) & 8.3 (1) & 8.12 (0.8) \\
- &  &HU &  MSE (sd) & 126.15 (20.6) & 113.96 (11.4) & 109.04 (7.8) \\
+ &  & $BIC_+$ &  MSE (sd) & 126.15 (20.6) & 113.96 (11.4) & 109.04 (7.8) \\
 & & & cpl (sd) & 13.25 (4.4) & 10.08 (2.6) & 8.12 (0.8) \\
@@
 & & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\
- &  &HU &MSE (sd) & 996.06 (323.7) & 1094.5 (621.8) & 1501.25 (905.9) \\
+ &  &$BIC_+$ &MSE (sd) & 996.06 (323.7) & 1094.5 (621.8) & 1501.25 (905.9) \\
 & & & cpl (sd) & 17.66 (5.4) & 16.98 (4.8) & 11.97 (4.9) \\ 
@@
 				\label{barplotMSEexfos}
-				\includegraphics[width=400px]{figures/MSEfinal.png}
+				\includegraphics[width=430px]{figures/MSEfinal.png}
 			\caption{MSE comparison on industrial dataset. Learning set : 3 000 products, validation set : 847 products}

### 2014-05-06 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@571 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 {\Large
-	{\sc Model-based variable selection for regression \\ with highly correlated variables.}
+	{\sc Model-based variable selection in regression \\ with highly correlated variables.}
 }
@@
 We propose to explicitly model the correlations by a family of linear regressions between the covariates, some covariates explaining others.
-It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable.
+It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model.
 It corresponds to a kind of variable selection preliminary step which has then to be followed by standard linear estimation methods including classical variables selection procedures for instance.
@@
 	increasing the chance to have correlated ones.  
-	 In such a context, variance of the estimators can lead to arbitrary results whereas interpretation and prediction are both strongly needed, depending on the context.
+	 In such a context, variance of the estimators can lead to arbitrary results leading to misleading interpretations whereas interpretation and prediction are both strongly needed, depending on the context.
 For instance, interpretation could be favoured to improve a situation instead of only predict problems, so robust estimation is a real stake.
@@
 	The Ordinary Least Squares (\textsc{OLS}) estimator is the minimum-variance linear unbiased estimator but it suffers from ill-conditioned matrices inversion when the covariates are highly correlated (section \ref{sectionOLS}).
-Therefore, some other methods try to improve the robustess of the estimator by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}.
+Some other methods try to improve the robustess of the estimator by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}
+	but many of these methods are not directly guided by the correlations whereas knowing the correlations between the covariates facilitates interpretation.
 
@@
 	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance tradeoff and get better prediction.
-	Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty.
-	But this penalty is not guided by the correlations.
-It is the same for each covariate and will be too large for independent covariates and/or too small for correlated ones.
-So the efficiency of such a method is limited.
+	Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty but
+	it is the same for each covariate no matter how much it is correlated or not to others.%and will be too large for independent covariates and/or too small for correlated ones.  
 	Moreover, coefficients tend to 0 but don't reach 0 so it gives difficult interpretations for large number of covariates.
@@
 	 Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
-	We see that none of the above methods takes explicitly the correlations into account, even if the clustering methods introduce a kind of relationship between the covariates.
+	We see that none of the above methods takes explicitly the correlations into account, even if the clustering methods may group the correlated covariates together.
 
@@
 % Principe de la méthode
-
- Instead of reducing the dimension without looking at correlations, Selvarclust \cite{maugis2009variable} proposes to focus on the correlations, giving an explicit structure between covariates by linear sub-regressions between them.
-But it is made in a clustering context, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even it is known to be often unstable \cite{miller2002subset}.
-We propose to develop this method for linear regression with a specifically adapted algorithm to find the structure of sub-regression.
+In the field of clustering, model-based explicit correlations by linear sub-regressions between covariates claims to obtain good results in both interpretation and estimation quality (like Selvarclust \cite{maugis2009variable}).
+But this is made in a clustering context only, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even it is known to be often unstable \cite{miller2002subset}.
+We propose to transpose this method for linear regression with a specifically adapted algorithm to find the structure of sub-regression.
 
@@
 \end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_r}$.
+\caption{OLS and OLS combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}_r$.
 Sometimes $\hat{S}$ gives better results than $S$ because $S$ is penalized by the fact that it relies only on covariates not in the true model.
@@
 \end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_r}$. }\label{YX2linLASSO}
+\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}_r$. }\label{YX2linLASSO}
 \end{table}
@@
 \end{tabular} 
-\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X_2}$.}\label{YX2linenet}
+\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}_r$.}\label{YX2linenet}
 \end{table}
@@
 \end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X_r}$.}\label{YX2linstep}
+\caption{Stepwise  combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}_r$.}\label{YX2linstep}
 \end{table}

### 2014-05-06 www/article13.tex
up intro
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@572 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	increasing the chance to have correlated ones.  
-	 In such a context, variance of the estimators can lead to arbitrary results leading to misleading interpretations whereas interpretation and prediction are both strongly needed, depending on the context.
+		Many estimators rely on the inversion of matrix that will be ill-conditioned in such a context, increasing the variance of these estimators and leading to misleading interpretations and reduced prediction efficiency.
+		%whereas interpretation and prediction are both strongly needed, depending on the context.
 For instance, interpretation could be favoured to improve a situation instead of only predict problems, so robust estimation is a real stake.
@@
 %bibliographie	
-	The Ordinary Least Squares (\textsc{OLS}) estimator is the minimum-variance linear unbiased estimator but it suffers from ill-conditioned matrices inversion when the covariates are highly correlated (section \ref{sectionOLS}).
-Some other methods try to improve the robustess of the estimator by introducing some bias, selecting only some covariates, grouping some covariates, {\it etc.}
-	but many of these methods are not directly guided by the correlations whereas knowing the correlations between the covariates facilitates interpretation.
+	We know the minimum-variance linear unbiased estimator for linear regression that is the Ordinary Least Squares (\textsc{OLS}) estimator but it suffers from ill-conditioned matrices inversion when the covariates are highly correlated (section \ref{sectionOLS}).
+	Many traditional try to reduce the variance introducing some bias to improve the bias-variance tradeoff and get better prediction by selection of only some covariates, grouping of some covariates, {\it etc.}
 
@@
 
-	 Penalized methods try to reduce the variance introducing some bias to improve the bias-variance tradeoff and get better prediction.
 	 Ridge regression \cite{marquardt1975ridge} proposes a biased estimator that can be written in terms of a parametric $L_2$ penalty but
@@
 
-	Real datasets often imply many irrelevant variables so variable selection would be favoured when possible to reduce the dimension and obtain models small enough to be understood.
-Moreover, selection may keep only uncorrelated covariates and thus resolve the correlations problem.
+	Real datasets often imply many irrelevant variables so variable selection should be favoured when possible to reduce the dimension and obtain models small enough to be understood.
+Moreover, selection may keep only uncorrelated covariates and thus resolve the correlations problem.	Variable selection methods may add some bias by deleting some relevant covariates but reduce the variance by the dimension reduction.
 	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a parametric $L_1$ penalty to shrink some coefficients exactly to zero.
@@
 	Another way of improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}) to reduce the dimension and also the correlations if correlated covariates are in the same clusters.
+The bias is added by the dimension reduction inherent to the coefficients clustering.
 	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the regression coefficients no longer as fixed effect parameters but as unobserved independent random variables with grouped coefficients following a Gaussian Mixture distribution.
-The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than the number of covariates.
+The idea is that if the model have a small number of groups of covariates and the mixture have few enough components the model will have a number of parameters to estimate significantly lower than the number of covariates.
 In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on the coefficients estimator.
@@
 	 Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
+It is again a probable increase of the bias by variable selection to reduce the variance by dimension reduction.
+	  %\textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) takes into account correlations between endogenous covariates but without explicit expression of the correlations and thus a reduced interpretation potential.
 	We see that none of the above methods takes explicitly the correlations into account, even if the clustering methods may group the correlated covariates together.
+	These methods are not directly guided by the correlations whereas knowing the correlations between the covariates facilitates interpretation.
 
@@
 % Principe de la méthode
-In the field of clustering, model-based explicit correlations by linear sub-regressions between covariates claims to obtain good results in both interpretation and estimation quality (like Selvarclust \cite{maugis2009variable}).
-But this is made in a clustering context only, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even it is known to be often unstable \cite{miller2002subset}.
+Outside of the linear regression field, model-based explicit correlations by linear sub-regressions between covariates  claims  to obtain good results in both interpretation and estimation quality  \cite{maugis2009variable}.
+But this is made in a clustering context only, with only irrelevant covariates being dependent from relevant ones (no sub-regressions between relevant or between irrelevant covariates) and the algorithm used to find the structure is a stepwise-like algorithm without protection against correlations \cite{raftery2006variable} even if it is known to be often unstable \cite{miller2002subset}.
 We propose to transpose this method for linear regression with a specifically adapted algorithm to find the structure of sub-regression.
@@
   The idea is that if we know explicitly the correlations, we could use this knowledge to avoid the problem. 
+	We use the correlations as new information to reduce the variance without adding any bias.  
   More precisely, we model the correlations with a system of linear sub-regressions for the joint distribution of the covariates to model the structure of the correlations.
@@
 Thus we have an explained variable selection guided by the correlations and it improves interpretation, not only prediction.
+This marginal model still is the true model but in a different probability space, without bias and with reduced variance for the estimators.
+The model depends only on independent covariates because it is a marginal model on these covariates.
  This can be viewed as a pretreatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
@@
 	\end{equation}
-	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I}_n)$ the noise of the regression, with $\boldsymbol{I}_n$ the $n$-sized identity matrix and $\sigma_Y >0$.
+	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables (that is a sub-matrix of $\tilde{\boldsymbol{X}}$ the $n\times \tilde{p}$ matrix of provided covariates), $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I}_n)$ the noise of the regression, with $\boldsymbol{I}_n$ the $n$-sized identity matrix and $\sigma_Y >0$.
 The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{OLS}): %As shown in section \ref{sectionOLS},
@@
 	Another problem is that matrix inversion requires $n\geq p$. 	
-As a running example, we look at a simple case with $p=5$ variables defined by four independent scaled Gaussian $\mathcal{N}(0,1)$ named $\boldsymbol{x}_1,\boldsymbol{x}_2$ and $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$.
+\paragraph{Running example:} we look at a simple case with $p=5$ variables defined by four independent scaled Gaussian $\mathcal{N}(0,1)$ named $\boldsymbol{x}_1,\boldsymbol{x}_2$ and $\boldsymbol{x}_3=\boldsymbol{x}_1+\boldsymbol{x}_2+\boldsymbol{\varepsilon}_3$ where $\boldsymbol{\varepsilon}_3\sim{\mathcal{N}(\boldsymbol{0},\sigma_3^2\boldsymbol{I}_n)}$.
 We also define another couple $\boldsymbol{x}_4,\boldsymbol{x}_5$ of covariates that are {\it i.i.d.
@@
  Then $\boldsymbol{X}_r$ is the $n\times p_r$ submatrix of $0\leq p_r <p$ redundent covariates and $\boldsymbol{X}_f$ the $n\times (p-p_r)$ submatrix of the free (independent) covariates.
-
-In our running example, $\boldsymbol{X}_r=\boldsymbol{x}_3$ and $\boldsymbol{X}_f=\{\boldsymbol{x}_1,\boldsymbol{x}_2,\boldsymbol{x}_4,\boldsymbol{x}_5 \}$
 
@@
 		where $\boldsymbol{\alpha}_j \in \mathcal{R}^{(p-p_r)}$ are the sparse vectors of the regression coefficients between the covariates (each sub-regression freely implies different covariates). 
-In our running example, $p_r=1$ and $\boldsymbol{\alpha}_3=(1,1,0,0)'$
 \\
@@
 
-%ancienne intro
+The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation} \cite{TIMM}.
+  	Here we suppose the $\boldsymbol{\varepsilon}_j$ independent but in other cases \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) takes into account correlations between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) and could be used to estimate the $\boldsymbol{\alpha}_j$.
+
+\paragraph{In the running example:}$\boldsymbol{X}_r=\boldsymbol{x}_3$, $\boldsymbol{X}_f=\{\boldsymbol{x}_1,\boldsymbol{x}_2,\boldsymbol{x}_4,\boldsymbol{x}_5 \}$, $p_r=1$ and $\boldsymbol{\alpha}_3=(1,1,0,0)'$
 
@@
 Now we know $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$ by the structure of sub-regressions, we are able to define a marginal regression model $P(\boldsymbol{Y}|\boldsymbol{X}_f,S)$ based on the reduced set of independent covariates $\boldsymbol{X}_f$ without significant information loss.
+We use the information of the correlations structure to rewrite the true model without bias in the marginal space defined by the independent covariates.
  	\\
@@
 	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
-	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_r^*) \textrm{ and } \boldsymbol{\beta}_r^*=\boldsymbol{0}\label{modexpl}
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*\label{modexpl}
 \end{eqnarray}
+ This model is still the true model and OLS estimator will still give an unbiased estimator, but its variance will be reduced by both dimension reduction and decorrelation (variables in $\boldsymbol{X}_f$ are independent so the matrix $\boldsymbol{X}_f'\boldsymbol{X}_f$ will be well-conditioned).
+So the information given by the structure $S$ allows to reduce the variance without adding bias, by simple marginalization.
+\\
+Nevertheless, to be able to compare the bias-variance tradeoff, we can see this model as a variable pre-selection independent of the response $\boldsymbol{Y}$.
 We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by selecting $\boldsymbol{X}_f$ because of the correlations given by $S$.
-It is a variable pre-selection independent of the response $\boldsymbol{Y}$.
-\\
-Our running example gives here $\boldsymbol{Y}= 2\boldsymbol{x}_1+2\boldsymbol{x}_2+\boldsymbol{x}_4+\boldsymbol{x}_5+\boldsymbol{\varepsilon}_3 +\boldsymbol{\varepsilon}_Y$
+So we also get the model
+\begin{equation}
+\boldsymbol{Y}_{|\boldsymbol{X},S}=\boldsymbol{X}\boldsymbol{\beta}^*+\boldsymbol{\varepsilon}_Y^* \textrm{ where }\boldsymbol{\beta}^*=(\boldsymbol{\beta}_f^*,\boldsymbol{\beta}_r^*) \textrm{ and } \boldsymbol{\beta}_r^*=\boldsymbol{0}
+\end{equation}
+	for which OLS estimator of the coefficients may be biased.
+
+\paragraph{Running example:} $\boldsymbol{Y}_{|\boldsymbol{X}_f}= 2\boldsymbol{x}_1+2\boldsymbol{x}_2+\boldsymbol{x}_4+\boldsymbol{x}_5+\boldsymbol{\varepsilon}_3 +\boldsymbol{\varepsilon}_Y$
 \subsection{Strategy of use: pretreatment before classical estimation/selection methods}\label{interpretation}
 
-As a pretreatment, it allows usage of any method in a second time to estimate $\boldsymbol{\beta}^*$, even with variable selection methods like LASSO or stepwise \cite{seber2012linear}.
-
-After selection and estimation we will obtain a model with { \it two steps of variable selection}: the decorrelation step (coerced selection associated to redundant information defined in $S$) and the classical selection step, with different meanings for obtained zeros in $\hat{\boldsymbol{\beta}}^*_f$ (irrelevant covariates) and for $\hat{\boldsymbol{\beta}}^*_r=0$ (redundant information).
+As a pretreatment, the model allows usage of any method in a second time to estimate $\boldsymbol{\beta}_f^*$, even with variable selection methods like LASSO or a best subset algorithm like stepwise \cite{seber2012linear}.
+However, we always have $\boldsymbol{X}_r=\boldsymbol{0}$
+
+After selection and estimation we will obtain a model with { \it two steps of variable selection}: the decorrelation step by marginalization(coerced selection associated to redundant information defined in $S$) and the classical selection step, with different meanings for obtained zeros in $\hat{\boldsymbol{\beta}}^*_f$ (irrelevant covariates) and for $\hat{\boldsymbol{\beta}}^*_r=0$ (redundant information).
  Thus we are able to distinguish the reasons of selection and consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.
@@
 
-The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}.
-Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when it is basic (recursive \textsc{SEM}).
-  	Some methods take into account correlations but they only consider covariances between residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or  endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}) without explicit expression of the correlations and thus a smaller interpretation potential.
-  	\\ Here we suppose independence between the residuals but in other cases it remains the possibility to use such methods to estimate the $\boldsymbol{\alpha}_j$ and $\sigma_j$.
-
 Moreover, the uncrossing constraint (partition of $\boldsymbol{X}$) guarantee to keep a simple structure easily interpretable (no cycles and no chain-effect) and straightforward readable.
@@
 It's just a compromise between numerical issues caused by correlations for estimation and selection versus increased variability due to structural hypothesis.
-We play on the bias-variance tradeoff.
+We just play on the traditional bias-variance tradeoff.
 

### 2014-05-06 www/article13.tex
up texte
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@573 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Thus we have an explained variable selection guided by the correlations and it improves interpretation, not only prediction.
-This marginal model still is the true model but in a different probability space, without bias and with reduced variance for the estimators.
+This marginal model still is the true model but in a different probability space, and OLS still gives an estimator without bias but with a reduced variance for the estimators.
 The model depends only on independent covariates because it is a marginal model on these covariates.
@@
 \subsection{A by-product model: marginal regression with decorrelated covariates}
-Now we know $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$ by the structure of sub-regressions, we are able to define a marginal regression model $P(\boldsymbol{Y}|\boldsymbol{X}_f,S)$ based on the reduced set of independent covariates $\boldsymbol{X}_f$ without significant information loss.
+Now we know $P(\boldsymbol{X}_r|\boldsymbol{X}_f,S)$ by the structure of sub-regressions, we are able to define a marginal regression model $P(\boldsymbol{Y}|\boldsymbol{X}_f,S)$ based on the reduced set of independent covariates $\hat{\boldsymbol{\beta}}_f$ without significant information loss.
 We use the information of the correlations structure to rewrite the true model without bias in the marginal space defined by the independent covariates.
@@
 \\
-Nevertheless, to be able to compare the bias-variance tradeoff, we can see this model as a variable pre-selection independent of the response $\boldsymbol{Y}$.
+Nevertheless, to be able to compare the bias-variance tradeoff, we can see this model as a variable pre-selection independent of the response in $\boldsymbol{Y}_{|\boldsymbol{X}}$.
 We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by selecting $\boldsymbol{X}_f$ because of the correlations given by $S$.

### 2014-05-15 pkg/CorReg/R/searchZ.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@602 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -38,7 +38,10 @@ searchZ<-function(X=X,Z=NULL,Bic_null_vect=NULL,candidates=-1,reject=0,methode=1
   if(is.null(Bic_null_vect)){
      Bic_null_vect=density_estimation(X=X,nbclustmax=10,verbose=FALSE,detailed=FALSE,mclust=TRUE)$BIC_vect
   }
-  if(is.null(nbini) | nbini<1 ){
+  if(!is.null(nbini)){
+     if (nbini<1){nbini=NULL}
+  }
+  if(is.null(nbini)){
      if(is.null(Z)){
         Z=matrix(0,ncol=ncol(X),nrow=ncol(X))
      }
@@ -74,7 +77,7 @@ searchZ<-function(X=X,Z=NULL,Bic_null_vect=NULL,candidates=-1,reject=0,methode=1
      }
     
      for(i in 1:nbini){
-        if(Wini){
+        if(Wini){#only if no Z provided
            Z=Winitial(W=W,X=X,p1max=p1max,Bic_null_vect=Bic_null_vect,p2max=p2max)
         }
         if(reject==0){#relax mode

### 2014-05-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@603 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
-	In the followings we suppose $\sigma^2_j\neq 0$, then $\boldsymbol{X}_f'\boldsymbol{X}_f$ and $\boldsymbol{X}'\boldsymbol{X}$ are of full rank (but the later is ill-conditioned for small values of $\sigma^2_j$)  .
-
+	In the followings we suppose $\sigma^2_j\neq 0$, then $\boldsymbol{X}_f'\boldsymbol{X}_f$ and $\boldsymbol{X}'\boldsymbol{X}$ are of full rank (but the later is ill-conditioned for small values of $\sigma^2_j$).
+	\\
+
+Our full generative model is a $p$-sized Gaussian mixture model of $K$ distinct components and 
+%	\begin{equation}
+%	f(\boldsymbol{X}|K,S)=f(\boldsymbol{X}_f|K,S)f(\boldsymbol{X}_r|\boldsymbol{X}_f,S)
+%	\end{equation}	
+%	
+	can be seen as a $\mathbf{SR}$ model defined by Maugis \cite{maugis2009variable}.
+In this section, $S$ will denote the set of variable as in the paper from Maugis and we call Gaussian mixtures the Gaussian mixtures with at least two distinct components.
+The equivalence with Maugis's model is defined by:
+	$\boldsymbol{X}_r=\boldsymbol{y}^{S^c}$ and $\boldsymbol{X}_f=\boldsymbol{y}^R$.
+We have supposed independence between variables in $\boldsymbol{X}_f$ so the identifiability theorem from Maugis tells that our model is identifiable if variables in $\boldsymbol{X}_f$ are Gaussian mixtures (what we supposed in section \ref{sectionfullgen}).
+	\\
+
+%First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there must be at least one Gaussian mixture on the right of each sub-regression. 
+We define $\boldsymbol{X}^G \subsetneq \boldsymbol{X}_f$ containing Gaussian variables and we note the Gaussian mixtures $\boldsymbol{X}^{G^c}\neq \emptyset$ its complement in $\boldsymbol{X}_f$.
+We suppose that variables in $\boldsymbol{X}_r$ are all Gaussian mixtures.
+It implies that $\forall j  \in I_r,\exists i \in I_f^j $ so that $\boldsymbol{X}^i \subset \boldsymbol{X}^{G^c} $ since any linear combination of Gaussian variable would only give a Gaussian (so each sub-regression contain at least one Gaussian mixture as a regressor).
+\\
 	We introduce the matricial notation
-		$\boldsymbol{X}_r=\boldsymbol{X}_f\boldsymbol{\alpha} + \boldsymbol{\varepsilon}$
-		we note $\Theta$ the parameter of the model. $\boldsymbol{\alpha}$ is the $(p-p_r)\times p_r$ matrix whose columns are the $\boldsymbol{\alpha}_j$.
-		\\ We want to know if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|\tilde{S},\tilde{\Theta})$ does imply $(S,\Theta)=(\tilde{S},\tilde{\Theta})$.
+		$\boldsymbol{X}_r=\boldsymbol{X}_f\boldsymbol{\alpha} + \boldsymbol{\varepsilon}$ where
+		 $\boldsymbol{\alpha}$ is the $(p-p_r)\times p_r$ matrix whose columns are the $\boldsymbol{\alpha}_j$ and $\boldsymbol{\varepsilon}$ is the $n\times p_r$ matrix whose columns are the $\boldsymbol{\varepsilon}_j$
+		%\\We note $\Theta$ the parameter of the model.We want to know if $P(\boldsymbol{X}|S,\Theta)=P(\boldsymbol{X}|\tilde{S},\tilde{\Theta})$ does imply $(S,\Theta)=(\tilde{S},\tilde{\Theta})$.
+
+The theorem from Maugis guarantee that a sub-regression between Gaussian mixtures is identifiable in terms of which one is regressed by others.
+		\begin{eqnarray}
+%			\forall j \in I_r, \boldsymbol{X}^j_{|\boldsymbol{X}^G,\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^G\boldsymbol{\alpha}_j^G+\boldsymbol{X}^{G^c}\boldsymbol{\alpha}_j^{G^c}+ \boldsymbol{\varepsilon}_j \\
+%			\boldsymbol{X}^j_{|\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^{G^c}\boldsymbol{\alpha}_j^{G^c} + \tilde{\boldsymbol{\varepsilon}}_j \textrm{ is identifiable where} \\
+%			\tilde{\boldsymbol{\varepsilon}}_j&=&\boldsymbol{X}^{G}\boldsymbol{\alpha}_j^{G}+ \boldsymbol{\varepsilon}_j \textrm{ is Gaussian. 
+		 \boldsymbol{X}_{r|\boldsymbol{X}^G,\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^G\boldsymbol{\alpha}_G+\boldsymbol{X}^{G^c}\boldsymbol{\alpha}_{G^c}+ \boldsymbol{\varepsilon} \\
+			\boldsymbol{X}_{r|\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^{G^c}\boldsymbol{\alpha}_{G^c} + \tilde{\boldsymbol{\varepsilon}} \textrm{ is identifiable where} \\
+			\tilde{\boldsymbol{\varepsilon}}_j&=&\boldsymbol{X}^{G}\boldsymbol{\alpha}_j^{G}+ \boldsymbol{\varepsilon}_j \textrm{ is Gaussian.}  
+		\end{eqnarray}
+%		We have the unique (Gaussian are allowed only on the right of the sub-regressions) decomposition $f(\boldsymbol{X}_r|\boldsymbol{X}_f)=f(\boldsymbol{X}_r|\boldsymbol{X}^G,\boldsymbol{X}^{G^c})=f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G}) $
+%		where  $f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})$ is identifiable. Moreover, $\boldsymbol{X}^{G}$ is a set (empty or not) of independent Gaussian variables, whose parameters are identifiable.
+%		So the structure is identifiable.
+%		\begin{eqnarray}
+%			f(\boldsymbol{X})&=&f(\boldsymbol{X}\setminus\boldsymbol{X}^{G},\boldsymbol{X}^{G})=f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c},\boldsymbol{X}^{G})f(\boldsymbol{X}^{G^c},\boldsymbol{X}^{G}) \\
+%			&=& f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c},\boldsymbol{X}^{G})f(\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G}) \\
+%		\end{eqnarray}
+%		
+%		\begin{equation}
+%			
+%		\end{equation}
+	%It can also be extended to the $(S,R,U,W)$ model with variables of $\boldsymbol{X}_f$ independent from $\boldsymbol{X}_r$ defining ...
+	The presence of two Gaussian mixtures on the right of the sub-regression gives the identifiability of it (even with also Gaussian variables) by the strict increase of the component number when combining Gaussian mixtures.
+	So a sub-regression is identifiable if it has only Gaussian mixtures (even only one) on the right or if it has at least two Gaussian mixtures (even with additional Gaussian variables)
 
@@
 %	
-If all the variables in $\boldsymbol{X}$ are Gaussian mixtures with at least two distinct components, then the model of sub-regression is identifiable (from the identifiability of the SR model by Maugis \cite{maugis2009variable}).
-
-But making this hypothesis on $\boldsymbol{X}_r$ is sufficient. 
-In the followings, we call Gaussian mixture the Gaussian mixtures with at least two distinct components, and Gaussian mixtures with equal components are called Gaussian (in fact they are).
-
-First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there must be at least one Gaussian mixture on the right of each sub-regression.
-So we have $\boldsymbol{X}^G \subsetneq \boldsymbol{X}_f$ containing Gaussian variables and we note the Gaussian mixtures $\boldsymbol{X}^{G^c}\neq \emptyset$ its complement in $\boldsymbol{X}_f$.
-
-The theorem from Maugis guarantee that a sub-regression between Gaussian mixtures is identifiable.
-		\begin{eqnarray}
-			\forall j \in I_r, \boldsymbol{X}^j_{|\boldsymbol{X}^G,\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^G\boldsymbol{\alpha}_j^G+\boldsymbol{X}^{G^c}\boldsymbol{\alpha}_j^{G^c}+ \boldsymbol{\varepsilon}_j \\
-			\boldsymbol{X}^j_{|\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^{G^c}\boldsymbol{\alpha}_j^{G^c} + \tilde{\boldsymbol{\varepsilon}}_j \textrm{ is identifiable where} \\
-			\tilde{\boldsymbol{\varepsilon}}_j&=&\boldsymbol{X}^{G}\boldsymbol{\alpha}_j^{G}+ \boldsymbol{\varepsilon}_j \textrm{ is Gaussian. And we have the unique decomposition} \\
-			f(\boldsymbol{X}^j|\boldsymbol{X}^G,\boldsymbol{X}^{G^c})&=&f(\boldsymbol{X}^j|\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G})
-		\end{eqnarray}
-		We have the unique (Gaussian are allowed only on the right of the sub-regressions) decomposition $f(\boldsymbol{X}_r|\boldsymbol{X}_f)=f(\boldsymbol{X}_r|\boldsymbol{X}^G,\boldsymbol{X}^{G^c})=f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G}) $
-		where  $f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})$ is identifiable. Moreover, $\boldsymbol{X}^{G}$ is a set (empty or not) of independent Gaussian variables, whose parameters are identifiable.
-		So the structure is identifiable.
+%If all the variables in $\boldsymbol{X}$ are Gaussian mixtures with at least two distinct components, then the model of sub-regression is identifiable (from the identifiability of the SR model by Maugis \cite{maugis2009variable}).
+%		
+%But making this hypothesis on $\boldsymbol{X}_r$ is sufficient. 
+%In the followings, we call Gaussian mixture the Gaussian mixtures with at least two distinct components, and Gaussian mixtures with equal components are called Gaussian (in fact they are).
 

### 2014-05-15 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@604 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	%It can also be extended to the $(S,R,U,W)$ model with variables of $\boldsymbol{X}_f$ independent from $\boldsymbol{X}_r$ defining ...
+
+Here, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
+	\begin{eqnarray}
+			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
+			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j  \label{densitydroite}\\
+			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
+						\mathcal{GM}(\bigotimes_{\substack{i \in I_f^j  }}\boldsymbol{\pi}_i \ ; 
+			                         \bigoplus_{\substack{i \in I_f^j  }} \alpha_{i,j} \boldsymbol{\mu}_i \ ; 
+			                         \boldsymbol{\sigma}_j^2+\bigoplus_{\substack{i \in I_f^j  }}\alpha_{i,j}^2 \boldsymbol{\sigma}_i^2 )\label{densitygauche}          
+		\end{eqnarray}
+		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
+
 	The presence of two Gaussian mixtures on the right of the sub-regression gives the identifiability of it (even with also Gaussian variables) by the strict increase of the component number when combining Gaussian mixtures.

### 2014-05-16 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@605 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals.
-For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$.
+For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$, variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$.
 The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign.
-$S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_f^j)=5$.
-	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.
+$\forall j \in I_r, p_1^j=2$ (sub-regressions of length 2) and we have $p_r=16$ sub-regressions.
+	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
 	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$ (see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
+	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
+This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
 		\subsection{Finding the structure}
@@
 
-We observe that \textsc{CorReg} is better than classical methods especially with the Hierarchical Uniform law.
-When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods, so it is robust.
-We also observe in table \ref{YXlinstep} that simple methods like stepwise (here from the package \textsc{lars}) can give good results in prediction.
-
-When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
-This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$& $P(S)$ & indicator &OLS  &    \textsc{CorReg} $\hat S$ + OLS & \textsc{CorReg} $S$ + OLS\\ 
-\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & $BIC$ & MSE (sd) & 262627.57 (732019) & 5928332.4 (49005690.2) & 262627.57 (732019) \\
-& & & cpl (sd) & 30 (0) & 29.99 (0.1) & 30 (0) \\
- &  & $BIC_+$ &MSE (sd) & 262627.57 (732019) & 10381962.64 (90962496.9) & 262627.57 (732019) \\
-& & & cpl (sd) & 30 (0) & 30 (0) & 30 (0) \\
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 510747.53 (2287539.8) & 635.42 (335.2) & 610.32 (424.9) \\
-& & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
- &  & $BIC_+$ &MSE (sd) & 510747.53 (2287539.8) & 603.24 (415.6) & 610.32 (424.9) \\
-& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25 (0) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 178323.95 (1426610.2) & 180.02 (44.9) & 141.03 (27.8) \\
-& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
- &  & $BIC_+$ & MSE (sd) & 178323.95 (1426610.2) & 330.31 (134.8) & 141.03 (27.8) \\
-& & & cpl (sd) & 30 (0) & 21 (0) & 9 (0) \\
-\hline
-\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & $BIC$ &MSE (sd) & 528.08 (228.6) & 886.54 (364.3) & 528.08 (228.6) \\
-& & & cpl (sd) & 41 (0) & 36.8 (1.7) & 41 (0) \\
- &  & $BIC_+$ &MSE (sd) & 528.08 (228.6) & 542.29 (235) & 528.08 (228.6) \\
-& & & cpl (sd) & 41 (0) & 40.87 (0.4) & 41 (0) \\
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &MSE (sd) & 612.72 (291.5) & 239.17 (89.4) & 200.32 (42.6) \\	
-& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
- &  & $BIC_+$ &MSE (sd) & 612.72 (291.5) & 207.6 (51.3) & 200.32 (42.6) \\
-& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25 (0) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 555.44 (262.5) & 128.98 (18.1) & 121.08 (11.9) \\
-& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
- &  & $BIC_+$ &MSE (sd) & 555.44 (262.5) & 171.9 (31.1) & 121.08 (11.9) \\
-& & &cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
-\hline
-\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & $BIC$ & MSE (sd) & 167.71 (20.6) & 323.44 (124.1) & 167.71 (20.6) \\
-& & & cpl (sd) & 41 (0) & 38.17 (1.3) & 41 (0) \\
- &  & $BIC_+$ &MSE (sd) & 167.71 (20.6) & 167.98 (20.8) & 167.71 (20.6) \\  
-& & & cpl (sd) & 41 (0) & 40.99 (0.1) & 41 (0) \\
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ &  MSE (sd) & 168.68 (22.4) & 158.51 (51.6) & 133.49 (12.2) \\
-& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
- &  & $BIC_+$ &  MSE (sd) & 168.68 (22.4) & 137.03 (20.3) & 133.49 (12.2) \\
-& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25 (0) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ &  MSE (sd) & 173.25 (22.9) & 112.8 (10.7) & 110.63 (6.9) \\
-& & & cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
- &  & $BIC_+$ & MSE (sd) & 173.25 (22.9) & 127.4 (11.9) & 110.63 (6.9) \\
-& & & cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
-\hline
-\end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_tout_MSE_NB.png} 
+			\caption{Comparison of the MSE between OLS and CorReg+OLS}\label{graphMixmod}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_tout_compl_NB.png} 
+			\caption{Comparison of the compexities between OLS and CorReg+OLS} \label{compareMSEexfos}
+   \end{minipage}
+\end{figure}
+
+%\caption{OLS and OLS combined with constrained \textsc{CorReg}.
 $\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
@@
 Without selection, all models have min$(n,p)$ non-zero coefficients.} \label{YXlinOLS}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&  $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$ + LASSO& \textsc{CorReg} $S$ + LASSO\\ 
-\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & $BIC$ &MSE (sd) & 1246.35 (350.5) & 1433.98 (526.7) & 1246.35 (350.5) \\
-& & & cpl (sd) & 17.84 (5.5) & 15.79 (5.6) & 17.84 (5.5) \\
- &  & $BIC_+$ &MSE (sd) & 1246.35 (350.5) & 1248.84 (341.4) & 1246.35 (350.5) \\
-& & & cpl (sd) & 17.84 (5.5) & 17.63 (5.6) & 17.84 (5.5) \\ 
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 &16& $BIC$ & MSE (sd) & 712.79 (405.2) & 566.32 (226.4) & 554.01 (257.1) \\
-& & & cpl (sd) & 16.68 (4.4) & 15.35 (4.3) & 16.14 (4.5) \\
- &  & $BIC_+$ & MSE (sd) & 712.79 (405.2) & 551.92 (242.4) & 554.01 (257.1) \\
-& & & cpl (sd) & 16.68 (4.4) & 15.96 (4.4) & 16.14 (4.5) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 216.12 (128.3) & 157.78 (39.4) & 147.04 (28.4) \\
-& & & cpl (sd) & 11.23 (4.4) & 8.56 (1.8) & 7.91 (1.3) \\
- &  & $BIC_+$ & MSE (sd) & 216.12 (128.3) & 178.88 (68.9) & 147.04 (28.4) \\
-& & & cpl (sd) & 11.23 (4.4) & 9.86 (3) & 7.91 (1.3) \\
-\hline
-\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & $BIC$ &MSE (sd) & 658.38 (221.4) & 872.32 (239.4) & 658.38 (221.4) \\
-& & & cpl (sd) & 28.2 (6) & 22.95 (5.8) & 28.2 (6) \\
- &  & $BIC_+$ &MSE (sd) & 658.38 (221.4) & 652.72 (211.9) & 658.38 (221.4) \\
-& & & cpl (sd) & 28.2 (6) & 28.36 (5.9) & 28.2 (6) \\
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &MSE (sd) & 274.34 (101.7) & 268.76 (106.3) & 225.64 (67.5) \\	
-& & & cpl (sd) & 22.14 (4.1) & 19.52 (2.9) & 20.47 (2.8) \\
- &  & $BIC_+$ &MSE (sd) & 274.34 (101.7) & 233.8 (78.9) & 225.64 (67.5) \\
-& & & cpl (sd) & 22.14 (4.1) & 20.19 (2.8) & 20.47 (2.8) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 165.6 (73.7) & 128.34 (18.8) & 123.73 (12.8) \\
-& & & cpl (sd) & 12.93 (6.4) & 8.51 (1.5) & 8.09 (1.1) \\
- &  & $BIC_+$ &MSE (sd) & 165.6 (73.7) & 135.84 (24.1) & 123.73 (12.8) \\
-& & &cpl (sd) & 12.93 (6.4) & 10.17 (3.1) & 8.09 (1.1) \\
-\hline
-\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & $BIC$ &  MSE (sd) & 183.33 (31.1) & 357.15 (133.4) & 183.33 (31.1) \\
-& & & cpl (sd) & 37.78 (2.4) & 32.93 (3.7) & 37.78 (2.4) \\
- &  & $BIC_+$ &  MSE (sd) & 183.33 (31.1) & 183.78 (31.6) & 183.33 (31.1) \\
-& & &  cpl (sd) & 37.78 (2.4) & 37.75 (2.4) & 37.78 (2.4) \\
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ & MSE (sd) & 148.83 (18.6) & 164.44 (54.3) & 139.57 (16.1) \\
-& & & cpl (sd) & 25.22 (4) & 21.72 (2) & 22.3 (1.7) \\
- &  & $BIC_+$ &  MSE (sd) & 148.83 (18.6) & 142.97 (22.1) & 139.57 (16.1) \\
-& & & cpl (sd) & 25.22 (4) & 22.24 (1.8) & 22.3 (1.7) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ & MSE (sd) & 124.29 (19.9) & 113.36 (11) & 111.54 (7.2) \\
-& & & cpl (sd) & 13.18 (5.9) & 8.68 (1.2) & 8.53 (1) \\
- &  & $BIC_+$ & MSE (sd) & 124.29 (19.9) & 118.33 (12) & 111.54 (7.2) \\
-& & & cpl (sd) & 13.18 (5.9) & 11.02 (2.9) & 8.53 (1) \\
-\hline
-\end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.
-$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
-\textsc{CorReg} logically wins.
-LASSO is better than OLS but is improved by \textsc{CorReg} even for large values of $n$.}\label{YXlinLASSO}
-\end{table}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/lar_tout_MSE_NB.png} 
+			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}\label{graphMixmod}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/lar_tout_compl_NB.png} 
+			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} \label{compareMSEexfos}
+   \end{minipage}
+\end{figure}
 

### 2014-05-16 www/article13.tex
up curves
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@606 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 $\forall j \in I_r, p_1^j=2$ (sub-regressions of length 2) and we have $p_r=16$ sub-regressions.
-	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
+The datasets were then scaled so that covariates $X_r$ don't have a greater variance or mean.
+	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
 	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$ (see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
@@
 This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
+	We compare different methods with and without CorReg as a pretreatment. All the results are provided by the CorReg package.
 		\subsection{Finding the structure}
@@
 \end{tabular} 
-\caption{Results of the Markov chain with  constraint $\hat{p}_f\leq 5$. Mean observed and standard deviation (sd). } \label{compZvrai}
+\caption{Results of the Markov chain: Mean observed and standard deviation (sd). } \label{compZvrai}
 \end{table}
 
+\begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/BIC_p2.png} 
+			\caption{Quality of the subregressions found with classical $BIC$ criterion}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/BICSTAR_P2.png} 
+			\caption{Quality of the subregressions found with our $BIC_+$ criterion} 
+   \end{minipage}
+\end{figure}
+
 \clearpage
-\subsection{Results on prediction}
-	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 \label{tableMSEsimtout}	
-We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure).
-The datasets used here were those from table \ref{compZvrai}.
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_tout_MSE_NB.png} 
-			\caption{Comparison of the MSE between OLS and CorReg+OLS}\label{graphMixmod}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_tout_compl_NB.png} 
-			\caption{Comparison of the compexities between OLS and CorReg+OLS} \label{compareMSEexfos}
+\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_f}$ (best case for us)}	 \label{tableMSEsimgauche}
+\begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X1_MSE_NB.png} 
+			\caption{Comparison of the MSE between OLS and CorReg+OLS}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
+			\caption{Comparison of the compexities between OLS and CorReg+OLS} 
    \end{minipage}
@@
 	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X1_MSE_NB.png} 
+			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X1_compl_NB.png} 
+			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} 
+   \end{minipage}
+\end{figure}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_X1_MSE_NB.png} 
+			\caption{Comparison of the MSE between elasticnet and CorReg+elasticnet}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_X1_compl_NB.png} 
+			\caption{Comparison of the compexities between elasticnet and CorReg+elasticnet} 
+   \end{minipage}
+\end{figure}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_X1_MSE_NB.png} 
+			\caption{Comparison of the MSE between stepwise and CorReg+stepwise}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_X1_compl_NB.png} 
+			\caption{Comparison of the compexities between stepwise and CorReg+stepwise} 
+   \end{minipage}
+\end{figure}
+
+\clearpage
+\subsection{Results on prediction}
+	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 	
+We then try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure).
+The datasets used here were those from table \ref{compZvrai}.
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_tout_MSE_NB.png} 
+			\caption{Comparison of the MSE between OLS and CorReg+OLS}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_tout_compl_NB.png} 
+			\caption{Comparison of the compexities between OLS and CorReg+OLS} 
+   \end{minipage}
+\end{figure}
+
+%\caption{OLS and OLS combined with constrained \textsc{CorReg}.
+$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$.
+\textsc{CorReg} logically wins.
+When $n<p$ the dataset was reduced to $p=n$ automatically by a $QR$ decomposition as the lm function of R does.
+Without selection, all models have min$(n,p)$ non-zero coefficients.} \label{YXlinOLS}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
 			\includegraphics[height=180px,width=245px]{figures/res_article/lar_tout_MSE_NB.png} 
-			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}\label{graphMixmod}
+			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}
 	\end{minipage} \
@@
 			\includegraphics[height=180px,width=240px]{figures/res_article/lar_tout_compl_NB.png} 
-			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} \label{compareMSEexfos}
-   \end{minipage}
-\end{figure}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&  $P(S)$&indicator &Elasticnet  &    \textsc{CorReg} $\hat S$ + Elasticnet& \textsc{CorReg} $S$ + Elasticnet\\ 
-\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & $BIC$ &MSE (sd) & 1 326.54 (388.4) & 1 356.98 (331.4) & 1 326.54 (388.4) \\
-& & & cpl (sd) & 12.14 (5.1) & 12.5 (5.2) & 12.14 (5.1) \\
- &  & $BIC_+$ & MSE (sd) & 1326.54 (388.4) & 1307.96 (356.6) & 1326.54 (388.4) \\
-& & & cpl (sd) & 12.14 (5.1) & 12.27 (5.1) & 12.14 (5.1) \\
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 1 400.56 (1598.2) & 668.57 (274.6) & 653.59 (283.7) \\
-& & & cpl (sd) & 13.86 (6.8) & 14.31 (5.7) & 14.83 (5.3) \\
- &  & $BIC_+$ &MSE (sd) & 1 400.56 (1598.2) & 643.25 (277.9) & 653.59 (283.7) \\
-& & & cpl (sd) & 13.86 (6.8) & 14.83 (5.6) & 14.83 (5.3) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 855.74 (582.6) & 181.08 (51.9) & 146.79 (32.7) \\
-& & & cpl (sd) & 15.57 (6.3) & 11.19 (2.4) & 8.37 (1.3) \\
- &  & $BIC_+$ & MSE (sd) & 855.74 (582.6) & 311.57 (163.9) & 146.79 (32.7) \\
-& & & cpl (sd) & 15.57 (6.3) & 14.87 (3.9) & 8.37 (1.3) \\
-\hline
-\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & $BIC$ &MSE (sd) & 738.94 (254.5) & 914.56 (283.7) & 738.94 (254.5) \\
-& & & cpl (sd) & 25.85 (8.7) & 20.97 (7.8) & 25.85 (8.7) \\
- &  & $BIC_+$ &MSE (sd) & 738.94 (254.5) & 751.06 (262.2) & 738.94 (254.5) \\
-& & & cpl (sd) & 25.85 (8.7) & 25.43 (8.8) & 25.85 (8.7) \\
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &	MSE (sd) & 516.36 (226.6) & 276.6 (133.6) & 228.85 (89.3) \\
-& & & cpl (sd) & 26.72 (7.4) & 20.49 (3.8) & 21.8 (3.3) \\
- &  & $BIC_+$ &MSE (sd) & 516.36 (226.6) & 239.77 (103.7) & 228.85 (89.3) \\
-& & & cpl (sd) & 26.72 (7.4) & 21.57 (3.4) & 21.8 (3.3) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 328.7 (146.3) & 130.61 (20.6) & 124.01 (15.4) \\
-& & & cpl (sd) & 24.7 (7.1) & 9.95 (1.5) & 8.22 (1.2) \\
- &  & $BIC_+$ &MSE (sd) & 328.7 (146.3) & 169.13 (40.2) & 124.01 (15.4) \\
-& & &cpl (sd) & 24.7 (7.1) & 16.04 (3.8) & 8.22 (1.2) \\
-\hline
-\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & $BIC$ & MSE (sd) & 175.93 (25.9) & 355.79 (143.8) & 175.93 (25.9) \\
-& & & cpl (sd) & 39.71 (1.7) & 34.11 (4.6) & 39.71 (1.7) \\
- &  & $BIC_+$ &  MSE (sd) & 175.93 (25.9) & 176.2 (26.3) & 175.93 (25.9) \\
-& & &  cpl (sd) & 39.71 (1.7) & 39.69 (1.7) & 39.71 (1.7) \\
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ &  MSE (sd) & 173.65 (25.1) & 164.96 (59) & 139.51 (17.6) \\
-& & &  cpl (sd) & 35.54 (4.2) & 22.58 (2.2) & 23.26 (1.9) \\
- &  & $BIC_+$ &  MSE (sd) & 173.65 (25.1) & 143.04 (23.2) & 139.51 (17.6) \\
-& & & cpl (sd) & 35.54 (4.2) & 23.16 (1.9) & 23.26 (1.9) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ &  MSE (sd) & 161.92 (25.2) & 113.61 (10.9) & 111.46 (7.2) \\
-& & &  cpl (sd) & 30.78 (6.4) & 9.25 (1.2) & 8.67 (1) \\
- &  & $BIC_+$ &  MSE (sd) & 161.92 (25.2) & 127.44 (13.3) & 111.46 (7.2) \\
-& & &cpl (sd) & 30.78 (6.4) & 17.85 (2.7) & 8.67 (1) \\
-\hline
-\end{tabular} 
-\caption{Elasticnet combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}$. \textsc{CorReg} logically wins. LASSO was better.}\label{YXlinenet}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$ + Stepwise& \textsc{CorReg} $S$ + Stepwise\\ 
-\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & $BIC$ & MSE (sd) & 1 919.43 (861.8) & 2 014.13 (683.6) & 1 919.43 (861.8) \\
-& & &cpl (sd) & 22.8 (3.3) & 19.34 (4.9) & 22.8 (3.3) \\
- &  & $BIC_+$ &MSE (sd) & 1 919.43 (861.8) & 1 885.39 (814) & 1 919.43 (861.8) \\
-& & & cpl (sd) & 22.8 (3.3) & 22.73 (3.1) & 22.8 (3.3) \\
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 696.46 (492.9) & 662.07 (285.9) & 660.93 (354) \\
-& & & cpl (sd) & 15.9 (3.8) & 15.1 (3.7) & 15.76 (3.7) \\
- &  & $BIC_+$ &MSE (sd) & 696.46 (492.9) & 655.94 (325.7) & 660.93 (354) \\
-& & & cpl (sd) & 15.9 (3.8) & 15.72 (3.6) & 15.76 (3.7) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 394.93 (356) & 155.18 (36.4) & 147.81 (30.3) \\
-& & &cpl (sd) & 16.1 (6.3) & 8.26 (1.7) & 7.77 (1.3) \\
- &  & $BIC_+$ &  MSE (sd) & 394.93 (356) & 189.49 (79) & 147.81 (30.3) \\
-& & & cpl (sd) & 16.1 (6.3) & 10.23 (3) & 7.77 (1.3) \\
-\hline
-\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & $BIC$ &MSE (sd) & 745.88 (241.7) & 962.12 (319.3) & 745.88 (241.7) \\
-& & & cpl (sd) & 27.19 (4.6) & 22.9 (4.5) & 27.19 (4.6) \\
- &  & $BIC_+$ &MSE (sd) & 745.88 (241.7) & 749.08 (244.2) & 745.88 (241.7) \\
-& & & cpl (sd) & 27.19 (4.6) & 27.21 (4.5) & 27.19 (4.6) \\
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &	MSE (sd) & 279.82 (151.1) & 290.13 (117.3) & 239.61 (69.5) \\
-& & & cpl (sd) & 21.24 (5.2) & 18.13 (2.8) & 19.23 (2.7) \\
- &  & $BIC_+$ &MSE (sd) & 279.82 (151.1) & 252.35 (108.5) & 239.61 (69.5) \\
-& & & cpl (sd) & 21.24 (5.2) & 18.97 (2.7) & 19.23 (2.7) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 199.3 (158.7) & 126.81 (17.4) & 124.19 (13) \\
-& & & cpl (sd) & 14.31 (6.2) & 8.16 (1.2) & 8.04 (1) \\
- &  & $BIC_+$ &MSE (sd) & 199.3 (158.7) & 136.32 (25.2) & 124.19 (13) \\
-& & &cpl (sd) & 14.31 (6.2) & 9.59 (2.6) & 8.04 (1) \\
-\hline
-\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & $BIC$ &   MSE (sd) & 193.26 (34.9) & 387.18 (151.6) & 193.26 (34.9) \\
-& & & cpl (sd) & 36.3 (2.5) & 30.94 (4.2) & 36.3 (2.5) \\
- &  & $BIC_+$ &   MSE (sd) & 193.26 (34.9) & 194.43 (37.3) & 193.26 (34.9) \\
-& & &  cpl (sd) & 36.3 (2.5) & 36.22 (2.7) & 36.3 (2.5) \\
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ &  MSE (sd) & 145.46 (20.1) & 165.74 (56.1) & 140.27 (17.5) \\
-& & & cpl (sd) & 23.44 (3.8) & 21.23 (2.1) & 21.93 (1.7) \\
- &  & $BIC_+$ &  MSE (sd) & 145.46 (20.1) & 143.74 (23.2) & 140.27 (17.5) \\
-& & & cpl (sd) & 23.44 (3.8) & 21.86 (1.7) & 21.93 (1.7) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ & MSE (sd) & 128.44 (18.4) & 113.27 (10.7) & 111.77 (7.2) \\
-& & &  cpl (sd) & 13.58 (5.2) & 8.59 (1.1) & 8.48 (1) \\
- &  & $BIC_+$ &  MSE (sd) & 128.44 (18.4) & 118.61 (11.5) & 111.77 (7.2) \\
-& & & cpl (sd) & 13.58 (5.2) & 10.75 (2.8) & 8.48 (1) \\
-\hline
-\end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}.
-$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$.
-\textsc{CorReg} logically wins but stepwise is quite good compared to elasticnet.}\label{YXlinstep}
-\end{table}
+			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} 
+   \end{minipage}
+\end{figure}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_tout_MSE_NB.png} 
+			\caption{Comparison of the MSE between elasticnet and CorReg+elasticnet}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_tout_compl_NB.png} 
+			\caption{Comparison of the compexities between elasticnet and CorReg+elasticnet} 
+   \end{minipage}
+\end{figure}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_tout_MSE_NB.png} 
+			\caption{Comparison of the MSE between stepwise and CorReg+stepwise}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_tout_compl_NB.png} 
+			\caption{Comparison of the compexities between stepwise and CorReg+stepwise} 
+   \end{minipage}
+\end{figure}
+
+We see that CorReg tends to give more parsimonious models and better predictions, even if the true model is not parsomious.
+We logically observe that when $n$ rises, all the models get better and the correlations cease to be a problem so the complete model starts to be better (CorReg does not allow the true model to be choosen).
 
@@
 We now try the method with a response depending only on variables in $\boldsymbol{X}_r$. The datasets used here were still those from \ref{compZvrai}.
-Depending only on $\boldsymbol{X}_r$ imply sparsity and impossibility to obtain the true model when using the true structure. \textsc{CorReg} is still better than classical methods.
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&$P(S)$ &indicator &OLS  &    \textsc{CorReg} $\hat S$ + OLS& \textsc{CorReg} $S$ + OLS\\ 
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 81781.54 (247281.7) & 507.1 (269.2) & 593.03 (428.3) \\
-& & & cpl (sd) & 30 (0) & 24.11 (1) & 25 (0) \\
- &  & $BIC_+$ &MSE (sd) & 81781.54 (247281.7) & 587.07 (425.6) & 593.03 (428.3) \\
-& & & cpl (sd) & 30 (0) & 24.82 (0.6) & 25 (0) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 74136.81 (297716.1) & 189.45 (48.1) & 144.51 (31.4) \\
-& & & cpl (sd) & 30 (0) & 13.85 (0.9) & 9 (0) \\
- &  & $BIC_+$ & MSE (sd) & 74136.81 (297716.1) & 340.53 (169.4) & 144.51 (31.4) \\
-& & & cpl (sd) & 30 (0) & 21 (0) & 9 (0) \\
-\hline
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &	MSE (sd) & 684.04 (438.1) & 190.97 (37.7) & 197.32 (40) \\
-& & & cpl (sd) & 41 (0) & 24.43 (0.7) & 25 (0) \\
- &  & $BIC_+$ &MSE (sd) & 684.04 (438.1) & 196.36 (40.5) & 197.32 (40) \\
-& & & cpl (sd) & 41 (0) & 24.99 (0.5) & 25 (0) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 596.32 (323) & 126.94 (15.5) & 119.39 (12.4) \\
-& & & cpl (sd) & 41 (0) & 11.45 (0.9) & 9 (0) \\
- &  & $BIC_+$ &MSE (sd) & 596.32 (323) & 168.03 (27.7) & 119.39 (12.4) \\
-& & &cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
-\hline
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ &  MSE (sd) & 168.35 (20.1) & 133.75 (12.4) & 135.08 (13.1) \\
-& & & cpl (sd) & 41 (0) & 24.53 (0.7) & 25 (0) \\
- &  & $BIC_+$ &  MSE (sd) & 168.35 (20.1) & 135.04 (13.1) & 135.08 (13.1) \\
-& & & cpl (sd) & 41 (0) & 25.01 (0.4) & 25 (0) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ &  MSE (sd) & 168.07 (21.4) & 109.61 (7.4) & 108.64 (7.2) \\
-& & &  cpl (sd) & 41 (0) & 10.02 (0.8) & 9 (0) \\
- &  & $BIC_+$ & MSE (sd) & 168.07 (21.4) & 124.23 (11.1) & 108.64 (7.2) \\
-& & & cpl (sd) & 41 (0) & 21 (0) & 9 (0) \\
-\hline
-\end{tabular} 
-\caption{OLS and OLS combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}_r$.
-Sometimes $\hat{S}$ gives better results than $S$ because $S$ is penalized by the fact that it relies only on covariates not in the true model.
-} \label{YX2linOLS}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&  $P(S)$ &indicator &LAR  &    \textsc{CorReg} $\hat S$ + LAR& \textsc{CorReg} $S$ + LAR\\ 
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 367.45 (198.4) & 298.32 (122.4) & 292.53 (102.2) \\
-& & & cpl (sd) & 14.37 (5.1) & 12.9 (3.8) & 12.84 (3.8) \\
- &  & $BIC_+$ &MSE (sd) & 367.45 (198.4) & 298.1 (119.1) & 292.53 (102.2) \\
-& & & cpl (sd) & 14.37 (5.1) & 12.81 (3.8) & 12.84 (3.8) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 273.12 (267.2) & 166.93 (49) & 152.45 (39.5) \\
-& & & cpl (sd) & 11.7 (5.8) & 8.7 (2.5) & 7.45 (1.4) \\
- &  & $BIC_+$ & MSE (sd) & 273.12 (267.2) & 175.68 (57.5) & 152.45 (39.5) \\
-& & & cpl (sd) & 11.7 (5.8) & 9.19 (2.9) & 7.45 (1.4) \\
-\hline
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &	MSE (sd) & 189.64 (52.5) & 171.97 (45.9) & 176.58 (46.8) \\
-& & & cpl (sd) & 14.15 (3.9) & 13.4 (2.8) & 13.5 (3.1) \\
- &  & $BIC_+$ &MSE (sd) & 189.64 (52.5) & 174.36 (45.3) & 176.58 (46.8) \\
-& & & cpl (sd) & 14.15 (3.9) & 13.57 (3) & 13.5 (3.1) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 163.88 (67.4) & 124.53 (17) & 121.99 (16.6) \\
-& & & cpl (sd) & 13.06 (7.2) & 8.29 (1.6) & 7.73 (1.2) \\
- &  & $BIC_+$ &MSE (sd) & 163.88 (67.4) & 133.7 (26.4) & 121.99 (16.6) \\
-& & &cpl (sd) & 13.06 (7.2) & 9.65 (3.3) & 7.73 (1.2) \\
-\hline
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ & MSE (sd) & 129.65 (14.8) & 127.11 (12.1) & 127.46 (12.2) \\
-& & &  cpl (sd) & 14.86 (3.1) & 13.97 (2.5) & 14.04 (2.5) \\
- &  & $BIC_+$ &  MSE (sd) & 129.65 (14.8) & 127.46 (12.2) & 127.46 (12.2) \\
-& & & cpl (sd) & 14.86 (3.1) & 14.02 (2.5) & 14.04 (2.5) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ &  MSE (sd) & 121.21 (20.1) & 109.66 (7.9) & 109.04 (7.8) \\
-& & & cpl (sd) & 12.84 (5.2) & 8.42 (1.1) & 8.11 (0.8) \\
- &  & $BIC_+$ &  MSE (sd) & 121.21 (20.1) & 112.56 (11.2) & 109.04 (7.8) \\
-& & & cpl (sd) & 12.84 (5.2) & 9.77 (2.3) & 8.11 (0.8) \\
-\hline
-\end{tabular} 
-\caption{LASSO (with LAR) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}_r$. }\label{YX2linLASSO}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&  $P(S)$ &indicator & Elasticnet  &    \textsc{CorReg} $\hat S$ + Elasticnet& \textsc{CorReg} $S$ + Elasticnet\\ 
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 499.32 (218.8) & 311.41 (137.4) & 305.09 (123.5) \\
-& & & cpl (sd) & 14.08 (6.1) & 11.75 (4.8) & 11.49 (4.7) \\
- &  & $BIC_+$ &MSE (sd) & 499.32 (218.8) & 307.6 (128.7) & 305.09 (123.5) \\
-& & & cpl (sd) & 14.08 (6.1) & 11.63 (4.7) & 11.49 (4.7) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 691.55 (503.2) & 193.45 (54.2) & 150.17 (35) \\
-& & & cpl (sd) & 12.54 (7.2) & 10.26 (2.7) & 7.81 (1.4) \\
- &  & $BIC_+$ & MSE (sd) & 691.55 (503.2) & 279.21 (117.3) & 150.17 (35) \\
-& & & cpl (sd) & 12.54 (7.2) & 13 (4.3) & 7.81 (1.4) \\
-\hline
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &	MSE (sd) & 282.05 (134.8) & 180.64 (47.6) & 181.49 (48.6) \\
-& & & cpl (sd) & 20.37 (6.3) & 13.68 (4) & 14.05 (4) \\
- &  & $BIC_+$ &MSE (sd) & 282.05 (134.8) & 180.84 (47.5) & 181.49 (48.6) \\
-& & & cpl (sd) & 20.37 (6.3) & 14.03 (4) & 14.05 (4) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 308.13 (142.9) & 129.39 (19.6) & 122.71 (17.8) \\
-& & & cpl (sd) & 22.94 (7.7) & 9.37 (2.1) & 7.91 (1.4) \\
- &  & $BIC_+$ &MSE (sd) & 308.13 (142.9) & 159.14 (30.4) & 122.71 (17.8) \\
-& & &cpl (sd) & 22.94 (7.7) & 15.42 (4) & 7.91 (1.4) \\
-\hline
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ & MSE (sd) & 150.14 (19.7) & 127.48 (14) & 128.21 (15) \\
-& & &  cpl (sd) & 25.6 (5.4) & 14.73 (3.8) & 14.73 (3.9) \\
- &  & $BIC_+$ &  MSE (sd) & 150.14 (19.7) & 128.21 (14.9) & 128.21 (15) \\
-& & & cpl (sd) & 25.6 (5.4) & 14.7 (3.9) & 14.73 (3.9) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ & MSE (sd) & 157.18 (23.5) & 110.24 (8) & 109.24 (7.8) \\
-& & &  cpl (sd) & 30.41 (6.7) & 8.96 (1.3) & 8.3 (1) \\
- &  & $BIC_+$ &  MSE (sd) & 157.18 (23.5) & 123.8 (12.5) & 109.24 (7.8) \\
-& & & cpl (sd) & 30.41 (6.7) & 17.18 (3.3) & 8.3 (1) \\
-\hline
-\end{tabular} 
-\caption{Elasticnet (with Elasticnet) combined with constrained \textsc{CorReg}.$\boldsymbol{Y}$  depends on all variables in $\boldsymbol{X}_r$.}\label{YX2linenet}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$&  $P(S)$ &indicator &Stepwise  &    \textsc{CorReg} $\hat S$ + stepwise & \textsc{CorReg} $S$ + stepwise \\ 
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &MSE (sd) & 394.59 (454.6) & 341.47 (153.2) & 351.14 (148) \\
-& & & cpl (sd) & 13.09 (3.2) & 12.9 (2.9) & 12.92 (3) \\
- &  & $BIC_+$ &MSE (sd) & 394.59 (454.6) & 353.99 (157) & 351.14 (148) \\
-& & & cpl (sd) & 13.09 (3.2) & 12.83 (2.9) & 12.92 (3) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & MSE (sd) & 410.11 (343.5) & 163.05 (45.4) & 152.08 (38.1) \\
-& & & cpl (sd) & 15.8 (6.5) & 8.29 (2.2) & 7.34 (1.3) \\
- &  & $BIC_+$ & MSE (sd) & 410.11 (343.5) & 178.94 (69.7) & 152.08 (38.1) \\
-& & & cpl (sd) & 15.8 (6.5) & 8.89 (3) & 7.34 (1.3) \\
-\hline
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &	MSE (sd) & 180.67 (52.7) & 173.02 (38.3) & 174.84 (36.8) \\
-& & & cpl (sd) & 13.56 (3.7) & 12.78 (2.4) & 13.07 (2.4) \\
- &  & $BIC_+$ &MSE (sd) & 180.67 (52.7) & 175.3 (38.1) & 174.84 (36.8) \\
-& & & cpl (sd) & 13.56 (3.7) & 13.01 (2.4) & 13.07 (2.4) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ &MSE (sd) & 188.91 (88.5) & 124.38 (17.4) & 122.38 (16.7) \\
-& & & cpl (sd) & 14.74 (6.6) & 8.07 (1.6) & 7.65 (1.2) \\
- &  & $BIC_+$ &MSE (sd) & 188.91 (88.5) & 134.22 (23.6) & 122.38 (16.7) \\
-& & &cpl (sd) & 14.74 (6.6) & 9.5 (2.8) & 7.65 (1.2) \\
-\hline
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ &  MSE (sd) & 128.18 (13.7) & 127.43 (13.7) & 127.71 (13.3) \\
-& & &  cpl (sd) & 13.78 (2.8) & 13.51 (2) & 13.59 (2.1) \\
- &  & $BIC_+$ & MSE (sd) & 128.18 (13.7) & 127.71 (13.3) & 127.71 (13.3) \\
-& & & cpl (sd) & 13.78 (2.8) & 13.59 (2.1) & 13.59 (2.1) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ & MSE (sd) & 126.15 (20.6) & 109.47 (8) & 109.04 (7.8) \\
-& & & cpl (sd) & 13.25 (4.4) & 8.3 (1) & 8.12 (0.8) \\
- &  & $BIC_+$ &  MSE (sd) & 126.15 (20.6) & 113.96 (11.4) & 109.04 (7.8) \\
-& & & cpl (sd) & 13.25 (4.4) & 10.08 (2.6) & 8.12 (0.8) \\
-\hline
-\end{tabular} 
-\caption{Stepwise  combined with constrained \textsc{CorReg}. $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}_r$.}\label{YX2linstep}
-\end{table}
+Depending only on $\boldsymbol{X}_r$ implies sparsity and impossibility to obtain the true model when using the true structure.
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X2_MSE_NB.png} 
+			\caption{Comparison of the MSE between OLS and CorReg+OLS}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X2_compl_NB.png} 
+			\caption{Comparison of the compexities between OLS and CorReg+OLS} 
+   \end{minipage}
+\end{figure}
+\textsc{CorReg} is still better than OLS for strong correlations and limited values of $n$. 
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X2_MSE_NB.png} 
+			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X2_compl_NB.png} 
+			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} 
+   \end{minipage}
+\end{figure}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_X2_MSE_NB.png} 
+			\caption{Comparison of the MSE between elasticnet and CorReg+elasticnet}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_X2_compl_NB.png} 
+			\caption{Comparison of the compexities between elasticnet and CorReg+elasticnet} 
+   \end{minipage}
+\end{figure}
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_X2_MSE_NB.png} 
+			\caption{Comparison of the MSE between stepwise and CorReg+stepwise}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_X2_compl_NB.png} 
+			\caption{Comparison of the compexities between stepwise and CorReg+stepwise}
+   \end{minipage}
+\end{figure}
 

### 2014-05-16 www/article13.tex
up robust
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@607 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Robustess of the model}
- We have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
+% We have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
 Variables in $\boldsymbol{X_r}$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $\boldsymbol{X_f}$ .
@@
 One example of non-linear structure found with \textsc{CorReg} on real datasets is given in section \ref{sectionexfos}.
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-\hline
-\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
-\hline
-$n$ & $p_r$ & $P(S)$&Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
-\hline
-30 & 16 & $BIC$ & 0.4313 & 2.7769 & 9.96 & 2.45 & 5.95 & 3.5 & 29.42  \\
-& & &(0.0391) & (0.0883) & (1.3175) & (1.3808) & (1.3056) & (1.453) & (6.2639) \\
- &  & $BIC_+$ & 0.4313 & 4.9079 & 7.28 & 1.25 & 8.63 & 7.38 & 5.77  \\ 
-& & &(0.0391) & (0.4353) & (1.5769) & (0.9987) & (1.5548) & (1.6316) & (6.1297) \\ 
-\hline
-\end{tabular} 
-\caption{Results of the Markov chain for non linear structure ($\log$ and square). Mean observed and standard deviation (sd). } \label{compZnonlin}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline 
-$n$ & $p_r$& $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
-\hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
-30 & 16 & $BIC$ &MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
-& & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\
- &  &$BIC_+$ &MSE (sd) & 996.06 (323.7) & 1094.5 (621.8) & 1501.25 (905.9) \\
-& & & cpl (sd) & 17.66 (5.4) & 16.98 (4.8) & 11.97 (4.9) \\ 
-\hline
-\end{tabular} 
-\caption{ $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$. \textsc{CorReg} is not too far from LASSO even if it stays behind. } \label{resYnonlin}
-\end{table}
+%\begin{table}[h!]
+%\centering
+%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+%\hline
+%\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
+%\hline
+%$n$ & $p_r$ & $P(S)$&Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
+%\hline
+%30 & 16 & $BIC$ & 0.4313 & 2.7769 & 9.96 & 2.45 & 5.95 & 3.5 & 29.42  \\
+%& & &(0.0391) & (0.0883) & (1.3175) & (1.3808) & (1.3056) & (1.453) & (6.2639) \\
+% &  & $BIC_+$ & 0.4313 & 4.9079 & 7.28 & 1.25 & 8.63 & 7.38 & 5.77  \\ 
+%& & &(0.0391) & (0.4353) & (1.5769) & (0.9987) & (1.5548) & (1.6316) & (6.1297) \\ 
+%\hline
+%\end{tabular} 
+%\caption{Results of the Markov chain for non linear structure ($\log$ and square). Mean observed and standard deviation (sd). } \label{compZnonlin}
+%\end{table}
+%
+%
+%\begin{table}[h!]
+%\centering
+%\begin{tabular}{|c|c|c|c|c|c|c|}
+%\hline 
+%$n$ & $p_r$& $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
+%\hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
+%30 & 16 & $BIC$ &MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
+%& & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\
+% &  &$BIC_+$ &MSE (sd) & 996.06 (323.7) & 1094.5 (621.8) & 1501.25 (905.9) \\
+%& & & cpl (sd) & 17.66 (5.4) & 16.98 (4.8) & 11.97 (4.9) \\ 
+%\hline
+%\end{tabular} 
+%\caption{ $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$. \textsc{CorReg} is not too far from LASSO even if it stays behind. } \label{resYnonlin}
+%\end{table}
+
+We have generated a non-linear structure to test the robustess of the model.
+$\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as previously but with random signs for the components means.
+$\boldsymbol{X}_r=\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2-2\boldsymbol{X}_3+ \varepsilon$.
+The matrix $\boldsymbol{X}$ is then scaled and we get $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\varepsilon_Y$.
+We let $a$ vary between $0$ and $10$ to increase progressively the non-linear part of the sub-regression.
+
+ \begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/robust_S.png} 
+			\caption{Quality of the structure found when the paramater $a$ increases}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/robust_MSE.png} 
+			\caption{MSE on the main regression.}
+   \end{minipage}
+\end{figure}
 

### 2014-05-19 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@608 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust.
-Gaussian case is just a special case ($K_j=1$) of Gaussian mixture so it is included in our hypothesis but identifiability of $S$ requires to have at least one Gaussian mixture with at least two distinct components in each sub-regression (derived from the identifiability of the SR model in \cite{maugis2009variable}, more details in the Appendices \ref{preuveident}).
+Gaussian case is just a special case ($K_j=1$) of Gaussian mixture so it is included in our hypothesis but identifiability of $S$ requires to have Gaussian mixtures with at least two distinct components in each sub-regression (derived from the identifiability of the SR model in \cite{maugis2009variable}, more details in the Appendices \ref{preuveident}).
 
@@
 	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
-	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$ (see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
+	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$.
+%(see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
 	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
@@
 		\subsubsection{Results on $S$}	\label{compZ}
-In table \ref{compZvrai} we compare found structures in different contexts with both Uniform and Hierarchical Uniform {\it a priori} law on $P(S)$, noted respectively $BIC$ and $BIC_+$. 
+%In table \ref{compZvrai} 
+We compare found structures in different contexts with both Uniform and Hierarchical Uniform {\it a priori} law on $P(S)$, noted respectively $BIC$ and $BIC_+$. 
 We see that usage of $BIC_+$ gives sparser models even with $\max (\hat{p}_f^j)=5$.
@@
 The datasets used for $BIC$ and $BIC_+$ are the same to keep the comparison  meaningful.
-These datasets and $\hat{S}$ are those used for tables
- \ref{YXlinOLS} to \ref{YX2linstep}.
+%These datasets and $\hat{S}$ are used for   %tables \ref{YXlinOLS} to \ref{YX2linstep}.
 
@@
 			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
-			\caption{Comparison of the compexities between OLS and CorReg+OLS} 
+			\caption{Comparison of the complexities between OLS and CorReg+OLS} 
    \end{minipage}

### 2014-05-20 pkg/CorReg/R/fillmiss.R
add fillmiss
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@613 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,20 +1,22 @@
-# ' Fill the missing values in the dataset
-# ' 
-# ' @param X the dataset (matrix) with missing values
-# ' @param Z the structure associated to X. Can be a matrix of zeros if non structure.
-# ' @param res_mixmod the best results found by mixmod if already computed
-# ' @param mixmod boolean to say if the function has to use mixmod hypothesis or just the observed mean.
-# ' @param nbclustmax the max number of cluster for mixmod.
-# ' @param X1 booleant to say if dependent variables on the right will be filled based on the structure
-# ' @param Bt the matrix used for X1 if X1=TRUE
+#' Fill the missing values in the dataset
+#' 
+#' @param X the dataset (matrix) with missing values
+#' @param Z the structure associated to X. Can be a matrix of zeros if non structure.
+#' @param res_mixmod the best results found by mixmod if already computed
+#' @param mixmod boolean to say if the function has to use mixmod hypothesis or just the observed mean.
+#' @param nbclustmax the max number of cluster for mixmod.
+#' @param X1 boolean to say if dependent variables on the right will be filled based on the structure
+#' @param Bt the matrix used for X1 if X1=TRUE
+#' @param mclust boolean indicates wether to use Mixmod or mclust
+#' @export
 
-fillmiss<-function(X=X,Z=NULL,mixmod=F,B=NULL,Bt=NULL,res_mixmod=res_mixmod,nbclustmax=10,X1=T,mclust=F){
-   quimank=which(is.na(X),arr.ind=T)
+fillmiss<-function(X=X,Z=NULL,mixmod=FALSE,B=NULL,Bt=NULL,res_mixmod=res_mixmod,nbclustmax=10,X1=TRUE,mclust=FALSE){
+   quimank=which(is.na(X),arr.ind=TRUE)
    p=ncol(X)
    if(is.null(Z)){
      Z=matrix(0,ncol=p,nrow=p)
    }
-   quiou=WhoIs(Z=Z,I1=T,I2=T,I3=T)
+   quiou=WhoIs(Z=Z,I1=TRUE,I2=TRUE,I3=TRUE)
    if(is.null(B)){
      B=hatB(Z=Z,X=X)
    } 

### 2014-05-20 www/article13.tex
up miss
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@614 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{table}
+\clearpage
 
@@
 			\caption{Quality of the subregressions found with our $BIC_+$ criterion} 
+   \end{minipage}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{minipage}[l]{.48\linewidth}
+			\includegraphics[height=180px,width=245px]{figures/res_article/bon_gauche.png} 
+			\caption{Number of $\boldsymbol{X_r}$ covariates found with both $BIC$ (dotted) and $BIC_+$ (plain)}
+	\end{minipage} \
+   \begin{minipage}[r]{.48\linewidth}
+			\includegraphics[height=180px,width=240px]{figures/res_article/faux_gauche.png} 
+			\caption{Number of covariates wrongly set in $\boldsymbol{X}_r$ with both $BIC$ (dotted) and $BIC_+$ (plain)} 
    \end{minipage}

### 2014-05-22 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@620 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-05-19 www/article13.tex
@@
 		\subsubsection{Results on $S$}	\label{compZ}
-%In table \ref{compZvrai} 
-We compare found structures in different contexts with both Uniform and Hierarchical Uniform {\it a priori} law on $P(S)$, noted respectively $BIC$ and $BIC_+$. 
-We see that usage of $BIC_+$ gives sparser models even with $\max (\hat{p}_f^j)=5$.
-Moreover, this constraint is not active because observed complexities are smaller (this constraint only serves to accelerate the walk by reducing the dimension of $\mathcal{S}$ because each configuration was computed a hundred times), meaning that the stronger penalty implied by $BIC_+$ really is efficient.
-The datasets used for $BIC$ and $BIC_+$ are the same to keep the comparison  meaningful.
-%These datasets and $\hat{S}$ are used for   %tables \ref{YXlinOLS} to \ref{YX2linstep}.
-
- It is also notable that $BIC_+$ has a greater computational cost than $BIC$. 
-	We notice that the MCMC is faster when there are numerous correlations (rejecting more candidates). 
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-\hline
-\multicolumn{3}{|c}{Configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
-\hline
-$n$ & $p_r$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_r$ & $\Delta compl$ \\
-\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
-30 & 0 & $BIC$ &0.4104 & 3.2428 & 0 & 5.43 & 0 & -5.43 & 22.55  \\
-& & & (0.0275) & (0.3711) & (0) & (1.9346) & (0) & (1.9346) & (8.0884) \\
- &  & $BIC_+$ &0.4104 & 8.2338 & 0 & 0.53 & 0 & -0.53 & 2.27  \\
-& & & (0.0275) & (0.9045) & (0) & (0.7844) & (0) & (0.7844) & (3.3024) \\
-\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
-30 & 16 & $BIC$ &0.4182 & 2.7735 & 10.96 & 5.93 & 4.98 & -0.95 & 38.16  \\
-& & & (0.0329) & (0.1498) & (1.9844) & (2.0313) & (1.9948) & (0.9987) & (6.499) \\
- &  & $BIC_+$ &0.4182 & 4.1876 & 11.61 & 4.57 & 4.33 & -0.24 & 16.48  \\
-& & & (0.0329) & (0.2178) & (1.8743) & (1.9502) & (1.8752) & (0.4948) & (5.4892) \\
-\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
-30 & 32 & $BIC$ & 0.4456 & 2.9154 & 25.23 & 1.92 & 6.5 & 4.58 & 28  \\
-& & & (0.0429) & (0.1331) & (1.4761) & (1.0888) & (1.4668) & (0.9866) & (5.0831) \\
- &  & $BIC_+$ & 0.4456 & 4.0233 & 16.96 & 3.04 & 14.77 & 11.73 & 4.35  \\
-& & & (0.0429) & (0.1091) & (1.3993) & (1.3993) & (1.4692) & (0.5478) & (5.8833) \\
-\hline
-\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
-50 & 0 & $BIC$ &0.5229 & 4.7068 & 0 & 4.2 & 0 & -4.2 & 13.35  \\
-& & & (0.0519) & (0.5865) & (0) & (1.7233) & (0) & (1.7233) & (5.6468) \\
- &  & $BIC_+$ &0.5229 & 10.1198 & 0 & 0.13 & 0 & -0.13 & 0.32  \\
-& & & (0.0519) & (0.5541) & (0) & (0.3667) & (0) & (0.3667) & (0.9732) \\
-\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
-50 & 16 & $BIC$ &0.5205 & 3.3681 & 11.15 & 5.42 & 4.72 & -0.7 & 22.85  \\	
-& & & (0.0451) & (0.3123) & (1.93) & (1.9132) & (1.886) & (0.7317) & (5.7742) \\
- &  & $BIC_+$ &0.5205 & 4.909 & 11.42 & 4.59 & 4.45 & -0.14 & 7.55  \\
-& & & (0.0451) & (0.4556) & (1.9079) & (1.8968) & (1.8333) & (0.3487) & (4.0611) \\
-\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
-50 & 32 & $BIC$ & 0.5833 & 3.2683 & 28.17 & 1.38 & 3.7 & 2.32 & 12.74  \\
-& & & (0.0628) & (0.316) & (1.3711) & (0.9077) & (1.3143) & (0.8394) & (4.3359) \\
- &  & $BIC_+$ &0.5833 & 4.3599 & 17.27 & 2.73 & 14.6 & 11.87 & -2.61  \\
-& & & (0.0628) & (0.3729) & (1.1708) & (1.1708) & (1.2792) & (0.338) & (4.4854) \\
-\hline
-\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
-100 & 0 & $BIC$ & 0.9623 & 12.9373 & 0 & 2.83 & 0 & -2.83 & 6.23  \\
-& & & (0.077) & (1.7778) & (0) & (1.2953) & (0) & (1.2953) & (3.1999) \\
- &  & $BIC_+$ & 0.9623 & 20.9817 & 0 & 0.01 & 0 & -0.01 & 0.02  \\
-& & & (0.077) & (1.9421) & (0) & (0.1) & (0) & (0.1) & (0.2) \\
-\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
-100 & 16 & $BIC$ & 1.1223 & 6.9647 & 11.67 & 4.8 & 4.25 & -0.55 & 12.58  \\
-& & & (0.1122) & (0.5473) & (2.0003) & (2.0646) & (1.956) & (0.7833) & (3.9471) \\
- &  & $BIC_+$ & 1.1223 & 8.8486 & 12.04 & 3.95 & 3.88 & -0.07 & 3.75  \\
-& & & (0.1122) & (0.7174) & (1.9223) & (1.9404) & (1.9137) & (0.2564) & (2.2625) \\
-\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
-100 & 32 & $BIC$ & 1.4343 & 5.9626 & 30.14 & 0.84 & 1.61 & 0.77 & 6.96  \\
-& & & (0.2528) & (0.3136) & (1.3928) & (0.8495) & (1.2941) & (0.7086) & (3.0975) \\
- &  & $BIC_+$ & 1.4343 & 7.3741 & 17.49 & 2.51 & 14.26 & 11.75 & -3.76  \\
-& & & (0.2528) & (0.2748) & (1.1849) & (1.1849) & (1.2441) & (0.4794) & (4.4859) \\
-\hline
-\end{tabular} 
-\caption{Results of the Markov chain: Mean observed and standard deviation (sd). } \label{compZvrai}
-\end{table}
 
@@
 \clearpage
+\subsection{Results on prediction}
+
 \subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_f}$ (best case for us)}	 \label{tableMSEsimgauche}
@@
 \clearpage
-\subsection{Results on prediction}
 	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 	

### 2014-05-26 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@621 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{lmodern}
+\usepackage{subfigure}
 \usepackage[table]{xcolor}
@@
 	\smallskip
-	At every moment, coherence between $I_f$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_f^j=|I_f^j|$, $I_r=\{j |p_f^j>0 \}$, $p_r= |I_r|$ .
+	At every moment, coherence between $I_f$ and others parts of $S$ can be done by $\forall 1\leq j\leq p :  p_f^j=|I_f^j|$, $I_r=\{j |p_f^j>0 \}$ and $p_r= |I_r|$ .
 
@@
 and the output will be the best structure in terms of $P(S|\boldsymbol{X})$ which weights each candidate. Practically speaking, \textsc{CorReg} returns the best structure seen during the walk.
-Numerical results (Section 4) illustrates the efficiency of the walk when the true model really contains a linear structure or no structure at all (Table (\ref{compZvrai})) and when the structure is not linear (Table \ref{compZnonlin})).
+Numerical results (Section 4) illustrates the efficiency of the walk when the true model contains structures with various strength (Figure \ref{reshatZ}) and an example with a non-linear structure (Figure \ref{resnonlin}).
 
@@
 Section \ref{compZ} shows the results obtained in terms of $\hat{S}$.
-Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods.
-Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals.
-For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$, variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$.
+Sections \ref{reshatY} and \ref{resrobust} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods respectively for general case and extreme cases.
+Figures gives both mean and inter-quartile intervals of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals.
+For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$ and those of the sub-regressions are successively set to $(0.1,0.3,0.5,0.7,0.99)$.
+Variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$.
 The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign.
@@
 This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
-	We compare different methods with and without CorReg as a pretreatment. All the results are provided by the CorReg package.
+	We compare different methods with and without CorReg as a pretreatment.
+All the results are provided by the CorReg package.
+In the following figures, the thickness represent various values of $n$, the thicker the greater, in $(30,50,100,400)$ so we start from $n<p$.
 		\subsection{Finding the structure}
@@
 But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning.
-To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
-			Global indicators :
+To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found $\hat{S}$:
 			\begin{itemize}
@@
 				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_r|-TL$
-				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple
-				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
+				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple.
+				%\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
 			\end{itemize}
+
 		\subsubsection{Results on $S$}	\label{compZ}
-
-\begin{figure}[h!]
+Here we just take a look at $TL$ and $WL$, that fully describe the quality of the variable selection that will be made in the marginal model.
+$ML$ and $\Delta p_r$ can be deduced knowing the value of $p_r$ (16 in these cases).
+\begin{figure}[h!] \label{reshatZ}
 	\begin{minipage}[l]{.48\linewidth}
@@
 \end{figure}
+We can clearly observe that $BIC_+$ find less sub-regressions when $R^2$ is small and give less wrong sub-regressions than $BIC$ for every $R^2$ but when the strength of the sub-regression increases, $BIC_+$ find as much good sub-regressions as the $BIC$ so the objective is reached.
+We found the sub-regression only when they are significant and we reduce the noise.
 
 \clearpage
-\subsection{Results on prediction}
-
-\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_f}$ (best case for us)}	 \label{tableMSEsimgauche}
-\begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X1_MSE_NB.png} 
-			\caption{Comparison of the MSE between OLS and CorReg+OLS}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
-			\caption{Comparison of the complexities between OLS and CorReg+OLS} 
-   \end{minipage}
-\end{figure}
-
+\subsection{Results on prediction}\label{reshatY}
+
+	%\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 	
+We first try the method with a response variable $Y$ depending on all covariates (\textsc{CorReg} reduces the dimension and thus can't give the true model if there is a structure).
+The datasets used here were the same as for Figure \ref{reshatZ}.
+
+\begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_tout_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_tout_compl_NB.png} 
+	}
+	\caption{Comparison of OLS (dotted) and CorReg + OLS (plain) when $Y$ depends on all covariates. Thickness represents $n=30,50,100,400$.}\label{toutOLS}
+\end{figure}
+
+ Figure \ref{toutOLS} (left) shows that CorReg improves the prediction power of OLS in a significant way for small values of $n$ and/or significant linear structures.
+This advantage then shrinks when $n$ explodes because the matrix to invert become well-conditioned and the true model does depend on all covariates (not allowed in the marginal model).
+We also see that (Figure \ref{toutOLS} right) the model provided by CorReg is more parsimonious and that for small values of $n$ and weak sub-regressions, the number of found sub-regressions is reduced so the marginal model tends to keep more covariates.
 %\caption{OLS and OLS combined with constrained \textsc{CorReg}.
@@
 \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X1_MSE_NB.png} 
-			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X1_compl_NB.png} 
-			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} 
-   \end{minipage}
-\end{figure}
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_X1_MSE_NB.png} 
-			\caption{Comparison of the MSE between elasticnet and CorReg+elasticnet}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_X1_compl_NB.png} 
-			\caption{Comparison of the compexities between elasticnet and CorReg+elasticnet} 
-   \end{minipage}
-\end{figure}
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_X1_MSE_NB.png} 
-			\caption{Comparison of the MSE between stepwise and CorReg+stepwise}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_X1_compl_NB.png} 
-			\caption{Comparison of the compexities between stepwise and CorReg+stepwise} 
-   \end{minipage}
-\end{figure}
-
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/lar_tout_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/lar_tout_compl_NB.png} 
+	}
+	\caption{Comparison of LASSO (dotted) and CorReg + LASSO (plain) when $Y$ depends on all covariates. Thickness represents $n=30,50,100,400$.}\label{toutlar}
+\end{figure}
+
+\begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_tout_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_tout_compl_NB.png} 
+	}
+	\caption{Comparison of elasticnet (dotted) and CorReg + elasticnet (plain) when $Y$ depends on all covariates. Thickness represents $n=30,50,100,400$.}\label{toutelast}
+\end{figure}
+
+\begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_tout_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_tout_compl_NB.png} 
+	}
+	\caption{Comparison of stepwise (dotted) and CorReg + stepwise (plain) when $Y$ depends on all covariates. Thickness represents $n=30,50,100,400$.}\label{toutstepwise}
+\end{figure}
+Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} show that CorReg used as a pretreatment improves the results obtained by variable selection methods and tends to give a more parsimonious model.
+This is remarkable because the true model depends on all covariates, so it shows that other variable selection methods are really penalized by the correlations between the covariates.
+
+ \begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/ridge_tout_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/ridge_tout_compl_NB.png} 
+	}
+	\caption{Comparison of ridge (dotted) and CorReg + ridge (plain) when $Y$ depends on all covariates. Thickness represents $n=30,50,100,400$.}\label{toutridge}
+\end{figure}
+Figure \ref{toutridge} shows that the ridge regression is efficient in prediction when confronted to correlated covariates.
+It is directly made to improve the conditioning and keeps all the covariates (as the true model) so it logically gives better predictions than CorReg which delete some covariates.
+But we notice that CorReg pretreatment before the ridge regression gives MSE that are really near form those from the ridge regression alone (inter-quartile intervals are essentially mingled) and with a parsimonious model.
+Ridge regression alone gives a full model (40 covariates even with only 30 individuals) and will give models too complex to be interpreted.
+A model with 40 covariates is already difficult to interpret and it get worse when $p$ rises.
+So even the ridge regression in its most favorable case (True model depending on all the covariates) takes benefits (for interpretation) form the CorReg pretreatment.
 \clearpage
-	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 	
-We then try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure).
-The datasets used here were those from table \ref{compZvrai}.
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_tout_MSE_NB.png} 
-			\caption{Comparison of the MSE between OLS and CorReg+OLS}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_tout_compl_NB.png} 
-			\caption{Comparison of the compexities between OLS and CorReg+OLS} 
-   \end{minipage}
-\end{figure}
-
+
+\subsection{Extreme cases and Robustess of the model} \label{resrobust}
+\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_f}$ (best case for us)}	 \label{tableMSEsimdroit}
+The datasets used here were the same as for Figure \ref{reshatZ}.
+
+ \begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X1_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
+	}
+	\caption{Comparison of OLS (dotted) and CorReg + OLS (plain) when $Y$ depends on $X_f$ only. Thickness represents $n=30,50,100,400$.}\label{X1OLS}
+\end{figure}
 %\caption{OLS and OLS combined with constrained \textsc{CorReg}.
@@
 \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/lar_tout_MSE_NB.png} 
-			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/lar_tout_compl_NB.png} 
-			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} 
-   \end{minipage}
-\end{figure}
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_tout_MSE_NB.png} 
-			\caption{Comparison of the MSE between elasticnet and CorReg+elasticnet}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_tout_compl_NB.png} 
-			\caption{Comparison of the compexities between elasticnet and CorReg+elasticnet} 
-   \end{minipage}
-\end{figure}
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_tout_MSE_NB.png} 
-			\caption{Comparison of the MSE between stepwise and CorReg+stepwise}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_tout_compl_NB.png} 
-			\caption{Comparison of the compexities between stepwise and CorReg+stepwise} 
-   \end{minipage}
-\end{figure}
-
-We see that CorReg tends to give more parsimonious models and better predictions, even if the true model is not parsomious.
-We logically observe that when $n$ rises, all the models get better and the correlations cease to be a problem so the complete model starts to be better (CorReg does not allow the true model to be choosen).
-
-\clearpage
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X1_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X1_compl_NB.png} 
+	}
+	\caption{Comparison of LASSO (dotted) and CorReg + LASSO (plain) when $Y$ depends on $X_f$ only. Thickness represents $n=30,50,100,400$.}\label{X1lasso}
+\end{figure}
+Figures \ref{X1OLS} and \ref{X1lasso} lead to the same comparison.
+Even if the true model needs variable selection and depends only on independent covariates, if the dataset contained correlated but irrelevant covariates then classical variable selection methods gain to be associated with a pretreatment by CorReg.
+
+\begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
+			\includegraphics[height=180px,width=245px]{figures/res_article/ridge_X1_MSE_NB.png} 
+	} \quad
+   	\subfigure[Complexity]{
+			\includegraphics[height=180px,width=240px]{figures/res_article/ridge_X1_compl_NB.png} 
+	}
+	\caption{Comparison of ridge (dotted) and CorReg + ridge (plain) when $Y$ depends on $X_f$ only. Thickness represents $n=30,50,100,400$.}\label{X1ridge}
+\end{figure}
+Figure \ref{X1ridge} shows that ridge regression is penalized when the true model is parsimonious and that CorReg can help to improve the results in such a case.
+
 	\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_r}$ (worst case for us)}	 \label{tableMSEsimgauche}
-We now try the method with a response depending only on variables in $\boldsymbol{X}_r$. The datasets used here were still those from \ref{compZvrai}.
+We now try the method with a response depending only on variables in $\boldsymbol{X}_r$. The datasets used here were still the same as for Figure \ref{reshatZ}.
 Depending only on $\boldsymbol{X}_r$ implies sparsity and impossibility to obtain the true model when using the true structure.
@@
 			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X2_compl_NB.png} 
-			\caption{Comparison of the compexities between OLS and CorReg+OLS} 
+			\caption{Comparison of the complexities between OLS and CorReg+OLS} 
    \end{minipage}
@@
 			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X2_compl_NB.png} 
-			\caption{Comparison of the compexities between LASSO and CorReg+LASSO} 
+			\caption{Comparison of the complexities between LASSO and CorReg+LASSO} 
    \end{minipage}
@@
 
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_X2_MSE_NB.png} 
-			\caption{Comparison of the MSE between elasticnet and CorReg+elasticnet}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_X2_compl_NB.png} 
-			\caption{Comparison of the compexities between elasticnet and CorReg+elasticnet} 
-   \end{minipage}
-\end{figure}
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_X2_MSE_NB.png} 
-			\caption{Comparison of the MSE between stepwise and CorReg+stepwise}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_X2_compl_NB.png} 
-			\caption{Comparison of the compexities between stepwise and CorReg+stepwise}
-   \end{minipage}
-\end{figure}
-
 \clearpage
-\subsection{Robustess of the model}
-% We have generated non linear structures (Tables \ref{compZnonlin} and \ref{resYnonlin}).
-Variables in $\boldsymbol{X_r}$ depends on the $\log$ or the square (randomly choosen with equiprobability) of a variable in $\boldsymbol{X_f}$ .
-Dependencies are still real but non linear.
-\textsc{CorReg} still found dependencies.
-One example of non-linear structure found with \textsc{CorReg} on real datasets is given in section \ref{sectionexfos}.
-%\begin{table}[h!]
-%\centering
-%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-%\hline
-%\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
-%\hline
-%$n$ & $p_r$ & $P(S)$&Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\
-%\hline
-%30 & 16 & $BIC$ & 0.4313 & 2.7769 & 9.96 & 2.45 & 5.95 & 3.5 & 29.42  \\
-%& & &(0.0391) & (0.0883) & (1.3175) & (1.3808) & (1.3056) & (1.453) & (6.2639) \\
-% &  & $BIC_+$ & 0.4313 & 4.9079 & 7.28 & 1.25 & 8.63 & 7.38 & 5.77  \\ 
-%& & &(0.0391) & (0.4353) & (1.5769) & (0.9987) & (1.5548) & (1.6316) & (6.1297) \\ 
-%\hline
-%\end{tabular} 
-%\caption{Results of the Markov chain for non linear structure ($\log$ and square). Mean observed and standard deviation (sd). } \label{compZnonlin}
-%\end{table}
-%
-%
-%\begin{table}[h!]
-%\centering
-%\begin{tabular}{|c|c|c|c|c|c|c|}
-%\hline 
-%$n$ & $p_r$& $P(S)$ &indicator &LASSO  &    \textsc{CorReg} $\hat S$& \textsc{CorReg} $S$\\ 
-%\hline %nonlin0.5global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_55_33.csv
-%30 & 16 & $BIC$ &MSE (sd) & 996.06 (323.7) & 1100.17 (391.4) & 1501.25 (905.9) \\
-%& & &cpl (sd) & 17.66 (5.4) & 14.7 (4.7) & 11.97 (4.9) \\
-% &  &$BIC_+$ &MSE (sd) & 996.06 (323.7) & 1094.5 (621.8) & 1501.25 (905.9) \\
-%& & & cpl (sd) & 17.66 (5.4) & 16.98 (4.8) & 11.97 (4.9) \\ 
-%\hline
-%\end{tabular} 
-%\caption{ $\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$. \textsc{CorReg} is not too far from LASSO even if it stays behind. } \label{resYnonlin}
-%\end{table}
+\subsubsection{non-linear case}
 
@@
 
- \begin{figure}[h!]
+ \begin{figure}[h!] \label{resnonlin}
 	\begin{minipage}[l]{.48\linewidth}
@@
 Elasticnet with \textsc{CorReg} is $13.51\%$ better than OLS.
-		\begin{figure}[h]
-			\centering
-				\label{barplotMSEexfos}
-				\includegraphics[width=430px]{figures/MSEfinal.png}
-			\caption{MSE comparison on industrial dataset. Learning set : 3 000 products, validation set : 847 products}
-		\end{figure}		
+%		\begin{figure}[h]
+%			\centering
+%				\label{barplotMSEexfos}
+%				\includegraphics[width=430px]{figures/MSEfinal.png}
+%			\caption{MSE comparison on industrial dataset. Learning set : 3 000 products, validation set : 847 products}
+%		\end{figure}		
 		\begin{table}[h!]
@@
 	\hline 
-	OLS & 115.63 & 206 \\ 
-	\hline 
-	\textsc{CorReg} + OLS & 109.59&130 \\ 
-	\hline  
-	LASSO & 106.84 & 21 \\ 
-	\hline 
-	\textsc{CorReg} + LASSO & 102.45 &24 \\ 
-	\hline 
-	elasticnet & 110.81 & 78\\ 
-	\hline 
-	\textsc{CorReg} + elasticnet & 100 &40 \\ 
+	OLS & 14.03 & 206 \\ 
+	\hline 
+	\textsc{CorReg} + OLS & 13.30 & 130 \\ 
+	\hline   
+	LASSO & 12.96 & 21 \\ 
+	\hline 
+	\textsc{CorReg} + LASSO & 12.77 & 24 \\ 
+	\hline 
+	elasticnet & 13.52 & 78\\ 
+	\hline 
+	\textsc{CorReg} + elasticnet & \textbf{12.15} & 40 \\ 
+	\hline 
+	ridge & 13.09 & 206\\ 
+	\hline 
+	\textsc{CorReg} + ridge & 12.69 & 130 \\ 
 	\hline 

### 2014-05-26 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@622 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{subfigure}
+\usepackage{placeins}
 \usepackage[table]{xcolor}
@@
 The model depends only on independent covariates because it is a marginal model on these covariates.
- This can be viewed as a pretreatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
-This pretreatment is specifically done to decorrelate the covariates, unlike methods described above.
+ This can be viewed as a pre-treatment on the dataset that will be followed then by any other tools for estimation and dimension reduction without suffering from correlations.
+This pre-treatment is specifically done to decorrelate the covariates, unlike methods described above.
 The linear structure is obtained by a MCMC algorithm optimizing the penalized likelihood of the joint distribution on the covariates, independently from the response variable.
@@
 Nevertheless, to be able to compare the bias-variance tradeoff, we can see this model as a variable pre-selection independent of the response in $\boldsymbol{Y}_{|\boldsymbol{X}}$.
-We note that it is simply a linear regression on some of the original covariates so we only made a pretreatment on the dataset by selecting $\boldsymbol{X}_f$ because of the correlations given by $S$.
+We note that it is simply a linear regression on some of the original covariates so we only made a pre-treatment on the dataset by selecting $\boldsymbol{X}_f$ because of the correlations given by $S$.
 So we also get the model
@@
 \paragraph{Running example:} $\boldsymbol{Y}_{|\boldsymbol{X}_f}= 2\boldsymbol{x}_1+2\boldsymbol{x}_2+\boldsymbol{x}_4+\boldsymbol{x}_5+\boldsymbol{\varepsilon}_3 +\boldsymbol{\varepsilon}_Y$
-\subsection{Strategy of use: pretreatment before classical estimation/selection methods}\label{interpretation}
-
-As a pretreatment, the model allows usage of any method in a second time to estimate $\boldsymbol{\beta}_f^*$, even with variable selection methods like LASSO or a best subset algorithm like stepwise \cite{seber2012linear}.
+\subsection{Strategy of use: pre-treatment before classical estimation/selection methods}\label{interpretation}
+
+As a pre-treatment, the model allows usage of any method in a second time to estimate $\boldsymbol{\beta}_f^*$, even with variable selection methods like LASSO or a best subset algorithm like stepwise \cite{seber2012linear}.
 However, we always have $\boldsymbol{X}_r=\boldsymbol{0}$
@@
 
-	\subsection{Illustration of the tradeoff conveyed by the pretreatment}	
-	We compare the OLS estimator on $\boldsymbol{X}$ defined in section \ref{sectionOLS} with the estimator obtained by the pretreatment that is $\boldsymbol{X}_f$ selection.
+	\subsection{Illustration of the tradeoff conveyed by the pre-treatment}	
+	We compare the OLS estimator on $\boldsymbol{X}$ defined in section \ref{sectionOLS} with the estimator obtained by the pre-treatment that is $\boldsymbol{X}_f$ selection.
 
@@
 This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
-	We compare different methods with and without CorReg as a pretreatment.
+	We compare different methods with and without CorReg as a pre-treatment.
 All the results are provided by the CorReg package.
@@
 $ML$ and $\Delta p_r$ can be deduced knowing the value of $p_r$ (16 in these cases).
-\begin{figure}[h!] \label{reshatZ}
-	\begin{minipage}[l]{.48\linewidth}
+\begin{figure}[h!]
+	\subfigure[with $BIC$ criterion]{
 			\includegraphics[height=180px,width=245px]{figures/res_article/BIC_p2.png} 
-			\caption{Quality of the subregressions found with classical $BIC$ criterion}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
+	} \quad
+   	\subfigure[with $BIC_+$ criterion]{
 			\includegraphics[height=180px,width=240px]{figures/res_article/BICSTAR_P2.png} 
-			\caption{Quality of the subregressions found with our $BIC_+$ criterion} 
-   \end{minipage}
-\end{figure}
+	}
+	\caption{Quality of the found subregressions (dotted = True Left, plain = Wrong Left). Thickness represents $n=30,50,100,400$.}\label{reshatZ}
+\end{figure}
+
 We can clearly observe that $BIC_+$ find less sub-regressions when $R^2$ is small and give less wrong sub-regressions than $BIC$ for every $R^2$ but when the strength of the sub-regression increases, $BIC_+$ find as much good sub-regressions as the $BIC$ so the objective is reached.
@@
 
-\clearpage
+\FloatBarrier
 \subsection{Results on prediction}\label{reshatY}
@@
 
- Figure \ref{toutOLS} (left) shows that CorReg improves the prediction power of OLS in a significant way for small values of $n$ and/or significant linear structures.
+ Figure \ref{toutOLS} (a) shows that CorReg improves the prediction power of OLS in a significant way for small values of $n$ and/or significant linear structures.
 This advantage then shrinks when $n$ explodes because the matrix to invert become well-conditioned and the true model does depend on all covariates (not allowed in the marginal model).
-We also see that (Figure \ref{toutOLS} right) the model provided by CorReg is more parsimonious and that for small values of $n$ and weak sub-regressions, the number of found sub-regressions is reduced so the marginal model tends to keep more covariates.
+We also see that (Figure \ref{toutOLS} (b)) the model provided by CorReg is more parsimonious and that for small values of $n$ and weak sub-regressions, the number of found sub-regressions is reduced so the marginal model tends to keep more covariates.
 %\caption{OLS and OLS combined with constrained \textsc{CorReg}.
@@
 \end{figure}
-Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} show that CorReg used as a pretreatment improves the results obtained by variable selection methods and tends to give a more parsimonious model.
+Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} show that CorReg used as a pre-treatment improves the results obtained by variable selection methods and tends to give a more parsimonious model.
 This is remarkable because the true model depends on all covariates, so it shows that other variable selection methods are really penalized by the correlations between the covariates.
@@
 It is directly made to improve the conditioning and keeps all the covariates (as the true model) so it logically gives better predictions than CorReg which delete some covariates.
-But we notice that CorReg pretreatment before the ridge regression gives MSE that are really near form those from the ridge regression alone (inter-quartile intervals are essentially mingled) and with a parsimonious model.
+But we notice that CorReg pre-treatment before the ridge regression gives MSE that are really near form those from the ridge regression alone (inter-quartile intervals are essentially mingled) and with a parsimonious model.
 Ridge regression alone gives a full model (40 covariates even with only 30 individuals) and will give models too complex to be interpreted.
 A model with 40 covariates is already difficult to interpret and it get worse when $p$ rises.
-So even the ridge regression in its most favorable case (True model depending on all the covariates) takes benefits (for interpretation) form the CorReg pretreatment.
-\clearpage
+So even the ridge regression in its most favorable case (True model depending on all the covariates) takes benefits (for interpretation) form the CorReg pre-treatment.
+\FloatBarrier
 
@@
 Figures \ref{X1OLS} and \ref{X1lasso} lead to the same comparison.
-Even if the true model needs variable selection and depends only on independent covariates, if the dataset contained correlated but irrelevant covariates then classical variable selection methods gain to be associated with a pretreatment by CorReg.
+Even if the true model needs variable selection and depends only on independent covariates, if the dataset contained correlated but irrelevant covariates then classical variable selection methods gain to be associated with a pre-treatment by CorReg.
 
@@
 
+\FloatBarrier
 	\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_r}$ (worst case for us)}	 \label{tableMSEsimgauche}
@@
 Depending only on $\boldsymbol{X}_r$ implies sparsity and impossibility to obtain the true model when using the true structure. 
-
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
+\begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
 			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X2_MSE_NB.png} 
-			\caption{Comparison of the MSE between OLS and CorReg+OLS}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
+	} \quad
+   	\subfigure[Complexity]{
 			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X2_compl_NB.png} 
-			\caption{Comparison of the complexities between OLS and CorReg+OLS} 
-   \end{minipage}
-\end{figure}
-\textsc{CorReg} is still better than OLS for strong correlations and limited values of $n$. 
- \begin{figure}[h!]
-	\begin{minipage}[l]{.48\linewidth}
+	}
+	\caption{Comparison of OLS (dotted) and CorReg + OLS (plain) when $Y$ depends on $X_r$ only. Thickness represents $n=30,50,100,400$.}\label{X2OLS}
+\end{figure}
+
+Figure \ref{X2OLS} reveals that \textsc{CorReg} is still better than OLS for strong correlations and limited values of $n$.
+When $n$ rises, we found the sub-regression and then delete relevant covariates so CorReg does not improve the results and increases the MSE.
+but for strong correlations, the error shrinks as the model tends to be less identifiable and switching variables is not a problem anymore even with large values of $n$.
+
+\begin{figure}[h!]
+	\subfigure[MSE on a validation subset]{
 			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X2_MSE_NB.png} 
-			\caption{Comparison of the MSE between LASSO and CorReg+LASSO}
-	\end{minipage} \
-   \begin{minipage}[r]{.48\linewidth}
+	} \quad
+   	\subfigure[Complexity]{
 			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X2_compl_NB.png} 
-			\caption{Comparison of the complexities between LASSO and CorReg+LASSO} 
-   \end{minipage}
-\end{figure}
-
-\clearpage
+	}
+	\caption{Comparison of LASSO (dotted) and CorReg + LASSO (plain) when $Y$ depends on $X_r$ only. Thickness represents $n=30,50,100,400$.}\label{X2LASSO}
+\end{figure}
+Figure \ref{X2LASSO} shows that LASSO naturally tends to keep $\boldsymbol{X_r}$ and thus is better in such a true model.
+So CorReg is not always the better solution and it is recommended to compare classical and marginal model.
+This model choice is only between two models so it avoids multiple tests issues and computational cost explosion.
+We suggest to always compute the complete model and the marginal one and then to compare them with the more pertinent criterion (AIC, cross-validation, validation sample, {\it etc.
+}) according to the context (size of the datasets for example).
+
+It is notable that the structure does not depend on $\boldsymbol{Y}$ so it can be used for several studies and has an intrinsic value for interpretation even when the marginal model is not kept after confrontation with the classical one.
+CorReg is both a pre-study and a pre-treatment
+
+\FloatBarrier
 \subsubsection{non-linear case}
@@
 
- \begin{figure}[h!] \label{resnonlin}
+ \begin{figure}[h!] 
 	\begin{minipage}[l]{.48\linewidth}
 			\includegraphics[height=180px,width=245px]{figures/res_article/robust_S.png} 
-			\caption{Quality of the structure found when the paramater $a$ increases}
+			\caption{Frequency of true $\boldsymbol{X}_r$ found (plain)   when the paramater $a$ increases}\label{resnonlin}
 	\end{minipage} \
    \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/robust_MSE.png} 
-			\caption{MSE on the main regression.}
+			\includegraphics[height=180px,width=240px]{figures/res_article/robust_MSE.png} \label{MSEnonlin}
+			\caption{MSE on the main regression for OLS(thick) and LASSO (thin) used both with (plain) or without CorReg (dotted).}
    \end{minipage}
 \end{figure}
-
-	\clearpage	
+Figure \ref{MSEnonlin} illustrates the advantage of using CorReg even with non-linear structures.
+Figure \ref{resnonlin} shows that the MCMC have more difficulties to find a linear structure as the non-linear part of the sub-regression increases but the model is quite robust (efficient for small values of $a$)
+
+	\FloatBarrier	
 \section{Numerical results on real datasets} \label{sectionrealcase}
@@
 			\item 205 variables from the whole process to explain it.
-			\item The stakes : a hundred euros per ton (for information: Dunkerque's  site aims to produce up to 7.5 millions tons a year)
 		\end{itemize}
@@
 \begin{figure}[h!]
-	\begin{minipage}[l]{.30\linewidth}
-			\includegraphics[height=150px,width=150px]{figures/mixmod.png} 
-			\caption{Example of non-Gaussian real variable easily modeled by a Gaussian mixture}\label{graphMixmod}
+	\begin{minipage}[t]{.30\linewidth}
+			\includegraphics[height=150px,width=150px]{figures/res_article/nb_comp_X_exfo.png}%{figures/mixmod.png} 
+			\caption{Distribution of the number of components found for each covariate.}\label{graphMixmod}
 	\end{minipage} \hfill
-	\begin{minipage}[c]{.30\linewidth}
-			\includegraphics[height=150px, width=150px]{figures/histR2exfos.png} 
-			\caption{$R^2_{adj}$ of the 76 sub-regressions.}
+	\begin{minipage}[t]{.30\linewidth}
+			\includegraphics[height=150px, width=150px]{figures/res_article/gaussianmixture_exfo.png}%{figures/histR2exfos.png} $R^2_{adj}$ of the 76 sub-regressions.
+			\caption{Example of non-Gaussian real variable easily modeled by a Gaussian mixture.}
 	\end{minipage} \hfill
-   \begin{minipage}[r]{.30\linewidth}
+   \begin{minipage}[t]{.30\linewidth}
 			\includegraphics[height=150px,width=150px]{figures/correlexfoshist.png} 
@@
 Industrial variables are naturally highly correlated as the width and the weight of a steel slab ($\rho=0.905$), the temperature before and after some tool ($\rho=0.983$), the  roughness of both faces of the product ($\rho= 0.919$), a mean and a max ($\rho=0.911$).
-\textsc{CorReg} also found more complex structures describing physical models, like   Width = f (Mean.flow , Mean.speed.CC) even if the true Physcial model is not linear : Width = flow / (speed * thickness) (here thickness is constant).
-Non linear regulation models used to optimize the process were also found (but are confidential).
+
+	The objective here is not only to predict non-quality but to understand and then avoid it.
+CorReg provides an automatic method without any a priori and is combined with variable selection methods.
+So it allows to obtain in a small amount of time some indications on the source of the problem, and to use human resources efficiently.
+When quality crises occurs, time is extremely precious so automation is a real stake.
+
+		\begin{figure}[h!]
+		\centering
+			\includegraphics[height=150px, width=150px]{figures/histR2exfos.png} 
+			\caption{$R^2_{adj}$ of the 76 sub-regressions.}
+\end{figure}
+
+	\textsc{CorReg} found the above correlations but it also found more complex structures describing physical models, like   Width = f (Mean flow , Mean speed) even if the true Physcial model is not linear : Width = flow / (speed * thickness) (here thickness is constant).
+Non-linear regulation models used to optimize the process were also found (but are confidential).
 These first results are easily understandable and meet metallurgists expertise.
+Sub-regressions with small values of $R^2$ are associated with non-linear model (chemical kinetics for example).
 			The algorithm gives a structure of $p_r=76$ subregressions with a mean of $\bar{\boldsymbol{p}}_f=5.17$ regressors.
 In $\boldsymbol{X}_f$ the number of $\rho>0.7$ is $\textbf{79.33\%}$ smaller than in $\boldsymbol{X}$.
-
-			It is now time to look at the predictive results (Figure \ref{compareMSEexfos}).
-				The best model found when not using \textsc{CorReg} is given by the LASSO.
+	\\
+
+			It is now time to look at the predictive results (Table \ref{Res_exfos}). We see that CorReg improves the results for each method tested in terms of prediction.
+			We get parsimonious models automatically in a small amount of time (several hours but able to work during the night or the week-end)
+%				The best model found when not using \textsc{CorReg} is given by the LASSO.
 But when using \textsc{CorReg} elasticnet produces a better model in terms of prediction.
@@
 $14$ non-zero coefficients are common between the two models.
-				Elasticnet alone get a model with 78 parameters that is improved by $9.75\%$ in prediction when used with \textsc{CorReg}.
+%				Elasticnet alone get a model with 78 parameters that is improved by $9.75\%$ in prediction when used with \textsc{CorReg}.
 When using LASSO with \textsc{CorReg} we obtain a model with 24 non-zero coefficients that is $4.11\%$ better than LASSO alone.
@@
 Elasticnet with \textsc{CorReg} is $13.51\%$ better than OLS.
-%		\begin{figure}[h]
+%%		\begin{figure}[h]
 %			\centering
@@
 %		\end{figure}
+
 		\begin{table}[h!]
 \centering
-\begin{tabular}{|c|c|c|}
-	\hline 
-	Model & MSE & Complexity (with intercept) \\ 
-	\hline 
-	OLS & 14.03 & 206 \\ 
-	\hline 
-	\textsc{CorReg} + OLS & 13.30 & 130 \\ 
-	\hline   
-	LASSO & 12.96 & 21 \\ 
-	\hline 
-	\textsc{CorReg} + LASSO & 12.77 & 24 \\ 
-	\hline 
-	elasticnet & 13.52 & 78\\ 
-	\hline 
-	\textsc{CorReg} + elasticnet & \textbf{12.15} & 40 \\ 
-	\hline 
-	ridge & 13.09 & 206\\ 
-	\hline 
-	\textsc{CorReg} + ridge & 12.69 & 130 \\ 
+\begin{tabular}{|c c|c|c|}
+	\hline 
+	Method& indicator& With CorReg & without CorReg \\ 
+	\hline 
+	OLS & MSE & 13.30 & 14.03 \\
+		& (complexity)& (130) & (206) \\
+	\hline
+	LASSO & MSE & 12.77 & 12.96 \\
+		& (complexity)& (24) & (21) \\
+	\hline
+	Elasticnet & MSE & \textbf{12.15} & 13.52 \\
+		& (complexity)& (40) & (78) \\
+	\hline
+	Ridge & MSE & 12.69 & 13.09 \\
+		& (complexity)& (130) & (206) \\
 	\hline
 \end{tabular} 
-\caption{Results obtained on a validation sample.}	
+\caption{Results obtained on a validation sample (847 individuals).}\label{Res_exfos}
 \end{table}
@@
 		In terms of interpretation, the main regression comes with the family of regression so it gives a better understanding of the consequences of corrective actions on the whole process.
-It typically permits to determine the \textit{tuning parameters} whereas LASSO would point variables we can't directly act on.	So it becomes easier to take corrective actions on the process to reach the goal.
-The stakes are so important that even a little improvement leads to consequent benefits, and we don't even talk of the impact on the market shares that is even more important.
-		\clearpage
-
+It typically permits to determine the \textit{tuning parameters} whereas variable selection alone would point variables we can't directly act on.	So it becomes easier to take corrective actions on the process to reach the goal.
+The stakes are so important that even a little improvement leads to consequent benefits, and we don't even talk about the impact on the market shares that is even more important.
+		\FloatBarrier
 		\subsection{Production case study}
-This second example is about a phenomenon that impacts the productivity of a steel plan.
-We have :
+This second example is about a phenomenon that impacts the productivity of a steel plant.
+We have:
 		\begin{itemize}
@@
 			\item $p=145$ variables from the whole process to explain it but only $n=100$ individuals.
-			\item The stakes : $20\%$ of productivity to gain
+			\item The stakes : $20\%$ of productivity to gain on a specific product with high added value.
 		\end{itemize}
@@
 \end{figure}   			
-	Here $n<p$ so we only compare the leave-one-out cross-validation MSE.
-	\textsc{CorReg} improves LASSO by $5.24\%$ and elasticnet by $8.60\%$. \textsc{CorReg} combined with LASSO gives the best result but it is only a leave-on-out MSE.
+%	Here $n<p$ so we only compare the leave-one-out cross-validation MSE.
+%	\textsc{CorReg} improves LASSO by $5.24\%$ and elasticnet by $8.60\%$. \textsc{CorReg} combined with LASSO gives the best result but it is only a leave-on-out MSE.
+%\begin{table}[h!]
+%\centering
+%\begin{tabular}{|c|c|c|}
+%	\hline 
+%	Model & MSE & Complexity (with intercept) \\ 
+%	\hline 
+%	LASSO & 105.54 & 34 \\ 
+%	\hline 
+%	\textsc{CorReg} + LASSO & 100 & 18 \\ 
+%	\hline 
+%	elasticnet & 129.94 & 13 \\ 
+%	\hline 
+%	\textsc{CorReg} + elasticnet & 118.76 & 21 \\ 
+%	\hline 
+%\end{tabular} 
+%\caption{Results obtained with leave-one out cross-validation. $n=100, p=145$.}	
+%\end{table}
+The response variable was binary but $n$ was too small compared to $p$ to use logistic regression so we have considered $\boldsymbol{Y}$ as a continuous variable and then made imputation by $1$ when $\hat{\boldsymbol{Y}}>0.5$ and by $0$ else.
+
 \begin{table}[h!]
 \centering
-\begin{tabular}{|c|c|c|}
-	\hline 
-	Model & MSE & Complexity (with intercept) \\ 
-	\hline 
-	LASSO & 105.54 & 34 \\ 
-	\hline 
-	\textsc{CorReg} + LASSO & 100 & 18 \\ 
-	\hline 
-	elasticnet & 129.94 & 13 \\ 
-	\hline 
-	\textsc{CorReg} + elasticnet & 118.76 & 21 \\ 
+\begin{tabular}{|c c|c|c|}
+	\hline 
+	Method& indicator& With CorReg & without CorReg \\ 
+	\hline
+	OLS & well-classified & 100& 56 \\
+		& MSE (leave-one-out) & & \\
+		& complexity & 91& 100 \\
+	\hline 
+		LASSO & well-classified &93 &93 \\
+		& MSE (leave-one-out) & 0.106 & 0.120\\
+		& complexity & 27&34\\
+	\hline 
+		Elasticnet & well-classified &84 &87 \\
+		& MSE (leave-one-out) &0.140 &0.148\\
+		& complexity &10 &13\\
+	\hline 
+		Ridge & well-classified &88 &85 \\
+		& MSE (leave-one-out) & &\\
+		& complexity &91 &146\\
 	\hline 
@@
 \end{table}
+
 In this precise case, \textsc{CorReg} found a structure that helped to decorrelate covariates in interpretation and to find the relevant part of the process to optimize.
+This product is made by a long process that requires several steel plants so it was necessary to point out the steel plant where the problem occurred.
 
@@
 
-	This paper only treats linear regression but such a pretreatment could be used for logistic regression, {\it etc.}
+	This paper only treats linear regression but as a pre-treatment, CorReg could be used for logistic regression, {\it etc.}
 	So the subject is still wide opened.
@@
 		\end{eqnarray}
+		So a sufficient condition for identifiability is to have at least one Gaussian mixture in each sub-regression.
 %		We have the unique (Gaussian are allowed only on the right of the sub-regressions) decomposition $f(\boldsymbol{X}_r|\boldsymbol{X}_f)=f(\boldsymbol{X}_r|\boldsymbol{X}^G,\boldsymbol{X}^{G^c})=f(\boldsymbol{X}_r|\boldsymbol{X}^{G^c})f(\boldsymbol{X}^{G}) $
@@
 
-Here, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
-	\begin{eqnarray}
-			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
-			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j  \label{densitydroite}\\
-			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
-						\mathcal{GM}(\bigotimes_{\substack{i \in I_f^j  }}\boldsymbol{\pi}_i \ ; 
-			                         \bigoplus_{\substack{i \in I_f^j  }} \alpha_{i,j} \boldsymbol{\mu}_i \ ; 
-			                         \boldsymbol{\sigma}_j^2+\bigoplus_{\substack{i \in I_f^j  }}\alpha_{i,j}^2 \boldsymbol{\sigma}_i^2 )\label{densitygauche}          
-		\end{eqnarray}
-		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
-
-	The presence of two Gaussian mixtures on the right of the sub-regression gives the identifiability of it (even with also Gaussian variables) by the strict increase of the component number when combining Gaussian mixtures.
-	So a sub-regression is identifiable if it has only Gaussian mixtures (even only one) on the right or if it has at least two Gaussian mixtures (even with additional Gaussian variables)
+%Here, $\otimes$ and $\oplus$ denote respectively the kronecker product and sum.
+%	\begin{eqnarray}
+%			\forall j \in I_r : \  \boldsymbol{X}^j_{|\boldsymbol{X}_f}&\sim & \mathcal{N}(\boldsymbol{X}_f\boldsymbol{\alpha}_j,\sigma^2_j\boldsymbol{I}_n) \label{densitycondgauche}\\
+%			\forall j \in I_f : \boldsymbol{X}^j_{|S} &\sim& \mathcal{GM}(\boldsymbol{\pi}_j;\boldsymbol{\mu}_j;\boldsymbol{\sigma}^2_j) \textrm{ with } \boldsymbol{\pi}_j,\boldsymbol{\mu}_j,\boldsymbol{\sigma}^2_j \textrm{ vectors of size } K_j  \label{densitydroite}\\
+%			\textrm{And we obtain, } \forall j \in I_r : \  \boldsymbol{X}^j&\sim & 
+%						\mathcal{GM}(\bigotimes_{\substack{i \in I_f^j  }}\boldsymbol{\pi}_i \ ; 
+%			                         \bigoplus_{\substack{i \in I_f^j  }} \alpha_{i,j} \boldsymbol{\mu}_i \ ; 
+%			                         \boldsymbol{\sigma}_j^2+\bigoplus_{\substack{i \in I_f^j  }}\alpha_{i,j}^2 \boldsymbol{\sigma}_i^2 )\label{densitygauche}          
+%		\end{eqnarray}
+%		So when we compare (\ref{densitydroite}) and (\ref{densitygauche}) we see that the number of components in $I_r$ variables differs when subregressions are of length $>1$ (almost 2 predictors) with multiple-class predictors (so the kronecker product is effective).
+%	
+%	The presence of two Gaussian mixtures on the right of the sub-regression gives the identifiability of it (even with also Gaussian variables) by the strict increase of the component number when combining Gaussian mixtures.
+%	So a sub-regression is identifiable if it has only Gaussian mixtures (even only one) on the right or if it has at least two Gaussian mixtures (even with additional Gaussian variables)
+%	
+%
 

### 2014-05-27 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@623 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up bv MSE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@624 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 When quality crises occurs, time is extremely precious so automation is a real stake.
+The combinatorial aspect of the sub-regression models makes it impossible to do manually.
 
@@
 	\begin{minipage}[c]{.30\linewidth}
-			\includegraphics[height=150px, width=150px]{figures/R2corregBVBI.png} 
-			\caption{$R^2_{adj}$ of the 67 sub-regressions.}
+				\includegraphics[height=150px,width=150px]{figures/histcorrelBVBI.png} 
+			\caption{Histogram of correlations in $\boldsymbol{X}$.} 
 	\end{minipage} \hfill
    \begin{minipage}[r]{.30\linewidth}
-			\includegraphics[height=150px,width=150px]{figures/histcorrelBVBI.png} 
-			\caption{Histogram of correlations in $\boldsymbol{X}$.} 
+			\includegraphics[height=150px, width=150px]{figures/R2corregBVBI.png} 
+			\caption{$R^2_{adj}$ of the 55 sub-regressions.}\label{R2bv}
    \end{minipage}
 \end{figure}   	
-%	Here $n<p$ so we only compare the leave-one-out cross-validation MSE.
-%	\textsc{CorReg} improves LASSO by $5.24\%$ and elasticnet by $8.60\%$. \textsc{CorReg} combined with LASSO gives the best result but it is only a leave-on-out MSE.
-%\begin{table}[h!]
-%\centering
-%\begin{tabular}{|c|c|c|}
-%	\hline 
-%	Model & MSE & Complexity (with intercept) \\ 
-%	\hline 
-%	LASSO & 105.54 & 34 \\ 
-%	\hline 
-%	\textsc{CorReg} + LASSO & 100 & 18 \\ 
-%	\hline 
-%	elasticnet & 129.94 & 13 \\ 
-%	\hline 
-%	\textsc{CorReg} + elasticnet & 118.76 & 21 \\ 
-%	\hline 
-%\end{tabular} 
-%\caption{Results obtained with leave-one out cross-validation. $n=100, p=145$.}	
-%\end{table}
+CorReg found 55 sub-regressions as shown in Figure \ref{R2bv}. One of them seems to be weak $R^2=0.17$ but is not linear (points out a link between diameter of a coil and some shape indicator).
+
 The response variable was binary but $n$ was too small compared to $p$ to use logistic regression so we have considered $\boldsymbol{Y}$ as a continuous variable and then made imputation by $1$ when $\hat{\boldsymbol{Y}}>0.5$ and by $0$ else.
@@
 	OLS & well-classified & 100& 56 \\
-		& MSE (leave-one-out) & & \\
+		& MSE (leave-one-out)& 1.95& 51 810\\
 		& complexity & 91& 100 \\
@@
 		LASSO & well-classified &93 &93 \\
-		& MSE (leave-one-out) & 0.106 & 0.120\\
+		& MSE (leave-one-out)& 0.106 & 0.120\\
 		& complexity & 27&34\\
@@
 		Elasticnet & well-classified &84 &87 \\
-		& MSE (leave-one-out) &0.140 &0.148\\
+		& MSE (leave-one-out)&0.140 &0.148\\
 		& complexity &10 &13\\
@@
 		Ridge & well-classified &88 &85 \\
-		& MSE (leave-one-out) & &\\
+		& MSE (leave-one-out)& 0.179 & 0.177\\
 		& complexity &91 &146\\

### 2014-05-27 www/article13.tex
up non lin
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@625 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \FloatBarrier
-\subsubsection{non-linear case}
+\subsubsection{Non-linear case}
 
@@
 $\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as previously but with random signs for the components means.
-$\boldsymbol{X}_r=\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2-2\boldsymbol{X}_3+ \varepsilon$.
+$\boldsymbol{X}_r=\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2+\boldsymbol{X}_3+ \varepsilon$.
 The matrix $\boldsymbol{X}$ is then scaled and we get $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\varepsilon_Y$.
@@
    \begin{minipage}[r]{.48\linewidth}
-			\includegraphics[height=180px,width=240px]{figures/res_article/robust_MSE.png} \label{MSEnonlin}
-			\caption{MSE on the main regression for OLS(thick) and LASSO (thin) used both with (plain) or without CorReg (dotted).}
+			\includegraphics[height=180px,width=240px]{figures/res_article/robust_MSE.png} 
+			\caption{MSE on the main regression for OLS(thick) and LASSO (thin) used both with (plain) or without CorReg (dotted).}\label{MSEnonlin}
    \end{minipage}

### 2014-06-10 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@626 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@631 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@635 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
-	% Some recent variants of the \textsc{LASSO} do exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
+	% Some recent variants of the \textsc{LASSO} do exist for the choice of the penalization coefficient like the adaptive \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
 But
@@
 \begin{eqnarray}
+		P(\boldsymbol{Y}|\boldsymbol{X}_f)&=& \int_{\boldsymbol{X}_r}P(\boldsymbol{Y}|\boldsymbol{X}_r,\boldsymbol{X}_f)P(\boldsymbol{X}_r|\boldsymbol{X}_f) d \boldsymbol{X} \\
 	\boldsymbol{Y}_{|\boldsymbol{X}_f,S}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{Trueexpl} \\
@@
 	\includegraphics[width=500px]{figures/MQEexplOLSp5.png}\label{MQE1}
-	\caption{MSE of OLS and CorReg (dotted) estimators for varying $(1-R^2)$ of the sub-regression, $n$ and $\sigma_Y$.}
+	\caption{MSE of OLS (plain) and CorReg (dotted) estimators for varying $(1-R^2)$ of the sub-regression, $n$ and $\sigma_Y$.}
 \end{figure} 
@@
 	The walk follows a time-homogeneous Markov Chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we just compute the probabilities when we need them).
-	At each step the markov chain moves with probabiliy:
+	At each step the markov chain moves with probability:
 	\begin{eqnarray}
@@
 			\includegraphics[height=180px,width=245px]{figures/res_article/robust_S.png} 
-			\caption{Frequency of true $\boldsymbol{X}_r$ found (plain)   when the paramater $a$ increases}\label{resnonlin}
+			\caption{Evolution of the quality of $\hat{S}$ when the paramater $a$ increases}\label{resnonlin}
 	\end{minipage} \

### 2014-06-11 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@637 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \section{Numerical results on simulated datasets} \label{sectionsimul}
+	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$. (see figure \ref{toutOLS}).
+	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
+This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
 
@@
 	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
-	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$.
-%(see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
-	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients.
-This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
+
 	We compare different methods with and without CorReg as a pre-treatment.

### 2014-06-11 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@638 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		\begin{figure}[h!]
-	\begin{minipage}[l]{.30\linewidth}
-			\includegraphics[width=150px]{figures/correlbeforeafter.png} 
-			\caption{Correlations between the covariates in $\boldsymbol{X}$ (upper) and $\hat{\boldsymbol{X}}_f$ (lower).}
+	\begin{minipage}[t]{.30\linewidth}
+			\includegraphics[height=150px,width=150px]{figures/nbcompBV.png}%{figures/mixmod.png} 
+			\caption{Distribution of the number of components found for each covariate.}\label{graphMixmod}
 	\end{minipage} \hfill
-	\begin{minipage}[c]{.30\linewidth}
+	\begin{minipage}[t]{.30\linewidth}
+			\includegraphics[height=150px, width=150px]{figures/GMcriseBV.png}%{figures/histR2exfos.png} $R^2_{adj}$ of the 76 sub-regressions.
+			\caption{Another example of non-Gaussian real variable easily modeled by a Gaussian mixture.}
+	\end{minipage} \hfill
+   \begin{minipage}[t]{.30\linewidth}
 			\includegraphics[height=150px,width=150px]{figures/histcorrelBVBI.png} 
-			\caption{Histogram of correlations in $\boldsymbol{X}$.} 
-	\end{minipage} \hfill
-   \begin{minipage}[r]{.30\linewidth}
+			\caption{Histogram of correlations in $\boldsymbol{X}$.} \label{compareMSEexfos}
+   \end{minipage}
+\end{figure}
+
+CorReg found 55 sub-regressions as shown in Figure \ref{R2bv}. One of them seems to be weak $R^2=0.17$ but is not linear (points out a link between diameter of a coil and some shape indicator).
+
+\begin{figure}
+\centering
 	\includegraphics[height=150px, width=150px]{figures/R2corregBVBI.png} 
 			\caption{$R^2_{adj}$ of the 55 sub-regressions.}\label{R2bv}
-   \end{minipage}
-\end{figure}   	
-CorReg found 55 sub-regressions as shown in Figure \ref{R2bv}. One of them seems to be weak $R^2=0.17$ but is not linear (points out a link between diameter of a coil and some shape indicator).
-
+\end{figure}
 The response variable was binary but $n$ was too small compared to $p$ to use logistic regression so we have considered $\boldsymbol{Y}$ as a continuous variable and then made imputation by $1$ when $\hat{\boldsymbol{Y}}>0.5$ and by $0$ else.

### 2014-06-17 pkg/CorReg/R/density_estimation.R
matricial shaped output
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@648 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -6,10 +6,11 @@
 #' @param nbclustmin min number of clusters in the gaussian mixtures
 #' @param verbose verbose or not
 #' @param detailed boolean to give the details of the mixtures found
+#' @param matshape boolean to give the detail in matricial shape
 #' @param max boolean. Use an heuristic to shrink nbclustmax according to the number of individuals in the dataset
 #' @param mclust boolean. Use mclust instead of Rmixmod
 #' @param nbini number of initial points for Rmixmod
-density_estimation<-function(X=X,nbclustmax=10,nbclustmin=1,verbose=FALSE,detailed=FALSE,max=TRUE,mclust=TRUE,nbini=20){
+density_estimation<-function(X=X,nbclustmax=10,nbclustmin=1,verbose=FALSE,detailed=FALSE,max=TRUE,mclust=TRUE,nbini=20,matshape=FALSE){
   #X est la matrice sans constante
    X=1*as.matrix(X)
   n=nrow(X)
@@ -37,7 +38,11 @@ density_estimation<-function(X=X,nbclustmax=10,nbclustmin=1,verbose=FALSE,detail
         prop=res[6][1]#proportions
         meansvect=c(res[6][2])#means
         varvect=unlist(res[6][3])#variances
-        detailsmat[[i]]=cbind(prop,meansvect,varvect,i)
+        if(matshape){
+           detailsmat=rbind(detailsmat,cbind(prop,meansvect,varvect,i))
+        }else{
+           detailsmat[[i]]=cbind(prop,meansvect,varvect,i)
+        }      
       }
     }
   }else{#on utilise mclust
@@ -59,7 +64,11 @@ density_estimation<-function(X=X,nbclustmax=10,nbclustmin=1,verbose=FALSE,detail
             prop=res[6][1]#proportions
             meansvect=c(res[6][2])#means
             varvect=unlist(res[6][3])#variances
-            detailsmat[[i]]=cbind(prop,meansvect,varvect,i)
+            if(matshape){
+               detailsmat=rbind(detailsmat,cbind(prop,meansvect,varvect,i))
+            }else{
+               detailsmat[[i]]=cbind(prop,meansvect,varvect,i)
+            }
          }
       }else{
          nbclust[i]=res$parameters$variance$G
@@ -68,7 +77,11 @@ density_estimation<-function(X=X,nbclustmax=10,nbclustmin=1,verbose=FALSE,detail
            prop=res$parameters$pro#proportions
[…]

### 2014-06-17 pkg/CorReg/R/Estep.R
E step for missing values (structure)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@649 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,27 @@
+#' Imputation of missing values knowing alpha (E step of the EM)
+#' 
+#' @param Ir the set of indices of the covariates on the left (computed once before EM to be efficient)
+#' @param Zc profil (sum) colonne de Z
+#' @param alpha matrix of the coefficients (p+1)xp
+#' @export
+Estep<-function(X=X,alpha=alpha,M=M,Z=Z,mixmod=mixmod,Zc=Zc){
+   #X bouge, mais pas alpha ni Z ni M ni mixmod
+   #on liste tout ce qu'il faut (on verra après comment extraire ça pour optimiser)
+   quimank=which(is.na(X),arr.ind=TRUE)# apriori on stockera les deux formats (M et quimank en permanence) et on fera pareil pour Z
+   nbmank=nrow(quimank)
+   Zc=colSums(Z)
+   #peut-être entrer quimank et écrire uniquement dans le Cpp M au format Eigen creux (sans contact avec R)
+   
+   #on remplit
+   for (miss in 1:nbmank){#pour chaque valeur manquante
+      miss=quimank[miss,]
+      if(Zc[miss[2]]!=0){#trou à gauche
+         #imputation par moyenne du mélange gaussien résultant de l'intégration de la sous-regression
+         X[miss[1],miss[2]]=alpha[1,miss[2]]+as.matrix(X[miss[1],-miss[2]])%*%as.matrix(alpha[-1,quimank[miss,2]][-quimank[miss,2]])
+
+      }else{#trou à droite
+         
+      }
+   }
+   return(X)
+}

### 2014-06-17 pkg/CorReg/R/mixture_generator.R
bug in B generation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@650 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -75,10 +75,14 @@ mixture_generator<-function(n=130,
     B[1,list_X2]=rpois(R,lambdapois)*(rep(-1,R)+2*rbinom(R,1,positive)) # 1ere ligne de B = ?????
     for(j in list_X2){#remplissage aléatoire de max_compl éléments de B sur chaque colonne à gauche
       if(is.null(Z)){
-         B[sample(qui[-c(list_X2-1)],size=max_compl),j]=(1/max_compl)*max(1,rpois(max_compl,lambdapois))*(rep(-1,max_compl)+2*rbinom(max_compl,1,positive))
+         loc=(1/max_compl)*rpois(max_compl,lambdapois)*(rep(-1,max_compl)+2*rbinom(max_compl,1,positive))
+         loc[loc==0]=1
+         B[sample(qui[-c(list_X2-1)],size=max_compl),j]=loc
       }else{
          compl_loc=colSums(Z)[j]
-         B[sample(qui[-c(list_X2-1)],size=compl_loc),j]=(1/compl_loc)*max(1,rpois(compl_loc,lambdapois))*(rep(-1,compl_loc)+2*rbinom(compl_loc,1,positive))
+         loc=(1/compl_loc)*rpois(compl_loc,lambdapois)*(rep(-1,compl_loc)+2*rbinom(compl_loc,1,positive))
+         loc[loc==0]=1
+         B[sample(qui[-c(list_X2-1)],size=compl_loc),j]=loc
       }
     }
     #ajout de G

### 2014-06-23 pkg/CorReg/R/Estep.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@666 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,26 +1,77 @@
 #' Imputation of missing values knowing alpha (E step of the EM)
-#' 
+#' @param X the dataset with missing values
+#' @param M binary matrix X-sized (1=missing) 
 #' @param Ir the set of indices of the covariates on the left (computed once before EM to be efficient)
 #' @param Zc profil (sum) colonne de Z
-#' @param alpha matrix of the coefficients (p+1)xp
-#' @export
-Estep<-function(X=X,alpha=alpha,M=M,Z=Z,mixmod=mixmod,Zc=Zc){
+#' @param mixmod the matrix from mixmod
+#' @param Zc p-sized vector to see if Z column is empty or not
+#' @param alpha matrix of the coefficients (p+1)xp format Matrix
+#'
+Estep<-function(X=X,alpha=alpha,M=NULL,Z=NULL,mixmod=mixmod,Zc=Zc,X1=FALSE){
    #X bouge, mais pas alpha ni Z ni M ni mixmod
-   #on liste tout ce qu'il faut (on verra après comment extraire ça pour optimiser)
+   require(Matrix)
    quimank=which(is.na(X),arr.ind=TRUE)# apriori on stockera les deux formats (M et quimank en permanence) et on fera pareil pour Z
-   nbmank=nrow(quimank)
-   Zc=colSums(Z)
-   #peut-être entrer quimank et écrire uniquement dans le Cpp M au format Eigen creux (sans contact avec R)
    
+   p=ncol(X)
+   n=nrow(X)
+   #on liste tout ce qu'il faut (on verra après comment extraire ça pour optimiser)
+   nbmank=nrow(quimank)
+   M=sparseMatrix(i=quimank[,1],j=quimank[,2],dims=c(n,p))
+   Z=as(Matrix(Z,sparse=TRUE),"nsparseMatrix")
+   Zc=Matrix(colSums(Z))
+   X=fillmiss(X=X,X1=FALSE,mixmod=FALSE)
+   
+   if(is.null(mixmod)){
+      mixmod=density_estimation(X=X,detailed=TRUE,matshape=TRUE)
+   }
+   #peut-être entrer quimank et écrire uniquement dans le Cpp M au format Eigen creux (sans contact avec R)
    #on remplit
    for (miss in 1:nbmank){#pour chaque valeur manquante
       miss=quimank[miss,]
       if(Zc[miss[2]]!=0){#trou à gauche
          #imputation par moyenne du mélange gaussien résultant de l'intégration de la sous-regression
[…]

### 2014-06-23 pkg/CorReg/R/Gibbs_X_ij_IF.R

@@ -1,0 +1,22 @@
+#' Gibbs sampler for redundant covariates
+#' @param X est la ligne concernée (X[i,]) 
+#' @param components est le vecteur des classes pour i (vecteur creux de taille p avec pr elements non nuls)
+#' @param mixmod is mixmod$details
+Gibbs_X_ij_IF<-function(Z=Z,X=X,p=p,mui=mui,sigmai=sigmai,alpha=alpha,mixmod=mixmod,j=j,components=components){
+   Sigma_j_reste=rep(0,times=p-1)
+   Sigma_reste_reste=matrix(0,nrow=p-1,ncol=p-1)
+   for(i in (1:p)[-j]){
+      if(i!=j){
+         #variable à droite de la manquante (impossible car elle est à droite) donc Z[i,j]=0
+         if(Z[j,i]!=0){#variable à gauche de la manquante 
+            Sigma_j_reste[i]=alpha[-1][j,i]*sigmai[i]
+         }#sinon indépendance car pas à droite non plus donc rien à faire car initialisation à 0
+      
+        
+      }
+   }
+   
+   mu=mui[j]+Sigma_j_reste%*%solve(Sigma_reste_reste)%*%(X[-j]-mui[-j])
+   sigma=0
+   return(rnorm(1,mean=mu,sd=sigma))
+}

### 2014-06-23 pkg/CorReg/R/muiZ.R

@@ -1,0 +1,33 @@
+
+# ' @param alpha the matrix of coefficients of the sub-regressions
+muiZ<-function(mixmod=mixmod,components=components,Z=Z,Zc=Zc,alpha=alpha){
+   
+   mui=components#en C on les prendra en entrée pour gagner une initialisation à chaque fois
+   sigmai=components
+   for(i in 1:p){
+      mui[i]=mixmod[[i]][components[i],2]
+      sigmai[i]=mixmod[[i]][components[i],3]#attention c'est une variance
+   }
+   Sigma=0*Z# à intialiser au début en creux symétrique et à passer en argument
+   for(i in 1:p){
+      for(j in 1:i){
+         if(i==j){#variance sur la diagonale
+            if(Zc[i]>0){#variable à gauche
+               Sigma[i,i]=sigmai[i]+alpha[-1][,j]^2%*%sigmai#les alphas nuls font le tri tout seul
+            }else{#variable à droite
+               Sigma[i,i]=sigmai[i]
+            }
+         }else if (Zc[i]!=0){#si une à gauche
+            if(Zc[j]!=0){#Les deux sont à gauche
+               #veiller à ne pas stocker des 0 en testant avant d'écrire ?
+               Sigma[i,j]=(alpha[-1][,i]*alpha[-1][,j])%*%sigmai#attention premier produit hadamard
+               Sigma[j,i]=Sigma[i,j]
+            }else{#j est à droite
+               Sigma[i,j]=alpha[-1][j,i]*sqrt(sigmai[j])
+               Sigma[j,i]=Sigma[i,j]
+            }
+         }#sinon, deux distinctes droite alors covariance nulle
+      }
+   }
+   return(list(mui=mui,sigmai=sigmai,Sigma=Sigma))
+}

### 2014-06-23 pkg/CorReg/R/Gibbs_X_ij_IR.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@666 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,6 @@
+#' Gibbs sampler for redundant covariates
+#' 
+Gibbs_X_ij_IR<-function(X=X,sigma=sigma,alpha=alpha){
+   mu=alpha[1]+X%*%alpha[-1]
+   return(rnorm(1,mean=mu,sd=sigma))
+}

### 2014-06-25 pkg/CorReg/R/Gibbs.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@669 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,53 @@
+Gibbs<-function(last=FALSE,nbit=1,mixmod=mixmod,X=X,comp_vect=comp_vect,missrow=missrow,quimiss=quimiss,
+                Z=Z,Zc=Zc,alpha=alpha,sigma_IR=sigma_IR,nbclust_vect=nbclust_vect,Ir=Ir){
+   for(iter in 1:nbit){
+      missrow=1
+      resmui=muiZ(mixmod = mixmod,components=comp_vect[missrow,],Z=Z,Zc=Zc,alpha=alpha,Ir=Ir,sigma_IR=sigma_IR)#on commence ligne 1
+      mui=resmui$mui
+      sigmai=resmui$sigmai
+      Sigma=resmui$Sigma
+      for(i in 1:nbmiss){
+#          print(paste("i",i))
+         miss=quimiss[i,]
+         if(miss[1]!=missrow){
+            missrow=miss[1]
+            resmui=muiZ(mixmod = mixmod,components=comp_vect[missrow,],Z=Z,Zc=Zc,alpha=alpha,Ir=Ir,sigma_IR=sigma_IR)#mise à jour
+            mui=resmui$mui
+            sigmai=resmui$sigmai
+            Sigma=resmui$Sigma
+         }
+#          print(paste("i2",i))
+#          print(Ir)
+#          print(miss[2])
+#          print(miss[2]%in%Ir)
+         if(miss[2]%in%Ir){#redundant covariate missing
+#             print("left")
+            
+            X[miss[1],miss[2]]=Gibbs_X_ij_IR(X=X[miss[1],],sigma=sigma_IR[miss[2]],alpha=alpha[,miss[2]])
+         }else{#missing right
+#             print("right")
+            X[miss[1],miss[2]]=Gibbs_X_ij_IF(Z=Z,X=X[miss[1],],mui=mui,sigmai=sigmai,Sigma=Sigma,alpha=alpha,mixmod=mixmod,j=miss[2],components=comp_vect[miss[1],])
+         }
+#          print(paste("i3",i))
+      }
+      if(!last | (iter!=nbit)){
+         #Imputation des classes
+         for(j in 1:p){
+#             print(paste("j",j))
+            if(nbclust_vect[j]>1){
+               for (i in 1:rowmiss){
+#                   print("a")
+                  vect=rmultinom(1, 1, tik(x=X[i,j],nbclust=nbclust_vect[j],mixmod=mixmod[[j]]))
+#                   print("b")
[…]

### 2014-06-25 pkg/CorReg/R/Gibbs_X_ij_IR.R

@@ -1,5 +1,5 @@
-#' Gibbs sampler for redundant covariates
-#' 
+# ' Gibbs sampler for redundant covariates
+# ' 
 Gibbs_X_ij_IR<-function(X=X,sigma=sigma,alpha=alpha){
    mu=alpha[1]+X%*%alpha[-1]
    return(rnorm(1,mean=mu,sd=sigma))

### 2014-06-25 pkg/CorReg/R/SEM.R

@@ -1,0 +1,49 @@
+SEM<-function(nbit_gibbs=1,nbit_SEM=50,warm=10,mixmod=mixmod,X=X,comp_vect=comp_vect,missrow=missrow,quimiss=quimiss,
+              Z=Z,Zc=Zc,alpha=alpha,sigma_IR=sigma_IR,nbclust_vect=nbclust_vect,Ir=Ir,Xout=FALSE,alphaout=TRUE){
+   last=FALSE
+   result=list()
+   
+   #initialisation
+   if(is.null(alpha)){
+      alpha=hatB(Z = Z,X=X)
+   }
+   if(is.null(Zc)){
+      Zc=colSums(Z)
+   }
+   if(is.null(Ir)){
+      Ir=which(Zc!=0)
+   }
+   if(is.null(comp_vect)){
+      comp_vect=matrix(rep(1,times=p*rowmiss),ncol=p)
+   }
+   if(is.null(sigma_IR)){
+      sigma_IR=rep(0,times=p)#résidus des regressions (fixés avec alpha donc stables pour gibbs)
+      for (j in Ir){
+         sigma_IR[j]=sd(X[,j]-X%*%alpha[-1,j])
+      }
+   }
+   
+   
+   for (i in 1:(nbit_SEM+warm)){
+      if(i==(nbit_SEM+warm)){last=TRUE}
+      #SE step
+      resgibbs2=Gibbs(last=last,nbit=nbit_gibbs,mixmod=mixmod,X=X,comp_vect=comp_vect,missrow=missrow,quimiss=quimiss,
+                     Z=Z,Zc=Zc,alpha=alpha,sigma_IR=sigma_IR,nbclust_vect=nbclust_vect,Ir=Ir)
+      comp_vect=resgibbs2$comp_vect
+      X=resgibbs2$X
+      #step M
+      resM=Mstep(Z=Z,X=X,sigma_IR=sigma_IR,Ir=Ir)
+      alpha=resM$alpha
+      sigma_IR=resM$sigma_IR
+      if(i>warm){
+         if(i==warm+1){
+            if(alphaout){result$alpha=alpha/nbit_SEM}
+            if(Xout){result$X=X/nbit_SEM}
+         }else{
+            if(alphaout){result$alpha=result$alpha+alpha/nbit_SEM}#optimiser au format creux
+            if(Xout){result$X=result$X+X/nbit_SEM}#optimiser en ne modifiant que les manquants
+         }   
+      }
+   }
+   return(result)
+}

### 2014-06-25 pkg/CorReg/R/muiZ.R

@@ -1,33 +1,31 @@
 
 # ' @param alpha the matrix of coefficients of the sub-regressions
-muiZ<-function(mixmod=mixmod,components=components,Z=Z,Zc=Zc,alpha=alpha){
-   
+muiZ<-function(mixmod=mixmod,components=components,Z=Z,Zc=Zc,alpha=alpha,Ir=Ir,sigma_IR=sigma_IR){ 
    mui=components#en C on les prendra en entrée pour gagner une initialisation à chaque fois
    sigmai=components
-   for(i in 1:p){
-      mui[i]=mixmod[[i]][components[i],2]
-      sigmai[i]=mixmod[[i]][components[i],3]#attention c'est une variance
+   for(i in (1:p)[-Ir]){#vaiable à droite donc mixmod
+         mui[i]=mixmod[[i]][components[i],2]
+         sigmai[i]=mixmod[[i]][components[i],3]#attention c'est une variance
    }
-   Sigma=0*Z# à intialiser au début en creux symétrique et à passer en argument
-   for(i in 1:p){
-      for(j in 1:i){
-         if(i==j){#variance sur la diagonale
-            if(Zc[i]>0){#variable à gauche
-               Sigma[i,i]=sigmai[i]+alpha[-1][,j]^2%*%sigmai#les alphas nuls font le tri tout seul
-            }else{#variable à droite
-               Sigma[i,i]=sigmai[i]
-            }
-         }else if (Zc[i]!=0){#si une à gauche
-            if(Zc[j]!=0){#Les deux sont à gauche
-               #veiller à ne pas stocker des 0 en testant avant d'écrire ?
-               Sigma[i,j]=(alpha[-1][,i]*alpha[-1][,j])%*%sigmai#attention premier produit hadamard
-               Sigma[j,i]=Sigma[i,j]
-            }else{#j est à droite
-               Sigma[i,j]=alpha[-1][j,i]*sqrt(sigmai[j])
-               Sigma[j,i]=Sigma[i,j]
-            }
-         }#sinon, deux distinctes droite alors covariance nulle
+   for(j in Ir){
+      mui[j]=alpha[1,j]+t(as.matrix(mui))%*%(alpha[-1,j])#intercept plus combi linéaire
+      sigmai[j]=t(as.matrix(sigmai))%*%((alpha[-1,j])^2)+sigma_IR[j]^2#attention c'est une variance
+   }
[…]

### 2014-06-25 pkg/CorReg/R/tik.R

@@ -1,0 +1,21 @@
+# x est le point étudié
+#nbclust le nombre de classes sur xj
+#mixmod le tableau mixmod pour xj
+tik<-function(x=x,nbclust=nbclust,mixmod=mixmod){
+   prob=rep(0,times=nbclust)
+   for(i in 1:nbclust){
+      prob[i]=mixmod[i,1]*dnorm(x,mean = mixmod[i,2],sd=sqrt(mixmod[i,3]))
+      if(is.na(prob[i])){
+         print("NA in tik")
+         prob[i]=0
+      }
+   }
+   sumprob=sum(prob)
+   if(sumprob==0){
+      prob[1]=1
+   }else{
+      prob=prob/sumprob
+   }
+   
+   return(prob)
+}

### 2014-06-25 pkg/CorReg/R/Estep.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@669 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -21,8 +21,8 @@ Estep<-function(X=X,alpha=alpha,M=NULL,Z=NULL,mixmod=mixmod,Zc=Zc,X1=FALSE){
    Zc=Matrix(colSums(Z))
    X=fillmiss(X=X,X1=FALSE,mixmod=FALSE)
    
-   if(is.null(mixmod)){
-      mixmod=density_estimation(X=X,detailed=TRUE,matshape=TRUE)
+   if(is.null(mixmod)){print("ok")
+      mixmod=density_estimation(X=X,detailed=TRUE,matshape=TRUE)$details
    }
    #peut-être entrer quimank et écrire uniquement dans le Cpp M au format Eigen creux (sans contact avec R)
    #on remplit

### 2014-06-25 pkg/CorReg/R/Gibbs_X_ij_IF.R

@@ -1,22 +1,22 @@
-#' Gibbs sampler for redundant covariates
-#' @param X est la ligne concernée (X[i,]) 
-#' @param components est le vecteur des classes pour i (vecteur creux de taille p avec pr elements non nuls)
-#' @param mixmod is mixmod$details
-Gibbs_X_ij_IF<-function(Z=Z,X=X,p=p,mui=mui,sigmai=sigmai,alpha=alpha,mixmod=mixmod,j=j,components=components){
-   Sigma_j_reste=rep(0,times=p-1)
-   Sigma_reste_reste=matrix(0,nrow=p-1,ncol=p-1)
-   for(i in (1:p)[-j]){
-      if(i!=j){
-         #variable à droite de la manquante (impossible car elle est à droite) donc Z[i,j]=0
-         if(Z[j,i]!=0){#variable à gauche de la manquante 
-            Sigma_j_reste[i]=alpha[-1][j,i]*sigmai[i]
-         }#sinon indépendance car pas à droite non plus donc rien à faire car initialisation à 0
-      
-        
-      }
+# ' Gibbs sampler for redundant covariates
+# '  @param X est la ligne concernée (X[i,]) 
+# ' @param components est le vecteur des classes pour i (vecteur creux de taille p avec pr elements non nuls)
+# ' @param mixmod is mixmod$details
+Gibbs_X_ij_IF<-function(Z=Z,X=X,p=p,mui=mui,sigmai=sigmai,Sigma=Sigma,alpha=alpha,mixmod=mixmod,j=j,components=components){
+   Sigma_j_reste=Sigma[j,-j]
+   Sigma_reste_reste=Sigma[-j,-j]
+   prodmat=Sigma_j_reste%*%solve(Sigma_reste_reste)
+   mu=mui[j]+prodmat%*%(X[-j]-mui[-j])
+   sigma=sigmai[j]-prodmat%*%Sigma_j_reste;
+   sigma=as.numeric(sigma)
+#    print(paste("sigma",sigma))
+   if(as.numeric(sigma)<=0){
+      print(paste("sigmas<0",sigma, j));
+      sigma=-as.numeric(sigma)
+      stop
    }
-   
-   mu=mui[j]+Sigma_j_reste%*%solve(Sigma_reste_reste)%*%(X[-j]-mui[-j])
-   sigma=0
-   return(rnorm(1,mean=mu,sd=sigma))
[…]

### 2014-07-01 pkg/CorReg/R/victory_int.R
graphical comparison of curves
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@675 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,4 @@
+victory_int<-function(x=x,y=y,col=c("black","red","green"),pos=10){
+   quimin=apply(y,1,which.min)
+   points(x,rep(pos,times=length(x)),col=col[quimin],lwd="2",pch=15)
+}

### 2014-07-18 pkg/CorReg/R/correg.R
up comment
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@704 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -379,7 +379,6 @@ if(explnew){compl=TRUE;expl=TRUE}
       res$pred$AIC=mon_AIC(theta=res$pred$A,Y=Y,X=X,intercept=intercept) 
       
     }
-print("ok")
     #nouveau prédictif####
     if (prednew) {
        if(length(I2)<2 & select=="adalasso"){

### 2014-07-30 pkg/CorReg/R/victory_int.R
up last rect
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@732 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -4,7 +4,7 @@ victory_int<-function(x=x,y=y,col=c("black","red","green"),alpha=0.2,what=which.
    minloc=quimin[1]
    xdeb=x[1]
    for(i in 2:length(x)){
-      if(quimin[i]!=minloc){
+      if(quimin[i]!=minloc & i!=length(x)){
          couleur_loc=c(t(col2rgb(col[quimin[i-1]]))/255)
          couleur_loc=rgb(red=couleur_loc[1],green=couleur_loc[2],blue=couleur_loc[3],alpha)
          rect(xdeb, lim[3]-1, (x[i-1]+x[i])/2, lim[4]+1, border = couleur_loc, col = couleur_loc)

### 2014-08-01 pkg/CorReg/R/correg.R
bu with ridge corrected
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@736 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -313,8 +313,13 @@ if(explnew){compl=TRUE;expl=TRUE}
          A_inj=rep(0,times=ncol(Xtilde))
          A_inj[which(respike$gnet.scale!=0)]=OLS(X=Xtilde[,respike$gnet.scale!=0],Y=as.numeric(Ytilde),intercept=intercept)$beta 
       }else{#ridge
-        ridge_pred = linearRidge(Ytilde~0+.,data=data.frame(Xtilde))
-        A_inj=coef(ridge_pred)
+         if(ncol(Xtilde)>1){
+            ridge_pred = linearRidge(Ytilde~0+.,data=data.frame(Xtilde))
+            A_inj=coef(ridge_pred)
+         }else{
+            ridge_pred = OLS(X = Xtilde,Y=Ytilde,intercept=FALSE)#ridge has no sens on only one covariate
+            A_inj=ridge_pred$beta
+         }
       }
       A_pred = rep(0, times = ncol(X) + intercept)
       A_pred[I2 + intercept] = A_inj

### 2014-08-01 pkg/CorReg/R/matplot_zone.R
upgrade
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@738 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -10,9 +10,10 @@
 #' @param xlab (like in matplot)
 #' @param ylab (like in matplot)
 #' @param lwd  (like in matplot)
+#' @param lty  (like in matplot)
 #' @param main the main title (like in matplot)
 #' @export
-matplot_zone<-function(x=x,y=y,col=1:6,alpha=0.2,what=which.min,ylim=NULL,xlim=NULL,type="p",xlab=NULL,ylab="NULL",main=NULL,lwd=lwd){
+matplot_zone<-function(x=x,y=y,col=1:6,alpha=0.2,what=which.min,ylim=NULL,xlim=NULL,type="p",xlab=NULL,ylab="NULL",main=NULL,lwd=lwd,lty = 1:5){
    matplot(x,y,ylim=ylim,type=type,xlab=xlab,ylab=ylab,main=main,lwd=lwd,xlim=xlim)
    victory_int(x=x,y=y,col=col,what=what,alpha=alpha)
    matplot(x,y,ylim=ylim,add=TRUE,col=col,type=type,xlab=xlab,ylab=ylab,main=main,lwd=lwd,xlim=xlim)

### 2014-08-01 pkg/CorReg/R/Terminator.R
up warning terminator
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@739 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -7,7 +7,7 @@
 Terminator<-function(target="Sarah Connor", wrath=0.1,diag=0,Z=NULL){
    if(target[1]=="Sarah Connor" ){
       print("I'll be back !")
-   }else if (target=="bender"){
+   }else if (target[1]=="bender"){
       Bender()
    }else if(diag>0){
       quidiag=cbind(rep(1:n,length.out=max(n,p)),rep(1:p,length.out=max(n,p)))

### 2014-08-05 pkg/CorReg/R/recursive_tree.R
up trees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@750 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -7,8 +7,11 @@
 #'@param index to give a number to the plot
 #'@param print boolean to print the tree parameters
 #'@param plot boolean to plot the tree
+#'@param main the main title if plot=TRUE
+#'@param sub the subtitle (if NULL it is automatically added)
+#'@param lang the language for the automatic subtitle in the plot
 #'
-recursive_tree<-function(data=data,Y="Y",modele=NULL,kill=NULL,index=NULL,print=TRUE,plot=TRUE){
+recursive_tree<-function(data=data,Y="Y",modele=NULL,kill=NULL,index=NULL,print=TRUE,plot=TRUE,main=NULL,sub=NULL,lang=c("en","fr")){
    if (is.null(modele)){
       modele=colnames(data)
       modele=modele[modele!=Y]   
@@ -18,12 +21,24 @@ recursive_tree<-function(data=data,Y="Y",modele=NULL,kill=NULL,index=NULL,print=
    }
    
    formule=as.formula(paste(Y," ~",paste(modele,collapse="+")))
-   arbre=rpart(formule,data);par( xpd=NA)
+   arbre=rpart(formule,data)
    if(plot){
-      sub=paste("oui a gauche, non a droite. Moyenne (global", round(arbre$frame$yval[1],digits=3),") et effectif (global ", arbre$frame$n[1],") \n la hauteur indique la significativite")
-      plot(arbre,sub=sub)
+      opar<-par()
+      par( xpd=NA)
+      if(is.null(sub)){
+         lang=lang[1]
+         if(lang=="fr"){
+            sub=paste("oui a gauche, non a droite. Moyenne (global", round(arbre$frame$yval[1],digits=3),") et effectif (global ", arbre$frame$n[1],") ")
+            vertical= "la hauteur indique la significativite"        
+         }else if (lang=="en"){
+            sub=paste("True left, False right. Mean (global", round(arbre$frame$yval[1],digits=3),") and effectives (global ", arbre$frame$n[1],") ")
+            vertical=" heights indicates significativity"         
+         }
+      }
+      plot(arbre)
       text(arbre, use.n=TRUE)
-      #title(main = titre,ylab=vertical,xlab = paste(infos,"\n Oui Ã  gauche, non Ã  droite.",reponse), col.main = "red", col.lab = gray(.5),cex.main = 1.2, cex.lab = 1.0, font.main = 4, font.lab = 3
[…]

### 2014-08-05 pkg/CorReg/R/rforge.R
upgrading function
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@751 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,6 @@
+#' to upgrade a package to the lastest version on R-forge
+#' @param package the name of the package to upgrade
+#' @export
+rforge<-function(package="CorReg"){
+   install.packages(package, repos="http://R-Forge.R-project.org")
+}

### 2014-08-05 pkg/CorReg/R/rforge.R
up rforge params
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@752 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,6 +1,6 @@
 #' to upgrade a package to the lastest version on R-forge
-#' @param package the name of the package to upgrade
+#' @param package vector of the name of the packages to upgrade
 #' @export
-rforge<-function(package="CorReg"){
-   install.packages(package, repos="http://R-Forge.R-project.org")
+rforge<-function(packages=c("CorReg"),repos="http://R-Forge.R-project.org"){
+   install.packages(packages, repos=repos)
 }

### 2014-08-05 pkg/CorReg/R/rforge.R
up rforge
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@753 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,5 +1,5 @@
-#' to upgrade a package to the lastest version on R-forge
-#' @param package vector of the name of the packages to upgrade
+#' Upgrades a package to the lastest version on R-forge
+#' @param packages vector of the name of the packages to upgrade
 #' @export
 rforge<-function(packages=c("CorReg"),repos="http://R-Forge.R-project.org"){
    install.packages(packages, repos=repos)

### 2014-08-06 pkg/CorReg/R/rforge.R
up bug Terminator diag
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@755 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,6 +1,8 @@
 #' Upgrades a package to the lastest version on R-forge
-#' @param packages vector of the name of the packages to upgrade
+#' @param package name of the packages to upgrade
 #' @export
-rforge<-function(packages=c("CorReg"),repos="http://R-Forge.R-project.org"){
-   install.packages(packages, repos=repos)
+rforge<-function(package="CorReg"){
+   detach(unload=TRUE)
+   install.packages(package, repos="http://R-Forge.R-project.org")
 }
+

### 2014-08-06 pkg/CorReg/R/Terminator.R
up bug Terminator diag
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@755 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -10,6 +10,8 @@ Terminator<-function(target="Sarah Connor", wrath=0.1,diag=0,Z=NULL){
    }else if (target[1]=="bender"){
       Bender()
    }else if(diag>0){
+      n=nrow(target)
+      p=ncol(target)
       quidiag=cbind(rep(1:n,length.out=max(n,p)),rep(1:p,length.out=max(n,p)))
       target[quidiag]=NA
       for(j in 2:diag){

### 2014-08-21 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@807 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{subfigure}
\usepackage{placeins}
\usepackage[table]{xcolor}
 \graphicspath{{figures/}}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
%%\author{Clément THERY, Christophe BIERNACKI, Gaétan LORIDANT}
%\title{Model-based variable selection for regression with highly correlated variables.}

%%%% debut macro %%%%
\newenvironment{changemargin}[2]{\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{0pt}%
\setlength{\rightmargin}{0pt}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setlength{\parsep}{0pt plus 1pt}%
\addtolength{\leftmargin}{#1}%
\addtolength{\rightmargin}{#2}%
}\item }{\end{list}}
%%%% fin macro %%%%

\newtheorem{hyp}{Hypothesis}

\definecolor{darkgreen}{rgb}{0,0.4,0}
	 \definecolor{darkred}{rgb}{0.75,0,0}
	 \definecolor{darkblue}{rgb}{0,0,0.4}

\begin{document}
\begin{center}
{\Large
	{\sc Model-Based Variable Decorrelation in Linear Regression}
}
\bigskip

  Clément Théry$^{1}$ \& Christophe Biernacki$^{2}$ \& Gaétan Loridant$^{3}$
\bigskip

{\it
$^{1}$ ArcelorMittal, Université Lille 1, CNRS, Inria, clement.thery@arcelormittal.com

$^{2}$ Université Lille 1, CNRS, Inria, christophe.biernacki@math.univ-lille1.fr

$^{3}$ Etudes Industrielles ArcelorMittal Dunkerque, gaetan.loridant@arcelormittal.com\textbf{}
}
\end{center}
\bigskip

{\bf Abstract.} Linear regression outcomes (estimates, prevision) are known to be damaged by highly correlated covariates. However most modern datasets are expected to mechanically convey more and more highly correlated covariates due to the global increase of the amount of variables they contain. We propose to explicitly model such correlations by a family of linear regressions between the covariates. The structure of correlations is found with an {\sc mcmc} algorithm aiming at optimizing a specific {\sc bic} criterion. This hierarchical-like approach leads to a joint probability distribution on both the initial response variable and the linearly explained covariates. Then, marginalisation on the linearly explained covariates produces a parsimonious correlation-free regression model from which classical procedures for estimating regression coefficient, including any variable selection procedures, can be plugged.
Both simulated and real-life datasets from steel industry, where correlated variables are frequent, highlight that this proposed covariable pretreatment-like method has two essential beneficits: First, it offers a real readability of the linear links between covariates; Second, it improves significantly efficiency of classical estimation/selection methods which are performed after.
An {\sc r} package (\textsc{CorReg}), available on the {\sc cran}, implements this new method.
\smallskip

{\bf Keywords.} Regression, correlations, steel industry, variable selection, generative models, model selection

\section{Introduction}\label{sec:intro}
%la régression et ses problèmes
Linear regression is a very standard and efficient method providing a predictive model with a good interpretability even for non-statisticians. Therefore, it is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}: Astronomy \cite{isobe1990linear}, Sociology \cite{longford2012revision}, Industry (real datasets of the present paper), \dots
With the rise of informatics, datasets contain more and more covariates leading high variance estimates, misleading interpretations and poor prediction accuracy for two different, quite related, reasons. First, the number of covariates leads to more complex models. Second, the number of covariates mechanically increases the chance to have correlated ones. Both situations are quite intricate and appear technically in a similar manner in the estimate procedure through ill-conditioned matrices. As a consequence, explicit break-up between them is not very common, most proposed methods focusing on the seminal question of the number of covariates selection. Originality of this paper is to distinguish explicitly them by proposing a specific decorrelation step followed by any classical variable selection step chosen by the practitioner. The decorrelation step can be viewed as a variable selection step but it is focused only on correlations, not on the number of covariables.

\vspace{3mm}

%bibliographie

Reducing the variance induced by the large number of covariables can be reached by targeting a better bias-variance trade off. Relating methods are numerous and continue to generate a lot of work. Most traditional directions are shrinkage, including variable selection, and also variable clustering.

Ridge regression \cite{marquardt1975ridge} is a shrinkage method which proposes possibly biased estimator that can be written in terms of a parametric $L_2$ penalty. It does not select variables since coefficients tend to 0 but don't reach 0, leading to difficult interpretations for a large number of covariates.	Since real datasets may imply many irrelevant variables, variable selection should be prefered for more interpretable models. Variable selection methods may add also some bias by deleting some relevant covariates but it may reduced drastically the variance by the dimension reduction.
As an emblematic method, the Least Absolute Shrinkage and Selection Operator (\textsc{lasso} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a parametric $L_1$ penalty to shrink some coefficients exactly to zero. But, like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates. It only produces a parsimonious model, that is a gain for interpretation but only half the way from our point of view. In particular, \textsc{lasso} is also known to face consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates. So the quality of interpretation is compromised. Elastic net \cite{zou2005regularization} is a method developed to reach a compromise between Ridge regression and the \textsc{lasso} by a linear combination of $L_1$ and $L_2$ penalties. But, since it is based on the grouping effect, correlated covariates get similar coefficients and are selected together.

Another way for improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{oscar} \cite{bondell2008simultaneous}) to reduce dimension and also correlations if correlated covariates are in the same clusters. A possible bias is added by the dimension reduction inherent to the coefficients clustering.
The CLusterwise Effect REgression (\textsc{clere} \cite{yengo2012variable}) describes the regression coefficients no longer as fixed effect parameters but as unobserved independent random variables with grouped coefficients following a Gaussian Mixture distribution. The idea is that if the model has a small number of groups of covariates, the model will have a number of free parameters significantly lower than the number of covariates. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on the coefficients estimator. 
Spike and Slab variable selection \cite{ishwaran2005spike} also relies on a Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute the Ordinary Least Squares estimate (\textsc{ols}) but has no specific protection against correlations issues.

\vspace{3mm}

% Principe de la méthode
None of the above methods takes explicitly the correlations into account, even if the clustering methods may group the correlated covariates together. However, modeling explicitly linear correlation between variables already exists in statistics. In Gaussian model-based clustering, \cite{maugis2009variable} consider that some irrelevant covariates for clustering are in linear regression with some relevant ones. The algorithm used to find the structure is a stepwise-like algorithm \cite{raftery2006variable} even if it is known to be often unstable \cite{miller2002subset}. We propose to transpose this method for linear regression with a specifically adapted algorithm to find the structure of sub-regression.

The idea of the present paper is that if we know explicitly the correlations, we could use this knowledge to avoid this specific problem. Correlations are thus new information to reduce the variance without adding any bias. More precisely, correlations are modeled through a system of linear sub-regressions between covariates. The set of covariates which are {\it never} at the place of a response variable in these sub-regressions is finally the greatest set of orthogonal covariates. Marginalizing over the dependent co-variables leads then to a linear regression (in relation to the initial response variable) with only orthogonal covariates. This marginalization step can be viewed also as a variable selection step but guided only by the correlations between covariates. Advantages of this approach is twofold. First, it improves interpretation through a good readability of dependency between covariables. Second, this marginal model is still a ``true'' model provided that both the initial regression model and all the sub-regressions are ``true''. As a consequence, the associated {\sc ols} will preserve an unbiased estimate but with a possibly reduced variance comparing to the {\sc ols} with the full regression model. The fact that the variance decreases depends on the residual variances involved in the sub-regressions: The more the sub-regressions are marked, the less will be the variance of associated {\sc ols}. In fact, any other estimation method than {\sc ols} can be plugged after the marginalization step. Indeed, it can be viewed as a pretreatment against correlation which can be chained after with dimension reduction methods, without no more suffering from correlations this time.
The sub-linear structure is obtained by a {\sc mcmc} algorithm optimizing a specific {\sc bic} criterion associated to the joint distribution on the covariates, regardless of the initial response variable. This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{cran}.

 	%plan
This paper will first present in Section~\ref{sec:model} the modelisation of the correlations between covariates by sub-regressions and the by-product marginal regression model. Section 3 is devoted to describe the {\sc mcmc} random walk used to find the structure of sub-regressions.
Some numerical results on simulated datasets (Section \ref{sectionsimul}) and real industrial datasets (Section \ref{sectionrealcase}) are then conducted to quantify the added value of our approach. Concluding remarks, including perspectives, are then given in Section~\ref{conclusion}.

\section{Model-based approach for selecting uncorrelated covariates}
\label{sec:model}

\subsection{Sub-regressions between covariates}%\label{sectionOLS}
The classical linear regression model can be written
\begin{equation}
		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
	\end{equation}
with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon} \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension. The $d\times 1$ vector $\boldsymbol{\beta}$ gathers the coefficients of the regression\footnote{Usually a constant is included as one of the regressors. For example we can take $\boldsymbol{X}^1=(1,\ldots,1)'$. The corresponding element of $\boldsymbol{\beta}$ is then the intercept $\beta_1$.}, that can be estimated by Ordinary Least Squares (\textsc{ols}): %As shown in section \ref{sectionOLS}, 
	\begin{equation}
		\hat{\boldsymbol{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y} \label{eq:OLS}
	\end{equation}
	It is an unbiased estimate with variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\boldsymbol{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}. \label{eq:varOLS}
	\end{equation}
	This estimate requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which can lead to great variance estimates, so unstable estimates, if it is ill-conditioned. Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grows (in absolute value)  \cite{hoerl1970ridge}. At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular. Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariables higher.

\vspace{3mm}

We focus now on an original manner to solve the covariates correlation problem. The covariates number problem will be solved at a second stage by standard methods, once only decorrelated covariates will be identified. The proposed method relies on the two following hypotheses.

\begin{hyp}\label{H1}
In order to take into account the covariates correlation problem, we make the hypothesis correlation between covariates is {\it only} the consequence that some covariates {\it linearly} depend on some other covariates. More precisely, there are $d_{r}\geq 0$ such ``sub-regressions'', each sub-regression $j=1,\ldots,d_{r}$ having the covariate $\boldsymbol{X}^{J_{r}^j}$ as {\it response} variable ($J_{r}^j\in\{1,\ldots,p\}$ and $J_{r}^j\neq J_{r}^{j'}$ if $j\neq j'$) and having the $d_p^j>0$ covariates $\boldsymbol{X}^{J_{p}^j}$  as {\it predictor} variables ($J_{p}^j\subset\{1,\ldots,d\} \backslash J_{r}^j$ and $d_p^j=|J_{p}^j|$ the cardinal of $J_{p}^j$):
\begin{equation}
\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j};\boldsymbol{\alpha}_j,\sigma^2_j=\boldsymbol{X}^{J_{p}^j}\boldsymbol{\alpha}_j+\boldsymbol{\varepsilon}_j, \label{eq:SR}
\end{equation}
where $\boldsymbol{\alpha}_j\in{\mathbb{R}^{d_r^j}}$ (${\alpha}_j^h\neq 0$ for all $j=1,\ldots,d_r$ and $h=1,\ldots,d_p^j$) and $\boldsymbol{\varepsilon}_j \sim\mathcal{N}_n(\boldsymbol{0},\sigma^2_j\boldsymbol{I})$.
\end{hyp}

\begin{hyp}\label{H2}
In addition, we make the complementary hypothesis that the responses covariables and the predictor covariables are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariables and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates. This new assumption allows to obtain very simple sub-regressions sequences, discarding hierarchical ones, in particular uninteresting cyclic sub-regressions. However it is not too much restrictive since any hierarchical (but non-cyclic) sequence of sub-regressions can be agglomerated into a non-hierarchical sequence of sub-regressions, even if it may implies to partially loose information through variance increase in the new non-hierarchical sub-regressions.
\end{hyp}

\paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions. As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariables in $\boldsymbol{J}_p$ or not. An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}. The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\boldsymbol{\sigma}^2_1,\ldots,\boldsymbol{\sigma}^2_{d_r})$ denotes the corresponding global variance.

\paragraph{Remarks}
\begin{itemize}
\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
\item As a consequence of Hypothesis~\ref{H1}, ``free'' covariables $\boldsymbol{X}_f$ are {\it all} decorrelated.
\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals. Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}. 
\end{itemize}

\subsection{Marginal regression with decorrelated covariates}
The aim is now to use the model of linear sub-regressions $\boldsymbol{S}$ (that we assume to be known in this part) between some covariates of $\boldsymbol{X}$ to obtain a linear regression on $\boldsymbol{Y}$ relying on only uncorrelated variables $\boldsymbol{X}_f$.  The way to proceed is to marginalize the joint distribution of $\{(\boldsymbol{Y},\boldsymbol{X}_f) |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ to obtain the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ depending only on uncorrelated variables $\boldsymbol{X}_f$:
\begin{equation}\label{eq:marginal}
\mathbb{P}(\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}) = \int_{{\mathbb{R}^{d_r}}}\mathbb{P}(\boldsymbol{Y}| \boldsymbol{X}_f,\boldsymbol{X}_r,\boldsymbol{S};\boldsymbol{\beta},\sigma_Y^2) \mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma^2}) d\boldsymbol{X}_r.
\end{equation}
We need the following new hypothesis.

\begin{hyp}\label{H3}
We assume that all errors $\boldsymbol{\varepsilon}$ and $\boldsymbol{\varepsilon}_j$ ($j=1,\ldots,d_r$) are {\it mutually independent}. It implies in particular that conditional response covariates $\{\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j\}$, with distribution defined in (\ref{eq:SR}), are {\it mutually independent}:
\begin{equation}\label{eq:H3}
\mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma^2}) = \prod_{j=1}^{d_r} \mathbb{P}(\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j).
\end{equation}
\end{hyp}

\vspace{3mm}

Decomposing $\boldsymbol{\beta}=(\boldsymbol{\beta}_r',\boldsymbol{\beta}_f')'$ by $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the response and to the free covariates, we can rewrite (\ref{regressionsimple}):
\begin{equation}
			\boldsymbol{Y}{|\boldsymbol{X},\boldsymbol{S}};\boldsymbol{\beta},\sigma_Y^2=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon}_Y. \label{eq:MainR}
\end{equation}
Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$:
\begin{eqnarray}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl}
\end{eqnarray}
Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariables $\boldsymbol{X}_f$. This decorrelation process has also acted like a specific covariable selection process-like because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$. These two statements are expected to decrease the variance of further estimates of $\boldsymbol{\beta}$. However, the counterpart is twofold. First, this regression has a higher residual variance than the initial one since it is now $\sigma^{2*}_Y=\sigma^2_Y+\sum_{j \in I_r}\beta_{j}^2\sigma^2_j$ instead of $\sigma^2_Y$. Second, covariable selection being equivalent to set $\hat{\boldsymbol{\beta}}_r=\boldsymbol{0}$, it implies possibly biased estimates of $\boldsymbol{\beta}_r$. As a conclusion, we are faced to a typical {\it bias-variance trade off}. We will illustrate it in the next section in the case of the {\sc ols} estimate.

\vspace{3mm}

In practice, the strategy we propose is to rely estimate of  $\hat{\boldsymbol{\beta}}$ upon Equation~(\ref{eq:modexpl}). The practitioner can choose any estimate of its choice, like {\sc ols} or any variable selection procedure like {\sc lasso}. In other words, it is possible to see $(\ref{eq:modexpl})$ as a kind of {\it pretreatment} for decorrelating covariates, while assuming nothing on the subsequent estimate process.

In the following, we will denote by {\sc CorReg} the new proposed strategy.

	\subsection{Illustration of the bias-variance trade off with {\sc ols}}	\label{sec:tradeoff}

\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*=(\boldsymbol{\beta}_r^{*'},\boldsymbol{\beta}_f^{*'})'$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_r^*={\bf 0}$, (\ref{eq:modexpl}) can be rewritten
\begin{equation}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
\end{equation}
The {\sc ols} estimate of $\boldsymbol{\beta}^*$ is then given by
	\begin{equation}
		\hat{\boldsymbol{\beta}}^*_f=\left(\boldsymbol{X}_f'\boldsymbol{X}_f \right) ^{-1}\boldsymbol{X}_f'\boldsymbol{Y} \quad \mbox{and} \quad \hat{\boldsymbol{\beta}}^*_r = {\bf 0}.
	\end{equation}
As usual, the {\sc ols} estimate $\hat{\boldsymbol{\beta}}^*$ of $\boldsymbol{\beta}^*$ is unbiased but, however, contrary to $\hat{\boldsymbol{\beta}}$, it could be a biased estimate of $\boldsymbol{\beta}$ since
		\begin{equation}
			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
		\end{equation}
In return, its variance could be reduced compared to this one of $\hat{\boldsymbol{\beta}}$ given in (\ref{eq:varOLS}) as soon as values of $\sigma_j$ are small enough (it means strong correlations in sub-regressions) as we can see in the following expression
		\begin{equation}
			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}. \label{eq:varOLS*}
		\end{equation}
Indeed, no correlations between covariables $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditionned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}). This bias-variance trade off can be resumed by the Mean Squared Error (\textsc{mse}) associated to both estimates:
	\begin{eqnarray}
			\textsc{mse}(\hat{\boldsymbol{\beta}})&=& \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}),
			 \\
			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
	\end{eqnarray}

%\vspace{3mm}

\paragraph{Numerical illustration} We now illustrate the bias-variance trade off, through a numerical example. Let a simple case with $d=5$ covariables $\boldsymbol{X}=(\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{X}^4, \boldsymbol{X}^5)$ independently drawn by four independent Gaussian $\mathcal{N}_n(\boldsymbol{0},\boldsymbol{I})$. Thus, $J_f=\{1,2,4,5\}$ and $\boldsymbol{X}_f=(\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5)$. Let also $d_r=1$ response covariate $\boldsymbol{X}^3|\boldsymbol{X}^1,\boldsymbol{X}^2=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{\varepsilon}_1$ where $\boldsymbol{\varepsilon}_1\sim{\mathcal{N}_n(\boldsymbol{0},\sigma_1^2\boldsymbol{I})}$. Thus, $\boldsymbol{\alpha}_1=(1,1)'$, $\boldsymbol{J}_r=(3)$, $J_r=\{3\}$, $\boldsymbol{X}_r=(\boldsymbol{X}^3)$, $\boldsymbol{d}_p=(2)$, $\boldsymbol{J}_p=(\{1,2\})$, $\boldsymbol{X}^{\boldsymbol{J}_p^1}=(\boldsymbol{X}^1,\boldsymbol{X}^2)$ and $\boldsymbol{S}=((3),(\{1,2\}))$. Concerning now the regression with $\boldsymbol{Y}$, we define $\boldsymbol{Y}|\boldsymbol{X}=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{X}^3+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_Y$, where $\boldsymbol{\beta}=(1,1,1,1,1)'$ and $\sigma_Y \in \{10,20\}$. Finally, we deduce that $\boldsymbol{Y}|\boldsymbol{X}_f,\boldsymbol{S}=2\boldsymbol{X}^1+2\boldsymbol{X}^2+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_1+\boldsymbol{\varepsilon}_Y$
It is clear that $\boldsymbol{X}'\boldsymbol{X}$ will become more ill-conditioned as $\sigma_1$ gets smaller.

We compute now the theoretical {\sc mse} of the {\sc ols} estimates $\hat{\boldsymbol{\beta}}$ and $\hat{\boldsymbol{\beta}}^*$ for several values of $\sigma_1$ (strength of the sub-regression) and the sample size $n$. Figure~\ref{MQE1} displays the {\sc mse} evolution with the strength of the sub-regression expressed by a function of the standard coefficient of determination
	\begin{equation}
		1-R^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon}_1)}{\operatorname{Var}(\boldsymbol{X}_3)}=\frac{\sigma_1^2}{\sigma_1^2+2}.
	\end{equation}
The lower is the value of $1-R^2$, the larger is the strength of the sub-regression.

\begin{figure}[h!]
	\includegraphics[width=500px]{figures/MQEexplOLSp5.png}\label{MQE1}
	\caption{Values of $\textsc{mse}(\hat{\boldsymbol{\beta}})$ (plain) and of $\textsc{mse}(\hat{\boldsymbol{\beta}}^*)$ (dotted) when varying the strength $(1-R^2)$ of the sub-regression, and also the values $n$ and $\sigma_Y$.}
\end{figure}
It appears that, when the sub-regression is strong (low $1-R^2$), $\hat{\boldsymbol{\beta}}^*$ is a better estimate than $\hat{\boldsymbol{\beta}}$: the gain on {\sc mse} can even be very significant. This effect amplified by the $\sigma_Y$ increase but is reduced by the $n$ increases. Thus, the estimate $\hat{\boldsymbol{\beta}}^*$ should be particularly useful when some covariates are highly correlated, when also the sample sizes is small and when the residual variance of $\boldsymbol{Y}$ is large. It corresponds to expected difficult practical situations.

Further results will be provided in Section~\ref{sectionsimul} and~\ref{sectionrealcase}.

\section{Sub-regressions model selection}	\label{sec:estimateS}

The question we address now is twofold: which criterion to retain for selecting a sub-regression structure $\boldsymbol{S}$ and which strategy to adopt for exploring the large space of models $\boldsymbol{S}$. Obviously, $\boldsymbol{S}$ being very simply understood even by non statisticians, practitioners could easily transform their expert knowledge on the phenomenon at hand, if any, into a given structure $\boldsymbol{S}$. For instance, Structural Equations Models (\textsc{sem}), which are related to our model as already mentioned in Section~\ref{sec:intro}, are often used in social sciences and economy where a structure $\boldsymbol{S}$ is generally ``hand-made''. However, in the general situation, $\boldsymbol{S}$ has to estimated.  A standard method as graphical {\sc lasso} \cite{friedman2008sparse}, which searchs for a structure on the precision matrix (inverse of the variance-covariance matrix) by setting some coefficients of the precision matrix to zero, can not be applied since it is not designed to estimate oriented structures like $\boldsymbol{S}$.

It is important to note that selecting $\boldsymbol{S}$ only relies on $\boldsymbol{X}$, not on $\boldsymbol{Y}$.

\subsection{Designing two specific {\sc bic} criteria}

The Bayesian model selection paradigm consists of retaining the model $\boldsymbol{S}$ maximizing the posterior distribution \cite{raftery1995bayesian,andrieu1999joint,chipman2001practical}
\begin{eqnarray}
 \mathbb{P}(\boldsymbol{S}|\boldsymbol{X})&\propto & \mathbb{P}(\boldsymbol{X}|\boldsymbol{S})\mathbb{P}(\boldsymbol{S}) \\
	&=&\mathbb{P}(\boldsymbol{X}_r|\boldsymbol{X}_f,\boldsymbol{S})\mathbb{P}(\boldsymbol{X}_f|\boldsymbol{S})\mathbb{P}(\boldsymbol{S}).
\end{eqnarray}
In order to implement this paradigm, we need first to define the three probabilities which are in the right hand of the previous equation.

\paragraph{Defining $\mathbb{P}(\boldsymbol{X}_r|\boldsymbol{X}_f,\boldsymbol{S})$} corresponds to the integrated likelihood based on $\mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma^2})$. It can be approximated by a {\sc bic}-like approach \cite{Sch1978}
\begin{equation}
-2\ln \mathbb{P}(\boldsymbol{X}_r|\boldsymbol{X}_f,\boldsymbol{S}) \approx -2\ln \mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\hat{\boldsymbol{\alpha}},\hat{\boldsymbol{\sigma}}^2) + (|{\boldsymbol{\alpha}}| + |{\boldsymbol{\sigma}}^2|) \ln(n) = \mbox{{\sc bic}}_r(\boldsymbol{S}),
\end{equation}
where $\hat{\boldsymbol{\alpha}}$ and $\hat{\boldsymbol{\sigma}}^2$ designate respectively the Maximum Likelihood Estimates ({\sc mle}) of ${\boldsymbol{\alpha}}$ and ${\boldsymbol{\sigma}}^2$, and $|\boldsymbol{\psi}|$ designates the number of free continuous parameters associated to the space of any parameter $\boldsymbol{\psi}$.

\paragraph{Defining $\mathbb{P}(\boldsymbol{X}_f|\boldsymbol{S})$} It corresponds to the integrated likelihood based on a not yet defined distribution $\mathbb{P}(\boldsymbol{X}_f|\boldsymbol{S};\boldsymbol{\theta})$ on the uncorrelated covariates $\boldsymbol{X}_f$ and parameterized by $\boldsymbol{\theta}$. In this purpose, we need the following new hypothesis.

\begin{hyp}\label{H4}
All variates $\boldsymbol{X}^j\in\boldsymbol{X}_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
\begin{equation}
\mathbb{P}(\boldsymbol{X}_f^j|\boldsymbol{S};\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j) = \sum_{h=1}^{k_j} \pi_{hj} \mathcal{N}_n(\mu_{hj} . (1,\ldots,1)',\lambda_{hj}^2\boldsymbol{I}),
\end{equation}
where $\boldsymbol{\pi}_{j}=(\pi_{1j},\ldots,\pi_{k_jj})$ is the vector of mixing proportions with $\pi_{hj}>0$ and $\sum_{h=1}^{k_j}\pi_{hj}=1$, $\boldsymbol{\mu}_j=(\mu_{1j},\ldots,\mu_{k_jj})$  is the vector of centers and $\boldsymbol{\lambda}^2_j=(\lambda^2_{1j},\ldots,\lambda^2_{k_jj})$ is the vector of variances. We stack together all the mixture parameters in $\boldsymbol{\theta}=(\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j ; j \in J_f)$.
\end{hyp}

\vspace{3mm}

Noting $\hat{\boldsymbol{\theta}}$ the {\sc mle} of $\boldsymbol{\theta}$, the {\sc bic} approximation can then be used again:
\begin{equation}
-2\ln \mathbb{P}(\boldsymbol{X}_f|\boldsymbol{S}) \approx -2\ln \mathbb{P}(\boldsymbol{X}_f|\boldsymbol{S};\hat{\boldsymbol{\theta}}) + |{\boldsymbol{\theta}}| \ln(n) = \mbox{{\sc bic}}_f(\boldsymbol{S}).
\end{equation}

\paragraph{Defining $\mathbb{P}(\boldsymbol{S})$} The most standard choice consists of putting a uniform distribution on the model space of $\boldsymbol{S}$, this choice being noted $\mathbb{P}_U(\boldsymbol{S}) = |\boldsymbol{S}|^{-1}$, with $|\boldsymbol{S}|$ the space dimension of $\boldsymbol{S}$.

However, $\boldsymbol{S}$ being combinatorial, $|\boldsymbol{S}|$ is huge. It has two cumulated consequences: First, the {\it exact} probability $\mathbb{P}(\boldsymbol{S}|\boldsymbol{X})$ may be of the same order of magnitude for a large number of candidates  $\boldsymbol{S}$, including the best one; Second, the {\sc bic} {\it approximations} of this quantity may introduce additional confusion to wisely distinguish between model probabilities. In order to limit this problem, we propose to introduce some information in $\mathbb{P}(\boldsymbol{S})$ promoting simple models through the following {\it hierarchical} uniform distribution denoted by $\mathbb{P}_H(\boldsymbol{S})$:
\begin{eqnarray}
\mathbb{P}_H(\boldsymbol{S}) & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p) \\
 & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p,d_r,\boldsymbol{d}_p) \\
 & = & \mathbb{P}_U(\boldsymbol{J}_p|\boldsymbol{d}_p,\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{d}_p|\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{J}_r|d_r)\times \mathbb{P}_U(d_r) \\
 & = & \left[\prod_{j\in \boldsymbol{J}_r} \left(\begin{array}{c} d-d_r \\ d_p^j \end{array}\right) \right]^{-1} \times \left[d-d_r\right]^{-d_r} \times \left[\left(\begin{array}{c} d \\ d_r \end{array} \right)\right]^{-1} \times [d+1]^{-1},
\end{eqnarray}
where $\left(\begin{array}{c} a \\ b \end{array}\right)$ designs combination of $a$ things taken $b$ at a time without repetition and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand. $\mathbb{P}_H(\boldsymbol{S})$ gives decreasing probabilities to more complex models, provided that the following new hypothesis is verified:

\begin{hyp}\label{H5}
We set $d_r<d/2$ and also $d_p^j<d/2$ ($j=1,\ldots,d_r$).
\end{hyp}

These two thresholds are sufficiently large to be wholly realistic.

\paragraph{Final approximation of $\mathbb{P}(\boldsymbol{S}|\boldsymbol{X})$}
Merging the previous three expressions, it leads to the following two {\it global} {\sc bic} criteria, to be minimized, denoted by {\sc bic}$_U$ or {\sc bic}$_H$, depending on the choice of $\mathbb{P}_U(\boldsymbol{S})$ or $\mathbb{P}_H(\boldsymbol{S})$ respectively:
\begin{eqnarray}
\mbox{{\sc bic}}_U(\boldsymbol{S}) & = & \mbox{{\sc bic}}_r(\boldsymbol{S}) + \mbox{{\sc bic}}_f(\boldsymbol{S}) - 2\ln \mathbb{P}_U(\boldsymbol{S}) \\
\mbox{{\sc bic}}_H(\boldsymbol{S}) & = & \mbox{{\sc bic}}_r(\boldsymbol{S}) + \mbox{{\sc bic}}_f(\boldsymbol{S}) - 2\ln \mathbb{P}_H(\boldsymbol{S}).
\end{eqnarray}
In the following, we will denote by $\mbox{{\sc bic}}_*$ any of both $\mbox{{\sc bic}}_U$ and $\mbox{{\sc bic}}_H$. Numerical results in Section~\ref{sec:simS} will allow to compare behaviour of both criteria.

\paragraph{Remarks}
\begin{itemize}
\item Hypothesis~\ref{H4} is the keystone to define a full generative model on the whole covariables $\boldsymbol{X}$. On the one hand, the {\sc bic} criterion can be applied in this context, avoiding to use a cross-validation criterio which can be much more time-consuming. On the other hand, the great flexibility  of Gaussian mixture models \cite{mclachlan2004finite}, provided that the number of components $k_j$ has to be estimated, implies that Hypothesis~\ref{H4} is particularly weak in fact.
\item In practice, Gaussian mixture models are estimated only once for each variable $\boldsymbol{X}^j$ ($j=1,\ldots,d$). Thus, there is no combinatorial difficulty associated with them. An {\sc em} algorithm \cite{dempster1977maximum} will be used for estimating the mixture parameters and a classical {\sc bic} criterion \cite{Sch1978} will be used for selecting the different number of components $k_j$.
\item A sufficient condition for identifiability of the structure of sub-regressions $\boldsymbol{S}$ is that all sub-regressions contain at least two predictor covariates ($d_p^j\geq 2$ for all $j=1,\ldots,d_r$). In fact, if there exists some sub-regressions with only one regressor, identifiability holds for these sub-regressions only up to a permutation between the related response and predictor covariates. However, even in this case, full identifiability may occur thank's to constraints on response and predictor covariates given in Hypothesis~\ref{H1} and~\ref{H2}.
\item As any {\sc bic} criterion, the {\sc bic}$_U$ and {\sc bic}$_H$ criteria are consistent \cite{BIChuard}.
\item Even if it favors more parsimonious models, $\mathbb{P}_H(\boldsymbol{S})$ can be also viewed as a poor informative prior on $\boldsymbol{S}$ since it is a combination of non informative priors.
\end{itemize}

\subsection{Exploring the structure space with an {\sc mcmc} algorithm}
We present now an {\sc mcmc} algorithm devoting to minimize the criterion $\mbox{{\sc bic}}_*$ to find the optimal estimate of the structure $\boldsymbol{S}$. This Markov chain is regular and ergodic with a finite state space, thus it has a stationary distribution $\pi\propto\exp(-\mbox{{\sc bic}}_*)$ on the space $\mathcal{S}$ of $\boldsymbol{S}$ \cite{grinstead1997introduction}. Consequently, the chain is expected to be more concentrated around the mode of $\pi$, where the optimal value of $\boldsymbol{S}$ stands.

This algorithm alternates two steps: the definition of a neighbourhood $\mathcal{V}(\boldsymbol{S})$ around the current structure $\boldsymbol{S}$ distribution and then the generation of a new structure $\boldsymbol{S}^+$ belonging to this neighbourhood according to its posterior probability.

Note that the $\mathcal{S}$ has to be a {\it regular} space. It means that it has to verify Hypothesis~1, 2 and~5. In addition, we note below $\boldsymbol{S}$ and $\boldsymbol{S}^+$ the structures at the current and the next iteration of the algorithm, respectively.

	\subsubsection{Definition of a neighbourhood $\mathcal{V}(\boldsymbol{S})$}

We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ describe below:
	\begin{itemize}
	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariable is added to $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{\boldsymbol{J}}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
\end{equation}	
		\item {\bf Removing a sub-regression}: a sub-regression with only one predictor covariable is removed from $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})=(\boldsymbol{J}_r,\boldsymbol{J}_p)^{\{1,\ldots,d_r\}\backslash j}, j\in\{1,\ldots,d_r\} \right\}.
\end{equation}	
	\item {\bf Adding a predictor covariable}: a predictor covariable is added to one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{\boldsymbol{J}}_p^{j}=\tilde{\boldsymbol{J}}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}
	\item {\bf Removing a predictor covariable}: a predictor covariable is removed from one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{\boldsymbol{J}}_p^{j}=\tilde{\boldsymbol{J}}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}	
	\end{itemize}

	\subsubsection{Generation of a new structure $\boldsymbol{S}^+$}

We then generate $\boldsymbol{S}^+$ from the following transition probability defined in  $\mathcal{V}(\boldsymbol{S})$
	\begin{equation}
	\boldsymbol{S}^+ \sim \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
	\end{equation}

 \subsubsection{Detailed use of the algorithm}

\begin{itemize}
\item {\bf Estimate $\hat{\boldsymbol{S}}$}: we retain the structure $\hat{\boldsymbol{S}}$ having the lowest value of the criterion $\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}})$ during the walk.
\item {\bf Initialization}: the initial structure is randomly from a distribution taking into account the absolute value of the correlations.
\item {\bf Long versus short runs}: we prefer to run multiple short chains than to run a unique long chains \cite{gilks1996markov}.
\end{itemize}

\section{Numerical results on simulated datasets} \label{sectionsimul}

We now aim to access the numerical behaviour of the proposed strategy {\sc CorReg}, and its related estimation and model selection processes, through some simulated datasets. We will also evaluate robustness of the method in some disadvantageous situations as non linear correlations between covariates.

	\subsection{Experimental design}

We consider regressions on $\boldsymbol{Y}$ with $d=40$ covariates and with a $R^2$ value equal to $0.4$. Sub-regressions will have $R^2$ successively set to $(0.1,0.3,0.5,0.7,0.99)$. Variables in $\boldsymbol{X}_f$ arise from a Gaussian mixture model whose the number of components follows a Poisson's law of mean parameter equal to $5$. The coefficients of $\boldsymbol{\beta}$ and of the $\boldsymbol{\alpha}_j$'s are independently generated according to the same Poisson distribution but with a uniform random sign. All sub-regressions are of length two ($\forall j=1,\ldots,d_r, d_p^j=2$ and we have $d_r=16$ sub-regressions. The datasets are then scaled, so that covariates $\boldsymbol{X}_f$ for avoiding large distorsions for variances or for means.	Different sample sizes  $n\in (30,50,100,400)$ are chosen, thus considering experiments in both situations $n<p$ and $n>p$. In all figures, the thickness of the lines will represent various values of $n$, the thicker being the greater.

We used \textsc{Rmixmod} \cite{remi:hal-00919486} to estimate the Gaussian mixture densities of each covariate $\boldsymbol{X}$, settings being using at default values both for parameter and the number of components estimation. For each configuration, the {\sc mcmc} walk was launched on $10$ initial structures with 1~000 iterations each time. All the results are provided by the {\sc CorReg} package available on the {\sc cran}\footnote{http://cran.r-project.org/web/packages/CorReg/index.html}.

\vspace{3mm}

In the following, Section~\ref{sec:simS} evaluates the quality of the  procedures to estimate the structure $\hat{\boldsymbol{S}}$. Section~\ref{reshatY} and~\ref{resrobust} compare predictive performance of standard methods with/without \textsc{CorReg} in some ``standard'' and ``robustness'' cases, respectively.

		\subsection{Evaluation of $\mbox{\sc bic}_*$ to estimate $\boldsymbol{S}$}\label{sec:simS}

To evaluate the quality of the estimated structure $\hat{\boldsymbol{S}}$ structure retained both by $\mbox{\sc bic}_U$ and $\mbox{\sc bic}_H$ we define two complementary indicators to compare it to the true model $\boldsymbol{S}$.
			\begin{itemize}
				\item $T_r=|J_r \cap \hat{J}_r|$ (``True Response''): it corresponds to the number of estimate response covariables in $\hat{\boldsymbol{S}}$ which are {\it truly} response covariables in the true model $\boldsymbol{S}$.
				\item $W_r=|\hat{J}_r|-T_r$ (``Wrong Response''): it corresponds to the number of estimate response covariables in $\hat{\boldsymbol{S}}$ which are are {\it wrongfully} response covariables in the true model $\boldsymbol{S}$.
			\end{itemize}
The $T_r$ and $W_r$ quality values are displayed in Figure~\ref{reshatZ}(a) and~(b) for the $\mbox{\sc bic}_U$ and $\mbox{\sc bic}_H$ respectively. We observe that $\mbox{\sc bic}_H$ provides notably less wrong sub-regressions than $\mbox{\sc bic}_U$ for any strength $R^2$ of the true sub-regressions. In addition, $\mbox{\sc bic}_H$ need to have significantly strong sub-regressions ($R^2>0.4$) to detect them. However, this is not a problem since our {\sc CorReg} strategy is expected to involve only quite strong correlated covariables to have potential interest. Finally, we keep now the $\mbox{\sc bic}_H$ criterion as the best one for our purpose.

\begin{figure}[h!]
	\subfigure[$\mbox{\sc bic}_U$ criterion]{
			\includegraphics[height=180px,width=245px]{figures/res_article/BIC_p2.png} 
	} \quad
   	\subfigure[$\mbox{\sc bic}_H$ criterion]{
			\includegraphics[height=180px,width=240px]{figures/res_article/BICSTAR_P2.png} 
	}
	\caption{Average quality of the estimate subregressions $\hat{\boldsymbol{S}}$ (dotted for $T_r$, plain for $W_r$) obtained by (a) $\mbox{\sc bic}_U$ and (b) $\mbox{\sc bic}_H$. Thickness represents $n=30,50,100,400$. Inter-quartile intervals are also displayed.}\label{reshatZ}
\end{figure}

\FloatBarrier
\subsection{Evaluation of the prediction quality}\label{reshatY}

To evaluate the prediction quality of {\sc CorReg} as a pretreatment method, we consider three {\it scenarii}: First, the response variable $\boldsymbol{Y}$ depends on all covariates $\boldsymbol{X}$; Second, $\boldsymbol{Y}$ depends on only covariates $\boldsymbol{X}_f$; Finally, $\boldsymbol{Y}$ depends on only covariates $\boldsymbol{X}_r$. It correspond respectively to a {\it neutral} situation for {\sc CorReg}, a {\it favorable} one and an {\it unfavorable} one. %The true structure $\boldsymbol{S}}$ is assumed to be known through all this experiments.
The $\mbox{\sc bic}_H$ criterion is always used for the model selection.

\subsubsection{$\boldsymbol{Y}$ depends on all $\boldsymbol{X}$}

We compare different standard estimation methods ({\sc ols}, {\sc lasso}, ridge, stepwise) with and without {\sc CorReg} as a pre-treatment. When $n<p$, the {\sc ols} method is associated as usual with the Moore-Penrose generalized inverse \cite{katsikis2008fast}. In addition, when using penalized estimators for variable selection like {\sc lasso}, an {\sc ols} step is used for coefficient estimation after shrinkage for better estimation \cite{SAM10088}.	
When {\sc CorReg} is combined with a standard estimation method, {\sc ols} for instance, the whole method will be simply noted {\sc CorReg}+{\sc ols}. Results will be evaluated though a predictive Mean Squared Error value ({\sc mse}) on a validation sample of 1~000 individuals and also by the complexity of the regression in $\boldsymbol{Y}$ ({\it i.e.} its number of variables). Associated figures will display both mean and inter-quartile intervals of this predictive {\sc mse}.

\paragraph{Comparison 1} We compare {\sc ols} with {\sc CorReg}+{\sc ols}.
Figure~\ref{toutOLS} (a) shows that {\sc CorReg} improves significantly the prediction power of {\sc ols} for small values of $n$ and/or heavy sub-regression structures. This advantage then shrinks when $n$ increases because the matrix to invert becomes better-conditioned and since {\sc CorReg} does not allow to retrieve that $\boldsymbol{Y}$ depends on all $\boldsymbol{X}$ because of the marginalization of some covariates implicated in the sub-regressions. Figure~\ref{toutOLS}~(b) illustrates also the model regression in $\boldsymbol{Y}$ retained by {\sc CorReg} is more parsimonious, provided that sub-regressions are strong enough.

\begin{figure}[!ht]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_tout_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_tout_compl_NB.png} 
	}
	\caption{Comparison of {\sc ols} (dotted) and {\sc CorReg} + {\sc ols} (plain) when $\boldsymbol{Y}$ depends on all covariates ${\boldsymbol{X}}$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{toutOLS}
\end{figure}

\paragraph{Comparisons 2 to 4} We compare now three variable selections methods: {\sc lasso} with {\sc CorReg}+{\sc lasso}, elasticnet with {\sc CorReg}+electicnet and also stepwise with {\sc CorReg}+stepwise. Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} respectively display the results on the same manner as the previous {\sc ols} with {\sc CorReg}+{\sc ols}. We see that {\sc CorReg}, used as a pre-treatment, provides similar prediction accurary as the three variable selection methods (this prediction is often better for small datasets) but with much more parsimonious regression models on $\boldsymbol{Y}$. This is remarkable because the true model depends on all covariates, so it highlight that other variable selection methods may be really penalized by the correlations between the covariates.

\begin{figure}[h!]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/lar_tout_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/lar_tout_compl_NB.png} 
	}
	\caption{Comparison of {\sc lasso} (dotted) and {\sc CorReg} + {\sc lasso} (plain) when $\boldsymbol{Y}$ depends on all covariates ${\boldsymbol{X}}$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{toutlar}
\end{figure}

\begin{figure}[h!]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/elasticnet_tout_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/elasticnet_tout_compl_NB.png} 
	}
	\caption{Comparison of elasticnet (dotted) and {\sc CorReg} + elasticnet (plain)  when $\boldsymbol{Y}$ depends on all covariates ${\boldsymbol{X}}$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{toutelast}
\end{figure}

\begin{figure}[h!]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/stepwise_tout_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/stepwise_tout_compl_NB.png} 
	}
	\caption{Comparison of stepwise (dotted) and {\sc CorReg} + stepwise (plain)  when $\boldsymbol{Y}$ depends on all covariates ${\boldsymbol{X}}$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{toutstepwise}
\end{figure}

\paragraph{Comparison 5} We compare now ridge with {\sc CorReg}+ridge. Figure~\ref{toutridge} shows that the ridge regression is efficient in prediction when confronted to correlated covariates. It is directly made to improve the conditioning and keeps all the covariates (as the true model) so it logically gives better predictions than {\sc CorReg} which removes some covariates. Nevertheless, we notice that the {\sc CorReg} pre-treatment before the ridge regression gives {\sc mse} values that are really close to the ridge regression ones (inter-quartile intervals are very mingled) but with drastically more parsimonious models. Indeed, ridge regression alone gives a full model (40 covariates even with only 30 individuals) and will give models too complex to be easily interpreted. Thus, the combination {\sc CorReg}+ridge provideshigh benefits for interpretation whle preserving good prediction accuracy.

 \begin{figure}[h!]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/ridge_tout_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/ridge_tout_compl_NB.png} 
	}
	\caption{Comparison of ridge (dotted) and {\sc CorReg} + ridge (plain)  when $\boldsymbol{Y}$ depends on all covariates ${\boldsymbol{X}}$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{toutridge}
\end{figure}

%\FloatBarrier

\subsubsection{$\boldsymbol{Y}$ depends only on $\boldsymbol{X}_f$}	 \label{tableMSEsimdroit}

Figures~\ref{X1OLS} and~\ref{X1lasso} display results for, respectively, {\sc ols} with {\sc CorReg}+{\sc ols} and {\sc lasso} with {\sc CorReg}+{\sc lasso}. Even if the true model involved correlated but irrelevant covariates then classical variable selection methods gain to be associated with a pre-treatment by {\sc CorReg}. Figure~\ref{X1ridge} displays results for {\sc ridge} with {\sc CorReg}+{\sc ridge}. It shows the accuracy prediction of ridge regression is heavily penalized since the true model is parsimonious. In that case, {\sc CorReg} can significantly help to improve the results.

 \begin{figure}[h]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X1_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
	}
	\caption{Comparison of {\sc ols} (dotted) and {\sc CorReg} + {\sc ols} (plain) when $\boldsymbol{Y}$ depends only on $\boldsymbol{X}_f$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{X1OLS}
\end{figure}

\begin{figure}[h]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X1_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X1_compl_NB.png} 
	}
	\caption{Comparison of {\sc lasso} (dotted) and {\sc CorReg} + {\sc lasso} (plain) when $\boldsymbol{Y}$ depends only on $\boldsymbol{X}_f$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{X1lasso}
\end{figure}

\begin{figure}[h]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/ridge_X1_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/ridge_X1_compl_NB.png} 
	}
	\caption{Comparison of ridge (dotted) and {\sc CorReg} + ridge (plain) when $\boldsymbol{Y}$ depends only on $\boldsymbol{X}_f$y. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{X1ridge}
\end{figure}

\subsubsection{$\boldsymbol{Y}$ depends only on $\boldsymbol{X}_r$}	 \label{tableMSEsimgauche}
We now try the method with a response depending only on variables in $\boldsymbol{X}_r$.
Depending only on $\boldsymbol{X}_r$ implies sparsity and impossibility to obtain the true model, even if using the true structure $\boldsymbol{S}$.

Figure~\ref{X2OLS} compares {\sc ols} with {\sc CorReg}+{\sc ols}. It reveals that \textsc{CorReg} is still better than {\sc ols} for strong correlations and limited values of $n$. When $n$ rises, the sub-regression are detected and relevant covariates are removed. As a conseuqnce, {\sc CorReg} can not improve the results and increases the {\sc mse}. However, for strong correlations, the error shrinks as the model tends to be less identifiable and switching variables is not a problem anymore even with large values of $n$.

Figure~\ref{X2LASSO} compares {\sc lasso} with {\sc CorReg}+{\sc lasso}. It shows that {\sc lasso} naturally tends to keep $\boldsymbol{X_r}$ and thus is better bevause it corresponds to the true model. So {\sc CorReg} is almost always the worst method. In such a case, it is recommended to compare both {\sc lasso} with {\sc CorReg}+{\sc lasso} with a model choice criterion. Note that this model choice is only between two models so it avoids multiple tests issues and computational cost explosion. Thus, we suggest to always compute the ``with {\sc CorReg}'' and the ``without {\sc CorReg}'' solutions and then to compare them with the more pertinent criterion (AIC, cross-validation, validation sample, {\it etc.}) according to the context (size of the datasets for example).

\paragraph{Remark}
Since the structure $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$, it can be interesting for interpretation in cases when the ``with {\sc CorReg}'' solutions are not kept in favor of ``without {\sc CorReg}'' solutions. {\sc CorReg} can then be seen both not only like a pre-treatment but also like a pre-study.

\begin{figure}[h!]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X2_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X2_compl_NB.png} 
	}
	\caption{Comparison of {\sc ols} (dotted) and {\sc CorReg} + {\sc ols} (plain) when $\boldsymbol{Y}$ depends only on $\boldsymbol{X}_r$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{X2OLS}
\end{figure}

\begin{figure}[h!]
	\subfigure[{\sc mse} on a validation subset]{
			\includegraphics[height=180px,width=245px]{figures/res_article/lar_X2_MSE_NB.png} 
	} \quad
   	\subfigure[Complexity]{
			\includegraphics[height=180px,width=240px]{figures/res_article/lar_X2_compl_NB.png} 
	}
	\caption{Comparison of {\sc lasso} (dotted) and {\sc CorReg} + {\sc lasso} (plain) when $\boldsymbol{Y}$ depends only on $\boldsymbol{X}_r$. Thickness represents $n=30,50,100,400$.  Inter-quartile intervals are also displayed.}\label{X2LASSO}
\end{figure}

\subsection{Robustess study through a non-linear case} \label{resrobust}

%\FloatBarrier

We have generated a non-linear sub-regression to test the robustess of our model. $\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as before. We design a unique, possibly non linear, sub-regression $\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2+\boldsymbol{X}_3+ \varepsilon_1$. The matrix $\boldsymbol{X}$ is then scaled and we set $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\varepsilon_Y$. We let $a$ vary between $0$ and $10$ to increase progressively the non-linear part of the sub-regression.
Figure~\ref{resnonlin} shows that the {\sc mcmc} algorithm has more difficulties to find a linear structure (by  $\mbox{{\sc bic}}_H$ as the non-linear part of the sub-regression increases with $a$. But the model is quite robust, preserving good efficient for small values of $a$.
In addition, Figure~\ref{MSEnonlin} illustrates the advantage of using {\sc CorReg}, even with non-linear sub-regressions, concerning the quality of the {\sc mse}.

 \begin{figure}[h!] 
	\begin{minipage}[l]{.48\linewidth}
			\includegraphics[height=180px,width=245px]{figures/res_article/robust_S.png} 
			\caption{Evolution of the quality of $\hat{\boldsymbol{S}}$ when the paramater $a$ increases.}\label{resnonlin}
	\end{minipage} \
   \begin{minipage}[r]{.48\linewidth}
			\includegraphics[height=180px,width=240px]{figures/res_article/robust_MSE.png} 
			\caption{{\sc mse} on the main regression for {\sc ols} (thick) and {\sc lasso} (thin) used both with (plain) or without {\sc CorReg} (dotted).}\label{MSEnonlin}
   \end{minipage}
\end{figure}

	\FloatBarrier	
\section{Numerical results on two real datasets} \label{sectionrealcase}
\subsection{Quality case study in steel industry} \label{sectionexfos}
This work takes place in steel industry context, with a quality oriented objective. Indeed, the purpose is to understand and to prevent quality problems on finished products, knowing the whole process. The correlations between involved features can be strong here because many parameters of the whole process are highly correlated (physical laws, process rules, {\it etc.}). 
We have a quality parameter (confidential) as response variable $\boldsymbol{Y}$ and $d=205$ variables from the whole process to explain it. We get a training set of $n=3\;000$ products described by these $205$ variables from the industrial process and also a validation sample of $847$ products.

The objective here is not only to predict non-quality but to understand and then to avoid it. {\sc CorReg} provides an automatic method without any {\it a priori} and can be combined with any variable selection methods. So it allows to obtain, in a small amount of time (several hours for this dataset), some indications on the source of the problem, and to use human resources efficiently. When quality crises occur, time is extremely precious so automation is a real stake. The combinatorial aspect of the sub-regression models makes it impossible to do manually.

\vspace{3mm}

To illustrate that some industrial variables are naturally highly correlated, we can measure the correlation $\rho$ between some couple of variables. For instance, the width and the weight of a steel slab gives $|\rho|=0.905$, the temperature before and after some tool gives $|\rho|=0.983$, the  roughness of both faces of the product gives $|\rho|= 0.919$ and a particular mean and a particular max gives $|\rho|=0.911$. For an overview of correlations, Figure~\ref{fig:graphCorr.quality}(a) gives an histogram of $\rho$ where we can see that, however, many other variables are not so highly correlated.

\textsc{CorReg} estimated a structure of $d_r=76$ sub-regressions with a mean of $\bar{\boldsymbol{d}}_p=5.17$ predictors. In the resulting uncorrelated covariate set $\boldsymbol{X}_f$ the number of values $|\rho|>0.7$ is $79.33\%$ smaller than in $\boldsymbol{X}$. Indeed, Figure~\ref{fig:graphCorr.quality}(b) displays the histogram of adjusted $R^2$ value ($R^2_{adj}$) and we can see that essentially large values of $R^2_{adj}$ are present. When we have a look at a more detailed level, we can see also that \textsc{CorReg} has been able non only to retrieve the above correlations (the width and the weight of a steel slab, {\it etc.}) but also to detect more complex structures describing physical models, like the width in function of the mean flow and the mean speed, even if the true physical model is not linear since ``width = flow / (speed * thickness)'' (here thickness is constant). Non-linear regulation models used to optimize the process were also found (but are confidential). These first results are easily understandable and meet metallurgists expertise.  Sub-regressions with small values of $R^2$ are associated with non-linear model (chemical kinetics for example).

\begin{figure}[h!]
\begin{center}
			\includegraphics[height=150px, width=150px]{figures/correlexfoshist.png}
			\includegraphics[height=150px,width=150px]{figures/histR2exfos.png}
\end{center}
\vspace{-5mm}
			\centerline{(a) \hspace{130px} (b)}
			\caption{Quality case study: (a) Histogram of correlations $\rho$ in $\boldsymbol{X}$, (b) histogram of the adjusted $R^2_{adj}$ for the $d_r=76$ sub-regressions.}\label{fig:graphCorr.quality}
\end{figure}

\vspace{3mm}

Note that the uncorrelated variables can be very well-modeled by parsimonious Gaussian mixtures as it is illustrated by Figure~\ref{fig:graphMixmod.quality}(a). In particular, the number of components is quite moderate as seen in Figure~\ref{fig:graphMixmod.quality}(b).

		\begin{figure}[h!]
\begin{center}
			\includegraphics[height=150px, width=150px]{figures/res_article/gaussianmixture_exfo.png}
			\includegraphics[height=150px,width=150px]{figures/res_article/nb_comp_X_exfo.png}
\end{center}
\vspace{-5mm}
			\centerline{(a) \hspace{130px} (b)}
			\caption{Quality case study: (a) Example of a non-Gaussian real variable easily modeled by a Gaussian mixture, (b) distribution of the number of components found for each covariate.}\label{fig:graphMixmod.quality}
\end{figure}

%\begin{figure}[h!]
%   \begin{minipage}[t]{.30\linewidth}
%			\includegraphics[height=150px,width=150px]{figures/correlexfoshist.png} 
%			\caption{Histogram of correlations in $\boldsymbol{X}$.} \label{compareMSEexfos}
%   \end{minipage}
%   \begin{minipage}[t]{.30\linewidth}
%			\includegraphics[height=150px, width=150px]{figures/res_article/gaussianmixture_exfo.png}%{figures/histR2exfos.png} $R^2_{adj}$ of the 76 sub-regressions.
%			\caption{Example of non-Gaussian real variable easily modeled by a Gaussian mixture.}
%	\end{minipage} \hfill
%	\begin{minipage}[t]{.30\linewidth}
%			\includegraphics[height=150px,width=150px]{figures/res_article/nb_comp_X_exfo.png}%{figures/mixmod.png} 
%			\caption{Distribution of the number of components found for each covariate.}\label{graphMixmod}
%	\end{minipage} \hfill
%\end{figure}   			
%
%	
%
%		
%		\begin{figure}[h!]
%		\centering
%			\includegraphics[height=150px, width=150px]{figures/histR2exfos.png} 
%			\caption{$R^2_{adj}$ of the 76 sub-regressions.}
%\end{figure} 
%

\vspace{3mm}

Table~\ref{Res_exfos} displays predictive results associated to different estimation methods with and without {\sc CorReg}. We can see that {\sc CorReg} improves the results for each method tested in terms of prediction, with generally a more parsimonious regression on $\boldsymbol{Y}$. In terms of interpretation, this regression gives a better understanding of the consequences of corrective actions on the whole process. It typically permits to determine the \textit{tuning parameters} whereas variable selection alone would point variables we can not directly act on.	So it becomes easier to take corrective actions on the process to reach the goal. The stakes are so important that even a little improvement leads to consequent benefits, and we do not even talk about the impact on the market shares that is even more important.

		\begin{table}[h!]
\centering
\begin{tabular}{llcc}
	\hline
	Method& Indicator& With {\sc CorReg} & Without {\sc CorReg} \\ 
	\hline \hline
	{\sc ols} & {\sc mse} & 13.30 & 14.03 \\
		& complexity& 130 & 206 \\
	\hline
	{\sc lasso} & {\sc mse} & 12.77 & 12.96 \\
		& complexity& 24 & 21 \\
	\hline
	elasticnet & {\sc mse} & \textbf{12.15} & 13.52 \\
		& complexity& 40 & 78 \\
	\hline
	ridge & {\sc mse} & 12.69 & 13.09 \\
		& complexity& 130 & 206 \\
	\hline
\end{tabular} 
\caption{Quality case study: Results obtained on a validation sample ($n=847$ individuals). In bold, the best {\sc mse} value.}\label{Res_exfos}
\end{table}

%		\FloatBarrier

		\subsection{Production case study}
This second example is about a phenomenon that impacts the productivity of a steel plant.
It is described by a (confidential)  response variable $\boldsymbol{Y}$ and $d=145$ covariables from the whole process to explain it but only $n=100$ individuals are present. The stake is to gain $20\%$ of productivity on a specific product with high added value.

Figure~\ref{fig:graphCorr.production}(a) shows that many variables are highly correlated. {\sc CorReg} found $d_r=55$ sub-regressions and corresponding $R_{adj}^2$ values are displayed in Figure~\ref{fig:graphCorr.production}(b). One of them seems to be weak ($R_{adj}^2=0.17$) but it corresponds in fact to a non-linear regression: It points out a link between diameter of a coil and some shape indicator. In this precise case, \textsc{CorReg} found a structure that helped to decorrelate covariates and to find the relevant part of the process to optimize. This product is made by a long process that requires several steel plants so it was necessary to point out the steel plant where the problem occurred.

\begin{figure}[h!]
\begin{center}
			\includegraphics[height=150px, width=150px]{figures/histcorrelBVBI.png}
			\includegraphics[height=150px,width=150px]{figures/R2corregBVBI.png}
\end{center}
\vspace{-5mm}
			\centerline{(a) \hspace{130px} (b)}
			\caption{Production case study: (a) Histogram of correlations $\rho$ in $\boldsymbol{X}$, (b) histogram of the adjusted $R^2_{adj}$ for the $d_r=55$ sub-regressions.}\label{fig:graphCorr.production}
\end{figure}

\vspace{3mm}

As in the previous quality case study, we note that the uncorrelated variables can be very well-modeled by parsimonious Gaussian mixtures as it is illustrated by Figure~\ref{fig:graphMixmod.production}(a). In particular, the number of components is really moderate as seen in Figure~\ref{fig:graphMixmod.production}(b).

\vspace{3mm}

		\begin{figure}[h!]
\begin{center}
			\includegraphics[height=150px, width=150px]{figures/GMcriseBV.png}
			\includegraphics[height=150px,width=150px]{figures/nbcompBV.png}
\end{center}
\vspace{-5mm}
			\centerline{(a) \hspace{130px} (b)}
			\caption{Production case study: (a) Example of a non-Gaussian real variable easily modeled by a Gaussian mixture, (b) distribution of the number of components found for each covariate.}\label{fig:graphMixmod.production}
\end{figure}

%		\begin{figure}[h!]
%	\begin{minipage}[t]{.30\linewidth}
%			\includegraphics[height=150px,width=150px]{figures/nbcompBV.png}%{figures/mixmod.png} 
%			\caption{Distribution of the number of components found for each covariate.}\label{graphMixmod}
%	\end{minipage} \hfill
%	\begin{minipage}[t]{.30\linewidth}
%			\includegraphics[height=150px, width=150px]{figures/GMcriseBV.png}%{figures/histR2exfos.png} $R^2_{adj}$ of the 76 sub-regressions.
%			\caption{Another example of non-Gaussian real variable easily modeled by a Gaussian mixture.}
%	\end{minipage} \hfill
%   \begin{minipage}[t]{.30\linewidth}
%			\includegraphics[height=150px,width=150px]{figures/histcorrelBVBI.png} 
%			\caption{Histogram of correlations in $\boldsymbol{X}$.} \label{compareMSEexfos}
%   \end{minipage}
%\end{figure} 
%

%\begin{figure}[h!]
%\centering
%	\includegraphics[height=150px, width=150px]{figures/R2corregBVBI.png} 
%			\caption{$R^2_{adj}$ of the 55 sub-regressions.}\label{R2bv}
%\end{figure}
%The response variable was binary but $n$ was too small compared to $p$ to use logistic regression so we have considered $\boldsymbol{Y}$ as a continuous variable and then made imputation by $1$ when $\hat{\boldsymbol{Y}}>0.5$ and by $0$ else.

Table~\ref{Res_prod} displays predictive results associated to different estimation methods with and without {\sc CorReg}. Note that {\sc mse} is calculated though a leave-one-out method because of the small sample size. We can again see that {\sc CorReg} globally improves the results for each method tested in terms of prediction, with always a more parsimonious regression on $\boldsymbol{Y}$.

\begin{table}[h!]
\centering
\begin{tabular}{llcc}
	\hline 
	Method& Indicator& With {\sc CorReg} & Without {\sc CorReg} \\ 
	\hline\hline
	{\sc ols} &  {\sc mse}& 1.95& 51 810\\
		& complexity & 91& 100 \\
	\hline 
		{\sc lasso} & {\sc mse} & {\bf 0.106} & 0.120\\
		& complexity & 27&34\\
	\hline 
		elasticnet & {\sc mse} &0.140 &0.148\\
		& complexity &10 &13\\
	\hline 
		ridge & {\sc mse} & 0.179 & 0.177\\
		& complexity &91 &146\\
	\hline 
\end{tabular} 
\caption{Production case study: Results obtained with leave-one out cross-validation ($n=100, d=145$). {\sc mse} is calculated though a leave-one-out method because of the small sample size. In bold, the best {\sc mse} value.}	\label{Res_prod}
\end{table}

\FloatBarrier
\section{Conclusion and perspectives} \label{conclusion}
	We have seen that correlations can lead to serious estimation and variable selection problems in linear regression. In such a situation, it can be useful to explicitly model the structure between the covariates and to use this structure by simple probabilistic marginalization to avoid correlations issues. It has led to the so called \textsc{CorReg} method, which can be viewed as a variable pre-selection based on covariable correlation and which has to be then combined with any standard estimate and covariable selection procedure. The \textsc{CorReg} strategy is able to give not only efficiently prediction but also a better understanding of the phenomenon at hand thanks to the sub-regressions description of correlated covariables. Its strength is then its great interpretability of the model, composed of several short linear regression easily managed by non-statisticians. In particular, we have illustrated both advantages of \textsc{CorReg} in two industrial contexts related to the steel industry.

Future works we plan is to allow {\sc CorReg} to manage missing values. They are very common in industry for instance. Indeed, the structure can be used to estimate missing values in $\boldsymbol{X}$ thanks to the full generative model assumed in {\sc CorReg}. Another perspective would be to take back lost information (the residual of each sub-regression) to improve predictive efficiency when needed. It would only consists in a second step of linear regression between the residuals and would thus still be able to use any selection method.

\paragraph{Package} \textsc{CorReg} is accessible on {\sc cran}: http://cran.r-project.org/web/packages/CorReg/index.html

\paragraph{Acknowledgements}
We want to thank ArcelorMittal Atlantique \& Lorraine that has granted this work, given the chance to use \textsc{CorReg} on real datasets and authorized the package to be open-sourced licensed (\textsc{CeCILL}), especially Dunkerque's site where most of the work has been done.
\bibliography{biblio}{}
\bibliographystyle{plain}
%\section{Appendices}
%	\subsection{Identifiability of the structure} \label{preuveident}
%	The model presented above relies on a discrete structure $S$ between the covariates. But to find it we need identifiability property to insure the MCMC will asymptotically find the true model. Identifiability of the structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions? 
%	
%		If there are exact sub-regressions ($\sigma^2_j=0$), the structure won't be identifiable. But it only means that several structure will have the same likelihood and they will have the same interpretation. So it's not really a problem. Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete. \textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
%	In the followings we suppose $\sigma^2_j\neq 0$, then $\boldsymbol{X}_f'\boldsymbol{X}_f$ and $\boldsymbol{X}'\boldsymbol{X}$ are of full rank (but the later is ill-conditioned for small values of $\sigma^2_j$).
%	\\
%	
%Our full generative model is a $p$-sized Gaussian mixture model of $K$ distinct components and 
%
%	can be seen as a $\mathbf{SR}$ model defined by Maugis \cite{maugis2009variable}. In this section, $S$ will denote the set of variable as in the paper from Maugis and we call Gaussian mixtures the Gaussian mixtures with at least two distinct components. The equivalence with Maugis's model is defined by:
%	$\boldsymbol{X}_r=\boldsymbol{y}^{S^c}$ and $\boldsymbol{X}_f=\boldsymbol{y}^R$. We have supposed independence between variables in $\boldsymbol{X}_f$ so the identifiability theorem from Maugis tells that our model is identifiable if variables in $\boldsymbol{X}_f$ are Gaussian mixtures (what we supposed in section \ref{sectionfullgen}).
%	\\
%	
%	
%%First, we observe that if each variable in $\boldsymbol{X}_r$ is a Gaussian mixture, then there must be at least one Gaussian mixture on the right of each sub-regression. 
%We define $\boldsymbol{X}^G \subsetneq \boldsymbol{X}_f$ containing Gaussian variables and we note the Gaussian mixtures $\boldsymbol{X}^{G^c}\neq \emptyset$ its complement in $\boldsymbol{X}_f$.
%We suppose that variables in $\boldsymbol{X}_r$ are all Gaussian mixtures. It implies that $\forall j  \in I_r,\exists i \in I_f^j $ so that $\boldsymbol{X}^i \subset \boldsymbol{X}^{G^c} $ since any linear combination of Gaussian variable would only give a Gaussian (so each sub-regression contain at least one Gaussian mixture as a regressor).
%\\
%	We introduce the matricial notation
%		$\boldsymbol{X}_r=\boldsymbol{X}_f\boldsymbol{\alpha} + \boldsymbol{\varepsilon}$ where
%		 $\boldsymbol{\alpha}$ is the $(p-p_r)\times p_r$ matrix whose columns are the $\boldsymbol{\alpha}_j$ and $\boldsymbol{\varepsilon}$ is the $n\times p_r$ matrix whose columns are the $\boldsymbol{\varepsilon}_j$
%
%		
%The theorem from Maugis guarantee that a sub-regression between Gaussian mixtures is identifiable in terms of which one is regressed by others.
%		\begin{eqnarray}
%		 \boldsymbol{X}_{r|\boldsymbol{X}^G,\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^G\boldsymbol{\alpha}_G+\boldsymbol{X}^{G^c}\boldsymbol{\alpha}_{G^c}+ \boldsymbol{\varepsilon} \\
%			\boldsymbol{X}_{r|\boldsymbol{X}^{G^c}}&=& \boldsymbol{X}^{G^c}\boldsymbol{\alpha}_{G^c} + \tilde{\boldsymbol{\varepsilon}} \textrm{ is identifiable where} \\
%			\tilde{\boldsymbol{\varepsilon}}_j&=&\boldsymbol{X}^{G}\boldsymbol{\alpha}_j^{G}+ \boldsymbol{\varepsilon}_j \textrm{ is Gaussian.}  
%		\end{eqnarray}
%		So a sufficient condition for identifiability is to have at least one Gaussian mixture in each sub-regression.
%
%
%	
%	\subsection{The \textsc{CorReg} package}
%\subsubsection{Alternative neighbourhoods for the MCMC}
%	We have here at each step $|\mathcal{V}_{S,j}|=p$ candidates but some other constraints can be added on the definition of $\mathcal{S}$ and will consequently modify the size of the neighbourhood (for example a maximum complexity for the internal regressions or the whole structure, a maximum number of internal regressions, {\it etc.}). \textsc{CorReg} allows to modify this neighbourhood to better fit users constraints. Relaxation (column-wise and row-wise) is optional but gives more stability to the number of feasible candidates at each step and allows to modify several parts of $I_f$ in only one step when needed. Hence it improves efficiency by a significant reinforcement of the irreductibility of the Markov chain. Rejecting candidates instead of doing the relaxation steps will  however reduce the number of evaluated candidates and thus accelerate the walk. So it can be used for a warming phase when $n$ is great and time is missing.
%	
%	The hierchical uniform hypothesis made above for $P(S)$ implies $p_r<\frac{p}{2}$ and $p_f^j<\frac{p}{2}$ so candidates may be rejected to satisfy this hypothesis. Stronger constraints on $p_r$ and/or $p_f^j$ can be given in \textsc{CorReg} if relevant.
%	
%If the algorithm did not have time to converge (stationnarity), it can be continued with a few step for which the neighboorhood would only contain smaller candidates (in terms of complexity). It is equivalent to ask for each element in $I_f$ if the criterion $P(S|\boldsymbol{X})$ would be better without it. Thus it can be seen as a final cleaning step. But in fact, it's just continuing the MCMC with a reduced neighbourhood.

\end{document}

### 2014-08-21 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@808 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Then, marginalisation on the linearly explained covariates produces a parsimonious correlation-free regression model from which classical procedures for estimating regression coefficient, including any variable selection procedures, can be plugged.
-Both simulated and real-life datasets from steel industry, where correlated variables are frequent, highlight that this proposed covariable pretreatment-like method has two essential beneficits: First, it offers a real readability of the linear links between covariates; Second, it improves significantly efficiency of classical estimation/selection methods which are performed after.
+Both simulated and real-life datasets from steel industry, where correlated variables are frequent, highlight that this proposed covariates pretreatment-like method has two essential benefits: First, it offers a real readability of the linear links between covariates; Second, it improves significantly efficiency of classical estimation/selection methods which are performed after.
 An {\sc r} package (\textsc{CorReg}), available on the {\sc cran}, implements this new method.
@@
 Originality of this paper is to distinguish explicitly them by proposing a specific decorrelation step followed by any classical variable selection step chosen by the practitioner.
-The decorrelation step can be viewed as a variable selection step but it is focused only on correlations, not on the number of covariables.
+The decorrelation step can be viewed as a variable selection step but it is focused only on correlations, not on the number of covariates.
 
@@
 
-Reducing the variance induced by the large number of covariables can be reached by targeting a better bias-variance trade off.
+Reducing the variance induced by the large number of covariates can be reached by targeting a better bias-variance trade off.
 Relating methods are numerous and continue to generate a lot of work.
@@
 Ridge regression \cite{marquardt1975ridge} is a shrinkage method which proposes possibly biased estimator that can be written in terms of a parametric $L_2$ penalty.
-It does not select variables since coefficients tend to 0 but don't reach 0, leading to difficult interpretations for a large number of covariates.	Since real datasets may imply many irrelevant variables, variable selection should be prefered for more interpretable models.
+It does not select variables since coefficients tend to 0 but don't reach 0, leading to difficult interpretations for a large number of covariates.	Since real datasets may imply many irrelevant variables, variable selection should be preferred for more interpretable models.
 Variable selection methods may add also some bias by deleting some relevant covariates but it may reduced drastically the variance by the dimension reduction.
@@
 
-Another way for improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{oscar} \cite{bondell2008simultaneous}) to reduce dimension and also correlations if correlated covariates are in the same clusters.
+Another way for improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octagonal Shrinkage and Clustering Algorithm for Regression (\textsc{oscar} \cite{bondell2008simultaneous}) to reduce dimension and also correlations if correlated covariates are in the same clusters.
 A possible bias is added by the dimension reduction inherent to the coefficients clustering.
@@
 Advantages of this approach is twofold.
-First, it improves interpretation through a good readability of dependency between covariables.
+First, it improves interpretation through a good readability of dependency between covariates.
 Second, this marginal model is still a ``true'' model provided that both the initial regression model and all the sub-regressions are ``true''.
@@
 At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular.
-Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariables higher.
+Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariates higher.
 
@@
 \begin{hyp}\label{H2}
-In addition, we make the complementary hypothesis that the responses covariables and the predictor covariables are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariables and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates.
+In addition, we make the complementary hypothesis that the responses covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates.
 This new assumption allows to obtain very simple sub-regressions sequences, discarding hierarchical ones, in particular uninteresting cyclic sub-regressions.
@@
 \paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions.
-As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariables in $\boldsymbol{J}_p$ or not.
+As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariates in $\boldsymbol{J}_p$ or not.
 An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}.
@@
 \item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
-\item As a consequence of Hypothesis~\ref{H1}, ``free'' covariables $\boldsymbol{X}_f$ are {\it all} decorrelated.
+\item As a consequence of Hypothesis~\ref{H1}, ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
 \item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals.
@@
 \end{eqnarray}
-Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariables $\boldsymbol{X}_f$.
-This decorrelation process has also acted like a specific covariable selection process-like because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$.
+Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariates $\boldsymbol{X}_f$.
+This decorrelation process has also acted like a specific variable selection process-like because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$.
 These two statements are expected to decrease the variance of further estimates of $\boldsymbol{\beta}$.
@@
 First, this regression has a higher residual variance than the initial one since it is now $\sigma^{2*}_Y=\sigma^2_Y+\sum_{j \in I_r}\beta_{j}^2\sigma^2_j$ instead of $\sigma^2_Y$.
-Second, covariable selection being equivalent to set $\hat{\boldsymbol{\beta}}_r=\boldsymbol{0}$, it implies possibly biased estimates of $\boldsymbol{\beta}_r$.
-As a conclusion, we are faced to a typical {\it bias-variance trade off}.
+Second, variable selection being equivalent to set $\hat{\boldsymbol{\beta}}_r=\boldsymbol{0}$, it implies possibly biased estimates of $\boldsymbol{\beta}_r$.
+As a conclusion, we are faced with a typical {\it bias-variance trade off}.
 We will illustrate it in the next section in the case of the {\sc ols} estimate.
@@
 		\end{equation}
-Indeed, no correlations between covariables $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditionned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}).
+Indeed, no correlations between covariates $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditioned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}).
 This bias-variance trade off can be resumed by the Mean Squared Error (\textsc{mse}) associated to both estimates:
@@
 \paragraph{Numerical illustration} We now illustrate the bias-variance trade off, through a numerical example.
-Let a simple case with $d=5$ covariables $\boldsymbol{X}=(\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{X}^4, \boldsymbol{X}^5)$ independently drawn by four independent Gaussian $\mathcal{N}_n(\boldsymbol{0},\boldsymbol{I})$.
+Let a simple case with $d=5$ covariates $\boldsymbol{X}=(\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{X}^4, \boldsymbol{X}^5)$ independently drawn by four independent Gaussian $\mathcal{N}_n(\boldsymbol{0},\boldsymbol{I})$.
 Thus, $J_f=\{1,2,4,5\}$ and $\boldsymbol{X}_f=(\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5)$.
@@
 However, in the general situation, $\boldsymbol{S}$ has to estimated.
-A standard method as graphical {\sc lasso} \cite{friedman2008sparse}, which searchs for a structure on the precision matrix (inverse of the variance-covariance matrix) by setting some coefficients of the precision matrix to zero, can not be applied since it is not designed to estimate oriented structures like $\boldsymbol{S}$.
+A standard method as graphical {\sc lasso} \cite{friedman2008sparse}, which searches for a structure on the precision matrix (inverse of the variance-covariance matrix) by setting some coefficients of the precision matrix to zero, can not be applied since it is not designed to estimate oriented structures like $\boldsymbol{S}$.
 
@@
 \begin{itemize}
-\item Hypothesis~\ref{H4} is the keystone to define a full generative model on the whole covariables $\boldsymbol{X}$.
-On the one hand, the {\sc bic} criterion can be applied in this context, avoiding to use a cross-validation criterio which can be much more time-consuming.
+\item Hypothesis~\ref{H4} is the keystone to define a full generative model on the whole covariates $\boldsymbol{X}$.
+On the one hand, the {\sc bic} criterion can be applied in this context, avoiding to use a cross-validation criterion which can be much more time-consuming.
 On the other hand, the great flexibility  of Gaussian mixture models \cite{mclachlan2004finite}, provided that the number of components $k_j$ has to be estimated, implies that Hypothesis~\ref{H4} is particularly weak in fact.
@@
 In fact, if there exists some sub-regressions with only one regressor, identifiability holds for these sub-regressions only up to a permutation between the related response and predictor covariates.
-However, even in this case, full identifiability may occur thank's to constraints on response and predictor covariates given in Hypothesis~\ref{H1} and~\ref{H2}.
+However, even in this case, full identifiability may occur thanks to constraints on response and predictor covariates given in Hypothesis~\ref{H1} and~\ref{H2}.
 \item As any {\sc bic} criterion, the {\sc bic}$_U$ and {\sc bic}$_H$ criteria are consistent \cite{BIChuard}.
@@
 	\begin{itemize}
-	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariable is added to $\boldsymbol{S}$
+	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariate is added to $\boldsymbol{S}$
 \begin{equation}
@@
 \end{equation}	
-		\item {\bf Removing a sub-regression}: a sub-regression with only one predictor covariable is removed from $\boldsymbol{S}$
+		\item {\bf Removing a sub-regression}: a sub-regression with only one predictor covariate is removed from $\boldsymbol{S}$
 \begin{equation}
@@
 \end{equation}	
-	\item {\bf Adding a predictor covariable}: a predictor covariable is added to one sub-regression of $\boldsymbol{S}$
+	\item {\bf Adding a predictor covariate}: a predictor covariate is added to one sub-regression of $\boldsymbol{S}$
 \begin{equation}
@@
 \end{equation}
-	\item {\bf Removing a predictor covariable}: a predictor covariable is removed from one sub-regression of $\boldsymbol{S}$
+	\item {\bf Removing a predictor covariate}: a predictor covariate is removed from one sub-regression of $\boldsymbol{S}$
 \begin{equation}
@@
 All sub-regressions are of length two ($\forall j=1,\ldots,d_r, d_p^j=2$ and we have $d_r=16$ sub-regressions.
-The datasets are then scaled, so that covariates $\boldsymbol{X}_f$ for avoiding large distorsions for variances or for means.	Different sample sizes  $n\in (30,50,100,400)$ are chosen, thus considering experiments in both situations $n<p$ and $n>p$.
+The datasets are then scaled, so that covariates $\boldsymbol{X}_f$ for avoiding large distortions for variances or for means.	Different sample sizes  $n\in (30,50,100,400)$ are chosen, thus considering experiments in both situations $n<p$ and $n>p$.
 In all figures, the thickness of the lines will represent various values of $n$, the thicker being the greater.
@@
 			\begin{itemize}
-				\item $T_r=|J_r \cap \hat{J}_r|$ (``True Response''): it corresponds to the number of estimate response covariables in $\hat{\boldsymbol{S}}$ which are {\it truly} response covariables in the true model $\boldsymbol{S}$.
-				\item $W_r=|\hat{J}_r|-T_r$ (``Wrong Response''): it corresponds to the number of estimate response covariables in $\hat{\boldsymbol{S}}$ which are are {\it wrongfully} response covariables in the true model $\boldsymbol{S}$.
+				\item $T_r=|J_r \cap \hat{J}_r|$ (``True Response''): it corresponds to the number of estimate response covariates in $\hat{\boldsymbol{S}}$ which are {\it truly} response covariates in the true model $\boldsymbol{S}$.
+				\item $W_r=|\hat{J}_r|-T_r$ (``Wrong Response''): it corresponds to the number of estimate response covariates in $\hat{\boldsymbol{S}}$ which are are {\it wrongfully} response covariates in the true model $\boldsymbol{S}$.
 			\end{itemize}
@@
 In addition, $\mbox{\sc bic}_H$ need to have significantly strong sub-regressions ($R^2>0.4$) to detect them.
-However, this is not a problem since our {\sc CorReg} strategy is expected to involve only quite strong correlated covariables to have potential interest.
+However, this is not a problem since our {\sc CorReg} strategy is expected to involve only quite strong correlated covariates to have potential interest.
 Finally, we keep now the $\mbox{\sc bic}_H$ criterion as the best one for our purpose.
@@
 Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} respectively display the results on the same manner as the previous {\sc ols} with {\sc CorReg}+{\sc ols}.
-We see that {\sc CorReg}, used as a pre-treatment, provides similar prediction accurary as the three variable selection methods (this prediction is often better for small datasets) but with much more parsimonious regression models on $\boldsymbol{Y}$.
+We see that {\sc CorReg}, used as a pre-treatment, provides similar prediction accuracy as the three variable selection methods (this prediction is often better for small datasets) but with much more parsimonious regression models on $\boldsymbol{Y}$.
 This is remarkable because the true model depends on all covariates, so it highlight that other variable selection methods may be really penalized by the correlations between the covariates.
@@
 Indeed, ridge regression alone gives a full model (40 covariates even with only 30 individuals) and will give models too complex to be easily interpreted.
-Thus, the combination {\sc CorReg}+ridge provideshigh benefits for interpretation whle preserving good prediction accuracy.
+Thus, the combination {\sc CorReg}+ridge provides high benefits for interpretation while preserving good prediction accuracy.
 
@@
 When $n$ rises, the sub-regression are detected and relevant covariates are removed.
-As a conseuqnce, {\sc CorReg} can not improve the results and increases the {\sc mse}.
+As a consequence, {\sc CorReg} can not improve the results and increases the {\sc mse}.
 However, for strong correlations, the error shrinks as the model tends to be less identifiable and switching variables is not a problem anymore even with large values of $n$.
@@
 Figure~\ref{X2LASSO} compares {\sc lasso} with {\sc CorReg}+{\sc lasso}.
-It shows that {\sc lasso} naturally tends to keep $\boldsymbol{X_r}$ and thus is better bevause it corresponds to the true model.
+It shows that {\sc lasso} naturally tends to keep $\boldsymbol{X_r}$ and thus is better because it corresponds to the true model.
 So {\sc CorReg} is almost always the worst method.
@@
 Note that this model choice is only between two models so it avoids multiple tests issues and computational cost explosion.
-Thus, we suggest to always compute the ``with {\sc CorReg}'' and the ``without {\sc CorReg}'' solutions and then to compare them with the more pertinent criterion (AIC, cross-validation, validation sample, {\it etc.}) according to the context (size of the datasets for example).
+Thus, we suggest to always compute the ``with {\sc CorReg}'' and the ``without {\sc CorReg}'' solutions and then to compare them with the more pertinent criterion ({\sc AIC}, cross-validation, validation sample, {\it etc.}) according to the context (size of the datasets for example).
 
@@
 
-We have generated a non-linear sub-regression to test the robustess of our model.
+We have generated a non-linear sub-regression to test the robustness of our model.
 $\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as before.
@@
 This second example is about a phenomenon that impacts the productivity of a steel plant.
-It is described by a (confidential)  response variable $\boldsymbol{Y}$ and $d=145$ covariables from the whole process to explain it but only $n=100$ individuals are present.
+It is described by a (confidential)  response variable $\boldsymbol{Y}$ and $d=145$ covariates from the whole process to explain it but only $n=100$ individuals are present.
 The stake is to gain $20\%$ of productivity on a specific product with high added value.
@@
 In such a situation, it can be useful to explicitly model the structure between the covariates and to use this structure by simple probabilistic marginalization to avoid correlations issues.
-It has led to the so called \textsc{CorReg} method, which can be viewed as a variable pre-selection based on covariable correlation and which has to be then combined with any standard estimate and covariable selection procedure.
-The \textsc{CorReg} strategy is able to give not only efficiently prediction but also a better understanding of the phenomenon at hand thanks to the sub-regressions description of correlated covariables.
+It has led to the so called \textsc{CorReg} method, which can be viewed as a variable pre-selection based on covariates correlations and which has to be then combined with any standard estimate and variable selection procedure.
+The \textsc{CorReg} strategy is able to give not only efficiently prediction but also a better understanding of the phenomenon at hand thanks to the sub-regressions description of correlated covariates.
 Its strength is then its great interpretability of the model, composed of several short linear regression easily managed by non-statisticians.

### 2014-08-21 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@809 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	This estimate requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which can lead to great variance estimates, so unstable estimates, if it is ill-conditioned.
-Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grows (in absolute value)  \cite{hoerl1970ridge}.
+Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grow (in absolute value)  \cite{hoerl1970ridge}.
 At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular.
@@
 
-\paragraph{Package} \textsc{CorReg} is accessible on {\sc cran}: http://cran.r-project.org/web/packages/CorReg/index.html
+\paragraph{Package} {\tt CorReg} is accessible on {\sc cran}: http://cran.r-project.org/web/packages/CorReg/index.html
 

### 2014-08-21 www/article.tex
up sum
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@810 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{equation}
-		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
+		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}_Y \label{regressionsimple}
 	\end{equation}
-with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon} \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension.
+with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon}_Y \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension.
 The $d\times 1$ vector $\boldsymbol{\beta}$ gathers the coefficients of the regression\footnote{Usually a constant is included as one of the regressors.
@@
 \begin{hyp}\label{H2}
-In addition, we make the complementary hypothesis that the responses covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates.
+In addition, we make the complementary hypothesis that the response covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates.
 This new assumption allows to obtain very simple sub-regressions sequences, discarding hierarchical ones, in particular uninteresting cyclic sub-regressions.
@@
 \subsection{Marginal regression with decorrelated covariates}
-The aim is now to use the model of linear sub-regressions $\boldsymbol{S}$ (that we assume to be known in this part) between some covariates of $\boldsymbol{X}$ to obtain a linear regression on $\boldsymbol{Y}$ relying on only uncorrelated variables $\boldsymbol{X}_f$.
+The aim is now to use the model of linear sub-regressions $\boldsymbol{S}$ (that we assume to be known in this part) between some covariates of $\boldsymbol{X}$ to obtain a linear regression on $\boldsymbol{Y}$ relying only on uncorrelated variables $\boldsymbol{X}_f$.
 The way to proceed is to marginalize the joint distribution of $\{(\boldsymbol{Y},\boldsymbol{X}_f) |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ to obtain the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ depending only on uncorrelated variables $\boldsymbol{X}_f$:
@@
 \begin{hyp}\label{H3}
-We assume that all errors $\boldsymbol{\varepsilon}$ and $\boldsymbol{\varepsilon}_j$ ($j=1,\ldots,d_r$) are {\it mutually independent}.
+We assume that all errors $\boldsymbol{\varepsilon}_Y$ and $\boldsymbol{\varepsilon}_j$ ($j=1,\ldots,d_r$) are {\it mutually independent}.
 It implies in particular that conditional response covariates $\{\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j\}$, with distribution defined in (\ref{eq:SR}), are {\it mutually independent}:
@@
 
-Decomposing $\boldsymbol{\beta}=(\boldsymbol{\beta}_r',\boldsymbol{\beta}_f')'$ by $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the response and to the free covariates, we can rewrite (\ref{regressionsimple}):
+Decomposing $\boldsymbol{\beta}=(\boldsymbol{\beta}_r',\boldsymbol{\beta}_f')'$ by $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the responses and to the free covariates, we can rewrite (\ref{regressionsimple}):
 \begin{equation}
@@
 \end{equation}
-Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$:
+Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}_Y$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$:
 \begin{eqnarray}
-	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
+	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j)+  \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
 	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl}
@@
 Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariates $\boldsymbol{X}_f$.
-This decorrelation process has also acted like a specific variable selection process-like because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$.
+This decorrelation process has also acted like a specific variable selection process because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$.
 These two statements are expected to decrease the variance of further estimates of $\boldsymbol{\beta}$.
 However, the counterpart is twofold.
-First, this regression has a higher residual variance than the initial one since it is now $\sigma^{2*}_Y=\sigma^2_Y+\sum_{j \in I_r}\beta_{j}^2\sigma^2_j$ instead of $\sigma^2_Y$.
+First, this regression has a higher residual variance than the initial one since it is now $\sigma^{2*}_Y=\sigma^2_Y+\sum_{j =1}^{d_r}\beta_{J_r^j}^2\sigma^2_j$ instead of $\sigma^2_Y$.
 Second, variable selection being equivalent to set $\hat{\boldsymbol{\beta}}_r=\boldsymbol{0}$, it implies possibly biased estimates of $\boldsymbol{\beta}_r$.
@@
 		\begin{equation}
-			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
+			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
 		\end{equation}
@@
 		\begin{equation}
-			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}.
+			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}.
 \label{eq:varOLS*}
@@
 Thus, $\boldsymbol{\alpha}_1=(1,1)'$, $\boldsymbol{J}_r=(3)$, $J_r=\{3\}$, $\boldsymbol{X}_r=(\boldsymbol{X}^3)$, $\boldsymbol{d}_p=(2)$, $\boldsymbol{J}_p=(\{1,2\})$, $\boldsymbol{X}^{\boldsymbol{J}_p^1}=(\boldsymbol{X}^1,\boldsymbol{X}^2)$ and $\boldsymbol{S}=((3),(\{1,2\}))$.
-Concerning now the regression with $\boldsymbol{Y}$, we define $\boldsymbol{Y}|\boldsymbol{X}=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{X}^3+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_Y$, where $\boldsymbol{\beta}=(1,1,1,1,1)'$ and $\sigma_Y \in \{10,20\}$.
+Concerning now the regression with $\boldsymbol{Y}$, we define $\boldsymbol{Y}|\boldsymbol{X}=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{X}^3+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_Y=\boldsymbol{X}\boldsymbol{\beta}+\boldsymbol{\varepsilon}_Y$, where $\boldsymbol{\beta}=(1,1,1,1,1)'$ and $\sigma_Y \in \{10,20\}$.
 Finally, we deduce that $\boldsymbol{Y}|\boldsymbol{X}_f,\boldsymbol{S}=2\boldsymbol{X}^1+2\boldsymbol{X}^2+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_1+\boldsymbol{\varepsilon}_Y$
@@
 It appears that, when the sub-regression is strong (low $1-R^2$), $\hat{\boldsymbol{\beta}}^*$ is a better estimate than $\hat{\boldsymbol{\beta}}$: the gain on {\sc mse} can even be very significant.
-This effect amplified by the $\sigma_Y$ increase but is reduced by the $n$ increases.
+This effect is amplified by the $\sigma_Y$ increase but is reduced by the $n$ increases.
 Thus, the estimate $\hat{\boldsymbol{\beta}}^*$ should be particularly useful when some covariates are highly correlated, when also the sample sizes is small and when the residual variance of $\boldsymbol{Y}$ is large.
@@
 For instance, Structural Equations Models (\textsc{sem}), which are related to our model as already mentioned in Section~\ref{sec:intro}, are often used in social sciences and economy where a structure $\boldsymbol{S}$ is generally ``hand-made''.
-However, in the general situation, $\boldsymbol{S}$ has to estimated.
+However, in the general situation, $\boldsymbol{S}$ has to be estimated.
 A standard method as graphical {\sc lasso} \cite{friedman2008sparse}, which searches for a structure on the precision matrix (inverse of the variance-covariance matrix) by setting some coefficients of the precision matrix to zero, can not be applied since it is not designed to estimate oriented structures like $\boldsymbol{S}$.

### 2014-08-21 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@811 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ describe below:
+We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ described below:
 	\begin{itemize}

### 2014-08-22 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@812 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\paragraph{Defining $\mathbb{P}(\boldsymbol{S})$} The most standard choice consists of putting a uniform distribution on the model space of $\boldsymbol{S}$, this choice being noted $\mathbb{P}_U(\boldsymbol{S}) = |\boldsymbol{S}|^{-1}$, with $|\boldsymbol{S}|$ the space dimension of $\boldsymbol{S}$.
+\paragraph{Defining $\mathbb{P}(\boldsymbol{S})$} The most standard choice consists of putting a uniform distribution on the model space $\mathcal{S}$, this choice being noted $\mathbb{P}_U(\boldsymbol{S}) = |\mathcal{S}|^{-1}$, with $|\mathcal{S}|$ the space dimension of $\mathcal{S}$.
 
-However, $\boldsymbol{S}$ being combinatorial, $|\boldsymbol{S}|$ is huge.
+However, $\mathcal{S}$ being combinatorial, $|\mathcal{S}|$ is huge.
 It has two cumulated consequences: First, the {\it exact} probability $\mathbb{P}(\boldsymbol{S}|\boldsymbol{X})$ may be of the same order of magnitude for a large number of candidates  $\boldsymbol{S}$, including the best one; Second, the {\sc bic} {\it approximations} of this quantity may introduce additional confusion to wisely distinguish between model probabilities.
@@
  & = & \mathbb{P}_U(\boldsymbol{J}_p|\boldsymbol{d}_p,\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{d}_p|\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{J}_r|d_r)\times \mathbb{P}_U(d_r) \\
- & = & \left[\prod_{j\in \boldsymbol{J}_r} \left(\begin{array}{c} d-d_r \\ d_p^j \end{array}\right) \right]^{-1} \times \left[d-d_r\right]^{-d_r} \times \left[\left(\begin{array}{c} d \\ d_r \end{array} \right)\right]^{-1} \times [d+1]^{-1},
+ & = & \left[\prod_{j=1}^{d_r} \left(\begin{array}{c} d-d_r \\ d_p^j \end{array}\right) \right]^{-1} \times \left[d-d_r\right]^{-d_r} \times \left[\left(\begin{array}{c} d \\ d_r \end{array} \right)\right]^{-1} \times [d+1]^{-1},
 \end{eqnarray}

### 2014-08-22 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@813 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-We used \textsc{Rmixmod} \cite{remi:hal-00919486} to estimate the Gaussian mixture densities of each covariate $\boldsymbol{X}$, settings being using at default values both for parameter and the number of components estimation.
+We used \textsc{Rmixmod} \cite{remi:hal-00919486} to estimate the Gaussian mixture densities of each covariate $\boldsymbol{X}$, settings being used at default values both for parameter and the number of components estimation.
 For each configuration, the {\sc mcmc} walk was launched on $10$ initial structures with 1~000 iterations each time.

### 2014-08-22 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@814 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{eqnarray}
-where $\left(\begin{array}{c} a \\ b \end{array}\right)$ designs combination of $a$ things taken $b$ at a time without repetition and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand.
+where $\left(\begin{array}{c} a \\ b \end{array}\right)$ designs the number of b-element subsets of an a-element set and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand.
 $\mathbb{P}_H(\boldsymbol{S})$ gives decreasing probabilities to more complex models, provided that the following new hypothesis is verified:
@@
 
-This algorithm alternates two steps: the definition of a neighbourhood $\mathcal{V}(\boldsymbol{S})$ around the current structure $\boldsymbol{S}$ distribution and then the generation of a new structure $\boldsymbol{S}^+$ belonging to this neighbourhood according to its posterior probability.
+This algorithm alternates two steps: the definition of a neighbourhood $\mathcal{V}(\boldsymbol{S})$ around the current structure $\boldsymbol{S}$ and then the generation of a new structure $\boldsymbol{S}^+$ belonging to this neighbourhood according to its posterior probability.
 
@@
 \begin{equation}
-\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{\boldsymbol{J}}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
+\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{J}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
 \end{equation}	
@@
 \begin{equation}
-\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{\boldsymbol{J}}_p^{j}=\tilde{\boldsymbol{J}}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
+\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
 \end{equation}
@@
 \begin{equation}
-\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{\boldsymbol{J}}_p^{j}=\tilde{\boldsymbol{J}}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
+\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
 \end{equation}	
@@
 	\begin{equation}
-	\boldsymbol{S}^+ \sim \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
+	\mathbb{P}(\boldsymbol{S}^+|\mathcal{V}(\boldsymbol{S})) \sim \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
 	\end{equation}

### 2014-08-22 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@815 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{eqnarray}
-where $\left(\begin{array}{c} a \\ b \end{array}\right)$ designs the number of b-element subsets of an a-element set and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand.
+where $\left(\begin{array}{c} a \\ b \end{array}\right)$ means the number of b-element subsets of an a-element set and where all probabilities $\mathbb{P}_U(\cdot)$ denote uniform distribution on the related space at hand.
 $\mathbb{P}_H(\boldsymbol{S})$ gives decreasing probabilities to more complex models, provided that the following new hypothesis is verified:

### 2014-08-22 www/article.tex
up figure13
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@816 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@817 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{equation}	
-		\item {\bf Removing a sub-regression}: a sub-regression with only one predictor covariate is removed from $\boldsymbol{S}$
+		\item {\bf Removing a sub-regression}: a sub-regression  is removed from $\boldsymbol{S}$
 \begin{equation}
@@
 We see that {\sc CorReg}, used as a pre-treatment, provides similar prediction accuracy as the three variable selection methods (this prediction is often better for small datasets) but with much more parsimonious regression models on $\boldsymbol{Y}$.
-This is remarkable because the true model depends on all covariates, so it highlight that other variable selection methods may be really penalized by the correlations between the covariates.
+This is remarkable because the true model depends on all covariates, so it highlights that other variable selection methods may be really penalized by the correlations between the covariates.
 
@@
 	\begin{minipage}[l]{.48\linewidth}
-			\includegraphics[height=180px,width=245px]{figures/res_article/robust_S.png} 
+			\includegraphics[height=180px,width=245px]{figures/res_article/robust_S_art.png} 
 			\caption{Evolution of the quality of $\hat{\boldsymbol{S}}$ when the paramater $a$ increases.}\label{resnonlin}

### 2014-08-22 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@818 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}.
-The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\boldsymbol{\sigma}^2_1,\ldots,\boldsymbol{\sigma}^2_{d_r})$ denotes the corresponding global variance.
+The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\sigma^2_1,\ldots,\sigma^2_{d_r})$ denotes the corresponding global variance.
 
@@
 \item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
-\item As a consequence of Hypothesis~\ref{H1}, ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
+\item As a consequence of Hypothesis~\ref{H1} and~\ref{H2} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
 \item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals.
@@
 \begin{hyp}\label{H4}
-All variates $\boldsymbol{X}^j\in\boldsymbol{X}_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
+All variates $\boldsymbol{X}^j$ with $j \in J_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
 \begin{equation}
@@
 \end{equation}
-where $\boldsymbol{\pi}_{j}=(\pi_{1j},\ldots,\pi_{k_jj})$ is the vector of mixing proportions with $\pi_{hj}>0$ and $\sum_{h=1}^{k_j}\pi_{hj}=1$, $\boldsymbol{\mu}_j=(\mu_{1j},\ldots,\mu_{k_jj})$  is the vector of centers and $\boldsymbol{\lambda}^2_j=(\lambda^2_{1j},\ldots,\lambda^2_{k_jj})$ is the vector of variances.
+where $\boldsymbol{\pi}_{j}=(\pi_{1j},\ldots,\pi_{k_jj})$ is the vector of mixing proportions with $\forall 1\leq h\leq k_j , \pi_{hj}>0$ and $\sum_{h=1}^{k_j}\pi_{hj}=1$, $\boldsymbol{\mu}_j=(\mu_{1j},\ldots,\mu_{k_jj})$  is the vector of centers and $\boldsymbol{\lambda}^2_j=(\lambda^2_{1j},\ldots,\lambda^2_{k_jj})$ is the vector of variances.
 We stack together all the mixture parameters in $\boldsymbol{\theta}=(\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j ; j \in J_f)$.

### 2014-08-26 pkg/CorReg/R/meilleur_lars.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@824 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,7 +1,7 @@
 # 'On estime les coefficients du meilleur mod?le obtenu (meilleur au sens du crit?re choisi)
 # ' 
 # ' 
-meilleur_lars<-function(lars=lars,X=X,Y=Y,mode=c("MSE","BIC"),intercept=TRUE,K=NULL,groupe=NULL,Amax=NULL){
+meilleur_lars<-function(lars=lars,X=X,Y=Y,mode=c("MSE","BIC"),intercept=TRUE,K=NULL,groupe=NULL,Amax=NULL,OLS=TRUE){
   mode=mode[1]
   X=as.matrix(X)
   if(is.null(Amax)){Amax=ncol(X)+1}

### 2014-08-26 pkg/CorReg/R/Estep.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@824 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -7,7 +7,9 @@
 #' @param alpha matrix of the coefficients (p+1)xp format Matrix
 #' @param X1 boolean for alternative method
 #'
-Estep<-function(X=X,alpha=alpha,M=NULL,Z=NULL,mixmod=mixmod,Zc=Zc,X1=FALSE){
+Estep<-function(X=X,alpha=alpha,M=NULL,Z=NULL,mixmod=NULL,Zc=Zc,X1=FALSE){
+   p=ncol(X)
+   n=nrow(X)
    #X bouge, mais pas alpha ni Z ni M ni mixmod
    if(is.null(M)){
       quimank=which(is.na(X),arr.ind=TRUE)# apriori on stockera les deux formats (M et quimank en permanence) et on fera pareil pour Z
@@ -16,8 +18,7 @@ Estep<-function(X=X,alpha=alpha,M=NULL,Z=NULL,mixmod=mixmod,Zc=Zc,X1=FALSE){
       quimank=which(M!=0,arr.ind=TRUE)# apriori on stockera les deux formats (M et quimank en permanence) et on fera pareil pour Z
    }
    
-   p=ncol(X)
-   n=nrow(X)
+   
    #on liste tout ce qu'il faut (on verra après comment extraire ça pour optimiser)
    nbmank=nrow(quimank)
    Z=alpha[-1,];Z[Z!=0]=1

### 2014-08-26 pkg/CorReg/R/inkei.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@825 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,22 @@
+inkei <- function(length=35, squirt=4, thickness=4)
+{
+   if (thickness > 3){
+      cat("          ", rep("_", length), "\n", sep="")
+      
+      for (t in 1:floor(thickness/2))
+      {     cat("          ", rep(" ", length+t), "\\\n", sep="")}
+      
+      cat("          ", rep(" ", length+floor(thickness/2)), "-o", rep("~", squirt), "\n", sep="")
+      
+      for (t in 1:(floor(thickness/2)-1))
+      {
+         cat("          ", rep(" ", length+floor(thickness/2)-t+1), "/\n", sep="")
+      }
+   }
+   cat("          ", rep("_", length+1), "/\n", sep="")
+   cat("          \\
+  .... |...|
+  .... |...|
+ \\____/____/")
+}
+#copyright Jim Aloku

### 2014-08-26 pkg/CorReg/R/Terminator.R
up graph
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@826 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -4,11 +4,14 @@
 #' @param target the dataset (matrix or data.frame) in which missing values will be made
 #' @param diag if >0 it is the thickness of the diagonal band of missing values
 #' @param Z adjacency matrix to coerce a maximum of 1 missing value per sub-regression for each individual
-Terminator<-function(target="Sarah Connor", wrath=0.1,diag=0,Z=NULL){
+Terminator<-function(target=NULL, wrath=0.1,diag=0,Z=NULL){
+   if(is.null(target)){
+      target="Sarah Connor"
+   }
    if(target[1]=="Sarah Connor" ){
       print("I'll be back !")
    }else if (target[1]=="bender"){
-      Bender()
+      Bender(a=wrath,b=diag,c=Z)
    }else if(diag>0){
       n=nrow(target)
       p=ncol(target)

### 2014-08-26 pkg/CorReg/R/Gibbs_miss.R
up graph
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@826 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,37 @@
+Gibbs_miss<-function(X=X,alpha=alpha,nbit=1,Z=NULL,Zc=NULL,M=NULL,comp_vect=NULL,loglik_bool=FALSE,warm=0,Xout=FALSE,GibbsIR=TRUE,Ir=NULL,mixmod=NULL,sigma_IR=NULL){
+   p=ncol(X)
+   if(is.null(M)){
+      M=0*X
+      M[is.na(X)]=1
+   }
+   if(is.null(alpha)){
+      alpha=hatB(Z = Z,X=X)
+   }
+   if(is.null(Z)){
+      Z=matrix(0,ncol=p,nrow=p)
+      Z[alpha[-1,]!=0]=1
+   }
+   if(is.null(Zc)){
+      Zc=colSums(Z)
+   }
+   if(is.null(Ir)){
+      Ir=which(Zc!=0)
+   }
+   if(is.null(sigma_IR)){
+      sigma_IR=rep(0,times=p)#résidus des regressions (stables pour gibbs)
+      for (j in Ir){
+         sigma_IR[j]=sd(X[,j]-X%*%alpha[-1,j])
+      }
+   }
+   if(is.null(mixmod)){
+      mixmod=density_estimation(X = X,detailed = TRUE)
+   }
+   missrow=sum(M)#nb d'individus à trou
+   quimiss=which(M!=0,arr.ind = TRUE)
+   if(is.null(comp_vect)){
+      comp_vect=1+0*M
+   }
+   res=Gibbs(last=TRUE,M=M,nbit=nbit,warm=warm,mixmod=mixmod$details,X=X,comp_vect=comp_vect,missrow=missrow,quimiss=quimiss,
+                       Z=Z,Zc=Zc,alpha=alpha,sigma_IR=sigma_IR,nbclust_vect=mixmod$nbclust,Ir=Ir,loglik_bool=loglik_bool,Xout=Xout,GibbsIR=GibbsIR)
+   return(res)
+}

### 2014-08-26 pkg/CorReg/R/bender.R

@@ -1,5 +1,7 @@
-Bender<-function(phrase="CorReg rules !"){
-  cat( "\n
+Bender<-function(phrase="CorReg rules !",a=0,b=0,c=0){
+   if(!is.null(c) & !try(c<4) & is.null(try(dim(c)))){
+      inkei(a,b,c)
+   }else{cat( "\n
                   o
                   |
                 ,'~'.
@@ -32,5 +34,5 @@ Bender<-function(phrase="CorReg rules !"){
           .'    '.   '.
          (_CorReg____)____)
   ")
-  
+  } 
 }

### 2014-08-27 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@820 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@821 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@823 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@828 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions.
-As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariates in $\boldsymbol{J}_p$ or not.
+As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including those used as predictor covariates in $\boldsymbol{J}_p$.
 An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}.
@@
 			 \\
-			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
+			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j =1}^{d_r}\sigma_{j}\beta_{J_r^j} \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
 	\end{eqnarray}
@@
 \begin{equation}
-\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
+\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J_p}^{\{1,\dots,d_r \}\setminus \{j\}}=J_p^{\{1,\dots,d_r \}\setminus \{j\}} ,\tilde{J}_p^{j}=J_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \setminus J_p^j\right\}.
 \end{equation}
@@
 \begin{equation}
-\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
+\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r,\tilde{J_p}^{\{1,\dots,d_r \}\setminus \{j\}}=J_p^{\{1,\dots,d_r \}\setminus \{j\}} , \tilde{J}_p^{j}=J_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in  J_p^j \right\}.
 \end{equation}	

### 2014-08-28 www/article.tex
up nouvelles notations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@830 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{hyp}\label{H2}
-In addition, we make the complementary hypothesis that the response covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates.
+In addition, we make the complementary hypothesis that the response covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates of cardinal $d_f=d-d_r=|J_f|$.
 This new assumption allows to obtain very simple sub-regressions sequences, discarding hierarchical ones, in particular uninteresting cyclic sub-regressions.
@@
 
-Decomposing $\boldsymbol{\beta}=(\boldsymbol{\beta}_r',\boldsymbol{\beta}_f')'$ by $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the responses and to the free covariates, we can rewrite (\ref{regressionsimple}):
+Noting $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the responses and to the free covariates, we can rewrite (\ref{regressionsimple}):
 \begin{equation}
@@
 \begin{eqnarray}
-	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j)+  \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
-	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl}
+	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^*)+  \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*,\label{eq:modexpl}
 \end{eqnarray}
+where $\boldsymbol{\alpha}^* \in \mathbb{R}^{d_f}$ with $(\boldsymbol{\alpha}_j^*)_{J_p^j}=\boldsymbol{\alpha}_j $ and $(\boldsymbol{\alpha}_j^*)_{J_f\setminus J_p^j}=\{\bf 0 \} $. \\
 Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariates $\boldsymbol{X}_f$.
@@
 
-\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*=(\boldsymbol{\beta}_r^{*'},\boldsymbol{\beta}_f^{*'})'$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_r^*={\bf 0}$, (\ref{eq:modexpl}) can be rewritten
+\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_{J_r}^*=\boldsymbol{\beta}_{r}^*={\bf 0}$ and $\boldsymbol{\beta}_{J_f}^*=\boldsymbol{\beta}_{f}^*$, (\ref{eq:modexpl}) can be rewritten
 \begin{equation}
@@
 Let a simple case with $d=5$ covariates $\boldsymbol{X}=(\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{X}^4, \boldsymbol{X}^5)$ independently drawn by four independent Gaussian $\mathcal{N}_n(\boldsymbol{0},\boldsymbol{I})$.
-Thus, $J_f=\{1,2,4,5\}$ and $\boldsymbol{X}_f=(\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5)$.
+Thus, $J_f=\{1,2,4,5\}$, $d_f=4$ and $\boldsymbol{X}_f=(\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5)$.
 Let also $d_r=1$ response covariate $\boldsymbol{X}^3|\boldsymbol{X}^1,\boldsymbol{X}^2=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{\varepsilon}_1$ where $\boldsymbol{\varepsilon}_1\sim{\mathcal{N}_n(\boldsymbol{0},\sigma_1^2\boldsymbol{I})}$.
@@
 	\begin{equation}
-		1-R^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon}_1)}{\operatorname{Var}(\boldsymbol{X}_3)}=\frac{\sigma_1^2}{\sigma_1^2+2}.
+		1-R^2=\frac{\operatorname{Var}(\boldsymbol{\varepsilon}_1)}{\operatorname{Var}(\boldsymbol{X}^3)}=\frac{\sigma_1^2}{\sigma_1^2+2}.
 	\end{equation}

### 2014-08-28 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@831 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@832 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The aim is now to use the model of linear sub-regressions $\boldsymbol{S}$ (that we assume to be known in this part) between some covariates of $\boldsymbol{X}$ to obtain a linear regression on $\boldsymbol{Y}$ relying only on uncorrelated variables $\boldsymbol{X}_f$.
-The way to proceed is to marginalize the joint distribution of $\{(\boldsymbol{Y},\boldsymbol{X}_f) |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ to obtain the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ depending only on uncorrelated variables $\boldsymbol{X}_f$:
+The way to proceed is to marginalize the joint distribution of $\{(\boldsymbol{Y},\boldsymbol{X}_r) |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ to obtain the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$ depending only on uncorrelated variables $\boldsymbol{X}_f$:
 \begin{equation}\label{eq:marginal}
@@
 \end{eqnarray}
-where $\boldsymbol{\alpha}^* \in \mathbb{R}^{d_f}$ with $(\boldsymbol{\alpha}_j^*)_{J_p^j}=\boldsymbol{\alpha}_j $ and $(\boldsymbol{\alpha}_j^*)_{J_f\setminus J_p^j}=\{\bf 0 \} $. \\
+where $\boldsymbol{\alpha}^* \in \mathbb{R}^{d_f}$ with $(\boldsymbol{\alpha}_j^*)_{J_p^j}=\boldsymbol{\alpha}_j $ and $(\boldsymbol{\alpha}_j^*)_{J_f\setminus J_p^j}=\bf 0  $. \\
 Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariates $\boldsymbol{X}_f$.

### 2014-08-28 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@833 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{equation}\label{eq:marginal}
-\mathbb{P}(\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}) = \int_{{\mathbb{R}^{d_r}}}\mathbb{P}(\boldsymbol{Y}| \boldsymbol{X}_f,\boldsymbol{X}_r,\boldsymbol{S};\boldsymbol{\beta},\sigma_Y^2) \mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma^2}) d\boldsymbol{X}_r.
+\mathbb{P}(\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2) = \int_{{\mathbb{R}^{d_r}}}\mathbb{P}(\boldsymbol{Y}| \boldsymbol{X}_f,\boldsymbol{X}_r,\boldsymbol{S};\boldsymbol{\beta},\sigma_Y^2) \mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma}^2) d\boldsymbol{X}_r.
 \end{equation}
@@
 \begin{equation}\label{eq:H3}
-\mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma^2}) = \prod_{j=1}^{d_r} \mathbb{P}(\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j).
+\mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma}^2) = \prod_{j=1}^{d_r} \mathbb{P}(\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j).
 \end{equation}
@@
 \end{equation}
-Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}_Y$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$:
+Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}_Y$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2\}$:
 \begin{eqnarray}
-	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^*)+  \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
+	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^*)+  \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
 	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*,\label{eq:modexpl}
@@
 \begin{equation}
-	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
+	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
 \end{equation}

### 2014-09-03 www/article.tex
up tilde
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@834 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@836 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up deut part1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@837 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{eqnarray}
-where $\boldsymbol{\alpha}^* \in \mathbb{R}^{d_f}$ with $(\boldsymbol{\alpha}_j^*)_{J_p^j}=\boldsymbol{\alpha}_j $ and $(\boldsymbol{\alpha}_j^*)_{J_f\setminus J_p^j}=\bf 0  $. \\
+where $\boldsymbol{\alpha}^*_j \in \mathbb{R}^{d_f}$ with $(\boldsymbol{\alpha}_j^*)_{J_p^j}=\boldsymbol{\alpha}_j $ and $(\boldsymbol{\alpha}_j^*)_{J_f\setminus J_p^j}=\bf 0  $. \\
 Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariates $\boldsymbol{X}_f$.
@@
 	\begin{equation}
-	\mathbb{P}(\boldsymbol{S}^+|\mathcal{V}(\boldsymbol{S})) \sim \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
+	\mathbb{P}(\boldsymbol{S}^+|\mathcal{V}(\boldsymbol{S})) = \frac{\exp{(-\mbox{{\sc bic}}_*(\cdot))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}}.
 	\end{equation}
@@
 All sub-regressions are of length two ($\forall j=1,\ldots,d_r, d_p^j=2$ and we have $d_r=16$ sub-regressions.
-The datasets are then scaled, so that covariates $\boldsymbol{X}_f$ for avoiding large distortions for variances or for means.	Different sample sizes  $n\in (30,50,100,400)$ are chosen, thus considering experiments in both situations $n<p$ and $n>p$.
+The datasets are then scaled, so that covariates $\boldsymbol{X}_r$ avoid large distortions for variances or for means.	Different sample sizes  $n\in (30,50,100,400)$ are chosen, thus considering experiments in both situations $n<d$ and $n>d$.
 In all figures, the thickness of the lines will represent various values of $n$, the thicker being the greater.

### 2014-09-11 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@841 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 In order to take into account the covariates correlation problem, we make the hypothesis correlation between covariates is {\it only} the consequence that some covariates {\it linearly} depend on some other covariates.
-More precisely, there are $d_{r}\geq 0$ such ``sub-regressions'', each sub-regression $j=1,\ldots,d_{r}$ having the covariate $\boldsymbol{X}^{J_{r}^j}$ as {\it response} variable ($J_{r}^j\in\{1,\ldots,p\}$ and $J_{r}^j\neq J_{r}^{j'}$ if $j\neq j'$) and having the $d_p^j>0$ covariates $\boldsymbol{X}^{J_{p}^j}$  as {\it predictor} variables ($J_{p}^j\subset\{1,\ldots,d\} \backslash J_{r}^j$ and $d_p^j=|J_{p}^j|$ the cardinal of $J_{p}^j$):
+More precisely, there are $d_{r}\geq 0$ such ``sub-regressions'', each sub-regression $j=1,\ldots,d_{r}$ having the covariate $\boldsymbol{X}^{J_{r}^j}$ as {\it response} variable ($J_{r}^j\in\{1,\ldots,d\}$ and $J_{r}^j\neq J_{r}^{j'}$ if $j\neq j'$) and having the $d_p^j>0$ covariates $\boldsymbol{X}^{J_{p}^j}$  as {\it predictor} variables ($J_{p}^j\subset\{1,\ldots,d\} \backslash J_{r}^j$ and $d_p^j=|J_{p}^j|$ the cardinal of $J_{p}^j$):
 \begin{equation}
@@
 \end{equation}
-where $\boldsymbol{\alpha}_j\in{\mathbb{R}^{d_r^j}}$ (${\alpha}_j^h\neq 0$ for all $j=1,\ldots,d_r$ and $h=1,\ldots,d_p^j$) and $\boldsymbol{\varepsilon}_j \sim\mathcal{N}_n(\boldsymbol{0},\sigma^2_j\boldsymbol{I})$.
+where $\boldsymbol{\alpha}_j\in{\mathbb{R}^{d_r^j}}$ (${\alpha}_j^h\neq 0$ for all $j=1,\ldots,d_r$ and $h=1,\ldots,d_p^j$) and $\boldsymbol{\varepsilon}_j \sim\mathcal{N}_n(\boldsymbol{0},\sigma^2_j\boldsymbol{I})$.cf
 \end{hyp}

### 2014-09-11 pkg/CorReg/R/correg.R
up brian
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@842 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -7,7 +7,7 @@
 #' @import mclust 
 #' @import ridge
 #' @import MASS
-#' @import parcor
+# ' @import parcor
 #' @import  clere
 #' @import spikeslab
 #' @import  rpart
@@ -61,6 +61,7 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, 
   if(select=="NULL"){
     returning=FALSE
   }
+  if(select=="adalasso"){require(parcor)}
   if(final){pred=TRUE}
 if(explnew){compl=TRUE;expl=TRUE}
   criterion = criterion[1]

### 2014-09-11 pkg/CorReg/src/rechercheZ_rejet.cpp

@@ -5,6 +5,7 @@
 #include <iostream>
 #include <string>
 #include <math.h>
+#include <Rmath.h>
 #include <stdlib.h>     /* exit, EXIT_FAILURE */
 using namespace Rcpp ;
 using namespace Eigen;
@@ -131,7 +132,7 @@ SEXP rechercheZ_rejet(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP 
   //liste candidats (couples [i,j])
     if (imet_tirage[0]==0)//methode de changement de la ligne et de la colonne
     {
-      k=rand() % p;//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
+      k=int(runif(0,p)[0]);//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
       for(i=0;i<p;i++)
       {
         if(i!=k)//on ne veut pas avoir de 1 sur la diagonale
@@ -147,7 +148,7 @@ SEXP rechercheZ_rejet(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP 
     }
     else if (imet_tirage[0]==-1)//seulement la colonne
     {
-      k=rand() % p;
+      k=int(runif(0,p)[0]);
       for(i=0;i<p;i++)
       {
         if(i!=k)
@@ -206,8 +207,8 @@ SEXP rechercheZ_rejet(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP 
     {
       for(i=0;i<nbcand;i++)
       {
-        rand1=rand() % p;//nombres aleatoire pour avoir le numero de la ligne
-        rand2=rand() % p;//nombres aleatoire pour avoir le numero de la colonne
+        rand1=int(runif(0,p)[0]);//nombres aleatoire pour avoir le numero de la ligne
+        rand2=int(runif(0,p)[0]);//nombres aleatoire pour avoir le numero de la colonne
         if (rand1==rand2)//on ne veut pas de 1 sur la diagonale
         {
           if(rand1<p-1)

### 2014-09-11 pkg/CorReg/src/rechercheZ.cpp
up brian
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@842 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -3,6 +3,7 @@
 #include <iostream>
 #include <string>
 #include <math.h>
+#include <Rmath.h>
 using namespace Rcpp ;
 using namespace Eigen;
 using namespace std;
@@ -19,6 +20,7 @@ SEXP rechercheZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP method
   const Map<MatrixXd> matX(as<Map<MatrixXd> >(X));//X
   const Map<VectorXd> Bic_vide_vect(as<Map<VectorXd> >(bic_vide_vect));//bic_vide_vect
   
+    
   Rcpp::NumericVector met_tirage(methode_tirage),met_newZ(methode_newZ),met_BIC(methode_BIC),rmax(Rmax),P2max(p2max),maxiter(Maxiter),Plot(plot),Best(best),Better(better),Random(random),Bla(bla),Nb_opt_max(nb_opt_max);
   typedef Rcpp::NumericVector::iterator vec_iterator;
   vec_iterator imet_tirage = met_tirage.begin(),imet_newZ=met_newZ.begin(),imet_BIC=met_BIC.begin(),irmax=rmax.begin(),ip2max=P2max.begin(),imaxiter=maxiter.begin(),iplot=Plot.begin(),ibest=Best.begin(),ibetter=Better.begin(),irandom=Random.begin(),ibla=Bla.begin(),inb_opt_max=Nb_opt_max.begin();
@@ -54,7 +56,8 @@ SEXP rechercheZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP method
   Eigen::VectorXd etape (imaxiter[0]);//vecteur qui stock le type de changement de chaque etapes
   bool station;//permet de savoir si on est stationnaire ou non
 
-    
+ 
+   
   //initialisation
   Zopt = MatrixXd::Zero(p,p);
 //somme a la main
@@ -115,7 +118,7 @@ SEXP rechercheZ(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP method
   //liste candidats (couples [i,j])
     if (imet_tirage[0]==0)//methode de changement de la ligne et de la colonne
     {
-      k=rand() % p;//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
+      k= int(runif(0,p)[0]);//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
[…]

### 2014-09-11 pkg/CorReg/src/rechercheZ_relax.cpp

@@ -4,6 +4,8 @@
 #include <iostream>
 #include <string>
 #include <math.h>
+#include <Rmath.h>
+
 using namespace Rcpp ;
 using namespace Eigen;
 using namespace std;
@@ -126,7 +128,7 @@ SEXP rechercheZ_relax(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP 
   //liste candidats (couples [i,j])
     if (imet_tirage[0]==0)//methode de changement de la ligne et de la colonne
     {
-      k=rand() % p;//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
+      k=int(runif(0,p)[0]);//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
       for(i=0;i<p;i++)
       {
         if(i!=k)//on ne veut pas avoir de 1 sur la diagonale
@@ -142,7 +144,7 @@ SEXP rechercheZ_relax(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP 
     }
     else if (imet_tirage[0]==-1)//seulement la colonne
     {
-      k=rand() % p;
+      k=int(runif(0,p)[0]);
       for(i=0;i<p;i++)
       {
         if(i!=k)
@@ -201,8 +203,8 @@ SEXP rechercheZ_relax(SEXP X,SEXP Z,SEXP bic_vide_vect,SEXP methode_tirage,SEXP 
     {
       for(i=0;i<nbcand;i++)
       {
-        rand1=rand() % p;//nombres aleatoire pour avoir le numero de la ligne
-        rand2=rand() % p;//nombres aleatoire pour avoir le numero de la colonne
+        rand1=int(runif(0,p)[0]);//nombres aleatoire pour avoir le numero de la ligne
+        rand2=int(runif(0,p)[0]);//nombres aleatoire pour avoir le numero de la colonne
         if (rand1==rand2)//on ne veut pas de 1 sur la diagonale
         {
           if(rand1<p-1)

### 2014-09-11 pkg/CorReg/src/rechercheZ_sparse_relax.cpp

@@ -4,6 +4,8 @@
 #include <iostream>
 #include <string>
 #include <math.h>
+#include <Rmath.h>
+
 using namespace Rcpp ;
 using namespace Eigen;
 using namespace std;
@@ -164,7 +166,7 @@ SEXP rechercheZ_sparse_relax(SEXP X,SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP bic_vid
   //liste candidats (couples [i,j])
     if (imet_tirage[0]==0)//methode de changement de la ligne et de la colonne
     {
-      k=rand() % p;//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
+      k=int(runif(0,p)[0]);//on tire un numero aleatoire pour savoir quel est la ligne et la colonne de candidats
       for(i=0;i<p;i++)
       {
         if(i!=k)//on ne veut pas avoir de 1 sur la diagonale
@@ -180,7 +182,7 @@ SEXP rechercheZ_sparse_relax(SEXP X,SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP bic_vid
     }
     else if (imet_tirage[0]==-1)//seulement la colonne
     {
-      k=rand() % p;
+      k=int(runif(0,p)[0]);
       for(i=0;i<p;i++)
       {
         if(i!=k)
@@ -195,8 +197,8 @@ SEXP rechercheZ_sparse_relax(SEXP X,SEXP Zi,SEXP Zj,SEXP Si,SEXP Sj,SEXP bic_vid
     {
       for(i=0;i<nbcand;i++)
       {
-        rand1=rand() % p;//nombres aleatoire pour avoir le numero de la ligne
-        rand2=rand() % p;//nombres aleatoire pour avoir le numero de la colonne
+        rand1=int(runif(0,p)[0]);//nombres aleatoire pour avoir le numero de la ligne
+        rand2=int(runif(0,p)[0]);//nombres aleatoire pour avoir le numero de la colonne
         if (rand1==rand2)//on ne veut pas de 1 sur la diagonale
         {
           if(rand1<p-1)

### 2014-09-26 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@867 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@869 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 		\begin{equation}
-			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
+			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^* \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
 		\end{equation}
@@
 
+		\section*{Appendix}\label{sectionident}
+		\section{Identifiability}
+		\subsection{Definition}
+		We call identifiability:
+\begin{eqnarray}
+	\nexists (\boldsymbol{S},\tilde{\boldsymbol{S}}) \in \mathcal{S}_d\times \mathcal{S}_d \textrm{ with } \boldsymbol{S} \neq \tilde{\boldsymbol{S}} \textrm{ and }
+	\mathbb{P}(\boldsymbol{X};\boldsymbol{S})=\mathbb{P}(\boldsymbol{X};\tilde{\boldsymbol{S}})
+ \end{eqnarray}
+ To avoid label-switching consideration, we suppose here (without loss of generality) that $\boldsymbol{J}_r$ is ordered by ascending order of the labels of the covariates.
+	Hence identifiability is paired with the hypotheses we made on $ \mathcal{S}_d$.
+It is not sufficient to find a structure of linear sub-regression, the structure also has to verify hypotheses \ref{H1} to \ref{H3} (uncrossing rule + dependencies exhaustively described by the structure and then independence between the conditional response covariates).
+As a consequence, the covariance between two covariates is not null if and only if these covariates are linked by some sub-regressions.
+
+		\subsection{Sufficient condition for identifiability}
+
+\paragraph{Identifability criterion:} The model $\boldsymbol{S}$ is identifiable if
+			\begin{equation}
+				\forall j \in \{1,\dots,d_r\},  d_p^j>1\label{identcondition}.
+			\end{equation}
+That is to have at least two regressors in each sub-regression.\\
+
+To prove the sufficiency of this condition for identifiability we rely on the following lemma.
+
+\paragraph{Lemma:} With $\boldsymbol{X}$ and $\boldsymbol{S}$ following hypotheses \ref{H1} to \ref{H3}, covariance between two distinct covariates does differ from $0$ in only two cases:
+\begin{enumerate}
+	\item One of the two variables is a regressor of the other in a sub-regression	
+	\begin{eqnarray}
+		j \in \{1,\dots,d_r\}, i \in J_p^j \textrm{ then } \operatorname{cov}(\boldsymbol{X}^{i},\boldsymbol{X}^{J_r^j})\neq 0
+	\end{eqnarray}
+	\item Both variables are regressed by a common covariate in their respective sub-regression:
+	\begin{eqnarray}
+		\exists k \in J_f, \exists (i,j)\in \{1,\dots,d_r\}\times \{1,\dots,d_r\} \textrm{ with } i\neq j,&&  k\in J_p^i \textrm{ and } k\in J_p^j \\
+		\textrm{ then } \operatorname{cov}(\boldsymbol{X}^{J_r^i},\boldsymbol{X}^{J_r^j})&\neq &0 \nonumber
+	\end{eqnarray}
+\end{enumerate}
+
+\paragraph{proof of the lemma:}
+The two cases lead immediately to non-zero covariance so we just look at other combinations of covariates.
+\begin{itemize}
+	\item if $\exists (i,j)\in \{1,\dots,d_r\}\times \{1,\dots,d_r\}, \operatorname{cov}(\boldsymbol{X}^{J_r^i},\boldsymbol{X}^{J_r^j})\neq 0$ then hypothesis \ref{H2} (uncrossing rule) guarantee that the two covariates are not in a same sub-regression so the covariance must come from the noises of the sub-regression but hypothesis \ref{H3} say that they are independent.
+The only remaining case is then the second case of the lemma: common covariate in the sub-regressions.
+	\item if $(i,j)\in J_f\times J_f$ then  $\operatorname{cov}(\boldsymbol{X}^i,\boldsymbol{X}^j)=0 $ because covariates in $\boldsymbol{X}_f$ are orthogonal (by hypotheses \ref{H1} and \ref{H2}).
+	\item if $j \in \{1,\dots,d_r\}, i \in J_f$ and $\operatorname{cov}(\boldsymbol{X}^{J_r^j},\boldsymbol{X}^i)\neq 0$ then $i\in J_p^j$ by hypotheses \ref{H1} and \ref{H3} and equation \ref{eq:H3}.
+\end{itemize}
+$\square$
+
+\paragraph{proof of the identifiability criterion:} We suppose that \ref{identcondition} is verified and the model is not identifiable:
+\begin{eqnarray}
+	\exists (\boldsymbol{S},\tilde{\boldsymbol{S}}) \in \mathcal{S}_d\times \mathcal{S}_d \textrm{ with } \boldsymbol{S} \neq \tilde{\boldsymbol{S}} \textrm{ and }
+	\mathbb{P}(\boldsymbol{X};\boldsymbol{S})=\mathbb{P}(\boldsymbol{X};\tilde{\boldsymbol{S}})
+ \end{eqnarray}
+ $\tilde{\boldsymbol{S}}=(\tilde{\boldsymbol{J}}_r,\tilde{\boldsymbol{J}}_p)$ contains $\tilde{d}_r$ sub-regressions and is characterized by $\tilde{\boldsymbol{J}}_r=(\tilde{J}_r^1,\dots,\tilde{J}_r^{\tilde{d}_r}),\tilde{\boldsymbol{J}}_p=(\tilde{J}_p^1,\dots,\tilde{J}_p^{\tilde{d}_r})$.\\
+Because $\boldsymbol{S}\neq \tilde{\boldsymbol{S}} $ we have $\boldsymbol{J}_r\neq \tilde{\boldsymbol{J}}_r$ or $\boldsymbol{J}_p\neq \tilde{\boldsymbol{J}}_p$.
+\begin{itemize}
+\item If $\boldsymbol{J}_r = \tilde{\boldsymbol{J}}_r$ and $\boldsymbol{S}\neq \tilde{\boldsymbol{S}}$ then one sub-regression as a predictor that stands only in one of the two structures.
+We suppose (without loss of generality) that $\exists j \in \{1,\dots,d_r\}$ for which
+$\exists i \in J_p^j$ with $i \notin \tilde{J}_p^j$ so $\operatorname{cov}_{S}(\boldsymbol{X}^{J_r^j},\boldsymbol{X}^{i})\neq 0$ and $\operatorname{cov}_{\tilde{S}}(\boldsymbol{X}^{J_r^j},\boldsymbol{X}^{i})= 0$ (from the lemma) so the two structure don't give the same joint distribution, leading to a contradiction.
+\item If $\boldsymbol{J}_r\neq \tilde{\boldsymbol{J}}_r$ then one of the two models has a sub-regression that is not in the other. We suppose (without loss of generality) that
+ $\exists J_r^j \in J_r$ with $J_r^j \notin \tilde{J}_r$ then $J_r^j\in \tilde{J_f}$ (recall $J_f=\{1,\dots,d \}\setminus J_r$).
+We note that $J_r^j \in J_r$ means $\exists k_1\neq k_2, \{k_1,k_2 \}\subset J_p^j \subset J_f$.
+Then $\operatorname{cov}_{\boldsymbol{S}}(\boldsymbol{X}^{J_r^j},\boldsymbol{X}^{k_1})\neq 0$ and $\operatorname{cov}_{\boldsymbol{S}}(\boldsymbol{X}^{J_r^j},\boldsymbol{X}^{k_2})\neq 0 $ so by the lemma $k_1$ and $k_2$ are responses variables in $\tilde{\boldsymbol{S}}$: $\exists (l_1,l_2) \in \{1,\dots,\tilde{d}_r \}\times \{1,\dots,\tilde{d}_r \}, \tilde{J}_r^{l_1}=k_1,\tilde{J}_r^{l_2}=k_2$ and $J_r^j$ is a regressor of $k_1$ and $k_2$: $J_r^j \in J_p^{l_1}, J_r^j \in J_p^{l_2}$ thus $\operatorname{cov}_{\tilde{\boldsymbol{S}}}(\boldsymbol{X}^{k_1},\boldsymbol{X}^{k_2})\neq 0$ that is not possible because $ \{k_1,k_2 \}\subset J_p^j \subset J_f$ and covariates in $\boldsymbol{X}^{J_f}$ are orthogonal by hypotheses.
+\end{itemize}
+
+Finally, condition \ref{identcondition} is sufficient for identifiability of $\boldsymbol{S}$.
+  	$\square$
+
+\paragraph{Remark:} Because sub-regressions with at least two regressors are identifiable, 
+the only non-identifiable sub-regressions could be those with only one regressor, leading only to pairwise correlations that can be seen directly in the correlation matrix.
+Such sub-regression can be permuted without any impact on interpretation so such trivial sub-regression are not a problem even if they may occur with real datasets.
+One more thing: exact sub-regression with at least two sub-regressors are identifiable with our hypotheses.
+
 \end{document}

### 2014-11-03 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@902 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
-\item As a consequence of Hypothesis~\ref{H1} and~\ref{H2} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
+%\item As a consequence of Hypothesis~\ref{H1} and~\ref{H2} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
 \item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals.
@@
 In other words, it is possible to see $(\ref{eq:modexpl})$ as a kind of {\it pretreatment} for decorrelating covariates, while assuming nothing on the subsequent estimate process.
-
+\paragraph{Remark}
+\begin{itemize}
+%\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
+\item As a consequence of Hypothesis~\ref{H1} to~\ref{H3} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated (see the Lemma in the Appendix about identifiability).
+%\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals.
+Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
+%\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}.
+\end{itemize} 
 In the following, we will denote by {\sc CorReg} the new proposed strategy.
@@
 			 \\
-			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j =1}^{d_r}\sigma_{j}\beta_{J_r^j} \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
+			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^* \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
 	\end{eqnarray}
@@
 \begin{hyp}\label{H4}
-All variates $\boldsymbol{X}^j$ with $j \in J_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
+All covariates $\boldsymbol{X}^j$ with $j \in J_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
 \begin{equation}
@@
 $\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as before.
-We design a unique, possibly non linear, sub-regression $\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2+\boldsymbol{X}_3+ \varepsilon_1$.
-The matrix $\boldsymbol{X}$ is then scaled and we set $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\varepsilon_Y$.
+We design a unique, possibly non linear, sub-regression $\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2+\boldsymbol{X}_3+ \boldsymbol{\varepsilon}_1$.
+The matrix $\boldsymbol{X}$ is then scaled and we set $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\boldsymbol{\varepsilon}_Y$.
 We let $a$ vary between $0$ and $10$ to increase progressively the non-linear part of the sub-regression.

### 2014-11-28 pkg/CorReg/R/CorReg-package.R
up bug 0 inj pred select
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@949 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -37,12 +37,12 @@
 #'    bicopt_vrai=BicZ(X=X_appr,Z=TrueZ,Bic_null_vect=Bic_null_vect)
 #'    sum(bicopt_vect);sum(bicopt_vrai)
 #'    
+#'    #Structure comparison
+#'    compZ=compare_struct(trueZ=TrueZ,Zalgo=hatZ)#qualitative comparison
+#'    
 #'    #interpretation of found and true structure ordered by increasing R2
 #'    readZ(Z=hatZ,crit="R2",X=X_appr,output="all",order=1)# <NA>line : name of subregressed covariate
 #'    readZ(Z=TrueZ,crit="R2",X=X_appr,output="all",order=1)# <NA>line : name of subregressed covariate
-#'    
-#'    #Structure comparison
-#'    compZ=compare_struct(trueZ=TrueZ,Zalgo=hatZ)#qualitative comparison
 #'    
 #'    #Regression coefficients estimation
 #'     select="NULL"

### 2014-11-28 pkg/CorReg/R/meilleur_lars.R

@@ -40,8 +40,10 @@ meilleur_lars<-function(lars=lars,X=X,Y=Y,mode=c("MSE","BIC"),intercept=TRUE,K=N
       qui=which.min(vrais_vect)
       MSE=vrais_vect[qui]
       qui=which(coef[qui,]!=0)
-      A[c(intercept,qui+intercept)]=OLS(X=as.matrix(X[,qui]),Y=Y,intercept=intercept)$beta
-    }
+      if(length(qui)>0){
+         A[c(intercept,qui+intercept)]=OLS(X=as.matrix(X[,qui]),Y=Y,intercept=intercept)$beta
+      }#sinon ne garde rien donc on laisse A sur sa valeur initiale nulle
+   }
     return(list(A=A,CVMSE=MSE))    
   }
 }

### 2014-12-05 pkg/CorReg/R/CorReg-package.R
up imports
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@960 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -62,3 +62,4 @@
 #'   abline(h=MSE_complete,col="red")
 #'    }
 NULL
+#la puissance CorReg!!!!!!!!!

### 2014-12-05 pkg/CorReg/R/correg.R

@@ -1,19 +1,19 @@
 #' Estimates the response variable using a structure
-#'  @import Rcpp
-# '   @import Rmixmod
+#' @import Rcpp
+#' @import Rmixmod
 #' @import lars
 #' @import elasticnet
 #' @import Matrix
-# ' @import mclust 
+#' @import mclust 
 #' @import ridge
 #' @import MASS
 # ' @import parcor
-# ' @import  clere
+# '  @import  clere
 # ' @import spikeslab
 #' @import  rpart
-# ' @import  stats
 #' @import corrplot
 #' @import mvtnorm
+# ' @import rtkpp
 #' @useDynLib CorReg
 #' @export
 #' @param B the (p+1)xp matrix associated to Z and that contains the parameters of the sub-regressions
@@ -41,6 +41,7 @@
 #' @param g number of group of variables for clere
 #' @param explnew select the number of sub-regression to take into account (by AIC on the corresponding final model)
 #' 
+#Attention cette fonction dégage une puissance phénoménale (it's over 9000!)
 correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, explnew=FALSE,
                 pred = FALSE,prednew=FALSE,
                 select = "lar",

### 2014-12-08 pkg/CorReg/R/correg.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@969 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -35,20 +35,16 @@
 #' @param Y_test response for the validation sample
 #' @param intercept boolean. If FALSE intercept will be set to 0 in each model.
 #' @param alpha Coefficients of the explicative model to coerce the predictive step. if not NULL explicative step is not computed.
-#' @param prednew alternate optimisation for predictive
-#' @param nbalter number of alternance for prednew
-#' @param deltamin criterion to stop alternance 
 #' @param g number of group of variables for clere
-#' @param explnew select the number of sub-regression to take into account (by AIC on the corresponding final model)
 #' 
 #Attention cette fonction dégage une puissance phénoménale (it's over 9000!)
 correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, explnew=FALSE,
-                pred = FALSE,prednew=FALSE,
+                pred = FALSE,
                 select = "lar",
                 criterion = c("MSE", "BIC"),
                 X_test = NULL, Y_test = NULL, intercept = TRUE, 
                 K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=FALSE,final=FALSE,
-                nbalter=10,deltamin=0.01,alpha=NULL,g=5) 
+                alpha=NULL,g=5) 
 {
   res = list()
   X = 1*as.matrix(X)
@@ -184,92 +180,7 @@ if(explnew){compl=TRUE;expl=TRUE}
     res$expl$BIC = BicTheta(X = X, Y = Y, intercept = intercept, beta = A_expl)
     res$expl$AIC=mon_AIC(theta=res$expl$A,Y=Y,X=X,intercept=intercept) 
     
-    #nouvel explicatif
-    if(explnew){#selection sur explicatif####
-      qui = WhoIs(Z = Z)
-      I1 = qui$I1
-      I2 = qui$I2
-      R2_vect=R2Z(Z=Z,X=X,adj=TRUE)#evaluation des R2
-      R2_vect=R2_vect[R2_vect!=0]
-      #tri par R2 => nouvel I2 ordonné
-      I2=I2[order(R2_vect,decreasing=FALSE)]#on commence par garder ce qui est mal expliqué par le reste (donc forte perte et peu de corrélations)
-      #initialisation du résultat final (par le modèle complet déjà calculé, + AIC associé)
-      res$expl2$A=res$expl$A
-      resopt=res$expl$A
-      AICopt=mon_AIC(theta=resopt,Y=Y,X=X,intercept=intercept)
-      for (iexplnew in 1:length(I2)){
[…]

### 2014-12-08 pkg/CorReg/R/correg.R
up cleaning correg
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@970 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -30,7 +30,6 @@
 #' @param groupe a vector to define the groups used for cross-validation (to obtain a reproductible result)
 #' @param Amax the maximum number of covariates in the final model
 #' @param returning boolean : second predictive step (selection on I1 knowing I2 coefficients)
-#' @param final boolean : recompute estimators without selection on the remaining parameters of the predictive model
 #' @param X_test validation sample
 #' @param Y_test response for the validation sample
 #' @param intercept boolean. If FALSE intercept will be set to 0 in each model.
@@ -43,7 +42,7 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, 
                 select = "lar",
                 criterion = c("MSE", "BIC"),
                 X_test = NULL, Y_test = NULL, intercept = TRUE, 
-                K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=FALSE,final=FALSE,
+                K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=FALSE,
                 alpha=NULL,g=5) 
 {
   res = list()
@@ -60,8 +59,6 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, 
     returning=FALSE
   }
   if(select=="adalasso"){require(parcor)}
-  if(final){pred=TRUE}
-if(explnew){compl=TRUE;expl=TRUE}
   criterion = criterion[1]
   if (is.null(Amax)) {
     Amax = ncol(X) + 1
@@ -305,24 +302,5 @@ if(explnew){compl=TRUE;expl=TRUE}
        res$expl2$A = res$compl$A
     }
   }
-   if(pred & final){
-      res$final$A=res$pred$A
-      A_pred=res$pred$A
-      if(intercept){
-         quifinal=which(A_pred[-1]!=0)
-      }else{
-         quifinal=which(A_pred!=0)
-      }
-      if(length(quifinal)==0){
-         res$final=res$pred
-      }else{
-         resA_final=correg(X=X[,quifinal],Y=Y,returning=F,final=F,groupe=groupe,K=K,intercept=intercept,criterion=criterion,select=select,compl=TRUE,expl=FALSE,pred=FALSE,explnew=FALSE)
-         quifinal=which(A_pred!=0)
-         res$final$A[quifinal]=resA_final$compl$A
[…]

### 2014-12-09 pkg/CorReg/R/correg.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@979 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -45,7 +45,17 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, 
                 K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=FALSE,
                 alpha=NULL,g=5) 
 {
-  res = list()
+  if(is.null(X)){
+     dat<- data.frame(t=seq(0, 2*pi, by=0.1) )
+     xhrt <- function(t) 16*sin(t)^3
+     yhrt <- function(t) 13*cos(t)-5*cos(2*t)-2*cos(3*t)-cos(4*t)
+     dat$y=yhrt(dat$t)
+     dat$x=xhrt(dat$t) 
+     with(dat, polygon(x,y, col="red"))   
+     points(c(10,-10, -15, 15), c(-10, -10, 10, 10), pch=169, font=5)
+     title(main="I Love CorReg !")
+  }
+   res = list()
   X = 1*as.matrix(X)
   K = abs(K)
   K = min(K, nrow(X))

### 2014-12-09 pkg/CorReg/R/correg.R
I love CorReg !
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@980 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -37,7 +37,7 @@
 #' @param g number of group of variables for clere
 #' 
 #Attention cette fonction dégage une puissance phénoménale (it's over 9000!)
-correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, explnew=FALSE,
+correg<-function (X = NULL, Y = NULL, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, explnew=FALSE,
                 pred = FALSE,
                 select = "lar",
                 criterion = c("MSE", "BIC"),
@@ -50,10 +50,12 @@ correg<-function (X = X, Y = Y, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, 
      xhrt <- function(t) 16*sin(t)^3
      yhrt <- function(t) 13*cos(t)-5*cos(2*t)-2*cos(3*t)-cos(4*t)
      dat$y=yhrt(dat$t)
-     dat$x=xhrt(dat$t) 
+     dat$x=xhrt(dat$t)
+     with(dat, plot(x,y, type="l"))  
      with(dat, polygon(x,y, col="red"))   
      points(c(10,-10, -15, 15), c(-10, -10, 10, 10), pch=169, font=5)
      title(main="I Love CorReg !")
+     return("I Love CorReg !")
   }
    res = list()
   X = 1*as.matrix(X)

### 2014-12-09 pkg/CorReg/R/WhoIs.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@981 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,7 +1,7 @@
 #' Give the partition implied by a structure
 # ' 
 #' @param Z the structure (square matrix)
-#' @param I1,I2,I3 wanted outputs
+#' @param I1,I2,I3 wanted outputs (I1= explicative covariates, I2=redundant covariates, I3=independent covariates)
 # ' 
 #'@export 
 WhoIs<-function(Z=Z,I3=F,I2=T,I1=T){

### 2014-12-24 pkg/CorReg/R/correg.R
up compl2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1007 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -35,7 +35,7 @@
 #' @param intercept boolean. If FALSE intercept will be set to 0 in each model.
 #' @param alpha Coefficients of the explicative model to coerce the predictive step. if not NULL explicative step is not computed.
 #' @param g number of group of variables for clere
-#' 
+#' @param compl2 boolean to compute regression (OLS only) upon [X_f,epsilon] instead of [X_f,X_r]
 #Attention cette fonction dégage une puissance phénoménale (it's over 9000!)
 correg<-function (X = NULL, Y = NULL, Z = NULL, B = NULL, compl = TRUE, expl = FALSE, explnew=FALSE,
                 pred = FALSE,
@@ -43,7 +43,7 @@ correg<-function (X = NULL, Y = NULL, Z = NULL, B = NULL, compl = TRUE, expl = F
                 criterion = c("MSE", "BIC"),
                 X_test = NULL, Y_test = NULL, intercept = TRUE, 
                 K = 10, groupe = NULL, Amax = NULL, lambda = 1,returning=FALSE,
-                alpha=NULL,g=5) 
+                alpha=NULL,g=5,compl2=FALSE) 
 {
   if(is.null(X)){
      dat<- data.frame(t=seq(0, 2*pi, by=0.1) )
@@ -126,6 +126,7 @@ correg<-function (X = NULL, Y = NULL, Z = NULL, B = NULL, compl = TRUE, expl = F
     res$compl$BIC = BicTheta(X = X, Y = Y, intercept = intercept, 
                              beta = res$compl$A)
     res$compl$AIC=mon_AIC(theta=res$compl$A,Y=Y,X=X,intercept=intercept) 
+   res$compl$CVMSE = CVMSE(X = X, Y = Y, intercept = intercept, K = K, groupe = groupe)
   }
 #explicatif
   if (sum(Z) != 0 & (expl | pred)) {
@@ -188,6 +189,7 @@ correg<-function (X = NULL, Y = NULL, Z = NULL, B = NULL, compl = TRUE, expl = F
     }
     res$expl$BIC = BicTheta(X = X, Y = Y, intercept = intercept, beta = A_expl)
     res$expl$AIC=mon_AIC(theta=res$expl$A,Y=Y,X=X,intercept=intercept) 
[…]

### 2015-02-08 www/soutenance.tex
up slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1046 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt]{beamer}
\usetheme{Malmoe}
\usepackage[utf8]{inputenc}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
  \usepackage{color}
\let\oldtabular=\tabular
\def\tabular{\small\oldtabular}
\setcounter{tocdepth}{1}
%  \usetheme{Warsaw}
 \graphicspath{{figures/}}
  \author{Clément Théry, Christophe Biernacki, Gaétan Loridant}\institute{ArcelorMittal Dunkerque, Université de Lille 1,équipe M$\Theta$dal Inria}
\title{CorReg}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
%\logo{} 
%\institute{} 
%\date{} 
%\subject{} 
\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\tableofcontents
\end{frame}

\section{Context}
	\subsection{Industrial context}
		\begin{frame}
				  \begin{enumerate}
				\item Steel industry databases.
				\item Goal: To understand and prevent quality problems on finished product, knowing the whole process, \underline{without a priori}.
			\end{enumerate}
	        \begin{center}
	          \begin{tabular}{ccc}
	         \includegraphics[width=70px,height=70px]{liquid.jpg} & \includegraphics[width=70px,height=70px]{Brame1.jpg} & \includegraphics[width=70px,height=70px]{Brame.jpg} \\
	          	\includegraphics[width=70px,height=70px]{ecras_moy.jpg} & \includegraphics[width=70px,height=70px]{tcc2.jpg} & \includegraphics[width=70px,height=70px]{bobines.jpg}
	          \end{tabular}
	        \end{center}
		\end{frame}
	\subsection{Statistical context}
		\begin{frame}{Regression}
			\begin{equation}
				\boldsymbol{Y}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}\label{regressionsimple}
			\end{equation}		
		\end{frame}

		\begin{frame}{OLS}
		\begin{figure}[h!]
	\centering
	\includegraphics[width=250px]{figures/OLS_geometric_interpretation.png}
	\caption{Multiple linear regression with Ordinary Least Squares seen as a projection on the $d-$dimensional hyperplane spanned by the regressors $\boldsymbol{X}$. Public domain image.} \label{geomOLS}
	\end{figure}
		\end{frame}

		\begin{frame}{OLS}
		$\boldsymbol{\beta}$ can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{ols}), that is the unbiased maximum likelihood estimator \cite{saporta2006probabilites,dodge2004analyse}: %As shown in section \ref{sectionOLS}, 
	\begin{equation}
		\boldsymbol{\hat{\beta}}_{OLS}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}\label{betaOLS}
	\end{equation}
	with variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\boldsymbol{\beta}}_{OLS})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}. \label{eq:varOLS}
	\end{equation}
	 In fact it is the Best Linear Unbiased Estimator (BLUE).
	 The theoretical \textsc{mse} is given by
	\begin{equation}
	\textsc{mse}(\hat{\boldsymbol{\beta}}_{OLS})= \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}). \nonumber 
	\end{equation}
		\end{frame}

		\begin{frame}{Ridge}
		Alternative ridge : bien mais pas de sélection donc inacceptable
		\end{frame}

		\begin{frame}{Ridge}
		Alternative ridge : bien mais pas de sélection donc inacceptable
		\end{frame}

		\begin{frame}{LASSO}
		Alternative lasso (et autres) : bien mais problèmes en cas de corrélations
		\end{frame}
		\begin{frame}{LASSO}
		Liste des alternatives
		\end{frame}
		\begin{frame}{ SEM }
		 Modélisation de la structure mais à la main et aucun impact sur l'estimation
		 \end{frame}

		\begin{frame}{ Selvarclust }
			Semble très bien mais n'aboutit pas vers la régression donc on le prolonge en CorReg
		\end{frame}
\section{Proposed Models}
	\begin{frame}{•}

	\end{frame}
	\subsection{Marginal model}
		\begin{frame}{•}

		\end{frame}
	\subsection{Plug-in model}
		\begin{frame}{•}

		\end{frame}
\section{Structure estimation}
	\begin{frame}{•}

	\end{frame}

\section{Results}
	\subsection{Simulation results}
		\begin{frame}{•}

		\end{frame}
	\subsection{Industrial results}
		\begin{frame}{•}

		\end{frame}
\section{Missing values}
	\begin{frame}{•}
	Modèle génératif complet avec dépendances
	\end{frame}
	\begin{frame}{•}
	Explosion des mélanges 
	\end{frame}
	\begin{frame}{•}
	SEM avec Gibbs 
	\end{frame}
	\begin{frame}{•}
	Bic pondéré 
	\end{frame}
	\begin{frame}{•}
	Résultats pourris 
	\end{frame}
\section{Tools}
	\begin{frame}{•}
		Excel, fonctions graphiques, arbres de décision
	\end{frame}

\begin{frame}
\bibliographystyle{apalike}
%\bibliographystyle{plain}
%\nocite{*}
\bibliography{biblio}
\end{frame}

\end{document}

### 2015-02-08 www/soutenance.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1047 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[utf8]{inputenc}
-\usepackage[french]{babel}
 \usepackage[T1]{fontenc}
@@
 \usepackage{graphicx}
-  \usepackage{color}
+  %\usepackage{color}
+  \usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc} 
+\usepackage[english]{babel}
+\usepackage{listings} 
+\usepackage[toc,page]{appendix}
+\usepackage{array}%,multirow,makecell}
+\usepackage{makecell}
+\usepackage{amssymb}
+\usepackage{subfigure}
+\usepackage{placeins}
+\usepackage{graphicx}
+%\usepackage{framed}
+%\usepackage[tikz]{bclogo}
+\usepackage{hyperref}
+\usepackage{url}
+\usepackage{lmodern}
+ % \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}%[usenames,dvipsnames,svgnames,table]
+
 \let\oldtabular=\tabular
@@
 %\subject{}
+
+%%%% debut macro %%%%
+\newenvironment{changemargin}[2]{\begin{list}{}{%
+\setlength{\topsep}{0pt}%
+\setlength{\leftmargin}{0pt}%
+\setlength{\rightmargin}{0pt}%
+\setlength{\listparindent}{\parindent}%
+\setlength{\itemindent}{\parindent}%
+\setlength{\parsep}{0pt plus 1pt}%
+\addtolength{\leftmargin}{#1}%
+\addtolength{\rightmargin}{#2}%
+}\item }{\end{list}}
+%%%% fin macro %%%%
+
 \begin{document}
@@
 			\end{equation}
+			where 	$\boldsymbol{\varepsilon}\sim \mathcal{N}(0,\sigma_Y^2)$	
 		\end{frame}
@@
 	\centering
-	\includegraphics[width=250px]{figures/OLS_geometric_interpretation.png}
+	\includegraphics[width=220px]{figures/OLS_geometric_interpretation.png}
 	\caption{Multiple linear regression with Ordinary Least Squares seen as a projection on the $d-$dimensional hyperplane spanned by the regressors $\boldsymbol{X}$.
@@
 
+\begin{frame}{Running example}
+
+%$d=5$ with four independent scaled Gaussian 
+$\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5  \sim \mathcal{N}(0,1)$  and 
+$\boldsymbol{X}^3=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{\varepsilon}_1$ where 
+$\boldsymbol{\varepsilon}_1\sim{\mathcal{N}(\boldsymbol{0},\sigma_1^2\boldsymbol{I}_n)}$. \\
+Two {\it scenarii} for $\boldsymbol{Y}$:\\
+ $\boldsymbol{\beta}=(1,1,1,1,1)'$ and $\sigma_Y \in \{10,20\}$.
+ \\
+It is clear that $\boldsymbol{X}'\boldsymbol{X}$ will become more ill-conditioned as $\sigma_1$ gets smaller. $R^2$ stands for the coefficient of determination which is here:
+	\begin{equation}\label{defR2}
+	R^2=1-\frac{\operatorname{Var}(\boldsymbol{\varepsilon}_1)}{\operatorname{Var}(\boldsymbol{X}^3)}
+	\end{equation}
+
+\end{frame}
+
+	\begin{frame}
+	\begin{figure}
+	 \centering
+	  \includegraphics[width=250px]{figures/MQEreel/OLScompl.png}
+	  \caption{Evolution of observed Mean Squared error on $\hat{\boldsymbol{\beta}}_{OLS}$ with the strength of the correlations for various sample sizes and strength of regression.
+$d=5$ covariates (running example).
+} \label{MQEOLScompl}
+	\end{figure}	
+\end{frame}		
 		\begin{frame}{Ridge}

### 2015-02-08 www/soutenance.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1048 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 			\end{equation}
-			where 	$\boldsymbol{\varepsilon}\sim \mathcal{N}(0,\sigma_Y^2)$	
+			where 	$\boldsymbol{\varepsilon}\sim \mathcal{N}(0,\sigma_Y^2\boldsymbol{I}_n)$	
 		\end{frame}
@@
 \end{frame}		
-		\begin{frame}{Ridge}
-		Alternative ridge : bien mais pas de sélection donc inacceptable
+		\begin{frame}{Ridge Regression}
+		\cite{hoerl1970ridge,marquardt1975ridge} proposes a possibly biased estimator for $\boldsymbol{\beta}$ that can be written in terms of a parametric $L_2$ penalty:
+	\begin{equation}
+		 \boldsymbol{\hat{\beta}}=\operatorname{argmin}_{\boldsymbol{\beta}} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel \boldsymbol{\beta} \parallel_2^2\leq \eta \textrm{ with } \eta>0
+	\end{equation}
+	But this penalty is not guided by correlations. 
+	The solution of the ridge regression is given by
+	\begin{equation}
+		 \hat{\boldsymbol{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} -\lambda\boldsymbol{I}_n\right) ^{-1}\boldsymbol{X}'\boldsymbol{Y}\label{betaridge}
+	\end{equation}
+ Methods do exist to automatically choose a good value for $\lambda$ \cite{cule2013ridge,er2013systematic} and a R package called {\tt ridge} is on {\sc cran} \cite{packageridge}.
 		\end{frame}
 
-		\begin{frame}{Ridge}
-		Alternative ridge : bien mais pas de sélection donc inacceptable
+		\begin{frame}{Ridge Regression}
+ \begin{figure}
+	 \centering
+	  \includegraphics[width=250px]{figures/MQEreel/ridgecompl.png}
+	  \caption{Evolution of observed Mean Squared error on $\hat{\boldsymbol{\beta}}_{ridge}$ with the strength of the correlations for various sample sizes and strength of regression.
+$d=5$ covariates.
+} \label{MQEridgecompl}
+	\end{figure}
 			\end{frame}
@@
 		\begin{frame}{LASSO}
-		Alternative lasso (et autres) : bien mais problèmes en cas de corrélations
+		The Least Absolute Shrinkage and Selection Operator ({\sc lasso}, \cite{tibshirani1996regression} and \cite{tibshiranilasso}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty to obtain zeros in $\hat{\boldsymbol{\beta}}$ instead of the $L_2$ penalty of the ridge regression:
+		\begin{equation}
+		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda \textrm{ with } \lambda>0 .
+\nonumber
+		\end{equation}
+
 		\end{frame}
 		\begin{frame}{LASSO}
-		Liste des alternatives
+		\begin{figure}[h!]
+			\centering
+			\includegraphics[width=150px]{figures/lasso.png} 
+			\caption{Geometric view of the Penalty for the \textsc{lasso} (left) compared to ridge regression (right) as shown in the book from Hastie \cite{hastie2009elements}} \label{lassogeom}
+		\end{figure}
+		Figure shows the contour of error (red) and constraint function (blue).
+		  The axis stands for the regression coefficients.
 		\end{frame}
+
+\begin{frame}
+
+		Here again we have to choose a value for $\lambda$.
+	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) algorithm offers a very efficient way to obtain the whole \textsc{lasso} path.
+It can be used through the {\tt lars } package on {\sc cran} (\cite{packagelars}).
+\end{frame}
+
+\begin{frame}
+ \begin{figure}
+	 \centering
+	  \includegraphics[width=250px]{figures/MQEreel/larcompl.png}
+	  \caption{Evolution of observed Mean Squared error on $\hat{\boldsymbol{\beta}}_{lar}$ with the strength of the correlations for various sample sizes and strength of regression.
+$d=5$ covariates.
+} \label{MQElarcompl}
+	\end{figure}	
+\end{frame}
+
 		\begin{frame}{ SEM }

### 2015-02-08 www/soutenance.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1049 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{frame}
-
-		Here again we have to choose a value for $\lambda$.
-	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) algorithm offers a very efficient way to obtain the whole \textsc{lasso} path.
-It can be used through the {\tt lars } package on {\sc cran} (\cite{packagelars}).
-\end{frame}
-
-\begin{frame}
  \begin{figure}
@@
 	\end{figure}	
+	{\tt lars } package on {\sc cran} (\cite{packagelars}).
 \end{frame}
@@
 \section{Proposed Models}
-	\begin{frame}{•}
-
+	\begin{frame}{Hypothesis 1}
+	There are $d_{r}\geq 0$ ``sub-regressions'', each sub-regression $j=1,\ldots,d_{r}$ having the covariate $\boldsymbol{X}^{J_{r}^j}$ as {\it response} variable ($J_{r}^j\in\{1,\ldots,p\}$ and $J_{r}^j\neq J_{r}^{j'}$ if $j\neq j'$) and having the $d_p^j>0$ covariates $\boldsymbol{X}^{J_{p}^j}$  as {\it predictor} variables ($J_{p}^j\subset\{1,\ldots,d\} \backslash J_{r}^j$ and $d_p^j=|J_{p}^j|$ the cardinal of $J_{p}^j$):
+\begin{equation}
+\boldsymbol{X}^{J_{r}^j}%|\boldsymbol{X}^{J_{p}^j};\boldsymbol{\alpha}_j,\sigma^2_j
+=\boldsymbol{X}^{J_{p}^j}\boldsymbol{\alpha}_j+\boldsymbol{\varepsilon}_j, \label{eq:SR}
+\end{equation}
+where $\boldsymbol{\alpha}_j\in{\mathbb{R}^{d_r^j}}$ (${\alpha}_j^h\neq 0$ for all $j=1,\ldots,d_r$ and $h=1,\ldots,d_p^j$) and $\boldsymbol{\varepsilon}_j \sim\mathcal{N}_n(\boldsymbol{0},\sigma^2_j\boldsymbol{I})$.
+	\end{frame}
+	\begin{frame}{Hypothesis 2}
+	the response covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates of cardinal $d_f=d-d_r=|J_f|$.
+We call this hypothesis the \underline{uncrossing rule}.
+Then:
+	\begin{equation}
+			\boldsymbol{Y} %{|\boldsymbol{X},\boldsymbol{S}};\boldsymbol{\beta},\sigma_Y^2
+			=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon}_Y. \label{eq:MainR}
+\end{equation}
+	\end{frame}
+	\begin{frame}{Hypotheses 3}
+	We assume that all errors $\boldsymbol{\varepsilon}_Y$ and $\boldsymbol{\varepsilon}_j$ ($j=1,\ldots,d_r$) are {\it mutually independent}.
+It implies in particular that conditional response covariates $\{\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j\}$ are {\it mutually independent}:
+\begin{equation}\label{eq:H3}
+\mathbb{P}(\boldsymbol{X}_r | \boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma}^2) = \prod_{j=1}^{d_r} \mathbb{P}(\boldsymbol{X}^{J_{r}^j}|\boldsymbol{X}^{J_{p}^j},\boldsymbol{S};\boldsymbol{\alpha}_j,\sigma^2_j).
+\end{equation}
 	\end{frame}
 	\subsection{Marginal model}
-		\begin{frame}{•}
-
+		\begin{frame}{Marginal model}
+		 We obtain  for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2\}$:
+\begin{eqnarray}
+	\boldsymbol{Y}%|\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2
+	&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^*)+  \sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
+	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*,\label{eq:modexpl}
+\end{eqnarray}
+where $\boldsymbol{\alpha}^*_j \in \mathbb{R}^{d_f}$ with $(\boldsymbol{\alpha}_j^*)_{J_p^j}=\boldsymbol{\alpha}_j $ and $(\boldsymbol{\alpha}_j^*)_{J_f\setminus J_p^j}=\bf 0 $.
+We define $\boldsymbol{\alpha}^* \in \mathbb{R}^{(d_f \times d_r)}$ to use more compact notations:
+\begin{eqnarray}
+	\boldsymbol{X}_r %|\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\alpha},\boldsymbol{\sigma}^2
+	&=&\boldsymbol{X}_f\boldsymbol{\alpha}^*+\boldsymbol{\varepsilon} \nonumber \\
+	\boldsymbol{Y}%|\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2
+	&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \boldsymbol{\alpha}^*\boldsymbol{\beta_{r}})+ \boldsymbol{\varepsilon}\boldsymbol{\beta_{r}}+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl2}
+\end{eqnarray} 
+Where $\boldsymbol{\varepsilon}$ is the $n\times d_r$ matrix whose columns are the $\boldsymbol{\varepsilon}_j$, the noises of the sub-regressions.
 		\end{frame}
 	\subsection{Plug-in model}
-		\begin{frame}{•}
+		\begin{frame}{Plug-in model}
+%\fcolorbox{jaunefonce}{jauneclair}{\begin{minipage}{1\textwidth}
+\begin{equation}
+	\boldsymbol{\varepsilon}_Y^*%|\boldsymbol{\varepsilon};\boldsymbol{\beta}_r,\sigma_Y^2
+	=\boldsymbol{\varepsilon}\boldsymbol{\beta}_r+\boldsymbol{\varepsilon}_Y. \label{eq:regplugin} 
+\end{equation}
+%\end{minipage}}
+Then the Best Linear Unbiased Estimator ({\sc BLUE}) for $\boldsymbol{\beta}_r$ is given ({\sc mle} estimator) by:
+\begin{equation}
+  \hat{\boldsymbol{\beta}_r}=(\boldsymbol{\varepsilon}'\boldsymbol{\varepsilon})^{-1}\boldsymbol{\varepsilon}'\boldsymbol{\varepsilon}_Y^* . \label{olsplugin}
+\end{equation}
+    And we have the following estimators:
+	\begin{eqnarray}
+		\hat{\boldsymbol{\varepsilon}}&=& \boldsymbol{X}_r-\boldsymbol{X}_f\hat{\boldsymbol{\alpha}}^* \nonumber \textrm{ and} \\
+		\hat{\boldsymbol{\varepsilon}}_Y^*&=&\boldsymbol{Y}-\boldsymbol{X}_f\hat{\boldsymbol{\beta}}_f^* \nonumber
+	\end{eqnarray}
+	that we can use by plug-in.
+	\end{frame}
+	\begin{frame}{Plug-in model}
+	\begin{eqnarray}
+		\hat{\boldsymbol{\beta}}_r^{\varepsilon}&=&(\hat{\boldsymbol{\varepsilon}}'\hat{\boldsymbol{\varepsilon}})^{-1}\hat{\boldsymbol{\varepsilon}}'\hat{\boldsymbol{\varepsilon}}_Y^*  \nonumber 
+	\end{eqnarray}
+that depends on all covariates in $\boldsymbol{X}$ and relies on the estimated coefficients of sub-regressions $\hat{\boldsymbol{\alpha}}^*$ and on the estimate $\hat{\boldsymbol{\beta}}_f^*$ of the coefficients in the marginal model.
+Then we can estimate $\boldsymbol{Y}$ by:
+\begin{eqnarray}
+	\hat{\boldsymbol{Y}}_{plug-in}&=&\boldsymbol{X}_f\hat{\boldsymbol{\beta}}^*_f + \hat{\boldsymbol{\varepsilon}}\hat{\boldsymbol{\beta}}_{r}^{\varepsilon}. \label{hatYplugin} 
+\end{eqnarray}
+We can improve estimation of $\boldsymbol{\beta}_f$ (in terms of bias) by doing an additional identification step.	We know that $\boldsymbol{\beta}^*_f=\boldsymbol{\beta}_f+\boldsymbol{\alpha}^*\boldsymbol{\beta}_r $ so we naturally define the following estimator:
+			\begin{equation}
+			\hat{\boldsymbol{\beta}}_f^{\varepsilon}=\hat{\boldsymbol{\beta}}^*_f-\hat{\boldsymbol{\alpha}}^*\hat{\boldsymbol{\beta}}_{r}^{\varepsilon}. \nonumber 
+			\end{equation}
+		\end{frame}
 
+		\begin{frame}{Marginal properties}
+		biased
 		\end{frame}
+		\begin{frame}{Plug-in properties}
+		asymptotically unbiased
+		\end{frame}
+		\begin{frame}
+		\begin{figure}[h!]
+\centering
+	\includegraphics[width=280px]{figures/MQE_toutOLSp5col.png}
+	\caption{\textsc{mse} on $\hat{\boldsymbol{\beta}}$ of \textsc{ols} (plain red) and {\tt CorReg} marginal (blue dashed) and {\tt CorReg} plug-in (green dotted) estimators for varying $R^2$ of the sub-regression, $n$ and $\sigma_Y$.
+Results obtained on the running example with $d=5$ covariates.}\label{MQE2}
+\end{figure}
+		\end{frame}
+
+\begin{frame}{Lasso Consistency}
+Consistency issues of the {\sc lasso} are well known and Zhao \cite{Zhao2006MSC} gives a very simple example to illustrate it.
+		We have taken the same example to show how our method is better to find the true relevant covariates.
+		Here $d=3$ and $n=1\;000$.\\
+		We define $\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{\varepsilon}_Y, \boldsymbol{\varepsilon}_{1} \quad i.i.d. \sim \mathcal{N}(\boldsymbol{0},\boldsymbol{I}_n)$ and then \\
+		\begin{eqnarray}
+		\boldsymbol{X}^3%|\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{S};\sigma_1^2
+		&=&\frac{2}{3}\boldsymbol{X}^1+\frac{2}{3}\boldsymbol{X}^2+\frac{1}{3}\boldsymbol{\varepsilon}_1 \textrm{ and}  \nonumber \\
+		\boldsymbol{Y}%|\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{S};\sigma_Y^2
+		&=&2\boldsymbol{X}^1+3\boldsymbol{X}^2+\boldsymbol{\varepsilon}_Y. \nonumber 
+		\end{eqnarray}
+\end{frame}		
+	\begin{frame}{Lasso Consistency}
+True $\boldsymbol{S}$ was found $991$ times on $1\;000$ tries.% (model is not identifiable because $\boldsymbol{X}^j$ are all Gaussian).
+		\begin{table}[h!]
+		\resizebox{\linewidth}{!}{%
+		 \begin{tabular}{|c|c|c|c|}
+		\hline 
+		 & Classical {\sc lasso} & {\tt CorReg} marginal + {\sc lasso}& {\tt CorReg} full plug-in + {\sc lasso}\\ 
+		\hline 
+		True $\boldsymbol{S}$ &  1.003303 (0.046) & \textbf{1.002273} (0.046)& \textbf{1.002812} (0.046)\\ 
+		\hline 
+		$\hat{\boldsymbol{S}}$ & 1.003303 (0.046)& 1.017622 (0.17)& \textbf{1.002812} (0.046)\\ 
+		\hline 
+		\end{tabular} 
+		}
+		\caption{\textsc{mse} observed on a validation sample (1 000 individuals) and their standard deviation (between brackets).}\label{MSEconsistlasso}
+		\end{table}
+
+We look at the consistency  that is the real stake:
+		\begin{table}[h!]	
+		\resizebox{\linewidth}{!}{%
+		\begin{tabular}{|c|c|c|c|}
+		\hline 
+		 & Classical {\sc lasso} & {\tt CorReg}  marginal + {\sc lasso}& {\tt CorReg}  full plug-in  + {\sc lasso} \\ 
+		\hline 
+		True $\boldsymbol{S}$ &  0 & \textbf{1000} & \textbf{835} \\ 
+		\hline 
+		$\hat{\boldsymbol{S}}$ & 0 & \textbf{991} & \textbf{829} \\ 
+		\hline 
+		\end{tabular} 
+		}
+		\caption{Number of consistent models found on $1\;000$ tries.}\label{testidentifiableG}
+		\end{table}	
+\end{frame}
+
 \section{Structure estimation}

### 2015-02-12 pkg/CorReg/R/inkei.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1050 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,6 +1,7 @@
 inkei <- function(length=35, squirt=4, thickness=4)
 {
-   if (thickness > 3){
+   if (thickness <= 3){thickness=4}
+   
       cat("          ", rep("_", length), "\n", sep="")
       
       for (t in 1:floor(thickness/2))
@@ -12,7 +13,7 @@ inkei <- function(length=35, squirt=4, thickness=4)
       {
          cat("          ", rep(" ", length+floor(thickness/2)-t+1), "/\n", sep="")
       }
-   }
+   
    cat("          ", rep("_", length+1), "/\n", sep="")
    cat("          \\
   .... |...|

### 2015-02-12 pkg/CorReg/R/showdata.R
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1050 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -2,14 +2,15 @@
 #'
 #'@param X the dataset to analyse
 #'@param what indicates what to plot
+#'@param pch for missing, symbol to plot (can set pch="." for large datasets)
 #'@export
 
-showdata<-function(X=X,what=c("miss","correl")){
+showdata<-function(X=X,what=c("miss","correl"),pch=7){
    what=what[1]
    if(what=="miss"){
       M=which(is.na(X),arr.ind=T)
       if(nrow(M)>1){
-         plot(M[,c(2,1)],pch=7)
+         plot(M[,c(2,1)],pch=pch)
          title("Missing values in the dataset")  
       }else{
          print("No missing values")

### 2015-02-12 www/soutenance.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1051 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 	\subsection{Statistical context}
-		\begin{frame}{Regression}
+		\begin{frame}{Linear Regression}
 			\begin{equation}
@@
 
-		\begin{frame}{OLS}
-		\begin{figure}[h!]
-	\centering
-	\includegraphics[width=220px]{figures/OLS_geometric_interpretation.png}
-	\caption{Multiple linear regression with Ordinary Least Squares seen as a projection on the $d-$dimensional hyperplane spanned by the regressors $\boldsymbol{X}$.
-Public domain image.} \label{geomOLS}
-	\end{figure}
-		\end{frame}
-
-		\begin{frame}{OLS}
+		\begin{frame}{MLE}
 		$\boldsymbol{\beta}$ can be estimated by $\hat{\boldsymbol{\beta}}$ with Ordinary Least Squares (\textsc{ols}), that is the unbiased maximum likelihood estimator \cite{saporta2006probabilites,dodge2004analyse}: %As shown in section \ref{sectionOLS},
@@
 		\end{frame}
-
+		\begin{frame}{OLS}
+		\begin{figure}[h!]
+	\centering
+	\includegraphics[width=220px]{figures/OLS_geometric_interpretation.png}
+	\caption{Multiple linear regression with Ordinary Least Squares seen as a projection on the $d-$dimensional hyperplane spanned by the regressors $\boldsymbol{X}$.
+Public domain image.} \label{geomOLS}
+	\end{figure}
+		\end{frame}
 \begin{frame}{Running example}
@@
 
-		\begin{frame}{ SEM }
-		 Modélisation de la structure mais à la main et aucun impact sur l'estimation
+		\begin{frame}{Simultaneous Equation Modeling }
+		 \begin{eqnarray}
+		 	\boldsymbol{X}_{n,d}&=&\boldsymbol{X}_{n,d}\boldsymbol{\beta}_{d,d} + \boldsymbol{\varepsilon}_{n,d} \nonumber
+		 \end{eqnarray}
+		 No real "response variable", no real estimators for recursive {\sc sem}, no real variable selection (structure finder)
 		 \end{frame}
 
-		\begin{frame}{ Selvarclust }
-			Semble très bien mais n'aboutit pas vers la régression donc on le prolonge en CorReg
+		\begin{frame}{ {\tt Selvarclust} }
+			\begin{equation}
+	\boldsymbol{X}^{U}%|\boldsymbol{X}^{R};\boldsymbol{\alpha},\boldsymbol{\Omega}
+	=a+\boldsymbol{X}^R \boldsymbol{\alpha}+ \boldsymbol{\varepsilon}  \nonumber 
+\end{equation}
+	\begin{itemize}
+	\item Structure based on stepwise \cite{raftery2006variable,miller2002subset}
+	\item Made only for clustering purpose.
+\end{itemize}
+
 		\end{frame}
@@
 	\begin{frame}{Hypothesis 2}
-	the response covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates of cardinal $d_f=d-d_r=|J_f|$.
+	The response covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates of cardinal $d_f=d-d_r=|J_f|$.
 We call this hypothesis the \underline{uncrossing rule}.
@@
 	\end{frame}
+
 	\begin{frame}{Hypotheses 3}
@@
 
-		\begin{frame}{Marginal properties}
-		biased
+		\begin{frame}{Marginal model's properties}
+				\begin{equation}
+			\mathbb{E}(\hat{\boldsymbol{\beta}}_{f}^*)=\boldsymbol{\beta}_{f}+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^* \textrm{ and }\mathbb{E}(\hat{\boldsymbol{\beta}}_{r}^*)=\boldsymbol{0}  \nonumber
+		\end{equation}
+
+		\begin{equation}
+			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}.
+\label{eq:varOLS*}
+		\end{equation}
 		\end{frame}
-		\begin{frame}{Plug-in properties}
-		asymptotically unbiased
+
+\begin{frame}{Plug-in model's properties}
+		Asymptotically unbiased (consistent estimators + continuous mapping theorem)
+\begin{eqnarray}
+	\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^{\varepsilon})
+	&=& \sigma^2_Y\big[\hat{\boldsymbol{\varepsilon}}'\hat{\boldsymbol{\varepsilon}}[\hat{\boldsymbol{\varepsilon}}'(\boldsymbol{I}_n-\boldsymbol{H}_f)\hat{\boldsymbol{\varepsilon}}]^{-1}\hat{\boldsymbol{\varepsilon}}'\hat{\boldsymbol{\varepsilon}}\big]^{-1} \nonumber \\
+	&=&\sigma^2_Y\big[(\boldsymbol{X}_r^{\varepsilon})'\boldsymbol{X}_r^{\varepsilon}\big]^{-1}
+\nonumber
+\end{eqnarray}
+
+\begin{eqnarray}
+		\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^{\varepsilon})
+		&=&\sigma_Y^2\big[(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}+\hat{\boldsymbol{\alpha}}^*\big[(\boldsymbol{X}_r^{\varepsilon})'\boldsymbol{X}_r^{\varepsilon}\big]^{-1}(\hat{\boldsymbol{\alpha}}^*)'\big].\nonumber \label{varpluginf}
+	\end{eqnarray}
+
 \end{frame}
+
+\begin{frame}
+	\fontsize{8}{8}\selectfont{\begin{eqnarray}
+	&&\operatorname{Var}_{ols}(\hat{\boldsymbol{\beta}}_{f})= \nonumber \\ 
+	&& \sigma_Y^2\big[ 
+(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}+(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{X}_r
+[\boldsymbol{X}_r'\boldsymbol{X}_r-\boldsymbol{X}_r'\boldsymbol{X}_f(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{X}_r]^{-1}
+\boldsymbol{X}_r'\boldsymbol{X}_f(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\big] , \nonumber 
+%	\nonumber ,\\
+%	\operatorname{Var}_{ols}(\hat{\boldsymbol{\beta}}_{r})= 
+%	\nonumber .
+\end{eqnarray}}
+\begin{equation}
+\qquad = \sigma_Y^2\big[ 
+(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}+\underbrace{(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{X}_r}_{\hat{\boldsymbol{\alpha}}^*_{ols}}
+[ \boldsymbol{X}_r'(\boldsymbol{I}_n-\boldsymbol{H}_f)\boldsymbol{X}_r]^{-1}
+\underbrace{\boldsymbol{X}_r'\boldsymbol{X}_f(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}}_{(\hat{\boldsymbol{\alpha}}^*_{ols})'}\big] \nonumber
+\end{equation}
+\begin{eqnarray}
+	\operatorname{Var}_{ols}(\hat{\boldsymbol{\beta}}_{r})&=&
+	\sigma_Y^2\big[ \boldsymbol{X}_r'\boldsymbol{X}_r-\boldsymbol{X}_r'\boldsymbol{X}_f(\boldsymbol{X}_f'\boldsymbol{X}_f)^{-1}\boldsymbol{X}_f'\boldsymbol{X}_r\big]^{-1} \\ \nonumber
+	&=&\sigma_Y^2\big[ \boldsymbol{X}_r'(\boldsymbol{I}_n-\boldsymbol{H}_f)\boldsymbol{X}_r\big]^{-1}. \nonumber
+\end{eqnarray}
+\end{frame}
+
 		\begin{frame}
@@
 True $\boldsymbol{S}$ was found $991$ times on $1\;000$ tries.% (model is not identifiable because $\boldsymbol{X}^j$ are all Gaussian).
-		\begin{table}[h!]
-		\resizebox{\linewidth}{!}{%
-		 \begin{tabular}{|c|c|c|c|}
-		\hline 
-		 & Classical {\sc lasso} & {\tt CorReg} marginal + {\sc lasso}& {\tt CorReg} full plug-in + {\sc lasso}\\ 
-		\hline 
-		True $\boldsymbol{S}$ &  1.003303 (0.046) & \textbf{1.002273} (0.046)& \textbf{1.002812} (0.046)\\ 
-		\hline 
-		$\hat{\boldsymbol{S}}$ & 1.003303 (0.046)& 1.017622 (0.17)& \textbf{1.002812} (0.046)\\ 
-		\hline 
-		\end{tabular} 
-		}
-		\caption{\textsc{mse} observed on a validation sample (1 000 individuals) and their standard deviation (between brackets).}\label{MSEconsistlasso}
-		\end{table}
-
 We look at the consistency  that is the real stake:
@@
 \section{Structure estimation}
-	\begin{frame}{•}
+	\begin{frame}
+	\fontsize{8}{8}\selectfont{
+	\begin{displaymath}G=\left(	\begin{array}{ccccc}
+	0 & 0 & 1 & 0 & 0 \\ 
+	0 & 0 & 1 & 0 & 0 \\ 
+	0 & 0 & 0 & 0 & 0 \\ 
+	0 & 0 & 0 & 0 & 0 \\ 
+	0 & 0 & 0 & 0 & 0
+	\end{array} \right)
+\end{displaymath}
+}
+	\begin{figure}[h!]
+	\centering
+	\includegraphics[width=200px]{figures/graphetoy.png} 
+	\caption{The bipartite graph associated to the running example. $\boldsymbol{X}_r$ is green and $\boldsymbol{X}_f$ is blue.}
+	\end{figure}
+	\end{frame}
 
+\begin{frame}{Hypothesis 4: Full generative model}
+
+All covariates $\boldsymbol{X}^j$ with $j \in J_f$ are mutually independent and arise from the following Gaussian mixture of $K_j$ components
+\begin{equation}
+\forall 1\leq i \leq n, \mathbb{P}(x_{i,j}|\boldsymbol{S};\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\Sigma}_j) = \sum_{h=1}^{K_j} \pi_{j,h} \Phi(x_{i,j};\mu_{j,h} ,\Sigma_{j,h}), \nonumber
+\end{equation}
+where $\boldsymbol{\pi}_{j}=(\pi_{j,1},\ldots,\pi_{j,K_j})$ is the vector of mixing proportions with $\forall 1\leq h\leq k_j , \pi_{j,h}>0$ and $\sum_{h=1}^{K_j}\pi_{j,h}=1$, $\boldsymbol{\mu}_j=(\mu_{j,1},\ldots,\mu_{j,K_j})$  is the vector of centres and $\boldsymbol{\Sigma}_j=(\Sigma_{j,1},\ldots,\Sigma_{j,K_j})$ is the vector of variances and $\Phi$ is the Gaussian density function.
+\\We stack together all these mixture parameters in $\boldsymbol{\theta}=(\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\Sigma}_j ; j \in J_f)$.
+We now have a full generative model on $\boldsymbol{X}$.
+\end{frame}
+
+\begin{frame}{Penalized BIC}
+we propose to introduce some information in $\mathbb{P}(\boldsymbol{S})$ promoting simple models through the following {\it hierarchical} uniform distribution denoted by $\mathbb{P}_H(\boldsymbol{S})$:
+\fontsize{10}{10}\selectfont{\begin{eqnarray}
+\mathbb{P}_H(\boldsymbol{S}) & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p) \nonumber \\
+ & = & \mathbb{P}_H(\boldsymbol{J}_r,\boldsymbol{J}_p,d_r,\boldsymbol{d}_p)  \nonumber \\
+ & = & \mathbb{P}_U(\boldsymbol{J}_p|\boldsymbol{d}_p,\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{d}_p|\boldsymbol{J}_r,d_r) \times \mathbb{P}_U(\boldsymbol{J}_r|d_r)\times \mathbb{P}_U(d_r)  \nonumber \\
+ & = & \left[\prod_{j=1}^{d_r} \left(\begin{array}{c} d-d_r \\ d_p^j \end{array}\right) \right]^{-1} \times \left[d-d_r\right]^{-d_r} \times \left[\left(\begin{array}{c} d \\ d_r \end{array} \right)\right]^{-1} \times [d+1]^{-1} \nonumber
+\end{eqnarray}
+}
+\end{frame}
+\subsection{Markov chain}
+\begin{frame}
+		At each step the Markov chain moves with probability:
+	\begin{eqnarray}
+			\forall \boldsymbol{S} \in \mathcal{S}_d, \forall \boldsymbol{S}^+\in \mathcal{V}(\boldsymbol{S}): %\mathcal{P}_{\boldsymbol{S},\boldsymbol{S}^+}=
+			\mathbb{P}(\boldsymbol{S}^+|\mathcal{V}(\boldsymbol{S}))&=&\frac{\exp{(-\mbox{{\sc bic}}_*(\boldsymbol{S}^+))}}{\sum_{\tilde{\boldsymbol{S}}\in \mathcal{V}(\boldsymbol{S})}\exp{(-\mbox{{\sc bic}}_*(\tilde{\boldsymbol{S}}))}} \nonumber
+	\end{eqnarray}
+	where $\mathcal{V}(\boldsymbol{S})$ is a neighbourhood and $\mathcal{S}_d$ is the set of feasible structures within $d$ variables.
+\end{frame}
+
+\begin{frame}
+	For each step $(q)$, starting from $\boldsymbol{S} \in \mathcal{S}_d$ we define a neighbourhood:
+		\begin{eqnarray}
+		\mathcal{V}(\boldsymbol{S})&=& \{\boldsymbol{S} \}\cup \{ \boldsymbol{S}^{(i,j)} \in \mathcal{S}_d|(i,j) \in \mathcal{A}_{(q)}\} \nonumber 
+	\end{eqnarray}	
+	where $\mathcal{A}_{(q)}$ is a set of couples $(i,j) \in \{1,\dots,d \}^2$ with $i\neq j$ drawn at the step $(q)$ according to a strategy defined below and corresponding to the directed edge of the graph to modify (add or remove).
+	And we have for  $\tilde{\boldsymbol{S}}=\boldsymbol{S}^{(i,j)}$:
+	\begin{eqnarray}
+		\forall (k,l)\neq (i,j), \quad	\tilde{\boldsymbol{G}}_{k,l}&=&\boldsymbol{G}_{k,l} \nonumber  \\
+		\tilde{\boldsymbol{G}}_{i,j}&=&1-\boldsymbol{G}_{i,j}  \nonumber 
+	\end{eqnarray}
+	where $\tilde{\boldsymbol{G}}$ is the adjacency matrix associated to $\tilde{\boldsymbol{S}}$. Any strategy can be chosen for $\mathcal{A}_{(q)}$, from uniform distribution to specific heuristics.
+\end{frame}
+\subsection{Constraint Relaxation}
+\begin{frame}
+New definition of $\tilde{\boldsymbol{G}}$:
+		Modification of the selected directed edge $(i,j)$ on the graph:
+	\begin{eqnarray}
+		\tilde{\boldsymbol{G}}_{i,j}&=&1-\boldsymbol{G}_{i,j} \textrm{ as usual and} \nonumber \\
+		\forall k \neq i, l\neq j, \quad	\tilde{\boldsymbol{G}}_{k,l}&=&\boldsymbol{G}_{k,l} \nonumber 
+\end{eqnarray}	
+Column-wise relaxation :		
+newly predictive covariate cannot be regressed anymore:
+		\begin{eqnarray}
+		\forall k \in \{1,\dots,d\}\setminus \{i\}, \tilde{\boldsymbol{G}}_{k,j}&=&\boldsymbol{G}_{i,j}\boldsymbol{G}_{k,j}  \nonumber  \\
+				\end{eqnarray}
+Row-wise relaxation: newly regressed covariate cannot be predictive anymore:
+		\begin{eqnarray}
+		\forall l \in \{1,\dots,d\}\setminus \{j\}, \tilde{\boldsymbol{G}}_{i,l}&=&\boldsymbol{G}_{i,j}\boldsymbol{G}_{i,l} \textrm{ (row-wise relaxation)} \nonumber 
+	\end{eqnarray} 
+\end{frame}
+
+\begin{frame}
+\begin{figure}
+	\begin{minipage}[l]{.45\linewidth}
+\includegraphics[width=.90\linewidth]{figures/identifiable/identifiabilite001.png} 
+	\end{minipage}
+	\begin{minipage}[c]{.45\linewidth}
+		$\boldsymbol{G}=\left( \begin{array}{ccccc}
+		0 & 0 & 0 & 1 & 0 \\ 
+		0 & 0 & 0 & 1 & 1 \\ 
+		0 & 0 & 0 & 0 & 1 \\ 
+		0 & 0 & 0 & 0 & 0 \\ 
+		0 & 0 & 0 & 0 & 0
+		\end{array} \right)$
+	\end{minipage}
+	\caption{We start from a structure $\boldsymbol{S}=((4,5),(\{1,2\},\{2,3\}))$ and its associated matrix $G$}\label{ident1}
+\end{figure}
+\end{frame}
+\begin{frame}		
+\begin{figure}
+	\begin{minipage}[c]{.45\linewidth}
+\includegraphics[width=.90\linewidth]{figures/identifiable/identifiabilite002.png} 
+	\end{minipage}
+	\begin{minipage}[c]{.45\linewidth}
+		$\left( \begin{array}{ccccc}
+		0 & 0 & 0 & 1 & 0 \\ 
+		0 & 0 & 0 & 1 & 1 \\ 
+		0 & 0 & 0 & 0 & 1 \\ 
+		0 & 0 & 0 & 0 & 0 \\ 
+		0 & \textcolor{red}{\textbf{1}}& 0 & 0 & 0
+		\end{array} \right)$
+	\end{minipage}
+	\caption{We want to define the candidate $\tilde{\boldsymbol{S}}=\boldsymbol{S}^{(5,2)}$ and its associated matrix but the structure obtained would not be feasible (breaking the uncrossing rule).
+}\label{ident2}
+\end{figure}
+\end{frame}
+\begin{frame}		
+\begin{figure}
+	\begin{minipage}[c]{.45\linewidth}
+\includegraphics[width=.90\linewidth]{figures/identifiable/identifiabilite003.png} 
+	\end{minipage}
+	\begin{minipage}[c]{.45\linewidth}
+		$\left( \begin{array}{ccccc}
+		0 & 0 & 0 & 1 & \textcolor{green}{\textbf{0}} \\ 
+		0 & 0 & 0 & 1 & \textcolor{green}{\textbf{0}} \\ 
+		0 & 0 & 0 & 0 & \textcolor{green}{\textbf{0}} \\ 
+		0 & 0 & 0 & 0 & \textcolor{green}{\textbf{0}} \\ 
+		0 & \textcolor{red}{\textbf{1}}& 0 & 0 & \textcolor{green}{\textbf{0}}
+		\end{array} \right)$
+	\end{minipage}
+	\caption{Column-wise relaxation: newly predictive covariate cannot be regressed anymore. }\label{ident3}
+\end{figure}
+\end{frame}
+\begin{frame}		
+\begin{figure}
+	\begin{minipage}[c]{.45\linewidth}
+\includegraphics[width=.90\linewidth]{figures/identifiable/identifiabilite004.png} 
+	\end{minipage}
+	\begin{minipage}[c]{.45\linewidth}
+		$\tilde{\boldsymbol{G}}=\left( \begin{array}{ccccc}
+		0 & 0 & 0 & 1 & 0 \\ 
+		\textcolor{green}{\textbf{0}} & \textcolor{green}{\textbf{0}} & \textcolor{green}{\textbf{0}} & \textcolor{green}{\textbf{0}} & \textcolor{green}{\textbf{0}} \\ 
+		0 & 0 & 0 & 0 & 0 \\ 
+		0 & 0 & 0 & 0 & 0 \\ 
+		0 & \textcolor{red}{\textbf{1}}& 0 & 0 & 0
+		\end{array} \right)$
+	\end{minipage}
+	\caption{Row-wise relaxation: newly predictive covariate cannot be regressed anymore. }\label{ident4}
+\end{figure}	
+\end{frame}
+\begin{frame}
+\begin{figure}
+	\begin{minipage}[c]{.45\linewidth}
+\includegraphics[width=.90\linewidth]{figures/identifiable/identifiabilite005.png} 
+	\end{minipage}
+	\begin{minipage}[c]{.45\linewidth}
+		$\tilde{\boldsymbol{G}}=\left( \begin{array}{ccccc}
+		0 & 0 & 0 & 1 & 0 \\ 
+		0 &0 & 0 & 0 & 0 \\ 
+		0 & 0 & 0 & 0 & 0 \\ 
+		0 & 0 & 0 & 0 & 0 \\ 
+		0 & 1& 0 & 0 & 0
+		\end{array} \right)$
+	\end{minipage}
+	\caption{We get a feasible candidate $\tilde{\boldsymbol{S}}=((2,4),(\{5\},\{1\}))$ that does differ from $\boldsymbol{S}$ in many points. }\label{ident5}
+\end{figure}
+\end{frame}
+\begin{frame}
+\begin{figure}
+\centering
+\includegraphics[width=250px]{figures/identifiable/identifiabilite006.png} 
+	\caption{All this modifications are made in only one step in the MCMC, meaning an increased scope for the neighbourhoods.}\label{ident6}
+\end{figure}
+\end{frame}
+
+\subsection{Pruning}
+\begin{frame}{Pruning}
+	Every sub-graph of a bipartite graph is bipartite thus every sub-graph can be reached. We propose an heuristic change in the strategy with:
+			\begin{equation}
+				\mathcal{A}_{(q)}=\{(i,J_r^j), i \in J_f, j \in \{1,\dots,d_r \}: i \in J_p^j \}. \nonumber 
+			\end{equation}
+\end{frame}
+\subsection{Initializations}
+\begin{frame}
+	\begin{center}
+		\begin{figure}[h!]
+		\centering
+		\includegraphics[width=250px]{figures/typeinit.png} %mars 2013 analyse plan ini marelax
+		\caption{Evolution of the $\mbox{{\sc bic}}$ (criterion to minimize in the MCMC) for each method.}\label{Bicini}
+		\end{figure}
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\begin{center}
+	\begin{figure}[h!]
+	\centering
+		\includegraphics[width=250px]{figures/courbes_BICmoyen_marelax_nbinibis.png} 
+		\caption{Comparison of distinct number of correlation-based initializations for the MCMC. Dark blue=1, red=10.}\label{nbini}
+	\end{figure}
+	\end{center}
+		In the following, the chain was launched with twenty initializations each time, based on the correlation matrix.
+\end{frame}
+
+\begin{frame}
+\begin{center}
+	\begin{figure}[h!]
+	\centering
+\includegraphics[width=300px]{figures/complexitycompareMCMC.png} 		
+\caption{Comparison of complexity evolution with or without constraint relaxation.}\label{comparecomplrelax}
+	\end{figure}
+	\end{center}
+	\end{frame}
+
+\begin{frame}	
+\begin{center}
+	\begin{figure}[h!]
+	\centering
+\includegraphics[width=300px]{figures/BICcompareMCMCrelax.png} 		
+\caption{Comparison of $\mbox{{\sc bic}}$ evolution with or without constraint relaxation.}\label{compareBICrelax}
+	\end{figure}
+	\end{center}
 \end{frame}
@@
 	\subsection{Simulation results}
-		\begin{frame}{•}
 
+		\begin{frame}
+			\begin{figure}[h!]
+	 \includegraphics[width=250px]{figures/predhatS/tout/MSEpredOLSbeta_zonetout.png}
+	\caption{Comparison of the \textsc{mse} on $\hat{\boldsymbol{\beta}}$, plain red=classical (complete) model, dashed blue=marginal model, dotted green=plug-in model. {\sc ols} estimators}
+\end{figure}
 		\end{frame}
+				\begin{frame}
+			\begin{figure}[h!]
+	 \includegraphics[width=250px]{figures/predhatS/tout/MSEpredlarbeta_zonetout.png}
+	\caption{Comparison of the \textsc{mse} on $\hat{\boldsymbol{\beta}}$, plain red=classical (complete) model, dashed blue=marginal model, dotted green=plug-in model. {\sc lasso} estimators}
+\end{figure}
+		\end{frame}
+		\subsection{Non-linear cases}
+
+\begin{frame}
+We define $%\boldsymbol{X}_r|\boldsymbol{X}_f;\boldsymbol{\alpha},\boldsymbol{\sigma}^2=
+\boldsymbol{X}^7%|\boldsymbol{X}^{\{1,2,3\}};a,\sigma_1^2
+=a(\boldsymbol{X}^1)^2+\boldsymbol{X}^2+\boldsymbol{X}^3+ \boldsymbol{\varepsilon}_1$. The matrix $\boldsymbol{X}$ is then scaled before doing $$\boldsymbol{Y}%|\boldsymbol{X};\sigma_Y^2
+=\sum_{i=1}^7\boldsymbol{X}^i+\boldsymbol{\varepsilon}_Y.$$ We let $a$ vary between $0$ and $10$ to increase progressively the non-linear part of the sub-regression.
+Once again, simulations has been made 100 times and the \textsc{mse} were computed with 1 000 individuals validation samples.
+\end{frame}
+
+		\begin{frame}
+\begin{figure}[h!] 
+	\subfigure[Evolution of the quality of $\hat{\boldsymbol{S}}$ when the paramater $a$ increases]{
+			\includegraphics[height=180px,width=230px]{figures/robust_S.png} 
+			\label{resnonlin}
+	} 
+   \subfigure[\textsc{mse} on the main regression for \textsc{ols}(thick) and \textsc{lasso} (thin) used both with (plain) or without {\tt CorReg} (dotted).]{
+			\includegraphics[height=180px,width=230px]{figures/robustY.png} \label{MSEnonlin}
+   }
+   \caption{Non-linear case analysis.}
+\end{figure}		\end{frame}
+
+		\begin{frame}
+\begin{figure}[h!]  
+   \subfigure[\textsc{mse} on the main regression for \textsc{ols}(thick) and \textsc{lasso} (thin) used both with (plain) or without {\tt CorReg} (dotted).]{
+			\includegraphics[height=180px,width=230px]{figures/robustY.png} \label{MSEnonlin}
+   }
+\end{figure}		\end{frame}
+
 	\subsection{Industrial results}
-		\begin{frame}{•}
+		\begin{frame}
+		We have :
+		\begin{itemize}
+			\item a quality parameter (confidential) as response variable,
+			\item $d=205$ variables from the whole process to explain it.
+		\end{itemize}
+We get a training set of $n=3\;000$ products described by these $205$ variables from the industrial process and also a validation sample of $847$ products.\\
 
+The objective here is not only to predict non-quality but to understand and then to avoid it.
 		\end{frame}
+
+	\begin{frame}
+	\begin{figure}[h!]
+\begin{center}
+			\includegraphics[height=150px, width=150px]{figures/res_article/gaussianmixture_exfo.png}
+			\includegraphics[height=150px, width=150px]{figures/res_article/nb_comp_X_exfo.png}
+\end{center}		
+			\caption{Quality case study: (left) Example of a non-Gaussian real variable easily modeled by a Gaussian mixture, (right) distribution of the number of components found for each covariate.}\label{fig:graphMixmod.quality}
+\end{figure}  	
+	\end{frame}
+
+		\begin{frame}
+		\begin{figure}[h!]
+\begin{center}
+			\includegraphics[width=220px, height=80px]{figures/histR2exfos.png}
+\end{center}
+			\caption{Quality case study: histogram of the adjusted $R^2_{adj}$ for the $d_r=76$ sub-regressions.}
+\end{figure}
+		 The width and the weight of a steel slab gives $|\rho|=0.905$, the temperature before and after some tool gives $|\rho|=0.983$, the  roughness of both faces of the product gives $|\rho|= 0.919$ and a particular mean and a particular max gives $|\rho|=0.911$.
+		\end{frame}
+
+\begin{frame}
+		\begin{table}[h!]
+\centering
+\begin{tabular}{llcc}
+	\hline
+	Method& Indicator& {\tt CorReg}'s reduced model)& complete model \\ 
+	\hline \hline
+	{\sc ols} & {\sc mse} & 13.30 & 14.03 \\
+		& complexity& 130 & 206 \\
+	\hline
+	{\sc lasso} & {\sc mse} & 12.77 & 12.96 \\
+		& complexity& 24 & 21 \\
+	\hline
+	elasticnet & {\sc mse} & \textbf{12.15} & 13.52 \\
+		& complexity& 40 & 78 \\
+	\hline
+	ridge & {\sc mse} & 12.69 & 13.09 \\
+		& complexity& 130 & 206 \\
+	\hline
+\end{tabular} 
+\caption{Quality case study: Results obtained on a validation sample ($n=847$ individuals). In bold, the best {\sc mse} value.}\label{Res_exfos}
+\end{table}
+\end{frame}
+
 \section{Missing values}
-	\begin{frame}{•}
-	Modèle génératif complet avec dépendances
+	\begin{frame}{Hope}
+		Full generative model + explicit dependencies = missing values management
 	\end{frame}
-	\begin{frame}{•}
-	Explosion des mélanges 
+	\begin{frame}
+	The number of component $k$ can be huge (combinations of all the components of the covariates in $\boldsymbol{X}_f$). In fact we have 
+	\begin{equation}
+		K=\prod_{j\in J_f}K_j  \nonumber 
+	\end{equation}
+	where $K_j$ is the number of components of the Gaussian mixture followed by $\boldsymbol{X}_j$ as defined in Hypothesis 4.
+It is clear that $K$ can really explode even if some components may be identical (but it happens with zero probability).
+For instance, if $\boldsymbol{X}$ contains only 10 independent Gaussian mixtures and with only 2 components each, then $\boldsymbol{X}$ will have up to $K=2^{10}=$ 1 024 components.
+And if these mixtures have 3 components each, then it rises up to $K=$ 59 049 components.
 	\end{frame}
-	\begin{frame}{•}
-	SEM avec Gibbs 
+
+	\begin{frame}{Problem}
+	\begin{itemize}
+		\item Too much components
+		\item The likelihood may not be linear
+	\end{itemize}
+	EM is not possible in practice.
 	\end{frame}
-	\begin{frame}{•}
-	Bic pondéré 
+	\begin{frame}{Stochastic EM}
+		It is possible to simplify the problem by using Stochastic EM.\\
+		But the number of components stays a problem.
 	\end{frame}
-	\begin{frame}{•}
-	Résultats pourris 
+\begin{frame}{Solution}
+	Gibbs Sampling for stochastic imputation in the Stochastic EM
+\end{frame}
+
+	\begin{frame}
+ We define
+			\begin{equation}
+			c_j=\frac{1}{n}\sum_{i=1}^nc_{i,j},  \nonumber 
+\end{equation}						where $c_{i,j}$ is the number of parameters to estimate in $\mathbb{P}(x_{i,j}|\boldsymbol{X}_i\setminus \boldsymbol{X}_i^j)$.
+If all predictors $\boldsymbol{X}^{J_p^j}_i$ are observed then we have to estimate $d_p^j$ coefficients and an intercept and the variance of the residual so  $c_{i,j}=d_p^j+2$.
+But for each missing predictor, we have to estimate the parameters of its distribution (Gaussian mixture with proportion, mean and variance to estimate for each component) so we have $$c_{i,j}=d_p^j+2 + \sum_{\substack{l \in J_p^j \\ M_{i,l}=1}} (3 K_l-1)$$  (the sum of the proportions is 1 so one estimation is useless for each mixture).
+	\end{frame}
+	\begin{frame}{Results}
+	Very slow and too much variability
 	\end{frame}
 \section{Tools}
-	\begin{frame}{•}
-		Excel, fonctions graphiques, arbres de décision
+\begin{frame}
+\begin{figure}
+\centering
+\includegraphics[width=280px]{figures/captureexcel.PNG} 
+	\caption{Screenshot of the Graphical User Interface of {\tt CorReg} in Excel.}\label{captureexcel}
+\end{figure}
+\end{frame}
+
+	\begin{frame}
+		\begin{figure}
+			\centering
+				\includegraphics[width=250px]{figures/arbretoy.png} 
+			\caption{Regression tree obtain with the package {\tt CorReg}
+			 (graphical layer on top of  the {\tt rpart} package) on the running example.}\label{arbretoy}
+		\end{figure}
+		arbres de décision, showdata, Conan, {\it etc.}
 	\end{frame}
@@
 \end{frame}
-
+\begin{frame}
+\begin{center}
+\huge Merci pour ce moment\dots
+\end{center} 
+\end{frame}
 \end{document}

### 2015-02-14 pkg/CorReg/R/Conan.R
up Conan files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@1055 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,98 @@
+#' Conan the destructor
+#' @export
+#' @description Removes missing values (rows and column to obtain a large full matrix)
+#' @param X the dataset (matrix)
+#' @param nbstep number of cutting steps (may remove several rows or columns at each step)
+#' @param sd remove constant covariates
+#' @param verbose to print the result
+#' @param coercing vector of the covariates to keep (names or index)
+#' @param Xout boolean to export or not the reduced matrix (if not, indices are sufficient)
+Conan<-function(X=X,nbstep=Inf,std=FALSE,verbose=FALSE,coercing=NULL,Xout=TRUE){
+   steploc=1 
+   nloc=nrow(X)
+   ploc=ncol(X)
+   n=nloc
+   p=ploc
+   individus_restants=1:n
+   variables_restantes=1:p
+   if(!is.null(coercing)){#on a des variables à garder  
+      #on va purger les individus manquants sur ces variables (pas le choix)
+      for(i in 1:length(coercing)){
+         if(is.numeric(coercing[i])){
+            j=coercing[i]
+         }else{
+            j=which(names(X)==coercing[i])[1] #si doublon on prend le premier
+         }
+         if(anyDuplicated(c(1:ploc,j))>0){#si on ne tape pas à côté
+            quimank=which(is.na(X[,j]))
+            if(length(quimank)>0){
+               X=X[-quimank,]
+               individus_restants=individus_restants[-quimank]
+            }
+         }
+      }
+      nloc=nrow(X)
+      ploc=ncol(X)
+   }
+   #nbmank/prod(dim(basefull))
+   if(nbstep==Inf){
+      nbstep=max(c(nloc,ploc))
+   }
+   if(verbose){print(dim(X))}
+   if(std){#nettoyage des constantes si demandé
+      sd_loc<-function(vect){
+         return(sd(vect[!is.na(vect)]))
+      }
+      quiconst=which(apply(X,2,sd_loc)==0)
+      if(length(quiconst)>0){
+         X=X[,-quiconst]
+         variables_restantes=variables_restantes[-quiconst]      
+      }   
+   }
+   if(verbose){print(dim(X))}
+   quimank=which(is.na(X),arr.ind = TRUE)
+   nbmank=dim(quimank)[1]
+   M=Matrix::Matrix(0,nrow = nrow(X),ncol=ncol(X))
+   M[quimank]=1 
+   while (steploc<=nbstep & nbmank>0 & nloc>0 & ploc>0){
[…]
